### 2019-03-08 article.tex
First version

\documentclass{article}

% Bibliography
\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

% Use 'It was found that A is B (Name 1234)' style
\setcitestyle{authoryear,open={},close={}}

% Affiliations
\usepackage{authblk}
\title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
\author[1]{Rich\`el J.C. Bilderbeek}
\author[1]{?Rampal S. Etienne}
\author[2]{Frans Bianchi}
\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}
\affil[2]{Frans' Institute, University of Groningen, Groningen, The Netherlands}

% Use double spacing
\usepackage{setspace}
\doublespacing

\usepackage{listings}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage{pgf}
\usepackage{bm}
\usepackage{multirow}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{array}
\usepackage{booktabs}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\usepackage{longtable}

\usepackage{tkz-graph}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}

% sidewaysfigure
\usepackage{rotating}

% Style of listings
% From http://r.789695.n4.nabble.com/How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
\usepackage{fancyvrb} 
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
\lstdefinestyle{mystyle}{
  language=R,% set programming language
  basicstyle=\ttfamily\small,% basic font style
  commentstyle=\color{gray},% comment style
  % numbers=left,% display line numbers on the left side
  numberstyle=\scriptsize,% use small line numbers
  numbersep=10pt,% space between line numbers and code
  tabsize=2,% sizes of tabs
  showstringspaces=false,% do not replace spaces in strings by a certain character
  captionpos=b,% positioning of the caption below
  breaklines=true,% automatic line breaking
  escapeinside={(*}{*)},% escaping to LaTeX
  fancyvrb=true,% verbatim code is typset by listings
  extendedchars=false,% prohibit extended chars (chars of codes 128--255)
  alsoletter={.<-},% becomes a letter
  alsoother={$},% becomes other
  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},% other keywords
  deletekeywords={c}% remove keywords 
}
\lstset{style=mystyle}

% Adds numbered lines
\usepackage{lineno}
\linenumbers

% Rename 'Abstract' to 'Summary 
\usepackage[english]{babel}
\addto{\captionsenglish}{\renewcommand{\abstractname}{Summary}}

%comments
\newcommand{\frans}[1]{\textcolor{blue}{\textbf{[FB: #1]}}}
\newcommand{\richel}[1]{\textcolor{orange}{\textbf{[RB: #1]}}}
\newcommand{\rampal}[1]{\textcolor{green}{\textbf{[RSE: #1]}}}

\begin{document}

\maketitle

\begin{abstract}

  \textbf{1. }
    Epiitopiwef
    wefwe. \\
  \textbf{2. }
    ewfw
    ewfwf. \\
  \textbf{3. }
    ewfqef. \\
  \textbf{4. }
    wefwef. \\
\end{abstract}

{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, pirouette, R

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Thanks to Geert van den Bogaart for his help.
We thank the Netherlands 
Organization for Scientific Research (NWO) for financial support 
through a VICI grant awarded to RSE.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data Accessibility}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

All code is archived at \url{http://github.com/richelbilderbeek/someplace},
with DOI \url{https://doi.org/12.3456/zenodo.1234567}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

RJCB, FB and RSE conceived the idea for this research. 
RJCB wrote the code.
FB supplied the DNA alignment.
RJCB, FB and RSE contributed to writing the article.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Appendix section 1}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2019-03-08 article.tex
Results of 6 percent of proteome

@@
 
+There are thousands of MHC2 alleles known.
+For this research, we picked the three alleles that are most abundant,
+accoring to \richel{reference here}, 
+which are DRB4*0101, 
+HLA-DPA10103-DPB10402 
+\richel{
+  we decided for DPB1*0402, but there are multiple alleles matching.
+  I just picked the first one of the alleles matching, which are:
+  HLA-DPA10103-DPB10402,
+  HLA-DPA10104-DPB10402,
+  HLA-DPA10105-DPB10402,
+  HLA-DPA10106-DPB10402,
+  HLA-DPA10107-DPB10402,
+  HLA-DPA10108-DPB10402,
+  HLA-DPA10109-DPB10402,
+  HLA-DPA10110-DPB10402,
+  HLA-DPA10201-DPB10402,
+  HLA-DPA10202-DPB10402,
+  HLA-DPA10203-DPB10402,
+  HLA-DPA10204-DPB10402,
+  HLA-DPA10301-DPB10402,
+  HLA-DPA10302-DPB10402,
+  HLA-DPA10303-DPB10402,
+  HLA-DPA10401-DPB10402
+}
+and DQA1*0501/DQB1*0301,
+that are present in respectively 31, 41 and 59 percent 
+\richel{fix numbers}
+of the current human population.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}
+\centering
+  \begin{tabular}{| r | c | c | c | c | c | c |}
+    \hline
+    \textbf{Allele}       & \textbf{i} & \textbf{m} & \textbf{o} & \textbf{I} & \textbf{M} & \textbf{O} \\ 
+    \hline
+    DRB4*0101             & 3752 & 2397 & 62024 & 1061 & 854  & 11952 \\
+    HLA-DPA10103-DPB10402 & 4335 & 2328 & 69040 & 478  & 923  & 4936 \\
+    DQA1*0501/DQB1*0301   & 3321 & 1329 & 45461 & 1492 & 1922 & 28515 \\
+    \hline 
+  \end{tabular}
+  \caption{
+    Results \richel{These are preliminary, 6 percent of proteome}
+  }
+  \label{tab:results}
+\end{table}
 

### 2019-03-08 article.tex
Results for 6 percent in nicer table

@@
   \caption{
-    Results \richel{These are preliminary, 6 percent of proteome}
+    \richel{These are preliminary, 6 percent of proteome}
+    Location and binding of amino acids for the alleles.
+    i: non-binding and inside, 
+    m: non-binding and transmembrane, 
+    o: non-binding and outside, 
+    I: binding and inside, 
+    M: binding and transmembrane, 
+    O: binding and outside
   }

### 2019-03-08 article.tex
Use generated LaTex table

@@
 
-  \textbf{1. }
-    Epiitopiwef
-    wefwe. \\
-  \textbf{2. }
-    ewfw
-    ewfwf. \\
-  \textbf{3. }
-    ewfqef. \\
-  \textbf{4. }
-    wefwef. \\
+About a fourth of the human proteome is anchored by transmembrane helices (TMHs) to
+lipid membranes. It, therefore, comes as no surprise that
+bioinformatics analysis predicts an over-representation of TMHs among strong MHC
+class I (MHC-I) binders. Published peptide elution studies confirm that TMHs are indeed
+presented by MHC-I. This raises the question how membrane proteins are processed
+for MHC-I (cross-)presentation, with current research focusing on soluble antigens. The
+presentation of membrane-buried peptides is likely important in health and disease, as
+TMHs are considerably conserved and their presentation might prevent escape muta-
+tions by pathogens. Therefore, it could contribute to the disease correlations described
+for many human leukocyte antigen haplotypes.
+
 \end{abstract}
 
-{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, pirouette, R
+{\bf Keywords:} antigen presentation, antigen cross-presentation, membrane proteins, bioinformatics, adaptive
+immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\cite{bianchi2017transmembrane}
 
@@
   }
-  \label{tab:results}
+  \label{tab:results_old}
 \end{table}
+
+\input{table_imoimo.latex}
+% has label tab:results
 
@@
 
-Thanks to Geert van den Bogaart for his help.
+Thanks to Geert van den Bogaart for his help \richel{reword}.
 We thank the Netherlands 

### 2019-03-08 article.tex
Delete old handcrafted table

@@
 
-\begin{table}
-\centering
-  \begin{tabular}{| r | c | c | c | c | c | c |}
-    \hline
-    \textbf{Allele}       & \textbf{i} & \textbf{m} & \textbf{o} & \textbf{I} & \textbf{M} & \textbf{O} \\ 
-    \hline
-    DRB4*0101             & 3752 & 2397 & 62024 & 1061 & 854  & 11952 \\
-    HLA-DPA10103-DPB10402 & 4335 & 2328 & 69040 & 478  & 923  & 4936 \\
-    DQA1*0501/DQB1*0301   & 3321 & 1329 & 45461 & 1492 & 1922 & 28515 \\
-    \hline 
-  \end{tabular}
-  \caption{
-    \richel{These are preliminary, 6 percent of proteome}
-    Location and binding of amino acids for the alleles.
-    i: non-binding and inside, 
-    m: non-binding and transmembrane, 
-    o: non-binding and outside, 
-    I: binding and inside, 
-    M: binding and transmembrane, 
-    O: binding and outside
-  }
-  \label{tab:results_old}
-\end{table}
-
 \input{table_imoimo.latex}

### 2019-03-08 article.tex
Article is getting there

@@
 \usepackage{authblk}
-\title{pirouette: Determining the error BEAST2 makes in inferring a phylogeny}
+\title{
+  Transmembrane helices are also 
+  an overlooked source of major histocompatibility complex class II epitopes
+  and evolutionary more conserved than expected by change
+}
 \author[1]{Rich\`el J.C. Bilderbeek}
@@
 
-About a fourth of the human proteome is anchored by transmembrane helices (TMHs) to
-lipid membranes. It, therefore, comes as no surprise that
-bioinformatics analysis predicts an over-representation of TMHs among strong MHC
-class I (MHC-I) binders. Published peptide elution studies confirm that TMHs are indeed
-presented by MHC-I. This raises the question how membrane proteins are processed
-for MHC-I (cross-)presentation, with current research focusing on soluble antigens. The
-presentation of membrane-buried peptides is likely important in health and disease, as
-TMHs are considerably conserved and their presentation might prevent escape muta-
-tions by pathogens. Therefore, it could contribute to the disease correlations described
-for many human leukocyte antigen haplotypes.
-
+Transmembrane helices (TMHs) are an overlooked source of 
+major histocompatibility complex (MHC) class I epitopes.
+It is unknown why this is the case and 
+whether MHC-II also is an overlooked source of epitopes.
+This study investigates if there is an evolutionary conservation 
+of a pathogen's TMHs and if this is matches 
+with MHC-II's over/equally/under-presenting TMHs.
+We find that ...
 \end{abstract}
 
-{\bf Keywords:} antigen presentation, antigen cross-presentation, membrane proteins, bioinformatics, adaptive
-immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2
+{\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
+adaptive immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\cite{bianchi2017transmembrane}
+For MHC-1, we know that epitopes derived from transmembrane helices (TMHs) are 
+over-presented by all human leukocyte antigen (HLA)-A and most HLA-B super 
+types [\cite{bianchi2017transmembrane}].
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+A hypothesis offered by that same paper is that the presentation of 
+especially TMHs may have an evolutionary advantage for the (human) host.
+The argumentation given is that the presentation of parts of TMHs
+is likelier to detect pathogenic proteins, as TMHs are already conserved
+due to the functional requirement of being able to span a membrane.
+Additionally, there are fewer escape mutations possible for the pathogen,
+as relatively less of these will result in a functional TMH.
+
+However, that argumentation can also go the other way: as there are less
+polypeptide fragments from (the less amount of) possible TMHs, 
+it is likelier that host and pathogen
+share the same protein sequence, 
+resulting the pathogen avoiding detection.
+
+Within the immune response, MHC-I is unique in [doing something].
+The other most important participant in the immune response in MHC-II,
+which is more involved in detecting bacterial pathogens, 
+e.g. tuberculosis, [and others].
+
+From the earlier research done on MHC-I, it naturally follows to check
+if MHC-II are likelier to present TMHs than expected by chance.
+
+Extending the previous study on MHC-I, we also measure the evolutionary
+conservation of TMH epitopes in Mycoplasma pathogens.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-There are thousands of MHC2 alleles known.
-For this research, we picked the three alleles that are most abundant,
-accoring to \richel{reference here}, 
-which are DRB4*0101, 
-HLA-DPA10103-DPB10402 
+Transmembrane helices and strong MHC-II-binding peptides
+were predicted for a tuberculosis reference proteome 
+from \url{https://www.ebi.ac.uk/reference_proteomes}
+using the \verb;epitopeome; 
 \richel{
-  we decided for DPB1*0402, but there are multiple alleles matching.
+  I'd enjoy a better name
+}
+R package [\cite{epitopeome}].
+We picked the three MHC-II alleles that are most abundant 
+in the current human population, 
+which are DRB4*0101, HLA-DPA10103-DPB10402 
+\richel{
+  We decided for DPB1*0402, but there are multiple alleles matching.
   I just picked the first one of the alleles matching, which are:
-  HLA-DPA10103-DPB10402,
-  HLA-DPA10104-DPB10402,
-  HLA-DPA10105-DPB10402,
-  HLA-DPA10106-DPB10402,
-  HLA-DPA10107-DPB10402,
-  HLA-DPA10108-DPB10402,
-  HLA-DPA10109-DPB10402,
-  HLA-DPA10110-DPB10402,
-  HLA-DPA10201-DPB10402,
-  HLA-DPA10202-DPB10402,
-  HLA-DPA10203-DPB10402,
-  HLA-DPA10204-DPB10402,
-  HLA-DPA10301-DPB10402,
-  HLA-DPA10302-DPB10402,
-  HLA-DPA10303-DPB10402,
-  HLA-DPA10401-DPB10402
+  HLA-DPA10x-DPB10402, where x equals 103-110, 201-204, 301-303 or 401
 }
-and DQA1*0501/DQB1*0301,
-that are present in respectively 31, 41 and 59 percent 
-\richel{fix numbers}
-of the current human population.
+and DQA1*0501/DQB1*0301 \richel{reference here}.
+The 5\% peptides with the lowest IC50 values were defined as binders.
+We then simply counted the number of amino acids that were present inside the
+cell, within the membrane or outside of the cell, as well as if it was part 
+of a strong MHC-II binding site, as shown in figue \ref{tab:results}.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The \verb;epitopeome; R package \cite{epitopeome} binds together 
+the \verb;tmhmm; [\cite{tmhmm}] and \verb;netmhc2pan; [\cite{netmhc2pan}] 
+R packages. \verb;tmhmm; provides an R interface to
+TMHMM [\cite{krogh2001predicting}, \cite{sonnhammer1998hidden}], a tool
+to predict where membrane proteins' amino acids are located within the
+membrance.   
+\verb;netmhc2pan; provides an R interface to
+NetMHC2pan [\cite{jensen2018improved}], a tool
+to predict MHC-II binding to proteins.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 % has label tab:results
+\richel{
+  I'd enjoy 
+  (1) a row with 'expected by chance', 
+  (2) using percentages instead,
+  (3) merging inside and outside
+}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Acknowledgements}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Data Accessibility}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Authors' contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % MEE style
@@
 \bibliography{article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Appendix section 1}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2019-03-08 article.tex
First incomplete draft of Methods, some fictional results, etc

@@
   an overlooked source of major histocompatibility complex class II epitopes
-  and evolutionary more conserved than expected by change
+  and evolutionary more conserved than expected by chance
 }
@@
 
-The \verb;epitopeome; R package \cite{epitopeome} binds together 
-the \verb;tmhmm; [\cite{tmhmm}] and \verb;netmhc2pan; [\cite{netmhc2pan}] 
+The \verb;epitopeome; R package \cite{epitopeome} binds \richel{pun intended} 
+together the \verb;tmhmm; [\cite{tmhmm}] and \verb;netmhc2pan; [\cite{netmhc2pan}] 
 R packages. \verb;tmhmm; provides an R interface to
@@
 
+The evolutionary conservation of TMHs was measured from a
+DNA alignment of multiple transmembrane proteins in multitiple
+species of the mycoplasma bacterial family
+\richel{how obtained exactly? how alignment done?}.
+
+Using the \verb;tmhprot; R package [\cite{tmhprot}], the DNA alignment is split
+into two alignments, one for the TMH parts, another for the non-TMH
+parts. Each alignment was tested by the \verb;mcbette; 
+R package [\cite{mcbette}] to select the Bayesian inference model with
+the highest evidence (a.k.a. the marginal likelihood) using the nested
+sampling approach as described in \cite{maturana2018model},
+using the popular Bayesian phylogenetic tool 
+BEAST2 [\cite{bouckaert2014beast}] in the back-end.
+\verb;mcbette; used a set of 40 candidate inference models, 
+consisting of all combinations of 
+4 site models (JC, HKY, TN, GTR), 
+2 clock models (strict, RLN) and 
+5 tree priors (Yule, BD, CBS, CCP, CEP).
+
+For each alignment, the inference model with the highest evidence
+is used in a Bayesian inference. From each Bayesian posterior,
+the parameter estimates regarding mutation rates (including clock rate)
+were obtained and compared to determine which realized mutation rate is lower.
+The Markov chain Monte Carlo was set up in such a way that the effective sample
+size for the likelihood of the inference model is above the recommended value
+of 200 [\cite{bouckaert2014beast}].
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Table \ref{tab:results} shows the location and binding strength for the
+tuberculosis proteome.
 
@@
 }
+
+The inference model with the highest evidence in the
+TMH-only alignment was [yet unknown] and [also yet unknown]
+for the non-TMH alignment. Individual model weights are shown
+in tables \ref{tab:evidences_tmh} 
+and \ref{tab:evidences_non_tmh}.
+
+The Bayesian inference resulted in [a distribution of mutation rates],
+as shown in [absent figure].
+The ESSes of the Bayesian parameter estimates was above 200, exact values
+are shown in tables \ref{tab:esses_tmh} and \ref{tab:esses_non_tmh}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Appendix section 1}
+\section{Model selection}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlllrr}
+  \hline
+ & Site model & Clock model & Tree prior & log(evidence) & Weight \\ 
+  \hline
+1 & JC & Strict & Yule & -8201.109 & 0.002 \\ 
+  2 & JC & Strict & BD & -8203.346 & 0.000 \\ 
+  3 & JC & Strict & CCP & -8195.927 & 0.278 \\ 
+  4 & JC & Strict & CEP & -8194.976 & 0.720 \\ 
+   \hline
+\end{tabular}
+\caption{
+  Evidences for the TMH-only alignment
+  \richel{These are just example values}
+} 
+\label{tab:evidences_tmh}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rlllrr}
+  \hline
+ & Site model & Clock model & Tree prior & log(evidence) & Weight \\ 
+  \hline
+1 & JC & Strict & Yule & -8201.109 & 0.002 \\ 
+  2 & JC & Strict & BD & -8203.346 & 0.000 \\ 
+  3 & JC & Strict & CCP & -8195.927 & 0.278 \\ 
+  4 & JC & Strict & CEP & -8194.976 & 0.720 \\ 
+   \hline
+\end{tabular}
+\caption{
+  Evidences for the non-TMH alignment
+  \richel{These are just example values}
+} 
+\label{tab:evidences_non_tmh}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Inference strength}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrrr}
+  \hline
+ & posterior & likelihood & prior & treeLikelihood & TreeHeight & YuleModel & birthRate \\ 
+  \hline
+1 & 10001 & 10001 & 10001 & 10001 & 9913 & 10001 & 9540 \\ 
+   \hline
+\end{tabular}
+\caption{
+  ESSes of the TMH-only alignment
+  \richel{These are just example values}
+} 
+\label{tab:esses_tmh}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrrr}
+  \hline
+ & posterior & likelihood & prior & treeLikelihood & TreeHeight & YuleModel & birthRate \\ 
+  \hline
+1 & 10001 & 10001 & 10001 & 10001 & 9913 & 10001 & 9540 \\ 
+   \hline
+\end{tabular}
+\caption{
+  ESSes of the non-TMH alignment
+  \richel{These are just example values}
+} 
+\label{tab:esses_non_tmh}
+\end{table}
+
 \end{document}

### 2019-03-08 article.tex
Start of discussion

@@
 
+We conclude that MHC-II binds to TMH peptides with a higher/lower/equal
+probability than expected by chance.
+
+We conclude that the evolutionary conservation if the TMH parts of membrane
+proteins is higher/less/equal compare to its non-TMH counterparts.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We compared the mutation rates between the TMH and non-TMH part of
+multiple mycoplasma species. Where we expect no variation 
+in mutation rate for every TMH amino acid,
+\richel{
+  we can test this, but unsure if that would make sense
+}
+we know that non-TMH part will have regions of different evolutionary
+conservation: functional domains, especially in protein-protein
+interactions will be strongly conserved, due to an even more constrained
+set of peptides that enable a certain function.
 
@@
 Thanks to Geert van den Bogaart for his help \richel{reword}.
+Thanks to Maxim for his help \richel{reword}.
 We thank the Netherlands 
@@
 
-RJCB, FB and RSE conceived the idea for this research. 
+RJCB and FB conceived the idea for this research. 
 RJCB wrote the code.

### 2019-03-08 article.tex
EOD

@@
 of a pathogen's TMHs and if this is matches 
-with MHC-II's over/equally/under-presenting TMHs.
+with MHC-II's binding to TMHs more/less/equal as expected by chance.
 We find that ...

### 2019-03-22 article.tex
Add notes
Feedback Frans

--- 2019-03-08 article.tex
@@
 
-Transmembrane helices (TMHs) are an overlooked source of 
-major histocompatibility complex (MHC) class I epitopes.
-It is unknown why this is the case and 
-whether MHC-II also is an overlooked source of epitopes.
-This study investigates if there is an evolutionary conservation 
-of a pathogen's TMHs and if this is matches 
-with MHC-II's over/equally/under-presenting TMHs.
-We find that ...
+Transmembrane helices (TMHs) are an overrepresented source of epitopes on major histocompatibility complex (MHC) class I due to a preference for hydrophobic for the majority of HLA-I haplotypes.
+It is unknown why there is such a preference and whether  this is the case for MHC-II haplotypes.
+This study investigates if there is an evolutionary conservation of a pathogen’s TMHs and if this correlates with the affinity for binding to MHC-II.
+
+TMHs more/less/equal as expected by chance. We ﬁnd that ...
+
 \end{abstract}
@@
 
-For MHC-1, we know that epitopes derived from transmembrane helices (TMHs) are 
-over-presented by all human leukocyte antigen (HLA)-A and most HLA-B super 
-types [\cite{bianchi2017transmembrane}].
+For MHC-1, it was found that epitopes derived from transmembrane helices (TMHs)
+are over-presented by all human leukocyte antigen (HLA)-A and most HLA-B super 
+types [\cite{bianchi2017transmembrane}]. One explanation is that the 
+presentation of especially TMHs may have an evolutionary advantage for 
+the (human) host. The argumentation for this is that the TMHs are more 
+conserved due to the functional requirement of being able to span a lipid 
+bilayer. Additionally, there are limited escape mutations possible for the 
+pathogen, as many will result in a non-functional TMH. 
+However, that argumentation can also be countered as there are limited unique 
+polypeptide fragments and TMHs are a conserved throughout all species, it is 
+likelier that host and pathogen share the same unique epitope sequences, 
+resulting in the pathogen avoiding detection due to the negative T-cell 
+selection process in the thymus.
 
-A hypothesis offered by that same paper is that the presentation of 
-especially TMHs may have an evolutionary advantage for the (human) host.
-The argumentation given is that the presentation of parts of TMHs
-is likelier to detect pathogenic proteins, as TMHs are already conserved
-due to the functional requirement of being able to span a membrane.
-Additionally, there are fewer escape mutations possible for the pathogen,
-as relatively less of these will result in a functional TMH.
-
-However, that argumentation can also go the other way: as there are less
-polypeptide fragments from (the less amount of) possible TMHs, 
-it is likelier that host and pathogen
-share the same protein sequence, 
-resulting the pathogen avoiding detection.
-
-Within the immune response, MHC-I is unique in [doing something].
-The other most important participant in the immune response in MHC-II,
-which is more involved in detecting bacterial pathogens, 
-e.g. tuberculosis, [and others].
-
-From the earlier research done on MHC-I, it naturally follows to check
-if MHC-II are likelier to present TMHs than expected by chance.
-
+If presentation of TMH’s on MHC-I would bring an evolutionary advantage 
+in the recognition of pathogens by the immune, 
+it would follow that this is equally important for MHC-II. 
+MHC-II, which is necessary for the activation of B-cells after the detection 
+of bacterial pathogens, e.g.  tuberculosis, by dendritic cells or [...]. 
+From the earlier research done on MHC-I, it naturally follows to check if 
+also MHC-II is more likely to present TMHs than expected by chance. 
 Extending the previous study on MHC-I, we also measure the evolutionary 
-conservation of TMH epitopes in Mycoplasma pathogens.
+conservation of TMH epitopes in Mycoplasma pathogens to adress if 
+presentation of TMH derived epitopes has an evolutionary advantage.
 
@@
 
+\richel{
+  Note that most bacteria are opportunistic pathogens.
+  Note that most bacteria are generalists.
+  Note that most bacteria have different cell membranes (and walls), that
+  may have different functional constraints than a human cell membrane
+}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-03-22 article.tex
Wrapped
TeX is nice

@@
 \author[1]{Rich\`el J.C. Bilderbeek}
-\author[1]{?Rampal S. Etienne}
 \author[2]{Frans Bianchi}
-\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands}
+\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of 
+Groningen, Groningen, The Netherlands}
 \affil[2]{Frans' Institute, University of Groningen, Groningen, The Netherlands}
@@
 % Style of listings
-% From http://r.789695.n4.nabble.com/How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
+% From http://r.789695.n4.nabble.com/
+%   How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
 \usepackage{fancyvrb} 
@@
   commentstyle=\color{gray},% comment style
-  % numbers=left,% display line numbers on the left side
   numberstyle=\scriptsize,% use small line numbers
@@
   tabsize=2,% sizes of tabs
-  showstringspaces=false,% do not replace spaces in strings by a certain character
+  showstringspaces=false,
   captionpos=b,% positioning of the caption below
@@
   alsoother={$},% becomes other
-  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},% other keywords
+  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},
   deletekeywords={c}% remove keywords 
@@
 
-% Adds numbered lines
-\usepackage{lineno}
-\linenumbers
-
-% Rename 'Abstract' to 'Summary 
-\usepackage[english]{babel}
-\addto{\captionsenglish}{\renewcommand{\abstractname}{Summary}}
+% Adds numbered lines. Not yet...
+% \usepackage{lineno}
+% \linenumbers
 
@@
 
-Transmembrane helices (TMHs) are an overrepresented source of epitopes on major histocompatibility complex (MHC) class I due to a preference for hydrophobic for the majority of HLA-I haplotypes.
-It is unknown why there is such a preference and whether  this is the case for MHC-II haplotypes.
-This study investigates if there is an evolutionary conservation of a pathogen’s TMHs and if this correlates with the affinity for binding to MHC-II.
+Transmembrane helices (TMHs) are an overrepresented source of epitopes on major 
+histocompatibility complex (MHC) class I due to a preference for hydrophobic 
+for the majority of HLA-I haplotypes. It is unknown why there is such a 
+preference and whether  this is the case for MHC-II haplotypes. This study 
+investigates if there is an evolutionary conservation of a pathogen’s TMHs 
+and if this correlates with the affinity for binding to MHC-II.
 
@@
 The \verb;epitopeome; R package \cite{epitopeome} binds \richel{pun intended} 
-together the \verb;tmhmm; [\cite{tmhmm}] and \verb;netmhc2pan; [\cite{netmhc2pan}] 
+together the \verb;tmhmm; [\cite{tmhmm}] and \verb;netmhc2pan; 
+[\cite{netmhc2pan}] 
 R packages. \verb;tmhmm; provides an R interface to
@@
   \hline
- & posterior & likelihood & prior & treeLikelihood & TreeHeight & YuleModel & birthRate \\ 
+ & posterior & likelihood & prior & treeLikelihood & TreeHeight & YuleModel & 
+birthRate \\ 
   \hline

### 2020-04-07 article.tex
Add figure 1
Fix layout
EOD
Fine abstract

@@
 \newcommand{\richel}[1]{\textcolor{orange}{\textbf{[RB: #1]}}}
-\newcommand{\rampal}[1]{\textcolor{green}{\textbf{[RSE: #1]}}}
 
@@
 Transmembrane helices (TMHs) are an overrepresented source of epitopes on major 
-histocompatibility complex (MHC) class I due to a preference for hydrophobic 
-for the majority of HLA-I haplotypes. It is unknown why there is such a 
-preference and whether  this is the case for MHC-II haplotypes. This study 
-investigates if there is an evolutionary conservation of a pathogen’s TMHs 
-and if this correlates with the affinity for binding to MHC-II.
-
-TMHs more/less/equal as expected by chance. We ﬁnd that ...
+histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
+It is unknown why there is such a preference exists and 
+whether this is the case for MHC-II haplotypes. 
+This study shows that also MHC-II binds to polypeptides derived from TMHs 
+more often than expected by chance. It is unknown if a signal of
+natural selection is present in the proteome coding for TMHs.
+This study shows that in the mutation rate in TMHs of [a pathogen] 
+are [less/equal/more], hinting at 
+evolutionary [conservation/neutrality/disruptive selection].
+Our findings suggest that developing vaccines targeting the TMHs of
+pathogens is a [good/neutral/bad] idea.
 
@@
 presentation of TMH derived epitopes has an evolutionary advantage.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Hypotheses}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{itemize}
+  \item $\mathcal{H}_1$: MHC-II binds as often to TMHs as the cytosolic and extracellular
+    parts of a pathogen's transmembrane proteins
+  \item $\mathcal{H}_2$: The TMHs of [a pathogen] have the same mutation rate 
+    as the cytosolic and extracellular
+    parts of its transmembrane proteins
+\end{itemize}
 
@@
 
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{figure_1.png}
+  \label{fig:1}
+\end{figure}
+
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{figure_1_5.png}
+  \label{fig:1_5}
+\end{figure}
+
+\begin{figure}[ht]
+  \includegraphics[width=\textwidth]{figure_3.png}
+  \label{fig:3}
+\end{figure}
+
 \input{table_imoimo.latex}
@@
 
-Thanks to Geert van den Bogaart for his help \richel{reword}.
-Thanks to Maxim for his help \richel{reword}.
-We thank the Netherlands 
-Organization for Scientific Research (NWO) for financial support 
-through a VICI grant awarded to RSE.
+We thanks to Geert van den Bogaart for his wisdom.
 
@@
 RJCB wrote the code.
-FB supplied the DNA alignment.
-RJCB, FB and RSE contributed to writing the article.
+RJCB and FB wrote the article.
 

### 2020-04-07 create_article.sh

@@ -1,9 +1,5 @@
 #!/bin/bash
 
-# cd ../figures
-# ./create_figures
-# cd ../article
-
 pdflatex article.tex --shell-escape
 bibtex article.aux
 pdflatex article.tex --shell-escape

### 2020-04-07 to_frans.sh

@@ -1,6 +1,5 @@
 #!/bin/bash
 pandoc article.tex -o article.odt
-pandoc article.tex -o article.doc
 pandoc article.tex -o article.docx
 pandoc article.tex -o article.md
 pandoc article.tex -o article.txt

### 2020-04-07 to_80_chars.sh
Update script

@@ -1,4 +1,11 @@
 #!/bin/bash
+#
+# Convert article.tex to have 80 characters per line
+#
+# Usage:
+#
+#   ./to_80_chars.sh
+#
 fold -w 80 -s article.tex > article_new.tex
 cp article_new.tex article.tex
 rm article_new.tex

### 2020-04-07 article.tex
EOD

@@
 pathogen, as many will result in a non-functional TMH.
+
 However, that argumentation can also be countered as there are limited unique 
@@
 resulting in the pathogen avoiding detection due to the negative T-cell 
-selection process in the thymus.
+selection process in the thymus \richel{Frans, jij beweerde dat bacterien
+en mensen compleet verschillende TMH sequenties hebben. Als ik dat goed
+onthouden hebt, klopt deze paragraaf dus niet}
 

### 2020-04-09 article.tex
Add hypothesis

@@
 {\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
-adaptive immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2
+adaptive immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2,
+COVID-19
 
@@
 \begin{itemize}
-  \item $\mathcal{H}_1$: MHC-II binds as often to TMHs as the cytosolic and extracellular
+  \item $\mathcal{H}_1$: MHC-I has the same percentage of epitopes overlapping
+    (that is 5.73 to 5.59 percent, with a confidence interval of 99 percent) 
+    with TMHs in COVID-19 as in 
+4 + ((117 / 135) * 2) = 5.73
+4 + ((107 / 135) * 2) = 5.59
+
+  \item $\mathcal{H}_2$: MHC-II binds as often to TMHs as the cytosolic and extracellular
     parts of a pathogen's transmembrane proteins
-  \item $\mathcal{H}_2$: The TMHs of [a pathogen] have the same mutation rate 
+  \item $\mathcal{H}_3$: The TMHs of [a pathogen] have the same mutation rate 
     as the cytosolic and extracellular

### 2020-04-09 article.tex
Hypotheses

@@
 For MHC-1, it was found that epitopes derived from transmembrane helices (TMHs)
-are over-presented by all human leukocyte antigen (HLA)-A and most HLA-B super 
-types [\cite{bianchi2017transmembrane}]. One explanation is that the 
+are over-presented by all human leukocyte antigen (HLA)-A and 
+most HLA-B super types (\cite{bianchi2017}). 
+One explanation is that the 
 presentation of especially TMHs may have an evolutionary advantage for 
@@
   \item $\mathcal{H}_1$: MHC-I has the same percentage of epitopes overlapping
-    (that is 5.73 to 5.59 percent, with a confidence interval of 99 percent) 
-    with TMHs in COVID-19 as in 
-4 + ((117 / 135) * 2) = 5.73
-4 + ((107 / 135) * 2) = 5.59
-
-  \item $\mathcal{H}_2$: MHC-II binds as often to TMHs as the cytosolic and extracellular
-    parts of a pathogen's transmembrane proteins
-  \item $\mathcal{H}_3$: The TMHs of [a pathogen] have the same mutation rate 
-    as the cytosolic and extracellular
-    parts of its transmembrane proteins
+    with TMHs in Homo sapiens as in COVID-19
+  \item $\mathcal{H}_2$: MHC-II has the same percentage of epitopes overlapping
+    with TMHs as expected by chance
+  \item $\mathcal{H}_3$: The TMHs of COVID-19 have the same mutation rate 
+    as the cytosolic and extracellular parts of its transmembrane proteins
 \end{itemize}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{MHC-I}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We used the same analysis as \cite{bianchi2017},
+except that instead of a human reference proteome,
+we used the proteome of the first sequenced COVID-19 strain (\cite{wu2020new}),
+with a GenBank ID of MN908947.3, as can be downloaded 
+from \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947}.
+
+Bianchi and colleagues obtained a distribution of 
+percentages of MHC-I epitopes overlap with TMHs in Homo sapiens
+for the different HLA haplotypes, with has an average of 5.3\%.
+We obtained a similar distribution of percentages of MHC-I epitopes that 
+overlap with TMHs for the different HLA haplotypes, but then applied to
+COVID-19.
+
+We compare these distribution using a two-sample 
+Kolmogorov-Smirnov (KS) test. The KS test determines if two samples
+are derived from a same distribution, without making assumptions
+regarding the shape of that distribution. We will reject
+the null hypothesis that MHC-I has the same percentage of epitopes 
+overlapping with TMHs in Homo sapiens as in COVID-19 when 
+the KS statistic follows the relationship as show in 
+figure \ref{fig:ks}, for an $\alpha$ of $0.05$.
+is below
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.4\textwidth]{pics/ks.png}
+  \caption{
+    Condition of a Kolmogorov-Smirnov (KS) statistic to rejects a null
+    hypothesis. $D_{n,m}$ is the KS statistic, $n = m$ is the size
+    of the samples, in this case, the number of HLA haplotypes ($n = m = 13$).
+    $\alpha = 0.05$ is the significance level.
+  }
+  \label{fig:ks}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{MHC-2}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Transmembrane helices and strong MHC-II-binding peptides
@@
 from \url{https://www.ebi.ac.uk/reference_proteomes}
-using the \verb;epitopeome; 
-\richel{
-  I'd enjoy a better name
-}
-R package [\cite{epitopeome}].
+using the \verb;epitopeome; R package (\cite{epitopeome}).
 We picked the three MHC-II alleles that are most abundant 
@@
 
-The \verb;epitopeome; R package \cite{epitopeome} binds \richel{pun intended} 
-together the \verb;tmhmm; [\cite{tmhmm}] and \verb;netmhc2pan; 
-[\cite{netmhc2pan}] 
+The \verb;epitopeome; R package (\cite{epitopeome}) binds 
+together the \verb;tmhmm; (\cite{tmhmm}) and \verb;netmhc2pan; 
+(\cite{netmhc2pan}) 
 R packages. \verb;tmhmm; provides an R interface to
-TMHMM [\cite{krogh2001predicting}, \cite{sonnhammer1998hidden}], a tool
+TMHMM (\cite{krogh2001predicting, sonnhammer1998hidden}), a tool
 to predict where membrane proteins' amino acids are located within the
@@
 \verb;netmhc2pan; provides an R interface to
-NetMHC2pan [\cite{jensen2018improved}], a tool
+NetMHC2pan (\cite{jensen2018improved}), a tool
 to predict MHC-II binding to proteins.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evolutionary conservation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-Using the \verb;tmhprot; R package [\cite{tmhprot}], the DNA alignment is split
+Using the \verb;tmhprot; R package (\cite{tmhprot}), the DNA alignment is split
 into two alignments, one for the TMH parts, another for the non-TMH
 parts. Each alignment was tested by the \verb;mcbette; 
-R package [\cite{mcbette}] to select the Bayesian inference model with
+R package (\cite{mcbette}) to select the Bayesian inference model with
 the highest evidence (a.k.a. the marginal likelihood) using the nested
@@
 using the popular Bayesian phylogenetic tool 
-BEAST2 [\cite{bouckaert2014beast}] in the back-end.
+BEAST2 (\cite{bouckaert2014beast}) in the back-end.
 \verb;mcbette; used a set of 40 candidate inference models, 
@@
 size for the likelihood of the inference model is above the recommended value
-of 200 [\cite{bouckaert2014beast}].
+of 200 (\cite{bouckaert2014beast}).
 
@@
 \section{Results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{MHC-I}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{MHC-II}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evolutionary conservation}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-09 view.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+qpdfview article.pdf

### 2020-04-09 edit.sh
Show PDF if TeX is edited

@@ -8,3 +8,4 @@
 #
 
 mousepad article.tex
+make

### 2020-04-09 to_frans.sh
Can 'make frans'

@@ -4,4 +4,4 @@ pandoc article.tex -o article.docx
 pandoc article.tex -o article.md
 pandoc article.tex -o article.txt
 pandoc article.tex -o article.rtf
-zip bbbq.zip article.pdf article.odt article.doc article.docx article.md article.txt article.rtf article.tex article.bib mee.bst
+zip bbbq.zip article.pdf article.odt article.docx article.md article.txt article.rtf article.tex article.bib mee.bst

### 2020-04-09 ks/ks.Rmd
Can create simulated KS figure

---
title: "Kolmogorov-Smirnov"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(testthat)
library(ggplot2)
```

Human samples:

```{r}
set.seed(314)
# n: sample size, equals the number of HLA haplotypes
n <- 13
# f denotes 'fraction', a value between 0.0 and 1.0
f_human <- rnorm(n = n, mean = 0.053, sd = 0.01)
f_covid <- rnorm(n = n, mean = runif(n = 1, min = 0.03, max = 0.1), sd = 0.01)
expect_true(all(f_human > 0.0))
expect_true(all(f_human < 1.0))
expect_true(all(f_covid > 0.0))
expect_true(all(f_covid < 1.0))
```

Plot:

```{r}
df <- tibble::tibble(
  proteome = rep(c("human", "covid"), each = n),
  haplotype = rep(LETTERS[1:n], time = 2),
  f = c(f_human, f_covid)
)
df$proteome <- as.factor(df$proteome)
df$haplotype <- as.factor(df$haplotype)
ggplot(df, aes(x = haplotype, y = f, fill = proteome)) + 
  scale_fill_manual(values = c("human" = "#ffffff", "covid" = "#cccccc")) +
  geom_col(position = position_dodge(), color = "#000000") + xlab("HLA haplotype") + 
  ylab("% epitopes overlapping \nwith transmembrane helix") +
  geom_hline(yintercept = mean(f_human), lty = "dashed") +
  geom_hline(yintercept = mean(f_covid), lty = "dotted") +
  labs(
    title = "SIMULATED DATA!",
    caption = paste0(
      "Dashed line: average in humans: ", formatC(mean(f_human), digits = 2),". ",
      "Dotted line: average in COVID-19: ", formatC(mean(f_covid), digits = 2),"."  
    )
  ) + ggsave("fig_1.png", width = 7, height = 7)
```

Test:

### 2020-04-09 ks/render.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+Rscript -e 'rmarkdown::render("ks.Rmd", output_file = "ks.html")'

### 2020-04-09 ks/ks.Rmd
Can test KS

@@
 ```{r}
-set.seed(314)
 # n: sample size, equals the number of HLA haplotypes
@@
 
-Plot:
+Create table:
 
@@
 df$haplotype <- as.factor(df$haplotype)
+```
+
+Show table:
+
+```{r}
+knitr::kable(df)
+```
+
+Save table:
+
+```{r}
+sink("table_1.latex")
+  xtable::print.xtable(
+    xtable::xtable(
+      df,
+      caption = paste0("Percentage of epitopes overlapping \nwith transmembrane helix"),
+      digits = 2
+    ),
+    include.rownames = FALSE
+  )
+sink()
+```
+
+Test:
+
+```{r}
+t <- ks.test(x = f_human, y = f_covid)
+print(t)
+```
+
+Stats:
+
+```{r}
+alpha <- 0.05
+if (t$p.value <= alpha) {
+  print("Reject that the two samples are taken from a same distribution") 
+} else {
+  print("Accept that the two samples are taken from a same distribution")
+}
+```
+
+Plot:
+
+```{r}
 ggplot(df, aes(x = haplotype, y = f, fill = proteome)) + 
@@
 ```
-
-Test:

### 2020-04-09 article.tex
Show PDF if TeX is edited
Make figure 1 and table

@@
 \begin{figure}[!htbp]
-  \includegraphics[width=0.4\textwidth]{pics/ks.png}
+  \includegraphics[width=0.4\textwidth]{ks/fig_1.png}
   \caption{
@@
 
+\iffalse
 	Transmembrane helices and strong MHC-II-binding peptides
@@
 	to predict MHC-II binding to proteins.
+\fi
 
@@
 
+\iffalse
 	The evolutionary conservation of TMHs was measured from a
@@
 	of 200 (\cite{bouckaert2014beast}).
+\fi
 
@@
 
+\iffalse
 	Table \ref{tab:results} shows the location and binding strength for the
@@
 	are shown in tables \ref{tab:esses_tmh} and \ref{tab:esses_non_tmh}.
+\fi
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix
+\section{Supplementary materials}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Model selection}
+\subsection{Binding}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlllrr}
-  \hline
- & Site model & Clock model & Tree prior & log(evidence) & Weight \\ 
-  \hline
-1 & JC & Strict & Yule & -8201.109 & 0.002 \\ 
-  2 & JC & Strict & BD & -8203.346 & 0.000 \\ 
-  3 & JC & Strict & CCP & -8195.927 & 0.278 \\ 
-  4 & JC & Strict & CEP & -8194.976 & 0.720 \\ 
-   \hline
-\end{tabular}
-\caption{
-  Evidences for the TMH-only alignment
-  \richel{These are just example values}
-} 
-\label{tab:evidences_tmh}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rlllrr}
-  \hline
- & Site model & Clock model & Tree prior & log(evidence) & Weight \\ 
-  \hline
-1 & JC & Strict & Yule & -8201.109 & 0.002 \\ 
-  2 & JC & Strict & BD & -8203.346 & 0.000 \\ 
-  3 & JC & Strict & CCP & -8195.927 & 0.278 \\ 
-  4 & JC & Strict & CEP & -8194.976 & 0.720 \\ 
-   \hline
-\end{tabular}
-\caption{
-  Evidences for the non-TMH alignment
-  \richel{These are just example values}
-} 
-\label{tab:evidences_non_tmh}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Inference strength}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & posterior & likelihood & prior & treeLikelihood & TreeHeight & YuleModel & 
-birthRate \\ 
-  \hline
-1 & 10001 & 10001 & 10001 & 10001 & 9913 & 10001 & 9540 \\ 
-   \hline
-\end{tabular}
-\caption{
-  ESSes of the TMH-only alignment
-  \richel{These are just example values}
-} 
-\label{tab:esses_tmh}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrrrrr}
-  \hline
- & posterior & likelihood & prior & treeLikelihood & TreeHeight & YuleModel & 
-birthRate \\ 
-  \hline
-1 & 10001 & 10001 & 10001 & 10001 & 9913 & 10001 & 9540 \\ 
-   \hline
-\end{tabular}
-\caption{
-  ESSes of the non-TMH alignment
-  \richel{These are just example values}
-} 
-\label{tab:esses_non_tmh}
-\end{table}
+\input{ks/table_1.latex}
 

### 2020-04-09 ks/ks.Rmd
Make figure 1 and table

@@
       caption = paste0("Percentage of epitopes overlapping \nwith transmembrane helix"),
-      digits = 2
+      digits = 3
     ),

### 2020-04-09 ks/render.sh

@@ -1,2 +1,2 @@
 #!/bin/bash
-Rscript -e 'rmarkdown::render("ks.Rmd", output_file = "ks.html")'
+Rscript -e 'rmarkdown::render("ks.Rmd")'

### 2020-04-09 ks/ks.Rmd
Use 'write.csv' instead of 'sink', as it renders better
Watermark artificial data

@@
 ```{r}
-sink("table_1.latex")
-  xtable::print.xtable(
-    xtable::xtable(
-      df,
-      caption = paste0("Percentage of epitopes overlapping \nwith transmembrane helix"),
-      digits = 3
-    ),
-    include.rownames = FALSE
-  )
-sink()
+write.csv(df, "table_1.csv", row.names = FALSE)
 ```
@@
   labs(
-    title = "SIMULATED DATA!",
     caption = paste0(

### 2020-04-09 ks/render.sh

@@ -1,2 +1,5 @@
 #!/bin/bash
 Rscript -e 'rmarkdown::render("ks.Rmd")'
+composite -tile -dissolve 5% -gravity center watermark.jpeg fig_1.png fig_1_watermarked.png
+
+

### 2020-04-09 article.tex
Part 1 is nice
Watermark artificial data
Only MHC-I for now

@@
 \begin{figure}[!htbp]
-  \includegraphics[width=0.4\textwidth]{ks/fig_1.png}
+  \includegraphics[width=0.4\textwidth]{pics/ks.png}
   \caption{
@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ks/fig_1_watermarked.png}
+  \caption{
+    \% epitopes overlapping with transmembrane helix
+    for a human and COVID-19 proteome
+  }
+  \label{fig:1}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\iffalse
 	We conclude that MHC-II binds to TMH peptides with a higher/lower/equal
@@
 	proteins is higher/less/equal compare to its non-TMH counterparts.
+\fi
 
@@
 
+\iffalse
 	We compared the mutation rates between the TMH and non-TMH part of
@@
 	}
+\fi
 

### 2020-04-09 article.tex
Completed 1st part

@@
 
-Transmembrane helices (TMHs) are an overrepresented source of epitopes on major 
+Transmembrane helices (TMHs) in the human proteome
+are an overrepresented source of epitopes on major 
 histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
-It is unknown why there is such a preference exists and 
-whether this is the case for MHC-II haplotypes. 
-This study shows that also MHC-II binds to polypeptides derived from TMHs 
-more often than expected by chance. It is unknown if a signal of
-natural selection is present in the proteome coding for TMHs.
-This study shows that in the mutation rate in TMHs of [a pathogen] 
-are [less/equal/more], hinting at 
+This finding may be a methodological artifact and using the proteome
+of an actual pathogen nullifies this finding.
+Additionally, it is unknown if TMH-II is also likelier to present
+TMHs than expected by chance. 
+Lastly, if the immune system if more tailored to detect polypeptides
+derived from TMHs, can we confirm this by detecting the evolutionary
+conservation of this behavior?
+This study shows that MHC-I [also has/does not have] more
+epitopes derived from a TMH for a pathogen proteome, when compared with
+a host proteome.
+Additionally, MHC-II binds to polypeptides derived from TMHs 
+[less/equally/more] often than expected by chance.
+Lastly, this study shows that in the mutation rate in TMHs of a pathogen
+is [lower/as much/higher], hinting at 
 evolutionary [conservation/neutrality/disruptive selection].
-Our findings suggest that developing vaccines targeting the TMHs of
-pathogens is a [good/neutral/bad] idea.
+Our findings suggest that the immune system is [less/neutral/more]
+vigilant to TMHs than expected by chance and this has left [a/no]
+signal in the evolutionary history of the pathogen.
 
@@
 
-For MHC-1, it was found that epitopes derived from transmembrane helices (TMHs)
+For MHC-1, it was found that epitopes derived 
+from human transmembrane helices (TMHs)
 are over-presented by all human leukocyte antigen (HLA)-A and 
 most HLA-B super types (\cite{bianchi2017}). 
-One explanation is that the 
-presentation of especially TMHs may have an evolutionary advantage for 
-the (human) host. The argumentation for this is that the TMHs are more 
+One explanation is that the presentation of TMHs 
+may have an evolutionary advantage for 
+the (human) host, as TMHs may be more 
 conserved due to the functional requirement of being able to span a lipid 
-bilayer. Additionally, there are limited escape mutations possible for the 
-pathogen, as many will result in a non-functional TMH.
+bilayer. 
+Due to this, pathogens have a lower chance to develop an escape mutation,
+as many mutations will result in a disfunctional TMH.
 
-However, that argumentation can also be countered as there are limited unique 
-polypeptide fragments and TMHs are a conserved throughout all species, it is 
-likelier that host and pathogen share the same unique epitope sequences, 
-resulting in the pathogen avoiding detection due to the negative T-cell 
-selection process in the thymus \richel{Frans, jij beweerde dat bacterien
-en mensen compleet verschillende TMH sequenties hebben. Als ik dat goed
-onthouden hebt, klopt deze paragraaf dus niet}
-
-If presentation of TMH’s on MHC-I would bring an evolutionary advantage 
+If presentation of TMHs on MHC-I would bring an evolutionary advantage 
 in the recognition of pathogens by the immune, 
 it would follow that this is equally important for MHC-II.
-MHC-II, which is necessary for the activation of B-cells after the detection 
-of bacterial pathogens, e.g.  tuberculosis, by dendritic cells or [...]. 
-From the earlier research done on MHC-I, it naturally follows to check if 
-also MHC-II is more likely to present TMHs than expected by chance. 
-Extending the previous study on MHC-I, we also measure the evolutionary 
-conservation of TMH epitopes in Mycoplasma pathogens to adress if 
-presentation of TMH derived epitopes has an evolutionary advantage.
+
+If the human immune system is more prone to detect polypeptides
+derived from a pathogen's TMHs, one may expect to find an evolutionary
+signal in fast-evolving pathogen. 
+We can assume that only after the moment that COVID-19 spilled over 
+to humans, it's the human immune system that is an influential component 
+of its evolutionary selection.
+Using the extensive amount of time-dated COVID-19 sequences,
+we can quantify to which extent selection in COVID-19 occurs
+in the transmembrane part of its proteins.
 
@@
 
+Figure \ref{fig:1} shows the percentages of epitopes overlapping 
+with TMHs for a human and COVID-19 proteome.
+The KS test to determine if these percentages are sampled from
+a same distribution resulted in a p-value of \richel{unknown},
+which makes us reject/accept the hypothesis that these percentages
+are sampled from a same distribution.
+
 \begin{figure}[!htbp]
@@
 
+We found that the percentages of epitopes overlapping 
+with TMHs for a human and COVID-19 proteome are 
+[similar/different]. In other words, the
+epitopes that MHC-I presents are [as/not as] likely 
+to be derived from TMH within either a human host and its viral pathogen.
+
 \iffalse
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We concluded that the
+epitopes that MHC-I presents are [as/not as] likely 
+to be derived from TMH within either a human host and its viral pathogen.
+Because the full COVID-19 only has 4 TMHs \richel{check}, the percentages
+of HMC-I epitopes being part of a TMH are likelier to be affected by
+stochasticity. We chose to use COVID-19 regardless, as the thousands
+of its time-dated genomic sequences are ideal for determining the 
+evolutionary conservation of MHC-I detecting TMHs.
 

### 2020-04-09 ks/ks.Rmd
Use percentages in fig 1

@@
 title: "Kolmogorov-Smirnov"
-output: html_document
+author: Richel J.C. Bilderbeek
+output: pdf_document
 ---
@@
 
-```{r}
+```{r, message=FALSE}
 library(testthat)
+library(tidyr)
 library(ggplot2)
@@
 
-Create table:
+Create tidy table:
 
@@
 
+Create wide (untidy) table:
+
+```{r}
+df_wide <- spread(df, proteome, f)
+```
+
 Show table:
@@
 ```{r}
-knitr::kable(df)
+knitr::kable(df_wide, caption = "Fraction of epitopes overlapping with transmembrane helix")
 ```
@@
 ```{r}
-write.csv(df, "table_1.csv", row.names = FALSE)
+write.csv(df_wide, "table_1.csv", row.names = FALSE)
 ```
@@
   geom_col(position = position_dodge(), color = "#000000") + xlab("HLA haplotype") + 
-  ylab("% epitopes overlapping \nwith transmembrane helix") +
+  ylab("Epitopes overlapping \nwith transmembrane helix") +
+  scale_y_continuous(labels = scales::percent_format(accuracy = 3)) +
   geom_hline(yintercept = mean(f_human), lty = "dashed") +
@@
     caption = paste0(
-      "Dashed line: average in humans: ", formatC(mean(f_human), digits = 2),". ",
-      "Dotted line: average in COVID-19: ", formatC(mean(f_covid), digits = 2),"."  
+      "Dashed line: average in humans: ", formatC(100.0 * mean(f_human), digits = 2),"%. ",
+      "Dotted line: average in COVID-19: ", formatC(100.0 * mean(f_covid), digits = 2),"%."  
     )

### 2020-04-09 ks/ks.Rmd
Use wide table with percentages

@@
 df_wide <- spread(df, proteome, f)
+df_wide$covid <- formatC(100.0 * df_wide$covid, digits = 3)
+df_wide$human <- formatC(100.0 * df_wide$human, digits = 3)
 ```
@@
 ```{r}
-knitr::kable(df_wide, caption = "Fraction of epitopes overlapping with transmembrane helix")
+knitr::kable(df_wide, caption = "Percentage of epitopes overlapping with transmembrane helix")
 ```

### 2020-04-09 article.tex
Spell check

@@
 }
-\author[1]{Rich\`el J.C. Bilderbeek}
+\author[1]{Richèl J.C. Bilderbeek}
 \author[2]{Frans Bianchi}
@@
 histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
-This finding may be a methodological artifact and using the proteome
-of an actual pathogen nullifies this finding.
+It is unknown if this same overrepresentation
+is also the case for the proteome of a pathogen.
 Additionally, it is unknown if TMH-II is also likelier to present
@@
 {\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
-adaptive immunity, transmembrane domain, epitopes, T lymphocyte, MHC-2,
-COVID-19
+adaptive immunity, transmembrane domain, epitopes, T lymphocyte, 
+MHC-1, MHC-I, MHC-2, MHC-II, COVID-19
 
@@
 
-For MHC-1, it was found that epitopes derived 
+For MHC-I, it was found that epitopes derived 
 from human transmembrane helices (TMHs)
@@
 	cell, within the membrane or outside of the cell, as well as if it was part 
-	of a strong MHC-II binding site, as shown in figue \ref{tab:results}.
+	of a strong MHC-II binding site, as shown in figure \ref{tab:results}.
 
@@
 	to predict where membrane proteins' amino acids are located within the
-	membrance.   
+	membrane.   
 	\verb;netmhc2pan; provides an R interface to
@@
 	The evolutionary conservation of TMHs was measured from a
-	DNA alignment of multiple transmembrane proteins in multitiple
+	DNA alignment of multiple transmembrane proteins in multiple
 	species of the mycoplasma bacterial family
@@
 Because the full COVID-19 only has 4 TMHs \richel{check}, the percentages
-of HMC-I epitopes being part of a TMH are likelier to be affected by
+of MHC-I epitopes being part of a TMH are likelier to be affected by
 stochasticity. We chose to use COVID-19 regardless, as the thousands
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgements}
+\section{Acknowledgments}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\begin{table}
   \input{ks/table_1.latex}
+  \caption{Percentage of MHC-I epitopes overlapping with transmembrane helix}
+  \label{table:1}
+\end{table}
 

### 2020-04-09 ks/ks.Rmd
Spell check
Render KS to supmat

@@
 
+Test:
+
+```{r}
+t <- ks.test(x = f_human, y = f_covid)
+print(t)
+```
+
+Stats:
+
+```{r}
+alpha <- 0.05
+if (t$p.value <= alpha) {
+  verdict <- "Reject that the two samples are taken from a same distribution"
+} else {
+  verdict <- "Accept that the two samples are taken from a same distribution"
+}
+print(verdict)
+```
+
 Create tidy table:
@@
 
-Create wide (untidy) table:
+## Stats output
 
 ```{r}
-df_wide <- spread(df, proteome, f)
-df_wide$covid <- formatC(100.0 * df_wide$covid, digits = 3)
-df_wide$human <- formatC(100.0 * df_wide$human, digits = 3)
+names(t)
+df_t_wide <- data.frame(
+  statistic = t$statistic,
+  p_value = t$p.value,
+  alternative = t$alternative,
+  method = t$method,
+  data_name = t$data.name,
+  alpha = alpha,
+  verdict = verdict
+)
+df_t <- t(df_t_wide)
+values <- df_t[,1]
+param_names <- rownames(df_t)
+df_t <- data.frame(
+  parameter = param_names,
+  value = values
+)
+
+knitr::kable(df_t, row.names = FALSE)
 ```
 
-Show table:
-
 ```{r}
-knitr::kable(df_wide, caption = "Percentage of epitopes overlapping with transmembrane helix")
+write.csv(df_t, "ks.csv", row.names = FALSE)
 ```
 
-Save table:
-
-```{r}
-write.csv(df_wide, "table_1.csv", row.names = FALSE)
-```
-
-Test:
-
-```{r}
-t <- ks.test(x = f_human, y = f_covid)
-print(t)
-```
-
-Stats:
-
-```{r}
-alpha <- 0.05
-if (t$p.value <= alpha) {
-  print("Reject that the two samples are taken from a same distribution") 
-} else {
-  print("Accept that the two samples are taken from a same distribution")
-}
-```
-
-Plot:
+## Plot
 
@@
 ```
+
+## Table
+
+Create wide (untidy) table:
+
+```{r}
+df_wide <- spread(df, proteome, f)
+df_wide$covid <- formatC(100.0 * df_wide$covid, digits = 3)
+df_wide$human <- formatC(100.0 * df_wide$human, digits = 3)
+```
+
+Show table:
+
+```{r}
+knitr::kable(df_wide, caption = "Percentage of MHC-I epitopes overlapping with transmembrane helix")
+```
+
+Save table:
+
+```{r}
+write.csv(df_wide, "table_1.csv", row.names = FALSE)
+```

### 2020-04-09 ks/render.sh

@@ -1,5 +1,5 @@
 #!/bin/bash
 Rscript -e 'rmarkdown::render("ks.Rmd")'
-composite -tile -dissolve 5% -gravity center watermark.jpeg fig_1.png fig_1_watermarked.png
+# composite -tile -dissolve 5% -gravity center watermark.jpeg fig_1.png fig_1_watermarked.png
 
 

### 2020-04-14 ks/render.sh
Use bbbq_1 as sub-repo

@@ -1,3 +1,0 @@
-#!/bin/bash
-Rscript -e 'rmarkdown::render("ks.Rmd")'
-

### 2020-04-14 zip.sh
Zip on the things Frans needs

@@ -1,7 +1,3 @@
 #!/bin/bash
-pandoc article.tex -o article.odt
 pandoc article.tex -o article.docx
-pandoc article.tex -o article.md
-pandoc article.tex -o article.txt
-pandoc article.tex -o article.rtf
-zip bbbq.zip article.pdf article.odt article.docx article.md article.txt article.rtf article.tex article.bib mee.bst
+zip bbbq.zip article.pdf article.docx

### 2020-04-14 article.tex
Add paragraphs
Render KS to supmat
Use bbbq_1 as sub-repo
Show simulated data more clearly
Add KS equation
Use KS equation

@@
 
+\paragraph{MHC-I}
+
 For MHC-I, it was found that epitopes derived 
@@
 
+\paragraph{MHC-II}
+
 If presentation of TMHs on MHC-I would bring an evolutionary advantage 
@@
 it would follow that this is equally important for MHC-II.
+
+\paragraph{Evolution}
 
@@
 are derived from a same distribution, without making assumptions
-regarding the shape of that distribution. We will reject
+regarding the shape of that distribution.
+
+We will reject
 the null hypothesis that MHC-I has the same percentage of epitopes 
 overlapping with TMHs in Homo sapiens as in COVID-19 when 
-the KS statistic follows the relationship as show in 
-figure \ref{fig:ks}, for an $\alpha$ of $0.05$.
-is below
+the KS statistic $D_{n,m}$ follows the relationship as shown in 
+equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
+and $n = m = 13$ HLA haplotypes.
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.4\textwidth]{pics/ks.png}
-  \caption{
-    Condition of a Kolmogorov-Smirnov (KS) statistic to rejects a null
-    hypothesis. $D_{n,m}$ is the KS statistic, $n = m$ is the size
-    of the samples, in this case, the number of HLA haplotypes ($n = m = 13$).
-    $\alpha = 0.05$ is the significance level.
-  }
-  \label{fig:ks}
-\end{figure}
+\begin{equation}
+   D_{n,m} > \frac{1}{\sqrt{n}} \cdot \sqrt{ -\ln(\frac{\alpha}{2}) \cdot \frac{1 + \frac{n}{m}}{2} }
+\  \label{eq:ks}
+\end{equation}
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ks/fig_1_watermarked.png}
+  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1_watermarked.png}
   \caption{
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Binding}
+\subsection{MHC-I}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{table}
-  \input{ks/table_1.latex}
-  \caption{Percentage of MHC-I epitopes overlapping with transmembrane helix}
+  \input{bbbq_1/bbbq_1_percentages.latex}
+  \caption{
+    Percentage of MHC-I epitopes overlapping with transmembrane helix.
+    \richel{This is simulated data}
+  }
   \label{table:1}
@@
 
+\begin{table}
+  \input{bbbq_1/bbbq_1_stats.latex}
+  \caption{
+    Kolmogorov-Smirnov test results
+    \richel{Done on the simulated data}
+  }
+  \label{table:ks}
+\end{table}
+
 \end{document}

### 2020-04-14 zip_all.sh
Can zip all

@@ -1,0 +1,2 @@
+#!/bin/bash
+zip bbbq_article_all.zip $(find . | egrep -v "\.zip$" | egrep -v "\.git")

### 2020-04-14 article.tex
Processed feedback Frans

@@
 \author[1]{Richèl J.C. Bilderbeek}
-\author[2]{Frans Bianchi}
-\affil[1]{Groningen Institute for Evolutionary Life Sciences, University of 
+\author[1]{Maxim Baranov}
+\author[1]{Geert van den Bogaart}
+\author[1]{Frans Bianchi}
+\affil[1]{GBB, University of 
 Groningen, Groningen, The Netherlands}
-\affil[2]{Frans' Institute, University of Groningen, Groningen, The Netherlands}
 
@@
 Transmembrane helices (TMHs) in the human proteome
-are an overrepresented source of epitopes on major 
+are an overrepresented potential source of epitopes on major 
 histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
 It is unknown if this same overrepresentation
-is also the case for the proteome of a pathogen.
+also holds for the proteome of a pathogen.
 Additionally, it is unknown if TMH-II is also likelier to present
@@
 
-For MHC-I, it was found that epitopes derived 
+For MHC-I, it was found that predicted epitopes derived 
 from human transmembrane helices (TMHs)
@@
 may have an evolutionary advantage for 
-the (human) host, as TMHs may be more 
-conserved due to the functional requirement of being able to span a lipid 
+the (human) host, as TMHs have a reduced variability 
+due to the functional requirement of being able to span a lipid 
 bilayer. 
@@
 If presentation of TMHs on MHC-I would bring an evolutionary advantage 
-in the recognition of pathogens by the immune, 
-it would follow that this is equally important for MHC-II.
+in the recognition of pathogens by the immune system, 
+it would follow that this is equally important for MHC-II, 
+especially as CD4+ T cell help is needed for a long lasting CD8+Tcell 
+response \cite{novy2007cd4}.
 
@@
 we can quantify to which extent selection in COVID-19 occurs
-in the transmembrane part of its proteins.
+in the transmembrane part of its proteome.
 
@@
 percentages of MHC-I epitopes overlap with TMHs in Homo sapiens
-for the different HLA haplotypes, with has an average of 5.3\%.
+for the different HLA haplotypes, with an average of 5.3\%.
 We obtained a similar distribution of percentages of MHC-I epitopes that 
@@
 
-We compare these distribution using a two-sample 
+We compare these distributions using a two-sample 
 Kolmogorov-Smirnov (KS) test. The KS test determines if two samples
-are derived from a same distribution, without making assumptions
+are derived from the same distribution, without making assumptions
 regarding the shape of that distribution.
@@
 	}
-	and DQA1*0501/DQB1*0301 \richel{reference here}.
+	and DQA1*0501/DQB1*0301 \richel{reference here}
+	\frans{
+	  Neem aan dat dit die ref is naar dat stukje van most common haplotype in human population. Misschien moeten we even de coverage erbij melden. 
+	  Greenbaum, Jason, et al.
+"Functional classification of class II human leukocyte antigen (HLA) molecules reveals seven different supertypes and a surprising degree of repertoire sharing across supertypes." Immunogenetics 63.6 (2011): 325-335.
+	}.
 	The 5\% peptides with the lowest IC50 values were defined as binders.
@@
 \iffalse
+
+	\frans{
+	  Ik heb zitten denken over de evolutie naast conservering is occurrence ook heel belangrijk. Moeten we de mate van conservering niet wegen ten opzichte van de occurrence.
+	  Wat ik daarmee bedoel is data als je kiest voor een mega geconseveerd epitope maar van dat type komen er maar 40 in een proteome voor versus een minder geconserveerd epitope maar daar zijn wel 100.000 varianten van dan heb je meer kans om 1 van die 100.000 te pakken toch?
+	}
 	The evolutionary conservation of TMHs was measured from a
@@
 	species of the mycoplasma bacterial family
-	\richel{how obtained exactly? how alignment done?}.
-
+	\richel{how obtained exactly? how alignment done?}
+	\frans{
+	Via entrez en blast 
+	  Ik he been transcript van Maxim ik zal dit proberen te vertalen naar een stukje tekst. 
+	}
+	.
 	Using the \verb;tmhprot; R package (\cite{tmhprot}), the DNA alignment is split
@@
 which makes us reject/accept the hypothesis that these percentages
-are sampled from a same distribution.
+are sampled from the same distribution.
 
@@
     \% epitopes overlapping with transmembrane helix
-    for a human and COVID-19 proteome
+    for a human and COVID-19 proteome.
+    \richel{
+      The Illuminati watermark indicates the data underlying this
+      figure has been simulated
+    }
   }
@@
 \iffalse
+	\frans{
+	  De TMH evolutie data van COVID is missachien interresanter. Ik zie het als volgt. 
+	  COVID is een mooi voorbeeld van een virus 
+	  Myco van een bacterie
+	  In de bacterie is duidelijk dat tenmisnte voor MHCII TMHs zijn overgrepresenteerd virus niet. 
+	  Dus verhaal kan zijn dat TMHs belangrijk voot bacteirele afweer niet zozeer voor virussen.
+
+	  Hiervoor moeten we MHC-I doen voor myco
+	  Kijken wat evolutie data doet voor virus dan evt kijken wat dit doet voor myco en vergelijken. 
+	}
+
 	We compared the mutation rates between the TMH and non-TMH part of

### 2020-04-14 article.tex
Use comments instead of iffalse

@@
 it would follow that this is equally important for MHC-II, 
-especially as CD4+ T cell help is needed for a long lasting CD8+Tcell 
+especially as CD4+ T cell help is needed for a long lasting CD8+ T cell 
 response \cite{novy2007cd4}.
@@
 
-\iffalse
-	Transmembrane helices and strong MHC-II-binding peptides
-	were predicted for a tuberculosis reference proteome 
-	from \url{https://www.ebi.ac.uk/reference_proteomes}
-	using the \verb;epitopeome; R package (\cite{epitopeome}).
-	We picked the three MHC-II alleles that are most abundant 
-	in the current human population, 
-	which are DRB4*0101, HLA-DPA10103-DPB10402 
-	\richel{
-	  We decided for DPB1*0402, but there are multiple alleles matching.
-	  I just picked the first one of the alleles matching, which are:
-	  HLA-DPA10x-DPB10402, where x equals 103-110, 201-204, 301-303 or 401
-	}
-	and DQA1*0501/DQB1*0301 \richel{reference here}
-	\frans{
-	  Neem aan dat dit die ref is naar dat stukje van most common haplotype in human population. Misschien moeten we even de coverage erbij melden. 
-	  Greenbaum, Jason, et al.
+	% Transmembrane helices and strong MHC-II-binding peptides
+	% were predicted for a tuberculosis reference proteome 
+	% from \url{https://www.ebi.ac.uk/reference_proteomes}
+	% using the \verb;epitopeome; R package (\cite{epitopeome}).
+	% We picked the three MHC-II alleles that are most abundant 
+	% in the current human population, 
+	% which are DRB4*0101, HLA-DPA10103-DPB10402 
+	% \richel{
+	%   We decided for DPB1*0402, but there are multiple alleles matching.
+	%   I just picked the first one of the alleles matching, which are:
+	%   HLA-DPA10x-DPB10402, where x equals 103-110, 201-204, 301-303 or 401
+	% }
+	% and DQA1*0501/DQB1*0301 \richel{reference here}
+	% \frans{
+	%   Neem aan dat dit die ref is naar dat stukje van most common haplotype in human population. Misschien moeten we even de coverage erbij melden. 
+	%   Greenbaum, Jason, et al.
 "Functional classification of class II human leukocyte antigen (HLA) molecules reveals seven different supertypes and a surprising degree of repertoire sharing across supertypes." Immunogenetics 63.6 (2011): 325-335.
-	}.
-	The 5\% peptides with the lowest IC50 values were defined as binders.
-	We then simply counted the number of amino acids that were present inside the
-	cell, within the membrane or outside of the cell, as well as if it was part 
-	of a strong MHC-II binding site, as shown in figure \ref{tab:results}.
+	% }.
+	% The 5\% peptides with the lowest IC50 values were defined as binders.
+	% We then simply counted the number of amino acids that were present inside the
+	% cell, within the membrane or outside of the cell, as well as if it was part 
+	% of a strong MHC-II binding site, as shown in figure \ref{tab:results}.
 
-	The \verb;epitopeome; R package (\cite{epitopeome}) binds 
-	together the \verb;tmhmm; (\cite{tmhmm}) and \verb;netmhc2pan; 
-	(\cite{netmhc2pan}) 
-	R packages. \verb;tmhmm; provides an R interface to
-	TMHMM (\cite{krogh2001predicting, sonnhammer1998hidden}), a tool
-	to predict where membrane proteins' amino acids are located within the
-	membrane.   
-	\verb;netmhc2pan; provides an R interface to
-	NetMHC2pan (\cite{jensen2018improved}), a tool
-	to predict MHC-II binding to proteins.
-\fi
+	% The \verb;epitopeome; R package (\cite{epitopeome}) binds 
+	% together the \verb;tmhmm; (\cite{tmhmm}) and \verb;netmhc2pan; 
+	% (\cite{netmhc2pan}) 
+	% R packages. \verb;tmhmm; provides an R interface to
+	% TMHMM (\cite{krogh2001predicting, sonnhammer1998hidden}), a tool
+	% to predict where membrane proteins' amino acids are located within the
+	% membrane.   
+	% \verb;netmhc2pan; provides an R interface to
+	% NetMHC2pan (\cite{jensen2018improved}), a tool
+	% to predict MHC-II binding to proteins.
 
@@
 
-\iffalse
+	% \frans{
+	%   Ik heb zitten denken over de evolutie naast conservering is occurrence ook heel belangrijk. Moeten we de mate van conservering niet wegen ten opzichte van de occurrence.
+	%   Wat ik daarmee bedoel is data als je kiest voor een mega geconseveerd epitope maar van dat type komen er maar 40 in een proteome voor versus een minder geconserveerd epitope maar daar zijn wel 100.000 varianten van dan heb je meer kans om 1 van die 100.000 te pakken toch?
+	% }
+	% The evolutionary conservation of TMHs was measured from a
+	% DNA alignment of multiple transmembrane proteins in multiple
+	% species of the mycoplasma bacterial family
+	% \richel{how obtained exactly? how alignment done?}
+	% \frans{
+	% Via entrez en blast 
+	%   Ik he been transcript van Maxim ik zal dit proberen te vertalen naar een stukje tekst. 
+	% }
+	% .
+	% Using the \verb;tmhprot; R package (\cite{tmhprot}), the DNA alignment is split
+	% into two alignments, one for the TMH parts, another for the non-TMH
+	% parts. Each alignment was tested by the \verb;mcbette; 
+	% R package (\cite{mcbette}) to select the Bayesian inference model with
+	% the highest evidence (a.k.a. the marginal likelihood) using the nested
+	% sampling approach as described in \cite{maturana2018model},
+	% using the popular Bayesian phylogenetic tool 
+	% BEAST2 (\cite{bouckaert2014beast}) in the back-end.
+	% \verb;mcbette; used a set of 40 candidate inference models, 
+	% consisting of all combinations of 
+	% 4 site models (JC, HKY, TN, GTR), 
+	% 2 clock models (strict, RLN) and 
+	% 5 tree priors (Yule, BD, CBS, CCP, CEP).
 
-	\frans{
-	  Ik heb zitten denken over de evolutie naast conservering is occurrence ook heel belangrijk. Moeten we de mate van conservering niet wegen ten opzichte van de occurrence.
-	  Wat ik daarmee bedoel is data als je kiest voor een mega geconseveerd epitope maar van dat type komen er maar 40 in een proteome voor versus een minder geconserveerd epitope maar daar zijn wel 100.000 varianten van dan heb je meer kans om 1 van die 100.000 te pakken toch?
-	}
-	The evolutionary conservation of TMHs was measured from a
-	DNA alignment of multiple transmembrane proteins in multiple
-	species of the mycoplasma bacterial family
-	\richel{how obtained exactly? how alignment done?}
-	\frans{
-	Via entrez en blast 
-	  Ik he been transcript van Maxim ik zal dit proberen te vertalen naar een stukje tekst. 
-	}
-	.
-	Using the \verb;tmhprot; R package (\cite{tmhprot}), the DNA alignment is split
-	into two alignments, one for the TMH parts, another for the non-TMH
-	parts. Each alignment was tested by the \verb;mcbette; 
-	R package (\cite{mcbette}) to select the Bayesian inference model with
-	the highest evidence (a.k.a. the marginal likelihood) using the nested
-	sampling approach as described in \cite{maturana2018model},
-	using the popular Bayesian phylogenetic tool 
-	BEAST2 (\cite{bouckaert2014beast}) in the back-end.
-	\verb;mcbette; used a set of 40 candidate inference models, 
-	consisting of all combinations of 
-	4 site models (JC, HKY, TN, GTR), 
-	2 clock models (strict, RLN) and 
-	5 tree priors (Yule, BD, CBS, CCP, CEP).
-
-	For each alignment, the inference model with the highest evidence
-	is used in a Bayesian inference. From each Bayesian posterior,
-	the parameter estimates regarding mutation rates (including clock rate)
-	were obtained and compared to determine which realized mutation rate is lower.
-	The Markov chain Monte Carlo was set up in such a way that the effective sample
-	size for the likelihood of the inference model is above the recommended value
-	of 200 (\cite{bouckaert2014beast}).
-\fi
+	% For each alignment, the inference model with the highest evidence
+	% is used in a Bayesian inference. From each Bayesian posterior,
+	% the parameter estimates regarding mutation rates (including clock rate)
+	% were obtained and compared to determine which realized mutation rate is lower.
+	% The Markov chain Monte Carlo was set up in such a way that the effective sample
+	% size for the likelihood of the inference model is above the recommended value
+	% of 200 (\cite{bouckaert2014beast}).
 
@@
 
-\iffalse
-	Table \ref{tab:results} shows the location and binding strength for the
-	tuberculosis proteome.
+	% Table \ref{tab:results} shows the location and binding strength for the
+	% tuberculosis proteome.
 
-	\begin{figure}[ht]
-	  \includegraphics[width=\textwidth]{figure_1.png}
-	  \label{fig:1}
-	\end{figure}
+	% \begin{figure}[ht]
+	%   \includegraphics[width=\textwidth]{figure_1.png}
+	%   \label{fig:1}
+	% \end{figure}
 
-	\begin{figure}[ht]
-	  \includegraphics[width=\textwidth]{figure_1_5.png}
-	  \label{fig:1_5}
-	\end{figure}
+	% \begin{figure}[ht]
+	%   \includegraphics[width=\textwidth]{figure_1_5.png}
+	%   \label{fig:1_5}
+	% \end{figure}
 
-	\begin{figure}[ht]
-	  \includegraphics[width=\textwidth]{figure_3.png}
-	  \label{fig:3}
-	\end{figure}
+	% \begin{figure}[ht]
+	%   \includegraphics[width=\textwidth]{figure_3.png}
+	%   \label{fig:3}
+	% \end{figure}
 
-	\input{table_imoimo.latex}
-	% has label tab:results
-	\richel{
-	  I'd enjoy 
-	  (1) a row with 'expected by chance', 
-	  (2) using percentages instead,
-	  (3) merging inside and outside
-	}
+	% \input{table_imoimo.latex}
+	% % has label tab:results
+	% \richel{
+	%   I'd enjoy 
+	%   (1) a row with 'expected by chance', 
+	%   (2) using percentages instead,
+	%   (3) merging inside and outside
+	% }
 
-	The inference model with the highest evidence in the
-	TMH-only alignment was [yet unknown] and [also yet unknown]
-	for the non-TMH alignment. Individual model weights are shown
-	in tables \ref{tab:evidences_tmh} 
-	and \ref{tab:evidences_non_tmh}.
+	% The inference model with the highest evidence in the
+	% TMH-only alignment was [yet unknown] and [also yet unknown]
+	% for the non-TMH alignment. Individual model weights are shown
+	% in tables \ref{tab:evidences_tmh} 
+	% and \ref{tab:evidences_non_tmh}.
 
-	The Bayesian inference resulted in [a distribution of mutation rates],
-	as shown in [absent figure].
-	The ESSes of the Bayesian parameter estimates was above 200, exact values
-	are shown in tables \ref{tab:esses_tmh} and \ref{tab:esses_non_tmh}.
-\fi
+	% The Bayesian inference resulted in [a distribution of mutation rates],
+	% as shown in [absent figure].
+	% The ESSes of the Bayesian parameter estimates was above 200, exact values
+	% are shown in tables \ref{tab:esses_tmh} and \ref{tab:esses_non_tmh}.
 
@@
 
-\iffalse
-	We conclude that MHC-II binds to TMH peptides with a higher/lower/equal
-	probability than expected by chance.
+	% We conclude that MHC-II binds to TMH peptides with a higher/lower/equal
+	% probability than expected by chance.
 
-	We conclude that the evolutionary conservation if the TMH parts of membrane
-	proteins is higher/less/equal compare to its non-TMH counterparts.
-\fi
+	% We conclude that the evolutionary conservation if the TMH parts of membrane
+	% proteins is higher/less/equal compare to its non-TMH counterparts.
 
@@
 
-\iffalse
-	\frans{
-	  De TMH evolutie data van COVID is missachien interresanter. Ik zie het als volgt. 
-	  COVID is een mooi voorbeeld van een virus 
-	  Myco van een bacterie
-	  In de bacterie is duidelijk dat tenmisnte voor MHCII TMHs zijn overgrepresenteerd virus niet. 
-	  Dus verhaal kan zijn dat TMHs belangrijk voot bacteirele afweer niet zozeer voor virussen.
+	% \frans{
+	%   De TMH evolutie data van COVID is missachien interresanter. Ik zie het als volgt. 
+	%   COVID is een mooi voorbeeld van een virus 
+	%   Myco van een bacterie
+	%   In de bacterie is duidelijk dat tenmisnte voor MHCII TMHs zijn overgrepresenteerd virus niet. 
+	%   Dus verhaal kan zijn dat TMHs belangrijk voot bacteirele afweer niet zozeer voor virussen. 
+	% 
+	%   Hiervoor moeten we MHC-I doen voor myco
+	%   Kijken wat evolutie data doet voor virus dan evt kijken wat dit doet voor myco en vergelijken. 
+	% }
+	% 
+	% We compared the mutation rates between the TMH and non-TMH part of
+	% multiple mycoplasma species. Where we expect no variation 
+	% in mutation rate for every TMH amino acid,
+	% \richel{
+	%   we can test this, but unsure if that would make sense
+	% }
+	% we know that non-TMH part will have regions of different evolutionary
+	% conservation: functional domains, especially in protein-protein
+	% interactions will be strongly conserved, due to an even more constrained
+	% set of peptides that enable a certain function.
 
-	  Hiervoor moeten we MHC-I doen voor myco
-	  Kijken wat evolutie data doet voor virus dan evt kijken wat dit doet voor myco en vergelijken. 
-	}
-
-	We compared the mutation rates between the TMH and non-TMH part of
-	multiple mycoplasma species. Where we expect no variation 
-	in mutation rate for every TMH amino acid,
-	\richel{
-	  we can test this, but unsure if that would make sense
-	}
-	we know that non-TMH part will have regions of different evolutionary
-	conservation: functional domains, especially in protein-protein
-	interactions will be strongly conserved, due to an even more constrained
-	set of peptides that enable a certain function.
-
-	\richel{
-	  Note that most bacteria are opportunistic pathogens.
-	  Note that most bacteria are generalists.
-	  Note that most bacteria have different cell membranes (and walls), that
-	  may have different functional constraints than a human cell membrane
-	}
-\fi
+	% \richel{
+	%   Note that most bacteria are opportunistic pathogens.
+	%   Note that most bacteria are generalists.
+	%   Note that most bacteria have different cell membranes (and walls), that
+	%   may have different functional constraints than a human cell membrane
+	% }
 

### 2020-04-15 article.tex
Introduction better

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Immune response}
+
+\richel{Check for correctedness}
+Our immune system fights invaders on a daily basis.
+These invaders can be fungi, bacteria or viruses.
+The innate immune response is its first general 
+and immediate strategy, where the acquired immune response
+needs time to develop its specialized and more effective
+combat forces.
+
+\paragraph{Immune response by MHC-I}
+
+\richel{Check for correctedness}
+All nucleated cells in humans present randomly sampled polypeptide
+fragments to the surroundings of the cell using MHC class I molecules.
+If a cell gets infected by a virus, also the virus' polypeptides
+will be presented at the cell surface. The viral antigen is detected 
+by cytotoxic T lymphocytes, which will kill the infected cell.
+
+\paragraph{Immune response by MHC-II}
+
+\richel{Check for correctedness}
+Bacteria and funghi do not infect a cell with their foreign DNA,
+but are detected by the proteins on their cell walls. MHC class II molecules 
+bind to specific polypeptide fragments. This binding allows T and B cells,
+which all express MHC-II, to detect a pathogen and trigger the response to
+kill the pathogen. Note that the foreign proteins of a viral envelope 
+can be detected in this way as well, which is commonly used in vaccination.
+
+\paragraph{Non-random epitopes}
+
+\richel{Check for correctedness}
+Any human's immune system detects only a selection of all possible
+polypeptide fragments. Each MHC molecule can only express (MHC-I)
+or detect (MHC-II) a couple of polypeptides. An increased detection
+is obtained by expressing a wide variety of
+MHCs, which is assured by the human leukocyte antigen gene complex (HLA).
+However, a different HLA will result in a different variety of MHCs,
+that will display different polypeptide fragments. Additionally, there
+may have been selection the HLA to display and or detect different
+polypeptide fragments.
 

### 2020-04-15 article.tex
Spellcheck

@@
 
-\richel{Check for correctedness}
+\richel{Check for correctness}
 Our immune system fights invaders on a daily basis.
@@
 
-\richel{Check for correctedness}
+\richel{Check for correctness}
 All nucleated cells in humans present randomly sampled polypeptide
@@
 
-\richel{Check for correctedness}
-Bacteria and funghi do not infect a cell with their foreign DNA,
+\richel{Check for correctness}
+Bacteria and fungi do not infect a cell with their foreign DNA,
 but are detected by the proteins on their cell walls. MHC class II molecules 
@@
 
-\richel{Check for correctedness}
+\richel{Check for correctness}
 Any human's immune system detects only a selection of all possible

### 2020-04-15 article.tex
Intro progress

@@
 
-\richel{Check for correctness}
+\richel{Check against stupidity}
 Our immune system fights invaders on a daily basis.
@@
 
-\richel{Check for correctness}
+\richel{Check against stupidity}
 All nucleated cells in humans present randomly sampled polypeptide
@@
 
-\richel{Check for correctness}
+\richel{Check against stupidity}
 Bacteria and fungi do not infect a cell with their foreign DNA,
@@
 
-\richel{Check for correctness}
+\richel{Check against stupidity}
 Any human's immune system detects only a selection of all possible
@@
 from human transmembrane helices (TMHs)
-are over-presented by all human leukocyte antigen (HLA)-A and 
+are over-presented by all HLA-A and 
 most HLA-B super types (\cite{bianchi2017}). 
@@
 
-\paragraph{Evolution}
+\paragraph{Evolutionary arms race}
 
-If the human immune system is more prone to detect polypeptides
-derived from a pathogen's TMHs, one may expect to find an evolutionary
-signal in fast-evolving pathogen. 
+\richel{Check against stupidity}
+The human immune system and human pathogen are in an evolutionary
+arms race, in which our immune systems is selected for the detection
+of pathogens, whereas pathogens are selected to avoid being detected.
+If the human immune system is more prone to display (and possibly 
+detect) polypeptides derived from TMHs, one may expect to find 
+an evolutionary signal in pathogens as well.
+
+\paragraph{Evolutionary signal of unknown strength}
+
+\richel{Check against stupidity}
+We do not know how viruses are selected \richel{so I should read the
+literature on that}. We predict that there is a strong 
+evolutionary conservation in regions that are essential for viral proteins'
+functions. Because these regions are rare in a viral proteome,
+and differ widely between pathogens, we expect that the human immune system
+is not tailored to detect these sites. TMHs, however, have less 
+functional constraints yet are more general. Because TMHs are more general,
+we expect it for the immune system to pay off to be more attentive towards
+TMHs. Because the strength of the evolutionary signal,
+are strong data set needs to be used.
+
+We expect that
+
 We can assume that only after the moment that COVID-19 spilled over 

### 2020-04-15 article.tex
Add Mycoplasma

@@
 
+\richel{Check against stupidity}
 For MHC-I, it was found that predicted epitopes derived 
@@
 as many mutations will result in a disfunctional TMH.
+Note that the mechanism by which a cell presents its TMHs is
+yet unknown.
 
@@
 
+\richel{Check against stupidity}
 If presentation of TMHs on MHC-I would bring an evolutionary advantage 
@@
 response \cite{novy2007cd4}. 
+The mechanism to detect foreign TMHs by MHC-II would be unknown, 
+similar to the discovery that MHC-I presents TMHs.
 
@@
 
-\paragraph{Evolutionary signal of unknown strength}
+\paragraph{Evolutionary signal of unknown strength for virus}
 
 \richel{Check against stupidity}
-We do not know how viruses are selected \richel{so I should read the
-literature on that}. We predict that there is a strong 
+Regarding human viruses, we do not know the extent to which the human
+immune systems selects on these, nor which loci are selected
+upon \richel{so I should read the literature on that}. 
+We predict that there is a strong 
 evolutionary conservation in regions that are essential for viral proteins'
@@
 we expect it for the immune system to pay off to be more attentive towards
-TMHs. Because the strength of the evolutionary signal,
-are strong data set needs to be used.
+TMHs. 
+Additionally, viruses are not only selected for their evasion of an 
+immune response, yet are selected to have a high reproductive value.
+Because the strength of the evolutionary signal is unknown,
+we use a strong data set to be able to detect it.
 
-We expect that
+\paragraph{Evolutionary signal of unknown strength for bacteria}
 
+\richel{Check against stupidity}
+Regarding human pathogenic bacteria, we do not know the extent to which the 
+human immune systems selects on these, nor which loci are selected
+upon \richel{so I should read the literature on that}.
+Because most bacterial pathogens are generalists, that is,
+can infect multiple hosts, we expect the selection pressure exerted
+by the human immune response to be weak.
+Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
+Due to this, the strength of the evolutionary signal is unknown,
+thus we need a strong data set to be able to detect it.
+
+\paragraph{Data sets}
+
+\richel{Check against stupidity}
+To detect an evolutionary signal for selection on the TMHs of viruses, 
+we use the many time-dated COVID-19 proteomes.
 We can assume that only after the moment that COVID-19 spilled over 
-to humans, it's the human immune system that is an influential component 
-of its evolutionary selection.
-Using the extensive amount of time-dated COVID-19 sequences,
-we can quantify to which extent selection in COVID-19 occurs
-in the transmembrane part of its proteome.
+to humans, it has mostly been the human immune system selecting against
+its detection, although this is only part of the selection for 
+having a high transmission rate.
+To detect an evolutionary signal for selection on the TMHs of bacteria, 
+we use the proteomes of Mycoplasma. Mycoplasma is \richel{something}.
 
@@
   \item $\mathcal{H}_1$: MHC-I has the same percentage of epitopes overlapping
-    with TMHs in Homo sapiens as in COVID-19
+    with TMHs in Homo sapiens as in COVID-19 as in Mycoplasma
   \item $\mathcal{H}_2$: MHC-II has the same percentage of epitopes overlapping
-    with TMHs as expected by chance
-  \item $\mathcal{H}_3$: The TMHs of COVID-19 have the same mutation rate 
-    as the cytosolic and extracellular parts of its transmembrane proteins
+    with TMHs as expected by chance in Homo sapiens and COVID-19 and Mycoplasma
+  \item $\mathcal{H}_3$: The TMHs of both COVID-19 and Mycoplasma
+    have the same mutation rate 
+    as the cytosolic and extracellular parts of their transmembrane proteins
 \end{itemize}
@@
 except that instead of a human reference proteome,
-we used the proteome of the first sequenced COVID-19 strain (\cite{wu2020new}),
-with a GenBank ID of MN908947.3, as can be downloaded 
-from \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947}.
+we use the proteome of the first sequenced COVID-19 strain (\cite{wu2020new},
+GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
+as well as a Mycoplasma \richel{that one} reference genome.
 
@@
 overlap with TMHs for the different HLA haplotypes, but then applied to
-COVID-19.
+COVID-19 and Mycoplasma.
 
-We compare these distributions using a two-sample 
+We compare the distributions of humans and each pathogen
+using a two-sample 
 Kolmogorov-Smirnov (KS) test. The KS test determines if two samples
@@
 the null hypothesis that MHC-I has the same percentage of epitopes 
-overlapping with TMHs in Homo sapiens as in COVID-19 when 
+overlapping with TMHs in Homo sapiens compared to each pathogen when 
 the KS statistic $D_{n,m}$ follows the relationship as shown in 
@@
 Figure \ref{fig:1} shows the percentages of epitopes overlapping 
-with TMHs for a human and COVID-19 proteome.
+with TMHs for a human, COVID-19 and Mycoplasma proteome.
 The KS test to determine if these percentages are sampled from
@@
     \% epitopes overlapping with transmembrane helix
-    for a human and COVID-19 proteome.
+    for a human, COVID-19 and Mycoplasma proteome.
     \richel{
-      The Illuminati watermark indicates the data underlying this
-      figure has been simulated
+      the data underlying this figure has been simulated
     }
@@
 
+We found that the percentages of epitopes overlapping 
+with TMHs for a human and Mycopasma proteome are 
+[similar/different]. In other words, the
+epitopes that MHC-I presents are [as/not as] likely 
+to be derived from TMH within either a human host and its bacterial pathogen.
+
 	% We conclude that MHC-II binds to TMH peptides with a higher/lower/equal
@@
 evolutionary conservation of MHC-I detecting TMHs.
+
+We concluded that the
+epitopes that MHC-I presents are [as/not as] likely 
+to be derived from TMH within either a human host and its bacterial pathogen.
+Because a bacterium does not infect a cell, thus its polypeptides
+will not be presented by MHC-I, this result is [unexpected/expected]
 

### 2020-04-15 article.tex
Spellcheck

@@
 \richel{Check against stupidity}
-All nucleated cells in humans present randomly sampled polypeptide
+All nucleated cells in humans present randomly sampled polypeptides
 fragments to the surroundings of the cell using MHC class I molecules.
@@
 We found that the percentages of epitopes overlapping 
-with TMHs for a human and Mycopasma proteome are 
+with TMHs for a human and Mycoplasma proteome are 
 [similar/different]. In other words, the

### 2020-04-15 article.tex
Add Myco
Add Mycobacterium reference proteome link

@@
 To detect an evolutionary signal for selection on the TMHs of bacteria, 
-we use the proteomes of Mycoplasma. Mycoplasma is \richel{something}.
+we use the proteomes of Mycobacterium. Mycobacterium is \richel{something}.
 
@@
   \item $\mathcal{H}_1$: MHC-I has the same percentage of epitopes overlapping
-    with TMHs in Homo sapiens as in COVID-19 as in Mycoplasma
+    with TMHs in Homo sapiens as in COVID-19 as in Mycobacterium
   \item $\mathcal{H}_2$: MHC-II has the same percentage of epitopes overlapping
-    with TMHs as expected by chance in Homo sapiens and COVID-19 and Mycoplasma
-  \item $\mathcal{H}_3$: The TMHs of both COVID-19 and Mycoplasma
+    with TMHs as expected by chance in Homo sapiens and COVID-19 and Mycobacterium
+  \item $\mathcal{H}_3$: The TMHs of both COVID-19 and Mycobacterium
     have the same mutation rate 
@@
 GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
-as well as a Mycoplasma \richel{that one} reference genome.
+and a reference Mycobacterium tuberculosis 
+proteome (\url{https://www.ebi.ac.uk/reference_proteomes}, UP000001584, 
+83332 MYCTU).
 
@@
 overlap with TMHs for the different HLA haplotypes, but then applied to
-COVID-19 and Mycoplasma.
+COVID-19 and Mycobacterium.
 
@@
 
+For MHC-II, we do the same analysis as MHC-I, except for
+MHC-II haplotypes.
+
 	% Transmembrane helices and strong MHC-II-binding peptides
 	% were predicted for a tuberculosis reference proteome 
-	% from \url{https://www.ebi.ac.uk/reference_proteomes}
 	% using the \verb;epitopeome; R package (\cite{epitopeome}).
@@
 	% DNA alignment of multiple transmembrane proteins in multiple
-	% species of the mycoplasma bacterial family
+	% species of the mycobacterium bacterial family
 	% \richel{how obtained exactly? how alignment done?}
@@
 Figure \ref{fig:1} shows the percentages of epitopes overlapping 
-with TMHs for a human, COVID-19 and Mycoplasma proteome.
+with TMHs for a human, COVID-19 and Mycobacterium proteome.
 The KS test to determine if these percentages are sampled from
@@
     \% epitopes overlapping with transmembrane helix
-    for a human, COVID-19 and Mycoplasma proteome.
+    for a human, COVID-19 and Mycobacterium proteome.
     \richel{
@@
 We found that the percentages of epitopes overlapping 
-with TMHs for a human and Mycoplasma proteome are 
+with TMHs for a human and Mycobacterium proteome are 
 [similar/different]. In other words, the
@@
 	% We compared the mutation rates between the TMH and non-TMH part of
-	% multiple mycoplasma species. Where we expect no variation 
+	% multiple mycobacterium species. Where we expect no variation 
 	% in mutation rate for every TMH amino acid,
@@
 \begin{table}
-  \input{bbbq_1/bbbq_1_stats.latex}
+  \input{bbbq_1/bbbq_1_stats_covid.latex}
   \caption{
@@
   }
-  \label{table:ks}
+  \label{table:ks_covid}
+\end{table}
+
+\begin{table}
+  \input{bbbq_1/bbbq_1_stats_myco.latex}
+  \caption{
+    Kolmogorov-Smirnov test results
+    \richel{Done on the simulated data}
+  }
+  \label{table:ks_myco}
 \end{table}

### 2020-04-15 article.tex
Add MHC-II results
Better labels

@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/fig_bbbq_2_watermarked.png}
+  \caption{
+    \% epitopes overlapping with transmembrane helix
+    for a human, COVID-19 and Mycobacterium proteome.
+    \richel{
+      the data underlying this figure has been simulated
+    }
+  }
+  \label{fig:2}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 	% \begin{figure}[ht]
-	%   \includegraphics[width=\textwidth]{figure_1.png}
+	%   \includegraphics[width=\textwidth]{bbbq_3/figure_1.png}
 	%   \label{fig:1}
@@
   }
-  \label{table:1}
+  \label{table:bbbq_1_percentages}
 \end{table}
@@
   \caption{
-    Kolmogorov-Smirnov test results
+    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-I
     \richel{Done on the simulated data}
   }
-  \label{table:ks_covid}
+  \label{table:bbbq_1_stats_covid}
 \end{table}
@@
   \caption{
-    Kolmogorov-Smirnov test results
+    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-I
     \richel{Done on the simulated data}
   }
-  \label{table:ks_myco}
+  \label{table:bbbq_1_stats_myco}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{MHC-II}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}
+  \input{bbbq_2/bbbq_2_percentages.latex}
+  \caption{
+    Percentage of MHC-II epitopes overlapping with transmembrane helix.
+    \richel{This is simulated data}
+  }
+  \label{table:bbbq_2_percentages}
+\end{table}
+
+\begin{table}
+  \input{bbbq_2/bbbq_2_stats_covid.latex}
+  \caption{
+    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-II
+    \richel{Done on the simulated data}
+  }
+  \label{table:bbbq_2_stats_covid}
+\end{table}
+
+\begin{table}
+  \input{bbbq_2/bbbq_2_stats_myco.latex}
+  \caption{
+    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-II
+    \richel{Done on the simulated data}
+  }
+  \label{table:bbbq_2_stats_myco}
 \end{table}

### 2020-04-15 article.tex
MHC-II

@@
 
+\richel{Check against stupidity}
 For MHC-II, we do the same analysis as MHC-I, except for
-MHC-II haplotypes.
-
-	% Transmembrane helices and strong MHC-II-binding peptides
-	% were predicted for a tuberculosis reference proteome 
-	% using the \verb;epitopeome; R package (\cite{epitopeome}).
-	% We picked the three MHC-II alleles that are most abundant 
-	% in the current human population, 
-	% which are DRB4*0101, HLA-DPA10103-DPB10402 
-	% \richel{
-	%   We decided for DPB1*0402, but there are multiple alleles matching.
-	%   I just picked the first one of the alleles matching, which are:
-	%   HLA-DPA10x-DPB10402, where x equals 103-110, 201-204, 301-303 or 401
-	% }
-	% and DQA1*0501/DQB1*0301 \richel{reference here}
-	% \frans{
-	%   Neem aan dat dit die ref is naar dat stukje van most common haplotype in human population. Misschien moeten we even de coverage erbij melden. 
+MHC-II haplotypes. We picked the top 21 MHC-II alleles that are most abundant 
+in the current human population \richel{reference here}
+%\frans{
+%  Neem aan dat dit die ref is naar dat stukje van most common haplotype in human population. 
+%  Misschien moeten we even de coverage erbij melden. 
 %  Greenbaum, Jason, et al.
 "Functional classification of class II human leukocyte antigen (HLA) molecules reveals seven different supertypes and a surprising degree of repertoire sharing across supertypes." Immunogenetics 63.6 (2011): 325-335.
-	% }.
-	% The 5\% peptides with the lowest IC50 values were defined as binders.
-	% We then simply counted the number of amino acids that were present inside the
-	% cell, within the membrane or outside of the cell, as well as if it was part 
-	% of a strong MHC-II binding site, as shown in figure \ref{tab:results}.
-
-	% The \verb;epitopeome; R package (\cite{epitopeome}) binds 
-	% together the \verb;tmhmm; (\cite{tmhmm}) and \verb;netmhc2pan; 
-	% (\cite{netmhc2pan}) 
-	% R packages. \verb;tmhmm; provides an R interface to
-	% TMHMM (\cite{krogh2001predicting, sonnhammer1998hidden}), a tool
-	% to predict where membrane proteins' amino acids are located within the
-	% membrane.   
-	% \verb;netmhc2pan; provides an R interface to
-	% NetMHC2pan (\cite{jensen2018improved}), a tool
-	% to predict MHC-II binding to proteins.
+%}.
+We chose to use 21 alleles to have a similar number of alleles as is done
+for MHC-I.
+Same as \cite{bianchi2017}, we defined the 5\% of polypeptides 
+with the lowest IC50 values as binders. We then counted the percentage
+of binders that have one or more amino acids that are part of a TMH.
 

### 2020-04-15 article.tex
EOD

@@
 equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
-and $n = m = 13$ HLA haplotypes.
+and $n = m$ equals the number of HLA haplotypes.
 
@@
 MHC-II haplotypes. We picked the top 21 MHC-II alleles that are most abundant 
-in the current human population \richel{reference here}
+in the current human population (\cite{greenbaum2011functional}).
 %\frans{
@@
 
-We thanks to Geert van den Bogaart for his wisdom.
+We thank the Center for Information Technology of the University 
+of Groningen for its support and for providing access to the Peregrine 
+high performance computing cluster.
 
@@
 \begin{table}
+  \input{mhc2_haplotypes.latex}
+  \caption{
+    Percentage of MHC-II haplotypes, from \cite{greenbaum2011functional}
+  }
+  \label{table:bbbq_2_percentages}
+\end{table}
+
+\begin{table}
   \input{bbbq_2/bbbq_2_percentages.latex}

### 2020-04-16 article.tex
Add new reference sent by Geert. Moved KS equation to end of appendix

@@
 We compare the distributions of humans and each pathogen
-using a two-sample 
-Kolmogorov-Smirnov (KS) test. The KS test determines if two samples
-are derived from the same distribution, without making assumptions
-regarding the shape of that distribution.
-
-We will reject
-the null hypothesis that MHC-I has the same percentage of epitopes 
-overlapping with TMHs in Homo sapiens compared to each pathogen when 
-the KS statistic $D_{n,m}$ follows the relationship as shown in 
-equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
-and $n = m$ equals the number of HLA haplotypes.
-
-\begin{equation}
-   D_{n,m} > \frac{1}{\sqrt{n}} \cdot \sqrt{ -\ln(\frac{\alpha}{2}) \cdot \frac{1 + \frac{n}{m}}{2} }
-\  \label{eq:ks}
-\end{equation}
+using a two-sample Kolmogorov-Smirnov (KS) test
+for a significance level $\alpha = 0.05$.
 
@@
 in the current human population (\cite{greenbaum2011functional}).
-%\frans{
-%  Neem aan dat dit die ref is naar dat stukje van most common haplotype in human population. 
-%  Misschien moeten we even de coverage erbij melden. 
-%  Greenbaum, Jason, et al.
-"Functional classification of class II human leukocyte antigen (HLA) molecules reveals seven different supertypes and a surprising degree of repertoire sharing across supertypes." Immunogenetics 63.6 (2011): 325-335.
-%}.
+\frans{Show coverage}
 We chose to use 21 alleles to have a similar number of alleles as is done
@@
 of binders that have one or more amino acids that are part of a TMH.
+
+\richel{Read up on \cite{arlehamn2016quantitative}}
 
@@
 \begin{table}
-  \input{mhc2_haplotypes.latex}
-  \caption{
-    Percentage of MHC-II haplotypes, from \cite{greenbaum2011functional}
-  }
-  \label{table:bbbq_2_percentages}
-\end{table}
-
-\begin{table}
   \input{bbbq_2/bbbq_2_percentages.latex}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{MHC-II haplotype occurrences}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  This is just a reminder, instead of new research. 
+  This subsection be deleted in the future.
+}
+
+\begin{table}
+  \input{mhc2_haplotypes.latex}
+  \caption{
+    Percentage of MHC-II haplotypes, from \cite{greenbaum2011functional}
+    \richel{
+      This is just a reminder, instead of new research. 
+      This table be deleted in the future.
+    }
+  }
+  \label{table:bbbq_2_percentages}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Kolmogorov-Smirnov}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  This is just a reminder, instead of new research. 
+  This subsection be deleted in the future.
+}
+
+The Kolmogorov-Smirnov (KS) test determines if two samples
+are derived from the same distribution, without making assumptions
+regarding the shape of that distribution.
+
+We will reject
+the null hypothesis that MHC-I has the same percentage of epitopes 
+overlapping with TMHs in Homo sapiens compared to each pathogen when 
+the KS statistic $D_{n,m}$ follows the relationship as shown in 
+equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
+and $n = m$ equals the number of HLA haplotypes.
+
+\begin{equation}
+   D_{n,m} > \frac{1}{\sqrt{n}} \cdot \sqrt{ -\ln(\frac{\alpha}{2}) \cdot \frac{1 + \frac{n}{m}}{2} }
+   \label{eq:ks}
+\end{equation}
+
 \end{document}

### 2020-04-16 article.tex
Progress

@@
 \begin{itemize}
-  \item $\mathcal{H}_1$: MHC-I has the same percentage of epitopes overlapping
-    with TMHs in Homo sapiens as in COVID-19 as in Mycobacterium
-  \item $\mathcal{H}_2$: MHC-II has the same percentage of epitopes overlapping
-    with TMHs as expected by chance in Homo sapiens and COVID-19 and Mycobacterium
-  \item $\mathcal{H}_3$: The TMHs of both COVID-19 and Mycobacterium
+  \item $\mathcal{H}_{1,1}$: MHC-I has the same percentage of epitopes overlapping
+    with TMHs in Homo sapiens as in COVID-19
+  \item $\mathcal{H}_{1,2}$: MHC-I has the same percentage of epitopes overlapping
+    with TMHs in Homo sapiens as in Mycobacterium
+  \item $\mathcal{H}_{2,1}$: MHC-II has the same percentage of epitopes overlapping
+    with TMHs as expected by chance in Homo sapiens
+  \item $\mathcal{H}_{2,2}$: MHC-II has the same percentage of epitopes overlapping
+    with TMHs as expected by chance in COVID-19
+  \item $\mathcal{H}_{2,3}$: MHC-II has the same percentage of epitopes overlapping
+    with TMHs as expected by chance in Mycobacterium
+  \item $\mathcal{H}_{3,1}$: The TMHs of COVID-19
+    have the same mutation rate 
+    as the cytosolic and extracellular parts of their transmembrane proteins
+  \item $\mathcal{H}_{3,2}$: The TMHs of Mycobacterium
     have the same mutation rate 
@@
 We chose to use 21 alleles to have a similar number of alleles as is done
-for MHC-I.
+for MHC-I \richel{Unsure if this was the actual reason}.
 Same as \cite{bianchi2017}, we defined the 5\% of polypeptides 
@@
 of binders that have one or more amino acids that are part of a TMH.
-
-\richel{Read up on \cite{arlehamn2016quantitative}}
 

### 2020-04-16 to_80_chars.sh
Add reference proteome

@@ -1,11 +1,0 @@
-#!/bin/bash
-#
-# Convert article.tex to have 80 characters per line
-#
-# Usage:
-#
-#   ./to_80_chars.sh
-#
-fold -w 80 -s article.tex > article_new.tex
-cp article_new.tex article.tex
-rm article_new.tex

### 2020-04-16 tmhs/edit.sh
Add edit script

@@ -1,0 +1,11 @@
+#!/bin/bash
+#
+# Start editing the R markdown file.
+#
+# Usage:
+#
+#   ./edit.sh
+#
+
+mousepad tmhs.Rmd
+make

### 2020-04-20 article.tex
Show TMHs of COVID-19
EOD
Show COVID-19 genome and proteome

@@
 \usepackage{amsfonts}
-\usepackage{array}
 \usepackage{array}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{COVID-19 genome and proteome}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  This is just a reminder, instead of new research. 
+  This subsection be deleted in the future.
+}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{pics/covid_genome_and_proteome.png}
+  \caption{
+    Overview of COVID-19 genome and proteome.
+    From \url{https://zhanglab.ccmb.med.umich.edu/COVID-19}
+  }
+  \label{fig:covid_genome_and_proteome}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{COVID-19 TMHs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
+  \caption{
+    Location of amino acids of reference COVID-19 proteome.
+    Note that ORF1a results in multiple proteins, 
+    as shown in figure \ref{fig:covid_genome_and_proteome}.
+    Legend: i = inside (cytosolic side), m = membrane, o = outside (extracellular side)
+    \richel{This is real data}
+  }
+  \label{fig:covid_locatome}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-II haplotype occurrences}

### 2020-04-20 article.tex
Introduce new hypotheses regarding hydrophobicity

@@
 
-\paragraph{Non-random epitopes}
+\paragraph{MHCs bind hydrophobic epitopes}
 
 \richel{Check against stupidity}
-Any human's immune system detects only a selection of all possible
-polypeptide fragments. Each MHC molecule can only express (MHC-I)
-or detect (MHC-II) a couple of polypeptides. An increased detection
+Any human's immune system detects only a fraction of all possible
+polypeptide fragments. 
+Already each MHC molecule can only bind a
+subset of all possible polypeptides. 
+The polypeptides that can bind are also likelier to by hydrophobic,
+as most TMHs have a hydrophobic binding pocket \richel{Reference needed}.
+There are two sources of hydrophobic polypeptides, which are 
+from membrane proteins and of soluble proteins.
+Membrane proteins have at least one TMH to anchor them to
+the membrane. TMHs are hydrophobic, as this is required to span the 
+hydrophobic cellular lipid membrane. Additionally,
+they often have a length of 23 amino acids to be able to span
+the membrane.
+Soluble proteins have (non-TMH) hydrophobic regions as well, 
+to guides the protein to achieve its 3D configuration.
+
+\paragraph{MHC-I presents hydrophobic regions more often}
+
+One might expect that the hydrophobic epitopes presented,
+for a same hydrophobicity, are as likely to stem from 
+membrane proteins TMHs or from soluble proteins hydrophobic (non-TMH) regions.
+For MHC-I, however, it is found that the 9-mers stemming
+from TMHs are presented more often than expected by chance \cite{bianchi2017}.
+For MHC-II, it is unknown which percentage of binders 
+is derived from TMHs for all 13-mers present in a proteome
+that have the same hydrophobicity \richel{$\mathcal{H}_{3,1}$}.
+
+\paragraph{HLAs increase detection range}
+
+An increased detection
 is obtained by expressing a wide variety of
@@
 it would follow that this is equally important for MHC-II, 
-especially as CD4+ T cell help is needed for a long lasting CD8+ T cell 
+especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
 response \cite{novy2007cd4}. 
@@
     with TMHs as expected by chance in Mycobacterium
-  \item $\mathcal{H}_{3,1}$: The TMHs of COVID-19
+  \item $\mathcal{H}_{3,1}$: MHC-II haplotypes will display epitopes
+    of a certain hydrophobicity, that is as likely to stem from
+    a TMH or a non-TMH 13-mer as expected by chance
+  \item $\mathcal{H}_{3,2}$: MHC-II haplotypes that have hydrophobic
+    binding pockets, will display more epitopes derived from (hydrophobic)
+    TMHs
+  \item $\mathcal{H}_{3,3}$: MHC-II haplotypes that have binding pockets
+    that are only partly hydrophobic, will display more epitopes derived 
+    from the flanking regions of TMHs
+  \item $\mathcal{H}_{4,1}$: The TMHs of COVID-19
     have the same mutation rate 
     as the cytosolic and extracellular parts of their transmembrane proteins
-  \item $\mathcal{H}_{3,2}$: The TMHs of Mycobacterium
+  \item $\mathcal{H}_{4,2}$: The TMHs of Mycobacterium
     have the same mutation rate 

### 2020-04-20 article.tex
Add refs what is known on selection

@@
 \documentclass{article}
+
+% Just to have them already
+% \begin{sloppypar}
+% \end{sloppypar}
 
@@
 
-\paragraph{MHC-I}
+\paragraph{MHC-I presents TMH-derived epitopes in humans more often}
 
@@
 yet unknown.
+
+\paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
+
+It is important that MHC-I presents both the polypeptides of the 
+(healthy) cell, as well as possible pathogen-derived polypeptides when the
+cell is infected. As described above, MHC-I presents TMH-derived epitopes 
+from the human host more often. It is unknown if MHC-I has the same
+dedication to present epitopes that stem from TMHs derived from
+proteins produced by pathogen, either would the pathogen
+be a virus \richel{$\mathcal{H}_{1,1}$} or 
+a bacterium \richel{$\mathcal{H}_{1,2}$}.
 
@@
 \richel{Check against stupidity}
-Regarding human viruses, we do not know the extent to which the human
-immune systems selects on these, nor which loci are selected
-upon \richel{so I should read the literature on that}. 
-We predict that there is a strong 
-evolutionary conservation in regions that are essential for viral proteins'
-functions. Because these regions are rare in a viral proteome,
+Regarding human viruses, we do not know the selection pressure
+exerted by the human immune system.
+We do know that this selection pressure is 
+limited for seasonal influenza \cite{han2019individual},
+but this study does not show how frequent selected upon 
+TMHs \richel{Use this study, if COVID-19 dataset is too weak}.
+For COVID-19, there is preliminary evidence that the strongest
+selection pressure is upon ORF3 (an ion channel and a viroporin)
+and ORF8 (unknown function). ORF3 has 3 TMHs that appear irrelevant for
+the natural selection of COVID-19.
+We predict that evolutionary conservation is strongest 
+for regions that are essential for viral proteins'
+functions (such as catalytic binding sites), instead of TMHs.
+Because these essential regions are rare in a viral proteome,
 and differ widely between pathogens, we expect that the human immune system
-is not tailored to detect these sites. TMHs, however, have less 
-functional constraints yet are more general. Because TMHs are more general,
+is not tailored to detect these sites, similar to the aforementioned
+influenza study. 
+TMHs, however, have less 
+functional constraints compared to catalytic regions,
+and as such are more general. 
+Because TMHs are more general,
 we expect it for the immune system to pay off to be more attentive towards
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{pics/covid_genome_and_proteome.png}
+  \includegraphics[width=0.9\textwidth]{pics/covid_genome_and_proteome.png}
   \caption{
     Overview of COVID-19 genome and proteome.
-    From \url{https://zhanglab.ccmb.med.umich.edu/COVID-19}
+    From 
+    \url{https://zhanglab.ccmb.med.umich.edu/COVID-19}
   }
@@
   }
-  \label{table:bbbq_2_percentages}
+  \label{table:mhc2_haplotypes}
 \end{table}

### 2020-04-20 article.tex
Hydrophobicity added

@@
 is derived from TMHs for all 13-mers present in a proteome
-that have the same hydrophobicity \richel{$\mathcal{H}_{3,1}$}.
+that have the same hydrophobicity 
+\richel{
+  $\mathcal{H}_{3,1}$
+}.
 
@@
 
-\paragraph{MHC-II}
+\paragraph{MHC-II is expected to present TMHs}
 
@@
 similar to the discovery that MHC-I presents TMHs.
+
+\paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
+
+It is unknown if MHC-II, like MHC-I, 
+present TMH-derived epitopes as as often, either 
+in humans \richel{$\mathcal{H}_{2,1}$},
+bacteria \richel{$\mathcal{H}_{2,2}$}
+or viruses \richel{$\mathcal{H}_{2,3}$}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{Will be moved to supplementary materials}
 

### 2020-04-20 hydrophobicity/hydrophobicity.Rmd
Start exploring hydrophobicity

---
title: "Hydrophobicity"
author: "Richel J.C. Bilderbeek"
date: "4/20/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2020-04-20 hydrophobicity/hydrophobicity.Rmd
Work on hydrophobicity

@@
 ```
+
+```{r}
+library(bbbq)
+library(testthat)
+```
+
+```{r}
+proteome_filename <- system.file(
+  "extdata", 
+  "UP000001584_83332_0.fasta", 
+  package = "bbbq")
+stopifnot(file.exists(proteome_filename))
+```
+
+```{r}
+proteome <- ape::read.FASTA(proteome_filename, type = "AA")
+```
+```{r}
+library(ape)
+first_protein <- as.character(proteome)[[1]]
+first_polypeptide <- paste0(first_protein[1:13], collapse = "")
+cat(first_polypeptide)
+```
+
+```{r}
+first_hydropbobicity <- Peptides::hydrophobicity(first_polypeptide)
+cat(first_hydropbobicity)
+```

### 2020-04-23 create.sh
'create.sh' only shows last pdflatex output

@@ -1,8 +1,7 @@
 #!/bin/bash
 
-pdflatex article.tex --shell-escape
-bibtex article.aux
-pdflatex article.tex --shell-escape
+pdflatex article.tex --shell-escape >/dev/null
+bibtex article.aux >/dev/null
+pdflatex article.tex --shell-escape >/dev/null
 pdflatex article.tex --shell-escape
 
-# pdfunite article.pdf ../figures/figures.pdf article_full.pdf

### 2020-04-23 article.tex
Simplify Makefile
Describe in silico TMH AASM determination

@@
 % \end{sloppypar}
+
+% Just try to parse, do not ask for input
+\nonstopmode
 
@@
 of pathogens, whereas pathogens are selected to avoid being detected.
-If the human immune system is more prone to display (and possibly 
-detect) polypeptides derived from TMHs, one may expect to find 
-an evolutionary signal in pathogens as well.
+If the human immune system is more prone to bind hydrophobic polypeptides, 
+most of which derived from TMHs, one may expect to find 
+an evolutionary signal in TMHs.
+
+\paragraph{Amino acid substitution models}
+
+One important assumption for such phylogenetic inference
+is the amino acid substitution model (AASM), which may be different
+in TMHs. An AASM encodes the relative rates at which amino acids
+change through evolutionary time. Some amino acids have more 
+DNA triplets encoding for them, which is reflected in the AASM.
+Already for decades, AASMs are devised and refined from whole proteomes,
+such as the Blosum62 (\cite{henikoff1992amino}), 
+Dayhoff (\cite{dayhoff197822}), JTT (\cite{jones1992rapid}), 
+and WAG (\cite{whelan2001general}) models. 
+That such a generalization
+is not always desired, AASMs are devised for 
+chloroplasts (the CPREV model, see \cite{adachi2000plastid})
+and mitochondria (the MTREV24 model, see \cite{adachi1996model}).
+In all AASM, TMHs are generalized out and therefore, 
+there is no specialized AASM for TMHS. It would, however,
+be possible to derive an AASM from in silico predictions. 
+An in silico prediction would bypass any effects of evolution.
+Another AASM could be constructed from humans, if we assume
+that there is no selection on TMHs, except for being functional.
 
@@
 limited for seasonal influenza \cite{han2019individual},
-but this study does not show how frequent selected upon 
-TMHs \richel{Use this study, if COVID-19 dataset is too weak}.
+but this study does not look into the evolution of TMHs
+\richel{Use this study, if COVID-19 dataset is too weak}.
 For COVID-19, there is preliminary evidence that the strongest
@@
 TMHs.
+For viral TMHs that are undetected by most human haplotypes, we
+expect these to be conserved, to remain avoiding detection.
+For viral TMHs that are detected by most human haplotypes, we
+expect there to be a higher mutation rate, to avoid detection.
 Additionally, viruses are not only selected for their evasion of an 
@@
 by the human immune response to be weak.
+For bacterial TMHs that are undetected by most human haplotypes, we
+expect these to be conserved, to remain avoiding detection.
+For bacterial TMHs that are detected by most human haplotypes, we
+expect there to be a higher mutation rate, to avoid detection.
 Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
@@
 \subsection{Evolutionary conservation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{In silico prediction of TMH AASM}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We use a Monte-Carlo approach to produce an AASM for TMHs.
+We simulate a random amino acid sequence with a length 23.
+Using TMHMM, we determine if a sequence encodes for a TMH.
+For a TMH sequence, we independently change 
+each locus to all other amino acids.
+Of each mutation, we count how often the new sequence remains a TMH.
+Doing this for many random amino acid sequences, we can determine
+the AASM. See figure \ref{fig:tmh_aasm_method} for a schematic
+overview of this method.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{tmh_aasm_method.png}
+  \caption{
+    In silico prediction of TMH AASM
+  }
+  \label{fig:tmh_aasm_method}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Baseline TMH mutation rate}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Detecting evolution}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-04-23 article.tex
Evolution method not sure

@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fig_tmh_mut_rate.png}
+  \caption{
+    TMH mutation rates for different levels of human detection in
+    two pathogens.
+    Each point denotes a TMH.
+    The horizontal axis denotes the percentage of Homo sapiens that
+    have an HLA that binds at least one polypeptide derived from the TMH.
+    The vertical axis denotes the mutation rate observed in the TMH, 
+    per generation.
+    At the bottom-left are undetected and conserved TMHs.
+    At the top-right are TMHs that are often detected with a higher mutation
+    rate, as to escape detection.
+    The solid lines are obtained from a best-fitting linear model.
+    The dashed horizontal line denotes the average TMH mutation rate 
+    per generation in humans.
+    If there is no selection, the solid lines are horizontal.
+    If there is selection, the solid lines go upwards when going from left to right.
+    We expect selection in viruses to be stronger, thus its line will be steeper.
+    \richel{
+      Stub
+    }
+  }
+  \label{fig:tmh_mut_rate}
+\end{figure}
+
 	% Table \ref{tab:results} shows the location and binding strength for the

### 2020-04-23 article.tex
EOD

--- 2020-04-23 article.tex
@@
 That such a generalization
-is not always desired, AASMs are devised for 
+is not always desired, AASMs are divised for 
 chloroplasts (the CPREV model, see \cite{adachi2000plastid})
@@
 \end{document}
+new research. 
+      This table be deleted in the future.
+    }
+  }
+  \label{table:mhc2_haplotypes}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Kolmogorov-Smirnov}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  This is just a reminder, instead of new research. 
+  This subsection be deleted in the future.
+}
+
+The Kolmogorov-Smirnov (KS) test determines if two samples
+are derived from the same distribution, without making assumptions
+regarding the shape of that distribution.
+
+We will reject
+the null hypothesis that MHC-I has the same percentage of epitopes 
+overlapping with TMHs in Homo sapiens compared to each pathogen when 
+the KS statistic $D_{n,m}$ follows the relationship as shown in 
+equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
+and $n = m$ equals the number of HLA haplotypes.
+
+\begin{equation}
+   D_{n,m} > \frac{1}{\sqrt{n}} \cdot \sqrt{ -\ln(\frac{\alpha}{2}) \cdot \frac{1 + \frac{n}{m}}{2} }
+   \label{eq:ks}
+\end{equation}
+
+\end{document}

### 2020-04-27 article.tex
EOD

--- 2020-04-23 article.tex
@@
 
+\paragraph{Use of protein data in phylogenetics}
+
+In DNA, select
+To detect the signature of selection \cite{murrell2015gene}
+
+From \cite{anisimova2009investigating}, I quote: 
+although most proteins evolve under purifying selection most of the time, 
+positive selection may affect some lineages, and during episodes of adaptive 
+evolution, only a small fraction of sites in the protein have the capacity 
+to increase the 'fitness' of the protein via AA 
+replacements (e.g., Gillespie 1991; Li 1997; Messier and Stewart 1997; 
+Pupko and Galtier 2002). 
+Thus, approaches assuming constant selective pressure over time and 
+over sites lack power in detecting genes affected by positive 
+selection (Yang and Bielawski 2000).
+
+Commonly, DNA is used in phylogenetics.
+Some argue that the proteome is superior \cite{diz2012proteomics}:
+
 \paragraph{Amino acid substitution models}
@@
 selection pressure is upon ORF3 (an ion channel and a viroporin)
-and ORF8 (unknown function). ORF3 has 3 TMHs that appear irrelevant for
-the natural selection of COVID-19.
+and ORF8 (unknown function) \cite{velazquez2020positive}. 
+Of these two proteins, ORF3 has 3 TMHs,
+yet selection was only detectable in non-TMHs.
 We predict that evolutionary conservation is strongest 

### 2020-04-27 markov_chain/markov_chain.R

@@ -1,0 +1,44 @@
+library(markovchain)
+
+# Let's assume all protein sequences are TMH
+# Let's assume that a sequence is detected by MHC is 50%
+
+transition_matrix <- matrix(c(0.5, 0.5, 0.5, 0.5), nrow = 2, byrow = TRUE)
+markov_chain <- new(
+  "markovchain",
+  states = c("Detected","Undetected"),
+  transitionMatrix = transition_matrix
+)
+
+markov_chain_as_net <- markovchain:::.getNet(markov_chain)
+igraph::plot.igraph(
+  markov_chain_as_net,
+  vertex.color = "white",
+  vertex.frame.color = "Black",
+  vertex.label.color = "Black",
+  vertex.size = 70,
+  edge.label = round(igraph::E(markov_chain_as_net)$weight / 100, 2),
+  edge.color = "Black",
+  edge.label.color = "Black",
+  edge.arrow.size = 0.1,
+  arrow.mode = 2,
+  edge.curved = 0.5 * pi,
+  layout = igraph::layout_on_grid
+)
+
+
+draw_markov_chain <- function() {
+  dot_text <- c(
+  "digraph markov_chain {",
+  "  a [label=detected];",
+  "  b [label=undetected];",
+  "  a -> a [label=0.7 tailport=nw headport=sw];",
+  "  a -> b [label=0.3 tailport=ne headport=nw];",
+  "  b -> a [label=0.2 tailport=sw headport=se];",
+  "  b -> b [label=0.8 tailport=se headport=ne];",
+  "  { rank=same; a b }",
+  "}"
+  )
+
+  dot -Tpng markov_chain.dot > markov_chain.png
+}

### 2020-04-27 sim_random_tmh/sim_random_tmh.R

@@ -1,0 +1,7 @@
+# How likely is it to simulate a random TMH?
+library(tmhmm)
+result <- tmhmm::run_tmhmm_on_sequence(
+  "QEKNWSALLTAVVIILTIAGNILVIMAVSLEKKLQ"
+)
+
+result

### 2020-04-28 article.tex
Improve test, before removing AASM

--- 2020-04-23 article.tex
@@
 Lastly, this study shows that in the mutation rate in TMHs of a pathogen
-is [lower/as much/higher], hinting at 
+is [lower/as much/higher] then expected just by chance, hinting at 
 evolutionary [conservation/neutrality/disruptive selection].
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 Our immune system fights invaders on a daily basis.
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 All nucleated cells in humans present randomly sampled polypeptides
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 Bacteria and fungi do not infect a cell with their foreign DNA,
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 Any human's immune system detects only a fraction of all possible
@@
 For MHC-I, however, it is found that the 9-mers stemming
-from TMHs are presented more often than expected by chance \cite{bianchi2017}.
+from TMHs are presented more often than expected by 
+chance \cite{bianchi2017}
+\richel{but also when compensating for hydrophobicity?}.
 For MHC-II, it is unknown which percentage of binders 
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 For MHC-I, it was found that predicted epitopes derived 
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 If presentation of TMHs on MHC-I would bring an evolutionary advantage 
@@
 
-\paragraph{Evolutionary arms race}
+\paragraph{Selection undetectable in whole proteome}
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 The human immune system and human pathogen are in an evolutionary
-arms race, in which our immune systems is selected for the detection
-of pathogens, whereas pathogens are selected to avoid being detected.
-If the human immune system is more prone to bind hydrophobic polypeptides, 
-most of which derived from TMHs, one may expect to find 
-an evolutionary signal in TMHs.
+arms race: our immune systems is selected for the detection
+of pathogens, whereas pathogens are selected to avoid detection.
+From a pathogen's point of view, however, this struggle 
+is of only minor importance:
+in seasonal influenza, for example, the selection pressure
+exerted by the immune system was only limited \cite{han2019individual}
+\richel{but this study does not look into the evolution of TMHs}.
+In general, on would expect that evolutionary selection is strongest 
+for regions that are important for viral proteins' functions:
+in COVID-19, for example, there is preliminary evidence that the strongest
+selection pressure is upon residues that changes its virulence \cite{velazquez2020positive}
+\richel{residue 251 of ORF3a (an ion channel and a 
+viroporin, residue is not part of TMH), and position 84 of ORF8 (unknown function)}.
+The part of a pathogen's proteome that is essential, 
+only accounts for small part of its full proteome.
+Also, these essential parts differ widely between pathogens.
+Because of this scarcity and variance in targets, 
+one expects that the human immune system
+is not tailored to detect these sites, 
+as hinted by upon by the aforementioned influenza study.
+
+\paragraph{Selection may be detectable in TMHs}
+
+TMHs, on the other hand, also have their function constraints, 
+yet can occur multiple time a pathogen's proteome.
+One can safely assume a pathogen's proteome contains multiple TMHs.
+Therefore, it may be beneficial for the host
+if its immune system would be more attentive towards TMHs.
+And maybe this has already happened: MHC-I already detects hydrophobic
+polypeptides. This feature, however, may also be caused by selection
+to detect hydrophobic regions in the soluble proteins of pathogens.
+It is unknown, when focussing on TMHs only, if a signal of selection
+can be detected.
+
+\paragraph{Use of protein data in phylogenetics}
+
+When using DNA sequences, one can use dS/dN rates 
+to detect the signature of selection \cite{murrell2015gene}.
+Using AA sequences, however, has its advantages,
+as its is closer to the actual phenotype
+selection acts upon: DNA may never be translated to RNA,
+or its RNA may never be transcribed \cite{diz2012proteomics}
+\richel{improve upon this point}.
+When using a preoteome in phylogenetic research,
+we know that the majority of proteins are selected to just 
+maintain their function most of the time, where
+is the timespans there is selection, only a few AAs
+can actually increase the 'fitness' of the 
+protein \cite{anisimova2009investigating}.
+There, when generalizing the dynamics of mostly purifying selection (to maintain
+a protein's function) and a short duration of positive selection,
+those genes that are selected cannot be detected \cite{yang2000statistical}.
 
@@
 That such a generalization
-is not always desired, AASMs are divised for 
+is not always desired, AASMs are devised for 
 chloroplasts (the CPREV model, see \cite{adachi2000plastid})
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 Regarding human viruses, we do not know the selection pressure
 exerted by the human immune system.
-We do know that this selection pressure is 
-limited for seasonal influenza \cite{han2019individual},
-but this study does not look into the evolution of TMHs
-\richel{Use this study, if COVID-19 dataset is too weak}.
-For COVID-19, there is preliminary evidence that the strongest
-selection pressure is upon ORF3 (an ion channel and a viroporin)
-and ORF8 (unknown function). ORF3 has 3 TMHs that appear irrelevant for
-the natural selection of COVID-19.
-We predict that evolutionary conservation is strongest 
-for regions that are essential for viral proteins'
-functions (such as catalytic binding sites), instead of TMHs.
-Because these essential regions are rare in a viral proteome,
-and differ widely between pathogens, we expect that the human immune system
-is not tailored to detect these sites, similar to the aforementioned
-influenza study. 
-TMHs, however, have less 
-functional constraints compared to catalytic regions,
-and as such are more general. 
-Because TMHs are more general,
-we expect it for the immune system to pay off to be more attentive towards
-TMHs.
 For viral TMHs that are undetected by most human haplotypes, we
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 Regarding human pathogenic bacteria, we do not know the extent to which the 
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 To detect an evolutionary signal for selection on the TMHs of viruses, 
@@
 
-\richel{Check against stupidity}
+\richel{Check for correctness}
 For MHC-II, we do the same analysis as MHC-I, except for

### 2020-04-28 markov_chain/markov_chain.R

@@ -1,126 +1,67 @@
-library(markovchain)
+library(bbbq)
 
-# Let's assume all protein sequences are TMH
+# By chance,
 # Let's assume that a sequence is detected by MHC is 50%
+transition_matrix <- matrix(rep(0.5, 4), nrow = 2)
+bbbq::plot_markov_chain(transition_matrix)
 
-show_igraph <- function(
-  transition_matrix = matrix(c(0.9, 0.1, 0.5, 0.5), nrow = 2, byrow = TRUE)
-) {
-  markov_chain <- new(
-    "markovchain",
-    states = c("Detected","Undetected"),
-    transitionMatrix = transition_matrix
-  )
+# By chance,
+# But if you are detected, you are probably hydrophobic.
+# After a mutation, you probably still are.
+# So self-loops are likelier.
+transition_matrix[1,1] <- 0.9
+transition_matrix[1,2] <- 0.1
+transition_matrix[2,1] <- 0.1
+transition_matrix[2,2] <- 0.9
+bbbq::plot_markov_chain(transition_matrix)
 
-  markov_chain_as_net <- markovchain:::.getNet(markov_chain)
-  igraph::plot.igraph(
-    markov_chain_as_net,
-    vertex.color = "white",
-    vertex.frame.color = "Black",
-    vertex.label.color = "Black",
-    vertex.size = 70,
-    edge.label = round(igraph::E(markov_chain_as_net)$weight / 100, 2),
-    edge.color = "Black",
-    edge.label.color = "Black",
-    edge.arrow.size = 0.1,
-    arrow.mode = 2,
-    edge.curved = 0.5 * pi,
-    layout = igraph::layout_on_grid
-  )
-}
-show_igraph()
-
-show_dow <- function(
-  transition_matrix = matrix(c(0.5, 0.5, 0.5, 0.5), nrow = 2)
-) {
-  dot_text <- c(
-  "digraph markov_chain {",
-  "  a [label=detected];",
-  "  b [label=undetected];",
-  "  a -> a [label=0.7 tailport=nw headport=sw];",
-  "  a -> b [label=0.3 tailport=ne headport=nw];",
-  "  b -> a [label=0.2 tailport=sw headport=se];",
-  "  b -> b [label=0.8 tailport=se headport=ne];",
-  "  { rank=same; a b }",
-  "}"
-  )
-  dot_filename <- tempfile()
-  writeLines(text = dot_text, con = dot_filename)
[…]

### 2020-04-28 article.tex
Mention transition counts

@@
 to detect hydrophobic regions in the soluble proteins of pathogens.
-It is unknown, when focussing on TMHs only, if a signal of selection
+It is unknown, when focusing on TMHs only, if a signal of selection
 can be detected.
@@
 
-When using DNA sequences, one can use dS/dN rates 
+When using DNA sequences, one can use $dS/dN$ rates 
 to detect the signature of selection \cite{murrell2015gene}.
@@
 \richel{improve upon this point}.
-When using a preoteome in phylogenetic research,
+When using a proteome in phylogenetic research,
 we know that the majority of proteins are selected to just 
 maintain their function most of the time, where
-is the timespans there is selection, only a few AAs
+is the time spans there is selection, only a few AAs
 can actually increase the 'fitness' of the 
@@
 those genes that are selected cannot be detected \cite{yang2000statistical}.
-
-\paragraph{Amino acid substitution models}
-
-One important assumption for such phylogenetic inference
-is the amino acid substitution model (AASM), which may be different
-in TMHs. An AASM encodes the relative rates at which amino acids
-change through evolutionary time. Some amino acids have more 
-DNA triplets encoding for them, which is reflected in the AASM.
-Already for decades, AASMs are devised and refined from whole proteomes,
-such as the Blosum62 (\cite{henikoff1992amino}), 
-Dayhoff (\cite{dayhoff197822}), JTT (\cite{jones1992rapid}), 
-and WAG (\cite{whelan2001general}) models. 
-That such a generalization
-is not always desired, AASMs are devised for 
-chloroplasts (the CPREV model, see \cite{adachi2000plastid})
-and mitochondria (the MTREV24 model, see \cite{adachi1996model}).
-In all AASM, TMHs are generalized out and therefore, 
-there is no specialized AASM for TMHS. It would, however,
-be possible to derive an AASM from in silico predictions. 
-An in silico prediction would bypass any effects of evolution.
-Another AASM could be constructed from humans, if we assume
-that there is no selection on TMHs, except for being functional.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{In silico prediction of TMH AASM}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We use a Monte-Carlo approach to produce an AASM for TMHs.
-We simulate a random amino acid sequence with a length 23.
-Using TMHMM, we determine if a sequence encodes for a TMH.
-For a TMH sequence, we independently change 
-each locus to all other amino acids.
-Of each mutation, we count how often the new sequence remains a TMH.
-Doing this for many random amino acid sequences, we can determine
-the AASM. See figure \ref{fig:tmh_aasm_method} for a schematic
-overview of this method.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{tmh_aasm_method.png}
-  \caption{
-    In silico prediction of TMH AASM
-  }
-  \label{fig:tmh_aasm_method}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Baseline TMH mutation rate}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Detecting evolution}
@@
 	% }
-	% .
-	% Using the \verb;tmhprot; R package (\cite{tmhprot}), the DNA alignment is split
-	% into two alignments, one for the TMH parts, another for the non-TMH
-	% parts. Each alignment was tested by the \verb;mcbette; 
-	% R package (\cite{mcbette}) to select the Bayesian inference model with
-	% the highest evidence (a.k.a. the marginal likelihood) using the nested
-	% sampling approach as described in \cite{maturana2018model},
-	% using the popular Bayesian phylogenetic tool 
-	% BEAST2 (\cite{bouckaert2014beast}) in the back-end.
-	% \verb;mcbette; used a set of 40 candidate inference models, 
-	% consisting of all combinations of 
-	% 4 site models (JC, HKY, TN, GTR), 
-	% 2 clock models (strict, RLN) and 
-	% 5 tree priors (Yule, BD, CBS, CCP, CEP).
-
-	% For each alignment, the inference model with the highest evidence
-	% is used in a Bayesian inference. From each Bayesian posterior,
-	% the parameter estimates regarding mutation rates (including clock rate)
-	% were obtained and compared to determine which realized mutation rate is lower.
-	% The Markov chain Monte Carlo was set up in such a way that the effective sample
-	% size for the likelihood of the inference model is above the recommended value
-	% of 200 (\cite{bouckaert2014beast}).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Poisson binomial distribution
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{transitions.png}
+  \caption{
+    Transitions in TMHs that occur more often then expected by chance,
+    for a confidence interval of 5 percent.
+    Table \ref{tab:transitions} shows the transition counts.
+    \richel{
+      stub figure
+    }
+  }
+  \label{fig:transitions}
+\end{figure}
+
+\begin{table}[!htbp]
+  \input{table_transitions.latex}
+  \caption{
+    Transitions counts, where the row indicates the source state,
+    and the column indicates the target state.
+    First number per cell is the observed number of this state transition,
+    where the second number is the expected number of this state transition
+    as predicted by chance.
+    An asterisk behind the observed count indicates that this count
+    is unlikely to be caused by chance only.
+    Figure \ref{fig:transitions} shows which transition counts are 
+    unlikely to be caused by chance only.
+  }
+  \label{tab:transitions}
+\end{table}
 

### 2020-04-29 article.tex
Improve abstract

@@
 histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
-It is unknown if this same overrepresentation
-also holds for the proteome of a pathogen.
-Additionally, it is unknown if TMH-II is also likelier to present
-TMHs than expected by chance. 
-Lastly, if the immune system if more tailored to detect polypeptides
-derived from TMHs, can we confirm this by detecting the evolutionary
-conservation of this behavior?
+It is unknown if the immune system is as able to detect those 
+TMHs present in pathogens.
+Additionally, it is unknown if TMH-II is also likelier to detect
+human and pathogen TMHs than expected by chance.
+It is unknown if natural selection is the cause that 
+the immune system is more attentive for TMHs
 This study shows that MHC-I [also has/does not have] more
@@
 [less/equally/more] often than expected by chance.
-Lastly, this study shows that in the mutation rate in TMHs of a pathogen
-is [lower/as much/higher] then expected just by chance, hinting at 
-evolutionary [conservation/neutrality/disruptive selection].
+Lastly, we show that natural selection on TMHs is [strong enough/still too weak]
+to detect.
 Our findings suggest that the immune system is [less/neutral/more]
-vigilant to TMHs than expected by chance and this has left [a/no]
+vigilant to TMHs than expected by chance and this has 
+left [a clear/a weak/no]
 signal in the evolutionary history of the pathogen.

### 2020-04-29 article.tex
HLA classification and epitope prediction

@@
 
-\richel{Check for correctness}
 Our immune system fights invaders on a daily basis.
@@
 
-\richel{Check for correctness}
 All nucleated cells in humans present randomly sampled polypeptides
@@
 If a cell gets infected by a virus, also the virus' polypeptides
-will be presented at the cell surface. The viral antigen is detected 
+will be presented at the cell surface. The foreign viral antigen is detected 
 by cytotoxic T lymphocytes, which will kill the infected cell.
@@
 
-\richel{Check for correctness}
-Bacteria and fungi do not infect a cell with their foreign DNA,
-but are detected by the proteins on their cell walls. MHC class II molecules 
-bind to specific polypeptide fragments. This binding allows T and B cells,
-which all express MHC-II, to detect a pathogen and trigger the response to
-kill the pathogen. Note that the foreign proteins of a viral envelope 
-can be detected in this way as well, which is commonly used in vaccination.
+All pathogens can be detected by the foreign proteins on 
+their (bacterial or fungal) cell walls or (viral) envelope.
+For bacteria and fungi, this is the main mode of their detection,
+as these do not infect a cell with their (foreign) DNA.
+T and B cells express MHC-II to detect foreign polypeptide fragments.
+An immune response is started when MHC-II detects a pathogen.
 
-\paragraph{MHCs bind hydrophobic epitopes}
+\paragraph{Epitope prediction}
 
-\richel{Check for correctness}
 Any human's immune system detects only a fraction of all possible
@@
 subset of all possible polypeptides.
+The classification of HLAs is based on algorithms that
+maximize the information content of each 
+classification, such as the presence of a certain amino acid at 
+the first (for MHC-II, \cite{southwood1998several})
+or second position (for MHC-I, \cite{lund2004definition}) of a polypeptide.
+However, this does not imply that all polypeptides for a classification
+pattern bind to the MHC, as there is no simple method to predict which 
+MHCs bind which epitopes.
+Using artificial neural networks, it became
+possible to make theoretical predictions \cite{nielsen2003reliable,nielsen2004improved},
+which are tested to be reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+
+??? 
 The polypeptides that can bind are also likelier to by hydrophobic,
@@
 
-\richel{Check for correctness}
 For MHC-I, it was found that predicted epitopes derived 
@@
 
-\richel{Check for correctness}
 If presentation of TMHs on MHC-I would bring an evolutionary advantage 
@@
 
-\richel{Check for correctness}
 The human immune system and human pathogen are in an evolutionary
@@
 
-\richel{Check for correctness}
 Regarding human viruses, we do not know the selection pressure
@@
 
-\richel{Check for correctness}
 Regarding human pathogenic bacteria, we do not know the extent to which the 
@@
 
-\richel{Check for correctness}
 To detect an evolutionary signal for selection on the TMHs of viruses, 
@@
 
-\richel{Check for correctness}
 For MHC-II, we do the same analysis as MHC-I, except for
@@
 will not be presented by MHC-I, this result is [unexpected/expected]
+
+We aimed our evolutionary experiment at TMHs, because these can
+be predicted well from a protein structure,
+are common structures and are present in all pathogens. 
+We could have done the same experiment on beta-turn,
+as also these can be predicted well \cite{petersen2010netturnp},
+are common structures and are present in all pathogens.
 

### 2020-04-29 article.tex
There is no empirical evidence that hydrophobic residues bind more often, lund2004 only uses hydrophobic residues for classification

@@
 Already each MHC molecule can only bind a
-subset of all possible polypeptides.
+subset of all possible polypeptides (but it has been observed that HLA-A and HLA-B have no overlap in which
+peptides they bind, \cite{lund2004definition}).
 The classification of HLAs is based on algorithms that
@@
 
-??? 
-The polypeptides that can bind are also likelier to by hydrophobic,
-as most TMHs have a hydrophobic binding pocket \richel{Reference needed}.
+\richel{This is nonsense: 'The polypeptides that can bind are also 
+likelier to by hydrophobic,
+as most TMHs have a hydrophobic binding pocket': \cite{lund2004definition}
+uses the hydrophobicity of some peptides for classification, but this
+does not mean the statement is true}
 There are two sources of hydrophobic polypeptides, which are 

### 2020-04-30 calc_p_transition/calc_p_transition.R
EOD

@@ -17,42 +17,15 @@
 # * Keep the ones that result in a TMH
 # * Of those, count which state these are
 # * Convert these counts to probabilities
+library(bbbq)
 library(testthat)
 
-#' @param s the string
-#' @param n the index of the character
-#' @param c the character to be placed at that index
-replace_nth_char <- function(s, n, c) {
-  letter <- c
-  # 1: The first n-1 chars,
-  # 2: The character to be replaces
-  # 3: The rest
-  lhs <- paste0('^([A-Z]{', n - 1, '})(.)([A-Z]+)?$')
-  rhs <- paste0('\\1', letter, '\\3')
-  gsub(lhs, rhs, s)
-}
-expect_equal("XBC", replace_nth_char("ABC", 1, "X"))
-expect_equal("AXC", replace_nth_char("ABC", 2, "X"))
-expect_equal("ABX", replace_nth_char("ABC", 3, "X"))
-
-get_adjacent_sequence <- function(aa_sequence = "VVIILTIAGNILVIMAVSLE")
-{
-  seqs <- rep(NA, nchar(aa_sequence) * 20)
-  i <- 1
-  for (locus in seq(1, nchar(aa_sequence))) {
-    for (aa in bbbq::get_amino_acids()) {
-      seqs[i] <- replace_nth_char(s = aa_sequence, n = locus, c = aa)
-      i <- i + 1
-    }
-  }
-  seqs[ seqs != aa_sequence ]
-}
 
 focal_sequence <- "VVIILTIAGNILVIMAVSLE"
 expect_true(bbbq::is_detected_by_mhc_1(focal_sequence))
 expect_false(bbbq::is_detected_by_mhc_2(focal_sequence))
 
-all_seqs <- get_adjacent_sequence(focal_sequence)
+all_seqs <- get_adjacent_sequences(focal_sequence)
 length(all_seqs) #380
 
 all_tmhs <- all_seqs[ tmhmm::are_tmhs(all_seqs) ]

### 2020-04-30 p_bind_per_hydrophobicity/p_bind_per_hydrophobicity.R

@@ -16,21 +16,6 @@ create_random_polypeptide <- function(n_aas)
   )
 }
 
-#' Create a random hydrophobic polypeptide
-#' @param n_aas number of amino acids
-create_random_hydrophobic_polypeptide <- function(n_aas)
-{
-  aas <- bbbq::get_amino_acids()
-  hs <- Peptides::hydrophobicity(aas)
-  prob <- hs - min(hs)
-  prob <- prob * prob
-  paste0(
-    sample(
-      x = aas, size = n_aas, replace = TRUE, prob = prob),
-    collapse = ""
-  )
-}
-
 create_random_extreme_polypeptide <- function(n_aas)
 {
   aas <- bbbq::get_amino_acids()

### 2020-05-04 article.tex
EOD
Research hydrophobicity vs TMH vs binds MHC-I vs binds MHC-II

@@
 It is unknown if MHC-II, like MHC-I, 
-present TMH-derived epitopes as as often, either 
+presents TMH-derived epitopes as often, either 
 in humans \richel{$\mathcal{H}_{2,1}$},
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Transition matrix from \cite{dang2010flu}.
 

### 2020-05-04 p_bind_per_hydrophobicity/p_bind_per_hydrophobicity.R

@@ -6,48 +6,174 @@
 #
 library(ggplot2)
 
-#' Create a dataset
-#' @param n number of polypeptides
-#' @param n_aas number of amino acids per polypeptide
-#' @param f function to create a polypeptide
-create_dataset <- function(
-  n,
-  n_aas,
-  f
-) {
-  df <- tibble::tibble(
-    hydrophobicity = rep(NA, n),
-    is_tmh = NA,
-    binds_mhc1 = NA,
-    binds_mhc2 = NA
-  )
-  df$hydrophobicity <- as.numeric(df$hydrophobicity)
-  df$is_tmh <- as.numeric(df$is_tmh)
-  df$binds_mhc1 <- as.numeric(df$binds_mhc1)
-  df$binds_mhc2 <- as.numeric(df$binds_mhc2)
-  for (i in seq(1, n)) {
-    polypeptide <- f(n_aas)
-    df$hydrophobicity[i] <- Peptides::hydrophobicity(polypeptide)
-    df$is_tmh[i] <- tmhmm::is_tmh(polypeptide)
-    df$binds_mhc1[i] <- bbbq::is_detected_by_mhc_1(polypeptide)
-    df$binds_mhc2[i] <- bbbq::is_detected_by_mhc_2(polypeptide)
+if (1 == 2) {
+  #' Create a dataset
+  #' @param n number of polypeptides
+  #' @param n_aas number of amino acids per polypeptide
+  #' @param f function to create a polypeptide
+  create_dataset <- function(
+    n,
+    n_aas,
+    f
+  ) {
+    df <- tibble::tibble(
+      hydrophobicity = rep(NA, n),
+      is_tmh = NA,
+      binds_mhc1 = NA,
+      binds_mhc2 = NA
+    )
+    df$hydrophobicity <- as.numeric(df$hydrophobicity)
+    df$is_tmh <- as.numeric(df$is_tmh)
+    df$binds_mhc1 <- as.numeric(df$binds_mhc1)
+    df$binds_mhc2 <- as.numeric(df$binds_mhc2)
+    for (i in seq(1, n)) {
+      polypeptide <- f(n_aas)
+      df$hydrophobicity[i] <- Peptides::hydrophobicity(polypeptide)
+      df$is_tmh[i] <- tmhmm::is_tmh(polypeptide)
+      df$binds_mhc1[i] <- bbbq::is_detected_by_mhc_1(polypeptide)
[…]

### 2020-05-04 article.tex
make figures in 'p_bind_per_hydrophobicity'

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Relation between hydrophobicity, TMH and MHC binding}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/binds_mhc1_vs_binds_mhc2.png}
+  \caption{
+    \richel{Used simulated data}
+  }
+  \label{fig:binds_mhc1_vs_binds_mhc2}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc1.png}
+  \caption{
+    \richel{Used simulated data}
+  }
+  \label{fig:hydrophobicity_vs_binds_mhc1}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc2.png}
+  \caption{
+    \richel{Used simulated data}
+  }
+  \label{fig:hydrophobicity_vs_binds_mhc2}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_is_tmh.png}
+  \caption{
+    \richel{Used simulated data}
+  }
+  \label{fig:hydrophobicity_vs_is_tmh}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc1.png}
+  \caption{
+    \richel{Used simulated data}
+  }
+  \label{fig:is_tmh_vs_binds_mhc1}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc2.png}
+  \caption{
+    \richel{Used simulated data}
+  }
+  \label{fig:is_tmh_vs_binds_mhc2}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{COVID-19 TMHs}

### 2020-05-05 article.tex
Add Frans' idea: not all epitopes are responded upon

@@
 
-When using DNA sequences, one can use $dS/dN$ rates 
+When using DNA sequences, one can use a skewed rate 
+in non- versus synonymous mutation
 to detect the signature of selection \cite{murrell2015gene}.
@@
 Transition matrix from \cite{dang2010flu}.
+
+We assume that if a polypeptide binds to an MHC, it is always presented and always
+leads to an immune response. We know this is not the case \richel{for example,
+I know this from Frans :-)}, but we do not know how common this is 
+\richel{read literature to find out}. The more often this phenomenon occurs,
+the more our analysis is weakened.
 
@@
   \caption{
-    \richel{Used simulated data}
+    \richel{Used randomly simulated polypeptides, results are real}
   }

### 2020-05-05 article.tex
MHC-I epitopes that are not immunogenic are non-existing or rare

@@
 [less/equally/more] often than expected by chance.
-Lastly, we show that natural selection on TMHs is [strong enough/still too weak]
+Lastly, using an unprecedented sharp method, 
+we show that natural selection on TMHs is [strong enough/still too weak]
 to detect.
@@
 Already each MHC molecule can only bind a
-subset of all possible polypeptides (but it has been observed that HLA-A and HLA-B have no overlap in which
+subset of all possible polypeptides (but it has been observed 
+that HLA-A and HLA-B have no overlap in which
 peptides they bind, \cite{lund2004definition}).
@@
 which are tested to be reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+
+We can predict epitopes.
+For MHC-I, \verb;epitope-prediction; is used \richel{Cite Johannes}.
+For MHC-II, NetMHCIIpan uses all haplotypes \richel{Cite Morten}.
+Although \cite{zhang2015accurate} looks like an even better
+method for MHC-II, they only take one haplotype into account.
 
@@
 We assume that if a polypeptide binds to an MHC, it is always presented and always
-leads to an immune response. We know this is not the case \richel{for example,
-I know this from Frans :-)}, but we do not know how common this is 
-\richel{read literature to find out}. The more often this phenomenon occurs,
-the more our analysis is weakened.
+leads to an immune response.
+The more often this assumption is violated, 
+the more our analysis will be weakened.
+Experiments in mice with MHC-I epitopes, however, indicate that (if the dose 
+administered was high enough) there are either none or rare
+exceptions \cite{sette1994relationship}.
 

### 2020-05-05 article.tex
Investigate the tools better

@@
 For MHC-I, \verb;epitope-prediction; is used \richel{Cite Johannes}.
-For MHC-II, NetMHCIIpan uses all haplotypes \richel{Cite Morten}.
+
+Alternatives are NetMHCcons \cite{karosiene2012netmhccons}.
+
+For MHC-II, NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
+uses all haplotypes.
+Also \cite{zhang2013predicting} predicts immunogenic T-cell epitopes.
+Also \cite{trolle2014nettepi} predicts immunogenic T-cell epitopes.
+
 Although \cite{zhang2015accurate} looks like an even better
 method for MHC-II, they only take one haplotype into account.
+
+According to \cite{lundegaard2011prediction}, in 2011, 
+NetMHCcons and NetMHCIIpan were the best.
 

### 2020-05-05 article.tex
HLAs described better

--- 2020-05-05 article.tex
@@
 
-\paragraph{Epitope prediction}
+\paragraph{Classification of HLA}
 
@@
 polypeptide fragments. 
-Already each MHC molecule can only bind a
-subset of all possible polypeptides (but it has been observed 
-that HLA-A and HLA-B have no overlap in which
-peptides they bind, \cite{lund2004definition}).
-The classification of HLAs is based on algorithms that
+Human MHCs are also called HLAs ('Human Leukocyte Antigens') and
+because this research uses human hosts only, we will use these terms
+interchangeably.
+Already each MHC molecule can only bind a, possible exclusive,
+subset of all possible polypeptides.
+For example, HLA-A and HLA-B have no overlap in which
+peptides they bind (\cite{lund2004definition}).
+The HLA region of humans is highly polymorphic, 
+making it hard to classify all of the many haplotypes.
+Classification of HLAs is based on algorithms that
 maximize the information content of each 
@@
 However, this does not imply that all polypeptides for a classification
-pattern bind to the MHC, as there is no simple method to predict which 
-MHCs bind which epitopes.
-Using artificial neural networks, it became
-possible to make theoretical predictions \cite{nielsen2003reliable,nielsen2004improved},
-which are tested to be reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+pattern bind to the MHC. Instead, software should be used that are
+tailored to predict if a polypeptide binds to an HLA.
 
-We can predict epitopes.
-For MHC-I, \verb;epitope-prediction; is used \richel{Cite Johannes}.
-For MHC-II, NetMHCIIpan uses all haplotypes \richel{Cite Morten}.
-Although \cite{zhang2015accurate} looks like an even better
-method for MHC-II, they only take one haplotype into account.
+\paragraph{Epitope prediction}
+
+It is helpful to be able to predict which polypeptides are immunogenic,
+in, among others, vaccine development \richel{Reference here}. 
+Already for a decade, synthesized polypeptide fragments are used 
+to experimentally determine which polypeptides
+are immunogenic \richel{Reference here}.
+This approach, however, is tedious and costly.
+It should come as no surprise that these results would be used
+to make in silico predictions.
+It may come as a surprise, however, that these predictions
+are actually reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+
+\paragraph{MHC-I epitope prediction}
+
+For MHC-I, there are multiple computational tools developed 
+to predict epitopes. 
+According to \cite{lundegaard2011prediction}, in 2011,
+from a set of multiple tools, 
+NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
+A tool not included in this review is \verb;epitope-prediction;,
+which uses a stabilized matrix method, as described in \cite{kim2009derivation}.
+The reason this tool is mentioned, is because it is used 
+in cite{bianchi2017}.
+
+\paragraph{MHC-II epitope prediction}
+
+Also for MHC-II, there are multiple computational tools developed 
+to predict epitopes,
+such as using a trained neural network (\cite{nielsen2003reliable})
+or a Gibbs sampling approach (\cite{nielsen2004improved}).
+According to \cite{lundegaard2011prediction}, in 2011,
+from a set of multiple tools, 
+NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
+gave rise to the most accurate predictions.
+Later tools are \cite{zhang2013predicting},
+\cite{trolle2014nettepi} and \cite{zhang2015accurate}.
+These later tools, although they look promising, have not been compared
+with one another \richel{Maybe cool to do so!}.
+
+\paragraph{Leftovers (to move or remove)}
 
@@
 to guides the protein to achieve its 3D configuration.
+
+\paragraph{TMH prediction}
+
+\richel{Or someplace else. Reorganise!}
 

### 2020-05-05 article.tex
Add TMH predicting methods

@@
 The reason this tool is mentioned, is because it is used 
-in cite{bianchi2017}.
+in \cite{bianchi2017}.
 
@@
 
-\paragraph{Leftovers (to move or remove)}
+\paragraph{TMHs}
 
-\richel{This is nonsense: 'The polypeptides that can bind are also 
-likelier to by hydrophobic,
-as most TMHs have a hydrophobic binding pocket': \cite{lund2004definition}
-uses the hydrophobicity of some peptides for classification, but this
-does not mean the statement is true}
-There are two sources of hydrophobic polypeptides, which are 
-from membrane proteins and of soluble proteins.
-Membrane proteins have at least one TMH to anchor them to
-the membrane. TMHs are hydrophobic, as this is required to span the 
+Transmembrane helices are conserved structures that span
+a cell membrane with an alpha helix.
+TMHs are hydrophobic, as this is required to span the 
 hydrophobic cellular lipid membrane. Additionally,
@@
 the membrane.
-Soluble proteins have (non-TMH) hydrophobic regions as well, 
-to guides the protein to achieve its 3D configuration.
+Polypeptide fragments derived from TMHs are among the most hydrophobic,
+together with the internals of soluble proteins, where the
+hydrophobicity parts guide the protein to achieve its 3D configuration.
+TMHs are general structures: 25 percent of the human proteome is
+anchored by at least one TMH. COVID-19 has 21 TMHs, 
+making up for \richel{?} percent of its entire genome.
 
@@
 
-\richel{Or someplace else. Reorganise!}
+There are multiple computational tools developed to detect which
+parts of membrane proteins are TMH (as well as which side is inside
+and outside).
+
+In 2001, multiple tools to do so have been compared \cite{moller2001evaluation},
+of which TMHMM \cite{krogh2001predicting} turned out to be the best.
+A later method such as \cite{kall2004combined} \richel{I should check}.
+Other later methods, such as ConPred II (\cite{arai2004conpred})
+and MetaTM (\cite{klammer2009metatm}) show promises, but
+appear to be unavailable \richel{Uh, both servers are down?!?!}.
 

### 2020-05-05 article.tex
PureseqTM is FOSS and better than TMHMM

@@
 There are multiple computational tools developed to detect which
-parts of membrane proteins are TMH (as well as which side is inside
-and outside).
-
+parts of membrane proteins are TMH
 In 2001, multiple tools to do so have been compared \cite{moller2001evaluation},
 of which TMHMM \cite{krogh2001predicting} turned out to be the best.
-A later method such as \cite{kall2004combined} \richel{I should check}.
-Other later methods, such as ConPred II (\cite{arai2004conpred})
-and MetaTM (\cite{klammer2009metatm}) show promises, but
-appear to be unavailable \richel{Uh, both servers are down?!?!}.
+Many other tools followed, such as 
+Phobius (\cite{kall2004combined}),
+ConPred II (\cite{arai2004conpred}),
+MEMSAT3 (\cite{jones2007improving})
+and MetaTM (\cite{klammer2009metatm}), which signals the importance
+of TMH predictions. 
+Unlike TMHMM, which is still up and running, 
+these later tools have already become obsolete.
+Contemporary tools that are running, such as
+the closed-source MemBrain \cite{feng2020topology} 
+and the FOSS tool PureseqTM \cite{wang2019efficient}.
+\richel{Will use TMHMM for starters (as it is already programmed),
+then use PureseqTM iff time}
 

### 2020-05-05 article.tex
Begin of method to determine evolution

@@
 \richel{Will use TMHMM for starters (as it is already programmed),
-then use PureseqTM iff time}
+then use PureseqTM if time}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evolutionary conservation}
+\subsubsection{Detecting evolution}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Introduction}
+
+To detect evolution, we determine the probability 
+that all of the observed TMH mutations have arisen by chance,
+with regards to being detected/presented by the immune system.
+If this probability is lower than 5\%, we will state 
+that this unlikely situation is most likely caused by natural selection.
+
+\input{markov_chain.latex}
 
@@
 exceptions \cite{sette1994relationship}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Detecting evolution}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-	% \frans{
-	%   Ik heb zitten denken over de evolutie naast conservering is occurrence ook heel belangrijk. Moeten we de mate van conservering niet wegen ten opzichte van de occurrence.
-	%   Wat ik daarmee bedoel is data als je kiest voor een mega geconseveerd epitope maar van dat type komen er maar 40 in een proteome voor versus een minder geconserveerd epitope maar daar zijn wel 100.000 varianten van dan heb je meer kans om 1 van die 100.000 te pakken toch?
-	% }
-	% The evolutionary conservation of TMHs was measured from a
-	% DNA alignment of multiple transmembrane proteins in multiple
-	% species of the mycobacterium bacterial family
-	% \richel{how obtained exactly? how alignment done?}
-	% \frans{
-	% Via entrez en blast 
-	%   Ik he been transcript van Maxim ik zal dit proberen te vertalen naar een stukje tekst. 
-	% }
 

### 2020-05-05 edit.sh
Pretty Markov chain

@@ -7,5 +7,5 @@
 #   ./edit.sh
 #
 
-mousepad article.tex
+mousepad article.tex article.bib markov_chain.latex
 make

### 2020-05-05 article.tex
Pretty Markov chain
Markov chain has a good caption

@@
 
+\begin{figure}
+  \centering
   \input{markov_chain.latex}
+  \caption{
+    State transitions in TMHs caused by mutations, with respect to the
+    detection by the immune system. Any TMH is either in a detected, $D$, or
+    undetected, $U$, state. Mutations cause state transitions with probability
+    $p$. For example, $p_{DU}$ denotes the probability that a detected
+    TMH mutations into an undetected one.
+  }
+  \label{fig:markov_chain}
+\end{figure}
 

### 2020-05-06 article.tex
Evolution method describe for two-state Markov chain

@@
 with regards to being detected/presented by the immune system.
-If this probability is lower than 5\%, we will state 
-that this unlikely situation is most likely caused by natural selection.
+To do so, we calculate the probability of observing the complete
+dataset, for which we need to calculate the probability 
+of each individual mutation.
+
+\paragraph{System depicted as a Markov chain}
+
+To calculate the probability all of our observations happened by chance,
+we view our experiment as a two-state Markov chain, as depicted in
+figure \ref{fig:markov_chain}. The two nodes depict the
+two possible detection state a polypeptide is in, which is
+either undetected $U$, or detected $D$ by the immune system.
+The edges depict the probability of a state transition by mutation,
+for example, $p_{DU}$ denotes the probability that a detected
+polypeptide mutates into an undetected one.
+We focus on polypeptides that are TMH and we assume a mutation that 
+renders the TMH non-functional as lethal.
 
@@
     $p$. For example, $p_{DU}$ denotes the probability that a detected
-    TMH mutations into an undetected one.
+    TMH mutates into an undetected one.
   }
@@
 
-Transition matrix from \cite{dang2010flu}.
+\paragraph{Statistics}
+
+When we have calculated the probability of each transition (as described below),
+we can simply count the number of observed transitions.
+Because the probability of transitions differ per polypeptide,
+the total system has the properties of a Poisson binomial distribution,
+where, for example, the expected number of transitions is the sum of individual
+probabilities. We calculate the probability of the observing our
+number of transitions.
+If this probability is lower than 5\%, we will state 
+that this unlikely situation is most likely caused by natural selection.
+
+\paragraph{Calculating one polypeptide's probability}
+
+We calculate the probability of a transition per polypeptide.
+For any polypeptide, we can generate all sequences that differ
+in only one amino acids. Of these sequences, we only keep the ones
+that result in a TMH. Of each of the remaining TMH sequences, we predict
+which mutants are detected by the immune system. Because not all 
+AA mutations are equally likely, we correct for the AA transition
+rates using the FLU transition matrix (\cite{dang2010flu}),
+which is a transition matrix derived from observed influenza AA mutations.
+
+\paragraph{Assumption that epitopes are immunogenic}
 

### 2020-05-06 edit.sh

@@ -7,5 +7,5 @@
 #   ./edit.sh
 #
 
-mousepad article.tex article.bib markov_chain.latex
+mousepad article.bib article.tex
 make

### 2020-05-06 figures_only_create.sh
Can view figures standalone

@@ -1,0 +1,5 @@
+#!/bin/bash
+pdflatex figures_only.tex
+pdflatex figures_only.tex
+pdflatex figures_only.tex
+

### 2020-05-06 figures_only_view.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+qpdfview figures_only.pdf

### 2020-05-06 figures_only_create.sh
Go towards a four-state Markov chain

@@ -2,4 +2,4 @@
 pdflatex figures_only.tex
 pdflatex figures_only.tex
 pdflatex figures_only.tex
-
+./figures_only_view.sh

### 2020-05-06 article.tex
Go towards a four-state Markov chain
Pretty four state Markov chain
Add Bonferroni correction

@@
   \centering
-  \input{markov_chain.latex}
+  \input{markov_chain_4_states.latex}
   \caption{
@@
 the total system has the properties of a Poisson binomial distribution,
-where, for example, the expected number of transitions is the sum of individual
-probabilities. We calculate the probability of the observing our
+from which we can calculate the probability of the observing our
 number of transitions.
-If this probability is lower than 5\%, we will state 
+
+\paragraph{Hypotheses testing}
+
+Of the complete system, we test for each of the 12 state transitions (that is,
+we do not test for mutations that result in the same state).
+For each transition rate, we know the expected and observed number
+of transitions. If the observed number of transitions is below the expected
+value, we test for finding that (low) number of transitions or less.
+If the observed number of transitions is above the expected
+value, we test for finding that (high) number of transitions or more.
+We use a significance level $\alpha$ of 5\%, because we have no prior
+evidence. 
+As we are testing 12 hypothesis, applying the Bonferroni correction results
+in $\alpha_c = \frac{\alpha}{12}$.
+If the probability of observing that few/many 
+transition is below $\alpha_c$, we will state 
 that this unlikely situation is most likely caused by natural selection.

### 2020-05-06 article.tex
Separate MHC-I and MHC-II, resulting in a 2 state Markov chain
Add Holm-Bonferroni correction, describe number of expected significant results

@@
 that all of the observed TMH mutations have arisen by chance,
-with regards to being detected/presented by the immune system.
-To do so, we calculate the probability of observing the complete
-dataset, for which we need to calculate the probability 
-of each individual mutation.
+with regards to being detected/presented by the immune system,
+per MHC class and haplotype.
+Because we use \richel{X} MHC-I and \richel{Y} MHC-II haplotypes
+for a significance level of 5\%, by chance we expect \richel{X+Y/20}
+times to find a haplotype that is suggested to have selection working 
+upon it.
 
@@
 either undetected $U$, or detected $D$ by the immune system.
+The $D$ can be either detection by an MHC-I or MHC-II haplotype.
 The edges depict the probability of a state transition by mutation,
@@
   \centering
-  \input{markov_chain_4_states.latex}
+  \input{markov_chain_2_states.latex}
   \caption{
@@
 evidence.
-As we are testing 12 hypothesis, applying the Bonferroni correction results
-in $\alpha_c = \frac{\alpha}{12}$.
+
+As we are testing 12 hypothesis, 
+we correct against type I errors (false positives), 
+by applying the Holm-Bonferroni correction, 
+resulting in \richel{here is just a normal Bonferroni correction, for now}
+$\alpha_c = \frac{\alpha}{12}$.
 If the probability of observing that few/many 
 transition is below $\alpha_c$, we will state 
-that this unlikely situation is most likely caused by natural selection.
+that this unlikely situation may be caused by natural selection.
 

### 2020-05-06 article.tex
Correct against many tests

@@
 per MHC class and haplotype.
+
+\paragraph{Correction against many tests}
+
 Because we use \richel{X} MHC-I and \richel{Y} MHC-II haplotypes
 for a significance level of 5\%, by chance we expect \richel{X+Y/20}
-times to find a haplotype that is suggested to have selection working 
-upon it.
+times to find a (MHC-I or MHC-II) haplotype 
+that is suggested to have selection working 
+upon it. Applying a binomial distribution ($p = 0.05$, $n = X+Y$, $\alpha = 0.05$),
+we will claim we detected selection when we found \richel{Z} or
+more examples of having a significant result.
 

### 2020-05-06 article.tex
Do evolution with human, viral and bacterial TMHs
Simplified hypotheses

@@
 \begin{itemize}
-  \item $\mathcal{H}_{1,1}$: MHC-I has the same percentage of epitopes overlapping
-    with TMHs in Homo sapiens as in COVID-19
-  \item $\mathcal{H}_{1,2}$: MHC-I has the same percentage of epitopes overlapping
-    with TMHs in Homo sapiens as in Mycobacterium
-  \item $\mathcal{H}_{2,1}$: MHC-II has the same percentage of epitopes overlapping
-    with TMHs as expected by chance in Homo sapiens
-  \item $\mathcal{H}_{2,2}$: MHC-II has the same percentage of epitopes overlapping
-    with TMHs as expected by chance in COVID-19
-  \item $\mathcal{H}_{2,3}$: MHC-II has the same percentage of epitopes overlapping
-    with TMHs as expected by chance in Mycobacterium
-  \item $\mathcal{H}_{3,1}$: MHC-II haplotypes will display epitopes
-    of a certain hydrophobicity, that is as likely to stem from
-    a TMH or a non-TMH 13-mer as expected by chance
-  \item $\mathcal{H}_{3,2}$: MHC-II haplotypes that have hydrophobic
-    binding pockets, will display more epitopes derived from (hydrophobic)
-    TMHs
-  \item $\mathcal{H}_{3,3}$: MHC-II haplotypes that have binding pockets
-    that are only partly hydrophobic, will display more epitopes derived 
-    from the flanking regions of TMHs
-  \item $\mathcal{H}_{4,1}$: The TMHs of COVID-19
-    have the same mutation rate 
-    as the cytosolic and extracellular parts of their transmembrane proteins
-  \item $\mathcal{H}_{4,2}$: The TMHs of Mycobacterium
-    have the same mutation rate 
-    as the cytosolic and extracellular parts of their transmembrane proteins
+  \item $\mathcal{H}_{1,VB}$: MHC-I has the same percentage of epitopes overlapping
+    with viral/bacterial TMHs as with human TMHs
+  \item $\mathcal{H}_{2,HVB}$: MHC-II has the same percentage of epitopes overlapping
+    with human/viral/bacterial TMHs as expected by chance
+  \item $\mathcal{H}_{3,HVB}$: The mutations observed in 
+    human/viral/bacterial TMHs are caused by chance.
+    Alternatively, these are caused by natural selection
+    induced by binding to HLAs
 \end{itemize}
@@
 with regards to being detected/presented by the immune system,
-per MHC class and haplotype.
+per MHC class and haplotype. We do this for human, viral and
+bacterial TMHs.
 

### 2020-05-06 article.tex
Describe which MHC-II haplotypes are chosen

@@
 
-\paragraph{Data sets}
-
-To detect an evolutionary signal for selection on the TMHs of viruses, 
-we use the many time-dated COVID-19 proteomes.
-We can assume that only after the moment that COVID-19 spilled over 
-to humans, it has mostly been the human immune system selecting against
-its detection, although this is only part of the selection for 
-having a high transmission rate.
-To detect an evolutionary signal for selection on the TMHs of bacteria, 
-we use the proteomes of Mycobacterium. Mycobacterium is \richel{something}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\paragraph{Data sets for TMH epitopes}
+
+To determine the percentages of epitopes overlapping
+with TMHs, we use two reference proteomes.
+Our viral AA sequence is represented by
+the proteome of the first sequenced COVID-19 strain (\cite{wu2020new},
+GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
+For bacteria, we use the reference genome of 
+Mycobacterium tuberculosis (\url{https://www.ebi.ac.uk/reference_proteomes}, UP000001584, 
+83332 MYCTU).
+
+\paragraph{Data sets for evolutionary selection}
+
+To detect an evolutionary signal for selection on binding
+TMHs polypeptides, we need a phylogenetic tree with 
+many time-dated proteomes.
+For humans, we use the 25000 proteomes used in influenza \richel{citation here}.
+For viruses, we use the many time-dated COVID-19 proteomes.
+We can assume that only after the moment that COVID-19 spilled over 
+to humans, it has mostly been the human immune system selecting against
+its detection, although this is only part of the selection for 
+having a high transmission rate.
+For bacteria, use the proteomes of Mycobacterium 
+\richel{Really? Pick better set, if possible!}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We used the same analysis as \cite{bianchi2017},
-except that instead of a human reference proteome,
-we use the proteome of the first sequenced COVID-19 strain (\cite{wu2020new},
-GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
-and a reference Mycobacterium tuberculosis 
-proteome (\url{https://www.ebi.ac.uk/reference_proteomes}, UP000001584, 
-83332 MYCTU).
-
+except that instead of a human reference proteome.
 Bianchi and colleagues obtained a distribution of 
@@
 overlap with TMHs for the different HLA haplotypes, but then applied to
-COVID-19 and Mycobacterium.
+our viral and bacterial proteome.
 
@@
 For MHC-II, we do the same analysis as MHC-I, except for
-MHC-II haplotypes. We picked the top 21 MHC-II alleles that are most abundant 
-in the current human population (\cite{greenbaum2011functional}).
-\frans{Show coverage}
-We chose to use 21 alleles to have a similar number of alleles as is done
-for MHC-I \richel{Unsure if this was the actual reason}.
-Same as \cite{bianchi2017}, we defined the 5\% of polypeptides 
+MHC-II haplotypes. We picked the MHC-II alleles
+that occur with a frequency of at least 20\% in
+the human population (\cite{greenbaum2011functional},
+see \ref{table:mhc2_haplotypes} for coverage),
+resulting in \richel{approx 8} HLAs.
+The threshold of 20\% is arbitrarily, yet chosen to result in 5 and 10 HLA variants.  
+As \cite{bianchi2017}, we define the 5\% of polypeptides 
 with the lowest IC50 values as binders. We then counted the percentage

### 2020-05-06 article.tex
Method is mostly written down

@@
 is derived from TMHs for all 13-mers present in a proteome
-that have the same hydrophobicity 
-\richel{
-  $\mathcal{H}_{3,1}$
-}.
+that have the same hydrophobicity.
 
@@
 
-We used the same analysis as \cite{bianchi2017},
-except that instead of a human reference proteome.
+To determine the percentages of MHC-I epitopes overlap
+with TMHs, we used the same 
+analysis as \cite{bianchi2017},
+except for using a viral and bacterial proteome (instead 
+of a human proteome, as used in the original study).
 Bianchi and colleagues obtained a distribution of 
@@
 
-For MHC-II, we do the same analysis as MHC-I, except for
+To determine the percentages of MHC-II epitopes overlap, 
+we do the same analysis as MHC-I, except for
 MHC-II haplotypes. We picked the MHC-II alleles
@@
 
-Figure \ref{fig:1} shows the percentages of epitopes overlapping 
-with TMHs for a human, COVID-19 and Mycobacterium proteome.
-The KS test to determine if these percentages are sampled from
-a same distribution resulted in a p-value of \richel{unknown},
-which makes us reject/accept the hypothesis that these percentages
+Figure \ref{fig:1} shows the percentages of MHC-I epitopes overlapping 
+with TMHs for our human, viral and bacterial proteome.
+
+The KS test to determine if MHC-I has the same percentage 
+of epitopes overlapping with viral TMHs, 
+compared to human TMHs \richel{$\mathcal{H}_{1,V}$},
+resulted in a p-value of \richel{unknown},
+which makes us [reject/accept] the hypothesis that these percentages
+are sampled from the same distribution.
+
+The KS test to determine if MHC-I has the same percentage 
+of epitopes overlapping with bacterial TMHs, 
+compared to human TMHs \richel{$\mathcal{H}_{1,B}$},
+resulted in a p-value of \richel{unknown},
+which makes us [reject/accept] the hypothesis that these percentages
 are sampled from the same distribution.
@@
   \caption{
-    \% epitopes overlapping with transmembrane helix
-    for a human, COVID-19 and Mycobacterium proteome.
+    Percentage of MHC-I epitopes overlapping with TMHs
+    for a human, viral and bacterial proteome.
     \richel{
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
+with TMHs for our human, viral and bacterial proteome.
 
@@
 
-	% Table \ref{tab:results} shows the location and binding strength for the
-	% tuberculosis proteome.
-
-	% \begin{figure}[ht]
-	%   \includegraphics[width=\textwidth]{bbbq_3/figure_1.png}
-	%   \label{fig:1}
-	% \end{figure}
-
-	% \begin{figure}[ht]
-	%   \includegraphics[width=\textwidth]{figure_1_5.png}
-	%   \label{fig:1_5}
-	% \end{figure}
-
-	% \begin{figure}[ht]
-	%   \includegraphics[width=\textwidth]{figure_3.png}
-	%   \label{fig:3}
-	% \end{figure}
-
-	% \input{table_imoimo.latex}
-	% % has label tab:results
-	% \richel{
-	%   I'd enjoy 
-	%   (1) a row with 'expected by chance', 
-	%   (2) using percentages instead,
-	%   (3) merging inside and outside
-	% }
-
-	% The inference model with the highest evidence in the
-	% TMH-only alignment was [yet unknown] and [also yet unknown]
-	% for the non-TMH alignment. Individual model weights are shown
-	% in tables \ref{tab:evidences_tmh} 
-	% and \ref{tab:evidences_non_tmh}.
-
-	% The Bayesian inference resulted in [a distribution of mutation rates],
-	% as shown in [absent figure].
-	% The ESSes of the Bayesian parameter estimates was above 200, exact values
-	% are shown in tables \ref{tab:esses_tmh} and \ref{tab:esses_non_tmh}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-	% \frans{
-	%   De TMH evolutie data van COVID is missachien interresanter. Ik zie het als volgt. 
-	%   COVID is een mooi voorbeeld van een virus 
-	%   Myco van een bacterie
-	%   In de bacterie is duidelijk dat tenmisnte voor MHCII TMHs zijn overgrepresenteerd virus niet. 
-	%   Dus verhaal kan zijn dat TMHs belangrijk voot bacteirele afweer niet zozeer voor virussen. 
-	% 
-	%   Hiervoor moeten we MHC-I doen voor myco
-	%   Kijken wat evolutie data doet voor virus dan evt kijken wat dit doet voor myco en vergelijken. 
-	% }
-	% 
-	% We compared the mutation rates between the TMH and non-TMH part of
-	% multiple mycobacterium species. Where we expect no variation 
-	% in mutation rate for every TMH amino acid,
-	% \richel{
-	%   we can test this, but unsure if that would make sense
-	% }
-	% we know that non-TMH part will have regions of different evolutionary
-	% conservation: functional domains, especially in protein-protein
-	% interactions will be strongly conserved, due to an even more constrained
-	% set of peptides that enable a certain function.
-
-	% \richel{
-	%   Note that most bacteria are opportunistic pathogens.
-	%   Note that most bacteria are generalists.
-	%   Note that most bacteria have different cell membranes (and walls), that
-	%   may have different functional constraints than a human cell membrane
-	% }
+\richel{
+  Note that most bacteria are opportunistic pathogens.
+  Note that most bacteria are generalists.
+  Note that most bacteria have different cell membranes (and walls), that
+  may have different functional constraints than a human cell membrane
+}
 

### 2020-05-06 article.tex
Describe artifical data sets

@@
 from human transmembrane helices (TMHs)
-are over-presented by all HLA-A and 
-most HLA-B super types (\cite{bianchi2017}). 
+are over-presented by all 5 HLA-A and 
+most of 8 HLA-B super types (\cite{bianchi2017}).
 One explanation is that the presentation of TMHs 
@@
 the (human) host, as TMHs have a reduced variability 
-due to the functional requirement of being able to span a lipid 
-bilayer. 
+due to the functional requirement of being able to span a lipid bilayer. 
 Due to this, pathogens have a lower chance to develop an escape mutation,
@@
 
+\paragraph{Extracting polypeptides}
+
+\richel{More detail here}.
+
 \paragraph{Correction against many tests}
 
-Because we use \richel{X} MHC-I and \richel{Y} MHC-II haplotypes
-for a significance level of 5\%, by chance we expect \richel{X+Y/20}
-times to find a (MHC-I or MHC-II) haplotype 
-that is suggested to have selection working 
-upon it. Applying a binomial distribution ($p = 0.05$, $n = X+Y$, $\alpha = 0.05$),
-we will claim we detected selection when we found \richel{Z} or
-more examples of having a significant result.
+Because we use 13 MHC-I and 7 MHC-II haplotypes
+for a significance level of 5\%, by chance we expect (5\% of $13 + 7$)
+one times to find a (MHC-I or MHC-II) haplotype 
+that is suggested to have selection working upon it.
+Applying a binomial distribution ($p = 0.05$, $n = 13+7$, $\alpha = 0.05$),
+we will claim we there is some form of selection when we find 3
+or more examples of having a significant result.
+However, because there are 2 hypotheses tested per haplotype,
+we will claim there is directional selection, 
+when, for a same hypothesis, we find 2 examples of having a significant result.
+
+% > pbinom(q = 2, size = 20, prob = 0.05, lower.tail = FALSE)
+% [1] 0.07548367
+%
+% pbinom(q = 3, size = 20, prob = 0.05, lower.tail = FALSE)
+% [1] 0.01590153
 
@@
 
-Of the complete system, we test for each of the 12 state transitions (that is,
+Of the complete system, we test for each of the 2 state transitions (that is,
 we do not test for mutations that result in the same state).
@@
 
-As we are testing 12 hypothesis, 
+As we are testing 2 hypotheses, 
 we correct against type I errors (false positives), 
@@
 resulting in \richel{here is just a normal Bonferroni correction, for now}
-$\alpha_c = \frac{\alpha}{12}$.
+$\alpha_c = \frac{\alpha}{2}$.
 If the probability of observing that few/many 
@@
 
+\paragraph{Measuring strength}
+
+To assess the strength of our analysis, 
+we have applied it to four simulated data sets,
+which are the product of using either MHC-I or MHC-II,
+as well as using two different selection scenarios.
+
+For both MHC-I and MHC-II, we use the haplotypes that have 
+the lowest phenotypic frequency, which is \richel{X} for MHC-I
+and DPB1*0501 (21.7\%) for MHC-II.
+
+The two selection scenarios are either no selection
+or a two-fold selection on what we would expect.
+The scenario without selection, the 'null' data set, acts 
+as a control, where we let the simulate observations occur by chance. 
+The 'rigged' data set is a data set in which we 
+let evolution bias the observations two-fold, that is, 
+we let mutations that are beneficial occur twice as often as expected by chance.
+
+All data sets consists out of $10^4$ observations.
+Each observation starts with a randomly generated source polypeptide fragment,
+as well a mutant. For the 'null' data set, the mutant is randomly select.
+For the 'rigged' data set, beneficial mutations are twice as likely to
+be selected. We define a beneficial mutation as a transition from
+being detected to being undetected.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Poisson binomial distribution
 
@@
 We found that the percentages of epitopes overlapping 
-with TMHs for a human and COVID-19 proteome are 
+with TMHs for a human and viral proteome are 
 [similar/different]. In other words, the
@@
 We found that the percentages of epitopes overlapping 
-with TMHs for a human and Mycobacterium proteome are 
+with TMHs for a human and bacterial proteome are 
 [similar/different]. In other words, the
@@
 
-	% We conclude that MHC-II binds to TMH peptides with a higher/lower/equal
-	% probability than expected by chance.
-
-	% We conclude that the evolutionary conservation if the TMH parts of membrane
-	% proteins is higher/less/equal compare to its non-TMH counterparts.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 to be derived from TMH within either a human host and its viral pathogen.
-Because the full COVID-19 only has 4 TMHs \richel{check}, the percentages
+Because the full COVID-19 has 21 TMHs, the percentages
 of MHC-I epitopes being part of a TMH are likelier to be affected by

### 2020-05-07 article.tex
Use correct spelling of my name

@@
 histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
-It is unknown if the immune system is as able to detect those 
-TMHs present in pathogens.
-Additionally, it is unknown if TMH-II is also likelier to detect
+It is unknown if the immune system is as able to detect 
+TMHs as found in pathogens.
+Additionally, it is unknown if MHC-II is also likelier to detect
 human and pathogen TMHs than expected by chance.
@@
 [less/equally/more] often than expected by chance.
-Lastly, using an unprecedented sharp method, 
+Lastly, using an innovative computation method, 
 we show that natural selection on TMHs is [strong enough/still too weak]
@@
 All nucleated cells in humans present randomly sampled polypeptides
-fragments to the surroundings of the cell using MHC class I molecules.
+fragments to the surroundings of the cell using the Major 
+Histocompatibility Complex (MHC) class I molecules.
 If a cell gets infected by a virus, also the virus' polypeptides
@@
 NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
-A tool not included in this review is \verb;epitope-prediction;,
-which uses a stabilized matrix method, as described in \cite{kim2009derivation}.
-The reason this tool is mentioned, is because it is used 
-in \cite{bianchi2017}.
+A tool developed later is \verb;epitope-prediction; \cite{bianchi2017},
+which uses a stabilized matrix method \cite{kim2009derivation}.
+In this study, we will use \verb;epitope-prediction; \cite{bianchi2017}.
 
@@
 \cite{trolle2014nettepi} and \cite{zhang2015accurate}.
-These later tools, although they look promising, have not been compared
-with one another \richel{Maybe cool to do so!}.
+In this study, we'll be using \richel{Pick FOSS tool here}
 
@@
 these later tools have already become obsolete.
-Contemporary tools that are running, such as
+Contemporary tools that are good for use, such as
 the closed-source MemBrain \cite{feng2020topology} 
 and the FOSS tool PureseqTM \cite{wang2019efficient}.
-\richel{Will use TMHMM for starters (as it is already programmed),
-then use PureseqTM if time}
+In this study, we will use TMHMM  for pragmatic reasons
+\richel{I would love to use PureseqTM instead, but only if there is time}.
 

### 2020-05-11 article.tex
There are approx 460 epitopes in original and 448 in recent proteome

@@
 
-\paragraph{Extracting polypeptides}
+\paragraph{Gene prediction}
 
 \richel{More detail here}.
+
+For gene prediction, we used \richel{X} as is recommended
+in a recent benchmark study for eukaryotes (\cite{scalzitti2020benchmark}).
+
+Although the closed source Vgas (\cite{zhang2019vgas}) appears promising,
+we opted for reproducible research and used Prodigal (\cite{hyatt2012gene})
+instead, which was just a bit better than prokka (\cite{seemann2014prokka}).
 

### 2020-05-11 n_epitopes_in_covid19/n_epitopes.R

@@ -1,0 +1,56 @@
+# How many epitopes does COVID-19 have?
+library(testthat)
+library(ggplot2)
+
+predict_n_epitopes <- function(fasta_filename) {
+  expect_true(file.exists(fasta_filename))
+  proteome <- protr::readFASTA(fasta_filename)
+
+  df <- tibble::tibble(
+    name = names(proteome),
+    n_epitopes = NA,
+    n_aas = NA
+  )
+  n_proteins <- length(proteome)
+  for (i in seq_len(n_proteins)) {
+    protein <- proteome[[i]]
+    n_aas <- nchar(protein)
+    n_epitopes <- nrow(EpitopePrediction::binders(protein))
+    df$n_aas[i] <- n_aas
+    df$n_epitopes[i] <- n_epitopes
+  }
+  df
+}
+
+
+df_original <- predict_n_epitopes(fasta_filename = "proteome.fasta")
+df_original$proteome <- "origin"
+df_recent <- predict_n_epitopes("recent_proteome.fasta")
+df_recent$proteome <- "recent"
+
+sum(df_original$n_aas)
+sum(df_recent$n_aas)
+sum(df_recent$n_aas) / sum(df_original$n_aas)
+
+sum(df_original$n_epitopes)
+sum(df_recent$n_epitopes)
+
+sum(df_recent$n_epitopes) / (sum(df_recent$n_aas) / sum(df_original$n_aas))
+
+df <- rbind(df_original, df_recent)
+df
+
+
+locatome_text <- tmhmm::run_tmhmm("proteome.fasta")
+
+locatome_lines <- locatome_text[
+  stringr::str_detect(
+    locatome_text,
+    pattern = "^>.*",
+    negate = TRUE
+  )
+]
+locatome_lines
+n_non_tmh <- sum(stringr::str_count(locatome_lines, pattern = "(O|o|i|I)"))
+n_tmh <- sum(stringr::str_count(locatome_lines, pattern = "(M|m)"))
+n_tmh / (n_tmh + n_non_tmh)

### 2020-05-12 article.tex
Need to investigate fitness landscape

--- 2020-05-07 article.tex
@@
 exerted by the immune system was only limited \cite{han2019individual}
-\richel{but this study does not look into the evolution of TMHs}.
-In general, on would expect that evolutionary selection is strongest 
-for regions that are important for viral proteins' functions:
-in COVID-19, for example, there is preliminary evidence that the strongest
-selection pressure is upon residues that changes its virulence \cite{velazquez2020positive}
-\richel{residue 251 of ORF3a (an ion channel and a 
-viroporin, residue is not part of TMH), and position 84 of ORF8 (unknown function)}.
-The part of a pathogen's proteome that is essential, 
-only accounts for small part of its full proteome.
-Also, these essential parts differ widely between pathogens.
+
+In general, on would hope that evolutionary selection results in
+an immune system that as most attentive for loci that are
+essential for a virus, as these will be most conserved.
+In COVID-19, for example, there is preliminary evidence that the strongest
+selection pressure is upon residues that changes its 
+virulence \cite{velazquez2020positive}.
+These loci, however, only account for a small part of a pathogen's proteome.
+Additionally, these essential parts differ widely between pathogens.
 Because of this scarcity and variance in targets, 
-one expects that the human immune system
+one can imagine that the human immune system 
 is not tailored to detect these sites, 
@@
 can be detected.
+
+\paragraph{Selection needs to be additive}
+
+Would a pathogen have only peptides that do not bind to
+any HLA, a pathogen would be undetected. In that case, a mutation
+that causes a pathogen to become detected by a host's immune
+system, may be selected against. Also, would a pathogen have only
+one epitope, there would be selection for a mutation to lose it.
+If there are many, say hundreds, of epitopes, the selection to lose or
+gain an epitope is expected to have little effect. The relation
+between the fitness of a pathogen and the number of epitopes is unknown.
+One could assume a two-state fitness landscape: being
+undetected, due to zero epitopes, or being detected, due to one or more
+epitopes. Alternatively, one may hypothesize a gradual fitness decrease per
+extra epitope.
+
+```
+[FIGURE HERE]
+```
+
+Would the pathogen have many epitopes, this study needs to 
+assume a gradual fitness landscape. Due to this, we did a simple 
+exploratory experiment (see 'Fitness landscape'\richel{TODO}) in which
+we counted the number of epitopes in COVID-19. Because the number
+of epitopes has decreased, we dared to assume a gradual fitness landscape
+and to proceed with the full experiment.
 
@@
 
-\paragraph{Extracting polypeptides}
-
-\richel{More detail here}.
+\paragraph{Gene prediction}
+
+Proteogenomics is the field dedicated to, among others,
+predict the proteins that arise from a DNA 
+sequence \cite{nesvizhskii2014proteogenomics}.
+For an ab initio gene prediction of a viral genome, 
+we searched the literature for the most recent software tool,
+which is Vgas (\cite{zhang2019vgas}). In this article, 
+Vgas is compared to tools, such as GeneMarkS (\cite{besemer2001genemarks}),
+Prodigal (\cite{hyatt2012gene}), GLIMMER (\cite{delcher1999improved}),
+RAST \cite{aziz2008rast} and prokka (\cite{seemann2014prokka}).
+We picked the FOSS tool that scored best in the (closed source) Vgas
+benchmark (\cite{zhang2019vgas}), which is Prodigal.
 
@@
 
+Tip: see \cite{bar2020sars} for COVID-19 in numbers.
+
 \richel{

### 2020-05-12 article.tex
Describe pilot experiment better

@@
 epitopes. Alternatively, one may hypothesize a gradual fitness decrease per
-extra epitope.
-
-```
-[FIGURE HERE]
-```
+extra epitope. Both fitness landscapes are depicted in 
+figure \ref{fig:fitness_landscape}.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fitness_landscape.png}
+  \caption{
+    Two conceptual fitness landscapes. At the left is a two-state
+    fitness landscape, in which the number of epitopes has a simple
+    effect: a pathogen's fitness is high if it has no epitopes,
+    as it is undetected by a host's immune system. 
+    For one or more epitopes, fitness is at a same lower values. 
+    In this fitness landscape, acquiring a second or third or more 
+    epitope, this has no effect on fitness.
+    At the right is a gradual fitness landscape, in which any increase in
+    the number of epitopes results in a lower fitness.
+  }
+  \label{fig:fitness_landscape}
+\end{figure}
 
@@
 assume a gradual fitness landscape. Due to this, we did a simple 
-exploratory experiment (see 'Fitness landscape'\richel{TODO}) in which
+exploratory experiment (see 'Fitness landscape', 
+subsection \ref{subsec:fitness_landscape}) in which
 we counted the number of epitopes in COVID-19. Because the number
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Fitness landscape}
+\label{subsec:fitness_landscape}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Reason to do before actual experiment}
+This exploratory experiment tests an assumptions 
+of the bigger evolutionary experiment. 
+If the assumption is incorrect,
+a different evolutionary experiment needs to be contrived.
+
+\paragraph{Method in general}
+For a given MHC-I haplotype, we count the number of epitopes in 
+the proteome of COVID-19, at the start of the 2019 pandemic (as
+described in \cite{wu2020new})
+and a random strain sequenced most recently at the day of the experiment. 
+Of both proteomes, we use use \verb;epitope-prediction; \cite{bianchi2017} 
+to count the number of predicted binders for the HLA-A-02:01 haplotype.
+
+\paragraph{Decision rule}
+If the number of epitopes has increased, we assume a two-state fitness landscape,
+in which the number of epitopes can increase neutrally by drift alone.
+In this case, the experiment will be cancelled and replaced by a 
+viable approach.
+Else, the bigger evolutionary experiment is done as described already.
+It may still be possible that the number of epitopes decreased due to drift,
+but this would be a proper result in confirming the null hypothesis
+of $\mathcal{H}_{3,HVB}$.
+
+\paragraph{Method detail}
+Because the proteome is only sequenced at the start of the pandemic,
+we use the DNA sequences of both strains, and infer the proteome
+of each, using Prodigal. Although the earlier strand has a published
+proteome, we infer it regardless, as to use the same pipeline.
+If needed, we tweak the Prodigal parameter settings up until
+the inferred proteome of the earlier strain with known proteome
+matches well enough, as concluded by eyeballing.
+The parameter settings used to infer a good proteome on the earlier
+strain, is then used on the genome of the more recent strain 
+to infer our recent proteome.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-I}

### 2020-05-12 article.tex
Put most exporatory study in supmat

@@
 between the fitness of a pathogen and the number of epitopes is unknown.
-One could assume a two-state fitness landscape: being
-undetected, due to zero epitopes, or being detected, due to one or more
-epitopes. Alternatively, one may hypothesize a gradual fitness decrease per
-extra epitope. Both fitness landscapes are depicted in 
-figure \ref{fig:fitness_landscape}.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fitness_landscape.png}
-  \caption{
-    Two conceptual fitness landscapes. At the left is a two-state
-    fitness landscape, in which the number of epitopes has a simple
-    effect: a pathogen's fitness is high if it has no epitopes,
-    as it is undetected by a host's immune system. 
-    For one or more epitopes, fitness is at a same lower values. 
-    In this fitness landscape, acquiring a second or third or more 
-    epitope, this has no effect on fitness.
-    At the right is a gradual fitness landscape, in which any increase in
-    the number of epitopes results in a lower fitness.
-  }
-  \label{fig:fitness_landscape}
-\end{figure}
-
-Would the pathogen have many epitopes, this study needs to 
-assume a gradual fitness landscape. Due to this, we did a simple 
+
+This study assumes a gradual fitness landscape, in which it always
+(slightly) beneficial for a pathogen to lose an epitope.
+Because this assumption needs to be correct for the research
+to be viable, we did an
 exploratory experiment (see 'Fitness landscape', 
-subsection \ref{subsec:fitness_landscape}) in which
-we counted the number of epitopes in COVID-19. Because the number
-of epitopes has decreased, we dared to assume a gradual fitness landscape
-and to proceed with the full experiment.
+subsection \ref{subsec:fitness_landscape}).
+We concluded that the assumption may be correct.
 
@@
 
-\paragraph{Reason to do before actual experiment}
+\paragraph{Introduction}
+One could assume a two-state fitness landscape: being
+undetected, due to zero epitopes, or being detected, due to one or more
+epitopes. Alternatively, one may hypothesize a gradual fitness decrease per
+extra epitope. Both fitness landscapes are depicted in 
+figure \ref{fig:fitness_landscape}.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{fitness_landscape.png}
+  \caption{
+    Two conceptual fitness landscapes. At the left is a two-state
+    fitness landscape, in which the number of epitopes has a simple
+    effect: a pathogen's fitness is high if it has no epitopes,
+    as it is undetected by a host's immune system. 
+    For one or more epitopes, fitness is at a same lower values. 
+    In this fitness landscape, acquiring a second or third or more 
+    epitope, this has no effect on fitness.
+    At the right is a gradual fitness landscape, in which any increase in
+    the number of epitopes results in a lower fitness.
+  }
+  \label{fig:fitness_landscape}
+\end{figure}
+
+\paragraph{Why}
 This exploratory experiment tests an assumptions 
@@
 
-\paragraph{Method in general}
+\paragraph{Method}
 For a given MHC-I haplotype, we count the number of epitopes in 
@@
 to count the number of predicted binders for the HLA-A-02:01 haplotype.
+We use that haplotype, as it is the haplotype used by default with the
+\verb;epitope-prediction; package.
 

### 2020-05-12 article.tex
Described exploratory study

@@
 Would a pathogen have only peptides that do not bind to
-any HLA, a pathogen would be undetected. In that case, a mutation
+any HLA, a pathogen would be undetected. In that case, one
+can imagine that a mutation
 that causes a pathogen to become detected by a host's immune
 system, may be selected against. Also, would a pathogen have only
-one epitope, there would be selection for a mutation to lose it.
-If there are many, say hundreds, of epitopes, the selection to lose or
-gain an epitope is expected to have little effect. The relation
-between the fitness of a pathogen and the number of epitopes is unknown.
+one epitope, one can imagine there would be selection for a mutation to lose it.
+If there are, however, already many epitopes present in the
+proteome, the selection to lose or gain an epitope is expected 
+to have less or no effect.
+
+\paragraph{Test that selection is additive}
 
@@
 subsection \ref{subsec:fitness_landscape}).
-We concluded that the assumption may be correct.
+We concluded that the assumption may be correct 
+\richel{Let's assume for now}.
 

### 2020-05-14 article.tex
Assume immunodominance, instead of having to prove the shape of the fitness landscape

@@
 
-\paragraph{Test that selection is additive}
-
-This study assumes a gradual fitness landscape, in which it always
-(slightly) beneficial for a pathogen to lose an epitope.
-Because this assumption needs to be correct for the research
-to be viable, we did an
-exploratory experiment (see 'Fitness landscape', 
-subsection \ref{subsec:fitness_landscape}).
-We concluded that the assumption may be correct 
-\richel{Let's assume for now}.
+\paragraph{Immunodominance}
+
+The effective number of epitopes, however,
+is much lower than the actual number of possible
+epitopes, due to immunodominance. 
+Immunodominance is a feature of the immune system,
+due to which some epitopes dominate in causing an immune response,
+where other (called subdominant) epitopes end up having no 
+effect \cite{akram2012immunodominance}.
+The same study mentions 20 factors that influence the strength
+of an immune response.
+The most relevant factor for this study is the effect of
+a mutation within an epitope, 
+such as the in vitro example described in \cite{berkhoff2004mutation},
+where one single mutation resulted in a 5-20\% weaker
+responese of virus-specific CD8+ T cells \richel{Fun experiment:
+do this in silico}.
+Due to immunodominance, we dare assume that there can be selection
+for immune system avoidance, as the effective number of loci may be
+low and effects of a single mutation noticable.
 
@@
 high performance computing cluster. 
+Additionally, we would like to thank Sci-Hub \cite{himmelstein2018sci}
+for allowing us to read paywalled articles while working from home.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Fitness landscape}
-\label{subsec:fitness_landscape}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\paragraph{Introduction}
-One could assume a two-state fitness landscape: being
-undetected, due to zero epitopes, or being detected, due to one or more
-epitopes. Alternatively, one may hypothesize a gradual fitness decrease per
-extra epitope. Both fitness landscapes are depicted in 
-figure \ref{fig:fitness_landscape}.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{fitness_landscape.png}
-  \caption{
-    Two conceptual fitness landscapes. At the left is a two-state
-    fitness landscape, in which the number of epitopes has a simple
-    effect: a pathogen's fitness is high if it has no epitopes,
-    as it is undetected by a host's immune system. 
-    For one or more epitopes, fitness is at a same lower values. 
-    In this fitness landscape, acquiring a second or third or more 
-    epitope, this has no effect on fitness.
-    At the right is a gradual fitness landscape, in which any increase in
-    the number of epitopes results in a lower fitness.
-  }
-  \label{fig:fitness_landscape}
-\end{figure}
-
-\paragraph{Why}
-This exploratory experiment tests an assumptions 
-of the bigger evolutionary experiment. 
-If the assumption is incorrect,
-a different evolutionary experiment needs to be contrived.
-
-\paragraph{Method}
-For a given MHC-I haplotype, we count the number of epitopes in 
-the proteome of COVID-19, at the start of the 2019 pandemic (as
-described in \cite{wu2020new})
-and a random strain sequenced most recently at the day of the experiment. 
-Of both proteomes, we use use \verb;epitope-prediction; \cite{bianchi2017} 
-to count the number of predicted binders for the HLA-A-02:01 haplotype.
-We use that haplotype, as it is the haplotype used by default with the
-\verb;epitope-prediction; package.
-
-\paragraph{Decision rule}
-If the number of epitopes has increased, we assume a two-state fitness landscape,
-in which the number of epitopes can increase neutrally by drift alone.
-In this case, the experiment will be cancelled and replaced by a 
-viable approach.
-Else, the bigger evolutionary experiment is done as described already.
-It may still be possible that the number of epitopes decreased due to drift,
-but this would be a proper result in confirming the null hypothesis
-of $\mathcal{H}_{3,HVB}$.
-
-\paragraph{Method detail}
-Because the proteome is only sequenced at the start of the pandemic,
-we use the DNA sequences of both strains, and infer the proteome
-of each, using Prodigal. Although the earlier strand has a published
-proteome, we infer it regardless, as to use the same pipeline.
-If needed, we tweak the Prodigal parameter settings up until
-the inferred proteome of the earlier strain with known proteome
-matches well enough, as concluded by eyeballing.
-The parameter settings used to infer a good proteome on the earlier
-strain, is then used on the genome of the more recent strain 
-to infer our recent proteome.
 

### 2020-05-14 article.tex
Works again

@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1_watermarked.png}
+  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/fig_bbbq_2_watermarked.png}
+  \includegraphics[width=\textwidth]{bbbq_2/fig_bbbq_2.png}
   \caption{
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{COVID-19 TMHs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
-  \caption{
-    Location of amino acids of reference COVID-19 proteome.
-    Note that ORF1a results in multiple proteins, 
-    as shown in figure \ref{fig:covid_genome_and_proteome}.
-    Legend: i = inside (cytosolic side), m = membrane, o = outside (extracellular side)
-    \richel{This is real data}
-  }
-  \label{fig:covid_locatome}
-\end{figure}
+%\subsection{COVID-19 TMHs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\begin{figure}[!htbp]
+%  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
+%  \caption{
+%    Location of amino acids of reference COVID-19 proteome.
+%    Note that ORF1a results in multiple proteins, 
+%    as shown in figure \ref{fig:covid_genome_and_proteome}.
+%    Legend: i = inside (cytosolic side), m = membrane, o = outside (extracellular side)
+%    \richel{This is real data}
+%  }
+%  \label{fig:covid_locatome}
+%\end{figure}
 

### 2020-05-14 p_bind_per_hydrophobicity/p_bind_per_hydrophobicity.R

@@ -5,6 +5,7 @@
 # H1: higher hydrophobicity equals higher chance to bind
 #
 library(ggplot2)
+library(dplyr)
 
 if (1 == 2) {
   #' Create a dataset
@@ -173,6 +174,8 @@ plot_hydrophobicity_vs_is_tmh(png_filename = "hydrophobicity_vs_is_tmh.png")
 plot_hydrophobicity_vs_binds_mhc1(png_filename = "hydrophobicity_vs_binds_mhc1.png")
 plot_hydrophobicity_vs_binds_mhc2(png_filename = "hydrophobicity_vs_binds_mhc2.png")
 
+df <- read.csv(system.file("extdata", "raw_data.csv", package = "bbbq"))
+df <- tibble::as_tibble(df)
 
 names(df)
 df_plot <- dplyr::mutate(df, bin = trunc(hydrophobicity * 2))
@@ -276,22 +279,24 @@ save_plots <- function(
     geom_boxplot() + ggsave(filenames[4])
 }
 
-save_plots(
-  df = create_dataset(n = 1000, n_aas = 20, f = create_random_hydrophobic_polypeptide),
-  filenames = c(
-    "p_bind_mhc1_per_hydrophobicity.png",
-    "p_bind_mhc2_per_hydrophobicity.png",
-    "hydrophobicity_per_mhc1.png",
-    "hydrophobicity_per_mhc2.png"
+if (1 == 2) {
+  save_plots(
+    df = create_dataset(n = 1000, n_aas = 20, f = create_random_hydrophobic_polypeptide),
+    filenames = c(
+      "p_bind_mhc1_per_hydrophobicity.png",
+      "p_bind_mhc2_per_hydrophobicity.png",
+      "hydrophobicity_per_mhc1.png",
+      "hydrophobicity_per_mhc2.png"
+    )
   )
-)
-
-save_plots(
-  df = create_dataset(n = 1000, n_aas = 20, f = create_random_tmh),
-  filenames = c(
-    "p_bind_mhc1_per_hydrophobicity_tmh.png",
-    "p_bind_mhc2_per_hydrophobicity_tmh.png",
-    "hydrophobicity_per_mhc1_tmh.png",
-    "hydrophobicity_per_mhc2_tmh.png"
+  
+  save_plots(
+    df = create_dataset(n = 1000, n_aas = 20, f = create_random_tmh),
+    filenames = c(
[…]

### 2020-05-14 view.sh

@@ -1,2 +1,2 @@
 #!/bin/bash
-qpdfview article.pdf
+evince article.pdf

### 2020-05-14 bbbq_article.tex
Separate LaTeX header from content

--- 2020-05-14 article.tex
@@
-\documentclass{article}
-
-% Just to have them already
-% \begin{sloppypar}
-% \end{sloppypar}
-
-% Just try to parse, do not ask for input
-\nonstopmode
-
-% Bibliography
-\usepackage{natbib}
-\bibpunct{(}{)}{;}{a}{}{;}
-
-% Use 'It was found that A is B (Name 1234)' style
-\setcitestyle{authoryear,open={},close={}}
-
-% Affiliations
-\usepackage{authblk}
-\title{
-  Transmembrane helices are also 
-  an overlooked source of major histocompatibility complex class II epitopes
-  and evolutionary more conserved than expected by chance
-}
-\author[1]{Richèl J.C. Bilderbeek}
-\author[1]{Maxim Baranov}
-\author[1]{Geert van den Bogaart}
-\author[1]{Frans Bianchi}
-\affil[1]{GBB, University of 
-Groningen, Groningen, The Netherlands}
-
-% Use double spacing
-\usepackage{setspace}
-\doublespacing
-
-\usepackage{listings}
-\usepackage{hyperref}
-\usepackage{todonotes}
-\usepackage{verbatim}
-\usepackage{pgf}
-\usepackage{bm}
-\usepackage{multirow}
-\usepackage{amsfonts}
-\usepackage{array}
-\usepackage{booktabs}
-\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
-\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
-\usepackage{longtable}
-
-\usepackage{tkz-graph}
-\usetikzlibrary{arrows,automata}
-\usetikzlibrary{calc}
-\usetikzlibrary{arrows.meta}
-
-% sidewaysfigure
-\usepackage{rotating}
-
-% Style of listings
-% From http://r.789695.n4.nabble.com/
-%   How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
-\usepackage{fancyvrb} 
-\definecolor{codegreen}{rgb}{0,0.6,0}
-\definecolor{codegray}{rgb}{0.5,0.5,0.5}
-\definecolor{codepurple}{rgb}{0.58,0,0.82}
-\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
-\lstdefinestyle{mystyle}{
-  language=R,% set programming language
-  basicstyle=\ttfamily\small,% basic font style
-  commentstyle=\color{gray},% comment style
-  numberstyle=\scriptsize,% use small line numbers
-  numbersep=10pt,% space between line numbers and code
-  tabsize=2,% sizes of tabs
-  showstringspaces=false,
-  captionpos=b,% positioning of the caption below
-  breaklines=true,% automatic line breaking
-  escapeinside={(*}{*)},% escaping to LaTeX
-  fancyvrb=true,% verbatim code is typset by listings
-  extendedchars=false,% prohibit extended chars (chars of codes 128--255)
-  alsoletter={.<-},% becomes a letter
-  alsoother={$},% becomes other
-  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},
-  deletekeywords={c}% remove keywords 
-}
-\lstset{style=mystyle}
-
-% Adds numbered lines. Not yet...
-% \usepackage{lineno}
-% \linenumbers
-
-%comments
-\newcommand{\frans}[1]{\textcolor{blue}{\textbf{[FB: #1]}}}
-\newcommand{\richel}[1]{\textcolor{orange}{\textbf{[RB: #1]}}}
-
-\begin{document}
-
-\maketitle
-
 \begin{abstract}
@@
 \bibliographystyle{mee}
-\bibliography{article}
+\bibliography{bbbq_article}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{equation}
-
-\end{document}

### 2020-05-14 create.sh

@@ -1,7 +1,8 @@
 #!/bin/bash
 
-pdflatex article.tex --shell-escape >/dev/null
-bibtex article.aux >/dev/null
-pdflatex article.tex --shell-escape >/dev/null
-pdflatex article.tex --shell-escape
+pdflatex bbbq_header.tex
+bibtex bbbq_header >/dev/null
+pdflatex bbbq_header.tex >/dev/null
+pdflatex bbbq_header.tex
+mv bbbq_header.pdf bbbq_article.pdf
 

### 2020-05-14 edit.sh

@@ -7,5 +7,5 @@
 #   ./edit.sh
 #
 
-mousepad article.bib article.tex
+mousepad bbbq_header.tex bbbq_article.bib bbbq_article.tex
 make

### 2020-05-14 view.sh

@@ -1,2 +1,2 @@
 #!/bin/bash
-evince article.pdf
+evince bbbq_article.pdf

### 2020-05-18 bbbq_article.tex
EOD
Fix makefile and references

@@
 Later tools are \cite{zhang2013predicting},
-\cite{trolle2014nettepi} and \cite{zhang2015accurate}.
-In this study, we'll be using \richel{Pick FOSS tool here}
+\cite{trolle2014nettepi}, \cite{zhang2015accurate}
+and the very recent MHCnuggets\cite{shao2020high}.
+In this study, we'll be using MHCnuggets, as it is the only FOSS tool available.
 
@@
 Due to this, pathogens have a lower chance to develop an escape mutation,
-as many mutations will result in a disfunctional TMH.
+as many mutations will result in a dysfunctional TMH.
 Note that the mechanism by which a cell presents its TMHs is
@@
 where one single mutation resulted in a 5-20\% weaker
-responese of virus-specific CD8+ T cells \richel{Fun experiment:
+response of virus-specific CD8+ T cells \richel{Fun experiment:
 do this in silico}.
@@
 for immune system avoidance, as the effective number of loci may be
-low and effects of a single mutation noticable.
+low and effects of a single mutation noticeable.
 

### 2020-05-18 bbbq_article.tex
Supmat fixed
Add dataset strength calculation

@@
 
-\begin{table}
+\begin{table}[!htbp]
   \input{bbbq_1/bbbq_1_percentages.latex}
@@
 
-\begin{table}
+\begin{table}[!htbp]
   \input{bbbq_1/bbbq_1_stats_covid.latex}
@@
 
-\begin{table}
+\begin{table}[!htbp]
   \input{bbbq_1/bbbq_1_stats_myco.latex}
@@
 
-\begin{table}
+\begin{table}[!htbp]
   \input{bbbq_2/bbbq_2_percentages.latex}
@@
 
-\begin{table}
+\begin{table}[!htbp]
   \input{bbbq_2/bbbq_2_stats_covid.latex}
@@
 
-\begin{table}
+\begin{table}[!htbp]
   \input{bbbq_2/bbbq_2_stats_myco.latex}
@@
 \end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Strength of COVID-19 dataset}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Our evolutionary experiment depends on the strength of the dataset.
+Here we do a back-on-the-envelope calculation to assess if the dataset
+is strong enough.
+
+\begin{itemize}
+	\item Genome size: $\approx 3 \cdot 10^4$ bases
+	\item Evolution rate: $\approx 10^{-3}$ per nucleotide per year (from \cite{bar2020sars})
+	\item Evolution rate per genome: $\approx 26$ per year (GISAID)
+	\item Evolution rate per nucleotide: $0.00088$ per nucleotide per year as $\frac{26}{\cdot 10^4}$ per nucleotide year
+	\item Mutation rate: $\approx 10^{-6}$ per nucleotide per cycle
+	\item Proteome size: 9701 (from reference proteome)
+	\item Size of proteome compared to genome: $\frac{9701}{3 \cdot 10^4} \approx \frac{1}{3}$ \richel{no/little non-coding DNA indeed}
+	\item Fraction of mutations that change AA: $\frac{20}{64}$ 
+	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
+	\item TMHs per proteome: 21 (from TMHMM)
+	\item AAs per TMHs: 23 (from Bianchi et al.)
+	\item AAs per proteome that is TMH: $21 \cdot 23 = 483$
+	\item Mutations in TMH per proteome year: $\approx 0.444$ (as $483 \cdot 0.00092$)
+	\item Number of sequenced proteomes: 5600 (GISAID, genomes)
+	\item Pandemic duration: $\approx \frac{1}{3}$ year
+	\item Expected number of observed TMH mutations: $\approx 800$ (as $0.444 \cdot 5600 \cdot \frac{1}{3}$)
+	\item Chance that a mutation may lose/acquire epitope binding: $\frac{1}{100}$ \richel{Just a guess, based on weak selection}
+	\item Expected number of observed TMH mutations that cause a transition: $\approx 8$ (as $800 \cdot \frac{1}{100}$)
+\end{itemize}
+
+Conclusion: weak, but a good
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Strength of Influenza A}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Our evolutionary experiment depends on the strength of the dataset.
+Here we do a back-on-the-envelope calculation to assess if the dataset
+is strong enough.
+
+\begin{itemize}
+	\item Genome size: $\approx 1.3 \cdot 10^4$ bases
+	\item Relative genome size compared to COVID-19: $\approx 0.4$ (as $\frac{1.3 \cdot 10^4}{3.0 \cdot 10^4}$)
+	\item Full genome sets: $\approx 3.2 \cdot 10^4$ (Influenza A)
+	\item Relative number of genome sets compared to COVID-19: $\approx 6.4$ as much (as $\frac{\approx 32 \cdot 10^3}{5 \cdot 10^3}$)
+	\item Relative time to accumulate mutations: $\approx 12$ times longer (average age $\approx$ 4 year, before 2015, already 8k sequences)
+	\item Relative chance of observing transitions: $\approx 30$ (as $0.4 \cdot 6.4 \cdot 12$)
+	\item Expected number of observed TMH mutations in COVID-19 that cause a transition: $\approx 8$
+	\item Expected number of observed TMH mutations that cause a transition: $\approx 240$ (as $8 \cdot 30$)
+\end{itemize}
+
+Conclusion: seems strong enough to us.
 
@@
 
-\begin{table}
+\begin{table}[!htbp]
   \input{mhc2_haplotypes.latex}

### 2020-05-18 n_epitopes_in_covid19/n_epitopes.R

@@ -2,6 +2,16 @@
 library(testthat)
 library(ggplot2)
 
+get_proteome_size <- function(fasta_filename) {
+  expect_true(file.exists(fasta_filename))
+  proteome <- protr::readFASTA(fasta_filename)
+  n_chars <- 0
+  for (protein in proteome) {
+    n_chars <- n_chars + nchar(protein)
+  }
+  n_chars
+}
+
 predict_n_epitopes <- function(fasta_filename) {
   expect_true(file.exists(fasta_filename))
   proteome <- protr::readFASTA(fasta_filename)

### 2020-05-19 bbbq_article.tex
COVID-19 TMHs is back in again, using pureseqtmr
Show number of TMHs, cite 'pureseqtmr'

@@
 
-%\begin{figure}[!htbp]
-%  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
-%  \caption{
-%    Location of amino acids of reference COVID-19 proteome.
-%    Note that ORF1a results in multiple proteins, 
-%    as shown in figure \ref{fig:covid_genome_and_proteome}.
-%    Legend: i = inside (cytosolic side), m = membrane, o = outside (extracellular side)
-%    \richel{This is real data}
-%  }
-%  \label{fig:covid_locatome}
-%\end{figure}
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
+  \caption{
+    Location of amino acids of reference COVID-19 proteome.
+    Note that ORF1a results in multiple proteins, 
+    as shown in figure \ref{fig:covid_genome_and_proteome}.
+    Legend: i = inside (cytosolic side), m = membrane, o = outside (extracellular side)
+    \richel{This is real data}
+  }
+  \label{fig:covid_topology}
+\end{figure}
+
+\begin{table}[!htbp]
+  \input{tmhs/covid_tmhs.latex}
+  \caption{
+    Number of TMHs per COVID-19 ORF,
+    using 'pureseqtmr' \cite{pureseqtmr}
+  }
+  \label{table:covid_tmhs}
+\end{table}
 

### 2020-05-19 covid_dataset_strength/covid_dataset_strength.Rmd
First start of determining the strength of the dataset

---
title: "COVID-19 dataset strength"
author: Richel J.C. Bilderbeek
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(testthat)
library(pureseqtmr)
```

Must be installed:

```{r}
expect_true(is_pureseqtm_installed())
```

```{r cache=TRUE}
fasta_filename <- "covid.fasta"
expect_true(file.exists(fasta_filename))
topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
```

```{r}
n_non_tmh_aas <- sum(stringr::str_count(topology$topology, "0"))
message("number of non-TMH amino acids, n_non_tmh_aas: ", n_non_tmh_aas)
```

```{r}
n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
message("number of TMH amino acids, n_tmh_aas: ", n_tmh_aas)
```

```{r}
n_aas <- n_non_tmh_aas + n_tmh_aas
message("proteome size, number of amino acids, n_aas: ", n_aas)
```

```{r}
perc_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
message("percentage of proteome that is TMH, perc_aas: ", perc_aas)
```

```{r}
n_tmhs <- sum(tally_tmhs(topology)$n_tmhs) 
message("number of TMHs, n_tmhs: ", n_tmhs)
```

### 2020-05-19 create.sh
Append PDF of COVID-19 dataset strength

@@ -4,5 +4,6 @@ pdflatex bbbq_header.tex
 bibtex bbbq_header >/dev/null
 pdflatex bbbq_header.tex >/dev/null
 pdflatex bbbq_header.tex
-mv bbbq_header.pdf bbbq_article.pdf
-
+pdftk bbbq_header.pdf covid_dataset_strength/covid_dataset_strength.pdf \
+  cat output bbbq_article.pdf
+rm bbbq_header.pdf

### 2020-05-19 bbbq_article.tex
Need to assess data set strength from script
Fix merge conflict (?) at Kolmogorov-Smirnoff test
EOD

@@
 	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
-	\item TMHs per proteome: 21 (from TMHMM)
+	\item TMHs per proteome: 26 (from table \ref{table:covid_tmhs})
 	\item AAs per TMHs: 23 (from Bianchi et al.)
@@
   \caption{
-    Location of amino acids of reference COVID-19 proteome.
+    Location of amino acids of reference COVID-19 
+    proteome (Uniprot ID: UP000464024).
     Note that ORF1a results in multiple proteins, 
     as shown in figure \ref{fig:covid_genome_and_proteome}.
-    Legend: i = inside (cytosolic side), m = membrane, o = outside (extracellular side)
-    \richel{This is real data}
+    Legend: 1 = located in the membrane, 0 = not located in the membrane.
+    Topology obtained using 'pureseqtmr' \cite{pureseqtmr}
   }
@@
 
-\end{document}
-new research. 
-      This table be deleted in the future.
-    }
-  }
-  \label{table:mhc2_haplotypes}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Kolmogorov-Smirnov}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{
-  This is just a reminder, instead of new research. 
-  This subsection be deleted in the future.
-}
-
-The Kolmogorov-Smirnov (KS) test determines if two samples
-are derived from the same distribution, without making assumptions
-regarding the shape of that distribution.
-
-We will reject
-the null hypothesis that MHC-I has the same percentage of epitopes 
-overlapping with TMHs in Homo sapiens compared to each pathogen when 
-the KS statistic $D_{n,m}$ follows the relationship as shown in 
-equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
-and $n = m$ equals the number of HLA haplotypes.
-
-\begin{equation}
-   D_{n,m} > \frac{1}{\sqrt{n}} \cdot \sqrt{ -\ln(\frac{\alpha}{2}) \cdot \frac{1 + \frac{n}{m}}{2} }
-   \label{eq:ks}
-\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Added PDFs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+These PDFs are added. These will be moved into the article when needed.

### 2020-05-19 covid_dataset_strength/covid_dataset_strength.Rmd
EOD

@@
 ```{r cache=TRUE}
-fasta_filename <- "covid.fasta"
+fasta_filename <- "UP000464024.fasta"
 expect_true(file.exists(fasta_filename))
@@
 n_aas <- n_non_tmh_aas + n_tmh_aas
-message("proteome size, number of amino acids, n_aas: ", n_aas)
+proteome_size <- n_aas
+message("number of amino acids, n_aas: ", n_aas)
+message("proteome size, proteome_size: ", proteome_size)
 ```
@@
 ```
+
+Get the proteome mutation rate, by converting it from the 
+known DNA mutation rate:
+
+```{r}
+genome_size <- 30000 # nucleotides
+
+dna_evolution_rate_per_genome_per_year <- 26 # From GISAID
+# (Same as bar2020sars, that esimates a DNA 
+# evolution rate per nucleotide per year at 0.001)
+```
+
+```{r}
+rel_size_proteome_to_genome <- 
+  proteome_size / genome_size
+message("rel_size_proteome_to_genome: ", rel_size_proteome_to_genome)
+```
+
+\begin{itemize}
+	\item Proteome size: 9701 (from reference proteome)
+	\item Size of proteome compared to genome: $\frac{9701}{3 \cdot 10^4} \approx \frac{1}{3}$ \richel{no/little non-coding DNA indeed}
+	\item Fraction of mutations that change AA: $\frac{20}{64}$ 
+	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
+	\item TMHs per proteome: 26 (from table \ref{table:covid_tmhs})
+	\item AAs per TMHs: 23 (from Bianchi et al.)
+	\item AAs per proteome that is TMH: $21 \cdot 23 = 483$
+	\item Mutations in TMH per proteome year: $\approx 0.444$ (as $483 \cdot 0.00092$)
+	\item Number of sequenced proteomes: 5600 (GISAID, genomes)
+	\item Pandemic duration: $\approx \frac{1}{3}$ year
+	\item Expected number of observed TMH mutations: $\approx 800$ (as $0.444 \cdot 5600 \cdot \frac{1}{3}$)
+	\item Chance that a mutation may lose/acquire epitope binding: $\frac{1}{100}$ \richel{Just a guess, based on weak selection}
+	\item Expected number of observed TMH mutations that cause a transition: $\approx 8$ (as $800 \cdot \frac{1}{100}$)
+\end{itemize}

### 2020-05-20 compare_proteomes/compare_proteomes.Rmd
Copy proteomes from 'downloads' folder

---
title: "COVID-19 dataset strength"
author: Richel J.C. Bilderbeek
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(testthat)
library(pureseqtmr)
```

Must be installed:

```{r}
expect_true(is_pureseqtm_installed())
```

```{r}
show_properties <- function(fasta_filename) {
  expect_true(file.exists(fasta_filename))
  topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
  n_proteins <- nrow(topology)
  n_non_tmh_aas <- sum(stringr::str_count(topology$topology, "0"))
  n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
  n_aas <- n_non_tmh_aas + n_tmh_aas
  testthat::expect_equal(
    n_aas,
    sum(nchar(protr::readFASTA(fasta_filename)))
  )
  perc_tmh_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
  n_tmhs <- sum(tally_tmhs(topology)$n_tmhs)

  tibble::tibble(
    n_aas = n_aas,
    n_proteins = n_proteins,
    n_non_tmh_aas = n_non_tmh_aas,
    n_tmh_aas = n_tmh_aas,
    perc_tmh_aas = perc_tmh_aas,
    n_tmhs = n_tmhs
  )
}

```

```{r}
t1 <- show_properties(fasta_filename = "proteome.fasta")
knitr::kable(t1)
```

```{r}
t2 <- show_properties(fasta_filename = "UP000464024.fasta")
knitr::kable(t2)
```

### 2020-05-20 bbbq_article.tex
COVID-19 dataset strength calculation looks solid to me, sent out for review by @bogeert and Frans

@@
 
-Our evolutionary experiment depends on the strength of the dataset.
-Here we do a back-on-the-envelope calculation to assess if the dataset
-is strong enough.
-
-\begin{itemize}
-	\item Genome size: $\approx 3 \cdot 10^4$ bases
-	\item Evolution rate: $\approx 10^{-3}$ per nucleotide per year (from \cite{bar2020sars})
-	\item Evolution rate per genome: $\approx 26$ per year (GISAID)
-	\item Evolution rate per nucleotide: $0.00088$ per nucleotide per year as $\frac{26}{\cdot 10^4}$ per nucleotide year
-	\item Mutation rate: $\approx 10^{-6}$ per nucleotide per cycle
-	\item Proteome size: 9701 (from reference proteome)
-	\item Size of proteome compared to genome: $\frac{9701}{3 \cdot 10^4} \approx \frac{1}{3}$ \richel{no/little non-coding DNA indeed}
-	\item Fraction of mutations that change AA: $\frac{20}{64}$ 
-	\item Evolution rate: $\approx 0.00092$ per amino acid per year (as $\frac{1}{3} \cdot \frac{20}{64} \cdot 0.00088$)
-	\item TMHs per proteome: 26 (from table \ref{table:covid_tmhs})
-	\item AAs per TMHs: 23 (from Bianchi et al.)
-	\item AAs per proteome that is TMH: $21 \cdot 23 = 483$
-	\item Mutations in TMH per proteome year: $\approx 0.444$ (as $483 \cdot 0.00092$)
-	\item Number of sequenced proteomes: 5600 (GISAID, genomes)
-	\item Pandemic duration: $\approx \frac{1}{3}$ year
-	\item Expected number of observed TMH mutations: $\approx 800$ (as $0.444 \cdot 5600 \cdot \frac{1}{3}$)
-	\item Chance that a mutation may lose/acquire epitope binding: $\frac{1}{100}$ \richel{Just a guess, based on weak selection}
-	\item Expected number of observed TMH mutations that cause a transition: $\approx 8$ (as $800 \cdot \frac{1}{100}$)
-\end{itemize}
-
-Conclusion: weak, but a good
+See at the end of this document.
 

### 2020-05-20 covid_dataset_strength/covid_dataset_strength.Rmd
Use reference proteome
COVID-19 dataset strength calculation looks solid to me, sent out for review by @bogeert and Frans

--- 2020-05-19 covid_dataset_strength/covid_dataset_strength.Rmd
@@
 
+## Goal
+
+To determine the strength of the COVID-19 dataset, using
+a back-of-the-envelope calculation.
+
+## Setup
+
+We load the libraries needed:
+
 ```{r, message=FALSE}
@@
 
-Must be installed:
+PureseqTM, which predicts TMHs, must be installed:
 
@@
 
-```{r cache=TRUE}
-fasta_filename <- "covid.fasta"
+## Abbreviations
+
+Abbr  |Full
+------|-----------
+AA    |Amino acid
+AAs   |Amino acids
+Abbr  |Abbreviation
+TMH   |Transmembrane helix
+TMHs  |Transmembrane helices
+
+## Method
+
+ * Obtain number of TMH AAs in proteome
+ * Convert genomic evolution rate to proteomic mutation rate
+ * Estimate the number of observed mutations that cause the pathogen
+   to lose/acquire an epitope
+
+### Obtain number of TMH AAs in proteome
+
+We start from a Uniprot reference proteome:
+
+```{r}
+fasta_filename <- "UP000464024.fasta"
 expect_true(file.exists(fasta_filename))
-topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
 ```
 
-```{r}
+We determine the topology:
+
+```{r cache=TRUE}
+topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
+n_proteins <- nrow(topology)
+n_tmhs <- sum(tally_tmhs(topology)$n_tmhs)
 n_non_tmh_aas <- sum(stringr::str_count(topology$topology, "0"))
-message("number of non-TMH amino acids, n_non_tmh_aas: ", n_non_tmh_aas)
+n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
+n_aas <- n_non_tmh_aas + n_tmh_aas
+f_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
+proteome_size <- n_aas
 ```
 
+From the topology, we learn:
+
+ * Proteome size: `r proteome_size` AAs
+ * Proteins: there are `r n_proteins` proteins   
+ * TMHs: there are `r n_tmhs` TMHs. 
+   Fraction of proteome that is TMH is `r f_aas`
+ * Amino acids: there are `r n_aas` AAs, 
+   of which are `r n_tmh_aas` TMH AAs (and `r n_non_tmh_aas` non-TMH AAs)
+
+### Convert genomic evolution rate to proteomic mutation rate
+
+Get the proteome mutation rate, by converting it from the 
+known DNA mutation rate.
+
 ```{r}
-n_tmh_aas <- sum(stringr::str_count(topology$topology, "1"))
-message("number of TMH amino acids, n_tmh_aas: ", n_tmh_aas)
+genome_size <- 29903
 ```
 
+We know from `wu2020new` (`https://www.ncbi.nlm.nih.gov/genbank/sars-cov-2-seqs/`) 
+that the reference genome size has `r genome_size` nucleotides.
+
 ```{r}
-n_aas <- n_non_tmh_aas + n_tmh_aas
-message("proteome size, number of amino acids, n_aas: ", n_aas)
+dna_evolution_rate_per_genome_per_year <- 26
 ```
 
+We know from GISAID (also `bar2020sars`) that COVID-19 evolves 
+with a rate of `r dna_evolution_rate_per_genome_per_year` mutations 
+per genome per year.
+
 ```{r}
-perc_aas <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
-message("percentage of proteome that is TMH, perc_aas: ", perc_aas)
+rel_size_proteome_to_genome <- proteome_size / genome_size
 ```
 
+The relative size of the proteome compared to the genome is
+`r rel_size_proteome_to_genome`, which is approximately one third,
+as three nucleotides code for one AA.
+
 ```{r}
-n_tmhs <- sum(tally_tmhs(topology)$n_tmhs) 
-message("number of TMHs, n_tmhs: ", n_tmhs)
+f_p_mut_synonymous <- 20 / 64
 ```
+
+The fraction of DNA mutations that does change the AA 
+is `r f_p_mut_synonymous`, estimated by taking the number of AAs per
+possible DNA triplet combinations.
+
+```{r}
+aa_evolution_rate_per_genome_per_year <- 
+  dna_evolution_rate_per_genome_per_year * f_p_mut_synonymous *
+  rel_size_proteome_to_genome
+```
+
+The proteomic evolution rate per proteome per year is then
+`r aa_evolution_rate_per_genome_per_year` amino acids per proteome per year.
+
+```{r}
+tmh_aa_evolution_rate_per_genome_per_year <- 
+  aa_evolution_rate_per_genome_per_year * f_aas
+```
+
+Because only `r f_aas` of all AAs is part of a TMH,
+we expect `r tmh_aa_evolution_rate_per_genome_per_year` TMH AAs
+changed per year.
+
+```{r}
+years_per_pandemic <- 0.5
+```
+
+The first COVID-19 sequence (`https://www.ncbi.nlm.nih.gov/nuccore/NC_045512`)
+was created in December 2019. Currenly, May 2020, 
+is `r years_per_pandemic` years later.
+
+```{r}
+f_avg_time_per_sequence <- 0.5
+```
+
+The COVID-19 sequences are obtained during that time, 
+instead of being collected today. Due to this, less mutations
+will be visible. Due to this, we assume that on average, sequences
+are collected halfway the pandemic, having a proportion of
+`r f_avg_time_per_sequence` of the mutation information.
+
+```{r}
+tmh_aa_evolution_rate_per_sequence <-
+  tmh_aa_evolution_rate_per_genome_per_year *
+  years_per_pandemic *
+  f_avg_time_per_sequence
+```
+
+Due to this, per proteome sequence, we 
+expect to observe `r tmh_aa_evolution_rate_per_sequence`
+TMH AA mutations per sequence.
+
+```{r}
+n_sequences <- 4329
+```
+
+There are `r n_sequences` COVID-19 sequences (`https://nextstrain.org/ncov`).
+
+```{r}
+n_tmh_aa_mutations <- tmh_aa_evolution_rate_per_sequence * n_sequences
+```
+
+Therefore, for the complete data set, we expect to 
+find `r n_tmh_aa_mutations` AA mutations in TMHs.
+
+##  Estimate the number of state transitions
+
+For the complete data set, we expect to 
+find `r n_tmh_aa_mutations` AA mutations in TMHs.
+
+Not all mutations, however, will cause the pathogen to lose or acquire
+an epitope.
+
+```{r}
+p_state_transition <- 0.01
+```
+
+Based upon nothing but our gut instinct, we estimate this
+chance at `r p_state_transition`.
+
+```{r}
+n_state_transitions <- p_state_transition * n_tmh_aa_mutations
+```
+
+So, for the full COVID-19 dataset, we expect to 
+find `r n_state_transitions` state transitions.
+
+## Conclusion
+
+```{r}
+draw_conclusion <- function(n_state_transitions) {
+  if (n_state_transitions < 1) return("useless")
+  if (n_state_transitions < 10) return("weak")
+  if (n_state_transitions < 20) return("fine")
+  if (n_state_transitions < 30) return("okay")
+  if (n_state_transitions < 50) return("good")
+  if (n_state_transitions < 100) return("great")
+  "awesome"
+}
+```
+
+Based upon nothing but our gut instinct, we base a conclusion on the
+number of transitions as shown in the code above.
+
+```{r}
+conclusion <- draw_conclusion(n_state_transitions)
+```
+
+From that we conclude that for our expected `r n_state_transitions` state
+transitions, that our dataset is `r conclusion`.

### 2020-05-21 p_transition/p_transition.Rmd
Run

---
title: "Chance of a transition"
author: Richel J.C. Bilderbeek
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

To determine the chance of mutation that loses or acquires an epitope,
while remaining TMH.

## Setup

We load the libraries needed:

```{r, message=FALSE}
library(testthat)
library(pureseqtmr)
```

PureseqTM, which predicts TMHs, must be installed:

```{r}
expect_true(is_pureseqtm_installed())
```

## Abbreviations

Abbr  |Full
------|-----------
AA    |Amino acid
AAs   |Amino acids
Abbr  |Abbreviation
TMH   |Transmembrane helix
TMHs  |Transmembrane helices

## Method

```{r}
mhc1_haplotype <- "HLA-A26-01"
```

We need to pick a haplotype. I will pick 'HLA-A26-01', because 50 percent
of its epitopes is a TMH. I got that info, 
from `find_tmhs_that_do_not_bind_to_mhc1/find_tmhs_that_do_not_bind_to_mhc1.csv`:

```
"mhc_1_haplotype","percentage of epitopes that are not TMH"
"HLA-A01-01",0.890000000000001
"HLA-A02-01",0
"HLA-A03-01",0.830000000000001
"HLA-A24-02",0.28
"HLA-A26-01",0.51
"HLA-B07-02",0.81
"HLA-B08-01",0.7
"HLA-B18-01",0.74
"HLA-B27-05",0.880000000000001
"HLA-B39-01",0.23
"HLA-B40-02",0.830000000000001
"HLA-B58-01",0.11
"HLA-B15-01",0.08
```

```{r}
n <- 2
ps <- rep(NA, n)
for (i in seq_len(n))
  protein_sequence <- bbbq::create_random_tmh(n_aas = 30)
  expect_true(pureseqtmr::is_tmh(protein_sequence))
  p <- calc_p_det_tmh_mut(
    protein_sequence = protein_sequence,
    mhc1_haplotype = mhc1_haplotype
  )
  ps[i] <- p
}
ps
```

### 2020-05-21 p_transition/p_transition.Rmd
Can make

@@
 ps <- rep(NA, n)
-for (i in seq_len(n))
+for (i in seq_len(n)) {
   protein_sequence <- bbbq::create_random_tmh(n_aas = 30)

### 2020-05-21 p_transition/p_transition.Rmd
Use MHC-I haplotype

--- 2020-05-21 p_transition/p_transition.Rmd
@@
 library(pureseqtmr)
+library(bbbq)
 ```
@@
 ```{r}
-mhc1_haplotype <- "HLA-A26-01"
+mhc_1_haplotype <- "HLA-A26-01"
 ```
@@
     protein_sequence = protein_sequence,
-    mhc1_haplotype = mhc1_haplotype
+    mhc_1_haplotype = mhc_1_haplotype
   )

### 2020-05-21 scripts/check_dead_links.sh
Check dead links

@@ -1,4 +1,11 @@
 #!/bin/bash
+#
+# Check dead links
+#
+# Usage:
+#
+#   ./script/check_dead_links.sh
+#
 
 # Collect all dead links in a file
 find . -name \*.md -exec markdown-link-check {} \; 2>/dev/null | egrep "[✖]" > broken.txt

### 2020-05-21 view.sh
Doc some scripts

@@ -1,2 +1,9 @@
 #!/bin/bash
+#
+# View the created PDF
+#
+# Usage:
+#
+#   ./view.sh
+#
 evince bbbq_article.pdf

### 2020-05-28 bbbq_article.tex
EOD

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{EpitopePrediction versus MHCnuggets}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+EpitopePrediction is a recent tool, used in \cite{bianchi2017},
+and can predict MHC-I epitopes.
+MHCnuggets is an even newer tool to predict both MHC-I and MHC-II epitopes.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{ep_vs_mhcn/ep_vs_mhcn.png}
+  Predicted IC50 (in nM) for EpitopePrediction and MHCnuggets
+  \caption{
+    \richel{Used randomly simulated polypeptides, results are real}
+  }
+  \label{fig:ep_vs_mhcn}
+\end{figure}
+
+\begin{table}[!htbp]
+  \input{ep_vs_mhcn/ep_vs_mhcn.latex}
+  \caption{
+    Correlation coefficients for different haplotypes
+    \richel{Used randomly simulated polypeptides, results are real}
+  }
+  \label{table:ep_vs_mhcn}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Relation between hydrophobicity, TMH and MHC binding}

### 2020-05-28 ep_vs_mhcn/ep_vs_mhcn.R

@@ -24,6 +24,8 @@ df <- tibble::as_tibble(
     stringsAsFactors = FALSE
   )
 )
+
+write.csv(df, "ic50s.csv")
 df
 n_rows <- nrow(df)
 for (i in seq_len(n_rows)) {
@@ -49,20 +51,18 @@ df_scatter <- tidyr::pivot_wider(
   names_from = tool,
   values_from = ic50
 )
-df_scatter$mhcn <- df_scatter$mhcn * 1000
 
 ggplot(df_scatter, aes(x = ep, y = mhcn, color = haplotype)) +
   geom_abline(slope = 1, intercept = 0, lty = "dashed") +
   geom_point() +
   scale_y_continuous(limits = c(0, max(df_scatter$mhcn))) +
-  scale_x_continuous(limits = c(0, max(df_scatter$ep))) +
-  geom_smooth(method = "lm") +
+  scale_x_continuous(limits = c(0, max(df_scatter$mhcn))) +
+  geom_smooth(method = "lm", alpha = 0.1) +
   xlab("IC50 predicted by EpitopePrediction") +
-  ylab("IC50 predicted by MHCnuggets (x1000)") +
+  ylab("IC50 predicted by MHCnuggets") +
   labs(
     caption = paste0(
       "Dashed line: x = y. ",
       "Blue line = fit to linear model, should ideally match the dashed line"
     )
   ) + ggsave("ep_vs_mhcn.png", width = 7, height = 7)
-

### 2020-06-08 ep_vs_mhcn/make.sh
Use 20 GB

@@ -9,7 +9,7 @@
 #SBATCH --nodes=1
 #SBATCH --ntasks-per-node=1
 #SBATCH --ntasks=1
-#SBATCH --mem=10G
+#SBATCH --mem=20G
 #SBATCH --job-name=make_data_ep_vs_mhcn
 #SBATCH --output=make_data_ep_vs_mhcn.log
 module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop

### 2020-06-08 ep_vs_mhcn/create_table.R
Only use long form

@@ -1,25 +1,0 @@
-library(ggplot2)
-library(testthat)
-
-expect_true(file.exists("ep_vs_mhcn.csv"))
-df <- readr::read_csv("ep_vs_mhcn.csv")
-
-# tool | haplotype | IC50
-# -----|-----------|------
-# EpitopPrediction |
-expect_true("haplotype" %in% names(df))
-expect_true("ic50" %in% names(df))
-expect_true("tool" %in% names(df))
-
-# Reshape from tidy data to wide data
-df_wide <- tidyr::pivot_wider(
-  df,
-  names_from = tool,
-  values_from = ic50
-)
-df_wide
-write.csv(df_wide, "ep_vs_mhcn_wide.csv", row.names = FALSE)
-
-
-expect_true(file.exists("ep_vs_mhcn_wide.csv"))
-

### 2020-06-08 ep_vs_mhcn/make.sh

@@ -10,8 +10,8 @@
 #SBATCH --ntasks-per-node=1
 #SBATCH --ntasks=1
 #SBATCH --mem=20G
-#SBATCH --job-name=make_data_ep_vs_mhcn
-#SBATCH --output=make_data_ep_vs_mhcn.log
+#SBATCH --job-name=ep_vs_mhcn
+#SBATCH --output=ep_vs_mhcn.log
 module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop
 make
 

### 2020-06-08 p_transition/make.sh
Add 'make.sh' for Peregrine

@@ -1,0 +1,17 @@
+#!/bin/bash
+# Bash script to run make with sbatch
+#
+# Usage:
+#
+#   sbatch make.sh
+#
+#SBATCH --time=240:00:00
+#SBATCH --nodes=1
+#SBATCH --ntasks-per-node=1
+#SBATCH --ntasks=1
+#SBATCH --mem=10G
+#SBATCH --job-name=p_transition
+#SBATCH --output=p_transition.log
+module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop
+make
+

### 2020-06-08 p_transition/make.sh
Load Pandoc

@@ -12,6 +12,6 @@
 #SBATCH --mem=10G
 #SBATCH --job-name=p_transition
 #SBATCH --output=p_transition.log
-module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop
+module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop Pandoc
 make
 

### 2020-06-08 p_bind_per_hydrophobicity/make.sh
Better log name

@@ -10,8 +10,8 @@
 #SBATCH --ntasks-per-node=1
 #SBATCH --ntasks=1
 #SBATCH --mem=10G
-#SBATCH --job-name=make_p_bind_per_hydrophobicity
-#SBATCH --output=make_p_bind_per_hydrophobicity.log
+#SBATCH --job-name=p_bind_per_hydrophobicity
+#SBATCH --output=p_bind_per_hydrophobicity.log
 module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop
 make
 

### 2020-06-08 p_bind_per_hydrophobicity/p_bind_per_hydrophobicity.Rmd
Try to render with Pandoc

---
title: "p_bind"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Calculate the relation between hydrophobicity
# and chance to bind to MHC-I.
#
# H0: there is no relation
# H1: higher hydrophobicity equals higher chance to bind
#
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(testthat))

plot_is_tmh_vs_binds_mhc1 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df_tally <- df %>%
    dplyr::group_by(is_tmh, binds_mhc1) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(is_tmh, binds_mhc1, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_x_discrete() +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Is TMH? Left = no, right = yes") +
      ggplot2::ylab("Binds to MHC-I? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_is_tmh_vs_binds_mhc2 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df_tally <- df %>%
    dplyr::group_by(is_tmh, binds_mhc2) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(is_tmh, binds_mhc2, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_x_discrete() +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Is TMH? Left = no, right = yes") +
      ggplot2::ylab("Binds to MHC-II? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_binds_mhc1_vs_binds_mhc2 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df_tally <- df %>%
    dplyr::group_by(binds_mhc1, binds_mhc2) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(binds_mhc1, binds_mhc2, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_x_discrete() +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Binds to MHC-I?? Left = no, right = yes") +
      ggplot2::ylab("Binds to MHC-II? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_hydrophobicity_vs_is_tmh <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df <- tibble::as_tibble(df)

  df_tally <- df %>%
    dplyr::mutate(bin = trunc(hydrophobicity * 2)) %>%
    dplyr::group_by(bin, is_tmh) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(bin, is_tmh, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Hydrophobicity x2") +
      ggplot2::ylab("Is TMH? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_hydrophobicity_vs_binds_mhc1 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df <- tibble::as_tibble(df)
  df_tally <- df %>%
    dplyr::mutate(bin = trunc(hydrophobicity * 2)) %>%
    dplyr::group_by(bin, binds_mhc1) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(bin, binds_mhc1, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Hydrophobicity x2") +
      ggplot2::ylab("Binds to MHC-I? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

plot_hydrophobicity_vs_binds_mhc2 <- function(
  df,
  png_filename = tempfile(fileext = ".png")
) {
  df <- tibble::as_tibble(df)

  df_tally <- df %>%
    dplyr::mutate(bin = trunc(hydrophobicity * 2)) %>%
    dplyr::group_by(bin, binds_mhc2) %>%
    dplyr::tally()
  ggplot2::ggplot(df_tally, ggplot2::aes(bin, binds_mhc2, fill = n)) +
    ggplot2::geom_tile() +
    ggplot2::geom_text(ggplot2::aes(label = n)) +
      ggplot2::scale_fill_gradient(low = "white", high = "red") +
      ggplot2::scale_y_discrete() +
      ggplot2::xlab("Hydrophobicity x2") +
      ggplot2::ylab("Binds to MHC-II? Down = no, up = yes") +
      ggplot2::ggsave(png_filename, width = 7, height = 7)
}

raw_data <- "p_bind_per_hydrophobicity.csv"
expect_true(file.exists(raw_data))
df <- read.csv(raw_data)

plot_is_tmh_vs_binds_mhc1(df = df, png_filename = "is_tmh_vs_binds_mhc1.png")
plot_is_tmh_vs_binds_mhc2(df = df, png_filename = "is_tmh_vs_binds_mhc2.png")
plot_binds_mhc1_vs_binds_mhc2(df = df, png_filename = "binds_mhc1_vs_binds_mhc2.png")
plot_hydrophobicity_vs_is_tmh(df = df, png_filename = "hydrophobicity_vs_is_tmh.png")
plot_hydrophobicity_vs_binds_mhc1(df = df, png_filename = "hydrophobicity_vs_binds_mhc1.png")
plot_hydrophobicity_vs_binds_mhc2(df = df, png_filename = "hydrophobicity_vs_binds_mhc2.png")
```

### 2020-06-08 p_bind_per_hydrophobicity/make.sh

@@ -12,6 +12,6 @@
 #SBATCH --mem=10G
 #SBATCH --job-name=p_bind_per_hydrophobicity
 #SBATCH --output=p_bind_per_hydrophobicity.log
-module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop
+module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop Pandoc
 make
 

### 2020-06-08 p_bind_per_hydrophobicity/create_peptides.R
Back to 10 peptides

@@ -4,7 +4,7 @@ library(bbbq)
 suppressMessages(library(dplyr))
 
 n_aas <- 40
-n_peptides_per_hydrophobicity <- 1000
+n_peptides_per_hydrophobicity <- 10
 
 df <- tibble::as_tibble(
   expand.grid(

### 2020-06-08 p_bind_per_hydrophobicity/create_dataset.R
Be silent

@@ -19,5 +19,4 @@ for (i in seq_len(n_rows)) {
   # df$binds_mhc2[i] <- bbbq::is_detected_by_mhc_2(sequence)
 }
 
-df
 readr::write_csv(df, "p_bind_per_hydrophobicity.csv")

### 2020-06-08 p_bind_per_hydrophobicity/create_peptides.R
Only 2 peptides for now

@@ -4,7 +4,7 @@ library(bbbq)
 suppressMessages(library(dplyr))
 
 n_aas <- 40
-n_peptides_per_hydrophobicity <- 10
+n_peptides_per_hydrophobicity <- 2
 
 df <- tibble::as_tibble(
   expand.grid(

### 2020-06-08 p_bind_per_hydrophobicity/make.sh
Only make data

@@ -12,6 +12,6 @@
 #SBATCH --mem=10G
 #SBATCH --job-name=p_bind_per_hydrophobicity
 #SBATCH --output=p_bind_per_hydrophobicity.log
-module load R Python/3.8.2-GCCcore-9.3.0 binutils ImageMagick X11 libX11 xprop Pandoc
+module load R Python/3.8.2-GCCcore-9.3.0
 make
 

### 2020-06-08 p_bind_per_hydrophobicity/make.sh
Make data on Peregrine, make figures on local

@@ -13,5 +13,5 @@
 #SBATCH --job-name=p_bind_per_hydrophobicity
 #SBATCH --output=p_bind_per_hydrophobicity.log
 module load R Python/3.8.2-GCCcore-9.3.0
-make
+make data
 

### 2020-06-08 p_bind_per_hydrophobicity/create_dataset.R
Peregrine creates data

@@ -8,7 +8,6 @@ df <- readr::read_csv(peptides_filename)
 df$is_tmh <- NA
 df$binds_mhc1 <- NA
 df$binds_mhc2 <- NA
-df
 
 n_rows <- nrow(df)
 

### 2020-06-09 p_bind_per_hydrophobicity/create_peptides.R
Use 1000 peptides per hydrophobicity

@@ -7,10 +7,10 @@ n_aas <- 40
 
 # 2: 30 secs
 # 10: 150 sec = 3 mins
-# 100: 1500 sec = 30 mins
-# 1000: 15000 sec = 300 mins = 5 hours
+# 100: 1500 sec = 30 mins indeed
+# 1000: 15000 sec = 300 mins = 5 hours predicted
 
-n_peptides_per_hydrophobicity <- 100
+n_peptides_per_hydrophobicity <- 1000
 
 df <- tibble::as_tibble(
   expand.grid(

### 2020-06-09 bbbq_article.tex
Do stuff on Peregrine (create data) and on local (create figures)
stub 'ic50s_per_haplotype' result

@@
 
-\begin{table}[!htbp]
-  \input{ep_vs_mhcn/ep_vs_mhcn.latex}
-  \caption{
-    Correlation coefficients for different haplotypes
+%\begin{table}[!htbp]
+%  \input{ep_vs_mhcn/ep_vs_mhcn.latex}
+%  \caption{
+%    Correlation coefficients for different haplotypes
+%    \richel{Used randomly simulated polypeptides, results are real}
+%  }
+%  \label{table:ep_vs_mhcn}
+%\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{IC50s per haplotype}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Here we have simulated \richel{some, say 100} random peptides.
+Per peptide, we measured the IC50 
+per MHC-I (figure \label{fig:ic50s_per_haplotype_mhc_1})
+and MHC-II  (figure \label{fig:ic50s_per_haplotype_mhc_2}) haplotype.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype_mhc_1.png}
+  Predicted IC50s (in nM) per MHC-I haplotype
+  \caption{
     \richel{Used randomly simulated polypeptides, results are real}
   }
-  \label{table:ep_vs_mhcn}
-\end{table}
+  \label{fig:ic50s_per_haplotype_mhc_1}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype_mhc_2.png}
+  Predicted IC50s (in nM) per MHC-II haplotype
+  \caption{
+    \richel{Used randomly simulated polypeptides, results are real}
+  }
+  \label{fig:ic50s_per_haplotype_mhc_2}
+\end{figure}
 

### 2020-06-15 ic50s_per_haplotype/create_dataset.R
EOD

@@ -1,43 +1,0 @@
-suppressMessages(library(ggplot2))
-suppressMessages(library(dplyr))
-library(testthat)
-
-peptides_filename <- "peptides.csv"
-expect_true(file.exists(peptides_filename))
-df <- readr::read_csv(peptides_filename)
-
-df <- rbind(
-  tidyr::expand_grid(peptide = df$peptide, mhc_class = "I", haplotype = bbbq::get_mhc1_haplotypes()),
-  tidyr::expand_grid(peptide = df$peptide, mhc_class = "II", haplotype = bbbq::get_mhc2_haplotypes())
-)
-df$ic50 <- NA
-
-n_rows <- nrow(df)
-for (i in seq_len(n_rows)) {
-  mhcnuggets_name <- mhcnuggetsr::to_mhcnuggets_name(df$haplotype[i])
-  if (!mhcnuggetsr::is_mhcnuggets_name(mhcnuggets_name)) {
-    stop(
-      "Not MHCnuggets name. \n",
-      "Haplotype: ", df$haplotype[i], "\n",
-      "mhcnuggets_name: ", mhcnuggets_name
-    )
-  }
-}
-
-
-for (i in seq_len(n_rows)) {
-  peptide <- df$peptide[i]
-  mhcnuggets_options <- mhcnuggetsr::create_mhcnuggets_options(
-    mhc_class =  df$mhc_class[i],
-    mhc = mhcnuggetsr::to_mhcnuggets_name(df$haplotype[i])
-  )
-  mhcnuggetsr::check_mhcnuggets_options(mhcnuggets_options)
-  ic50 <- mhcnuggetsr::predict_ic50(
-    mhcnuggets_options = mhcnuggets_options,
-    peptides = peptide
-  )$ic50
-  expect_equal(1, length(ic50))
-  df$ic50[i] <- ic50
-}
-
-readr::write_csv(df, "ic50s_per_haplotype.csv")

### 2020-06-15 ic50s_per_haplotype/create_figures.R

@@ -1,19 +1,0 @@
-# Get the distributions of IC50s for random peptides per haplotype
-suppressMessages(library(ggplot2))
-suppressMessages(library(dplyr))
-suppressMessages(library(testthat))
-
-df <- readr::read_csv("ic50s_per_haplotype.csv")
-
-n_peptides <- length(unique(df$peptide))
-n_aas <- nchar(df$peptide)[1]
-
-ggplot(df, aes(x = as.factor(haplotype), y = ic50)) +
-  geom_boxplot() +
-  xlab("Haplotype") +
-  ylab("IC50 (nM)") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  ggplot2::labs(
-    caption = glue::glue(
-      "Number of random peptides: {n_peptides}, peptide length: {n_aas} amino acids"
-    )
-  ) + ggplot2::ggsave("ic50s_per_haplotype.png", width = 7, height = 7)

### 2020-06-15 ic50s_per_haplotype/create_peptides.R

@@ -1,17 +1,0 @@
-# Creates peptides.csv
-
-library(bbbq)
-suppressMessages(library(dplyr))
-
-n_aas <- 9
-
-# 2: 1.5 minutes
-# 100: 4 hours
-# 1000: ?40 hours
-n_peptides <- 1000 # per haplotype
-
-t <- tibble::tibble(
-  peptide = replicate(n = n_peptides, bbbq::create_random_peptide(n_aas = n_aas))
-)
-
-readr::write_csv(t, "peptides.csv")

### 2020-06-15 ic50s_per_haplotype/make.sh

@@ -1,18 +1,0 @@
-#!/bin/bash
-# Bash script to run make with sbatch
-#
-# Usage:
-#
-#   sbatch make.sh
-#
-#SBATCH --time=240:00:00
-#SBATCH --nodes=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --ntasks=1
-#SBATCH --mem=20G
-#SBATCH --job-name=ic50s_per_haplotype
-#SBATCH --output=ic50s_per_haplotype.log
-module load R Python/3.8.2-GCCcore-9.3.0 ImageMagick X11 
-
-make
-

### 2020-06-22 covid_dataset_strength/covid_dataset_strength.Rmd
Use better estimated value

@@
 ```{r cache=TRUE}
-topology <- pureseqtmr::predict_proteome_topology(fasta_filename)
+topology <- pureseqtmr::predict_topology(fasta_filename)
 n_proteins <- nrow(topology)
@@
 
+This chance we call it `r p_state_transition`.
+
+Based upon simulations (see https://github.com/richelbilderbeek/bbbq_p_trans), we 
+estimate `r p_state_transition`:
+
+![](p_trans_0_02.png)
+
 ```{r}
-p_state_transition <- 0.01
+p_state_transition <- 0.0494
 ```
-
-Based upon nothing but our gut instinct, we estimate this
-chance at `r p_state_transition`.
 

### 2020-06-22 covid_dataset_strength/covid_dataset_strength.Rmd
Dataset is awesome

@@
 
+![](nextstrain.png)
 ```{r}
-dna_evolution_rate_per_genome_per_year <- 26
+dna_evolution_rate_per_genome_per_year <- 23.565
 ```
@@
 The first COVID-19 sequence (`https://www.ncbi.nlm.nih.gov/nuccore/NC_045512`)
-was created in December 2019. Currenly, May 2020, 
+was created in December 2019. Currenly, June 2020, 
 is `r years_per_pandemic` years later.
@@
 
+![](gisaid.png)
+
 ```{r}
-n_sequences <- 4329
+n_sequences <- 49790
 ```
 
-There are `r n_sequences` COVID-19 sequences (`https://nextstrain.org/ncov`).
+There are `r n_sequences` COVID-19 sequences (`https://www.gisaid.org/`).
 
@@
 
-This chance we call it `r p_state_transition`.
+![](p_trans_0_02.png)
 
-Based upon simulations (see https://github.com/richelbilderbeek/bbbq_p_trans), we 
-estimate `r p_state_transition`:
-
-![](p_trans_0_02.png)
+> This plot shows the chance of a state transition `p`,
+> per haplotype, based on simulated data
 
@@
 ```
+
+Based upon simulations (see https://github.com/richelbilderbeek/bbbq_p_trans), we 
+estimated this chance at `p_state_transition`:
 

### 2020-06-22 covid_dataset_strength/covid_dataset_strength.Rmd
We conclude that for our expected 139.3155572 state transitions, that the dataset is awesome

@@
 
-Based upon nothing but our gut instinct, we base a conclusion on the
-number of transitions as shown in the code above.
+Based on the expected number of transitions, `r n_state_transitions`, we draw a conslusion.
 
@@
 
-From that we conclude that for our expected `r n_state_transitions` state
-transitions, that our dataset is `r conclusion`.
+We conclude that for our expected `r n_state_transitions` state
+transitions, that the dataset is `r conclusion`.

### 2020-06-22 bbbq_article.tex
EOD
Can predict peptides

--- 2020-05-20 bbbq_article.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Strength of Influenza A}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Our evolutionary experiment depends on the strength of the dataset.
-Here we do a back-on-the-envelope calculation to assess if the dataset
-is strong enough.
-
-\begin{itemize}
-	\item Genome size: $\approx 1.3 \cdot 10^4$ bases
-	\item Relative genome size compared to COVID-19: $\approx 0.4$ (as $\frac{1.3 \cdot 10^4}{3.0 \cdot 10^4}$)
-	\item Full genome sets: $\approx 3.2 \cdot 10^4$ (Influenza A)
-	\item Relative number of genome sets compared to COVID-19: $\approx 6.4$ as much (as $\frac{\approx 32 \cdot 10^3}{5 \cdot 10^3}$)
-	\item Relative time to accumulate mutations: $\approx 12$ times longer (average age $\approx$ 4 year, before 2015, already 8k sequences)
-	\item Relative chance of observing transitions: $\approx 30$ (as $0.4 \cdot 6.4 \cdot 12$)
-	\item Expected number of observed TMH mutations in COVID-19 that cause a transition: $\approx 8$
-	\item Expected number of observed TMH mutations that cause a transition: $\approx 240$ (as $8 \cdot 30$)
-\end{itemize}
-
-Conclusion: seems strong enough to us.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{COVID-19 genome and proteome}
@@
   \label{fig:covid_genome_and_proteome}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{IC50s per haplotype}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Per MHC haplotype, we simulated random peptides and measured the
+predicted IC50 using MHCnuggets. 
+Here, we show the IC50 value per haplotype that
+is used to determine if a peptide binds to the haplotype's MHC.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
+  Predicted IC50s (in nM) per MHC-I haplotype
+  \caption{
+    \richel{Used randomly simulated polypeptides, results are real}
+  }
+  \label{fig:ic50s_per_haplotype}
 \end{figure}

### 2020-06-22 p_bind_per_hydrophobicity/create_dataset.R
Nice figures

@@ -1,21 +1,29 @@
-suppressMessages(library(ggplot2))
-suppressMessages(library(dplyr))
+library(dplyr, quietly = TRUE)
 library(testthat)
+library(readr)
 
-peptides_filename <- "peptides.csv"
-expect_true(file.exists(peptides_filename))
-df <- readr::read_csv(peptides_filename)
-df$is_tmh <- NA
-df$binds_mhc1 <- NA
-df$binds_mhc2 <- NA
+is_tmh_filename <- "is_tmh.csv"
+expect_true(file.exists(is_tmh_filename))
+is_binder_filename <- "is_binder.csv"
+expect_true(file.exists(is_binder_filename))
 
-n_rows <- nrow(df)
+is_tmh <- read_csv(
+  is_tmh_filename,
+   col_types = cols(
+     sequence = col_character(),
+     hydrophobicity = col_double(),
+     is_tmh = col_logical()
+   )
+)
+is_binder <- readr::read_csv(
+  is_binder_filename,
+   col_types = cols(
+     sequence = col_character(),
+     mhc_haplotype = col_character(),
+     percentile = col_double(),
+     is_binder = col_logical()
+   )
+)
 
-for (i in seq_len(n_rows)) {
-  sequence <- df$sequence[i]
-  df$is_tmh[i] <- pureseqtmr::is_tmh(sequence)
-  df$binds_mhc1[i] <- bbbq::is_detected_by_mhc_1(sequence)
-  # df$binds_mhc2[i] <- bbbq::is_detected_by_mhc_2(sequence)
-}
-
-readr::write_csv(df, "p_bind_per_hydrophobicity.csv")
+t <- dplyr::full_join(is_tmh, is_binder)
+readr::write_csv(t, "dataset.csv")

### 2020-06-22 p_bind_per_hydrophobicity/create_figure_hydrophobicity_vs_is_tmh.R

@@ -1,0 +1,27 @@
+# Calculate the relation between hydrophobicity
+# and chance to bind to MHC-I.
+#
+# H0: there is no relation
+# H1: higher hydrophobicity equals higher chance to bind
+#
+library(ggplot2, quietly = TRUE)
+library(dplyr, quietly = TRUE)
+library(testthat, quietly = TRUE)
+
+raw_data <- "is_tmh.csv"
+expect_true(file.exists(raw_data))
+df <- readr::read_csv(raw_data)
+
+df_tally <- df %>%
+  mutate(bin = trunc(hydrophobicity * 2)) %>%
+  group_by(bin, is_tmh) %>%
+  tally()
+
+ggplot(df_tally, aes(bin, is_tmh, fill = n)) +
+  geom_tile() +
+  geom_text(aes(label = n)) +
+    scale_fill_gradient(low = "white", high = "red") +
+    scale_y_discrete() +
+    xlab("Hydrophobicity x2") +
+    ylab("Is TMH? Down = no, up = yes") +
+    ggsave("hydrophobicity_vs_is_tmh.png", width = 7, height = 7)

### 2020-06-22 p_bind_per_hydrophobicity/create_figures.R

@@ -4,63 +4,72 @@
 # H0: there is no relation
 # H1: higher hydrophobicity equals higher chance to bind
 #
-suppressMessages(library(ggplot2))
-suppressMessages(library(dplyr))
-suppressMessages(library(testthat))
+library(ggplot2, quietly = TRUE)
+library(dplyr, quietly = TRUE)
+library(testthat, quietly = TRUE)
+library(bbbq, quietly = TRUE)
 
-plot_is_tmh_vs_binds_mhc1 <- function(
+plot_is_tmh_vs_is_binder <- function(
   df,
   png_filename = tempfile(fileext = ".png")
 ) {
+  mhc_class <- NA
+  if (df$mhc_haplotype[1] %in% bbbq::get_mhc1_haplotypes()) {
+    mhc_class <- "I"
+    testthat::expect_true(
+      all(df$mhc_haplotype %in% bbbq::get_mhc1_haplotypes())
+    )
+  }
+  if (df$mhc_haplotype[1] %in% bbbq::get_mhc2_haplotypes()) {
+    mhc_class <- "II"
+    testthat::expect_true(
+      all(df$mhc_haplotype %in% bbbq::get_mhc2_haplotypes())
+    )
+  }
+  testthat::expect_true(!is.na(mhc_class))
+  n_haplotypes <- length(unique(df$mhc_haplotype))
+  n_peptides <- length(unique(df$sequence))
+
+
   df_tally <- df %>%
-    dplyr::group_by(is_tmh, binds_mhc1) %>%
+    dplyr::group_by(is_tmh, is_binder) %>%
     dplyr::tally()
-  ggplot2::ggplot(df_tally, ggplot2::aes(is_tmh, binds_mhc1, fill = n)) +
+  ggplot2::ggplot(df_tally, ggplot2::aes(is_tmh, is_binder, fill = n)) +
     ggplot2::geom_tile() +
     ggplot2::geom_text(ggplot2::aes(label = n)) +
       ggplot2::scale_fill_gradient(low = "white", high = "red") +
       ggplot2::scale_x_discrete() +
       ggplot2::scale_y_discrete() +
       ggplot2::xlab("Is TMH? Left = no, right = yes") +
-      ggplot2::ylab("Binds to MHC-I? Down = no, up = yes") +
+      ggplot2::ylab(
+        glue::glue("Binds to MHC-{mhc_class}? Down = no, up = yes")
[…]

### 2020-06-22 p_bind_per_hydrophobicity/create_is_binder.R

@@ -1,0 +1,37 @@
+library(dplyr, quietly = TRUE)
+library(testthat)
+
+peptides_filename <- "peptides.csv"
+expect_true(file.exists(peptides_filename))
+t <- readr::read_csv(
+  peptides_filename,
+  col_types = readr::cols(
+    sequence = readr::col_character(),
+    hydrophobicity = readr::col_double()
+  )
+)
+
+mhc_haplotypes <- bbbq::get_mhc_haplotypes()
+if (!peregrine::is_on_peregrine()) {
+  mhc_haplotypes <- sample(mhc_haplotypes, size = 2)
+}
+
+t_binds <- tidyr::expand_grid(
+  sequence = t$sequence,
+  mhc_haplotype = bbbq::get_mhc_haplotypes(),
+  percentile = 0.02,
+  is_binder = NA
+)
+
+for (i in seq_len(nrow(t_binds))) {
+
+  t_binds$is_binder[i] <- bbbq::is_detected(
+    protein_sequence = t_binds$sequence[i],
+    mhc_haplotype = t_binds$mhc_haplotype[i],
+    fragment_length = 9,
+    percentile = t_binds$percentile[i]
+  )
+}
+t_binds
+
+readr::write_csv(t_binds, "is_binder.csv")

### 2020-06-22 p_bind_per_hydrophobicity/create_is_tmh.R

@@ -1,0 +1,20 @@
+library(dplyr, quietly = TRUE)
+library(testthat)
+
+peptides_filename <- "peptides.csv"
+expect_true(file.exists(peptides_filename))
+
+t_is_tmh <- readr::read_csv(
+  peptides_filename,
+  col_types = readr::cols(
+    sequence = readr::col_character(),
+    hydrophobicity = readr::col_double()
+  )
+)
+t_is_tmh$is_tmh <- NA
+
+for (i in seq_len(nrow(t_is_tmh))) {
+  t_is_tmh$is_tmh[i] <- pureseqtmr::is_tmh(t_is_tmh$sequence[i])
+}
+t_is_tmh
+readr::write_csv(t_is_tmh, "is_tmh.csv")

### 2020-06-22 p_bind_per_hydrophobicity/create_peptides.R

@@ -1,7 +1,7 @@
 # Creates peptides.csv
 
 library(bbbq)
-suppressMessages(library(dplyr))
+library(dplyr, quietly = TRUE)
 
 n_aas <- 40
 
@@ -10,7 +10,10 @@ n_aas <- 40
 # 100: 1500 sec = 30 mins indeed
 # 1000: 15000 sec = 300 mins = 5 hours predicted
 
-n_peptides_per_hydrophobicity <- 1000
+n_peptides_per_hydrophobicity <- 2
+if (peregrine::is_on_peregrine()) {
+  n_peptides_per_hydrophobicity <- 1000
+}
 
 df <- tibble::as_tibble(
   expand.grid(

### 2020-06-22 bbbq_article.tex
Improve

@@
 
-Transmembrane helices (TMHs) in the human proteome
-are an overrepresented potential source of epitopes on major 
-histocompatibility complex (MHC) class I for the majority of HLA-I haplotypes. 
-It is unknown if the immune system is as able to detect 
-TMHs as found in pathogens.
-Additionally, it is unknown if MHC-II is also likelier to detect
-human and pathogen TMHs than expected by chance.
-It is unknown if natural selection is the cause that 
-the immune system is more attentive for TMHs
+The human immune system is unexpectedly attentive towards
+human transmembrane helices (TMHs), as TMH-derived peptide fragments
+bind to histocompatibility complex (MHC) class I more often then expected
+by chance only.
+This finding hints that there has been selection upon detecting TMH-derived
+peptides in pathogens.
+If this signature of selection is present in MHC-II is unknown.
 This study shows that MHC-I [also has/does not have] more
@@
 a host proteome.
-Additionally, MHC-II binds to polypeptides derived from TMHs 
+Additionally, MHC-II binds to peptides derived from TMHs 
 [less/equally/more] often than expected by chance.
@@
 {\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
-adaptive immunity, transmembrane domain, epitopes, T lymphocyte, 
-MHC-1, MHC-I, MHC-2, MHC-II, COVID-19
+adaptive immunity, transmembrane domain, transmembrane helix, 
+epitopes, T lymphocyte, MHC-1, MHC-I, MHC-2, MHC-II, COVID-19
 
@@
 
-All nucleated cells in humans present randomly sampled polypeptides
+All nucleated cells in humans present randomly sampled peptides
 fragments to the surroundings of the cell using the Major 
 Histocompatibility Complex (MHC) class I molecules.
-If a cell gets infected by a virus, also the virus' polypeptides
+If a cell gets infected by a virus, also the virus' peptides
 will be presented at the cell surface. The foreign viral antigen is detected 
@@
 as these do not infect a cell with their (foreign) DNA.
-T and B cells express MHC-II to detect foreign polypeptide fragments.
+T and B cells express MHC-II to detect foreign peptide fragments.
 An immune response is started when MHC-II detects a pathogen.
@@
 Any human's immune system detects only a fraction of all possible
-polypeptide fragments. 
-Human MHCs are also called HLAs ('Human Leukocyte Antigens') and
-because this research uses human hosts only, we will use these terms
-interchangeably.
-Already each MHC molecule can only bind a, possible exclusive,
-subset of all possible polypeptides.
+peptide fragments.
+For humans, the MHC proteins are encoded by the
+HLA ('Human Leukocyte Antigens') genes.
+Each MHC complex can only bind a subset of all possible peptides.
 For example, HLA-A and HLA-B have no overlap in which
@@
 the first (for MHC-II, \cite{southwood1998several})
-or second position (for MHC-I, \cite{lund2004definition}) of a polypeptide.
-However, this does not imply that all polypeptides for a classification
-pattern bind to the MHC. Instead, software should be used that are
-tailored to predict if a polypeptide binds to an HLA.
+or second position (for MHC-I, \cite{lund2004definition}) of a peptide.
+However, this does not imply that when a peptide follows
+a classification pattern, that it will bind to the MHC. 
+Instead, software should be used to predict if the
+full peptide will do so.
 
@@
 
-It is helpful to be able to predict which polypeptides are immunogenic,
+It is helpful to be able to predict which peptides are immunogenic,
 in, among others, vaccine development \richel{Reference here}. 
-Already for a decade, synthesized polypeptide fragments are used 
-to experimentally determine which polypeptides
+Already for a decade, synthesized peptide fragments are used 
+to experimentally determine which peptides
 are immunogenic \richel{Reference here}.
 This approach, however, is tedious and costly.
-It should come as no surprise that these results would be used
-to make in silico predictions.
-It may come as a surprise, however, that these predictions
-are actually reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+Due to this, software was written that allows for silico 
+predictions, which are reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
 
@@
 which uses a stabilized matrix method \cite{kim2009derivation}.
-In this study, we will use \verb;epitope-prediction; \cite{bianchi2017}.
+The most recent FOSS tool available is MHCnuggets\cite{shao2020high},
+which we will use in this study
 
@@
 However, a different HLA will result in a different variety of MHCs,
-that will display different polypeptide fragments. Additionally, there
+that will display different peptide fragments. Additionally, there
 may have been selection the HLA to display and or detect different
-polypeptide fragments.
+peptide fragments.
 
@@
 
-It is important that MHC-I presents both the polypeptides of the 
-(healthy) cell, as well as possible pathogen-derived polypeptides when the
+It is important that MHC-I presents both the peptides of the 
+(healthy) cell, as well as possible pathogen-derived peptides when the
 cell is infected. As described above, MHC-I presents TMH-derived epitopes 
@@
 And maybe this has already happened: MHC-I already detects hydrophobic
-polypeptides. This feature, however, may also be caused by selection
+peptides. This feature, however, may also be caused by selection
 to detect hydrophobic regions in the soluble proteins of pathogens.
@@
 To detect an evolutionary signal for selection on binding
-TMHs polypeptides, we need a phylogenetic tree with 
+TMHs peptides, we need a phylogenetic tree with 
 many time-dated proteomes.
@@
 The threshold of 20\% is arbitrarily, yet chosen to result in 5 and 10 HLA variants.  
-As \cite{bianchi2017}, we define the 5\% of polypeptides 
+As \cite{bianchi2017}, we define the 5\% of peptides 
 with the lowest IC50 values as binders. We then counted the percentage
@@
 figure \ref{fig:markov_chain}. The two nodes depict the
-two possible detection state a polypeptide is in, which is
+two possible detection state a peptide is in, which is
 either undetected $U$, or detected $D$ by the immune system.
@@
 for example, $p_{DU}$ denotes the probability that a detected
-polypeptide mutates into an undetected one.
-We focus on polypeptides that are TMH and we assume a mutation that 
+peptide mutates into an undetected one.
+We focus on peptides that are TMH and we assume a mutation that 
 renders the TMH non-functional as lethal.
@@
 we can simply count the number of observed transitions.
-Because the probability of transitions differ per polypeptide,
+Because the probability of transitions differ per peptide,
 the total system has the properties of a Poisson binomial distribution,
@@
 
-\paragraph{Calculating one polypeptide's probability}
+\paragraph{Calculating one peptide's probability}
 
-We calculate the probability of a transition per polypeptide.
-For any polypeptide, we can generate all sequences that differ
+We calculate the probability of a transition per peptide.
+For any peptide, we can generate all sequences that differ
 in only one amino acids. Of these sequences, we only keep the ones
@@
 
-We assume that if a polypeptide binds to an MHC, it is always presented and always
+We assume that if a peptide binds to an MHC, it is always presented and always
 leads to an immune response.
@@
 All data sets consists out of $10^4$ observations.
-Each observation starts with a randomly generated source polypeptide fragment,
+Each observation starts with a randomly generated source peptide fragment,
 as well a mutant. For the 'null' data set, the mutant is randomly select.
@@
 to be derived from TMH within either a human host and its bacterial pathogen.
-Because a bacterium does not infect a cell, thus its polypeptides
+Because a bacterium does not infect a cell, thus its peptides
 will not be presented by MHC-I, this result is [unexpected/expected]
@@
   \caption{
-    \richel{Used randomly simulated polypeptides, results are real}
+    \richel{Used randomly simulated peptides, results are real}
   }

### 2020-06-23 bbbq_article.tex
Improved Methods quite a bit

@@
 \cite{trolle2014nettepi}, \cite{zhang2015accurate}
-and the very recent MHCnuggets\cite{shao2020high}.
+and the very recent MHCnuggets (\cite{shao2020high}).
 In this study, we'll be using MHCnuggets, as it is the only FOSS tool available.
@@
 of TMH predictions. 
-Unlike TMHMM, which is still up and running, 
-these later tools have already become obsolete.
-Contemporary tools that are good for use, such as
-the closed-source MemBrain \cite{feng2020topology} 
-and the FOSS tool PureseqTM \cite{wang2019efficient}.
-In this study, we will use TMHMM  for pragmatic reasons
-\richel{I would love to use PureseqTM instead, but only if there is time}.
+Contemporary tools are the closed-source MemBrain (\cite{feng2020topology}) 
+and the FOSS tool PureseqTM (\cite{wang2019efficient}).
+In this study, we will use PureseqTM.
 
@@
 especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
-response \cite{novy2007cd4}. 
+response (\cite{novy2007cd4}). 
 The mechanism to detect foreign TMHs by MHC-II would be unknown, 
@@
 in seasonal influenza, for example, the selection pressure
-exerted by the immune system was only limited \cite{han2019individual}
+exerted by the immune system was only limited (\cite{han2019individual}).
 
@@
 To determine the percentages of epitopes overlapping
-with TMHs, we use two reference proteomes.
-Our viral AA sequence is represented by
+with TMHs, we use three reference proteomes, which are
+a human, viral and bacterial one.
+The human reference proteome is \richel{here}, 
+which is the same as \cite{bianchi2017}.
+The viral reference proteome is represented by
 the proteome of the first sequenced COVID-19 strain (\cite{wu2020new},
 GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
-For bacteria, we use the reference genome of 
+For bacteria, the reference genome of 
 Mycobacterium tuberculosis (\url{https://www.ebi.ac.uk/reference_proteomes}, UP000001584, 
-83332 MYCTU).
+83332 MYCTU) is used.
 
-\paragraph{Data sets for evolutionary selection}
+\paragraph{Data set for evolutionary selection}
 
@@
 many time-dated proteomes.
-For humans, we use the 25000 proteomes used in influenza \richel{citation here}.
-For viruses, we use the many time-dated COVID-19 proteomes.
-We can assume that only after the moment that COVID-19 spilled over 
-to humans, it has mostly been the human immune system selecting against
-its detection, although this is only part of the selection for 
-having a high transmission rate.
-For bacteria, use the proteomes of Mycobacterium 
-\richel{Really? Pick better set, if possible!}.
+To do so, we use a viral pathogen, as these are detected by both
+MHC-I and MHC-II.
+To detect individual mutations, we need a high resolution dataset
+that also includes the evolutionary history of the virus.
+COVID-19 satisfies both requirements:
+as of 2020-06-22, it has 49790 proteomes, as provided by GISAID (\cite{shu2017gisaid}),
+as well as a provided evolutionary history by Nextstrain (\cite{hadfield2018nextstrain}).
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MHC-I}
+\subsection{Measuring TMH epitopes}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-To determine the percentages of MHC-I epitopes overlap
-with TMHs, we used the same 
-analysis as \cite{bianchi2017},
-except for using a viral and bacterial proteome (instead 
-of a human proteome, as used in the original study).
-Bianchi and colleagues obtained a distribution of 
-percentages of MHC-I epitopes overlap with TMHs in Homo sapiens
-for the different HLA haplotypes, with an average of 5.3\%.
-We obtained a similar distribution of percentages of MHC-I epitopes that 
-overlap with TMHs for the different HLA haplotypes, but then applied to
-our viral and bacterial proteome.
+To determine the percentages of MHC-I and MHC-II epitopes overlap
+with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
+To summarize: a proteome is converted into all its possible 9-mers. For each
+of these peptides, it is measured if it has been part of a TMH (at least
+1 residue), as well as if the peptide is a binder to an MHC-I haplotype.
+Because the percentage of TMH residues is known, a binomial test can be
+used to determine if TMH-derived peptides are as likely to be a binder
+as expected by chance (the null hypothesis), or that TMHs are indeed an
+overrepresented source of epitopes (as found by \cite{bianchi2017}).
 
-We compare the distributions of humans and each pathogen
-using a two-sample Kolmogorov-Smirnov (KS) test
-for a significance level $\alpha = 0.05$.
+This study differs in some aspects, described below in more detail,
+which is the definition of what a binder is,
+the addition of using MHC-II haplotypes, the software used,
+and the significance level to determine if TMH-derived peptides are as 
+likely to be a binder as expected by chance.
+These deviations are either a refinement of the previous method or
+a pragmatic choice made due to the extension of the original experiment.
+
+The definition of a binder differs from \cite{bianchi2017}:
+in this study a peptide is labeled an MHC-I binder if, within a haplotype, 
+any of its 9-mer peptides have an IC50 in the lowest 2\% of randomly simulated 
+peptides (see Supmat for values \richel{Put these values in supmat}), were the original study defined
+a binder as having an IC50 in the lowest 2\% of the peptides within a protein.
+% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
+
+The 13 MHC-I haplotypes used in this study are the same as 
+the ones in \cite{bianchi2017}.
+The MHC-II haplotypes used additionally are selected 
+to occur with a phenotypic frequency of at least 14\% in
+the human population (\cite{greenbaum2011functional},
+see \ref{table:mhc2_haplotypes} for coverage),
+resulting in 13 haplotypes.
+This frequency threshold is arbitrarily chosen to result in 
+the same number of haplotypes for each MHC class
+\richel{Add this to bbbq, there it is now 21}.
+When using an MHC-II haplotype, instead of using 9-mers, 13-mers are
+used, as these are the canonical MHC-II epitope size \richel{Reference here}.
+
+Instead of using the MHC-I-only epitope prediction 
+tool \verb;EpitopePrediction;, MHCnuggets is used, as it can predict
+both MHC-I and MHC-II epitopes.
+Additionally, instead of only using a human proteome, in this study
+also a viral and bacterial proteome are analyzed.
+
+Where the original study used a significance level of 1\% 
+for the binomial test to determine if TMH-derived peptides 
+are as likely to be a binder as expected by chance, this study
+uses 5\%, which is a common significance level for an exploratory study.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MHC-2}
+\subsection{Comparing amounts of TMH epitopes}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-To determine the percentages of MHC-II epitopes overlap, 
-we do the same analysis as MHC-I, except for
-MHC-II haplotypes. We picked the MHC-II alleles
-that occur with a frequency of at least 20\% in
-the human population (\cite{greenbaum2011functional},
-see \ref{table:mhc2_haplotypes} for coverage),
-resulting in \richel{approx 8} HLAs.
-The threshold of 20\% is arbitrarily, yet chosen to result in 5 and 10 HLA variants.  
-As \cite{bianchi2017}, we define the 5\% of peptides 
-with the lowest IC50 values as binders. We then counted the percentage
-of binders that have one or more amino acids that are part of a TMH.
+To determine if MHC-I has the same percentage of epitopes overlapping
+with viral/bacterial TMHs as with human TMHs \richel{$\mathcal{H}_{1,VB}$}, 
+we compare the distributions of the measured percentages 
+of TMH-overlapping peptides for all haplotypes,
+and compare the distributions between humans and each pathogen
+using a two-sample Kolmogorov-Smirnov (KS) test
+for a significance level $\alpha = 0.05$, 
+which is a common significance level for an exploratory study.
 
@@
 
-\paragraph{Gene prediction}
-
-Proteogenomics is the field dedicated to, among others,
-predict the proteins that arise from a DNA 
-sequence \cite{nesvizhskii2014proteogenomics}.
-For an ab initio gene prediction of a viral genome, 
-we searched the literature for the most recent software tool,
-which is Vgas (\cite{zhang2019vgas}). In this article, 
-Vgas is compared to tools, such as GeneMarkS (\cite{besemer2001genemarks}),
-Prodigal (\cite{hyatt2012gene}), GLIMMER (\cite{delcher1999improved}),
-RAST \cite{aziz2008rast} and prokka (\cite{seemann2014prokka}).
-We picked the FOSS tool that scored best in the (closed source) Vgas
-benchmark (\cite{zhang2019vgas}), which is Prodigal.
+% \paragraph{Gene prediction}
+% 
+% Proteogenomics is the field dedicated to, among others,
+% predict the proteins that arise from a DNA 
+% sequence \cite{nesvizhskii2014proteogenomics}.
+% For an ab initio gene prediction of a viral genome, 
+% we searched the literature for the most recent software tool,
+% which is Vgas (\cite{zhang2019vgas}). In this article, 
+% Vgas is compared to tools, such as GeneMarkS (\cite{besemer2001genemarks}),
+% Prodigal (\cite{hyatt2012gene}), GLIMMER (\cite{delcher1999improved}),
+% RAST \cite{aziz2008rast} and prokka (\cite{seemann2014prokka}).
+% We picked the FOSS tool that scored best in the (closed source) Vgas
+% benchmark (\cite{zhang2019vgas}), which is Prodigal.
 
@@
 
-Because we use 13 MHC-I and 7 MHC-II haplotypes
-for a significance level of 5\%, by chance we expect (5\% of $13 + 7$)
-one times to find a (MHC-I or MHC-II) haplotype 
-that is suggested to have selection working upon it.
-Applying a binomial distribution ($p = 0.05$, $n = 13+7$, $\alpha = 0.05$),
-we will claim we there is some form of selection when we find 3
-or more examples of having a significant result.
-However, because there are 2 hypotheses tested per haplotype,
-we will claim there is directional selection, 
-when, for a same hypothesis, we find 2 examples of having a significant result.
+Because we use 13 haplotypes per MHC class,
+for a significance level of 5\%, 
+there is a 14% chance to find one haplotype 
+that hints to have selection working upon it.
+% > pbinom(q = 1, size = 13, prob = 0.05, lower.tail = FALSE)
+% [1] 0.1354239
 
-% > pbinom(q = 2, size = 20, prob = 0.05, lower.tail = FALSE)
-% [1] 0.07548367
-%
-% pbinom(q = 3, size = 20, prob = 0.05, lower.tail = FALSE)
-% [1] 0.01590153
+The chance to find 2 haplotypes that hint to have selection working upon it
+is 2\%, which is below our significance level.
+% > pbinom(q = 2, size = 13, prob = 0.05, lower.tail = FALSE)
+% [1] 0.02450784
+
+Therefore, we will claim there is directional selection, 
+when, for a proteome, we find 2 haplotypes 
+that hint to have selection working upon it
 

### 2020-06-23 ic50s_per_haplotype/ic50s_per_haplotype.R
Create table

@@ -1,20 +1,27 @@
 library(bbbq)
 library(ggplot2, quietly = TRUE)
+library(dplyr)
 library(testthat)
 
 t <- tidyr::expand_grid(
   official_name = bbbq::get_mhc_haplotypes(),
   mhc_class = NA,
   mhcnuggets_name = NA,
-  peptide_length = c(9, 10),
+  peptide_length = NA,
   filename = NA,
   n_peptides = NA,
   percentile = 0.02,
   ic50 = NA
 )
 for (i in seq_len(nrow(t))) {
-  if (t$official_name[i] %in% bbbq::get_mhc1_haplotypes()) t$mhc_class[i] <- "I"
-  if (t$official_name[i] %in% bbbq::get_mhc2_haplotypes()) t$mhc_class[i] <- "II"
+  if (t$official_name[i] %in% bbbq::get_mhc1_haplotypes()) {
+    t$mhc_class[i] <- "I"
+    t$peptide_length[i] <- 9
+  }
+  if (t$official_name[i] %in% bbbq::get_mhc2_haplotypes()) {
+    t$mhc_class[i] <- "II"
+    t$peptide_length[i] <- 10
+  }
 }
 for (i in seq_len(nrow(t))) {
   t$mhcnuggets_name[i] <- mhcnuggetsr::to_mhcnuggets_name(t$official_name[i])
@@ -52,6 +59,18 @@ ggplot(t, aes(x = official_name, y = ic50, color = peptide_length)) +
   scale_y_continuous(name = "IC50") +
   scale_x_discrete(name = "Haplotype") +
   labs(
-    caption = glue::glue("Percentile: {percentile}, # peptides per IC50: {n_peptides}")
+    caption = glue::glue(
+      "Percentile: {percentile}, # peptides per IC50: {n_peptides}",
+      " , TODO: use 13 peptides for MHC-II"
+    )
   ) +
   ggsave("ic50s_per_haplotype.png", width = 7, height = 7)
+
+
+
+readr::write_csv(
+  x = t %>%
+  select(official_name, ic50) %>%
+  rename(haplotype = official_name),
+  "ic50s_per_haplotype.csv"
+)

### 2020-06-23 bbbq_article.tex
Show IC50s per haplotype

@@
 any of its 9-mer peptides have an IC50 in the lowest 2\% of randomly simulated 
-peptides (see Supmat for values \richel{Put these values in supmat}), were the original study defined
+peptides (see table \ref{table:ic50s_per_haplotype} for values), were the original study defined
 a binder as having an IC50 in the lowest 2\% of the peptides within a protein.
@@
 
+% Chance to find
+%            this amount of positives (sums to 1)
+%              |
+%              v
+% > dbinom(x = 0, size = 26, prob = 0.05)
+% [1] 0.2635201
+% > dbinom(x = 1, size = 26, prob = 0.05)
+% [1] 0.3606064
+% > dbinom(x = 2, size = 26, prob = 0.05)
+% [1] 0.2372411
+% > dbinom(x = 3, size = 26, prob = 0.05)
+% [1] 0.09989098
+% > dbinom(x = 4, size = 26, prob = 0.05)
+% [1] 0.03023017
+% > dbinom(x = 5, size = 26, prob = 0.05)
+% [1] 0.00700067
+%
+% Chance to find more than
+%            this amount of positives (excluding the value!)
+%              |
+%              v
+% > pbinom(q = 0, size = 26, prob = 0.05, lower.tail = FALSE)
+% [1] 0.7364799
+% > pbinom(q = 1, size = 26, prob = 0.05, lower.tail = FALSE)
+% [1] 0.3758735
+% > pbinom(q = 2, size = 26, prob = 0.05, lower.tail = FALSE)
+% [1] 0.1386324
+% > pbinom(q = 3, size = 26, prob = 0.05, lower.tail = FALSE)
+% [1] 0.0387414
+% > pbinom(q = 4, size = 26, prob = 0.05, lower.tail = FALSE)
+% [1] 0.008511231
+
 Because we use 13 haplotypes per MHC class,
 for a significance level of 5\%,
-there is a 14% chance to find one haplotype 
-that hints to have selection working upon it.
-% > pbinom(q = 1, size = 13, prob = 0.05, lower.tail = FALSE)
-% [1] 0.1354239
-
+for both directions (D to U, or U to D)
+we expect 5% times 26 equals 1.3 significant findings.
+
+\richel{Hmm, improve}
 The chance to find 2 haplotypes that hint to have selection working upon it
-is 2\%, which is below our significance level.
-% > pbinom(q = 2, size = 13, prob = 0.05, lower.tail = FALSE)
-% [1] 0.02450784
+is 4\%, which is below our significance level.
+% > pbinom(q = 3, size = 26, prob = 0.05, lower.tail = FALSE)
+% [1] 0.0387414
+
+Upon finding 3 haplotypes that hint to have selection working upon it,
+there is the additional chance that selection is working in different
+directions (i.e. in one case there is selection for losing detection,
+where in the other case, there is (unexpected) selection for gaining
+detection). Therefore, finding 3 haplotypes that hint to have selection working 
+upon it in the same direction
 
@@
 two possible detection state a peptide is in, which is
-either undetected $U$, or detected $D$ by the immune system.
+either undetected $U$, or detected $D$ by the immune system,
+where a detected peptide is a binder, that is, 
+a peptide with an IC50 in the lowest 2\%.
 The $D$ can be either detection by an MHC-I or MHC-II haplotype.
@@
 
-Of the complete system, we test for each of the 2 state transitions (that is,
-we do not test for mutations that result in the same state).
+Of the complete system, we test for each of the 2 observed 
+state transition rates (that is, we do not test for mutations that result in the same state)
+if these are likely to be caused by chance.
 For each transition rate, we know the expected and observed number
@@
 We use a significance level $\alpha$ of 5\%, because we have no prior
-evidence.
-
-As we are testing 2 hypotheses, 
+expectations.
+
+As we are testing 2 rates (and thus 2 hypotheses), 
 we correct against type I errors (false positives), 
-by applying the Holm-Bonferroni correction, 
-resulting in \richel{here is just a normal Bonferroni correction, for now}
-$\alpha_c = \frac{\alpha}{2}$.
+by applying the Bonferroni correction, 
+resulting in $\alpha_c = \frac{\alpha}{2}$.
 If the probability of observing that few/many 
@@
   \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
-  Predicted IC50s (in nM) per MHC-I haplotype
+  Predicted IC50s (in nM) per haplotype
   \caption{
@@
 \end{figure}
+
+\begin{table}[!htbp]
+  \input{ic50s_per_haplotype/ic50s_per_haplotype.latex}
+  \caption{
+    Predicted IC50s (in nM) per haplotype
+    \richel{Used randomly simulated peptides, results are real}
+  }
+  \label{table:ic50s_per_haplotype}
+\end{table}
 

### 2020-06-24 bbbq_article.tex
Move prediction software to Methods

@@
 
-\paragraph{MHC-I epitope prediction}
-
-For MHC-I, there are multiple computational tools developed 
-to predict epitopes. 
-According to \cite{lundegaard2011prediction}, in 2011,
-from a set of multiple tools, 
-NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
-A tool developed later is \verb;epitope-prediction; \cite{bianchi2017},
-which uses a stabilized matrix method \cite{kim2009derivation}.
-The most recent FOSS tool available is MHCnuggets\cite{shao2020high},
-which we will use in this study
-
-\paragraph{MHC-II epitope prediction}
-
-Also for MHC-II, there are multiple computational tools developed 
-to predict epitopes,
-such as using a trained neural network (\cite{nielsen2003reliable})
-or a Gibbs sampling approach (\cite{nielsen2004improved}).
-According to \cite{lundegaard2011prediction}, in 2011,
-from a set of multiple tools, 
-NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
-gave rise to the most accurate predictions.
-Later tools are \cite{zhang2013predicting},
-\cite{trolle2014nettepi}, \cite{zhang2015accurate}
-and the very recent MHCnuggets (\cite{shao2020high}).
-In this study, we'll be using MHCnuggets, as it is the only FOSS tool available.
-
 \paragraph{TMHs}
@@
 
-There are multiple computational tools developed to detect which
+There are multiple computational tools developed to predict which
 parts of membrane proteins are TMH
-In 2001, multiple tools to do so have been compared \cite{moller2001evaluation},
-of which TMHMM \cite{krogh2001predicting} turned out to be the best.
-Many other tools followed, such as 
-Phobius (\cite{kall2004combined}),
-ConPred II (\cite{arai2004conpred}),
-MEMSAT3 (\cite{jones2007improving})
-and MetaTM (\cite{klammer2009metatm}), which signals the importance
-of TMH predictions. 
-Contemporary tools are the closed-source MemBrain (\cite{feng2020topology}) 
-and the FOSS tool PureseqTM (\cite{wang2019efficient}).
-In this study, we will use PureseqTM.
 
@@
 bacterial TMHs.
-
-% \paragraph{Gene prediction}
-% 
-% Proteogenomics is the field dedicated to, among others,
-% predict the proteins that arise from a DNA 
-% sequence \cite{nesvizhskii2014proteogenomics}.
-% For an ab initio gene prediction of a viral genome, 
-% we searched the literature for the most recent software tool,
-% which is Vgas (\cite{zhang2019vgas}). In this article, 
-% Vgas is compared to tools, such as GeneMarkS (\cite{besemer2001genemarks}),
-% Prodigal (\cite{hyatt2012gene}), GLIMMER (\cite{delcher1999improved}),
-% RAST \cite{aziz2008rast} and prokka (\cite{seemann2014prokka}).
-% We picked the FOSS tool that scored best in the (closed source) Vgas
-% benchmark (\cite{zhang2019vgas}), which is Prodigal.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Prediction software used}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For this research, the scientific literature was explored 
+to find the most recent free and open source (FOSS) prediction software.
+This was done by searching for papers that (1) reference older
+prediction software, and (2) present a novel method to make predictions.
+As a starting point, a review paper was used.
+
+\paragraph{TMH prediction}
+
+There are multiple computational tools developed to predict which
+parts of membrane proteins are TMH
+In 2001, multiple tools to do so have been compared \cite{moller2001evaluation},
+of which TMHMM \cite{krogh2001predicting} turned out to be the best, yet
+this software has a restrictive software license and is nearly two
+decades old. Inspired on TMHMM, many other tools followed, such as 
+Phobius (\cite{kall2004combined}),
+ConPred II (\cite{arai2004conpred}),
+MEMSAT3 (\cite{jones2007improving})
+and MetaTM (\cite{klammer2009metatm}).
+The most recent FOSS tool is called PureseqTM (\cite{wang2019efficient}),
+which is used in this study.
+
+\paragraph{MHC-I epitope prediction}
+
+For MHC-I, there are multiple computational tools developed 
+to predict epitopes. 
+According to \cite{lundegaard2011prediction}, in 2011,
+from a set of multiple tools, 
+NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
+A tool developed later is \verb;epitope-prediction; \cite{bianchi2017},
+which uses a stabilized matrix method \cite{kim2009derivation}.
+The most recent FOSS tool available is MHCnuggets\cite{shao2020high},
+which we will use in this study
+
+\paragraph{MHC-II epitope prediction}
+
+Also for MHC-II, there are multiple computational tools developed 
+to predict epitopes,
+such as using a trained neural network (\cite{nielsen2003reliable})
+or a Gibbs sampling approach (\cite{nielsen2004improved}).
+According to \cite{lundegaard2011prediction}, in 2011,
+from a set of multiple tools, 
+NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
+gave rise to the most accurate predictions.
+Later tools are \cite{zhang2013predicting},
+\cite{trolle2014nettepi}, \cite{zhang2015accurate}
+and the very recent MHCnuggets (\cite{shao2020high}),
+the latter which we will use in this study.
+
+% \paragraph{Gene prediction}
+% 
+% Proteogenomics is the field dedicated to, among others,
+% predict the proteins that arise from a DNA 
+% sequence \cite{nesvizhskii2014proteogenomics}.
+% For an ab initio gene prediction of a viral genome, 
+% we searched the literature for the most recent software tool,
+% which is Vgas (\cite{zhang2019vgas}). In this article, 
+% Vgas is compared to tools, such as GeneMarkS (\cite{besemer2001genemarks}),
+% Prodigal (\cite{hyatt2012gene}), GLIMMER (\cite{delcher1999improved}),
+% RAST \cite{aziz2008rast} and prokka (\cite{seemann2014prokka}).
+% We picked the FOSS tool that scored best in the (closed source) Vgas
+% benchmark (\cite{zhang2019vgas}), which is Prodigal.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Prediction software written}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The R programming language is used for the complete 
+experiment, including the analysis.
+The complete experiment is bundled in the 'bbbq' R package,
+which is dependent on 'pureseqtmr', 'mhcnuggetsr' and 'mhcnpreds',
+as described below.
+
+To be able to call, from R, the TMH prediction software 'PureseqTM',
+which is written in C, the package 'pureseqtmr' has been developed. 
+'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
+Excluded are the features that are used by the 'PureseqTM' 
+developers to verify the correctness of their work.
+'pureseqtmr' has been submitted to CRAN.
+
+The R package 'mhcnuggetsr' is written to call the 
+'MHCnuggets' Python package.
+'mhcnuggetsr' allows to install 'MHCnuggets' and call it
+to predict IC50 values for multiple MHC haplotypes that 'MHCnuggets' 
+has already been trained upon.
+Excluded is the feature to train 'MHCnuggets' on new data and/or other
+haplotypes, as this study only needs IC50 predictions.
+'mhcnuggetsr' has been submitted to CRAN.
+
+To be able to use pre-calculated MHCnuggets
+predictions (i.e. a distribution of IC50 values per haplotype, 
+derived from random peptides), the 'mhcnpreds' R package has been 
+developed. In practice, this is nothing more than a big look-up table.
+Because the package is too big, it will not been submitted to CRAN.
+
+To reproduce the full experiment as done in this paper,
+the functions needed are bundled in the 'bbbq' R package.
+This package is too specific to be submitted to CRAN.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}

### 2020-06-24 bbbq_article.tex
EOD

@@
 Due to this, software was written that allows for silico 
-predictions, which are reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+predictions, which are reliable in 
+practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
 
@@
 making up for \richel{?} percent of its entire genome.
-
-\paragraph{TMH prediction}
-
 There are multiple computational tools developed to predict which
-parts of membrane proteins are TMH
+parts of membrane proteins are TMH.
 
@@
 
-\paragraph{Evolutionary signal of unknown strength for virus}
-
-Regarding human viruses, we do not know the selection pressure
-exerted by the human immune system.
-For viral TMHs that are undetected by most human haplotypes, we
-expect these to be conserved, to remain avoiding detection.
-For viral TMHs that are detected by most human haplotypes, we
-expect there to be a higher mutation rate, to avoid detection.
-Additionally, viruses are not only selected for their evasion of an 
-immune response, yet are selected to have a high reproductive value.
-Because the strength of the evolutionary signal is unknown,
-we use a strong data set to be able to detect it.
-
-\paragraph{Evolutionary signal of unknown strength for bacteria}
-
-Regarding human pathogenic bacteria, we do not know the extent to which the 
-human immune systems selects on these, nor which loci are selected
-upon \richel{so I should read the literature on that}.
-Because most bacterial pathogens are generalists, that is,
-can infect multiple hosts, we expect the selection pressure exerted
-by the human immune response to be weak.
-For bacterial TMHs that are undetected by most human haplotypes, we
-expect these to be conserved, to remain avoiding detection.
-For bacterial TMHs that are detected by most human haplotypes, we
-expect there to be a higher mutation rate, to avoid detection.
-Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
-Due to this, the strength of the evolutionary signal is unknown,
-thus we need a strong data set to be able to detect it.
+\paragraph{Evolutionary signal of unknown strength for pathogens}
+
+Up until today, it is unknown if an evolutionary signal 
+is present in a (human) pathogen's proteome that stems from avoiding
+detection by the human immune system.
+If such a signal is present, we expect to detect it most likely in
+viruses, as both MHC-I and MHC-II are used in protecting the
+human host. 
+Therefore, this study focuses in detecting this signal in viruses.
 
@@
     with human/viral/bacterial TMHs as expected by chance
-  \item $\mathcal{H}_{3,HVB}$: The mutations observed in 
-    human/viral/bacterial TMHs are caused by chance.
+  \item $\mathcal{H}_{3}$: The mutations observed in 
+    viral TMHs are caused by chance.
     Alternatively, these are caused by natural selection
-    induced by binding to HLAs
+    induced by (non-)binding to HLAs
 \end{itemize}

### 2020-06-24 p_bind_per_hydrophobicity/create_figure_hydrophobicity_vs_is_binder.R
Can make figures

@@ -1,0 +1,71 @@
+# Calculate the relation between hydrophobicity
+# and chance to bind to MHC-I.
+#
+# H0: there is no relation
+# H1: higher hydrophobicity equals higher chance to bind
+#
+library(ggplot2, quietly = TRUE)
+library(dplyr, quietly = TRUE)
+library(testthat, quietly = TRUE)
+library(bbbq, quietly = TRUE)
+
+raw_data <- "dataset.csv"
+expect_true(file.exists(raw_data))
+df <- readr::read_csv(raw_data)
+
+plot_hydrophobicity_vs_is_binder <- function(
+  df,
+  png_filename = tempfile(fileext = ".png")
+) {
+  mhc_class <- NA
+  if (df$mhc_haplotype[1] %in% bbbq::get_mhc1_haplotypes()) {
+    mhc_class <- "I"
+    testthat::expect_true(
+      all(df$mhc_haplotype %in% bbbq::get_mhc1_haplotypes())
+    )
+  }
+  if (df$mhc_haplotype[1] %in% bbbq::get_mhc2_haplotypes()) {
+    mhc_class <- "II"
+    testthat::expect_true(
+      all(df$mhc_haplotype %in% bbbq::get_mhc2_haplotypes())
+    )
+  }
+  testthat::expect_true(!is.na(mhc_class))
+
+  n_haplotypes <- length(unique(df$mhc_haplotype))
+  n_peptides <- length(unique(df$sequence))
+
+  df <- tibble::as_tibble(df)
+  df_tally <- df %>%
+    dplyr::mutate(bin = trunc(hydrophobicity * 2)) %>%
+    dplyr::group_by(bin, is_binder) %>%
+    dplyr::tally()
+  ggplot2::ggplot(df_tally, ggplot2::aes(bin, is_binder, fill = n)) +
+    ggplot2::geom_tile() +
+    ggplot2::geom_text(ggplot2::aes(label = n)) +
+      ggplot2::scale_fill_gradient(low = "white", high = "red") +
+      ggplot2::scale_y_discrete() +
+      ggplot2::xlab("Hydrophobicity x2") +
+      ggplot2::ylab(
+        glue::glue("Binds to MHC-{mhc_class}? Down = no, up = yes")
+      ) +
+      ggplot2::labs(
+        caption = glue::glue(
+          "# peptides: {n_peptides}, # haplotypes: {n_haplotypes}"
[…]

### 2020-06-24 p_bind_per_hydrophobicity/create_figure_is_tmh_vs_binds_mhc.R

@@ -9,6 +9,10 @@ library(dplyr, quietly = TRUE)
 library(testthat, quietly = TRUE)
 library(bbbq, quietly = TRUE)
 
+raw_data <- "dataset.csv"
+expect_true(file.exists(raw_data))
+df <- readr::read_csv(raw_data)
+
 plot_is_tmh_vs_is_binder <- function(
   df,
   png_filename = tempfile(fileext = ".png")
@@ -135,32 +139,15 @@ plot_hydrophobicity_vs_is_binder <- function(
       ggplot2::ggsave(png_filename, width = 7, height = 7)
 }
 
-raw_data <- "dataset.csv"
-expect_true(file.exists(raw_data))
-df <- readr::read_csv(raw_data)
-
 df_1 <- df %>% filter(mhc_haplotype %in% get_mhc1_haplotypes())
 df_2 <- df %>% filter(mhc_haplotype %in% get_mhc2_haplotypes())
 
 plot_is_tmh_vs_is_binder(
   df = df_1,
-  png_filename = "is_tmh_vs_is_binder_1.png"
+  png_filename = "is_tmh_vs_binds_mhc1.png"
 )
 plot_is_tmh_vs_is_binder(
   df = df_2,
-  png_filename = "is_tmh_vs_is_binder_2.png"
-)
-# plot_is_binder_vs_is_binder(df = df, png_filename = "is_binder_vs_is_binder.png")
-plot_hydrophobicity_vs_is_tmh(
-  df = df, png_filename = "hydrophobicity_vs_is_tmh.png"
-)
-
-plot_hydrophobicity_vs_is_binder(
-  df = df_1,
-  png_filename = "hydrophobicity_vs_is_binder_1.png"
-)
-plot_hydrophobicity_vs_is_binder(
-  df = df_2,
-  png_filename = "hydrophobicity_vs_is_binder_2.png"
+  png_filename = "is_tmh_vs_binds_mhc2.png"
 )
 

### 2020-07-16 covid_dataset_strength/covid_dataset_strength.Rmd
Taking unique sequences into account, the dataset is weak

@@
 Abbr  |Abbreviation
+Abbrs |Abbreviations
 TMH   |Transmembrane helix
@@
 topology <- pureseqtmr::predict_topology(fasta_filename)
+```
+
+Plot it:
+
+```{r}
+pureseqtmr::plot_topology(topology)
+```
+
+Extract data from the topology:
+
+```{r}
 n_proteins <- nrow(topology)
@@
 There are `r n_sequences` COVID-19 sequences (`https://www.gisaid.org/`).
+Not all of these are unique.
+We will determine the number of unique sequences
+from the percentage of unique sequence per protein.
+
+To do so, we read all COVID-19 sequences:
+
+```{r cache=TRUE}
+all_sequences_filename <- "../downloads/allprot0621.fasta"
+testthat::expect_true(file.exists(all_sequences_filename))
+sequences <- seqinr::read.fasta(
+  all_sequences_filename, 
+  seqtype = "AA",
+  as.string = TRUE
+)
+```
+
+Grouping them by protein:
 
 ```{r}
-n_tmh_aa_mutations <- tmh_aa_evolution_rate_per_sequence * n_sequences
+t <- tibble::tibble(
+  prot_name = as.character(
+    na.omit(
+      stringr::str_match(
+        names(sequences),
+        "^([:upper:][:alnum:]{0,5})\\|"
+      )[, 2]
+    )
+  ),
+  sequence = as.character(sequences)
+)
+library(dplyr)
+t$prot_name <- as.factor(t$prot_name)
+t_unique <- t %>% group_by(prot_name) %>% 
+  summarise(n = n(), n_distinct = n_distinct(sequence)) %>%
+  mutate(perc = n_distinct / n)
+perc_unique <- mean(t_unique$perc)
+```
+
+Of all proteins, on average, `r perc_unique` are unique.
+
+```{r}
+n_unique_sequences <- perc_unique * n_sequences
+```
+
+Therefor, there are `r n_unique_sequences` unique sequences.
+
+```{r}
+n_tmh_aa_mutations <- tmh_aa_evolution_rate_per_sequence * n_unique_sequences
 ```

### 2020-07-28 bbbq_article.tex
EOD

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evolutionary conservation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{transitions.png}
-  \caption{
-    Transitions in TMHs that occur more often then expected by chance,
-    for a confidence interval of 5 percent.
-    Table \ref{tab:transitions} shows the transition counts.
-    \richel{
-      stub figure
-    }
-  }
-  \label{fig:transitions}
-\end{figure}
-
-\begin{table}[!htbp]
-  \input{table_transitions.latex}
-  \caption{
-    Transitions counts, where the row indicates the source state,
-    and the column indicates the target state.
-    First number per cell is the observed number of this state transition,
-    where the second number is the expected number of this state transition
-    as predicted by chance.
-    An asterisk behind the observed count indicates that this count
-    is unlikely to be caused by chance only.
-    Figure \ref{fig:transitions} shows which transition counts are 
-    unlikely to be caused by chance only.
-  }
-  \label{tab:transitions}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Differences with Bianchi et al., 2017}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A part of this study does the same analysis as Bianchi et al., 2017.
+Here we describe the deviations, which are about the use of different
+software and the use of a different definition of what a binder is.
+
+\paragraph{TMH prediction software}
+
+The earlier study uses TMHMM, which has a restricted software license.
+Instead, we chose to use PureseqTM, which is free and open source.
+
+\paragraph{IC50 prediction software}
+
+The earlier study uses a hand-crafted method, that only works on MHC-I.
+Instead, we use MHCnuggets, which (1) free and open source, and (2) predicts
+IC50s on both MHC-I and MHC-II.
+
+\paragraph{Definition of what a binder is}
+
+The earlier study defines a peptide a binder (for a haplotype), 
+if \emp{within the peptide} in which it is found, 
+is within the peptides with the 2\% lowest IC50 values.
+This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
+where the binders are written to file.
+
+In this study, a peptide is defined a binder (for a haplotype), 
+if \emp{within a proteome} in which it is found, 
+is within the peptides with the 2\% lowest IC50 values.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-I}

### 2020-08-05 bbbq_article.tex
Show tables

@@
 
-\begin{table}[!htbp]
-  \input{bbbq_1/bbbq_1_percentages.latex}
-  \caption{
-    Percentage of MHC-I epitopes overlapping with transmembrane helix.
-    \richel{This is simulated data}
-  }
-  \label{table:bbbq_1_percentages}
-\end{table}
-
-\begin{table}[!htbp]
-  \input{bbbq_1/bbbq_1_stats_covid.latex}
-  \caption{
-    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-I
-    \richel{Done on the simulated data}
-  }
-  \label{table:bbbq_1_stats_covid}
-\end{table}
-
-\begin{table}[!htbp]
-  \input{bbbq_1/bbbq_1_stats_myco.latex}
-  \caption{
-    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-I
-    \richel{Done on the simulated data}
-  }
-  \label{table:bbbq_1_stats_myco}
-\end{table}
+\input{bbbq_1/table_tmh_binders_mhc1.latex}
+
+%\begin{table}[!htbp]
+%   \input{bbbq_1/bbbq_1_stats_covid.latex}
+%   \caption{
+%     Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-I
+%     \richel{Done on the simulated data}
+%   }
+%   \label{table:bbbq_1_stats_covid}
+% \end{table}
+
+% \begin{table}[!htbp]
+%   \input{bbbq_1/bbbq_1_stats_myco.latex}
+%   \caption{
+%     Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-I
+%     \richel{Done on the simulated data}
+%   }
+%   \label{table:bbbq_1_stats_myco}
+% \end{table}
 
@@
 
-\begin{table}[!htbp]
-  \input{bbbq_2/bbbq_2_percentages.latex}
-  \caption{
-    Percentage of MHC-II epitopes overlapping with transmembrane helix.
-    \richel{This is simulated data}
-  }
-  \label{table:bbbq_2_percentages}
-\end{table}
-
-\begin{table}[!htbp]
-  \input{bbbq_2/bbbq_2_stats_covid.latex}
-  \caption{
-    Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-II
-    \richel{Done on the simulated data}
-  }
-  \label{table:bbbq_2_stats_covid}
-\end{table}
-
-\begin{table}[!htbp]
-  \input{bbbq_2/bbbq_2_stats_myco.latex}
-  \caption{
-    Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-II
-    \richel{Done on the simulated data}
-  }
-  \label{table:bbbq_2_stats_myco}
-\end{table}
+\input{bbbq_1/table_tmh_binders_mhc2.latex}
+
+% \begin{table}[!htbp]
+%   \input{bbbq_2/bbbq_2_stats_covid.latex}
+%   \caption{
+%     Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-II
+%     \richel{Done on the simulated data}
+%   }
+%   \label{table:bbbq_2_stats_covid}
+% \end{table}
+
+% \begin{table}[!htbp]
+%   \input{bbbq_2/bbbq_2_stats_myco.latex}
+%   \caption{
+%     Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-II
+%     \richel{Done on the simulated data}
+%   }
+%   \label{table:bbbq_2_stats_myco}
+% \end{table}
 

### 2020-08-06 bbbq_article.tex
Show figure
Update names

@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1/fig_bbbq_1.png}
+  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc1.png}
   \caption{
@@
     for a human, viral and bacterial proteome.
-    \richel{
-      the data underlying this figure has been simulated
-    }
   }
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/fig_bbbq_2.png}
+  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc2.png}
   \caption{
@@
     for a human, COVID-19 and Mycobacterium proteome.
-    \richel{
-      the data underlying this figure has been simulated
-    }
   }
@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1/fig_ic50_distribution.png}
+  \caption{
+    Distribution of IC50 estimations per haplotype per target.
+    The IC50 value is plotted on a logarithmic scale.
+  }
+  \label{fig:1}
+\end{figure}
+
 \paragraph{Definition of what a binder is}
@@
 is within the peptides with the 2\% lowest IC50 values.
+
+\input{bbbq_1/table_ic50_binders.latex}
 

### 2020-08-06 ic50s_per_haplotype/ic50s_per_haplotype.R

@@ -10,7 +10,7 @@ t <- tidyr::expand_grid(
   peptide_length = NA,
   filename = NA,
   n_peptides = NA,
-  percentile = 0.02,
+  percentile = bbbq::get_ic50_percentile_binder(),
   ic50 = NA
 )
 for (i in seq_len(nrow(t))) {

### 2020-08-06 p_bind_per_hydrophobicity/create_is_binder.R

@@ -19,7 +19,7 @@ if (!peregrine::is_on_peregrine()) {
 t_binds <- tidyr::expand_grid(
   sequence = t$sequence,
   mhc_haplotype = bbbq::get_mhc_haplotypes(),
-  percentile = 0.02,
+  percentile = bbbq::get_ic50_percentile_binder(),
   is_binder = NA
 )
 

### 2020-08-31 bbbq_article.tex
Show percentage TMH expected, fix #39
Remove tedious evolution experiment

@@
     with human/viral/bacterial TMHs as expected by chance
-  \item $\mathcal{H}_{3}$: The mutations observed in 
-    viral TMHs are caused by chance.
-    Alternatively, these are caused by natural selection
-    induced by (non-)binding to HLAs
+  \item $\mathcal{H}_{3}$: TMH and non-TMH regions are equally conserved
+    in a proteome
 \end{itemize}
@@
 a human, viral and bacterial one.
-The human reference proteome is \richel{here}, 
+For a human reference proteome, we used the Uniprot human reference proteome
+from \url{https://www.uniprot.org/proteomes/UP000005640}), 
 which is the same as \cite{bianchi2017}.
@@
 GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
-For bacteria, the reference genome of 
-Mycobacterium tuberculosis (\url{https://www.ebi.ac.uk/reference_proteomes}, UP000001584, 
-83332 MYCTU) is used.
-
-\paragraph{Data set for evolutionary selection}
-
-To detect an evolutionary signal for selection on binding
-TMHs peptides, we need a phylogenetic tree with 
-many time-dated proteomes.
-To do so, we use a viral pathogen, as these are detected by both
-MHC-I and MHC-II.
-To detect individual mutations, we need a high resolution dataset
-that also includes the evolutionary history of the virus.
-COVID-19 satisfies both requirements:
-as of 2020-06-22, it has 49790 proteomes, as provided by GISAID (\cite{shu2017gisaid}),
-as well as a provided evolutionary history by Nextstrain (\cite{hadfield2018nextstrain}).
+For bacteria, we used the Uniprot Mycobacterium tuberculosis reference genome
+(\url{https://www.uniprot.org/proteomes/UP000001584}) is used.
+
+\paragraph{Data set for evolutionary conservation of TMHs}
+
+To detect the evolutionary conservation of TMHs,
+we use the 49790 (as of 2020-06-22) fully sequenced SARS-CoV-2 proteomes, 
+as provided by GISAID (\cite{shu2017gisaid}).
 
@@
 a pragmatic choice made due to the extension of the original experiment.
+Additionally, instead of only using a human proteome, in this study
+also a viral and bacterial proteome are analyzed.
 
 The definition of a binder differs from \cite{bianchi2017}:
-in this study a peptide is labeled an MHC-I binder if, within a haplotype, 
-any of its 9-mer peptides have an IC50 in the lowest 2\% of randomly simulated 
-peptides (see table \ref{table:ic50s_per_haplotype} for values), were the original study defined
+in this study a peptide is called a binder if, within a haplotype, 
+any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
+one thousand randomly simulated 
+peptides (see table \ref{table:ic50s_per_haplotype} for values), 
+whereas the original study defined
 a binder as having an IC50 in the lowest 2\% of the peptides within a protein.
@@
 The MHC-II haplotypes used additionally are selected 
-to occur with a phenotypic frequency of at least 14\% in
+to occur with a phenotypic frequency of at least 
+\richel{recalculate number}
+14\% in
 the human population (\cite{greenbaum2011functional},
 see \ref{table:mhc2_haplotypes} for coverage),
-resulting in 13 haplotypes.
-This frequency threshold is arbitrarily chosen to result in 
-the same number of haplotypes for each MHC class
-\richel{Add this to bbbq, there it is now 21}.
+resulting in 21 haplotypes.
 When using an MHC-II haplotype, instead of using 9-mers, 13-mers are
 used, as these are the canonical MHC-II epitope size \richel{Reference here}.
-
-Instead of using the MHC-I-only epitope prediction 
-tool \verb;EpitopePrediction;, MHCnuggets is used, as it can predict
-both MHC-I and MHC-II epitopes.
-Additionally, instead of only using a human proteome, in this study
-also a viral and bacterial proteome are analyzed.
 
@@
 uses 5\%, which is a common significance level for an exploratory study.
+
+Instead of using the membrane protein topology prediction
+tool \verb;TMHMM; \cite{krogh2001predicting}, PureseqTM \cite{wang2019efficient}
+is used instead.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Detecting evolution}
+\subsubsection{Evolutionary conservation of TMHs}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To detect evolution, we determine the probability 
-that all of the observed TMH mutations have arisen by chance,
-with regards to being detected/presented by the immune system,
-per MHC class and haplotype. We do this for human, viral and
-bacterial TMHs.
-
-\paragraph{Correction against many tests}
-
-% Chance to find
-%            this amount of positives (sums to 1)
-%              |
-%              v
-% > dbinom(x = 0, size = 26, prob = 0.05)
-% [1] 0.2635201
-% > dbinom(x = 1, size = 26, prob = 0.05)
-% [1] 0.3606064
-% > dbinom(x = 2, size = 26, prob = 0.05)
-% [1] 0.2372411
-% > dbinom(x = 3, size = 26, prob = 0.05)
-% [1] 0.09989098
-% > dbinom(x = 4, size = 26, prob = 0.05)
-% [1] 0.03023017
-% > dbinom(x = 5, size = 26, prob = 0.05)
-% [1] 0.00700067
-%
-% Chance to find more than
-%            this amount of positives (excluding the value!)
-%              |
-%              v
-% > pbinom(q = 0, size = 26, prob = 0.05, lower.tail = FALSE)
-% [1] 0.7364799
-% > pbinom(q = 1, size = 26, prob = 0.05, lower.tail = FALSE)
-% [1] 0.3758735
-% > pbinom(q = 2, size = 26, prob = 0.05, lower.tail = FALSE)
-% [1] 0.1386324
-% > pbinom(q = 3, size = 26, prob = 0.05, lower.tail = FALSE)
-% [1] 0.0387414
-% > pbinom(q = 4, size = 26, prob = 0.05, lower.tail = FALSE)
-% [1] 0.008511231
-
-Because we use 13 haplotypes per MHC class,
-for a significance level of 5\%,
-for both directions (D to U, or U to D)
-we expect 5% times 26 equals 1.3 significant findings.
-
-\richel{Hmm, improve}
-The chance to find 2 haplotypes that hint to have selection working upon it
-is 4\%, which is below our significance level.
-% > pbinom(q = 3, size = 26, prob = 0.05, lower.tail = FALSE)
-% [1] 0.0387414
-
-Upon finding 3 haplotypes that hint to have selection working upon it,
-there is the additional chance that selection is working in different
-directions (i.e. in one case there is selection for losing detection,
-where in the other case, there is (unexpected) selection for gaining
-detection). Therefore, finding 3 haplotypes that hint to have selection working 
-upon it in the same direction
-
-Therefore, we will claim there is directional selection, 
-when, for a proteome, we find 2 haplotypes 
-that hint to have selection working upon it
-
-\paragraph{System depicted as a Markov chain}
-
-To calculate the probability all of our observations happened by chance,
-we view our experiment as a two-state Markov chain, as depicted in
-figure \ref{fig:markov_chain}. The two nodes depict the
-two possible detection state a peptide is in, which is
-either undetected $U$, or detected $D$ by the immune system,
-where a detected peptide is a binder, that is, 
-a peptide with an IC50 in the lowest 2\%.
-The $D$ can be either detection by an MHC-I or MHC-II haplotype.
-The edges depict the probability of a state transition by mutation,
-for example, $p_{DU}$ denotes the probability that a detected
-peptide mutates into an undetected one.
-We focus on peptides that are TMH and we assume a mutation that 
-renders the TMH non-functional as lethal.
-
-\begin{figure}
-  \centering
-  \input{markov_chain_2_states.latex}
-  \caption{
-    State transitions in TMHs caused by mutations, with respect to the
-    detection by the immune system. Any TMH is either in a detected, $D$, or
-    undetected, $U$, state. Mutations cause state transitions with probability
-    $p$. For example, $p_{DU}$ denotes the probability that a detected
-    TMH mutates into an undetected one.
-  }
-  \label{fig:markov_chain}
-\end{figure}
+To detect the evolutionary conservation of TMHs,
+we create an alignment using msa, that uses msaClustalW in the back,
+create a consensus sequence, determine the topology of the consensus
+using the FLU transition matrix,
+measure the topology of it, then estimate the conservation values.
 
@@
 
-When we have calculated the probability of each transition (as described below),
-we can simply count the number of observed transitions.
-Because the probability of transitions differ per peptide,
-the total system has the properties of a Poisson binomial distribution,
-from which we can calculate the probability of the observing our
-number of transitions.
-
-\paragraph{Hypotheses testing}
-
-Of the complete system, we test for each of the 2 observed 
-state transition rates (that is, we do not test for mutations that result in the same state)
-if these are likely to be caused by chance.
-For each transition rate, we know the expected and observed number
-of transitions. If the observed number of transitions is below the expected
-value, we test for finding that (low) number of transitions or less.
-If the observed number of transitions is above the expected
-value, we test for finding that (high) number of transitions or more.
-We use a significance level $\alpha$ of 5\%, because we have no prior
-expectations.
-
-As we are testing 2 rates (and thus 2 hypotheses), 
-we correct against type I errors (false positives), 
-by applying the Bonferroni correction, 
-resulting in $\alpha_c = \frac{\alpha}{2}$.
-If the probability of observing that few/many 
-transition is below $\alpha_c$, we will state 
-that this unlikely situation may be caused by natural selection.
-
-\paragraph{Calculating one peptide's probability}
-
-We calculate the probability of a transition per peptide.
-For any peptide, we can generate all sequences that differ
-in only one amino acids. Of these sequences, we only keep the ones
-that result in a TMH. Of each of the remaining TMH sequences, we predict
-which mutants are detected by the immune system. Because not all 
-AA mutations are equally likely, we correct for the AA transition
-rates using the FLU transition matrix (\cite{dang2010flu}),
-which is a transition matrix derived from observed influenza AA mutations.
-
-\paragraph{Assumption that epitopes are immunogenic}
-
-We assume that if a peptide binds to an MHC, it is always presented and always
-leads to an immune response.
-The more often this assumption is violated, 
-the more our analysis will be weakened.
-Experiments in mice with MHC-I epitopes, however, indicate that (if the dose 
-administered was high enough) there are either none or rare
-exceptions \cite{sette1994relationship}.
-
-\paragraph{Measuring strength}
-
-To assess the strength of our analysis, 
-we have applied it to four simulated data sets,
-which are the product of using either MHC-I or MHC-II,
-as well as using two different selection scenarios.
-
-For both MHC-I and MHC-II, we use the haplotypes that have 
-the lowest phenotypic frequency, which is \richel{X} for MHC-I
-and DPB1*0501 (21.7\%) for MHC-II.
-
-The two selection scenarios are either no selection
-or a two-fold selection on what we would expect.
-The scenario without selection, the 'null' data set, acts 
-as a control, where we let the simulate observations occur by chance. 
-The 'rigged' data set is a data set in which we 
-let evolution bias the observations two-fold, that is, 
-we let mutations that are beneficial occur twice as often as expected by chance.
-
-All data sets consists out of $10^4$ observations.
-Each observation starts with a randomly generated source peptide fragment,
-as well a mutant. For the 'null' data set, the mutant is randomly select.
-For the 'rigged' data set, beneficial mutations are twice as likely to
-be selected. We define a beneficial mutation as a transition from
-being detected to being undetected.
+We compare the conservation values for TMH versus non-TMH locations,
+using a Kolmogorov-Smirnov test, for a significance level of 5 percent.
 
@@
 \subsection{Prediction software used}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\label{subsec:prediction_software_used}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[]
+  \begin{tabular}{lll}
+    Goal & Tool  \\ 
+    \hline
+    Predict topology                  & TMHMM \cite{krogh2001predicting} \\
+    Predict topology                  & PureseqTM \cite{wang2019efficient} \\
+    Predict epitopes MHC-I            & \verb;epitope-prediction; \cite{bianchi2017} \\
+    Predict epitopes MHC-II           & NetMHCIIpan \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
+    Predict epitopes MHC-I and MHC-II & MHCnuggets \cite{shao2020high} \\
+    Call TMHMM from R                 & tmhmm \\
+    Call PureseqTM from R            & pureseqtmr \\
+    Call NetMHCIIpan from R           & netmhc2pan \\
+    Call MHCnuggets from R            & mhcnuggetsr \\
+    Call TMHMM from R                 & tmhmm \\
+    \verb;epitope-prediction; LUT     & epiprepreds \\
+    NetMHCIIpan LUT                   & nmhc2ppreds \\
+    MHCnuggets LUT                    & mhcnpreds \\
+    Combine all                       & bbbq
+  \end{tabular}
+  \caption{
+    Overview of all software used in this research.
+    Software without a reference is written for this research.
+    'LUT' is an abbreviation for 'Look-up table'. 
+  }
+  \label{table:software_used}
+\end{table}
 
@@
 There are multiple computational tools developed to predict which
-parts of membrane proteins are TMH
+parts of membrane proteins are TMH.
 In 2001, multiple tools to do so have been compared \cite{moller2001evaluation},
-of which TMHMM \cite{krogh2001predicting} turned out to be the best, yet
-this software has a restrictive software license and is nearly two
-decades old. Inspired on TMHMM, many other tools followed, such as 
+of which TMHMM \cite{krogh2001predicting} turned out to be the best.
+TMHMM is used in the previous study \cite{bianchi2017},
+yet TMHMM itself has a restrictive software license and is nearly two
+decades old. 
+Inspired on TMHMM, many other tools followed, such as 
 Phobius (\cite{kall2004combined}),
@@
 which uses a stabilized matrix method \cite{kim2009derivation}.
-The most recent FOSS tool available is MHCnuggets\cite{shao2020high},
-which we will use in this study
+The most recent FOSS tool available is 
+MHCnuggets \cite{shao2020high} (which can do both MHC-I 
+and MHC-II prediction) which we will use in this study
 
@@
 
-% \paragraph{Gene prediction}
-% 
-% Proteogenomics is the field dedicated to, among others,
-% predict the proteins that arise from a DNA 
-% sequence \cite{nesvizhskii2014proteogenomics}.
-% For an ab initio gene prediction of a viral genome, 
-% we searched the literature for the most recent software tool,
-% which is Vgas (\cite{zhang2019vgas}). In this article, 
-% Vgas is compared to tools, such as GeneMarkS (\cite{besemer2001genemarks}),
-% Prodigal (\cite{hyatt2012gene}), GLIMMER (\cite{delcher1999improved}),
-% RAST \cite{aziz2008rast} and prokka (\cite{seemann2014prokka}).
-% We picked the FOSS tool that scored best in the (closed source) Vgas
-% benchmark (\cite{zhang2019vgas}), which is Prodigal.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The complete experiment is bundled in the 'bbbq' R package,
-which is dependent on 'pureseqtmr', 'mhcnuggetsr' and 'mhcnpreds',
+which is dependent on 'pureseqtmr', 
+'mhcnuggetsr' and 'mhcnpreds',
+'epitope-prediction' and 'epiprepreds',
+'netmhc2pan' and 'nmhc2ppreds'
 as described below.
@@
 
-To be able to use pre-calculated MHCnuggets
+To be able to use pre-calculated 
+\verb;epitope-prediction;, NetMHCIIpan and MHCnuggets
 predictions (i.e. a distribution of IC50 values per haplotype, 
-derived from random peptides), the 'mhcnpreds' R package has been 
-developed. In practice, this is nothing more than a big look-up table.
-Because the package is too big, it will not been submitted to CRAN.
+derived from random peptides), the 'epiprepreds', 'nmhc2ppreds' 
+and 'mhcnpreds' R package have been 
+developed. In practice, these are nothing more than a big look-up table.
+If \verb;epitope-prediction; is put onto CRAN, 'epiprepreds'
+will be submitted as well.
+If 'netmhc2pan' is allowed onto CRAN, 'nmhc2ppreds'
+will be submitted as well.
+If 'mhcnuggetsr' is put onto CRAN, 'mhcnpreds'
+will be submitted as well.
 
@@
 
+\input{bbbq_1/table_f_tmh.latex}
+
 \paragraph{IC50 prediction software}
@@
   \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
-  Predicted IC50s (in nM) per haplotype
-  \caption{
-    \richel{Used randomly simulated peptides, results are real}
-  }
+  Distribution of predicted IC50s (in nM) per haplotype
   \label{fig:ic50s_per_haplotype}

### 2020-09-02 bbbq_article.tex
Correctly cite msa, fix https://github.com/richelbilderbeek/bbbq_article/issues/68

@@
 To detect the evolutionary conservation of TMHs,
-we create an alignment using msa, that uses msaClustalW in the back,
-create a consensus sequence, determine the topology of the consensus
-using the FLU transition matrix,
-measure the topology of it, then estimate the conservation values.
+we first selected all unique protein sequences per
+gene from all 45k SARS-CoV-2 proteomes,
+resulting in around 500 unique protein sequences per gene.
+From the unique protein sequences, we made an alignment per gene,
+using the msa \cite{bodenhofer2015msa} R package, 
+using default settings, including the use of the default alignment tool
+ClustalW \cite{thompson1994clustal}.
+Per alignment, we used msa to determine the consensus sequence
+and the conservation scores.
+Per consensus sequences, the topology was determined using pureseqtmr.
+
+\richel{use the FLU transition matrix?}
 

### 2020-09-02 bbbq_article.tex
Remove 'Evolutionary signal of unknown strength for pathogens'

@@
 those genes that are selected cannot be detected \cite{yang2000statistical}.
-
-\paragraph{Evolutionary signal of unknown strength for pathogens}
-
-Up until today, it is unknown if an evolutionary signal 
-is present in a (human) pathogen's proteome that stems from avoiding
-detection by the human immune system.
-If such a signal is present, we expect to detect it most likely in
-viruses, as both MHC-I and MHC-II are used in protecting the
-human host. 
-Therefore, this study focuses in detecting this signal in viruses.
 

### 2020-09-02 bbbq_article.tex
Cite R packages I wrote

@@
 with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
-To summarize: a proteome is converted into all its possible 9-mers. For each
-of these peptides, it is measured if it has been part of a TMH (at least
-1 residue), as well as if the peptide is a binder to an MHC-I haplotype.
+To summarize: from a proteome all possible 9-mers are derived. For each
+of these peptides, it is measured if it has been part of a 
+TMH (at least overlapping 1 residue), 
+as well as if the peptide is a binder to an MHC-I haplotype.
 Because the percentage of TMH residues is known, a binomial test can be
@@
 We compare the conservation values for TMH versus non-TMH locations,
+for all membrane proteins combined,
 using a Kolmogorov-Smirnov test, for a significance level of 5 percent.
@@
     Predict epitopes MHC-I and MHC-II & MHCnuggets \cite{shao2020high} \\
-    Call TMHMM from R                 & tmhmm \\
-    Call PureseqTM from R            & pureseqtmr \\
-    Call NetMHCIIpan from R           & netmhc2pan \\
-    Call MHCnuggets from R            & mhcnuggetsr \\
-    Call TMHMM from R                 & tmhmm \\
-    \verb;epitope-prediction; LUT     & epiprepreds \\
-    NetMHCIIpan LUT                   & nmhc2ppreds \\
-    MHCnuggets LUT                    & mhcnpreds \\
-    Combine all                       & bbbq
+    Call TMHMM from R                 & tmhmm \cite{tmhmm} \\
+    Call PureseqTM from R             & pureseqtmr \cite{pureseqtmr} \\
+    Call NetMHCIIpan from R           & netmhc2pan \cite{netmhc2pan} \\
+    Call MHCnuggets from R            & mhcnuggetsr \cite{mhcnuggetsr} \\
+    \verb;epitope-prediction; LUT     & epiprepreds \cite{epiprepreds} \\
+    NetMHCIIpan LUT                   & nmhc2ppreds \cite{nmhc2ppreds} \\
+    MHCnuggets LUT                    & mhcnpreds \cite{mhcnpreds} \\
+    Combine all                       & bbbq \cite{bbbq}
   \end{tabular}
@@
 MHCnuggets \cite{shao2020high} (which can do both MHC-I 
-and MHC-II prediction) which we will use in this study
+and MHC-II prediction) which is used in this study
 
@@
 and the very recent MHCnuggets (\cite{shao2020high}),
-the latter which we will use in this study.
+the latter which is used in this study.
 

### 2020-09-02 bbbq_article.tex
Nice table of software

@@
 \begin{table}[]
-  \begin{tabular}{lll}
-    Goal & Tool  \\ 
+  \begin{tabular}{llll}
+    Goal & Tool & Reference \\ 
     \hline
-    Predict topology                  & TMHMM \cite{krogh2001predicting} \\
-    Predict topology                  & PureseqTM \cite{wang2019efficient} \\
-    Predict epitopes MHC-I            & \verb;epitope-prediction; \cite{bianchi2017} \\
-    Predict epitopes MHC-II           & NetMHCIIpan \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
-    Predict epitopes MHC-I and MHC-II & MHCnuggets \cite{shao2020high} \\
-    Call TMHMM from R                 & tmhmm \cite{tmhmm} \\
-    Call PureseqTM from R             & pureseqtmr \cite{pureseqtmr} \\
-    Call NetMHCIIpan from R           & netmhc2pan \cite{netmhc2pan} \\
-    Call MHCnuggets from R            & mhcnuggetsr \cite{mhcnuggetsr} \\
-    \verb;epitope-prediction; LUT     & epiprepreds \cite{epiprepreds} \\
-    NetMHCIIpan LUT                   & nmhc2ppreds \cite{nmhc2ppreds} \\
-    MHCnuggets LUT                    & mhcnpreds \cite{mhcnpreds} \\
-    Combine all                       & bbbq \cite{bbbq}
+    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
+    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
+    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
+    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
+    Predict epitopes MHC-I and MHC-II & MHCnuggets                & \cite{shao2020high} \\
+    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
+    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
+    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
+    Call MHCnuggets from R            & mhcnuggetsr               & \cite{mhcnuggetsr} \\
+    \verb;epitope-prediction; LUT     & epiprepreds               & \cite{epiprepreds} \\
+    NetMHCIIpan LUT                   & nmhc2ppreds               & \cite{nmhc2ppreds} \\
+    MHCnuggets LUT                    & mhcnpreds                 & \cite{mhcnpreds} \\
+    Sequence alignment                & ClustalW                  & \cite{thompson1994clustal} \\
+    Conservation scores               & msa                       & \cite{bodenhofer2015msa} \\
+    Combine all                       & bbbq                      & \cite{bbbq}
   \end{tabular}
@@
     Overview of all software used in this research.
-    Software without a reference is written for this research.
     'LUT' is an abbreviation for 'Look-up table'. 

### 2020-09-02 bbbq_article.tex
Removed irrelevant text, such as 'Strength of the dataset' and general info about SARS-CoV-2

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Strength of COVID-19 dataset}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-See at the end of this document.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{COVID-19 genome and proteome}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Tip: see \cite{bar2020sars} for COVID-19 in numbers.
-
-\richel{
-  This is just a reminder, instead of new research. 
-  This subsection be deleted in the future.
-}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{pics/covid_genome_and_proteome.png}
-  \caption{
-    Overview of COVID-19 genome and proteome.
-    From 
-    \url{https://zhanglab.ccmb.med.umich.edu/COVID-19}
-  }
-  \label{fig:covid_genome_and_proteome}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{IC50s per haplotype}
@@
 \end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Added PDFs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-These PDFs are added. These will be moved into the article when needed.

### 2020-09-02 create.sh

@@ -13,6 +13,6 @@ pdflatex bbbq_header.tex
 bibtex bbbq_header >/dev/null
 pdflatex bbbq_header.tex >/dev/null
 pdflatex bbbq_header.tex
-pdftk bbbq_header.pdf covid_dataset_strength/covid_dataset_strength.pdf \
-  cat output bbbq_article.pdf
-rm bbbq_header.pdf
+#pdftk bbbq_header.pdf covid_dataset_strength/covid_dataset_strength.pdf \
+#  cat output bbbq_article.pdf
+mv bbbq_header.pdf bbbq_article.pdf

### 2020-09-02 ic50s_per_haplotype/ic50s_per_haplotype.R
'ic50s_per_haplotype' shows IC50 predictions for different tools

@@ -1,67 +1,53 @@
 library(bbbq)
 library(ggplot2, quietly = TRUE)
-library(dplyr)
-library(testthat)
+library(dplyr, warn.conflicts = FALSE)
+library(testthat, warn.conflicts = FALSE)
 
 t <- tidyr::expand_grid(
   official_name = bbbq::get_mhc_haplotypes(),
   mhc_class = NA,
-  mhcnuggets_name = NA,
   peptide_length = NA,
-  filename = NA,
-  n_peptides = NA,
   percentile = bbbq::get_ic50_percentile_binder(),
+  ic50_prediction_tool = c("EpitopePrediction", "netmhc2pan", "mhcnuggetsr"),
   ic50 = NA
 )
+
+
+t$mhc_class[ t$official_name %in% bbbq::get_mhc1_haplotypes() ] <- "I"
+t$mhc_class[ t$official_name %in% bbbq::get_mhc2_haplotypes() ] <- "II"
+t$peptide_length[ t$official_name %in% bbbq::get_mhc1_haplotypes() ] <- bbbq::get_mhc1_peptide_length()
+t$peptide_length[ t$official_name %in% bbbq::get_mhc2_haplotypes() ] <- bbbq::get_mhc2_peptide_length()
+
+nrow(t)
+t <- t[-which(t$ic50_prediction_tool == "EpitopePrediction" & t$mhc_class == "II"), ]
+nrow(t)
+t <- t[-which(t$ic50_prediction_tool == "netmhc2pan" & t$mhc_class == "I"), ]
+nrow(t)
+
 for (i in seq_len(nrow(t))) {
-  if (t$official_name[i] %in% bbbq::get_mhc1_haplotypes()) {
-    t$mhc_class[i] <- "I"
-    t$peptide_length[i] <- 9
-  }
-  if (t$official_name[i] %in% bbbq::get_mhc2_haplotypes()) {
-    t$mhc_class[i] <- "II"
-    t$peptide_length[i] <- 10
-  }
-}
-for (i in seq_len(nrow(t))) {
-  t$mhcnuggets_name[i] <- mhcnuggetsr::to_mhcnuggets_name(t$official_name[i])
-}
-expect_true(all(mhcnuggetsr::are_mhcnuggets_names(t$mhcnuggets_name)))
-for (i in seq_len(nrow(t))) {
-  t$filename[i] <- mhcnpreds::get_lut_filename(
[…]

### 2020-09-03 ic50s_per_haplotype/ic50s_per_haplotype.R
Create LaTeX table in R

@@ -1,5 +1,5 @@
 library(bbbq)
-library(ggplot2, quietly = TRUE)
+library(ggplot2, warn.conflicts = FALSE, quietly = TRUE)
 library(dplyr, warn.conflicts = FALSE)
 library(testthat, warn.conflicts = FALSE)
 
@@ -40,17 +40,44 @@ t
 expect_true(all(t$percentile == t$percentile[1]))
 percentile <- t$percentile[1]
 
+title <- "IC50 values that define a binder"
+
+
+ic50_prediction_tool_colors <- c(
+  EpitopePrediction = "#555555",
+  mhcnuggetsr = "#AAAAAA",
+  netmhc2pan = "#FFFFFF"
+)
+
+
 ggplot(t, aes(x = official_name, y = ic50, fill = ic50_prediction_tool)) +
-  geom_col(position = "dodge") +
+  geom_col(position = "dodge", col = "black") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_fill_manual(values = ic50_prediction_tool_colors) +
   scale_y_log10(name = "IC50 (nM) threshold to be a binder") +
   scale_x_discrete(name = "Haplotype") +
   labs(
+    title = title,
     caption = glue::glue(
       "Percentile: {percentile}"
     )
   ) +
   ggsave("ic50s_per_haplotype.png", width = 7, height = 7)
+
+  ggplot(t, aes(x = official_name, y = ic50, fill = ic50_prediction_tool)) +
+  geom_col(position = "dodge", col = "black") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  scale_fill_manual(values = ic50_prediction_tool_colors) +
+  ggplot2::facet_grid(. ~ mhc_class, scales = "free_x") +
+  scale_y_log10(name = "IC50 (nM) threshold to be a binder") +
+  scale_x_discrete(name = "Haplotype") +
+  labs(
+    title = title,
+    caption = glue::glue(
+      "Percentile: {percentile}"
+    )
+  ) +
+  ggsave("ic50s_per_haplotype_per_mhc_class.png", width = 7, height = 7)
 
 
 
@@ -60,3 +87,11 @@ readr::write_csv(
   rename(haplotype = official_name),
   "ic50s_per_haplotype.csv"
 )
+
+knitr::kable(
+  t %>% dplyr::select(-percentile), "latex",
+  caption = glue::glue(
[…]

### 2020-09-03 ic50s_per_haplotype/ic50s_per_haplotype.R
Make better tables

@@ -88,10 +88,40 @@ readr::write_csv(
   "ic50s_per_haplotype.csv"
 )
 
+t
+t_wide <- tidyr::pivot_wider(
+  data = t %>% dplyr::select(official_name, ic50_prediction_tool, ic50, mhc_class),
+  names_from = ic50_prediction_tool,
+  values_from = ic50
+)
+t_wide
+
 knitr::kable(
-  t %>% dplyr::select(-percentile), "latex",
+  t_wide, "latex",
   caption = glue::glue(
     "{title}. Percentile: {percentile}"
   ),
   label = "ic50s_per_haplotype"
 ) %>% cat(., file = "ic50s_per_haplotype.latex")
+
+knitr::kable(
+  t_wide %>% dplyr::filter(mhc_class == "I") %>% dplyr::select(official_name, EpitopePrediction, mhcnuggetsr),
+  "latex",
+  caption = glue::glue(
+    "{title}, for MHC-I haplotypes. ",
+    "Peptide length: {bbbq::get_mhc1_peptide_length()}. ",
+    "Percentile: {percentile}."
+  ),
+  label = "ic50s_per_haplotype_1"
+) %>% cat(., file = "ic50s_per_haplotype_1.latex")
+
+knitr::kable(
+  t_wide %>% dplyr::filter(mhc_class == "II") %>% dplyr::select(official_name, netmhc2pan, mhcnuggetsr),
+  "latex",
+  caption = glue::glue(
+    "{title}, for MHC-II haplotypes. ",
+    "Peptide length: {bbbq::get_mhc2_peptide_length()}. ",
+    "Percentile: {percentile}."
+  ),
+  label = "ic50s_per_haplotype_2"
+) %>% cat(., file = "ic50s_per_haplotype_2.latex")

### 2020-09-03 bbbq_article.tex
Better tables

@@
 
+% Less pretty 
+% 
+% \begin{figure}[!htbp]
+%   \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
+%   \caption{
+%     IC50 values that define a binder
+%   }
+%   \label{fig:ic50s_per_haplotype}
+% \end{figure}
+
 \begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
-  Distribution of predicted IC50s (in nM) per haplotype
+  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype_per_mhc_class.png}
+  \caption{
+    IC50 values that define a binder
+  }
   \label{fig:ic50s_per_haplotype}
@@
 
-\begin{table}[!htbp]
-  \input{ic50s_per_haplotype/ic50s_per_haplotype.latex}
-  \caption{
-    Predicted IC50s (in nM) per haplotype
-    \richel{Used randomly simulated peptides, results are real}
-  }
-  \label{table:ic50s_per_haplotype}
-\end{table}
+\input{ic50s_per_haplotype/ic50s_per_haplotype_1.latex}
+
+\input{ic50s_per_haplotype/ic50s_per_haplotype_2.latex}
 

### 2020-09-03 bbbq_article.tex
Show conservation scores from bbbq_2, show IC50 values for binders per haplotype from the 'ic50s_per_haplotype' folder

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evolutionary conservation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_boxplot.png}
+  \caption{
+    Evolutionary conservation
+  }
+  \label{fig:evolutionary_conservation_boxplot}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1/fig_ic50_distribution.png}
-  \caption{
-    Distribution of IC50 estimations per haplotype per target.
-    The IC50 value is plotted on a logarithmic scale.
-  }
-  \label{fig:1}
-\end{figure}
-
 \paragraph{Definition of what a binder is}
@@
 is within the peptides with the 2\% lowest IC50 values.
-
-\input{bbbq_1/table_ic50_binders.latex}
+Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
+for a binder per haplotype.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evolutionary conservation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
+  \caption{
+    Evolutionary conservation
+  }
+  \label{fig:evolutionary_conservation_tmh}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_all.png}
+  \caption{
+    Evolutionary conservation
+  }
+  \label{fig:evolutionary_conservation_all}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{IC50s per haplotype}
+\label{subsec:ic50s_per_haplotype}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{COVID-19 TMHs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{tmhs/tmhs.png}
-  \caption{
-    Location of amino acids of reference COVID-19 
-    proteome (Uniprot ID: UP000464024).
-    Note that ORF1a results in multiple proteins, 
-    as shown in figure \ref{fig:covid_genome_and_proteome}.
-    Legend: 1 = located in the membrane, 0 = not located in the membrane.
-    Topology obtained using 'pureseqtmr' \cite{pureseqtmr}
-  }
-  \label{fig:covid_topology}
-\end{figure}
-
-\begin{table}[!htbp]
-  \input{tmhs/covid_tmhs.latex}
-  \caption{
-    Number of TMHs per COVID-19 ORF,
-    using 'pureseqtmr' \cite{pureseqtmr}
-  }
-  \label{table:covid_tmhs}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-II haplotype occurrences}

### 2020-09-03 bbbq_article.tex
Start taking shape

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Relation between hydrophobicity, TMH and MHC binding}
+%\subsection{Relation between hydrophobicity, TMH and MHC binding}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/binds_mhc1_vs_binds_mhc2.png}
-  \caption{
-    \richel{Used randomly simulated peptides, results are real}
-  }
-  \label{fig:binds_mhc1_vs_binds_mhc2}
-\end{figure}
+%\begin{figure}[!htbp]
+%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/binds_mhc1_vs_binds_mhc2.png}
+%  \caption{
+%    \richel{Used randomly simulated peptides, results are real}
+%  }
+%  \label{fig:binds_mhc1_vs_binds_mhc2}
+%\end{figure}
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc1.png}
-  \caption{
-    \richel{Used randomly simulated peptides, results are real}
-  }
-  \label{fig:hydrophobicity_vs_binds_mhc1}
-\end{figure}
+%\begin{figure}[!htbp]
+%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc1.png}
+%  \caption{
+%    \richel{Used randomly simulated peptides, results are real}
+%  }
+%  \label{fig:hydrophobicity_vs_binds_mhc1}
+%\end{figure}
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc2.png}
-  \caption{
-    \richel{Used randomly simulated peptides, results are real}
-  }
-  \label{fig:hydrophobicity_vs_binds_mhc2}
-\end{figure}
+%\begin{figure}[!htbp]
+%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc2.png}
+%  \caption{
+%    \richel{Used randomly simulated peptides, results are real}
+%  }
+%  \label{fig:hydrophobicity_vs_binds_mhc2}
+%\end{figure}
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_is_tmh.png}
-  \caption{
-    \richel{Used randomly simulated peptides, results are real}
-  }
-  \label{fig:hydrophobicity_vs_is_tmh}
-\end{figure}
+%\begin{figure}[!htbp]
+%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_is_tmh.png}
+%  \caption{
+%    \richel{Used randomly simulated peptides, results are real}
+%  }
+%  \label{fig:hydrophobicity_vs_is_tmh}
+%\end{figure}
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc1.png}
-  \caption{
-    \richel{Used randomly simulated peptides, results are real}
-  }
-  \label{fig:is_tmh_vs_binds_mhc1}
-\end{figure}
+%\begin{figure}[!htbp]
+%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc1.png}
+%  \caption{
+%    \richel{Used randomly simulated peptides, results are real}
+%  }
+%  \label{fig:is_tmh_vs_binds_mhc1}
+%\end{figure}
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc2.png}
-  \caption{
-    \richel{Used randomly simulated peptides, results are real}
-  }
-  \label{fig:is_tmh_vs_binds_mhc2}
-\end{figure}
+%\begin{figure}[!htbp]
+%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc2.png}
+%  \caption{
+%    \richel{Used randomly simulated peptides, results are real}
+%  }
+%  \label{fig:is_tmh_vs_binds_mhc2}
+%\end{figure}
 
@@
 
-\richel{
-  This is just a reminder, instead of new research. 
-  This subsection be deleted in the future.
-}
-
 \begin{table}[!htbp]
@@
     Percentage of MHC-II haplotypes, from \cite{greenbaum2011functional}
-    \richel{
-      This is just a reminder, instead of new research. 
-      This table be deleted in the future.
-    }
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{
-  This is just a reminder, instead of new research. 
-  This subsection be deleted in the future.
-}
 

### 2020-09-07 bbbq_article.tex
Use results from 'bbbq_1_fast'
Use all BBBQ2 figures

@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc1.png}
+  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc1_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc2.png}
+  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc2_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
+  \caption{
+    Conservation scores across the SARS-CoV-2 proteome
+    for proteins that have at least one TMH
+  }
+  \label{fig:evolutionary_conservation}
+\end{figure}
+
+\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{bbbq_2/scores_boxplot.png}
   \caption{
-    Evolutionary conservation
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    for proteins that have at least one TMH
+  }
+  \label{fig:evolutionary_conservation_boxplot}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_all.png}
+  \caption{
+    Conservation scores across the SARS-CoV-2 proteome
+    for all proteins
+  }
+  \label{fig:evolutionary_conservation}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_boxplot_all.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    for all proteins
   }

### 2020-09-07 bbbq_article.tex
Improve abstract, use SARS-CoV-2 and Mycobacterium tuberculosis instead of 'COVID/covid' and 'Mycobacterium/myco'

@@
 [less/equally/more] often than expected by chance.
-Lastly, using an innovative computation method, 
-we show that natural selection on TMHs is [strong enough/still too weak]
-to detect.
+Lastly, we show the TMHs are [more/less/equally] 
+conserved.
 Our findings suggest that the immune system is [less/neutral/more]
@@
 adaptive immunity, transmembrane domain, transmembrane helix, 
-epitopes, T lymphocyte, MHC-1, MHC-I, MHC-2, MHC-II, COVID-19
+epitopes, T lymphocyte, MHC-I, MHC-II, SARS-CoV-2
 
@@
 TMHs are general structures: 25 percent of the human proteome is
-anchored by at least one TMH. COVID-19 has 21 TMHs, 
+anchored by at least one TMH. SARS-CoV-2 has 21 TMHs, 
 making up for \richel{?} percent of its entire genome.
@@
 essential for a virus, as these will be most conserved.
-In COVID-19, for example, there is preliminary evidence that the strongest
+In SARS-CoV-2, for example, there is preliminary evidence that the strongest
 selection pressure is upon residues that changes its 
@@
 The viral reference proteome is represented by
-the proteome of the first sequenced COVID-19 strain (\cite{wu2020new},
+the proteome of the first sequenced SARS-CoV-2 strain (\cite{wu2020new},
 GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
@@
     \% epitopes overlapping with transmembrane helix
-    for a human, COVID-19 and Mycobacterium proteome.
+    for a human, SARS-CoV-2 (labeled 'covid') 
+    and Mycobacterium tuberculosis (labeled 'myco') proteome. 
   }
@@
 to be derived from TMH within either a human host and its viral pathogen.
-Because the full COVID-19 has 21 TMHs, the percentages
+Because the full SARS-CoV-2 has 21 TMHs, the percentages
 of MHC-I epitopes being part of a TMH are likelier to be affected by
-stochasticity. We chose to use COVID-19 regardless, as the thousands
+stochasticity. We chose to use SARS-CoV-2 regardless, as the thousands
 of its time-dated genomic sequences are ideal for determining the 
@@
 %   \caption{
-%     Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-I
+%     Kolmogorov-Smirnov test results comparing human and SARS-CoV-2 for MHC-I
 %     \richel{Done on the simulated data}
@@
 %   \caption{
-%     Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-I
+%     Kolmogorov-Smirnov test results comparing human and Mycobacterium tuberculosis for MHC-I
 %     \richel{Done on the simulated data}
@@
 %   \caption{
-%     Kolmogorov-Smirnov test results comparing human and COVID-19 for MHC-II
+%     Kolmogorov-Smirnov test results comparing human and SARS-CoV-2 for MHC-II
 %     \richel{Done on the simulated data}
@@
 %   \caption{
-%     Kolmogorov-Smirnov test results comparing human and Mycobacterium for MHC-II
+%     Kolmogorov-Smirnov test results comparing human and 
+      Mycobacterium tuberculosis for MHC-II
 %     \richel{Done on the simulated data}

### 2020-09-07 bbbq_article.tex
Better references, add percentage of TMH of SARS-CoV-2

@@
 
-Transmembrane helices are conserved structures that span
+Transmembrane helices are structurally conserved structures that span
 a cell membrane with an alpha helix.
@@
 anchored by at least one TMH. SARS-CoV-2 has 21 TMHs, 
-making up for \richel{?} percent of its entire genome.
+making up for 5.5 percent of its entire genome.
 There are multiple computational tools developed to predict which
@@
 from TMHs are presented more often than expected by 
-chance \cite{bianchi2017}
+chance (\cite{bianchi2017})
 \richel{but also when compensating for hydrophobicity?}.
@@
 
-An increased detection
-is obtained by expressing a wide variety of
+One of many strategies to increase the number of pathogenic
+peptides that can be detected is to express a wide variety of
 MHCs, which is assured by the human leukocyte antigen gene complex (HLA).
@@
 selection pressure is upon residues that changes its 
-virulence \cite{velazquez2020positive}.
+virulence (\cite{velazquez2020positive}).
 These loci, however, only account for a small part of a pathogen's proteome.
@@
 where other (called subdominant) epitopes end up having no 
-effect \cite{akram2012immunodominance}.
+effect (\cite{akram2012immunodominance}).
 The same study mentions 20 factors that influence the strength
@@
 a mutation within an epitope, 
-such as the in vitro example described in \cite{berkhoff2004mutation},
+such as the in vitro example described in (\cite{berkhoff2004mutation}),
 where one single mutation resulted in a 5-20\% weaker
@@
 in non- versus synonymous mutation
-to detect the signature of selection \cite{murrell2015gene}.
+to detect the signature of selection (\cite{murrell2015gene}).
 Using AA sequences, however, has its advantages,
@@
 selection acts upon: DNA may never be translated to RNA,
-or its RNA may never be transcribed \cite{diz2012proteomics}
+or its RNA may never be transcribed (\cite{diz2012proteomics})
 \richel{improve upon this point}.
@@
 can actually increase the 'fitness' of the 
-protein \cite{anisimova2009investigating}.
+protein (\cite{anisimova2009investigating}).
 There, when generalizing the dynamics of mostly purifying selection (to maintain
 a protein's function) and a short duration of positive selection,
-those genes that are selected cannot be detected \cite{yang2000statistical}.
+those genes that are selected cannot be detected (\cite{yang2000statistical}).
 
@@
 Instead of using the membrane protein topology prediction
-tool \verb;TMHMM; \cite{krogh2001predicting}, PureseqTM \cite{wang2019efficient}
+tool \verb;TMHMM; (\cite{krogh2001predicting}), PureseqTM (\cite{wang2019efficient})
 is used instead.
@@
 From the unique protein sequences, we made an alignment per gene,
-using the msa \cite{bodenhofer2015msa} R package, 
+using the msa (\cite{bodenhofer2015msa}) R package, 
 using default settings, including the use of the default alignment tool
-ClustalW \cite{thompson1994clustal}.
+ClustalW (\cite{thompson1994clustal}).
 Per alignment, we used msa to determine the consensus sequence
@@
 parts of membrane proteins are TMH.
-In 2001, multiple tools to do so have been compared \cite{moller2001evaluation},
-of which TMHMM \cite{krogh2001predicting} turned out to be the best.
-TMHMM is used in the previous study \cite{bianchi2017},
+In 2001, multiple tools to do so have been compared (\cite{moller2001evaluation}),
+of which TMHMM (\cite{krogh2001predicting}) turned out to be the best.
+TMHMM is used in the previous study (\cite{bianchi2017}),
 yet TMHMM itself has a restrictive software license and is nearly two
@@
 from a set of multiple tools, 
-NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
-A tool developed later is \verb;epitope-prediction; \cite{bianchi2017},
-which uses a stabilized matrix method \cite{kim2009derivation}.
+NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
+A tool developed later is \verb;epitope-prediction; (\cite{bianchi2017}),
+which uses a stabilized matrix method (\cite{kim2009derivation}).
 The most recent FOSS tool available is 
-MHCnuggets \cite{shao2020high} (which can do both MHC-I 
+MHCnuggets (\cite{shao2020high}) (which can do both MHC-I 
 and MHC-II prediction) which is used in this study
@@
 gave rise to the most accurate predictions.
-Later tools are \cite{zhang2013predicting},
-\cite{trolle2014nettepi}, \cite{zhang2015accurate}
+Later tools are 
+\richel{name of tool here} (\cite{zhang2013predicting}),
+\richel{name of tool here} (\cite{trolle2014nettepi}),
+\richel{name of tool here} (\cite{zhang2015accurate})
 and the very recent MHCnuggets (\cite{shao2020high}),
@@
 We could have done the same experiment on beta-turn,
-as also these can be predicted well \cite{petersen2010netturnp},
+as also these can be predicted well (\cite{petersen2010netturnp}),
 are common structures and are present in all pathogens.
@@
 high performance computing cluster. 
-Additionally, we would like to thank Sci-Hub \cite{himmelstein2018sci}
+Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
 for allowing us to read paywalled articles while working from home.

### 2020-09-07 tmhs/tmhs.R

@@ -1,0 +1,45 @@
+library(pureseqtmr, warn.conflicts = FALSE)
+library(testthat, warn.conflicts = FALSE)
+library(tidyr, warn.conflicts = FALSE)
+library(ggplot2, warn.conflicts = FALSE)
+
+expect_true(is_pureseqtm_installed())
+
+
+
+fasta_filename <- "proteome.fasta"
+expect_true(file.exists(fasta_filename))
+
+
+topology <- pureseqtmr::predict_topology(fasta_filename)
+
+n_tmh_aas <- sum(stringr::str_count(topology$topology, pattern = "1"))
+n_non_tmh_aas <- sum(stringr::str_count(topology$topology, pattern = "0"))
+f_tmh <- n_tmh_aas / (n_tmh_aas + n_non_tmh_aas)
+text <-
+readr::write_lines(
+  x = glue::glue(
+    "n_tmh_aas: {n_tmh_aas}\n",
+    "n_non_tmh_aas: {n_non_tmh_aas}\n",
+    "f_tmh: {f_tmh}\n"
+  ),
+  "f_tmh.txt"
+)
+
+# Create tally as CSV
+tmhs_tally <- tally_tmhs(topology)
+write.csv(tmhs_tally, "covid_tmhs.csv", row.names = FALSE)
+
+# Create tally as LaTeX
+knitr::kable(
+  tmhs_tally, "latex",
+  caption = paste0(
+    "Number of TMHs per SARS-CoV-2 protein"
+  ),
+  label = paste0("covid_tmhs")
+) %>% cat(., file = "covid_tmhs.latex")
+
+# Create plot
+plot_topology(topology) + ggsave("tmhs.png", width = 7, height = 7)
+
+

### 2020-09-07 bbbq_article.tex
Main text points to tables in supmat

@@
 with TMHs for our human, viral and bacterial proteome.
+See the supplementary materials (tables \ref{tab:tmh_binders_mhc1} 
+and \ref{tab:tmh_binders_mhc2}) for the TMH and epitope counts.
 
@@
 
-\input{bbbq_1/table_f_tmh.latex}
+% Percentage of spots and spots that overlap with a TMH
+\input{bbbq_1_fast/table_f_tmh.latex}
 
@@
 
-\input{bbbq_1/table_tmh_binders_mhc1.latex}
+% Percentage of MHC-I 9-mers overlapping with TMH. Values in brackets
+% show the number of binders that have at least one residue overlapping with a
+% TMH (first value)as well as the number of binders (second value). Percentile
+% used: 0.02
+% Label: tab:tmh_binders_mhc1
+\input{bbbq_1_fast/table_tmh_binders_mhc1.latex}
 
 %\begin{table}[!htbp]
-%   \input{bbbq_1/bbbq_1_stats_covid.latex}
+%   \input{bbbq_1_fast/bbbq_1_stats_covid.latex}
 %   \caption{
@@
 % \begin{table}[!htbp]
-%   \input{bbbq_1/bbbq_1_stats_myco.latex}
+%   \input{bbbq_1_fast/bbbq_1_stats_myco.latex}
 %   \caption{
@@
 
-\input{bbbq_1/table_tmh_binders_mhc2.latex}
+\input{bbbq_1_fast/table_tmh_binders_mhc2.latex}
 
 % \begin{table}[!htbp]
-%   \input{bbbq_2/bbbq_2_stats_covid.latex}
+%   \input{bbbq_1_fast/bbbq_2_stats_covid.latex}
 %   \caption{
@@
 % \begin{table}[!htbp]
-%   \input{bbbq_2/bbbq_2_stats_myco.latex}
+%   \input{bbbq_1_fast/bbbq_2_stats_myco.latex}
 %   \caption{

### 2020-09-07 bbbq_article.tex
More neutral wording

@@
 
-The earlier study uses a hand-crafted method, that only works on MHC-I.
-Instead, we use MHCnuggets, which (1) free and open source, and (2) predicts
+The earlier study uses a hand-crafted method, 
+that has been trained on MHC-I haplotypes only.
+This study uses MHCnuggets, which (1) free and open source, and (2) predicts
 IC50s on both MHC-I and MHC-II.

### 2020-09-08 bbbq_article.tex
Mention the use of FLU, progress https://github.com/richelbilderbeek/bbbq_article/issues/62
Create graph: haplotype freq versus %TMH epitopes, fix https://github.com/richelbilderbeek/bbbq_article/issues/42

@@
 ClustalW (\cite{thompson1994clustal}).
+Instead of the default and general BLOSUM62 amino acid transition matrix (\cite{henikoff1992amino}),
+the FLU transition matrix (\cite{dang2010flu}) was used,
+as the latter is based on amino acid transitions in another viral pathogen, influenza.
 Per alignment, we used msa to determine the consensus sequence
@@
 Per consensus sequences, the topology was determined using pureseqtmr.
-
-\richel{use the FLU transition matrix?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Relation between haplotype and TMH over-representation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/haplotype_perc_vs_f_tmh_binders.png}
+  \caption{
+    Relation between genotype frequency and the percentage
+    of binders that are TMH for the 21 MHC-II haplotypes
+  }
+  \label{fig:haplotype_perc_vs_f_tmh_binders}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/phenotype_freq_vs_f_tmh_binders.png}
+  \caption{
+    Relation between phenotype frequency and the percentage
+    of binders that are TMH for the 21 MHC-II haplotypes
+  }
+  \label{fig:phenotype_freq_vs_f_tmh_binders}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{IC50s per haplotype}

### 2020-09-08 haplotype_freq_vs_f_tmh_epitopes/create_figure.R

@@ -1,0 +1,47 @@
+library(dplyr, warn.conflicts = FALSE)
+library(ggplot2, warn.conflicts = FALSE)
+
+
+counts_filenames <- "../bbbq_1_fast/counts.csv"
+testthat::expect_true(file.exists(counts_filenames))
+
+counts_per_protein <- readr::read_csv(counts_filenames)
+
+counts <- counts_per_protein %>% dplyr::group_by(target, haplotype_id) %>%
+  summarise(
+    n_binders = sum(n_binders, na.rm = TRUE),
+    n_binders_tmh = sum(n_binders_tmh, na.rm = TRUE),
+    n_spots = sum(n_spots, na.rm = TRUE),
+    n_spots_tmh = sum(n_spots_tmh, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+counts$f_tmh_binders <- counts$n_binders_tmh / counts$n_binders
+
+haplotypes_lut <- bbbq::create_haplotypes_lut()
+
+counts_with_haplotypes <- merge(counts, haplotypes_lut)
+
+occurances <- rbind(
+  bbbq::get_mhc1_haplotype_occurance(),
+  bbbq::get_mhc2_haplotype_occurance()
+)
+
+t <- merge(counts_with_haplotypes, occurances)
+t$phenotype_freq <- t$phenotype_freq/ 100.0
+t$haplotype_perc <- t$haplotype_perc/ 100.0
+
+
+ggplot2::ggplot(t, ggplot2::aes(x = phenotype_freq, y = f_tmh_binders, color = target)) +
+  ggplot2::geom_point() +
+  ggplot2::geom_smooth() +
+  ggplot2::scale_y_continuous("Percentage binders that are TMH", labels = scales::percent) +
+  ggplot2::scale_x_continuous("Haplotype phenotypic frequency", labels = scales::percent) +
+  ggplot2::ggsave("phenotype_freq_vs_f_tmh_binders.png", width = 7, height = 7)
+
+ggplot2::ggplot(t, ggplot2::aes(x = haplotype_perc, y = f_tmh_binders, color = target)) +
+  ggplot2::geom_point() +
+  ggplot2::geom_smooth() +
+  ggplot2::scale_y_continuous("Percentage binders that are TMH", labels = scales::percent) +
+  ggplot2::scale_x_continuous("Haplotype genotypic frequency", labels = scales::percent) +
+  ggplot2::ggsave("haplotype_perc_vs_f_tmh_binders.png", width = 7, height = 7)

### 2020-09-08 bbbq_article.tex
Remove Kolomorov-Smirnov test explanation, fix #81

@@
 \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Kolmogorov-Smirnov}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-The Kolmogorov-Smirnov (KS) test determines if two samples
-are derived from the same distribution, without making assumptions
-regarding the shape of that distribution.
-
-We will reject
-the null hypothesis that MHC-I has the same percentage of epitopes 
-overlapping with TMHs in Homo sapiens compared to each pathogen when 
-the KS statistic $D_{n,m}$ follows the relationship as shown in 
-equation \ref{eq:ks}, for a significance level $\alpha = 0.05$
-and $n = m$ equals the number of HLA haplotypes.
-
-\begin{equation}
-   D_{n,m} > \frac{1}{\sqrt{n}} \cdot \sqrt{ -\ln(\frac{\alpha}{2}) \cdot \frac{1 + \frac{n}{m}}{2} }
-   \label{eq:ks}
-\end{equation}

### 2020-09-08 bbbq_article.tex
Mention NSA gap, fix #75

@@
 
+Evolutionary conservation scores are shown in 
+figures \ref{fig:evolutionary_conservation} (topology)
+and \ref{fig:evolutionary_conservation_boxplot} (boxplot)
+for proteins that have at least one TMH (see figures
+\ref{fig:evolutionary_conservation_all} and 
+\ref{fig:evolutionary_conservation_all_boxplot}
+for the scores for all proteins).
+
 \begin{figure}[!htbp]
@@
   \caption{
-    Conservation scores across the SARS-CoV-2 proteome
-    for proteins that have at least one TMH
+    Conservation scores for the aligned SARS-CoV-2 proteomes,
+    for proteins that have at least one TMH.
+    The (non-logarithmic) scale of the conservation score has been removed
+    for readability.
+    Conservation scores do not start at residue 1 when
+    an aligned protein sequence outflanks the consensus sequence 
+    at the N-terminal
   }
@@
     for proteins that have at least one TMH
-  }
-  \label{fig:evolutionary_conservation_boxplot}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_all.png}
-  \caption{
-    Conservation scores across the SARS-CoV-2 proteome
-    for all proteins
-  }
-  \label{fig:evolutionary_conservation}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_boxplot_all.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    for all proteins
   }
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
+  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/scores_all.png}
   \caption{
-    Evolutionary conservation
+    Conservation scores for the aligned SARS-CoV-2 proteomes,
+    for all proteins.
+    The (non-logarithmic) scale of the conservation score has been removed
+    for readability.
+    Conservation scores do not start at residue 1 when
+    an aligned protein sequence outflanks the consensus sequence 
+    at the N-terminal.
   }
-  \label{fig:evolutionary_conservation_tmh}
+  \label{fig:evolutionary_conservation_all}
 \end{figure}
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_all.png}
+  \includegraphics[width=\textwidth]{bbbq_2/scores_boxplot_all.png}
   \caption{
-    Evolutionary conservation
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    for all proteins
   }
-  \label{fig:evolutionary_conservation_all}
+  \label{fig:evolutionary_conservation_all_boxplot}
 \end{figure}

### 2020-09-09 bbbq_article.tex
Add normalized perc binders that is TMH to manuscript, fix https://github.com/richelbilderbeek/bbbq_article/issues/90

@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc1_normalized.png}
+  \caption{
+    Normalized proportion of MHC-I epitopes overlapping with TMHs
+    for a human, viral and bacterial proteome.
+  }
+  \label{fig:f_tmh_mhc1_normalized}
+\end{figure}
+
 % Percentage of MHC-I 9-mers overlapping with TMH. Values in brackets
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc2_normalized.png}
+  \caption{
+    Normalized proportion of MHC-II epitopes overlapping with TMHs
+    for a human, viral and bacterial proteome.
+  }
+  \label{fig:f_tmh_mhc2_normalized}
+\end{figure}
 

### 2020-09-10 bbbq_article.tex
Show epitope prediction by MHCnuggets

@@
     for a human, viral and bacterial proteome.
+    Epitope prediction by EpitopePrediction.
   }
@@
     for a human, SARS-CoV-2 (labeled 'covid') 
-    and Mycobacterium tuberculosis (labeled 'myco') proteome. 
+    and Mycobacterium tuberculosis (labeled 'myco') proteome,
+    Epitope prediction by NetMHCIIpan.
   }
@@
 \begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc1_grid.png}
+  \caption{
+    Percentage of MHC-I epitopes overlapping with TMHs
+    for a human, viral and bacterial proteome.
+    Epitope prediction by MHCnuggets.
+  }
+  \label{fig:1_mhcnuggets}
+\end{figure}
+
+\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc1_normalized.png}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc2_grid.png}
+  \caption{
+    \% epitopes overlapping with transmembrane helix
+    for a human, SARS-CoV-2 (labeled 'covid') 
+    and Mycobacterium tuberculosis (labeled 'myco') proteome,
+    Epitope prediction by MHCnuggets.
+  }
+  \label{fig:2_mhcnuggets}
+\end{figure}
 

### 2020-09-10 bbbq_article.tex
Figures 1 and 2 refer to tables

@@
 Figure \ref{fig:1} shows the percentages of MHC-I epitopes overlapping 
-with TMHs for our human, viral and bacterial proteome.
+with TMHs for our human, viral and bacterial proteome,
+when using EpitopePrediction to predict epitopes.
+See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
+for the TMH and epitope counts.
 
-The KS test to determine if MHC-I has the same percentage 
-of epitopes overlapping with viral TMHs, 
-compared to human TMHs \richel{$\mathcal{H}_{1,V}$},
-resulted in a p-value of \richel{unknown},
-which makes us [reject/accept] the hypothesis that these percentages
-are sampled from the same distribution.
+% The KS test to determine if MHC-I has the same percentage 
+% of epitopes overlapping with viral TMHs, 
+% compared to human TMHs \richel{$\mathcal{H}_{1,V}$},
+% resulted in a p-value of \richel{unknown},
+% which makes us [reject/accept] the hypothesis that these percentages
+% are sampled from the same distribution.
 
-The KS test to determine if MHC-I has the same percentage 
-of epitopes overlapping with bacterial TMHs, 
-compared to human TMHs \richel{$\mathcal{H}_{1,B}$},
-resulted in a p-value of \richel{unknown},
-which makes us [reject/accept] the hypothesis that these percentages
-are sampled from the same distribution.
+% The KS test to determine if MHC-I has the same percentage 
+% of epitopes overlapping with bacterial TMHs, 
+% compared to human TMHs \richel{$\mathcal{H}_{1,B}$},
+% resulted in a p-value of \richel{unknown},
+% which makes us [reject/accept] the hypothesis that these percentages
+% are sampled from the same distribution.
 
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc1_grid.png}
+  \includegraphics[height=0.9\textheight]{bbbq_1_fast/fig_f_tmh_mhc1_grid.png}
   \caption{
@@
     Epitope prediction by EpitopePrediction.
+    See table \ref{tab:tmh_binders_mhc1} for the TMH and epitope counts.
   }
@@
 with TMHs for our human, viral and bacterial proteome.
-See the supplementary materials (tables \ref{tab:tmh_binders_mhc1} 
-and \ref{tab:tmh_binders_mhc2}) for the TMH and epitope counts.
+See the supplementary materials (table \ref{tab:tmh_binders_mhc2}) 
+for the TMH and epitope counts.
 
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc2_grid.png}
+  \includegraphics[height=0.9\textheight]{bbbq_1_fast/fig_f_tmh_mhc2_grid.png}
   \caption{
@@
     Epitope prediction by NetMHCIIpan.
+    See table \ref{tab:tmh_binders_mhc2} for the TMH and epitope counts.
   }

### 2020-09-15 bbbq_article.tex
Remove MHCnuggets, fix #101

@@
     Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
-    Predict epitopes MHC-I and MHC-II & MHCnuggets                & \cite{shao2020high} \\
     Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
@@
     Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
-    Call MHCnuggets from R            & mhcnuggetsr               & \cite{mhcnuggetsr} \\
     \verb;epitope-prediction; LUT     & epiprepreds               & \cite{epiprepreds} \\
     NetMHCIIpan LUT                   & nmhc2ppreds               & \cite{nmhc2ppreds} \\
-    MHCnuggets LUT                    & mhcnpreds                 & \cite{mhcnpreds} \\
     Sequence alignment                & ClustalW                  & \cite{thompson1994clustal} \\
@@
 which uses a stabilized matrix method (\cite{kim2009derivation}).
-The most recent FOSS tool available is 
-MHCnuggets (\cite{shao2020high}) (which can do both MHC-I 
-and MHC-II prediction) which is used in this study
+The most recent and promising FOSS tool available now appears
+to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
+and MHC-II prediction.
 
@@
 gave rise to the most accurate predictions.
-Later tools are 
-\richel{name of tool here} (\cite{zhang2013predicting}),
-\richel{name of tool here} (\cite{trolle2014nettepi}),
-\richel{name of tool here} (\cite{zhang2015accurate})
-and the very recent MHCnuggets (\cite{shao2020high}),
-the latter which is used in this study.
+Later tools are NetTepi (\cite{trolle2014nettepi}),
+a genetic-algorithm-based ensemble method (\cite{zhang2015accurate})
+and (again) the very recent MHCnuggets (\cite{shao2020high}).
 
@@
 which is dependent on 'pureseqtmr', 
-'mhcnuggetsr' and 'mhcnpreds',
 'epitope-prediction' and 'epiprepreds',
@@
 
-The R package 'mhcnuggetsr' is written to call the 
-'MHCnuggets' Python package.
-'mhcnuggetsr' allows to install 'MHCnuggets' and call it
-to predict IC50 values for multiple MHC haplotypes that 'MHCnuggets' 
-has already been trained upon.
-Excluded is the feature to train 'MHCnuggets' on new data and/or other
-haplotypes, as this study only needs IC50 predictions.
-'mhcnuggetsr' has been submitted to CRAN.
-
 To be able to use pre-calculated 
-\verb;epitope-prediction;, NetMHCIIpan and MHCnuggets
+\verb;epitope-prediction; and NetMHCIIpan
 predictions (i.e. a distribution of IC50 values per haplotype, 
-derived from random peptides), the 'epiprepreds', 'nmhc2ppreds' 
-and 'mhcnpreds' R package have been 
+derived from random peptides), the 'epiprepreds' and 'nmhc2ppreds' 
+R packages have been 
 developed. In practice, these are nothing more than a big look-up table.
@@
 If 'netmhc2pan' is allowed onto CRAN, 'nmhc2ppreds'
-will be submitted as well.
-If 'mhcnuggetsr' is put onto CRAN, 'mhcnpreds'
 will be submitted as well.
@@
 
-The earlier study uses a hand-crafted method, 
-that has been trained on MHC-I haplotypes only.
-This study uses MHCnuggets, which (1) free and open source, and (2) predicts
-IC50s on both MHC-I and MHC-II.
+The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
+that has been trained on MHC-I haplotypes only,
+which is used here again. For MHC-II IC50 predictions, the
+'netmhc2pan' R package is used, which calls NetMHC2pan.
+Both are packages are free and open-source.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc1_grid.png}
-  \caption{
-    Percentage of MHC-I epitopes overlapping with TMHs
-    for a human, viral and bacterial proteome.
-    Epitope prediction by MHCnuggets.
-  }
-  \label{fig:1_mhcnuggets}
-\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1/fig_f_tmh_mhc2_grid.png}
-  \caption{
-    \% epitopes overlapping with transmembrane helix
-    for a human, SARS-CoV-2 (labeled 'covid') 
-    and Mycobacterium tuberculosis (labeled 'myco') proteome,
-    Epitope prediction by MHCnuggets.
-  }
-  \label{fig:2_mhcnuggets}
-\end{figure}
 
@@
 Per MHC haplotype, we simulated random peptides and measured the
-predicted IC50 using MHCnuggets. 
+predicted IC50 using \verb;epitope-prediction; (for MHC-I) 
+and netmhc2pan (for MHC-II). 
 Here, we show the IC50 value per haplotype that

### 2020-09-15 bbbq_article.tex
Use violin plot, fix #94

@@
 figures \ref{fig:evolutionary_conservation} (topology)
-and \ref{fig:evolutionary_conservation_boxplot} (boxplot)
+and \ref{fig:evolutionary_conservation_violin} (boxplot)
 for proteins that have at least one TMH (see figures
 \ref{fig:evolutionary_conservation_all} and 
-\ref{fig:evolutionary_conservation_all_boxplot}
+\ref{fig:evolutionary_conservation_all_violin}
 for the scores for all proteins).
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_boxplot.png}
+  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
   \caption{
@@
   }
-  \label{fig:evolutionary_conservation_boxplot}
+  \label{fig:evolutionary_conservation_violin}
 \end{figure}
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_boxplot_all.png}
+  \includegraphics[width=\textwidth]{bbbq_2/scores_violin_all.png}
   \caption{
@@
   }
-  \label{fig:evolutionary_conservation_all_boxplot}
+  \label{fig:evolutionary_conservation_all_violin}
 \end{figure}

### 2020-09-15 bbbq_article.tex
Show conservation scores grouped for all TMH proteins, fix #93

@@
 Evolutionary conservation scores are shown in 
-figures \ref{fig:evolutionary_conservation} (topology)
-and \ref{fig:evolutionary_conservation_violin} (boxplot)
-for proteins that have at least one TMH (see figures
+figures \ref{fig:evolutionary_conservation} (topology),
+\ref{fig:evolutionary_conservation_violin} (per 
+protein with at least 1 TMH, violin)
+and \ref{fig:evolutionary_conservation_merged} (grouped 
+from all proteins that have at least 1 TMH).
+See figures
 \ref{fig:evolutionary_conservation_all} and 
@@
   \label{fig:evolutionary_conservation}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    with conservation scores group from all proteins that have at least one TMH
+  }
+  \label{fig:evolutionary_conservation_merged}
 \end{figure}

### 2020-09-23 bbbq_article.tex
Process feedback @bogeert, progress https://github.com/richelbilderbeek/bbbq_article/issues/69, move text about HLA classification to https://github.com/richelbilderbeek/bbbq_article/issues/110

@@
 bind to histocompatibility complex (MHC) class I more often then expected
-by chance only.
-This finding hints that there has been selection upon detecting TMH-derived
+by chance only, which can be mostly explained by the low hydrophobicity
+of those residues.
+The physiological reason is yet unclear,
+yet this finding hints that there has been selection upon detecting TMH-derived
 peptides in pathogens.
@@
 signal in the evolutionary history of the pathogen.
+The impact of this is [revolutionary/awesome].
 
@@
 
-\paragraph{Immune response}
+% \paragraph{Immune response}
 
-Our immune system fights invaders on a daily basis.
-These invaders can be fungi, bacteria or viruses.
-The innate immune response is its first general 
-and immediate strategy, where the acquired immune response
+Our immune system fights microbial pathogens on a daily basis,
+such as fungi, bacteria or viruses.
+The acquired immune response
 needs time to develop its specialized and more effective
-combat forces.
+combat forces and learn to recognize the pathogens 
+by use of the Major Histocompatibility Complex (MHC).
 
-\paragraph{Immune response by MHC-I}
+% \paragraph{Immune response by MHC-I}
 
-All nucleated cells in humans present randomly sampled peptides
-fragments to the surroundings of the cell using the Major 
-Histocompatibility Complex (MHC) class I molecules.
-If a cell gets infected by a virus, also the virus' peptides
-will be presented at the cell surface. The foreign viral antigen is detected 
+Antigen-presenting cells such as dendritic cells
+use the MHC class I molecules to 
+present randomly sampled peptides fragments,
+as created by the immunoproteasome (\cite{chapiro2006destructive}).
+If a cell gets infected by an intracellular pathogen, also the virus' peptides
+will also be presented at the cell surface. The foreign viral antigen is detected 
 by cytotoxic T lymphocytes, which will kill the infected cell.
 
-\paragraph{Immune response by MHC-II}
+% \paragraph{Immune response by MHC-II, contrast with APC}
 
+\richel{Weak, add refs}
 All pathogens can be detected by the foreign proteins on 
@@
 For humans, the MHC proteins are encoded by the
-HLA ('Human Leukocyte Antigens') genes.
+HLA (Human Leukocyte Antigens) genes.
 Each MHC complex can only bind a subset of all possible peptides.
@@
 making it hard to classify all of the many haplotypes.
-Classification of HLAs is based on algorithms that
-maximize the information content of each 
-classification, such as the presence of a certain amino acid at 
-the first (for MHC-II, \cite{southwood1998several})
-or second position (for MHC-I, \cite{lund2004definition}) of a peptide.
-However, this does not imply that when a peptide follows
-a classification pattern, that it will bind to the MHC. 
-Instead, software should be used to predict if the
-full peptide will do so.
 

### 2020-09-23 bbbq_article.tex
Process feedback @bogeert, progress https://github.com/richelbilderbeek/bbbq_article/issues/69

@@
 
-\paragraph{Classification of HLA}
+% \paragraph{Classification of HLA}
 
@@
 HLA (Human Leukocyte Antigens) genes.
+There are three genes encoding for MHC-I, which are HLA-A, HLA-B and HLA-C,
+for MHC-II there are three major genes, which are HLA-DR, HLA-DQ, HLA-DP.
 Each MHC complex can only bind a subset of all possible peptides.
@@
 peptides they bind (\cite{lund2004definition}).
-The HLA region of humans is highly polymorphic, 
-making it hard to classify all of the many haplotypes.
+The HLA region of humans is highly polymorphic, with hundreds
+to thousands of alleles for a gene (\cite{marsh2010nomenclature}).
 
-\paragraph{Epitope prediction}
+% \paragraph{HLAs increase detection range}
+
+Because of these multiple and highly polymorphic MHC genes,
+the number of pathogenic peptides that can be detected is increased,
+as a wide variety of MHCs will be expressed,
+each presenting their own subset of epitopes.
+
+% \paragraph{Epitope prediction}
 
@@
 
-\paragraph{TMHs}
+% \paragraph{TMHs}
+
+As the immunoproteasome is a cytosolic protein complex, on expects it
+to cleave only cytosolic proteins. However, it has recently been
+discovered that transmembrane protein residues are also presented
+by MHC-I complexes (\cite{bianchi2017}). This discovery is based
+on both in vitro elution studies, as well as on in silico predictions,
+with a focus on transmembrane helices. 
+The intracellular pathway, in which transmembrane residues are loaded onto
+an MHC, however, is yet unknown.
 
@@
 TMHs are general structures: 25 percent of the human proteome is
-anchored by at least one TMH. SARS-CoV-2 has 21 TMHs, 
+anchored by at least one TMH. Also the pathogen SARS-CoV-2 has 21 TMHs, 
 making up for 5.5 percent of its entire genome.
 There are multiple computational tools developed to predict which
-parts of membrane proteins are TMH.
+parts of membrane proteins are TMH (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
 
-\paragraph{MHC-I presents hydrophobic regions more often}
-
-One might expect that the hydrophobic epitopes presented,
-for a same hydrophobicity, are as likely to stem from 
-membrane proteins TMHs or from soluble proteins hydrophobic (non-TMH) regions.
-For MHC-I, however, it is found that the 9-mers stemming
-from TMHs are presented more often than expected by 
-chance (\cite{bianchi2017})
-\richel{but also when compensating for hydrophobicity?}.
-For MHC-II, it is unknown which percentage of binders 
-is derived from TMHs for all 13-mers present in a proteome
-that have the same hydrophobicity.
-
-\paragraph{HLAs increase detection range}
-
-One of many strategies to increase the number of pathogenic
-peptides that can be detected is to express a wide variety of
-MHCs, which is assured by the human leukocyte antigen gene complex (HLA).
-However, a different HLA will result in a different variety of MHCs,
-that will display different peptide fragments. Additionally, there
-may have been selection the HLA to display and or detect different
-peptide fragments.
-
-\paragraph{MHC-I presents TMH-derived epitopes in humans more often}
+% \paragraph{MHC-I presents TMH-derived epitopes in humans more often}
 
@@
 are over-presented by all 5 HLA-A and 
-most of 8 HLA-B super types (\cite{bianchi2017}).
-One explanation is that the presentation of TMHs 
-may have an evolutionary advantage for 
-the (human) host, as TMHs have a reduced variability 
-due to the functional requirement of being able to span a lipid bilayer. 
-Due to this, pathogens have a lower chance to develop an escape mutation,
+most of 8 HLA-B super-types (\cite{bianchi2017}),
+due the high hydrophobicity as is required for a TMH to span
+the (hydrophobic) cell membrane, combined with the fact that the MHC
+binding cleft is hydrophobic as well.
+
+It is unknown if the presentation of hydrophobic/TMH residues
+is evolutionary selected for.
+One possible explanation is that TMHs have a reduced variability,
+imposed by the functional requirement of being able to span a lipid bilayer.
+Due to this, the TMHs that pathogens possess, 
+have a lower chance to develop an escape mutation,
 as many mutations will result in a dysfunctional TMH.
-Note that the mechanism by which a cell presents its TMHs is
-yet unknown.
 

### 2020-09-23 bbbq_article.tex
Incorrect: Geert enjoyed 'Use of protein data in phylogenetics', fix https://github.com/richelbilderbeek/bbbq_article/issues/112

@@
 
-\paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
+% \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
 
@@
 
-\paragraph{MHC-II is expected to present TMHs}
+% \paragraph{MHC-II is expected to present TMHs}
 
@@
 
-\paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
+% \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
 
@@
 
-\paragraph{Selection undetectable in whole proteome}
-
-The human immune system and human pathogen are in an evolutionary
-arms race: our immune systems is selected for the detection
-of pathogens, whereas pathogens are selected to avoid detection.
-From a pathogen's point of view, however, this struggle 
-is of only minor importance:
-in seasonal influenza, for example, the selection pressure
-exerted by the immune system was only limited (\cite{han2019individual}).
+% \paragraph{Selection undetectable in whole proteome}
 
@@
 
-\paragraph{Immunodominance}
-
-The effective number of epitopes, however,
-is much lower than the actual number of possible
-epitopes, due to immunodominance. 
-Immunodominance is a feature of the immune system,
-due to which some epitopes dominate in causing an immune response,
-where other (called subdominant) epitopes end up having no 
-effect (\cite{akram2012immunodominance}).
-The same study mentions 20 factors that influence the strength
-of an immune response.
-The most relevant factor for this study is the effect of
-a mutation within an epitope, 
-such as the in vitro example described in (\cite{berkhoff2004mutation}),
-where one single mutation resulted in a 5-20\% weaker
-response of virus-specific CD8+ T cells \richel{Fun experiment:
-do this in silico}.
-Due to immunodominance, we dare assume that there can be selection
-for immune system avoidance, as the effective number of loci may be
-low and effects of a single mutation noticeable.
-
 \paragraph{Use of protein data in phylogenetics}
@@
 as its is closer to the actual phenotype
-selection acts upon: DNA may never be translated to RNA,
-or its RNA may never be transcribed (\cite{diz2012proteomics})
+selection acts upon: DNA may never be transcribed to RNA,
+or its RNA may never be translated (\cite{diz2012proteomics})
 \richel{improve upon this point}.
@@
 
+The human immune system and human pathogen are in an evolutionary
+arms race: our immune systems is selected for the detection
+of pathogens, whereas pathogens are selected to avoid detection.
+From a pathogen's point of view, however, this struggle 
+is of only minor importance:
+in seasonal influenza, for example, the selection pressure
+exerted by the immune system was only limited (\cite{han2019individual}).
+
 \richel{

### 2020-09-23 bbbq_article.tex
Remember: immunodominance, note at #111. Remember: Selection needs to be additive, note at #113

@@
 can be detected.
-
-\paragraph{Selection needs to be additive}
-
-Would a pathogen have only peptides that do not bind to
-any HLA, a pathogen would be undetected. In that case, one
-can imagine that a mutation
-that causes a pathogen to become detected by a host's immune
-system, may be selected against. Also, would a pathogen have only
-one epitope, one can imagine there would be selection for a mutation to lose it.
-If there are, however, already many epitopes present in the
-proteome, the selection to lose or gain an epitope is expected 
-to have less or no effect.
 

### 2020-09-23 bbbq_article.tex
Move hypotheses, thanks Geert
Processed feedback, thanks @bogeert

@@
 
-\paragraph{Selection may be detectable in TMHs}
+% \paragraph{Selection may be detectable in TMHs}
 
@@
 
-\paragraph{Use of protein data in phylogenetics}
+% \paragraph{Use of protein data in phylogenetics}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Hypotheses}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{Will be moved to supplementary materials}
-
-\begin{itemize}
-  \item $\mathcal{H}_{1,VB}$: MHC-I has the same percentage of epitopes overlapping
-    with viral/bacterial TMHs as with human TMHs
-  \item $\mathcal{H}_{2,HVB}$: MHC-II has the same percentage of epitopes overlapping
-    with human/viral/bacterial TMHs as expected by chance
-  \item $\mathcal{H}_{3}$: TMH and non-TMH regions are equally conserved
-    in a proteome
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
@@
 
-\paragraph{Data sets for TMH epitopes}
+% \paragraph{Data sets for TMH epitopes}
 
 To determine the percentages of epitopes overlapping
-with TMHs, we use three reference proteomes, which are
-a human, viral and bacterial one.
+with TMHs, we use reference proteomes of 
+a human, viral and bacterial nature.
 For a human reference proteome, we used the Uniprot human reference proteome
 from \url{https://www.uniprot.org/proteomes/UP000005640}), 
-which is the same as \cite{bianchi2017}.
+which is the same as used previously (\cite{bianchi2017}).
 The viral reference proteome is represented by
@@
     for a human, viral and bacterial proteome.
-    Epitope prediction by EpitopePrediction.
+    The horizontal lines indicate the percentage as expected by chance.
+    \richel{Refer to a table with the numbers as expected by chance)}
     See table \ref{tab:tmh_binders_mhc1} for the TMH and epitope counts.
+    \richel{Epitope prediction by EpitopePrediction.}.
+    Note that for smaller proteomes a percentage of zero is likelier.
+    \richel{use of top 2\% binders}
   }
@@
   \caption{
-    \% epitopes overlapping with transmembrane helix
-    for a human, SARS-CoV-2 (labeled 'covid') 
-    and Mycobacterium tuberculosis (labeled 'myco') proteome,
-    Epitope prediction by NetMHCIIpan.
+    Percentage of MHC-II epitopes overlapping with TMHs
+    for a human, viral and bacterial proteome.
+    The horizontal lines indicate the percentage as expected by chance.
+    \richel{Refer to a table with the numbers as expected by chance)}
+    \richel{Epitope prediction by NetMHCIIpan.}
     See table \ref{tab:tmh_binders_mhc2} for the TMH and epitope counts.
+    Note that for smaller proteomes a percentage of zero is likelier.
+    \richel{use of top 2\% binders}
   }
@@
 \end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Hypotheses}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{As a reminder, to focus the research on}
+
+\begin{itemize}
+  \item $\mathcal{H}_{1,VB}$: MHC-I has the same percentage of epitopes overlapping
+    with viral/bacterial TMHs as with human TMHs
+  \item $\mathcal{H}_{2,HVB}$: MHC-II has the same percentage of epitopes overlapping
+    with human/viral/bacterial TMHs as expected by chance
+  \item $\mathcal{H}_{3}$: TMH and non-TMH regions are equally conserved
+    in a proteome
+\end{itemize}

### 2020-09-23 bbbq_article.tex
Reviewed supmat, progress #69

@@
     Normalized proportion of MHC-I epitopes overlapping with TMHs
-    for a human, viral and bacterial proteome.
+    for human, viral and bacterial proteomes.
+    Legend: covid = SARS-CoV-2, flua = influenza A, hepa = hepatitis A,
+    human = homo sapiens, myco = Mycobacterium tuberculosis, 
+    polio = poliovirus, rhino = rhinovirus
   }
@@
     Normalized proportion of MHC-II epitopes overlapping with TMHs
-    for a human, viral and bacterial proteome.
+    for human, viral and bacterial proteomes.
+    Legend: covid = SARS-CoV-2, flua = influenza A, hepa = hepatitis A,
+    human = homo sapiens, myco = Mycobacterium tuberculosis, 
+    polio = poliovirus, rhino = rhinovirus
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Relation between haplotype and TMH over-representation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/haplotype_perc_vs_f_tmh_binders.png}
-  \caption{
-    Relation between genotype frequency and the percentage
-    of binders that are TMH for the 21 MHC-II haplotypes
-  }
-  \label{fig:haplotype_perc_vs_f_tmh_binders}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/phenotype_freq_vs_f_tmh_binders.png}
-  \caption{
-    Relation between phenotype frequency and the percentage
-    of binders that are TMH for the 21 MHC-II haplotypes
-  }
-  \label{fig:phenotype_freq_vs_f_tmh_binders}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{IC50s per haplotype}
@@
 is used to determine if a peptide binds to the haplotype's MHC.
+Included in this figure is the MHCnuggets prediction tool,
+by \cite{shao2020high}.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Relation between haplotype frequency and TMH over-representation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+There is no relation between haplotype frequency and TMH over-representation.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/haplotype_perc_vs_f_tmh_binders.png}
+  \caption{
+    Relation between genotype frequency and the percentage
+    of binders that are TMH for the 21 MHC-II haplotypes
+  }
+  \label{fig:haplotype_perc_vs_f_tmh_binders}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/phenotype_freq_vs_f_tmh_binders.png}
+  \caption{
+    Relation between phenotype frequency and the percentage
+    of binders that are TMH for the 21 MHC-II haplotypes
+  }
+  \label{fig:phenotype_freq_vs_f_tmh_binders}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Hypotheses}

### 2020-09-23 bbbq_article.tex
Mention ClustalOmega and BLOSUM80

@@
 using the msa (\cite{bodenhofer2015msa}) R package, 
-using default settings, including the use of the default alignment tool
-ClustalW (\cite{thompson1994clustal}).
-Instead of the default and general BLOSUM62 amino acid transition matrix (\cite{henikoff1992amino}),
-the FLU transition matrix (\cite{dang2010flu}) was used,
-as the latter is based on amino acid transitions in another viral pathogen, influenza.
+set up to use ClustalOmega (\cite{thompson1994clustal})
+with the BLOSUM80 amino acid transition matrix (\cite{henikoff1992amino}).
 Per alignment, we used msa to determine the consensus sequence

### 2020-09-23 create_table_reference_proteomes/create.R
Create table with Uniprot IDs, progress https://github.com/richelbilderbeek/bbbq_article/issues/114

@@ -1,0 +1,24 @@
+# Create a table with the reference proteomes
+library(dplyr, warn.conflicts = FALSE)
+
+t <- tibble::tibble(
+  target = bbbq::get_target_names(),
+  name = NA,
+  uniprot_id = NA
+)
+
+for (i in seq_len(nrow(t))) {
+ t$name[i] <- bbbq::get_target_english_name(t$target[i])
+ t$uniprot_id[i] <- bbbq::get_uniprot_id(t$target[i])
+}
+
+readr::write_csv(t, "uniprot_ids.csv")
+
+knitr::kable(
+  t %>% dplyr::select(name, uniprot_id),
+  "latex",
+  caption = paste0(
+    "Overview of reference proteomes and Uniprot IDs"
+  ),
+  label = "uniprot_ids"
+) %>% cat(., file = "uniprot_ids.latex")

### 2020-09-23 bbbq_article.tex
Add table for reference proteomes, fix #114

@@
 a human, viral and bacterial nature.
-For a human reference proteome, we used the Uniprot human reference proteome
-from \url{https://www.uniprot.org/proteomes/UP000005640}), 
-which is the same as used previously (\cite{bianchi2017}).
-The viral reference proteome is represented by
-the proteome of the first sequenced SARS-CoV-2 strain (\cite{wu2020new},
-GenBank ID of MN908947.3, \url{https://www.ncbi.nlm.nih.gov/nuccore/MN908947})
-For bacteria, we used the Uniprot Mycobacterium tuberculosis reference genome
-(\url{https://www.uniprot.org/proteomes/UP000001584}) is used.
+An overview of Uniprot IDs is shown in table \ref{tab:uniprot_ids}.
+The human reference proteome is the same as used previously (\cite{bianchi2017}).
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Uniprot IDs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% tab:uniprot_ids
+\input{create_table_reference_proteomes/uniprot_ids.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-I}

### 2020-10-28 fix_table_tmh_binders_mhc.sh
Tables fit on paper

@@ -1,0 +1,59 @@
+#!/bin/bash
+
+# The LaTeX files
+#
+#  * bbbq_1_fast/table_tmh_binders_mhc1.latex
+#  * bbbq_1_fast/table_tmh_binders_mhc2.latex
+#
+# both are a table with more columns than fit on a page.
+# 
+# This script converts these to a format that fits on a page.
+#
+# The original format:
+#
+# ```
+# \begin{table}
+# \caption{\label{tab:tab:ic50_binders}IC50 values (in nM) per haplotype below which a peptide is considered a binder. Percentile used: 0.02}
+# \centering
+# \begin{tabular}[t]{l|r}
+#   .. data
+# \end{tabular}
+# \end{table}
+# ```
+
+# A sidewaystable is not prettty
+#
+# %\begin{sidewaystable}
+# %  \centering
+# %  ... centered table here
+# %\end{sidewaystable}
+#
+# sed -i 's/\\begin{table}/\\begin{sidewaystable}/' bbbq_1_fast/table_tmh_binders_mhc1.latex
+# sed -i 's/\\end{table}/\\end{sidewaystable}/' bbbq_1_fast/table_tmh_binders_mhc1.latex
+# sed -i 's/\\begin{table}/\\begin{sidewaystable}/' bbbq_1_fast/table_tmh_binders_mhc2.latex
+# sed -i 's/\\end{table}/\\end{sidewaystable}/' bbbq_1_fast/table_tmh_binders_mhc2.latex
+
+
+
+# 
+# % \scalebox{0.7}{
+# %  ... scaled table here
+# % }
+
+# Only tiny still does not fit
+# 
+# %{\tiny
+# %  ... something with a tiny font
+# %}
+#
+# sed -i 's/\\begin{table}/\\begin{table}{\\tiny/' bbbq_1_fast/table_tmh_binders_mhc1.latex
+# sed -i 's/\\end{table}/}\\end{table}/' bbbq_1_fast/table_tmh_binders_mhc1.latex
+# sed -i 's/\\begin{table}/\\begin{table}{\\tiny/' bbbq_1_fast/table_tmh_binders_mhc2.latex
+# sed -i 's/\\end{table}/}\\end{table}/' bbbq_1_fast/table_tmh_binders_mhc2.latex
+
+
[…]

### 2020-11-02 bbbq_article.tex
Tables fit on paper
Show 8k random sequences

@@
 Evolutionary conservation scores are shown in 
-figures \ref{fig:evolutionary_conservation} (topology),
-\ref{fig:evolutionary_conservation_violin} (per 
+figures \ref{fig:evolutionary_conservation_8k} (topology),
+\ref{fig:evolutionary_conservation_8k_violin} (per 
 protein with at least 1 TMH, violin)
-and \ref{fig:evolutionary_conservation_merged} (grouped 
+and \ref{fig:evolutionary_conservation_8k_merged} (grouped 
 from all proteins that have at least 1 TMH).
 See figures
-\ref{fig:evolutionary_conservation_all} and 
-\ref{fig:evolutionary_conservation_all_violin}
+\ref{fig:evolutionary_conservation_8k_all} and 
+\ref{fig:evolutionary_conservation_8k_all_violin}
 for the scores for all proteins).
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
+  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores.png}
   \caption{
     Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for proteins that have at least one TMH.
+    for proteins that have at least one TMH,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
     The (non-logarithmic) scale of the conservation score has been removed
@@
   }
-  \label{fig:evolutionary_conservation}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
+  \label{fig:evolutionary_conservation_8k}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_merged_violin.png}
   \caption{
@@
     residues in the SARS-CoV-2 proteome,
-    with conservation scores group from all proteins that have at least one TMH
-  }
-  \label{fig:evolutionary_conservation_merged}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
+    with conservation scores group from all proteins that have at least one TMH,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
+  }
+  \label{fig:evolutionary_conservation_8k_merged}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin.png}
   \caption{
@@
     residues in the SARS-CoV-2 proteome,
-    for proteins that have at least one TMH
-  }
-  \label{fig:evolutionary_conservation_violin}
+    for proteins that have at least one TMH,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
+  }
+  \label{fig:evolutionary_conservation_8k_violin}
 \end{figure}
@@
 % Label: tab:tmh_binders_mhc1
+%\begin{sidewaystable}
+%  \centering
+%  ... centered table here
+%\end{sidewaystable}
+
+% \scalebox{0.7}{
+%  ... scaled table here
+% }
+
+%{\tiny
+%  ... something with a tiny font
+%}
+
 \input{bbbq_1_fast/table_tmh_binders_mhc1.latex}
@@
 
+%\begin{sidewaystable}
+%  \centering
   \input{bbbq_1_fast/table_tmh_binders_mhc2.latex}
+%\end{sidewaystable}
 
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/scores_all.png}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Subset of random sequences}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores.png}
   \caption{
     Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for all proteins.
+    for proteins that have at least one TMH,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
+    The (non-logarithmic) scale of the conservation score has been removed
+    for readability.
+    Conservation scores do not start at residue 1 when
+    an aligned protein sequence outflanks the consensus sequence 
+    at the N-terminal
+  }
+  \label{fig:evolutionary_conservation_8k}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_merged_violin.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    with conservation scores group from all proteins that have at least one TMH,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
+  }
+  \label{fig:evolutionary_conservation_8k_merged}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    for proteins that have at least one TMH,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
+  }
+  \label{fig:evolutionary_conservation_8k_violin}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/8000_42/scores_all.png}
+  \caption{
+    Conservation scores for the aligned SARS-CoV-2 proteomes,
+    for all proteins,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
     The (non-logarithmic) scale of the conservation score has been removed
@@
   }
-  \label{fig:evolutionary_conservation_all}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_violin_all.png}
+  \label{fig:evolutionary_conservation_8k_all}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin_all.png}
   \caption{
@@
     residues in the SARS-CoV-2 proteome,
-    for all proteins
-  }
-  \label{fig:evolutionary_conservation_all_violin}
+    for all proteins,
+    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
+  }
+  \label{fig:evolutionary_conservation_8k_all_violin}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{All unique sequences}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
+  \caption{
+    Conservation scores for the aligned SARS-CoV-2 proteomes,
+    for proteins that have at least one TMH,
+    using all unique sequences \richel{which I doubt is valid}.
+    The (non-logarithmic) scale of the conservation score has been removed
+    for readability.
+    Conservation scores do not start at residue 1 when
+    an aligned protein sequence outflanks the consensus sequence 
+    at the N-terminal
+  }
+  \label{fig:evolutionary_conservation_all_unique}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    with conservation scores group from all proteins that have at least one TMH,
+    using all unique sequences \richel{which I doubt is valid}.
+  }
+  \label{fig:evolutionary_conservation_all_unique_merged}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    for proteins that have at least one TMH,
+    using all unique sequences \richel{which I doubt is valid}.
+  }
+  \label{fig:evolutionary_conservation_all_unique_violin}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
+  \caption{
+    Conservation scores for the aligned SARS-CoV-2 proteomes,
+    for proteins that have at least one TMH,
+    using all unique sequences \richel{which I doubt is valid}.
+    The (non-logarithmic) scale of the conservation score has been removed
+    for readability.
+    Conservation scores do not start at residue 1 when
+    an aligned protein sequence outflanks the consensus sequence 
+    at the N-terminal
+  }
+  \label{fig:evolutionary_conservation_all_unique}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    with conservation scores group from all proteins that have at least one TMH,
+    using all unique sequences \richel{which I doubt is valid}.
+  }
+  \label{fig:evolutionary_conservation_all_unique_merged}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    for proteins that have at least one TMH,
+    using all unique sequences \richel{which I doubt is valid}.
+  }
+  \label{fig:evolutionary_conservation_all_unique_violin}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/scores_all.png}
+  \caption{
+    Conservation scores for the aligned SARS-CoV-2 proteomes,
+    for all proteins,
+    using all unique sequences \richel{which I doubt is valid}.
+    The (non-logarithmic) scale of the conservation score has been removed
+    for readability.
+    Conservation scores do not start at residue 1 when
+    an aligned protein sequence outflanks the consensus sequence 
+    at the N-terminal.
+  }
+  \label{fig:evolutionary_conservation_all_unique_all}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_2/scores_violin_all.png}
+  \caption{
+    Evolutionary conservation scores of TMH versus non-TMH
+    residues in the SARS-CoV-2 proteome,
+    for all proteins,
+    using all unique sequences \richel{which I doubt is valid}.
+  }
+  \label{fig:evolutionary_conservation_all_unique_all_violin}
 \end{figure}

### 2020-12-03 bbbq_article.tex
Determine if test is run on a CI server, instead of just Travis

@@
 Lastly, we show the TMHs are [more/less/equally] 
-conserved.
+conserved by determining if SNPs are found in TMH more/less/equal
+than expected by chance.
 Our findings suggest that the immune system is [less/neutral/more]
@@
 To detect the evolutionary conservation of TMHs, 
-we first selected all unique protein sequences per
-gene from all 45k SARS-CoV-2 proteomes,
-resulting in around 500 unique protein sequences per gene.
-From the unique protein sequences, we made an alignment per gene,
-using the msa (\cite{bodenhofer2015msa}) R package, 
-set up to use ClustalOmega (\cite{thompson1994clustal})
-with the BLOSUM80 amino acid transition matrix (\cite{henikoff1992amino}).
-Per alignment, we used msa to determine the consensus sequence
-and the conservation scores.
-Per consensus sequences, the topology was determined using pureseqtmr.
-
-\paragraph{Statistics}
-
-We compare the conservation values for TMH versus non-TMH locations,
-for all membrane proteins combined,
-using a Kolmogorov-Smirnov test, for a significance level of 5 percent.
+we use the follow workflow (also shown
+in figure \ref{fig:snp_workflow}) to be able to count the
+number of SNPs that occur in TMHs, as well as the number of SNPs
+that occur in TMHs by chance.
+
+To do so, we use the 
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases:
+'gene' to find the gene names of membane proteins, 
+'dbSNP' for SNPs associated with those genes
+and 'protein' to obtain the sequence of proteins that SNPs act upon.
+The 'dbSNP' contains 2 billion (human-only) SNPs, although
+it is estimated that approx 10 percent is a false positive 
+\richel{see references on dbSNP Wikipedia page}.
+
+The \verb;rentrez; R package provides for doing so from script.
+
+The first query is a call to the 'gene' database for the 
+term 'membrane protein' (in all fields) for the organism Homo sapiens.
+This results in \richel{1126} gene IDs.
+The next query is a call to the 'gene' database 
+to obtain the gene names from the gene IDs.
+Per gene name, the 'dbSNP' NCBI database is queried for 
+SNPs associated with the gene name. The number of SNPs
+was limited to the first \richel{100} SNPs per gene,
+resulting in \richel{112600} SNPs.
+Per selected SNP, all variation (i.e. descriptions of the changes on DNA
+and/or RNA and/or protein level) are obtained from the 'dbSNP' database
+as well. Of these variations, only the variations that result in a single
+amino acid substitution are kept, resulting in \richel{???} protein changes.
+Per protein substitution (i.e. a protein name and an index of the changed residue), 
+the 'protein' NCBI database was queried for the
+protein sequence.
+Of each protein sequence, the protein topology is determined 
+using PureseqTM.
+From the topology and the known location of the amino acid substitution, 
+it is tallied if the change occurred within a TMH or not.
+Also, the percentage of the topology that is TMH is measured.
+Not all membrane proteins have a transmembrane helix, resulting
+in \richel{???} proteins with at least one TMH.
+
+\begin{figure}[!htbp]
+  \includegraphics[height=0.9\textheight]{snp_workflow.png}
+  \caption{
+    Workflow to determine the evolutionary conservation of TMHs
+    using human SNPs.
+  }
+  \label{fig:snp_workflow}
+\end{figure}
+
+% \paragraph{Introduction}
+% 
+% To detect the evolutionary conservation of TMHs,
+% we first selected all unique protein sequences per
+% gene from all 45k SARS-CoV-2 proteomes,
+% resulting in around 500 unique protein sequences per gene.
+% From the unique protein sequences, we made an alignment per gene,
+% using the msa (\cite{bodenhofer2015msa}) R package, 
+% set up to use ClustalOmega (\cite{thompson1994clustal})
+% with the BLOSUM80 amino acid transition matrix (\cite{henikoff1992amino}).
+% Per alignment, we used msa to determine the consensus sequence
+% and the conservation scores.
+% Per consensus sequences, the topology was determined using pureseqtmr.
+% 
+% \paragraph{Statistics}
+% 
+% We compare the conservation values for TMH versus non-TMH locations,
+% for all membrane proteins combined,
+% using a Kolmogorov-Smirnov test, for a significance level of 5 percent.
 

### 2020-12-16 bbbq_article.tex
For 30 SNPs per gene, we know that TMHs are evolutionarily conserved

@@
 Additionally, MHC-II binds to peptides derived from TMHs 
-[less/equally/more] often than expected by chance.
-Lastly, we show the TMHs are [more/less/equally] 
-conserved by determining if SNPs are found in TMH more/less/equal
-than expected by chance.
-Our findings suggest that the immune system is [less/neutral/more]
-vigilant to TMHs than expected by chance and this has 
-left [a clear/a weak/no]
-signal in the evolutionary history of the pathogen.
-The impact of this is [revolutionary/awesome].
+more often than expected by chance.
+Lastly, we show the TMHs are evolutionarily more conserved,
+as less SNPs are found in TMH than expected by chance.
+Our findings suggest that the immune system is more
+vigilant to TMHs than expected by chance
+and suggests that an evolutionary signal in the evolutionary history of a pathogen
+may be found.
 
@@
 To detect the evolutionary conservation of TMHs, 
-we use the follow workflow (also shown
+we use the following workflow (also shown
 in figure \ref{fig:snp_workflow}) to be able to count the
@@
 NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases:
-'gene' to find the gene names of membane proteins, 
-'dbSNP' for SNPs associated with those genes
+'gene' to find the gene names of membrane proteins, 
+'dbSNP' (\cite{sachidanandam2001map}) for SNPs associated with those genes
 and 'protein' to obtain the sequence of proteins that SNPs act upon.
 The 'dbSNP' contains 2 billion (human-only) SNPs, although
-it is estimated that approx 10 percent is a false positive 
-\richel{see references on dbSNP Wikipedia page}.
+it is estimated that approximately ten percent (ranging from 2 to 36 percent) 
+are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}).
 
@@
 
+Regarding the evolutionary conservation of TMHs using SNPs,
+again, it is estimated that approx 8.32 percent
+of SNPs is caused, due to sequence variations in
+highly similar duplicated sequences \cite{musumeci2010single}.
+We assume that these duplications occur as often in TMHs as in
+regions around these, hence we expect this not to affect our results.
+
 \richel{

### 2020-12-16 bbbq_article.tex
Quote dpSNP correctly

@@
 'gene' to find the gene names of membrane proteins, 
-'dbSNP' (\cite{sachidanandam2001map}) for SNPs associated with those genes
+'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
 and 'protein' to obtain the sequence of proteins that SNPs act upon.
@@
 was limited to the first \richel{100} SNPs per gene,
-resulting in \richel{112600} SNPs.
+resulting in \richel{112600} SNPs 
+\richel{Note that \cite{musumeci2010single} demonstrates for a way to get
+try SNPs, instead of SNDs, maybe I should use that}
+.
 Per selected SNP, all variation (i.e. descriptions of the changes on DNA

### 2020-12-16 bbbq_article.tex
Add sampling

@@
 If this signature of selection is present in MHC-II is unknown.
-This study shows that MHC-I [also has/does not have] more
+This study shows that MHC-I also has more
 epitopes derived from a TMH for a pathogen proteome, when compared with
@@
 adaptive immunity, transmembrane domain, transmembrane helix, 
-epitopes, T lymphocyte, MHC-I, MHC-II, SARS-CoV-2
+epitopes, T lymphocyte, MHC-I, MHC-II, SARS-CoV-2, evolutionary conservation
 
@@
 Regarding the evolutionary conservation of TMHs using SNPs,
-again, it is estimated that approx 8.32 percent
-of SNPs is caused, due to sequence variations in
-highly similar duplicated sequences \cite{musumeci2010single}.
+again, it is estimated that approximately ten percent
+of SNPs is a false positive that result from the methods to determine
+a SNP. One example is that sequence variations are incorrectly
+detected due to highly similar duplicated sequences \cite{musumeci2010single}.
 We assume that these duplications occur as often in TMHs as in
 regions around these, hence we expect this not to affect our results.
+
+We used a low number of SNPs, as was imposed by the
+NCBI API, which constrains its users to three calls per second.
+As a second factor, we limited the time querying the NCBI database
+for the duration of the project.
+The SNPs we did use, we picked those
+with the lowest SNP ID (which denotes the earliest discovery date) first.
+We expect this not to introduce a bias, so we expect that
+the SNPs we used are a representative sample of all SNPs.
 

### 2021-01-19 bbbq_article.tex
Add scripts
Add selenoproteins

@@
 In 2001, multiple tools to do so have been compared (\cite{moller2001evaluation}),
-of which TMHMM (\cite{krogh2001predicting}) turned out to be the best.
-TMHMM is used in the previous study (\cite{bianchi2017}),
-yet TMHMM itself has a restrictive software license and is nearly two
+of which TMHMM (\cite{krogh2001predicting}) turned out to be the best, 
+as is used in the previous study (\cite{bianchi2017}).
+TMHMM has a restrictive software license and is nearly two
 decades old.
-Inspired on TMHMM, many other tools followed, such as 
-Phobius (\cite{kall2004combined}),
-ConPred II (\cite{arai2004conpred}),
-MEMSAT3 (\cite{jones2007improving})
-and MetaTM (\cite{klammer2009metatm}).
-The most recent FOSS tool is called PureseqTM (\cite{wang2019efficient}),
-which is used in this study.
+Due to this, also PureseqTM (\cite{wang2019efficient}),
+is used in this study, which has been recently developed
+and has a free software license.
 
@@
 
+Both TMH prediction tools struggle with selenocysteine,
+the 21st amino acid incorporated in human proteins.
+The human reference proteome used contains 30 selenoproteins.
+TMHMM gives a clear error message, due to which we removed
+these selenoproteins from the analysis.
+How PureseqTM deals with selenocysteine is unclear,
+as it gives no error, nor is it mentioned anywhere,
+nor do the authors respond to emails.
+We assume PureseqTM ignores these residues.
+In either case, we think that the effect of the 30 selenoproteins
+on a proteome of 20600 proteins is only minor.
+
 \richel{
@@
 for a binder per haplotype.
+
+\paragraph{Unclear handing of sequences with U}
+
+Within the human reference proteome, there are 30 sequences that have a
+selenocysteine (with the symbol 'U').
+See \url{https://github.com/richelbilderbeek/bbbq_article/issues/158}.
+
+\paragraph{Which reference proteome?}
+
+See \url{https://github.com/richelbilderbeek/bbbq_article/issues/155}.
 

### 2021-01-29 bbbq_article.tex
EOD

@@
 
-Both TMH prediction tools struggle with selenocysteine,
-the 21st amino acid incorporated in human proteins.
-The human reference proteome used contains 30 selenoproteins.
-TMHMM gives a clear error message, due to which we removed
+We have removed the 25 selenoproteins found in the human
+reference proteome, as both prediction tools struggle with 
+selenocysteine, the 21st amino acid incorporated in human proteins.
+TMHMM gives a clear error message, forcing us to remove
 these selenoproteins from the analysis.
@@
 as it gives no error, nor is it mentioned anywhere,
-nor do the authors respond to emails.
-We assume PureseqTM ignores these residues.
-In either case, we think that the effect of the 30 selenoproteins
-on a proteome of 20600 proteins is only minor.
+nor do the authors respond to emails. Due to this, we removed the
+selenoproteins from our analysis.
+Another bug in PureseqTM is that it does not predict the topology
+of proteins that have less than three amino acids. 
+The TRDD1_HUMAN protein, short for 'T cell receptor delta diversity 1',
+is, however, two peptides long. We manually added the prediction
+that both peptides reside in the cytosol.
 

### 2021-02-04 bbbq_article.tex
Add MHC-II elution study

@@
 An overview of Uniprot IDs is shown in table \ref{tab:uniprot_ids}.
-The human reference proteome is the same as used previously (\cite{bianchi2017}).
+Instead of using the full human reference proteome, we
+use the representative human reference proteome 
+as used previously (\cite{bianchi2017}).
 
@@
 as provided by GISAID (\cite{shu2017gisaid}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Elution studies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To determine if epitopes derived from TMHs are presented at all,
+we started from epitopes sequenced from elution studies
+for MHC-I \cite{schellens2015comprehensive} 
+and MHC-II \cite{bergseng2015different}.
+For each epitope, the human representative reference proteome
+was searched for its origin.
+We kept only the epitopes of which were found exactly once
+on the proteome.
+We predicted the topology of the proteome
+and counted how often the epitopes overlapped
+with a predicted TMH.
+In this analysis, both TMHMM and 
+PureseqTM (see section 'Prediction software used') 
+were used to predict the topology.
+The full analysis can be found
+at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.
 
@@
 as well as if the peptide is a binder to an MHC-I haplotype.
+For MHC-II, 15-mers are used, as these are the most occurring
+epitope length \cite{bergseng2015different}.
+
 Because the percentage of TMH residues is known, a binomial test can be
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Elution studies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-I}

### 2021-02-04 bbbq_article.tex
Progress

@@
 any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
-one thousand randomly simulated 
-peptides (see table \ref{table:ic50s_per_haplotype} for values), 
+the epitopes within a proteome (see table \ref{table:ic50s_per_haplotype} for values), 
 whereas the original study defined
@@
 uses 5\%, which is a common significance level for an exploratory study.
-
-Instead of using the membrane protein topology prediction
-tool \verb;TMHMM; (\cite{krogh2001predicting}), PureseqTM (\cite{wang2019efficient})
-is used instead.
 
@@
 
+% tab:elution
+\input{../bbbq_article_issue_157/results/results.tex}
+
+Table \ref{tab:elution} shows the percentage of epitopes derived from a TMH
+found in the two elution studies, for the two different kind of topology
+prediction tools. Whatever the tool or MHC class, at least 100 epitopes
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{bbbq_1_fast/fig_f_tmh_mhc1_grid.png}
+  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc1_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{bbbq_1_fast/fig_f_tmh_mhc2_grid.png}
+  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc2_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc1_normalized.png}
+  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc1_normalized.png}
   \caption{
@@
 
-%\begin{table}[!htbp]
-%   \input{bbbq_1_fast/bbbq_1_stats_covid.latex}
-%   \caption{
-%     Kolmogorov-Smirnov test results comparing human and SARS-CoV-2 for MHC-I
-%     \richel{Done on the simulated data}
-%   }
-%   \label{table:bbbq_1_stats_covid}
-% \end{table}
-
-% \begin{table}[!htbp]
-%   \input{bbbq_1_fast/bbbq_1_stats_myco.latex}
-%   \caption{
-%     Kolmogorov-Smirnov test results comparing human and Mycobacterium tuberculosis for MHC-I
-%     \richel{Done on the simulated data}
-%   }
-%   \label{table:bbbq_1_stats_myco}
-% \end{table}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_fast/fig_f_tmh_mhc2_normalized.png}
+  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc2_normalized.png}
   \caption{
@@
 %  \centering
-  \input{bbbq_1_fast/table_tmh_binders_mhc2.latex}
+%  \input{../bbbq_1_smart/table_tmh_binders_mhc2.latex}
 %\end{sidewaystable}
-
-% \begin{table}[!htbp]
-%   \input{bbbq_1_fast/bbbq_2_stats_covid.latex}
-%   \caption{
-%     Kolmogorov-Smirnov test results comparing human and SARS-CoV-2 for MHC-II
-%     \richel{Done on the simulated data}
-%   }
-%   \label{table:bbbq_2_stats_covid}
-% \end{table}
-
-% \begin{table}[!htbp]
-%   \input{bbbq_1_fast/bbbq_2_stats_myco.latex}
-%   \caption{
-%     Kolmogorov-Smirnov test results comparing human and 
-      Mycobacterium tuberculosis for MHC-II
-%     \richel{Done on the simulated data}
-%   }
-%   \label{table:bbbq_2_stats_myco}
-% \end{table}
 

### 2021-02-08 bbbq_article.tex
Use bbbq_1_smart figures

@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc1_grid.png}
+  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc1_2_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc2_grid.png}
+  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc2_2_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc1_normalized.png}
+  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc1_2_normalized.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc2_normalized.png}
+  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc2_2_normalized.png}
   \caption{

### 2021-02-08 bbbq_article.tex
No need of 'bbbq_1_fast' anymore, use 'bbbq_1_smart'
Show number of binders
15-mers for MHC-II nicely shown
Show IC50 values
Show IC50 values that determine a binder per MHC class

@@
 % Percentage of spots and spots that overlap with a TMH
-\input{bbbq_1_fast/table_f_tmh.latex}
+\input{../bbbq_1_smart/table_f_tmh_2.latex}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{IC50s per haplotype}
+\label{subsec:ic50s_per_haplotype}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Per target proteome (i.e. human, SARS-CoV-2, Mycobacterium tuberculosis),
+we collected all 9-mers (for MHC-I) and 15-mers (for MHC-II),
+after removing the selenoproteins and proteins that are shorter
+than the epitope length.
+From these epitopes, per MHC haplotype,
+we predicted the IC50 using \verb;epitope-prediction; (for MHC-I)
+and MHCnuggets (for MHC-II). 
+Here, we show the IC50 value per haplotype that
+is used to determine if a peptide binds to the haplotype's MHC.
+
+% tab:ic50_binders_mhc1
+\input{../bbbq_1_smart/table_ic50_binders_mhc1_2.latex}
+
+% tab:ic50_binders_mhc2
+\input{../bbbq_1_smart/table_ic50_binders_mhc2_2.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-I}
@@
 
-% Percentage of MHC-I 9-mers overlapping with TMH. Values in brackets
-% show the number of binders that have at least one residue overlapping with a
-% TMH (first value)as well as the number of binders (second value). Percentile
-% used: 0.02
 % Label: tab:tmh_binders_mhc1
-%\begin{sidewaystable}
-%  \centering
-%  ... centered table here
-%\end{sidewaystable}
-
-% \scalebox{0.7}{
-%  ... scaled table here
-% }
-
-%{\tiny
-%  ... something with a tiny font
-%}
-
-\input{bbbq_1_fast/table_tmh_binders_mhc1.latex}
+\input{../bbbq_1_smart/table_tmh_binders_mhc1_2.latex}
 
@@
 
-%\begin{sidewaystable}
-%  \centering
-%  \input{../bbbq_1_smart/table_tmh_binders_mhc2.latex}
-%\end{sidewaystable}
+% Label: tab:tmh_binders_mhc2
+\input{../bbbq_1_smart/table_tmh_binders_mhc2_2.latex}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IC50s per haplotype}
-\label{subsec:ic50s_per_haplotype}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Per MHC haplotype, we simulated random peptides and measured the
-predicted IC50 using \verb;epitope-prediction; (for MHC-I) 
-and netmhc2pan (for MHC-II). 
-Here, we show the IC50 value per haplotype that
-is used to determine if a peptide binds to the haplotype's MHC.
-Included in this figure is the MHCnuggets prediction tool,
-by \cite{shao2020high}.
-
-% Less pretty 
-% 
-% \begin{figure}[!htbp]
-%   \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype.png}
-%   \caption{
-%     IC50 values that define a binder
-%   }
-%   \label{fig:ic50s_per_haplotype}
-% \end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=0.9\textwidth]{ic50s_per_haplotype/ic50s_per_haplotype_per_mhc_class.png}
-  \caption{
-    IC50 values that define a binder
-  }
-  \label{fig:ic50s_per_haplotype}
-\end{figure}
-
-\input{ic50s_per_haplotype/ic50s_per_haplotype_1.latex}
-
-\input{ic50s_per_haplotype/ic50s_per_haplotype_2.latex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\subsection{Relation between hydrophobicity, TMH and MHC binding}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%\begin{figure}[!htbp]
-%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/binds_mhc1_vs_binds_mhc2.png}
-%  \caption{
-%    \richel{Used randomly simulated peptides, results are real}
-%  }
-%  \label{fig:binds_mhc1_vs_binds_mhc2}
-%\end{figure}
-
-%\begin{figure}[!htbp]
-%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc1.png}
-%  \caption{
-%    \richel{Used randomly simulated peptides, results are real}
-%  }
-%  \label{fig:hydrophobicity_vs_binds_mhc1}
-%\end{figure}
-
-%\begin{figure}[!htbp]
-%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_binds_mhc2.png}
-%  \caption{
-%    \richel{Used randomly simulated peptides, results are real}
-%  }
-%  \label{fig:hydrophobicity_vs_binds_mhc2}
-%\end{figure}
-
-%\begin{figure}[!htbp]
-%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/hydrophobicity_vs_is_tmh.png}
-%  \caption{
-%    \richel{Used randomly simulated peptides, results are real}
-%  }
-%  \label{fig:hydrophobicity_vs_is_tmh}
-%\end{figure}
-
-%\begin{figure}[!htbp]
-%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc1.png}
-%  \caption{
-%    \richel{Used randomly simulated peptides, results are real}
-%  }
-%  \label{fig:is_tmh_vs_binds_mhc1}
-%\end{figure}
-
-%\begin{figure}[!htbp]
-%  \includegraphics[width=0.9\textwidth]{p_bind_per_hydrophobicity/is_tmh_vs_binds_mhc2.png}
-%  \caption{
-%    \richel{Used randomly simulated peptides, results are real}
-%  }
-%  \label{fig:is_tmh_vs_binds_mhc2}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{MHC-II haplotype occurrences}
@@
 \end{itemize}
+
+%
+% Notes to self
+%
+
+%\begin{sidewaystable}
+%  \centering
+%  ... centered table here
+%\end{sidewaystable}
+
+% \scalebox{0.7}{
+%  ... scaled table here
+% }
+
+%{\tiny
+%  ... something with a tiny font
+%}

### 2021-02-09 bbbq_article.tex
Remove evolutionary conservation

@@
 
-Evolutionary conservation scores are shown in 
-figures \ref{fig:evolutionary_conservation_8k} (topology),
-\ref{fig:evolutionary_conservation_8k_violin} (per 
-protein with at least 1 TMH, violin)
-and \ref{fig:evolutionary_conservation_8k_merged} (grouped 
-from all proteins that have at least 1 TMH).
-See figures
-\ref{fig:evolutionary_conservation_8k_all} and 
-\ref{fig:evolutionary_conservation_8k_all_violin}
-for the scores for all proteins).
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores.png}
-  \caption{
-    Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for proteins that have at least one TMH,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-    The (non-logarithmic) scale of the conservation score has been removed
-    for readability.
-    Conservation scores do not start at residue 1 when
-    an aligned protein sequence outflanks the consensus sequence 
-    at the N-terminal
-  }
-  \label{fig:evolutionary_conservation_8k}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_merged_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    with conservation scores group from all proteins that have at least one TMH,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-  }
-  \label{fig:evolutionary_conservation_8k_merged}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    for proteins that have at least one TMH,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-  }
-  \label{fig:evolutionary_conservation_8k_violin}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 of proteins that have less than three amino acids. 
-The TRDD1_HUMAN protein, short for 'T cell receptor delta diversity 1',
+The TRDD1\_HUMAN protein, short for 'T cell receptor delta diversity 1',
 is, however, two peptides long. We manually added the prediction
@@
 
+Table \ref{tab:uniprot_ids} shows an overview of the Uniprot IDs
+for the reference proteomes used. Note that for Homo sapiens,
+we did not use the full reference proteome, but a representative
+and smaller reference proteome instead.
+
 % tab:uniprot_ids
@@
 From these epitopes, per MHC haplotype,
-we predicted the IC50 using \verb;epitope-prediction; (for MHC-I)
+we predicted the IC50 (in nM) using \verb;epitope-prediction; (for MHC-I)
 and MHCnuggets (for MHC-II). 
 Here, we show the IC50 value per haplotype that
-is used to determine if a peptide binds to the haplotype's MHC.
+is used to determine if a peptide binds to the haplotype's MHC
+for MHC-I (see \ref{tab:ic50_binders_mhc1}) and 
+MHC-II (see \ref{tab:ic50_binders_mhc2}).
 
@@
     for human, viral and bacterial proteomes.
-    Legend: covid = SARS-CoV-2, flua = influenza A, hepa = hepatitis A,
-    human = homo sapiens, myco = Mycobacterium tuberculosis, 
-    polio = poliovirus, rhino = rhinovirus
+    Legend: covid = SARS-CoV-2,
+    human = homo sapiens, myco = Mycobacterium tuberculosis
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores.png}
-  \caption{
-    Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for proteins that have at least one TMH,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-    The (non-logarithmic) scale of the conservation score has been removed
-    for readability.
-    Conservation scores do not start at residue 1 when
-    an aligned protein sequence outflanks the consensus sequence 
-    at the N-terminal
-  }
-  \label{fig:evolutionary_conservation_8k}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_merged_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    with conservation scores group from all proteins that have at least one TMH,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-  }
-  \label{fig:evolutionary_conservation_8k_merged}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    for proteins that have at least one TMH,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-  }
-  \label{fig:evolutionary_conservation_8k_violin}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/8000_42/scores_all.png}
-  \caption{
-    Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for all proteins,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-    The (non-logarithmic) scale of the conservation score has been removed
-    for readability.
-    Conservation scores do not start at residue 1 when
-    an aligned protein sequence outflanks the consensus sequence 
-    at the N-terminal.
-  }
-  \label{fig:evolutionary_conservation_8k_all}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/8000_42/scores_violin_all.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    for all proteins,
-    from a subset of 8000 randomly selected SARS-CoV-2 sequences.
-  }
-  \label{fig:evolutionary_conservation_8k_all_violin}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{All unique sequences}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
-  \caption{
-    Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for proteins that have at least one TMH,
-    using all unique sequences \richel{which I doubt is valid}.
-    The (non-logarithmic) scale of the conservation score has been removed
-    for readability.
-    Conservation scores do not start at residue 1 when
-    an aligned protein sequence outflanks the consensus sequence 
-    at the N-terminal
-  }
-  \label{fig:evolutionary_conservation_all_unique}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    with conservation scores group from all proteins that have at least one TMH,
-    using all unique sequences \richel{which I doubt is valid}.
-  }
-  \label{fig:evolutionary_conservation_all_unique_merged}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    for proteins that have at least one TMH,
-    using all unique sequences \richel{which I doubt is valid}.
-  }
-  \label{fig:evolutionary_conservation_all_unique_violin}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores.png}
-  \caption{
-    Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for proteins that have at least one TMH,
-    using all unique sequences \richel{which I doubt is valid}.
-    The (non-logarithmic) scale of the conservation score has been removed
-    for readability.
-    Conservation scores do not start at residue 1 when
-    an aligned protein sequence outflanks the consensus sequence 
-    at the N-terminal
-  }
-  \label{fig:evolutionary_conservation_all_unique}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_merged_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    with conservation scores group from all proteins that have at least one TMH,
-    using all unique sequences \richel{which I doubt is valid}.
-  }
-  \label{fig:evolutionary_conservation_all_unique_merged}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_violin.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    for proteins that have at least one TMH,
-    using all unique sequences \richel{which I doubt is valid}.
-  }
-  \label{fig:evolutionary_conservation_all_unique_violin}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=0.7\textwidth,scale=0.7]{bbbq_2/scores_all.png}
-  \caption{
-    Conservation scores for the aligned SARS-CoV-2 proteomes,
-    for all proteins,
-    using all unique sequences \richel{which I doubt is valid}.
-    The (non-logarithmic) scale of the conservation score has been removed
-    for readability.
-    Conservation scores do not start at residue 1 when
-    an aligned protein sequence outflanks the consensus sequence 
-    at the N-terminal.
-  }
-  \label{fig:evolutionary_conservation_all_unique_all}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_2/scores_violin_all.png}
-  \caption{
-    Evolutionary conservation scores of TMH versus non-TMH
-    residues in the SARS-CoV-2 proteome,
-    for all proteins,
-    using all unique sequences \richel{which I doubt is valid}.
-  }
-  \label{fig:evolutionary_conservation_all_unique_all_violin}
-\end{figure}
 

### 2021-02-09 bbbq_article.tex
Use stored results

@@
 as well as if the peptide is a binder to an MHC-I haplotype.
-For MHC-II, 15-mers are used, as these are the most occurring
+For MHC-II, 14-mers are used, as these are the most occurring
 epitope length \cite{bergseng2015different}.
@@
 To detect the evolutionary conservation of TMHs, 
-we use the following workflow (also shown
-in figure \ref{fig:snp_workflow}) to be able to count the
-number of SNPs that occur in TMHs, as well as the number of SNPs
-that occur in TMHs by chance.
+we collect human mutations and 
+tally their predicted location.
+To be more precise, we collected single nucleotide
+polymorphisms (SNPs) within the human population
+that resulted in the substitution of one amino acid.
+As we know the sequence of the protein a SNP acts upon,
+we predict that protein's topology and tally if
+the SNP occurred in a TMH or not, 
+after which we use statistics to determine if there are
+more, less or an equal amount of SNPs in TMHs
+as a measure of evolutionary conservation.
+This workflow is discussed in more detail below and also
+depicted in figure \ref{fig:snp_workflow}).
 
-To do so, we use the 
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases:
+HIERO
+
+\paragraph{Data}
+As a data source, multiple
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases are used:
 'gene' to find the gene names of membrane proteins, 
@@
 and 'protein' to obtain the sequence of proteins that SNPs act upon.
-The 'dbSNP' contains 2 billion (human-only) SNPs, although
-it is estimated that approximately ten percent (ranging from 2 to 36 percent) 
+The 'dbSNP' contains 650 million 
+catalogued non-redundant humane variations (called RefSNPs,
+/url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
+It is estimated that approximately ten percent (ranging from 2 to 36 percent) 
 are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}).
@@
 \end{figure}
-
-% \paragraph{Introduction}
-% 
-% To detect the evolutionary conservation of TMHs,
-% we first selected all unique protein sequences per
-% gene from all 45k SARS-CoV-2 proteomes,
-% resulting in around 500 unique protein sequences per gene.
-% From the unique protein sequences, we made an alignment per gene,
-% using the msa (\cite{bodenhofer2015msa}) R package, 
-% set up to use ClustalOmega (\cite{thompson1994clustal})
-% with the BLOSUM80 amino acid transition matrix (\cite{henikoff1992amino}).
-% Per alignment, we used msa to determine the consensus sequence
-% and the conservation scores.
-% Per consensus sequences, the topology was determined using pureseqtmr.
-% 
-% \paragraph{Statistics}
-% 
-% We compare the conservation values for TMH versus non-TMH locations,
-% for all membrane proteins combined,
-% using a Kolmogorov-Smirnov test, for a significance level of 5 percent.
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc1_2_grid.png}
+  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc1_2_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/fig_f_tmh_mhc2_2_grid.png}
+  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc2_2_grid.png}
   \caption{
@@
 
+DRIPs.
+
+Translational staling.
+
+Trans-splicing for epitopes.
+
 \richel{
@@
 % Percentage of spots and spots that overlap with a TMH
-\input{../bbbq_1_smart/table_f_tmh_2.latex}
+\input{../bbbq_1_smart/results_14-mer/table_f_tmh_2.latex}
 
@@
 Per target proteome (i.e. human, SARS-CoV-2, Mycobacterium tuberculosis),
-we collected all 9-mers (for MHC-I) and 15-mers (for MHC-II),
+we collected all 9-mers (for MHC-I) and 14-mers (for MHC-II),
 after removing the selenoproteins and proteins that are shorter
@@
 % tab:ic50_binders_mhc1
-\input{../bbbq_1_smart/table_ic50_binders_mhc1_2.latex}
+\input{../bbbq_1_smart/results_14-mer/table_ic50_binders_mhc1_2.latex}
 
 % tab:ic50_binders_mhc2
-\input{../bbbq_1_smart/table_ic50_binders_mhc2_2.latex}
+\input{../bbbq_1_smart/results_14-mer/table_ic50_binders_mhc2_2.latex}
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc1_2_normalized.png}
+  \includegraphics[width=\textwidth]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc1_2_normalized.png}
   \caption{
@@
 % Label: tab:tmh_binders_mhc1
-\input{../bbbq_1_smart/table_tmh_binders_mhc1_2.latex}
+\input{../bbbq_1_smart/results_14-mer/table_tmh_binders_mhc1_2.latex}
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{../bbbq_1_smart/fig_f_tmh_mhc2_2_normalized.png}
+  \includegraphics[width=\textwidth]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc2_2_normalized.png}
   \caption{
@@
 % Label: tab:tmh_binders_mhc2
-\input{../bbbq_1_smart/table_tmh_binders_mhc2_2.latex}
+\input{../bbbq_1_smart/results_14-mer/table_tmh_binders_mhc2_2.latex}
 

### 2021-02-09 scripts/issue_172/issue_172.R

@@ -1,0 +1,23 @@
+# Determine the minimum and maximum epitope length for MHCnuggets
+t <- tibble::tibble(
+  peptide_length = seq(1, 20),
+  is_supported = NA
+)
+
+for (i in seq_len(nrow(t))) {
+  peptide_length <- t$peptide_length[i]
+  peptide <- paste0(rep("F", times = peptide_length), collapse = "")
+  t$is_supported[i] <- FALSE
+  tryCatch({
+    mhcnuggetsr::predict_ic50(
+      mhcnuggets_options = mhcnuggetsr::create_mhcnuggets_options(
+        mhc = mhcnuggetsr::get_trained_mhc_1_haplotypes()[1]
+      ),
+      peptides = peptide
+    )
+    t$is_supported[i] <- TRUE
+  }, error = function(e) {} # nolint
+  )
+}
+t
+knitr::kable(t)

### 2021-02-09 check_elution_study/check.R
Check MHC-I elution study at https://github.com/richelbilderbeek/bbbq_article_issue_157, see #157

@@ -1,39 +1,0 @@
-proteome_filename <- "UP000005640_9606.fasta.gz"
-yes_filename <- "TMH-Bcell-elution.txt"
-no_filename <- "non-TMH-Bcell-elution.txt"
-
-testthat::expect_true(file.exists(proteome_filename))
-testthat::expect_true(file.exists(yes_filename))
-testthat::expect_true(file.exists(no_filename))
-
-proteome <- seqinr::read.fasta(proteome_filename)
-
-proteome_sequences <- unlist(seqinr::getSequence(proteome, as.string = TRUE))
-
-# Use YES sequences
-peptide_sequences <- stringr::str_to_lower(readr::read_lines(yes_filename))
-length(peptide_sequences)
-unique_peptide_sequences <- sort(unique(peptide_sequences))
-length(unique_peptide_sequences)
-head(unique_peptide_sequences)
-
-hit_topologies <- rep(NA, length(unique_peptide_sequences))
-
-for (i in seq_len(length(unique_peptide_sequences))) {
-  message(i)
-  peptide_sequence <- unique_peptide_sequences[i]
-  indices <- stringr::str_which(string = proteome_sequences, pattern = peptide_sequence)
-  message(peptide_sequence, ": ", paste(indices, collapse = ","))
-  protein_sequence <- proteome_sequences[indices[1]]
-  topology <- pureseqtmr::predict_topology_from_sequence(stringr::str_to_upper(protein_sequence))
-  testthat::expect_equal(stringr::str_length(protein_sequence), stringr::str_length(topology))
-  where <- stringr::str_locate(string = protein_sequence, pattern = peptide_sequence)
-  hit_topologies[i] <- stringr::str_sub(topology, start = where[1,1], end = where[1,2])
-  message(hit_topologies[i])
-}
-hit_topologies
-
-hit <- stringr::str_which(string = hit_topologies, pattern = "1")
-no_hit <- stringr::str_which(string = hit_topologies, pattern = "1", negate = TRUE)
-length(hit)
-length(no_hit)

### 2021-02-09 bbbq_article.tex
Show pictures from SNPs

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Subset of random sequences}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_f_snps_found_and_expected_per_n_tmhs.png}
+  \caption{fig_f_snps_found_and_expected_per_n_tmhs}
+  \label{fig:f_snps_found_and_expected_per_n_tmhs}
+\end{figure}
+)
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_f_snps_found_and_expected.png}
+  \caption{fig_f_snps_found_and_expected}
+  \label{fig:f_snps_found_and_expected}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_n_proteins_per_gene_name.png}
+  \caption{fig_n_proteins_per_gene_name}
+  \label{fig:n_proteins_per_gene_name}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_n_snps_found_and_expected_per_n_tmhs.png}
+  \caption{fig_n_snps_found_and_expected_per_n_tmhs}
+  \label{fig:n_snps_found_and_expected_per_n_tmhs}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_n_tmhs_per_protein.png}
+  \caption{fig_n_tmhs_per_protein}
+  \label{fig:n_tmhs_per_protein}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_snp_rel_pos.png}
+  \caption{fig_snp_rel_pos}
+  \label{fig:snp_rel_pos}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_snps_per_gene_name_ncbi.png}
+  \caption{fig_snps_per_gene_name_ncbi}
+  \label{fig:snps_per_gene_name_ncbi}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_snps_per_gene_name.png}
+  \caption{fig_snps_per_gene_name}
+  \label{fig:snps_per_gene_name}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{../ncbi_peregrine/scripts/fig_snps_per_gene_name_processed.png}
+  \caption{fig_snps_per_gene_name_processed}
+  \label{fig:snps_per_gene_name_processed}
+\end{figure}
 

### 2021-02-09 bbbq_article.tex
Store most important results within this repo

--- 2021-02-09 bbbq_article.tex
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc1_2_grid.png}
+  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc2_2_grid.png}
+  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
   \caption{
@@
 % Percentage of spots and spots that overlap with a TMH
-\input{../bbbq_1_smart/results_14-mer/table_f_tmh_2.latex}
+\input{bbbq_1_smart_results/table_f_tmh_2.latex}
 
@@
 % tab:ic50_binders_mhc1
-\input{../bbbq_1_smart/results_14-mer/table_ic50_binders_mhc1_2.latex}
+\input{bbbq_1_smart_results/table_ic50_binders_mhc1_2.latex}
 
 % tab:ic50_binders_mhc2
-\input{../bbbq_1_smart/results_14-mer/table_ic50_binders_mhc2_2.latex}
+\input{bbbq_1_smart_results/table_ic50_binders_mhc2_2.latex}
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc1_2_normalized.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_normalized.png}
   \caption{
@@
 % Label: tab:tmh_binders_mhc1
-\input{../bbbq_1_smart/results_14-mer/table_tmh_binders_mhc1_2.latex}
+\input{bbbq_1_smart_results/table_tmh_binders_mhc1_2.latex}
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{../bbbq_1_smart/results_14-mer/fig_f_tmh_mhc2_2_normalized.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_normalized.png}
   \caption{
@@
 % Label: tab:tmh_binders_mhc2
-\input{../bbbq_1_smart/results_14-mer/table_tmh_binders_mhc2_2.latex}
+\input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Subset of random sequences}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_n_tmhs.png}
+  \caption{fig_f_snps_found_and_expected_per_n_tmhs}
+  \label{fig:f_snps_found_and_expected_per_n_tmhs}
+\end{figure}
+)
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
+  \caption{fig_f_snps_found_and_expected}
+  \label{fig:f_snps_found_and_expected}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_proteins_per_gene_name.png}
+  \caption{fig_n_proteins_per_gene_name}
+  \label{fig:n_proteins_per_gene_name}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_snps_found_and_expected_per_n_tmhs.png}
+  \caption{fig_n_snps_found_and_expected_per_n_tmhs}
+  \label{fig:n_snps_found_and_expected_per_n_tmhs}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_tmhs_per_protein.png}
+  \caption{fig_n_tmhs_per_protein}
+  \label{fig:n_tmhs_per_protein}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snp_rel_pos.png}
+  \caption{fig_snp_rel_pos}
+  \label{fig:snp_rel_pos}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_ncbi.png}
+  \caption{fig_snps_per_gene_name_ncbi}
+  \label{fig:snps_per_gene_name_ncbi}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name.png}
+  \caption{fig_snps_per_gene_name}
+  \label{fig:snps_per_gene_name}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_processed.png}
+  \caption{fig_snps_per_gene_name_processed}
+  \label{fig:snps_per_gene_name_processed}
+\end{figure}
 

### 2021-02-09 bbbq_1_smart_results/update.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+cp ../../bbbq_1_smart/results/*.* .
+

### 2021-02-09 ncbi_peregrine_results/update.sh

@@ -1,0 +1,6 @@
+#!/bin/bash
+cp ../../ncbi_peregrine/scripts/*.png .
+cp ../../ncbi_peregrine/scripts/*.csv .
+cp ../../ncbi_peregrine/scripts/*.latex .
+cp ../../ncbi_peregrine/scripts/*.tex .
+

### 2021-02-09 bbbq_article.tex
Added NCBI figs

@@
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_n_tmhs.png}
-  \caption{fig_f_snps_found_and_expected_per_n_tmhs}
+  \caption{
+    Percentage of SNPs found in TMHs,
+    separated per number of TMHs.
+    The dashed diagonal lines show the expected linear trend line
+    if TMHs and soluble protein regions are equally conserved.
+  }
   \label{fig:f_snps_found_and_expected_per_n_tmhs}
@@
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
-  \caption{fig_f_snps_found_and_expected}
+  \caption{
+    Percentage of SNPs found in TMHs.
+    The dashed diagonal line shows the expected linear trend line
+    if TMHs and soluble protein regions are equally conserved.
+  }
   \label{fig:f_snps_found_and_expected}
@@
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_proteins_per_gene_name.png}
-  \caption{fig_n_proteins_per_gene_name}
+  \caption{
+    Histogram of the number of proteins found per gene name.
+    Most often, a gene name is associated with one proteins. 
+  }
   \label{fig:n_proteins_per_gene_name}
@@
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_snps_found_and_expected_per_n_tmhs.png}
-  \caption{fig_n_snps_found_and_expected_per_n_tmhs}
+  \caption{
+    Number of SNPs found in TMHs,
+    separated per number of TMHs.
+    The dashed diagonal lines show the expected linear trend line
+    if TMHs and soluble protein regions are equally conserved.
+  }
   \label{fig:n_snps_found_and_expected_per_n_tmhs}
@@
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_tmhs_per_protein.png}
-  \caption{fig_n_tmhs_per_protein}
+  \caption{
+    Histogram of the number of TMHs predicted per protein,
+    for the integral membrane proteins used.
+  }
   \label{fig:n_tmhs_per_protein}
@@
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snp_rel_pos.png}
-  \caption{fig_snp_rel_pos}
+  \caption{
+    Distribution of the relative position of the SNPs used,
+    where a relative position of zero denotes the first amino
+    acid at the N-terminus, where a relative position of one
+    indicates the last residue at the C-terminus.
+  }
   \label{fig:snp_rel_pos}
@@
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_ncbi.png}
-  \caption{fig_snps_per_gene_name_ncbi}
+  \caption{
+    Total number of SNPs per gene name.
+  }
   \label{fig:snps_per_gene_name_ncbi}
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name.png}
-  \caption{fig_snps_per_gene_name}
-  \label{fig:snps_per_gene_name}
-\end{figure}
-
-\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_processed.png}
-  \caption{fig_snps_per_gene_name_processed}
+  \caption{
+    Number of SNPs per gene processed.
+  }
   \label{fig:snps_per_gene_name_processed}

### 2021-02-09 bbbq_article.tex
Remove 'haplotype_freq_vs_f_tmh_epitopes'

@@
 
-HIERO
+\paragraph{Data}
 
-\paragraph{Data}
 As a data source, multiple
@@
 catalogued non-redundant humane variations (called RefSNPs,
-/url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
 It is estimated that approximately ten percent (ranging from 2 to 36 percent) 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Relation between haplotype frequency and TMH over-representation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-There is no relation between haplotype frequency and TMH over-representation.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/haplotype_perc_vs_f_tmh_binders.png}
-  \caption{
-    Relation between genotype frequency and the percentage
-    of binders that are TMH for the 21 MHC-II haplotypes
-  }
-  \label{fig:haplotype_perc_vs_f_tmh_binders}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{haplotype_freq_vs_f_tmh_epitopes/phenotype_freq_vs_f_tmh_binders.png}
-  \caption{
-    Relation between phenotype frequency and the percentage
-    of binders that are TMH for the 21 MHC-II haplotypes
-  }
-  \label{fig:phenotype_freq_vs_f_tmh_binders}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Hypotheses}

### 2021-02-09 haplotype_freq_vs_f_tmh_epitopes/create_figure.R

@@ -1,47 +1,0 @@
-library(dplyr, warn.conflicts = FALSE)
-library(ggplot2, warn.conflicts = FALSE)
-
-
-counts_filenames <- "../bbbq_1_fast/counts.csv"
-testthat::expect_true(file.exists(counts_filenames))
-
-counts_per_protein <- readr::read_csv(counts_filenames)
-
-counts <- counts_per_protein %>% dplyr::group_by(target, haplotype_id) %>%
-  summarise(
-    n_binders = sum(n_binders, na.rm = TRUE),
-    n_binders_tmh = sum(n_binders_tmh, na.rm = TRUE),
-    n_spots = sum(n_spots, na.rm = TRUE),
-    n_spots_tmh = sum(n_spots_tmh, na.rm = TRUE),
-    .groups = "drop"
-  )
-
-counts$f_tmh_binders <- counts$n_binders_tmh / counts$n_binders
-
-haplotypes_lut <- bbbq::create_haplotypes_lut()
-
-counts_with_haplotypes <- merge(counts, haplotypes_lut)
-
-occurances <- rbind(
-  bbbq::get_mhc1_haplotype_occurance(),
-  bbbq::get_mhc2_haplotype_occurance()
-)
-
-t <- merge(counts_with_haplotypes, occurances)
-t$phenotype_freq <- t$phenotype_freq/ 100.0
-t$haplotype_perc <- t$haplotype_perc/ 100.0
-
-
-ggplot2::ggplot(t, ggplot2::aes(x = phenotype_freq, y = f_tmh_binders, color = target)) +
-  ggplot2::geom_point() +
-  ggplot2::geom_smooth() +
-  ggplot2::scale_y_continuous("Percentage binders that are TMH", labels = scales::percent) +
-  ggplot2::scale_x_continuous("Haplotype phenotypic frequency", labels = scales::percent) +
-  ggplot2::ggsave("phenotype_freq_vs_f_tmh_binders.png", width = 7, height = 7)
-
-ggplot2::ggplot(t, ggplot2::aes(x = haplotype_perc, y = f_tmh_binders, color = target)) +
-  ggplot2::geom_point() +
-  ggplot2::geom_smooth() +
-  ggplot2::scale_y_continuous("Percentage binders that are TMH", labels = scales::percent) +
-  ggplot2::scale_x_continuous("Haplotype genotypic frequency", labels = scales::percent) +
-  ggplot2::ggsave("haplotype_perc_vs_f_tmh_binders.png", width = 7, height = 7)

### 2021-02-09 bbbq_article.tex
Describe API calls

@@
 As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases are used:
-'gene' to find the gene names of membrane proteins, 
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases are used,
+which are 'gene', to find the gene names of membrane proteins, 
 'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
-and 'protein' to obtain the sequence of proteins that SNPs act upon.
+and 'protein', to obtain the sequence of proteins that SNPs act upon.
 The 'dbSNP' contains 650 million 
@@
 It is estimated that approximately ten percent (ranging from 2 to 36 percent) 
-are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}).
+are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}),
+yet if this has been corrected for (if possible) in the last decade is unknown.
 
-The \verb;rentrez; R package provides for doing so from script.
+\paragraph{Data retrieval}
+
+To retrieve the data from these databases the
+\verb;rentrez; R package (\cite{rentrez}) is used
+to calls the NCBI website's API. To provide for a 
+stable user experience for all users, 
+this API limits the user to 3 calls per second.
+Additionally, the API splits the result of a bigger
+query into multiple pages, each of which needs one API call.
+We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
+bigger queries of multiple (and delayed) API calls.
+
+\paragraph{Data retrieval}
 

### 2021-02-09 bbbq_article.tex
Rewrite TMH

@@
 
-Transmembrane helices are structurally conserved structures that span
-a cell membrane with an alpha helix.
-TMHs are hydrophobic, as this is required to span the 
-hydrophobic cellular lipid membrane. Additionally,
-they often have a length of 23 amino acids to be able to span
-the membrane.
-Polypeptide fragments derived from TMHs are among the most hydrophobic,
-together with the internals of soluble proteins, where the
-hydrophobicity parts guide the protein to achieve its 3D configuration.
-TMHs are general structures: 25 percent of the human proteome is
-anchored by at least one TMH. Also the pathogen SARS-CoV-2 has 21 TMHs, 
-making up for 5.5 percent of its entire genome.
-There are multiple computational tools developed to predict which
-parts of membrane proteins are TMH (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
+Transmembrane helices (TMHs) 
+are structurally conserved and hydrophobic 
+regions of a protein that span a (also hydrophobic) 
+cell membrane by an alpha helix of usually 23 amino acids.
+Together with the internals of soluble proteins,
+peptides derived from TMHs are the most hydrophobic
+of a proteome. In both cases, the hydrophobic regions
+help shape the protein in its correct 3D configuration.
+TMHs are common structures in the proteins of
+humans and pathogens. For example, 
+15-39\% of the proteins in humans
+is predicted to be a membrane protein (\cite{ahram2006estimation}).
+There are multiple tools to do in silico predictions 
+to determine the transmembrane regions of a protein 
+(\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
 
@@
     with human/viral/bacterial TMHs as expected by chance
-  \item $\mathcal{H}_{3}$: TMH and non-TMH regions are equally conserved
+  \item $\mathcal{H}_{3}$: TMHs and soluble protein regions are equally conserved
     in a proteome

### 2021-02-11 bbbq_article.tex
Add DRiPS, epitope fusion and all those other goodies for the discussion

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Elution studies}
+\subsection{Elution studies}\label{subsec:elution_studies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\paragraph{Elution studies}
+
+We assume that epitopes 
+derived from TMHs are actually presented by MHC-I and MHC-II.
+Our analysis (at subsection \ref{subsec:elution_studies}) 
+finds that 1.4\% (for MHC-I) and 4.0% (for MHC-II) 
+of the epitopes presented 
+overlap with a TMH.
+It is unsure if this a true finding or a false positive.
+One source of noise would stem from imperfect predictions
+by the topology prediction software used. 
+We have counteracted this by using two different topology prediction
+software and confirm these have similar results.
+Two biological sources of noise
+are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
+and peptide fusion \cite{delong2016pathogenic},
+two processes that both result in peptide fragments
+that did not originate from genome directly.
+Note however, that both elution studies have a
+methodological bias caused by mass spectronomy,
+as hydrophobic fragments are detected less 
+by mass spectronomy \richel{Look up wording and ref in Bianchi et al 2017}.
+
+\paragraph{\% epitopes overlapping with TMHs}
+
 We have removed the 25 selenoproteins found in the human
@@
 
-DRIPs.
+\paragraph{Evolutionary conservation}
+
+In our evolutionary experiment, we removed variations that were synonymous mutations (i.e.
+resulted in the same amino acid, from a different genetic code) 
+from our analysis.
+There is evidence, however, that these synonymous mutations
+do have an effect and may even be evolutionary selected 
+for (\cite{hunt2009silent}).
+As the possible effect of synonymous mutations is ignored by our
+topology prediction software, we do so as well.
 

### 2021-02-11 bbbq_article.tex
Use Bogaart-Shakespearean, fix #177
Discuss haplotype that presents TMH-derived epitopes from SARS-CoV-2 four times as much, thanks @fransbianchi, fix #71

@@
 
-We found that the percentages of epitopes overlapping 
-with TMHs for a human and viral proteome are 
-[similar/different]. In other words, the
-epitopes that MHC-I presents are [as/not as] likely 
-to be derived from TMH within either a human host and its viral pathogen.
-
-We found that the percentages of epitopes overlapping 
-with TMHs for a human and bacterial proteome are 
-[similar/different]. In other words, the
-epitopes that MHC-I presents are [as/not as] likely 
-to be derived from TMH within either a human host and its bacterial pathogen.
+\paragraph{General pattern in epitope presentation}
+
+We found a general pattern in the
+presentation of epitopes that overlap with a TMH.
+For humans, a bacterial and viral pathogen,
+as well as for MHC-I and MHC-II,
+we found that these are presented more often 
+than expected by chance.
+
+\paragraph{Interesting haplotype for SARS-CoV-2}
+
+One interesting finding is that the MHC-I 
+haplotype HLA-A*24:02 
+present four times as much epitopes derived from SARS-CoV-2 TMHs
+than expected by chance.
+This may help the immune system to detect a SARS-CoV-2 infection earlier,
+but we have not investigates if an effect can be found in clinical data.
 
@@
 
-We concluded that the
-epitopes that MHC-I presents are [as/not as] likely 
-to be derived from TMH within either a human host and its viral pathogen.
-Because the full SARS-CoV-2 has 21 TMHs, the percentages
-of MHC-I epitopes being part of a TMH are likelier to be affected by
-stochasticity. We chose to use SARS-CoV-2 regardless, as the thousands
-of its time-dated genomic sequences are ideal for determining the 
-evolutionary conservation of MHC-I detecting TMHs.
-
-We concluded that the
-epitopes that MHC-I presents are [as/not as] likely 
-to be derived from TMH within either a human host and its bacterial pathogen.
-Because a bacterium does not infect a cell, thus its peptides
-will not be presented by MHC-I, this result is [unexpected/expected]
-
-We aimed our evolutionary experiment at TMHs, because these can
-be predicted well from a protein structure,
-are common structures and are present in all pathogens. 
-We could have done the same experiment on beta-turn,
-as also these can be predicted well (\cite{petersen2010netturnp}),
-are common structures and are present in all pathogens.
-
-The human immune system and human pathogen are in an evolutionary
-arms race: our immune systems is selected for the detection
-of pathogens, whereas pathogens are selected to avoid detection.
-From a pathogen's point of view, however, this struggle 
-is of only minor importance:
-in seasonal influenza, for example, the selection pressure
-exerted by the immune system was only limited (\cite{han2019individual}).
-
-Regarding the evolutionary conservation of TMHs using SNPs,
-again, it is estimated that approximately ten percent
-of SNPs is a false positive that result from the methods to determine
-a SNP. One example is that sequence variations are incorrectly
-detected due to highly similar duplicated sequences \cite{musumeci2010single}.
-We assume that these duplications occur as often in TMHs as in
-regions around these, hence we expect this not to affect our results.
-
-We used a low number of SNPs, as was imposed by the
-NCBI API, which constrains its users to three calls per second.
-As a second factor, we limited the time querying the NCBI database
-for the duration of the project.
-The SNPs we did use, we picked those
-with the lowest SNP ID (which denotes the earliest discovery date) first.
-We expect this not to introduce a bias, so we expect that
-the SNPs we used are a representative sample of all SNPs.
-
-\paragraph{Elution studies}
+\subsection{Elution studies}
+
+\paragraph{False positives}
 
@@
 
-\paragraph{\% epitopes overlapping with TMHs}
+\subsection{\% epitopes overlapping with TMHs}
+
+\paragraph{Removal of selenoproteins}
 
@@
 selenoproteins from our analysis.
-Another bug in PureseqTM is that it does not predict the topology
+
+\paragraph{Topology prediction of TRDD1}
+
+PureseqTM does not predict the topology
 of proteins that have less than three amino acids. 
-The TRDD1\_HUMAN protein, short for 'T cell receptor delta diversity 1',
-is, however, two peptides long. We manually added the prediction
-that both peptides reside in the cytosol.
-
-\paragraph{Evolutionary conservation}
-
-In our evolutionary experiment, we removed variations that were synonymous mutations (i.e.
+The TRDD1 ('T cell receptor delta diversity 1') protein,
+however, is two peptides long. 
+We corrected for this by predicting that this protein
+is completely cytosolic.
+
+\paragraph{Bacteria have a different cell membrane}
+
+In this experiment we predicted epitopes that overlap with 
+TMHs from a human, bacterial and viral proteome,
+would these proteins be expressed in a human host.
+Bacteria, however have different cell membranes and cell walls, 
+hence different structural requirements for a TMH.
+Both topology prediction tools were trained to recognize
+human TMHs (\richel{Check}), thus we cannot be sure that
+the transmembrane regions predicted in bacterial proteins
+are actually part of an integral membrane protein.
+For the purpose of this study, we assume the 
+error in topology predictions to be unbiased way towards topology.
+In other words: that a bacterial TMH is incorrectly
+predicted to be absent just as often as it is incorrectly
+predicted to be present elsewhere.
+
+\subsection{Evolutionary conservation}
+
+\paragraph{synonymous mutations}
+
+In our evolutionary experiment, 
+we removed variations that were synonymous mutations (i.e.
 resulted in the same amino acid, from a different genetic code) 
@@
 
-Translational staling.
-
-Trans-splicing for epitopes.
-
-\richel{
-  Note that most bacteria are opportunistic pathogens.
-  Note that most bacteria are generalists.
-  Note that most bacteria have different cell membranes (and walls), that
-  may have different functional constraints than a human cell membrane
-}
+\paragraph{False positives in SNPs}
+
+Regarding the evolutionary conservation of TMHs using SNPs,
+again, it is estimated that approximately ten percent
+of SNPs is a false positive that result from the methods to determine
+a SNP. One example is that sequence variations are incorrectly
+detected due to highly similar duplicated sequences \cite{musumeci2010single}.
+We assume that these duplications occur as often in TMHs as in
+regions around these, hence we expect this not to affect our results.
+
+\paragraph{Low number of SNPs}
+
+The NCBI dbSNP database contains millions of SNPs.
+As the NCBI API constrains its users to three calls per second,
+we had to limited the extent of our analysis
+by downloading SNPs to approximately 300 
+hours (i.e. the length of a Christmas holiday).
+
+\paragraph{Selection of SNPs}
+
+We selected to analyze SNPs with the lowest SNP ID first.
+This means we have a bias for picking SNPs with
+an earlier discovery date.
+We expect this bias to be biologically irrelevant
+and we find no clear evidence of a bias, as
+shown in figure \ref{fig:n_tmhs_per_protein}.
+Still, the best way to interpret this figure
+is open to discussion.
+
+\subsection{Notes and/or to be removed}
+
+We concluded that the
+epitopes that MHC-I presents are [as/not as] likely 
+to be derived from TMH within either a human host and its bacterial pathogen.
+Because a bacterium does not infect a cell, thus its peptides
+will not be presented by MHC-I, this result is [unexpected/expected]
+
+We aimed our evolutionary experiment at TMHs, because these can
+be predicted well from a protein structure,
+are common structures and are present in all pathogens. 
+We could have done the same experiment on beta-turn,
+as also these can be predicted well (\cite{petersen2010netturnp}),
+are common structures and are present in all pathogens.
+
+The human immune system and human pathogen are in an evolutionary
+arms race: our immune systems is selected for the detection
+of pathogens, whereas pathogens are selected to avoid detection.
+From a pathogen's point of view, however, this struggle 
+is of only minor importance:
+in seasonal influenza, for example, the selection pressure
+exerted by the immune system was only limited (\cite{han2019individual}).
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_snps_found_and_expected_per_n_tmhs.png}
-  \caption{
-    Number of SNPs found in TMHs,
-    separated per number of TMHs.
-    The dashed diagonal lines show the expected linear trend line
-    if TMHs and soluble protein regions are equally conserved.
-  }
-  \label{fig:n_snps_found_and_expected_per_n_tmhs}
-\end{figure}
-
-\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_tmhs_per_protein.png}
@@
 %}
+
+% Instead of non-TMH use 'soluble protein region'
+% or 'cytosolic or extracellular protein region'

### 2021-02-11 bbbq_article.tex
Improve the conclusions

@@
 
-One interesting finding is that the MHC-I 
-haplotype HLA-A*24:02 
-present four times as much epitopes derived from SARS-CoV-2 TMHs
-than expected by chance.
+It is known that an individual's haplotype
+influences the dynamics of a pathogen (see,
+for example, \cite{eccleston2017host} for HIV).
+The bioinformatics pipeline developed in this
+study may be interesting to make in silico
+predictions of a SARS-CoV-2 infection.
+The most prominent example is the (MHC-I)
+haplotype HLA-A*24:02,
+which is predicted to present four times as much epitopes derived 
+from SARS-CoV-2 TMHs than expected by chance.
 This may help the immune system to detect a SARS-CoV-2 infection earlier,
-but we have not investigates if an effect can be found in clinical data.
+and change the infection's dynamics.
 

### 2021-02-11 bbbq_article.tex
Add main figure to the results

@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
+  \caption{
+    Percentage of SNPs found in TMHs.
+    Each point resembles one protein, with a predicted percentage of
+    TMH (x-axis) and an observed occurrence of SNPs being located
+    within a TMH (y-axis).
+    The dashed diagonal line shows the expected linear trend line
+    if TMHs and soluble protein regions are equally conserved.
+    The red line is a linear trend line that includes the membrane-associated
+    proteins at the origin. 
+    The blue line is a linear trend line that includes only the
+    transmembrane proteins.
+  }
+  \label{fig:f_snps_found_and_expected}
+\end{figure}
+
+SNPs occur more often in soluble than in transmembrane protein regions,
+as shown in figure \ref{fig:f_snps_found_and_expected}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
-  \caption{
-    Percentage of SNPs found in TMHs.
-    The dashed diagonal line shows the expected linear trend line
-    if TMHs and soluble protein regions are equally conserved.
-  }
-  \label{fig:f_snps_found_and_expected}
-\end{figure}
-
-\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_proteins_per_gene_name.png}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MHC-II haplotype occurrences}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}[!htbp]
-  \input{mhc2_haplotypes.latex}
-  \caption{
-    Percentage of MHC-II haplotypes, from \cite{greenbaum2011functional}
-  }
-  \label{table:mhc2_haplotypes}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Hypotheses}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{As a reminder, to focus the research on}
-
-\begin{itemize}
-  \item $\mathcal{H}_{1,VB}$: MHC-I has the same percentage of epitopes overlapping
-    with viral/bacterial TMHs as with human TMHs
-  \item $\mathcal{H}_{2,HVB}$: MHC-II has the same percentage of epitopes overlapping
-    with human/viral/bacterial TMHs as expected by chance
-  \item $\mathcal{H}_{3}$: TMHs and soluble protein regions are equally conserved
-    in a proteome
-\end{itemize}
-
 %

### 2021-02-12 bbbq_article.tex
Describe SNP results better

@@
 
-\paragraph{Data retrieval}
+\paragraph{Pipeline}
 
@@
 term 'membrane protein' (in all fields) for the organism Homo sapiens.
-This results in \richel{1126} gene IDs.
+This results in 1129 gene IDs (note that this number varies through time).
 The next query is a call to the 'gene' database 
@@
 Per gene name, the 'dbSNP' NCBI database is queried for 
-SNPs associated with the gene name. The number of SNPs
-was limited to the first \richel{100} SNPs per gene,
-resulting in \richel{112600} SNPs 
-\richel{Note that \cite{musumeci2010single} demonstrates for a way to get
-try SNPs, instead of SNDs, maybe I should use that}
-.
-Per selected SNP, all variation (i.e. descriptions of the changes on DNA
-and/or RNA and/or protein level) are obtained from the 'dbSNP' database
-as well. Of these variations, only the variations that result in a single
-amino acid substitution are kept, resulting in \richel{???} protein changes.
-Per protein substitution (i.e. a protein name and an index of the changed residue), 
-the 'protein' NCBI database was queried for the
+variations associated with the gene name. 
+The number of variations
+was limited to the first 250 variations per gene,
+resulting in 282250 variations. A variation need not always be a SNP,
+as dbSNP also catalogs other DNA alterations, such as, among others, insertions,
+deletions and tandem repeats.
+We select only the variations that result in a SNP for
+a single amino acid substitution.
+Per SNP, the 'protein' NCBI database was queried for the
 protein sequence.
@@
 using PureseqTM.
-From the topology and the known location of the amino acid substitution, 
-it is tallied if the change occurred within a TMH or not.
-Also, the percentage of the topology that is TMH is measured.
-Not all membrane proteins have a transmembrane helix, resulting
-in \richel{???} proteins with at least one TMH.
-
-\begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{snp_workflow.png}
-  \caption{
-    Workflow to determine the evolutionary conservation of TMHs
-    using human SNPs.
-  }
-  \label{fig:snp_workflow}
-\end{figure}
+From the topology and the known location of the SNP, 
+we score the location (i.e. transmembrane or soluble protein region) 
+where the change occurred.
 
@@
 
-SNPs occur more often in soluble than in transmembrane protein regions,
-as shown in figure \ref{fig:f_snps_found_and_expected}.
+We obtained 1129 gene names associated with the key phrase 'membrane protein'.
+These genes are linked to 5203 proteins, of which 2754
+are predicted to be transmembrane proteins.
+We obtained 4811 variations that resulted in an
+amino acids substitution, of which 2568 were located 
+in predicted transmembrane proteins.
+Per protein, we predicted the percentage of TMH
+and calculated the percentage of amino acids substitutions
+occurring in transmembrane regions.
+These points, one per protein, were plotted as depicted in figure
+\ref{fig:f_snps_found_and_expected}.
+We did a linear fit on the data, with and without
+membrane-associated proteins, and added a 95\%
+confidence interval.
+Also, we added the line \inlineequation{y = x},
+which indicates the linear fit that would be expected if
+SNPs occur just as likely within transmembrane and soluble
+protein domains. Already these trend lines
+hint that variations occur more often in soluble 
+than in transmembrane protein regions.
+
+To test whether SNPs occur just as likely in soluble as
+in transmembrane protein domains, we used a Poisson binomial
+distribution. We first calculated the expected number of SNPs
+found in TMHs, which is 61457 SNPs multiplied by 6.8\% 
+of all proteins being TMH, resulting in the null hypothesis
+of 4203 SNPs being in the TMH. 
+Note that including or excluding
+the membrane associated proteins 
+has no effect on this statistical test, as the chance 
+of a successful trial is zero (there is zero chance
+to find a SNP in TMHs if there are none).
+The actually number of SNPs found in TMHs
+was 3889, which is lower than the null hypothesis. 
+The chance to find, within TMHs, this amount or less SNPs 
+equals $1.6201 \cdot 10^{-9}$.
+Using a alpha value of $0.05$, as is common practice for an uninformed
+expectation, we concluded that there are significantly less SNPs
+in TMHs as expected by chance.
+The effect size is that per 1 SNP found in soluble protein
+domains, one finds 0.925 SNPs in TMHs.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_n_tmhs.png}
+  \caption{
+    Percentage of SNPs found in TMHs,
+    separated per number of TMHs.
+    The dashed diagonal lines show the expected linear trend line
+    if TMHs and soluble protein regions are equally conserved.
+  }
+  \label{fig:f_snps_found_and_expected_per_n_tmhs}
+\end{figure}
+)
+
+We decided to split this analysis based on the number of TMHs
+a protein has. We hypothesize that single-spanners (i.e. proteins
+with one TMH predicted) are less conserved, when compared to multi-spanners,
+as single spanners just need to span a membrane, where some multi-spanners
+have enzymatically active sites embedded within their TMHs and thus
+have an additional structural constraint.
+As can be seen in figure \ref{fig:f_snps_found_and_expected_per_n_tmhs}, 
+there is no clear pattern to be observed.
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_n_tmhs.png}
-  \caption{
-    Percentage of SNPs found in TMHs,
-    separated per number of TMHs.
-    The dashed diagonal lines show the expected linear trend line
-    if TMHs and soluble protein regions are equally conserved.
-  }
-  \label{fig:f_snps_found_and_expected_per_n_tmhs}
-\end{figure}
-)
-
-\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_proteins_per_gene_name.png}
@@
 
+\begin{figure}[!htbp]
+  \includegraphics[height=0.9\textheight]{snp_workflow.png}
+  \caption{
+    Workflow to determine the evolutionary conservation of TMHs
+    using human SNPs.
+  }
+  \label{fig:snp_workflow}
+\end{figure}
+
 %

### 2021-02-12 bbbq_article.tex
Use all software

@@
 
-\paragraph{Introduction}
+% \paragraph{Introduction}
 
@@
 
-\paragraph{Data}
+% \paragraph{Data}
 
@@
 
-\paragraph{Data retrieval}
+% \paragraph{Data retrieval}
 
@@
 
-\paragraph{Pipeline}
+% \paragraph{Pipeline}
 
@@
     Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
-    \verb;epitope-prediction; LUT     & epiprepreds               & \cite{epiprepreds} \\
-    NetMHCIIpan LUT                   & nmhc2ppreds               & \cite{nmhc2ppreds} \\
-    Sequence alignment                & ClustalW                  & \cite{thompson1994clustal} \\
-    Conservation scores               & msa                       & \cite{bodenhofer2015msa} \\
     Combine all                       & bbbq                      & \cite{bbbq}
@@
 
-\paragraph{TMH prediction}
+% \paragraph{TMH prediction}
 
@@
 
-\paragraph{MHC-I epitope prediction}
+% \paragraph{MHC-I epitope prediction}
 
@@
 
-\paragraph{MHC-II epitope prediction}
+% \paragraph{MHC-II epitope prediction}
 
@@
 The complete experiment is bundled in the 'bbbq' R package,
-which is dependent on 'pureseqtmr', 
-'epitope-prediction' and 'epiprepreds',
-'netmhc2pan' and 'nmhc2ppreds'
+which is dependent on 'tmhmm', 'pureseqtmr', 
+'epitope-prediction' and 'mhcnuggetsr'
 as described below.
+
+% \paragraph{tmhmm}
+
+The R package 'tmhmm' was developed to do the same topology
+predictions as the earlier study (that used 'TMHMM'), yet in a reproducible way.
+'TMHMM' has a restrictive software license and allows a user
+to download a pre-compiled executable after confirmation he/she
+is in academia. The R package respects this restriction
+and allows the user to install and use TMHMM from within R,
+as done in this study.
+'tmhmm' has been submitted to and is accepted by CRAN.
+
+% \paragraph{pureseqtmr}
 
@@
 developers to verify the correctness of their work.
-'pureseqtmr' has been submitted to CRAN.
-
-To be able to use pre-calculated 
-\verb;epitope-prediction; and NetMHCIIpan
-predictions (i.e. a distribution of IC50 values per haplotype, 
-derived from random peptides), the 'epiprepreds' and 'nmhc2ppreds' 
-R packages have been 
-developed. In practice, these are nothing more than a big look-up table.
-If \verb;epitope-prediction; is put onto CRAN, 'epiprepreds'
-will be submitted as well.
-If 'netmhc2pan' is allowed onto CRAN, 'nmhc2ppreds'
-will be submitted as well.
+'pureseqtmr' has been submitted to and is accepted by CRAN.
+
+% \paragraph{mhcnuggetsr}
+
+MHCnuggets is a free and open-source Python package to predict 
+epitope affinity for many MHC-I and MHC-II variants.
+The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
+from within R.
+Also 'mhcnuggetsr' has been submitted to and is accepted by CRAN.
+
+% \paragraph{bbbq}
 

### 2021-02-12 bbbq_article.tex
Clean up article, use 14-mers

@@
 proteins produced by pathogen, either would the pathogen
-be a virus \richel{$\mathcal{H}_{1,1}$} or 
-a bacterium \richel{$\mathcal{H}_{1,2}$}.
+be a virus or 
+a bacterium.
 
@@
 presents TMH-derived epitopes as often, either 
-in humans \richel{$\mathcal{H}_{2,1}$},
-bacteria \richel{$\mathcal{H}_{2,2}$}
-or viruses \richel{$\mathcal{H}_{2,3}$}.
+in humans,
+bacteria or viruses.
 
@@
 The MHC-II haplotypes used additionally are selected 
-to occur with a phenotypic frequency of at least 
-\richel{recalculate number}
-14\% in
+to occur with a phenotypic frequency of at least 14\% in
 the human population (\cite{greenbaum2011functional},
@@
 resulting in 21 haplotypes.
-When using an MHC-II haplotype, instead of using 9-mers, 13-mers are
-used, as these are the canonical MHC-II epitope size \richel{Reference here}.
-
-Where the original study used a significance level of 1\% 
-for the binomial test to determine if TMH-derived peptides 
-are as likely to be a binder as expected by chance, this study
-uses 5\%, which is a common significance level for an exploratory study.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparing amounts of TMH epitopes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-To determine if MHC-I has the same percentage of epitopes overlapping
-with viral/bacterial TMHs as with human TMHs \richel{$\mathcal{H}_{1,VB}$}, 
-we compare the distributions of the measured percentages 
-of TMH-overlapping peptides for all haplotypes,
-and compare the distributions between humans and each pathogen
-using a two-sample Kolmogorov-Smirnov (KS) test
-for a significance level $\alpha = 0.05$, 
-which is a common significance level for an exploratory study.
+When using an MHC-II haplotype, instead of using 9-mers, 14-mers are
+used, as these are the most common MHC-II epitope size,
+as found in the elution study (\cite{bergseng2015different}).
 

### 2021-02-12 bbbq_article.tex
Removed notes to self

@@
 
-\richel{Weak, add refs}
 All pathogens can be detected by the foreign proteins on 
@@
 It is helpful to be able to predict which peptides are immunogenic,
-in, among others, vaccine development \richel{Reference here}. 
+in, among others, vaccine development. 
 Already for a decade, synthesized peptide fragments are used
 to experimentally determine which peptides
-are immunogenic \richel{Reference here}.
+are immunogenic.
 This approach, however, is tedious and costly.
@@
 selection acts upon: DNA may never be transcribed to RNA,
-or its RNA may never be translated (\cite{diz2012proteomics})
-\richel{improve upon this point}.
+or its RNA may never be translated (\cite{diz2012proteomics}).
 When using a proteome in phylogenetic research,
@@
 
-% The KS test to determine if MHC-I has the same percentage 
-% of epitopes overlapping with viral TMHs, 
-% compared to human TMHs \richel{$\mathcal{H}_{1,V}$},
-% resulted in a p-value of \richel{unknown},
-% which makes us [reject/accept] the hypothesis that these percentages
-% are sampled from the same distribution.
-
-% The KS test to determine if MHC-I has the same percentage 
-% of epitopes overlapping with bacterial TMHs, 
-% compared to human TMHs \richel{$\mathcal{H}_{1,B}$},
-% resulted in a p-value of \richel{unknown},
-% which makes us [reject/accept] the hypothesis that these percentages
-% are sampled from the same distribution.
-
 \begin{figure}[!htbp]
@@
     The horizontal lines indicate the percentage as expected by chance.
-    \richel{Refer to a table with the numbers as expected by chance)}
-    See table \ref{tab:tmh_binders_mhc1} for the TMH and epitope counts.
-    \richel{Epitope prediction by EpitopePrediction.}.
-    Note that for smaller proteomes a percentage of zero is likelier.
-    \richel{use of top 2\% binders}
+    See table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
   }
@@
     The horizontal lines indicate the percentage as expected by chance.
-    \richel{Refer to a table with the numbers as expected by chance)}
-    \richel{Epitope prediction by NetMHCIIpan.}
-    See table \ref{tab:tmh_binders_mhc2} for the TMH and epitope counts.
+    See table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
     Note that for smaller proteomes a percentage of zero is likelier.
-    \richel{use of top 2\% binders}
   }
@@
 as hydrophobic fragments are detected less 
-by mass spectronomy \richel{Look up wording and ref in Bianchi et al 2017}.
+by mass spectronomy.
 
@@
 Both topology prediction tools were trained to recognize
-human TMHs (\richel{Check}), thus we cannot be sure that
+human TMHs, thus we cannot be sure that
 the transmembrane regions predicted in bacterial proteins
@@
 See \url{https://github.com/richelbilderbeek/bbbq_article/issues/158}.
-
-\paragraph{Which reference proteome?}
-
-See \url{https://github.com/richelbilderbeek/bbbq_article/issues/155}.
 

### 2021-02-12 bbbq_article.tex
Remove obsolete text, add braces around refs

@@
 predictions, which are reliable in 
-practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+practice (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
 
@@
 is evolutionary selected for.
-One possible explanation is that TMHs have a reduced variability,
-imposed by the functional requirement of being able to span a lipid bilayer.
+One reason to expect a reduced variability (and hence 
+evolutionary conservation) in TMHs, is
+that these are restricted in their evolution
+by the functional requirement to span a lipid bilayer.
 Due to this, the TMHs that pathogens possess, 
@@
 
-\paragraph{Data set for evolutionary conservation of TMHs}
-
-To detect the evolutionary conservation of TMHs,
-we use the 49790 (as of 2020-06-22) fully sequenced SARS-CoV-2 proteomes, 
-as provided by GISAID (\cite{shu2017gisaid}).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 we started from epitopes sequenced from elution studies
-for MHC-I \cite{schellens2015comprehensive} 
-and MHC-II \cite{bergseng2015different}.
+for MHC-I (\cite{schellens2015comprehensive}) 
+and MHC-II (\cite{bergseng2015different}).
 For each epitope, the human representative reference proteome
@@
 as well as if the peptide is a binder to an MHC-I haplotype.
-For MHC-II, 14-mers are used, as these are the most occurring
-epitope length \cite{bergseng2015different}.
-
-Because the percentage of TMH residues is known, a binomial test can be
-used to determine if TMH-derived peptides are as likely to be a binder
-as expected by chance (the null hypothesis), or that TMHs are indeed an
-overrepresented source of epitopes (as found by \cite{bianchi2017}).
+For MHC-II, 14-mers are used, as these are the most frequently occurring
+epitope length (\cite{bergseng2015different}).
 
@@
 any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
-the epitopes within a proteome (see table \ref{table:ic50s_per_haplotype} for values), 
-whereas the original study defined
-a binder as having an IC50 in the lowest 2\% of the peptides within a protein.
+the epitopes within a 
+\emph{proteome} (see tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+for values), whereas the original study defined
+a binder as having an IC50 in the lowest 2\% 
+of the peptides within a \emph{protein}.
 % See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
@@
 to occur with a phenotypic frequency of at least 14\% in
-the human population (\cite{greenbaum2011functional},
-see \ref{table:mhc2_haplotypes} for coverage),
+the human population (\cite{greenbaum2011functional}),
 resulting in 21 haplotypes.
@@
 to predict epitopes. 
-According to \cite{lundegaard2011prediction}, in 2011,
-from a set of multiple tools, 
+According to \cite{lundegaard2011prediction}, at that time,
 NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.

### 2021-02-12 bbbq_article.tex
First self-review

@@
 prediction tools. Whatever the tool or MHC class, at least 100 epitopes
+can be uniquely map to human representative reference proteome.
 
@@
 Figure \ref{fig:1} shows the percentages of MHC-I epitopes overlapping 
-with TMHs for our human, viral and bacterial proteome,
-when using EpitopePrediction to predict epitopes.
+with TMHs for our human, viral and bacterial proteome.
 See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
-for the TMH and epitope counts.
+for the exact TMH and epitope counts.
 
@@
 
-\paragraph{General pattern in epitope presentation}
+% \paragraph{General pattern in epitope presentation}
 
@@
 
-\paragraph{Interesting haplotype for SARS-CoV-2}
+% \paragraph{Interesting haplotype for SARS-CoV-2}
 
@@
 
-\paragraph{False positives}
+% \paragraph{False positives}
 
@@
 
-\paragraph{Removal of selenoproteins}
+% \paragraph{Removal of selenoproteins}
 
@@
 
-\paragraph{Topology prediction of TRDD1}
+% \paragraph{Topology prediction of TRDD1}
 
@@
 
-\paragraph{Bacteria have a different cell membrane}
+% \paragraph{Bacteria have a different cell membrane}
 
@@
 
-\paragraph{synonymous mutations}
+% \paragraph{synonymous mutations}
 
@@
 
-\paragraph{False positives in SNPs}
+% \paragraph{False positives in SNPs}
 
@@
 
-\paragraph{Low number of SNPs}
+% \paragraph{Low number of SNPs}
 
@@
 
-\paragraph{Selection of SNPs}
+% \paragraph{Selection of SNPs}
 
@@
 
-\subsection{Notes and/or to be removed}
-
-We concluded that the
-epitopes that MHC-I presents are [as/not as] likely 
-to be derived from TMH within either a human host and its bacterial pathogen.
-Because a bacterium does not infect a cell, thus its peptides
-will not be presented by MHC-I, this result is [unexpected/expected]
-
-We aimed our evolutionary experiment at TMHs, because these can
-be predicted well from a protein structure,
-are common structures and are present in all pathogens. 
-We could have done the same experiment on beta-turn,
-as also these can be predicted well (\cite{petersen2010netturnp}),
-are common structures and are present in all pathogens.
-
-The human immune system and human pathogen are in an evolutionary
-arms race: our immune systems is selected for the detection
-of pathogens, whereas pathogens are selected to avoid detection.
-From a pathogen's point of view, however, this struggle 
-is of only minor importance:
-in seasonal influenza, for example, the selection pressure
-exerted by the immune system was only limited (\cite{han2019individual}).
+% We concluded that the
+% epitopes that MHC-I presents are [as/not as] likely 
+% to be derived from TMH within either a human host and its bacterial pathogen.
+% Because a bacterium does not infect a cell, thus its peptides
+% will not be presented by MHC-I, this result is [unexpected/expected]
+% 
+% We aimed our evolutionary experiment at TMHs, because these can
+% be predicted well from a protein structure,
+% are common structures and are present in all pathogens. 
+% We could have done the same experiment on beta-turn,
+% as also these can be predicted well (\cite{petersen2010netturnp}),
+% are common structures and are present in all pathogens.
+% 
+% The human immune system and human pathogen are in an evolutionary
+% arms race: our immune systems is selected for the detection
+% of pathogens, whereas pathogens are selected to avoid detection.
+% From a pathogen's point of view, however, this struggle 
+% is of only minor importance:
+% in seasonal influenza, for example, the selection pressure
+% exerted by the immune system was only limited (\cite{han2019individual}).
 
@@
 
-\paragraph{TMH prediction software}
+% \paragraph{TMH prediction software}
 
@@
 
-\paragraph{IC50 prediction software}
+% \paragraph{IC50 prediction software}
 
@@
 
-\paragraph{Definition of what a binder is}
+% \paragraph{Definition of what a binder is}
 
@@
 
-\paragraph{Unclear handing of sequences with U}
+% \paragraph{Unclear handing of sequences with U}
 
@@
 See \url{https://github.com/richelbilderbeek/bbbq_article/issues/158}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Uniprot IDs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Table \ref{tab:uniprot_ids} shows an overview of the Uniprot IDs
-for the reference proteomes used. Note that for Homo sapiens,
-we did not use the full reference proteome, but a representative
-and smaller reference proteome instead.
-
-% tab:uniprot_ids
-\input{create_table_reference_proteomes/uniprot_ids.latex}
 
@@
 is used to determine if a peptide binds to the haplotype's MHC
-for MHC-I (see \ref{tab:ic50_binders_mhc1}) and 
-MHC-II (see \ref{tab:ic50_binders_mhc2}).
+for MHC-I (see table \ref{tab:ic50_binders_mhc1}) and 
+MHC-II (see table \ref{tab:ic50_binders_mhc2}).
 
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_ncbi.png}
-  \caption{
-    Total number of SNPs per gene name.
-  }
-  \label{fig:snps_per_gene_name_ncbi}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_processed.png}
-  \caption{
-    Number of SNPs per gene processed.
-  }
-  \label{fig:snps_per_gene_name_processed}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{snp_workflow.png}
-  \caption{
-    Workflow to determine the evolutionary conservation of TMHs
-    using human SNPs.
-  }
-  \label{fig:snp_workflow}
-\end{figure}
-
 %

### 2021-02-15 bbbq_article.tex
Rewrite of abstract, progress #179

@@
 
-The human immune system is unexpectedly attentive towards
-human transmembrane helices (TMHs), as TMH-derived peptide fragments
-bind to histocompatibility complex (MHC) class I more often then expected
-by chance only, which can be mostly explained by the low hydrophobicity
-of those residues.
-The physiological reason is yet unclear,
-yet this finding hints that there has been selection upon detecting TMH-derived
-peptides in pathogens.
-If this signature of selection is present in MHC-II is unknown.
-This study shows that MHC-I also has more
-epitopes derived from a TMH for a pathogen proteome, when compared with
-a host proteome.
-Additionally, MHC-II binds to peptides derived from TMHs 
-more often than expected by chance.
-Lastly, we show the TMHs are evolutionarily more conserved,
-as less SNPs are found in TMH than expected by chance.
-Our findings suggest that the immune system is more
-vigilant to TMHs than expected by chance
-and suggests that an evolutionary signal in the evolutionary history of a pathogen
-may be found.
+Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
+Because the peptide-binding grooves of most haplotypes 
+of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
+peptide fragments derived from transmembrane helices (TMHs) 
+are predicted to be presented more often than expected based on their abundance.
+However, the physiological reason of why membrane proteins might be 
+over-presented is unclear
+In this study, we show the TMHs are evolutionarily more conserved, 
+as relatively less single nucleotide polymorphisms (SNPs) 
+are present in TMH-coding chromosomal regions 
+compared to regions coding for extracellular and cytosolic protein regions. 
+Moreover, we show that the over-presentation of TMH-derived peptides is 
+likely general, as it is predicted for diverse microbial pathogens 
+and for both MHC-I and MHC-II.
+Thus, our findings suggest that T cells might respond 
+more to membrane proteins, because these are evolutionary more conserved.
+We speculate that TMHs might be less prone of the escape mutations 
+that enable pathogens to evade T cell responses.
 

### 2021-02-15 bbbq_article.tex
Remove keyword 'SARS-CoV-2', thanks Geert, progress #179

@@
 adaptive immunity, transmembrane domain, transmembrane helix, 
-epitopes, T lymphocyte, MHC-I, MHC-II, SARS-CoV-2, evolutionary conservation
+epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation
 

### 2021-02-15 bbbq_article.tex
Process feedback @bogeert. Thanks Geert! Progress #179

@@
 
-Our immune system fights microbial pathogens on a daily basis,
+Our immune system fights microbial pathogens, 
 such as fungi, bacteria or viruses. 
-The acquired immune response
-needs time to develop its specialized and more effective
-combat forces and learn to recognize the pathogens 
-by use of the Major Histocompatibility Complex (MHC).
+An important part of the acquired immune response, 
+that develops specialized and more effective recognition of pathogens, 
+are T cells which recognize peptides derived from the pathogens 
+by use of Major Histocompatibility Complexes (MHC).
 
@@
 
-Antigen-presenting cells such as dendritic cells
-use the MHC class I molecules to 
-present randomly sampled peptides fragments,
-as created by the immunoproteasome (\cite{chapiro2006destructive}).
-If a cell gets infected by an intracellular pathogen, also the virus' peptides
-will also be presented at the cell surface. The foreign viral antigen is detected 
-by cytotoxic T lymphocytes, which will kill the infected cell.
+% GvdB: Not really relevant.
+You don’t really need to explain the role of DCs.
+What is far more important is the concept of what an MHC is: a protein complex that contains a fold, called the peptide-binding groove, which presents peptides.
+Also immediately explain that MHC-I presents peptides of about 9-amino acids in length to cytolytic T cells, whereas MHC-II presents longer peptides to helper T cells.
+
+% Antigen-presenting cells such as dendritic cells
+% use the MHC class I molecules to 
+% present randomly sampled peptides fragments,
+% as created by the immunoproteasome (\cite{chapiro2006destructive}).
+% If a cell gets infected by an intracellular pathogen, also the virus' peptides
+% will also be presented at the cell surface. The foreign viral antigen is detected 
+% by cytotoxic T lymphocytes, which will kill the infected cell.
 
@@
 
-All pathogens can be detected by the foreign proteins on 
-their (bacterial or fungal) cell walls or (viral) envelope.
-For bacteria and fungi, this is the main mode of their detection,
-as these do not infect a cell with their (foreign) DNA.
-T and B cells express MHC-II to detect foreign peptide fragments.
-An immune response is started when MHC-II detects a pathogen.
+% GvdB: Also not really relevant
+
+% All pathogens can be detected by the foreign proteins on 
+% their (bacterial or fungal) cell walls or (viral) envelope.
+% For bacteria and fungi, this is the main mode of their detection,
+% as these do not infect a cell with their (foreign) DNA.
+% T and B cells express MHC-II to detect foreign peptide fragments.
+% An immune response is started when MHC-II detects a pathogen.
 
@@
 
-Any human's immune system detects only a fraction of all possible
+Any individual's immune system detects only a fraction of all possible
 peptide fragments.
@@
 The HLA region of humans is highly polymorphic, with hundreds 
-to thousands of alleles for a gene (\cite{marsh2010nomenclature}).
+to thousands of different alleles, 
+and each different HLA gene is called 
+an MHC haplotype (\cite{marsh2010nomenclature}).
 
@@
 each presenting their own subset of epitopes.
+It is believed that this improves immunity at the population level, 
+as mutations in a protein that disrupt a particular MHC presentation, 
+so-called escape mutations, 
+will not affect MHC presentation for all haplotypes \richel{Reference needed}.
 

### 2021-02-17 bbbq_article.tex
Process feedback @bogeert. Thanks Geert! Progress #179

@@
 
-It is helpful to be able to predict which peptides are immunogenic,
-in, among others, vaccine development. 
-Already for a decade, synthesized peptide fragments are used
-to experimentally determine which peptides
-are immunogenic.
-This approach, however, is tedious and costly.
-Due to this, software was written that allows for silico 
-predictions, which are reliable in 
-practice (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
+Much studies are aimed at determining which peptides are presented in MHC 
+and will result in an immune response, 
+as this will for instance aid the design of vaccines. 
+These studies have led to the development 
+of very reliable prediction algorithms 
+that allows for in silico predictions 
+of the binding affinities of peptides (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
 
@@
 
-As the immunoproteasome is a cytosolic protein complex, on expects it
-to cleave only cytosolic proteins. However, it has recently been
-discovered that transmembrane protein residues are also presented
-by MHC-I complexes (\cite{bianchi2017}). This discovery is based
-on both in vitro elution studies, as well as on in silico predictions,
-with a focus on transmembrane helices. 
-The intracellular pathway, in which transmembrane residues are loaded onto
-an MHC, however, is yet unknown.
+Using these prediction algorithms, 
+we recently showed that peptides derived 
+from transmembrane helices (TMHs) 
+are predicted to be more frequently presented by MHC-I 
+than expected based on their abundance (\cite{bianchi2017}). 
+Moreover, we showed that some well-known immunodominant peptides stem from TMHs.
 
-Transmembrane helices (TMHs) 
-are structurally conserved and hydrophobic 
-regions of a protein that span a (also hydrophobic) 
-cell membrane by an alpha helix of usually 23 amino acids.
-Together with the internals of soluble proteins,
-peptides derived from TMHs are the most hydrophobic
-of a proteome. In both cases, the hydrophobic regions
-help shape the protein in its correct 3D configuration.
-TMHs are common structures in the proteins of
-humans and pathogens. For example, 
-15-39\% of the proteins in humans
-is predicted to be a membrane protein (\cite{ahram2006estimation}).
-There are multiple tools to do in silico predictions 
-to determine the transmembrane regions of a protein 
-(\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
+This over-presentation is entirely explained by the fact 
+that the peptide-binding groove of most MHC-I haplotypes 
+is relatively hydrophobic, 
+and therefore hydrophobic peptides have a higher affinity 
+than hydrophilic peptides.
+
+Transmembrane helices (TMHs) are hydrophobic 
+as they need to span the hydrophobic lipid bilayer of cellular membranes, 
+and they consist of an alpha helix of usually 23 amino acids. 
+TMHs are common structures in the proteins of humans and pathogens. 
+For example, 
+15-39% 
+\richel{Why wide range? Issue 180}
+of the proteins in humans 
+are predicted to be a membrane protein 
+containing at least one TMH (\cite{ahram2006estimation}). 
+TMHs can also be predicted from a protein sequence 
+with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
 
@@
 
-For MHC-I, it was found that predicted epitopes derived 
-from human transmembrane helices (TMHs)
-are over-presented by all 5 HLA-A and 
-most of 8 HLA-B super-types (\cite{bianchi2017}),
-due the high hydrophobicity as is required for a TMH to span
-the (hydrophobic) cell membrane, combined with the fact that the MHC
-binding cleft is hydrophobic as well.
+%For MHC-I, it was found that predicted epitopes derived 
+%from human transmembrane helices (TMHs)
+%are over-presented by all 5 HLA-A and 
+%most of 8 HLA-B super-types (\cite{bianchi2017}),
+%due the high hydrophobicity as is required for a TMH to span
+%the (hydrophobic) cell membrane, combined with the fact that the MHC
+%binding cleft is hydrophobic as well.
+%
+%It is unknown if the presentation of hydrophobic/TMH residues
+%is evolutionary selected for.
+%One reason to expect a reduced variability (and hence 
+%evolutionary conservation) in TMHs, is
+%that these are restricted in their evolution
+%by the functional requirement to span a lipid bilayer.
+%Due to this, the TMHs that pathogens possess, 
+%have a lower chance to develop an escape mutation,
+%as many mutations will result in a dysfunctional TMH.
 
-It is unknown if the presentation of hydrophobic/TMH residues
-is evolutionary selected for.
-One reason to expect a reduced variability (and hence 
-evolutionary conservation) in TMHs, is
-that these are restricted in their evolution
+However, the physiological reason why peptides derived from TMHs 
+would be presented more often than peptides 
+stemming from soluble protein regions is unknown. 
+One possibility is that the presentation of 
+hydrophobic/TMH residues is evolutionary selected for, 
+because TMHs are less prone to undergo escape mutation. 
+One reason to expect such a reduced 
+variability (and hence evolutionary conservation) in TMHs, 
+is that these are restricted in their evolution 
 by the functional requirement to span a lipid bilayer. 
-Due to this, the TMHs that pathogens possess, 
-have a lower chance to develop an escape mutation,
-as many mutations will result in a dysfunctional TMH.
+Due to this, there are constraints in the types of amino acids 
+that can be present in the TMH \richel{ref}. 
+Therefore, the TMHs of pathogens 
+might have a lower chance to develop an escape mutations, 
+as many mutations will result in a dysfunctional TMH 
+and render the protein inactive.
 
+% GvdB:
+% The text below here is more discussion 
+% as it is already interpretation of the findings. 
+% I would add one short paragraph, 
+% where you reiterate the most important findings:
+% In this study, we show that ...(and than basically 
+% the abstract in slightly different words and slightly 
+% more detail). End with a statement of why it is 
+% important (vaccin development based on TMHs?)...
+
+% RJCB: This was suggested for removal, until the Methods:
 % \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
-
-It is important that MHC-I presents both the peptides of the 
-(healthy) cell, as well as possible pathogen-derived peptides when the
-cell is infected. As described above, MHC-I presents TMH-derived epitopes 
-from the human host more often. It is unknown if MHC-I has the same
-dedication to present epitopes that stem from TMHs derived from
-proteins produced by pathogen, either would the pathogen
-be a virus or 
-a bacterium.
-
+% 
+% It is important that MHC-I presents both the peptides of the 
+% (healthy) cell, as well as possible pathogen-derived peptides when the
+% cell is infected. As described above, MHC-I presents TMH-derived epitopes 
+% from the human host more often. It is unknown if MHC-I has the same
+% dedication to present epitopes that stem from TMHs derived from
+% proteins produced by pathogen, either would the pathogen
+% be a virus or 
+% a bacterium.
+% 
 % \paragraph{MHC-II is expected to present TMHs}
-
-If presentation of TMHs on MHC-I would bring an evolutionary advantage 
-in the recognition of pathogens by the immune system, 
-it would follow that this is equally important for MHC-II, 
-especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
-response (\cite{novy2007cd4}). 
-The mechanism to detect foreign TMHs by MHC-II would be unknown, 
-similar to the discovery that MHC-I presents TMHs.
-
+% 
+% If presentation of TMHs on MHC-I would bring an evolutionary advantage 
+% in the recognition of pathogens by the immune system, 
+% it would follow that this is equally important for MHC-II, 
+% especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
+% response (\cite{novy2007cd4}). 
+% The mechanism to detect foreign TMHs by MHC-II would be unknown, 
+% similar to the discovery that MHC-I presents TMHs.
+% 
 % \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
-
-It is unknown if MHC-II, like MHC-I, 
-presents TMH-derived epitopes as often, either 
-in humans,
-bacteria or viruses.
-
+% 
+% It is unknown if MHC-II, like MHC-I, 
+% presents TMH-derived epitopes as often, either 
+% in humans,
+% bacteria or viruses.
+% 
 % \paragraph{Selection undetectable in whole proteome}
-
-In general, on would hope that evolutionary selection results in
-an immune system that as most attentive for loci that are
-essential for a virus, as these will be most conserved.
-In SARS-CoV-2, for example, there is preliminary evidence that the strongest
-selection pressure is upon residues that changes its 
-virulence (\cite{velazquez2020positive}).
-These loci, however, only account for a small part of a pathogen's proteome.
-Additionally, these essential parts differ widely between pathogens.
-Because of this scarcity and variance in targets, 
-one can imagine that the human immune system 
-is not tailored to detect these sites, 
-as hinted by upon by the aforementioned influenza study.
-
+% 
+% In general, on would hope that evolutionary selection results in
+% an immune system that as most attentive for loci that are
+% essential for a virus, as these will be most conserved.
+% In SARS-CoV-2, for example, there is preliminary evidence that the strongest
+% selection pressure is upon residues that changes its 
+% virulence (\cite{velazquez2020positive}).
+% These loci, however, only account for a small part of a pathogen's proteome.
+% Additionally, these essential parts differ widely between pathogens.
+% Because of this scarcity and variance in targets, 
+% one can imagine that the human immune system 
+% is not tailored to detect these sites, 
+% as hinted by upon by the aforementioned influenza study.
+% 
 % \paragraph{Selection may be detectable in TMHs}
-
-TMHs, on the other hand, also have their function constraints, 
-yet can occur multiple time a pathogen's proteome.
-One can safely assume a pathogen's proteome contains multiple TMHs.
-Therefore, it may be beneficial for the host
-if its immune system would be more attentive towards TMHs.
-And maybe this has already happened: MHC-I already detects hydrophobic
-peptides. This feature, however, may also be caused by selection
-to detect hydrophobic regions in the soluble proteins of pathogens.
-It is unknown, when focusing on TMHs only, if a signal of selection
-can be detected.
-
+% 
+% TMHs, on the other hand, also have their function constraints, 
+% yet can occur multiple time a pathogen's proteome.
+% One can safely assume a pathogen's proteome contains multiple TMHs.
+% Therefore, it may be beneficial for the host
+% if its immune system would be more attentive towards TMHs.
+% And maybe this has already happened: MHC-I already detects hydrophobic
+% peptides. This feature, however, may also be caused by selection
+% to detect hydrophobic regions in the soluble proteins of pathogens.
+% It is unknown, when focusing on TMHs only, if a signal of selection
+% can be detected.
+% 
 % \paragraph{Use of protein data in phylogenetics}
-
-When using DNA sequences, one can use a skewed rate
-in non- versus synonymous mutation
-to detect the signature of selection (\cite{murrell2015gene}).
-Using AA sequences, however, has its advantages,
-as its is closer to the actual phenotype
-selection acts upon: DNA may never be transcribed to RNA,
-or its RNA may never be translated (\cite{diz2012proteomics}).
-When using a proteome in phylogenetic research,
-we know that the majority of proteins are selected to just
-maintain their function most of the time, where
-is the time spans there is selection, only a few AAs
-can actually increase the 'fitness' of the
-protein (\cite{anisimova2009investigating}).
-There, when generalizing the dynamics of mostly purifying selection (to maintain
-a protein's function) and a short duration of positive selection,
-those genes that are selected cannot be detected (\cite{yang2000statistical}).
+% 
+% When using DNA sequences, one can use a skewed rate
+% in non- versus synonymous mutation
+% to detect the signature of selection (\cite{murrell2015gene}).
+% Using AA sequences, however, has its advantages,
+% as its is closer to the actual phenotype
+% selection acts upon: DNA may never be transcribed to RNA,
+% or its RNA may never be translated (\cite{diz2012proteomics}).
+% When using a proteome in phylogenetic research,
+% we know that the majority of proteins are selected to just
+% maintain their function most of the time, where
+% is the time spans there is selection, only a few AAs
+% can actually increase the 'fitness' of the
+% protein (\cite{anisimova2009investigating}).
+% There, when generalizing the dynamics of mostly purifying selection (to maintain
+% a protein's function) and a short duration of positive selection,
+% those genes that are selected cannot be detected (\cite{yang2000statistical}).
 

### 2021-02-17 bbbq_article.tex
Process feedback @bogeert. Thanks Geert! Progress #179
Moved Introduction part to Discussion
Process feedback until 'Measuring TMH epitopes', progress #179

@@
 % GvdB:
-% The text below here is more discussion 
-% as it is already interpretation of the findings. 
-% I would add one short paragraph, 
+% Add one short paragraph, 
 % where you reiterate the most important findings:
@@
 
-% RJCB: This was suggested for removal, until the Methods:
-% \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
-% 
-% It is important that MHC-I presents both the peptides of the 
-% (healthy) cell, as well as possible pathogen-derived peptides when the
-% cell is infected. As described above, MHC-I presents TMH-derived epitopes 
-% from the human host more often. It is unknown if MHC-I has the same
-% dedication to present epitopes that stem from TMHs derived from
-% proteins produced by pathogen, either would the pathogen
-% be a virus or 
-% a bacterium.
-% 
-% \paragraph{MHC-II is expected to present TMHs}
-% 
-% If presentation of TMHs on MHC-I would bring an evolutionary advantage 
-% in the recognition of pathogens by the immune system, 
-% it would follow that this is equally important for MHC-II, 
-% especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
-% response (\cite{novy2007cd4}). 
-% The mechanism to detect foreign TMHs by MHC-II would be unknown, 
-% similar to the discovery that MHC-I presents TMHs.
-% 
-% \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
-% 
-% It is unknown if MHC-II, like MHC-I, 
-% presents TMH-derived epitopes as often, either 
-% in humans,
-% bacteria or viruses.
-% 
-% \paragraph{Selection undetectable in whole proteome}
-% 
-% In general, on would hope that evolutionary selection results in
-% an immune system that as most attentive for loci that are
-% essential for a virus, as these will be most conserved.
-% In SARS-CoV-2, for example, there is preliminary evidence that the strongest
-% selection pressure is upon residues that changes its 
-% virulence (\cite{velazquez2020positive}).
-% These loci, however, only account for a small part of a pathogen's proteome.
-% Additionally, these essential parts differ widely between pathogens.
-% Because of this scarcity and variance in targets, 
-% one can imagine that the human immune system 
-% is not tailored to detect these sites, 
-% as hinted by upon by the aforementioned influenza study.
-% 
-% \paragraph{Selection may be detectable in TMHs}
-% 
-% TMHs, on the other hand, also have their function constraints, 
-% yet can occur multiple time a pathogen's proteome.
-% One can safely assume a pathogen's proteome contains multiple TMHs.
-% Therefore, it may be beneficial for the host
-% if its immune system would be more attentive towards TMHs.
-% And maybe this has already happened: MHC-I already detects hydrophobic
-% peptides. This feature, however, may also be caused by selection
-% to detect hydrophobic regions in the soluble proteins of pathogens.
-% It is unknown, when focusing on TMHs only, if a signal of selection
-% can be detected.
-% 
-% \paragraph{Use of protein data in phylogenetics}
-% 
-% When using DNA sequences, one can use a skewed rate
-% in non- versus synonymous mutation
-% to detect the signature of selection (\cite{murrell2015gene}).
-% Using AA sequences, however, has its advantages,
-% as its is closer to the actual phenotype
-% selection acts upon: DNA may never be transcribed to RNA,
-% or its RNA may never be translated (\cite{diz2012proteomics}).
-% When using a proteome in phylogenetic research,
-% we know that the majority of proteins are selected to just
-% maintain their function most of the time, where
-% is the time spans there is selection, only a few AAs
-% can actually increase the 'fitness' of the
-% protein (\cite{anisimova2009investigating}).
-% There, when generalizing the dynamics of mostly purifying selection (to maintain
-% a protein's function) and a short duration of positive selection,
-% those genes that are selected cannot be detected (\cite{yang2000statistical}).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 To determine the percentages of epitopes overlapping
-with TMHs, we use reference proteomes of 
+with TMHs, we used reference proteomes of 
 a human, viral and bacterial nature.
@@
 Instead of using the full human reference proteome, we
-use the representative human reference proteome 
+used the representative human reference proteome 
 as used previously (\cite{bianchi2017}).
@@
 To determine if epitopes derived from TMHs are presented at all,
-we started from epitopes sequenced from elution studies
+we started from epitopes identified in elution studies
 for MHC-I (\cite{schellens2015comprehensive}) 
 and MHC-II (\cite{bergseng2015different}).
-For each epitope, the human representative reference proteome
+For each of these peptides, the human representative reference proteome
 was searched for its origin.
-We kept only the epitopes of which were found exactly once
-on the proteome.
+We kept only the epitopes which were uniquely present
+in the proteome.
 We predicted the topology of the proteome
@@
 
-To determine the percentages of MHC-I and MHC-II epitopes overlap
+To determine the percentages of MHC-I and MHC-II epitopes overlapping
 with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
 To summarize: from a proteome all possible 9-mers are derived. For each
-of these peptides, it is measured if it has been part of a 
-TMH (at least overlapping 1 residue), 
-as well as if the peptide is a binder to an MHC-I haplotype.
-For MHC-II, 14-mers are used, as these are the most frequently occurring
+of these peptides, it was determined if it was part of a 
+TMH (at least overlapping with 1 residue), 
+and if the peptide bound to an MHC-I haplotype.
+For MHC-II, 14-mers were used, as these are the most frequently occurring
 epitope length (\cite{bergseng2015different}).
 
-This study differs in some aspects, described below in more detail,
-which is the definition of what a binder is,
-the addition of using MHC-II haplotypes, the software used,
-and the significance level to determine if TMH-derived peptides are as 
-likely to be a binder as expected by chance.
-These deviations are either a refinement of the previous method or
+This study differs in some aspects from our previous work (\cite{bianchi2017}), 
+described below in more detail.
+The main differences are our definition of what a so-called binder is, 
+the inclusion of both MHC-I and MHC-II haplotypes, 
+the prediction software used, 
+and the significance level to determine if TMH-derived peptides will bind.
+These deviations are either a refinement of our previous approach or
 a pragmatic choice made due to the extension of the original experiment.
-Additionally, instead of only using a human proteome, in this study
-also a viral and bacterial proteome are analyzed.
+Additionally, instead of only using a human proteome, this study
+also includes a viral and bacterial proteome.
 
 The definition of a binder differs from \cite{bianchi2017}:
-in this study a peptide is called a binder if, within a haplotype, 
+in our current study a peptide was called a binder if, within a certain haplotype, 
 any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
@@
 % See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
+We believe our revised definition precludes bias of proteins 
+that give rise to no or only very few MHC epitopes
 
 The 13 MHC-I haplotypes used in this study are the same as 
-the ones in \cite{bianchi2017}.
-The MHC-II haplotypes used additionally are selected 
+we used in our previous study (\cite{bianchi2017}).
+The MHC-II haplotypes used additionally were selected 
 to occur with a phenotypic frequency of at least 14\% in
@@
 resulting in 21 haplotypes.
-When using an MHC-II haplotype, instead of using 9-mers, 14-mers are
+When using an MHC-II haplotype, instead of using 9-mers, 14-mers were
 used, as these are the most common MHC-II epitope size,
@@
 
+% RJCB: moved this from Introduction to Discussion.
+%
+% \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
+% 
+% It is important that MHC-I presents both the peptides of the 
+% (healthy) cell, as well as possible pathogen-derived peptides when the
+% cell is infected. As described above, MHC-I presents TMH-derived epitopes 
+% from the human host more often. It is unknown if MHC-I has the same
+% dedication to present epitopes that stem from TMHs derived from
+% proteins produced by pathogen, either would the pathogen
+% be a virus or 
+% a bacterium.
+% 
+% \paragraph{MHC-II is expected to present TMHs}
+% 
+% If presentation of TMHs on MHC-I would bring an evolutionary advantage 
+% in the recognition of pathogens by the immune system, 
+% it would follow that this is equally important for MHC-II, 
+% especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
+% response (\cite{novy2007cd4}). 
+% The mechanism to detect foreign TMHs by MHC-II would be unknown, 
+% similar to the discovery that MHC-I presents TMHs.
+% 
+% \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
+% 
+% It is unknown if MHC-II, like MHC-I, 
+% presents TMH-derived epitopes as often, either 
+% in humans,
+% bacteria or viruses.
+% 
+% \paragraph{Selection undetectable in whole proteome}
+% 
+% In general, on would hope that evolutionary selection results in
+% an immune system that as most attentive for loci that are
+% essential for a virus, as these will be most conserved.
+% In SARS-CoV-2, for example, there is preliminary evidence that the strongest
+% selection pressure is upon residues that changes its 
+% virulence (\cite{velazquez2020positive}).
+% These loci, however, only account for a small part of a pathogen's proteome.
+% Additionally, these essential parts differ widely between pathogens.
+% Because of this scarcity and variance in targets, 
+% one can imagine that the human immune system 
+% is not tailored to detect these sites, 
+% as hinted by upon by the aforementioned influenza study.
+% 
+% \paragraph{Selection may be detectable in TMHs}
+% 
+% TMHs, on the other hand, also have their function constraints, 
+% yet can occur multiple time a pathogen's proteome.
+% One can safely assume a pathogen's proteome contains multiple TMHs.
+% Therefore, it may be beneficial for the host
+% if its immune system would be more attentive towards TMHs.
+% And maybe this has already happened: MHC-I already detects hydrophobic
+% peptides. This feature, however, may also be caused by selection
+% to detect hydrophobic regions in the soluble proteins of pathogens.
+% It is unknown, when focusing on TMHs only, if a signal of selection
+% can be detected.
+% 
+% \paragraph{Use of protein data in phylogenetics}
+% 
+% When using DNA sequences, one can use a skewed rate
+% in non- versus synonymous mutation
+% to detect the signature of selection (\cite{murrell2015gene}).
+% Using AA sequences, however, has its advantages,
+% as its is closer to the actual phenotype
+% selection acts upon: DNA may never be transcribed to RNA,
+% or its RNA may never be translated (\cite{diz2012proteomics}).
+% When using a proteome in phylogenetic research,
+% we know that the majority of proteins are selected to just
+% maintain their function most of the time, where
+% is the time spans there is selection, only a few AAs
+% can actually increase the 'fitness' of the
+% protein (\cite{anisimova2009investigating}).
+% There, when generalizing the dynamics of mostly purifying selection (to maintain
+% a protein's function) and a short duration of positive selection,
+% those genes that are selected cannot be detected (\cite{yang2000statistical}).
+
 \subsection{Elution studies}

### 2021-02-17 bbbq_article.tex
Process feedback 'Results', 'Elution studies', progress #179

@@
 To detect the evolutionary conservation of TMHs, 
-we collect human mutations and 
-tally their predicted location.
+we collected human mutations and 
+tallied their predicted location.
 To be more precise, we collected single nucleotide
@@
 that resulted in the substitution of one amino acid.
-As we know the sequence of the protein a SNP acts upon,
-we predict that protein's topology and tally if
-the SNP occurred in a TMH or not, 
-after which we use statistics to determine if there are
+We then predicted that protein's topology and tallied if
+the SNP occurred within a TMH or not, 
+after which we used statistics to determine if there are
 more, less or an equal amount of SNPs in TMHs
 as a measure of evolutionary conservation.
-This workflow is discussed in more detail below and also
-depicted in figure \ref{fig:snp_workflow}).
+This workflow is discussed in more detail below.
 
@@
 As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases are used,
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used,
 which are 'gene', to find the gene names of membrane proteins, 
@@
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
-It is estimated that approximately ten percent (ranging from 2 to 36 percent) 
-are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}),
-yet if this has been corrected for (if possible) in the last decade is unknown.
+% GvdB: Not really relevant. As these are sequencing artefacts in the DNA, these should be randomly distributed and not affect the conclusions.
+% GvdB: In addition, you use the most frequent SNPs, which minimizes the chance of those sequencing artifacts.
+% It is estimated that approximately ten percent (ranging from 2 to 36 percent) 
+% are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}),
+% yet if this has been corrected for (if possible) in the last decade is unknown.
 
@@
 To retrieve the data from these databases the
-\verb;rentrez; R package (\cite{rentrez}) is used
-to calls the NCBI website's API. To provide for a 
+\verb;rentrez; R package (\cite{rentrez}) was used
+that calls the NCBI website's API. To provide for a 
 stable user experience for all users, 
@@
 
-The first query is a call to the 'gene' database for the 
-term 'membrane protein' (in all fields) for the organism Homo sapiens.
-This results in 1129 gene IDs (note that this number varies through time).
-The next query is a call to the 'gene' database 
+The first query was a call to the 'gene' database for the 
+term 'membrane protein' (in all fields) for the organism \textbf{Homo sapiens}.
+This resulted in 1129 gene IDs.
+The next query was a call to the 'gene' database 
 to obtain the gene names from the gene IDs.
@@
 was limited to the first 250 variations per gene,
-resulting in 282250 variations. A variation need not always be a SNP,
+resulting in 282250 variations. 
+A variation needs not always be a SNP,
 as dbSNP also catalogs other DNA alterations, such as, among others, insertions,
@@
 protein sequence.
-Of each protein sequence, the protein topology is determined 
+Of each protein sequence, the protein topology was determined 
 using PureseqTM.
 From the topology and the known location of the SNP, 
-we score the location (i.e. transmembrane or soluble protein region) 
+we score the location (i.e. TMH or soluble protein region) 
 where the change occurred.
@@
     Overview of all software used in this research.
-    'LUT' is an abbreviation for 'Look-up table'. 
   }
@@
 For this research, the scientific literature was explored 
-to find the most recent free and open source (FOSS) prediction software.
+to identify the most recent free and open source (FOSS) prediction software.
 This was done by searching for papers that (1) reference older
@@
 As a starting point, a review paper was used.
+For all software needed, we found no studies that compared contemporary tools 
+in their prediction quality.
 
@@
 There are multiple computational tools developed to predict which
-parts of membrane proteins are TMH.
-In 2001, multiple tools to do so have been compared (\cite{moller2001evaluation}),
+parts of a protein forms a TMH.
+In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
 of which TMHMM (\cite{krogh2001predicting}) turned out to be the best, 
 as is used in the previous study (\cite{bianchi2017}).
-TMHMM has a restrictive software license and is nearly two
+However, TMHMM has a restrictive software license and is nearly two
 decades old.
-Due to this, also PureseqTM (\cite{wang2019efficient}),
-is used in this study, which has been recently developed
+Therefore, PureseqTM (\cite{wang2019efficient}),
+was also used in this study, which has been more recently developed
 and has a free software license.
@@
 NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
-A tool developed later is \verb;epitope-prediction; (\cite{bianchi2017}),
-which uses a stabilized matrix method (\cite{kim2009derivation}).
-The most recent and promising FOSS tool available now appears
-to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
-and MHC-II prediction.
+We used the same tool as used in our earlier study, \verb;epitope-prediction; (\cite{bianchi2017}),
 
@@
 gave rise to the most accurate predictions.
-Later tools are NetTepi (\cite{trolle2014nettepi}),
-a genetic-algorithm-based ensemble method (\cite{zhang2015accurate})
-and (again) the very recent MHCnuggets (\cite{shao2020high}).
+The most recent and promising FOSS tool available now appears
+to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
+and MHC-II prediction, which we used for MHC-II predictions.
 
@@
 
-The R package 'tmhmm' was developed to do the same topology
-predictions as the earlier study (that used 'TMHMM'), yet in a reproducible way.
-'TMHMM' has a restrictive software license and allows a user
+The R package 'tmhmm' was developed to do the similar topology
+predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
+'TMHMM' has a restrictive software license \richel{ref} and allows a user
 to download a pre-compiled executable after confirmation he/she
@@
 
-To be able to call, from R, the TMH prediction software 'PureseqTM',
+To be able to call, from R, the TMH prediction software 'PureseqTM' \richel{ref},
 which is written in C, the package 'pureseqtmr' has been developed. 
@@
 MHCnuggets is a free and open-source Python package to predict 
-epitope affinity for many MHC-I and MHC-II variants.
+epitope affinity for many MHC-I and MHC-II variants \richel{ref}.
 The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
@@
 
-To reproduce the full experiment as done in this paper,
+To reproduce the full experiment presented in this paper,
 the functions needed are bundled in the 'bbbq' R package.
@@
 
-Table \ref{tab:elution} shows the percentage of epitopes derived from a TMH
-found in the two elution studies, for the two different kind of topology
-prediction tools. Whatever the tool or MHC class, at least 100 epitopes
-can be uniquely map to human representative reference proteome.
+% GvdB: A narrative was missing a bit. Why was this done? 
+To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
+we reanalyzed published peptide elution studies.
+Table \ref{tab:elution} shows the percentage of epitopes derived 
+from a TMH
+found in HNC-I and MHC-II elution studies \richel{refs}, 
+for the two topology prediction tools TMHMM and PureseqTM \richel{refs}. 
+Regardless of the prediction, 
+at least 100 epitopes were predicted to be derived from a TMH. 
+% GvdB: Also add a conclusion at the end of the paragraph.
+From these findings, we conclude that PureseqTM and TMHMM 
+result in prediction of a similar number of TMHs 
+and that peptides derived from TMHs have been identified for both MHC-I and -II.
 

### 2021-02-17 bbbq_article.tex
Process feedback. Thanks Geert! Progress #179

@@
 
-Figure \ref{fig:1} shows the percentages of MHC-I epitopes overlapping 
-with TMHs for our human, viral and bacterial proteome.
+We recently showed that peptides derived from TMHs of human proteins 
+were predicted to be over-presented in MHC-I. 
+Here, we wondered if this finding would be general 
+and if peptides stemming from TMHs of microbial pathogens 
+would also be over-presented.
+Figure \ref{fig:1} shows the percentages of MHC-I epitopes predicted to be overlapping 
+with TMHs for our human, viral and bacterial proteome \richel{GvdB: you didn't explain yet which virus and bacteria}.
 See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
 for the exact TMH and epitope counts.
+\richel{GvdB: Also provide a short conclusion}
 
@@
 
+\richel{GvdB: Again a rationale}
 Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
@@
 for the TMH and epitope counts.
+\richel{GvdB: And a conclusion}
 
@@
 
-We obtained 1129 gene names associated with the key phrase 'membrane protein'.
-These genes are linked to 5203 proteins, of which 2754
+In the final set of experiments, we addressed the question whether TMHs would be more evolutionary conserved than soluble proteins regions by comparing the occurrences of SNPs in the genome coding for TMHs and soluble protein regions.
+We obtained 1,129 gene names associated with the phrase 'membrane protein'.
+These genes are linked to 5,203 proteins
+\richel{GvdB: Why so many? Isoforms?}, of which 2,754
 are predicted to be transmembrane proteins.
-We obtained 4811 variations that resulted in an
-amino acids substitution, of which 2568 were located 
+We obtained 4,811 SNPs that resulted in an
+amino acids substitution, of which 2,568 were located 
 in predicted transmembrane proteins.
-Per protein, we predicted the percentage of TMH
+Per protein, we predicted the percentages of the gene coding for TMHs and soluble regions
 and calculated the percentage of amino acids substitutions
-occurring in transmembrane regions.
-These points, one per protein, were plotted as depicted in figure
+occurring in TMHs.
+Per protein, this percentage was plotted as a function of the percentage TMH in figure
 \ref{fig:f_snps_found_and_expected}.
-We did a linear fit on the data, with and without
-membrane-associated proteins, and added a 95\%
+We performed a linear regression analysis on the data, with and without
+membrane-associated proteins \richel{Dit is nog niet goed uitgelegd}, and calculated the 95\%
 confidence interval.
-Also, we added the line \inlineequation{y = x},
-which indicates the linear fit that would be expected if
-SNPs occur just as likely within transmembrane and soluble
-protein domains. Already these trend lines
-hint that variations occur more often in soluble 
+Since the regression curve clearly deviated from the line of equality,
+which would be expected if SNPs occur just as likely within transmembrane and soluble protein domains,
+this suggests that variations occur more often in soluble 
 than in transmembrane protein regions.
@@
 distribution. We first calculated the expected number of SNPs
-found in TMHs, which is 61457 SNPs multiplied by 6.8\% 
-of all proteins being TMH, resulting in the null hypothesis
-of 4203 SNPs being in the TMH. 
+found in TMHs, which is 61,457 SNPs \richel{GvdB: I don’t completely follow this. Based on what?} and multiplied this number by 6.8\% 
+of all proteins being a TMH, resulting in the null hypothesis
+of 4203 SNPs occurring in TMHs. 
 Note that including or excluding
@@
 of a successful trial is zero (there is zero chance
-to find a SNP in TMHs if there are none).
+to find a SNP in TMHs if there are none) \richel{GvdB: Also too complex for biologists. Explain in bit more words}.
 The actually number of SNPs found in TMHs
@@
 Using a alpha value of $0.05$, as is common practice for an uninformed
-expectation, we concluded that there are significantly less SNPs
+expectation, we conclude that there are significantly less SNPs
 in TMHs as expected by chance.
@@
     if TMHs and soluble protein regions are equally conserved.
+    \richel{GvdB: As discussed, I would simply plot 1 and multiple (>1) TMHs.}
   }
@@
 We decided to split this analysis based on the number of TMHs
-a protein has. We hypothesize that single-spanners (i.e. proteins
+a protein has \richel{Why? Provide brief rationale again. Thus TMH-TMH interactions or something}. We hypothesize that single-spanners (i.e. proteins
 with one TMH predicted) are less conserved, when compared to multi-spanners,
-as single spanners just need to span a membrane, where some multi-spanners
-have enzymatically active sites embedded within their TMHs and thus
-have an additional structural constraint.
+as single spanners just need to span a membrane, while multi-spanners
+might have interactions between their TMHs, 
+for example to accommodate active sites, and 
+thus might have additional structural constraints.
 As can be seen in figure \ref{fig:f_snps_found_and_expected_per_n_tmhs}, 
 there is no clear pattern to be observed. 
+\richel{GvdB: Rewrite after pooling >1 TMHs}
 

### 2021-02-17 bbbq_article.tex
All @bogeert feedback is in the TeX. Thanks Geert! Progress #179

@@
 from a TMH
-found in HNC-I and MHC-II elution studies \richel{refs}, 
+found in MHC-I and MHC-II elution studies \richel{refs}, 
 for the two topology prediction tools TMHMM and PureseqTM \richel{refs}. 
@@
 
-We found a general pattern in the
-presentation of epitopes that overlap with a TMH.
-For humans, a bacterial and viral pathogen,
-as well as for MHC-I and MHC-II,
-we found that these are presented more often 
-than expected by chance.
+From this study, two important conclusions can be drawn. 
+First, the MHC over-presentation of TMHs is likely a general feature 
+and predicted to occur for most haplotypes of both MHC-I and -II 
+and for humans as well as bacterial and viral pathogens. 
+Second, TMHs are more evolutionary conserved that soluble protein motifs, 
+at least in the human proteome.
 
@@
 influences the dynamics of a pathogen (see,
-for example, \cite{eccleston2017host} for HIV).
+for example, \cite{eccleston2017host} for HIV) \richel{GvdB: Too vague. How?}.
 The bioinformatics pipeline developed in this
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  GvdB:
+
+  Discussion section needs some work, 
+  as here interpretation of the results needs to be given. 
+  What is the consequence of the findings? Why is it important.
+
+  Now it is just a list of potential reasons why the results might be flawed. 
+  While this is in principle OK, 
+  it is also nitpicking as some things affect only a limited number of 
+  proteins (like selenomethionines) 
+  or you’d expect changes in the other direction (wrong prediction of TMHs for bacteria).
+
+  So offer interpretations. 
+  The 2nd part of the introduction would fit here well. I can also help with this.
+}
 
@@
 software and confirm these have similar results.
-Two biological sources of noise
+Two biological sources of noise \richel{GvdB: What do you mean by noise?}
 are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
@@
 
+\richel{GvdB: Sounds a but irrelevant to me. Do you really want to mention this in the discussion?}
 % \paragraph{Removal of selenoproteins}
-
+%
 We have removed the 25 selenoproteins found in the human
@@
 
+\richel{GvdB: Also irrelevant. Only 1 protein}
 % \paragraph{Topology prediction of TRDD1}
-
+%
 PureseqTM does not predict the topology
@@
 
+\richel{GvdB: Also not really relevant. This is just a minor thing, only relevant if we would see large differences between human and bacteria (but we see not)}
 % \paragraph{Bacteria have a different cell membrane}
-
+%
 In this experiment we predicted epitopes that overlap with 
@@
 
+\richel{GvdB: This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
 % \paragraph{synonymous mutations}
-
+%
 In our evolutionary experiment, 
@@
 
+\richel{GvdB: Also not really relevant}
 % \paragraph{False positives in SNPs}
-
+%
 Regarding the evolutionary conservation of TMHs using SNPs,
@@
 As the NCBI API constrains its users to three calls per second,
-we had to limited the extent of our analysis
-by downloading SNPs to approximately 300 
-hours (i.e. the length of a Christmas holiday).
-
+we had to limited the extent of our analysis.
+
+\richel{GvdB: This can go to the methods or results.}
 % \paragraph{Selection of SNPs}
-
+%
 We selected to analyze SNPs with the lowest SNP ID first.
@@
 for allowing us to read paywalled articles while working from home.
+\richel{GvdB: Should we include Maxim? He did the alignment of Myc tub. Let’s also ask Frans.}
 

### 2021-02-18 bbbq_article.tex
Removed obsolete text
Add results separated by number of spanners

@@
 
-% \paragraph{Immune response by MHC-I}
-
-% GvdB: Not really relevant.
-You don’t really need to explain the role of DCs.
-What is far more important is the concept of what an MHC is: a protein complex that contains a fold, called the peptide-binding groove, which presents peptides.
-Also immediately explain that MHC-I presents peptides of about 9-amino acids in length to cytolytic T cells, whereas MHC-II presents longer peptides to helper T cells.
-
-% Antigen-presenting cells such as dendritic cells
-% use the MHC class I molecules to 
-% present randomly sampled peptides fragments,
-% as created by the immunoproteasome (\cite{chapiro2006destructive}).
-% If a cell gets infected by an intracellular pathogen, also the virus' peptides
-% will also be presented at the cell surface. The foreign viral antigen is detected 
-% by cytotoxic T lymphocytes, which will kill the infected cell.
-
-% \paragraph{Immune response by MHC-II, contrast with APC}
-
-% GvdB: Also not really relevant
-
-% All pathogens can be detected by the foreign proteins on 
-% their (bacterial or fungal) cell walls or (viral) envelope.
-% For bacteria and fungi, this is the main mode of their detection,
-% as these do not infect a cell with their (foreign) DNA.
-% T and B cells express MHC-II to detect foreign peptide fragments.
-% An immune response is started when MHC-II detects a pathogen.
+% GvdB: Not really relevant. You don’t really need to explain the role of DCs. 
+% What is far more important is the concept of what an MHC is: 
+% a protein complex that contains a fold, called the peptide-binding groove, 
+% which presents peptides. Also immediately explain that MHC-I 
+% presents peptides of about 9-amino acids in length to cytolytic T cells, 
+% whereas MHC-II presents longer peptides to helper T cells.
 
@@
 with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
-
-% \paragraph{MHC-I presents TMH-derived epitopes in humans more often}
-
-%For MHC-I, it was found that predicted epitopes derived 
-%from human transmembrane helices (TMHs)
-%are over-presented by all 5 HLA-A and 
-%most of 8 HLA-B super-types (\cite{bianchi2017}),
-%due the high hydrophobicity as is required for a TMH to span
-%the (hydrophobic) cell membrane, combined with the fact that the MHC
-%binding cleft is hydrophobic as well.
-%
-%It is unknown if the presentation of hydrophobic/TMH residues
-%is evolutionary selected for.
-%One reason to expect a reduced variability (and hence 
-%evolutionary conservation) in TMHs, is
-%that these are restricted in their evolution
-%by the functional requirement to span a lipid bilayer.
-%Due to this, the TMHs that pathogens possess, 
-%have a lower chance to develop an escape mutation,
-%as many mutations will result in a dysfunctional TMH.
 
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_n_tmhs.png}
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
   \caption{
-    Percentage of SNPs found in TMHs,
-    separated per number of TMHs.
+    Percentage of SNPs found in TMHs per spanner.
     The dashed diagonal lines show the expected linear trend line
     if TMHs and soluble protein regions are equally conserved.
-    \richel{GvdB: As discussed, I would simply plot 1 and multiple (>1) TMHs.}
   }
-  \label{fig:f_snps_found_and_expected_per_n_tmhs}
+  \label{fig:f_snps_found_and_expected_per_spanner}
 \end{figure}
@@
 thus might have additional structural constraints.
-As can be seen in figure \ref{fig:f_snps_found_and_expected_per_n_tmhs}, 
-there is no clear pattern to be observed. 
-\richel{GvdB: Rewrite after pooling >1 TMHs}
+As can be seen in figure \ref{fig:f_snps_found_and_expected_per_spanner}, 
+both single- and multi-spanners are evolutionarily conserved.
 

### 2021-02-18 bbbq_article.tex
Range in percentage of TMH in human proteome is caused by the use of multiple TMH prediction tools. Fix #180

@@
 TMHs are common structures in the proteins of humans and pathogens. 
-For example, 
-15-39% 
-\richel{Why wide range? Issue 180}
-of the proteins in humans 
-are predicted to be a membrane protein 
-containing at least one TMH (\cite{ahram2006estimation}). 
+Different TMH prediction tools estimate
+that 15-39% of the proteins in humans 
+contain at least one TMH (\cite{ahram2006estimation}). 
 TMHs can also be predicted from a protein sequence 
@@
 We performed a linear regression analysis on the data, with and without
-membrane-associated proteins \richel{Dit is nog niet goed uitgelegd}, and calculated the 95\%
+membrane-associated proteins \richel{GvdB: explain clearer}, and calculated the 95\%
 confidence interval.
@@
   it is also nitpicking as some things affect only a limited number of 
-  proteins (like selenomethionines) 
+  proteins (like selenoproteins) 
   or you’d expect changes in the other direction (wrong prediction of TMHs for bacteria).
@@
   So offer interpretations. 
-  The 2nd part of the introduction would fit here well. I can also help with this.
+  The second part of the introduction would fit here well. I can also help with this.
 }

### 2021-02-19 bbbq_article.tex
Selenoproteins to supmat

@@
 
-To determine the percentages of epitopes overlapping
-with TMHs, we used reference proteomes of 
-a human, viral and bacterial nature.
-An overview of Uniprot IDs is shown in table \ref{tab:uniprot_ids}.
-Instead of using the full human reference proteome, we
-used the representative human reference proteome 
-as used previously (\cite{bianchi2017}).
+We used a human, viral and bacterial reference proteome to 
+determine the percentages of epitopes overlapping
+with TMHs.
+We used a human reference proteome
+with UniProt ID UP000005640\_9606, which
+is a representative set of the full proteome
+and used is our previous study (\cite{bianchi2017}).
+We used SARS-CoV-2 as our viral reference proteome
+with UniProt ID UP000464024 and
+Mycobacterium tuberculosis (MTb) 
+with UniProt ID UP000001584.
+We chose these pathogens based on their impact on the human population.
 
@@
 the epitopes within a 
-\emph{proteome} (see tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+\emph{proteome} (see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2}
 for values), whereas the original study defined
@@
 Figure \ref{fig:1} shows the percentages of MHC-I epitopes predicted to be overlapping 
-with TMHs for our human, viral and bacterial proteome \richel{GvdB: you didn't explain yet which virus and bacteria}.
+with TMHs for our human, viral and bacterial proteome.
 See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
@@
 
-\richel{GvdB: Sounds a but irrelevant to me. Do you really want to mention this in the discussion?}
-% \paragraph{Removal of selenoproteins}
-%
-We have removed the 25 selenoproteins found in the human
-reference proteome, as both prediction tools struggle with 
-selenocysteine, the 21st amino acid incorporated in human proteins.
-TMHMM gives a clear error message, forcing us to remove
-these selenoproteins from the analysis.
-How PureseqTM deals with selenocysteine is unclear,
-as it gives no error, nor is it mentioned anywhere,
-nor do the authors respond to emails. Due to this, we removed the
-selenoproteins from our analysis.
-
 \richel{GvdB: Also irrelevant. Only 1 protein}
@@
 for allowing us to read paywalled articles while working from home.
-\richel{GvdB: Should we include Maxim? He did the alignment of Myc tub. Let’s also ask Frans.}
+\richel{GvdB: Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}
 
@@
 
-% \paragraph{TMH prediction software}
-
-The earlier study uses TMHMM, which has a restricted software license.
-Instead, we chose to use PureseqTM, which is free and open source.
-
 % Percentage of spots and spots that overlap with a TMH
@@
 which is used here again. For MHC-II IC50 predictions, the
-'netmhc2pan' R package is used, which calls NetMHC2pan.
-Both are packages are free and open-source.
+'mhcnuggetsr' R package is used.
 
@@
 In this study, a peptide is defined a binder (for a haplotype), 
-if \emp{within a proteome} in which it is found, 
-is within the peptides with the 2\% lowest IC50 values.
+if within a \emp{proteome} in which it is found, 
+is within the peptides with the 2\% lowest IC50 values,
+where our previous study used the lowest 2\% IC50 values
+within each \emp{protein}.
 Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
@@
 
-% \paragraph{Unclear handing of sequences with U}
-
-Within the human reference proteome, there are 30 sequences that have a
-selenocysteine (with the symbol 'U').
-See \url{https://github.com/richelbilderbeek/bbbq_article/issues/158}.
+% \paragraph{Selenoproteins}
+
+Our previous study used the TMHMM web server
+to predict TMHs.
+The desktop version of TMHMM, however, gives an
+error message on the 25 selenoproteins found in the human
+reference proteome.
+For the sake of reproducible research, we use the desktop version (as
+we can call it from scripts) and, due to this, we removed the
+selenoproteins from this analysis.
 

### 2021-02-19 bbbq_article.tex
Moved mono- and di-peptides to supmat
Moved minor discussion point that software is tuned to predict TMHs in human to supmat
Moved noise in dbSNP to minor discussion in supmat
Improved 'Selection of SNPs'

@@
 variations associated with the gene name.
+
 The number of variations
@@
 a single amino acid substitution.
+
+% \paragraph{Selection of SNPs}
+%
+We selected to analyze SNPs with the lowest SNP ID first.
+This means we have a bias for picking SNPs with
+an earlier discovery date.
+We expect this bias to be biologically irrelevant.
+To verify this expectation, we determined the
+relative position of each SNP in a protein,
+as shown in figure \ref{fig:snp_rel_pos}.
+
 Per SNP, the 'protein' NCBI database was queried for the
@@
 
-\richel{GvdB: Also irrelevant. Only 1 protein}
-% \paragraph{Topology prediction of TRDD1}
-%
+\subsection{Evolutionary conservation}
+
+\richel{GvdB: This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
+% \paragraph{synonymous mutations}
+%
+In our evolutionary experiment, 
+we removed variations that were synonymous mutations (i.e.
+resulted in the same amino acid, from a different genetic code) 
+from our analysis.
+There is evidence, however, that these synonymous mutations
+do have an effect and may even be evolutionary selected 
+for (\cite{hunt2009silent}).
+As the possible effect of synonymous mutations is ignored by our
+topology prediction software, we do so as well.
+
+% \paragraph{Low number of SNPs}
+
+The NCBI dbSNP database contains millions of SNPs.
+As the NCBI API constrains its users to three calls per second,
+we had to limited the extent of our analysis.
+
+% We concluded that the
+% epitopes that MHC-I presents are [as/not as] likely 
+% to be derived from TMH within either a human host and its bacterial pathogen.
+% Because a bacterium does not infect a cell, thus its peptides
+% will not be presented by MHC-I, this result is [unexpected/expected]
+% 
+% We aimed our evolutionary experiment at TMHs, because these can
+% be predicted well from a protein structure,
+% are common structures and are present in all pathogens. 
+% We could have done the same experiment on beta-turn,
+% as also these can be predicted well (\cite{petersen2010netturnp}),
+% are common structures and are present in all pathogens.
+% 
+% The human immune system and human pathogen are in an evolutionary
+% arms race: our immune systems is selected for the detection
+% of pathogens, whereas pathogens are selected to avoid detection.
+% From a pathogen's point of view, however, this struggle 
+% is of only minor importance:
+% in seasonal influenza, for example, the selection pressure
+% exerted by the immune system was only limited (\cite{han2019individual}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acknowledgments}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We thank the Center for Information Technology of the University 
+of Groningen for its support and for providing access to the Peregrine 
+high performance computing cluster. 
+Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
+for allowing us to read paywalled articles while working from home.
+\richel{GvdB: Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Data Accessibility}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+All code is archived at \url{http://github.com/richelbilderbeek/someplace},
+with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Authors' contributions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+RJCB and FB conceived the idea for this research. 
+RJCB wrote the code.
+RJCB and FB wrote the article.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Bibliography
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% MEE style
+\bibliographystyle{mee}
+\bibliography{bbbq_article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\appendix
+\section{Supplementary materials}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Differences with Bianchi et al., 2017}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A part of this study does the same analysis as Bianchi et al., 2017.
+Here we describe the deviations, which are about the use of different
+software and the use of a different definition of what a binder is.
+
+% Percentage of spots and spots that overlap with a TMH
+\input{bbbq_1_smart_results/table_f_tmh_2.latex}
+
+% \paragraph{IC50 prediction software}
+
+The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
+that has been trained on MHC-I haplotypes only,
+which is used here again. For MHC-II IC50 predictions, the
+'mhcnuggetsr' R package is used.
+
+% \paragraph{Definition of what a binder is}
+
+The earlier study defines a peptide a binder (for a haplotype), 
+if \emp{within the peptide} in which it is found, 
+is within the peptides with the 2\% lowest IC50 values.
+This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
+where the binders are written to file.
+
+In this study, a peptide is defined a binder (for a haplotype), 
+if within a \emp{proteome} in which it is found, 
+is within the peptides with the 2\% lowest IC50 values,
+where our previous study used the lowest 2\% IC50 values
+within each \emp{protein}.
+Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
+for a binder per haplotype.
+
+% \paragraph{Selenoproteins}
+
+Our previous study used the TMHMM web server
+to predict TMHs.
+The desktop version of TMHMM, however, gives an
+error message on the 25 selenoproteins found in the human
+reference proteome.
+For the sake of reproducible research, we use the desktop version (as
+we can call it from scripts) and, due to this, we removed the
+selenoproteins from this analysis.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Minor methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 PureseqTM does not predict the topology
@@
 however, is two peptides long. 
-We corrected for this by predicting that this protein
-is completely cytosolic.
-
-\richel{GvdB: Also not really relevant. This is just a minor thing, only relevant if we would see large differences between human and bacteria (but we see not)}
+The R package \verb;pureseqtmr;, however, 
+predicts that mono- and di-peptides are cytosolic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Minor discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % \paragraph{Bacteria have a different cell membrane}
-%
+
 In this experiment we predicted epitopes that overlap with 
@@
 
-\subsection{Evolutionary conservation}
-
-\richel{GvdB: This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
-% \paragraph{synonymous mutations}
-%
-In our evolutionary experiment, 
-we removed variations that were synonymous mutations (i.e.
-resulted in the same amino acid, from a different genetic code) 
-from our analysis.
-There is evidence, however, that these synonymous mutations
-do have an effect and may even be evolutionary selected 
-for (\cite{hunt2009silent}).
-As the possible effect of synonymous mutations is ignored by our
-topology prediction software, we do so as well.
-
-\richel{GvdB: Also not really relevant}
 % \paragraph{False positives in SNPs}
-%
+
 Regarding the evolutionary conservation of TMHs using SNPs,
@@
 regions around these, hence we expect this not to affect our results.
-
-% \paragraph{Low number of SNPs}
-
-The NCBI dbSNP database contains millions of SNPs.
-As the NCBI API constrains its users to three calls per second,
-we had to limited the extent of our analysis.
-
-\richel{GvdB: This can go to the methods or results.}
-% \paragraph{Selection of SNPs}
-%
-We selected to analyze SNPs with the lowest SNP ID first.
-This means we have a bias for picking SNPs with
-an earlier discovery date.
-We expect this bias to be biologically irrelevant
-and we find no clear evidence of a bias, as
-shown in figure \ref{fig:n_tmhs_per_protein}.
-Still, the best way to interpret this figure
-is open to discussion.
-
-% We concluded that the
-% epitopes that MHC-I presents are [as/not as] likely 
-% to be derived from TMH within either a human host and its bacterial pathogen.
-% Because a bacterium does not infect a cell, thus its peptides
-% will not be presented by MHC-I, this result is [unexpected/expected]
-% 
-% We aimed our evolutionary experiment at TMHs, because these can
-% be predicted well from a protein structure,
-% are common structures and are present in all pathogens. 
-% We could have done the same experiment on beta-turn,
-% as also these can be predicted well (\cite{petersen2010netturnp}),
-% are common structures and are present in all pathogens.
-% 
-% The human immune system and human pathogen are in an evolutionary
-% arms race: our immune systems is selected for the detection
-% of pathogens, whereas pathogens are selected to avoid detection.
-% From a pathogen's point of view, however, this struggle 
-% is of only minor importance:
-% in seasonal influenza, for example, the selection pressure
-% exerted by the immune system was only limited (\cite{han2019individual}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We thank the Center for Information Technology of the University 
-of Groningen for its support and for providing access to the Peregrine 
-high performance computing cluster. 
-Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
-for allowing us to read paywalled articles while working from home.
-\richel{GvdB: Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Data Accessibility}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-All code is archived at \url{http://github.com/richelbilderbeek/someplace},
-with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Authors' contributions}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-RJCB and FB conceived the idea for this research. 
-RJCB wrote the code.
-RJCB and FB wrote the article.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Bibliography
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% MEE style
-\bibliographystyle{mee}
-\bibliography{bbbq_article}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\appendix
-\section{Supplementary materials}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Differences with Bianchi et al., 2017}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A part of this study does the same analysis as Bianchi et al., 2017.
-Here we describe the deviations, which are about the use of different
-software and the use of a different definition of what a binder is.
-
-% Percentage of spots and spots that overlap with a TMH
-\input{bbbq_1_smart_results/table_f_tmh_2.latex}
-
-% \paragraph{IC50 prediction software}
-
-The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
-that has been trained on MHC-I haplotypes only,
-which is used here again. For MHC-II IC50 predictions, the
-'mhcnuggetsr' R package is used.
-
-% \paragraph{Definition of what a binder is}
-
-The earlier study defines a peptide a binder (for a haplotype), 
-if \emp{within the peptide} in which it is found, 
-is within the peptides with the 2\% lowest IC50 values.
-This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
-where the binders are written to file.
-
-In this study, a peptide is defined a binder (for a haplotype), 
-if within a \emp{proteome} in which it is found, 
-is within the peptides with the 2\% lowest IC50 values,
-where our previous study used the lowest 2\% IC50 values
-within each \emp{protein}.
-Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
-for a binder per haplotype.
-
-% \paragraph{Selenoproteins}
-
-Our previous study used the TMHMM web server
-to predict TMHs.
-The desktop version of TMHMM, however, gives an
-error message on the 25 selenoproteins found in the human
-reference proteome.
-For the sake of reproducible research, we use the desktop version (as
-we can call it from scripts) and, due to this, we removed the
-selenoproteins from this analysis.
 
@@
 
+% \paragraph{Relative position of SNPs}
+
+To verify if SNPs were sampled uniformly
+over proteins, we show the distribution 
+of the relative position in figure \ref{fig:snp_rel_pos}.
+We find no clear evidence of a bias.
+
 %

### 2021-02-22 bbbq_article.tex
Add refs

@@
 predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
-'TMHMM' has a restrictive software license \richel{ref} and allows a user
+'TMHMM' has a restrictive software license (\cite{krogh2001predicting}) and allows a user
 to download a pre-compiled executable after confirmation he/she
@@
 
-To be able to call, from R, the TMH prediction software 'PureseqTM' \richel{ref},
+To be able to call, from R, the TMH prediction 
+software 'PureseqTM' (\cite{wang2019efficient}),
 which is written in C, the package 'pureseqtmr' has been developed. 
@@
 MHCnuggets is a free and open-source Python package to predict 
-epitope affinity for many MHC-I and MHC-II variants \richel{ref}.
+epitope affinity for many MHC-I and MHC-II variants (\cite{shao2020high}).
 The R package 'mhcnuggetsr' allows one to install and use MHCnuggets

### 2021-02-22 bbbq_article.tex
Process feedback

@@
 so-called escape mutations, 
-will not affect MHC presentation for all haplotypes \richel{Reference needed}.
+will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
 
@@
 Due to this, there are constraints in the types of amino acids 
-that can be present in the TMH \richel{ref}. 
+that can be present in the TMH (\cite{jones1994model}). 
 Therefore, the TMHs of pathogens 

### 2021-02-22 bbbq_article.tex
Process feedback Geert

@@
 from a TMH
-found in MHC-I and MHC-II elution studies \richel{refs}, 
-for the two topology prediction tools TMHMM and PureseqTM \richel{refs}. 
+found in MHC-I and MHC-II elution 
+studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} respectively), 
+for the two topology prediction tools TMHMM and 
+PureseqTM (\cite{krogh2001predicting} and \cite{wang2019efficient} respectively). 
 Regardless of the prediction, 
@@
 for the exact TMH and epitope counts.
-\richel{GvdB: Also provide a short conclusion}
+We found that also for pathogen epitopes derived from TMHs are
+over-presented. 
+The extent of the over-presentation is similar between the haplotypes:
+if a haplotype if prone to present TMH-derived epitopes from
+human proteins, it is also prone to present such epitopes of
+bacterial and viral proteins.
 
@@
 
-\richel{GvdB: Again a rationale}
+Where TMH-derived epitopes are over-presented by MHC-I,
+we wondered if this finding would also generalize to MHC-II.
 Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
@@
 for the TMH and epitope counts.
-\richel{GvdB: And a conclusion}
+We show that MHC-II has the same pattern of over-presentation
+as MHC-I, yet this over-presentation is more extreme.
+Also for MHC-II, we observe that the extent the which TMH-derived
+epitopes are presented is similar between the haplotypes, regardless
+of the protein source.
 

### 2021-02-22 bbbq_article.tex
Process feedback

@@
 We obtained 1,129 gene names associated with the phrase 'membrane protein'.
-These genes are linked to 5,203 proteins
-\richel{GvdB: Why so many? Isoforms?}, of which 2,754
-are predicted to be transmembrane proteins.
+These genes are linked to 5,203 protein isoforms, 
+of which 2,754 are predicted to be transmembrane proteins.
 We obtained 4,811 SNPs that resulted in an
@@
 in predicted transmembrane proteins.
-Per protein, we predicted the percentages of the gene coding for TMHs and soluble regions
-and calculated the percentage of amino acids substitutions
-occurring in TMHs.
-Per protein, this percentage was plotted as a function of the percentage TMH in figure
-\ref{fig:f_snps_found_and_expected}.
-We performed a linear regression analysis on the data, with and without
-membrane-associated proteins \richel{GvdB: explain clearer}, and calculated the 95\%
-confidence interval.
-Since the regression curve clearly deviated from the line of equality,
-which would be expected if SNPs occur just as likely within transmembrane and soluble protein domains,
-this suggests that variations occur more often in soluble 
-than in transmembrane protein regions.
-
-To test whether SNPs occur just as likely in soluble as
+Per protein, we calculated two percentages: 
+(1) the percentage of the gene coding for TMHs, 
+and (2) the percentage of SNPs occurring in TMHs.
+Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
+We performed a linear regression analyses on the data twice,
+one with and one without membrane-associated proteins,
+and added a 95\% confidence interval for both of these regressions.
+Would the confidence interval of the regression line 
+overlap with the line of equality,
+this would be expected if SNPs occur just as likely within TMHs and soluble protein domains.
+Our analysis shows that SNPs occur more often in soluble domains
+than in TMHs.
+The effect size is that per 1 SNP found in soluble protein
+domains, one finds 0.925 SNPs in TMHs.
+
+To statistically test whether SNPs occur just as likely in soluble as
 in transmembrane protein domains, we used a Poisson binomial
-distribution. We first calculated the expected number of SNPs
+distribution.
+
+We first calculated the expected number of SNPs
 found in TMHs, which is 61,457 SNPs \richel{GvdB: I don’t completely follow this. Based on what?} and multiplied this number by 6.8\% 
@@
 in TMHs as expected by chance.
-The effect size is that per 1 SNP found in soluble protein
-domains, one finds 0.925 SNPs in TMHs.
 

### 2021-02-23 bbbq_article.tex
Show the percentages of TMH

@@
 
-In the final set of experiments, we addressed the question whether TMHs would be more evolutionary conserved than soluble proteins regions by comparing the occurrences of SNPs in the genome coding for TMHs and soluble protein regions.
+In the final set of experiments, 
+we addressed the question whether TMHs 
+would be more evolutionary conserved than soluble proteins regions 
+by comparing the occurrences of SNPs in the genome coding 
+for TMHs and soluble protein regions. 
 We obtained 1,129 gene names associated with the phrase 'membrane protein'.
-These genes are linked to 5,203 protein isoforms, 
-of which 2,754 are predicted to be transmembrane proteins.
-We obtained 4,811 SNPs that resulted in an
-amino acids substitution, of which 2,568 were located 
+These genes are linked to 4,811 protein isoforms, 
+of which 2,568 are predicted to be transmembrane proteins.
+We obtained 38,882 SNPs that resulted in an
+amino acids substitution, of which 21,738 were located 
 in predicted transmembrane proteins.
@@
 distribution.
+We calculated the number of SNPs that are expected to be
+found in TMHs by chance, $E(n_{SNP\_TMH})$, when including
+and excluding soluble proteins (with symbols
+$E(n_{SNP\_TMH}^{+})$ and $E(n_{SNP\_TMH}^{-})$ respectively).
+The number of SNPs that are expected to be
+found in TMHs by chance when including soluble proteins,
+$E(n_{SNP\_TMH}^{+})$ 
+equals the number of (all) SNPs, $n_{SNP}^{+} = 38,882$, multiplied by the fraction
+of all proteins that are TMH, $f_{TMH}^{+} = 9.96\%$,
+resulting in 
+$E(n_{SNP\_TMH}^{+}) = n_{SNP}^{+} \cdot f_{TMH}^{+} \approx 3873$
+The number of SNPs that are expected to be
+found in TMHs by chance when excluding soluble proteins,
+$E(n_{SNP\_TMH}^{-})$ 
+equals the number of SNPs in TMHs, $n_{SNP}^{-} = 21,738$, 
+multiplied by the fraction of the TMPs that are TMH, $f_{TMH}^{-} = 18.66\%$,
+$E(n_{SNP\_TMH}^{-}) = n_{SNP}^{-} \cdot f_{TMH}^{-} \approx 4056$.
+The discrepancy between the two expectations
+is caused by the distribution of SNPs over the soluble and
+transmembrane proteins.
+See figure \ref{fig:f_tmh_ncbi} to see which percentages
+are TMH of the transmembrane proteins.
+See figure HIERO to see the distribution of SNPs over the
+soluble and transmembrane proteins.
 
@@
 \begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
+  \caption{
+    Distribution of the percentages of TMH of transmembrane proteins.
+  }
+  \label{fig:f_tmh_ncbi}
+\end{figure}
+
+\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_proteins_per_gene_name.png}

### 2021-02-24 bbbq_article.tex
Show number of SNPs in TMHs

@@
 is caused by the distribution of SNPs over the soluble and
-transmembrane proteins.
+transmembrane proteins: 
+there are 38,882 SNPs found in 4811 proteins (i.e. $\approx 8.08$ SNP per protein),
+of which 21,738 SNPs in 2568 integral membrane proteins (i.e. $\approx 8.46$
+SNP per integral membrane protein).
+
+\richel{How many SNPs in TMHs?}
 See figure \ref{fig:f_tmh_ncbi} to see which percentages
 are TMH of the transmembrane proteins.
-See figure HIERO to see the distribution of SNPs over the
-soluble and transmembrane proteins.
 

### 2021-02-24 bbbq_article.tex
Use TMP for 'transmembrane protein'

@@
 These genes are linked to 4,811 protein isoforms, 
-of which 2,568 are predicted to be transmembrane proteins.
+of which 2,568 are predicted to be transmembrane proteins (TMPs),
+where the others are predicted to be membrane-associated proteins.
 We obtained 38,882 SNPs that resulted in an
@@
 there are 38,882 SNPs found in 4811 proteins (i.e. $\approx 8.08$ SNP per protein),
-of which 21,738 SNPs in 2568 integral membrane proteins (i.e. $\approx 8.46$
-SNP per integral membrane protein).
+of which 21,738 SNPs in 2568 TMPs (i.e. $\approx 8.46$
+SNP per TMP).
 
@@
 the transmembrane regions predicted in bacterial proteins
-are actually part of an integral membrane protein.
+are actually part of a TMH.
 For the purpose of this study, we assume the 
@@
     Histogram of the number of TMHs predicted per protein,
-    for the integral membrane proteins used.
+    for the trans-membrane proteins used.
   }

### 2021-02-26 bbbq_article.tex
Add number of SNPs in NCBI database

@@
 The number of variations
-was limited to the first 250 variations per gene,
-resulting in 282250 variations. 
+was limited to the first 250 variations per gene (see figure
+\ref{fig:snps_per_gene_name_ncbi} for the number of SNPs
+within the dbSNP database for these gene names),
+resulting in \richel{282250} variations.
+
 A variation needs not always be a SNP,
@@
 by comparing the occurrences of SNPs in the genome coding 
-for TMHs and soluble protein regions. 
-We obtained 1,129 gene names associated with the phrase 'membrane protein'.
+for TMHs and soluble protein regions within membrane proteins. 
+We obtained 1,129 gene names associated with the phrase 'membrane protein',
+which are genes for both membrane-associated proteins (MAPs, no TMH) and 
+transmembrane proteins (TMPs, at least 1 TMH).
 These genes are linked to 4,811 protein isoforms, 
-of which 2,568 are predicted to be transmembrane proteins (TMPs),
-where the others are predicted to be membrane-associated proteins.
+of which 2,568 are predicted to be TMPs (hence 
+2,243 proteins are predicted to be MAPs).
 We obtained 38,882 SNPs that resulted in an
 amino acids substitution, of which 21,738 were located 
-in predicted transmembrane proteins.
+in predicted TMPs.
 Per protein, we calculated two percentages: 
-(1) the percentage of the gene coding for TMHs, 
+(1) the percentage of the protein coding for TMHs, 
 and (2) the percentage of SNPs occurring in TMHs.
@@
 We performed a linear regression analyses on the data twice,
-one with and one without membrane-associated proteins,
+one with and one without MAPs,
 and added a 95\% confidence interval for both of these regressions.
@@
 
-To statistically test whether SNPs occur just as likely in soluble as
-in transmembrane protein domains, we used a Poisson binomial
-distribution.
-We calculated the number of SNPs that are expected to be
-found in TMHs by chance, $E(n_{SNP\_TMH})$, when including
-and excluding soluble proteins (with symbols
-$E(n_{SNP\_TMH}^{+})$ and $E(n_{SNP\_TMH}^{-})$ respectively).
-The number of SNPs that are expected to be
-found in TMHs by chance when including soluble proteins,
-$E(n_{SNP\_TMH}^{+})$ 
-equals the number of (all) SNPs, $n_{SNP}^{+} = 38,882$, multiplied by the fraction
-of all proteins that are TMH, $f_{TMH}^{+} = 9.96\%$,
-resulting in 
-$E(n_{SNP\_TMH}^{+}) = n_{SNP}^{+} \cdot f_{TMH}^{+} \approx 3873$
-The number of SNPs that are expected to be
-found in TMHs by chance when excluding soluble proteins,
-$E(n_{SNP\_TMH}^{-})$ 
-equals the number of SNPs in TMHs, $n_{SNP}^{-} = 21,738$, 
-multiplied by the fraction of the TMPs that are TMH, $f_{TMH}^{-} = 18.66\%$,
-$E(n_{SNP\_TMH}^{-}) = n_{SNP}^{-} \cdot f_{TMH}^{-} \approx 4056$.
-The discrepancy between the two expectations
-is caused by the distribution of SNPs over the soluble and
-transmembrane proteins: 
-there are 38,882 SNPs found in 4811 proteins (i.e. $\approx 8.08$ SNP per protein),
-of which 21,738 SNPs in 2568 TMPs (i.e. $\approx 8.46$
-SNP per TMP).
+To statistically test whether SNPs occur 
+just as likely in soluble domains as
+in TMHs, we first calculated the number of SNPs that are expected to be
+found in TMHs of TMPs by chance, $E(n_{SNP\_TMH})$,
+equals the number of SNPs in TMHs, $n_{SNP} = 21,738$, 
+multiplied by the fraction of the TMPs that are TMH, $f_{TMH} = 18.66\%$,
+$E(n_{SNP\_TMH}) = n_{SNP} \cdot f_{TMH} \approx 4056$.
+
+The actually number of SNPs found in TMHs
+was $n_{SNP\_TMH} = 3,903$, which is lower than the expectation
+of $E(n_{SNP\_TMH}) \approx 4056$.
+To test whether this difference is significant,
+we used a binomial Poisson distribution, where the number of trails,
+$n$ equals the number of SNPs, which is 21,738. The probability of success
+for TMP $i$, $p\_i$, is the percentage of each TMP that is TMH.
+We calculated the probability of having 3,903 or less successes.
+We used the type I error cut-off value of $\alpha = 2.5\%$.
+The chance to find, within TMHs, this amount or less SNPs 
+equals \richel{recalculate}
+% $1.6201 \cdot 10^{-9}$.
+
+Using a alpha value of $0.05$, as is common practice for an uninformed
+expectation, we conclude that there are significantly less SNPs
+in TMHs as expected by chance.
 
@@
 are TMH of the transmembrane proteins.
-
-We first calculated the expected number of SNPs
-found in TMHs, which is 61,457 SNPs \richel{GvdB: I don’t completely follow this. Based on what?} and multiplied this number by 6.8\% 
-of all proteins being a TMH, resulting in the null hypothesis
-of 4203 SNPs occurring in TMHs. 
-Note that including or excluding
-the membrane associated proteins 
-has no effect on this statistical test, as the chance 
-of a successful trial is zero (there is zero chance
-to find a SNP in TMHs if there are none) \richel{GvdB: Also too complex for biologists. Explain in bit more words}.
-The actually number of SNPs found in TMHs
-was 3889, which is lower than the null hypothesis. 
-The chance to find, within TMHs, this amount or less SNPs 
-equals $1.6201 \cdot 10^{-9}$.
-Using a alpha value of $0.05$, as is common practice for an uninformed
-expectation, we conclude that there are significantly less SNPs
-in TMHs as expected by chance.
 
@@
 \begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_ncbi.png}
+  \caption{
+    Distribution of the number of SNPs per gene name in the NCBI database.
+  }
+  \label{fig:snps_per_gene_name_ncbi}
+\end{figure}
+
+\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}

### 2021-02-26 bbbq_article.tex
Update and add snps_per_gene_name_processed
@bogeert has his own color

@@
 within the dbSNP database for these gene names),
-resulting in \richel{282250} variations.
-
+resulting in 60,683 variations at the protein level.
 A variation needs not always be a SNP,
@@
 We select only the variations that result in a SNP for
-a single amino acid substitution.
+a single amino acid substitution, resulting in 38,882 SNPs.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{I suggest to move this section to the Supplementary Materials}
 
@@
 
+\richel{I suggest to move this section to the Supplementary Materials}
+
 The R programming language is used for the complete 
@@
 
-% GvdB: A narrative was missing a bit. Why was this done? 
+\geert{A narrative was missing a bit. Why was this done?}
 To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
@@
 at least 100 epitopes were predicted to be derived from a TMH. 
-% GvdB: Also add a conclusion at the end of the paragraph.
+\geert{Also add a conclusion at the end of the paragraph.}
 From these findings, we conclude that PureseqTM and TMHMM 
@@
 influences the dynamics of a pathogen (see,
-for example, \cite{eccleston2017host} for HIV) \richel{GvdB: Too vague. How?}.
+for example, \cite{eccleston2017host} for HIV) 
+\geert{Too vague. How?}.
 The bioinformatics pipeline developed in this
@@
 
-\richel{
-  GvdB:
-
+\geert{
   Discussion section needs some work, 
@@
 software and confirm these have similar results.
-Two biological sources of noise \richel{GvdB: What do you mean by noise?}
+Two biological sources of noise \geert{What do you mean by noise?}
 are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
@@
 
-\richel{GvdB: This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
+\geert{This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
 % \paragraph{synonymous mutations}
@@
 for allowing us to read paywalled articles while working from home.
-\richel{GvdB: Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}
+\geert{Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}
 
@@
 \begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_processed.png}
+  \caption{
+    Distribution of the number of protein variations and SNPs per gene name processed.
+  }
+  \label{fig:snps_per_gene_name_processed}
+\end{figure}
+
+\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}

### 2021-03-02 bbbq_article.tex
Add stats

@@
 term 'membrane protein' (in all fields) for the organism \textbf{Homo sapiens}.
-This resulted in 1129 gene IDs.
+This resulted in 1077 gene IDs
+\richel{
+  This was December 2020.
+  At 2021-03-01, one will get 1130 gene IDs.
+  Also, one of the gene IDs that was valid back then,
+  has been obsoleted.
+}.
 The next query was a call to the 'gene' database 
@@
 
+% \paragraph{Statistics}
+
+Table \ref{tab:snp_stats} shows the statistics for all
+SNPs, where tables \ref{tab:snp_stats_per_spanner_single}
+and \ref{tab:snp_stats_per_spanner_multi} show the
+statistics for only single-spanners and multi-spanners respectively.
+These probabilities are also displayed visually in 
+figures \ref{fig:ppoisbinom} (all proteins), \ref{fig:ppoisbinom_single} (single-spanners)
+and \ref{fig:ppoisbinom_multi} (multi-spanners).
+
+% Label: tab:snp_stats
+\input{ncbi_peregrine_results/snp_stats.tex}
+
+% Label: tab:snp_stats_per_spanner_single
+\input{ncbi_peregrine_results/snp_stats_per_spanner_single.tex}
+
+% Label: tab:snp_stats_per_spanner_multi
+\input{ncbi_peregrine_results/snp_stats_per_spanner_multi.tex}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/ppoisbinom.png}
+  \caption{
+    Probabilities for all SNPs
+  }
+  \label{fig:ppoisbinom}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/ppoisbinom_single.png}
+  \caption{
+    Probabilities for SNPs in single-spanners
+  }
+  \label{fig:ppoisbinom_single}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/ppoisbinom_multi.png}
+  \caption{
+    Probabilities for SNPs in multi-spanners
+  }
+  \label{fig:ppoisbinom_multi}
+\end{figure}
+
 %

### 2021-03-02 ncbi_peregrine_results/snp_stats_per_spanner_multi.tex
Add stats

\begin{table}

\caption{\label{tab:snp_stats_per_spanner_multi}Statistics for the multi-spanners. p = p value. n = number of SNPs in multi-spanners. n\_success = number of SNPs found in TMHs of multi-spanners (dashed blue line). E(n\_success) = expected number of SNPs to be found in TMHs of multi-spanners  (dashed red line). }
\centering
\begin{tabular}[t]{l|l}
\hline
parameter & value\\
\hline
p & 4.388709e-14\\
\hline
n & 13386\\
\hline
n\_success & 3377\\
\hline
E(n\_success) & 3767.26\\
\hline
\end{tabular}
\end{table}

### 2021-03-02 ncbi_peregrine_results/snp_stats_per_spanner_single.tex
Add stats

\begin{table}

\caption{\label{tab:snp_stats_per_spanner_single}Statistics for the single-spanners. p = p value. n = number of SNPs in single-spanners. n\_success = number of SNPs found in TMHs of single-spanners (dashed blue line). E(n\_success) = expected number of SNPs to be found in TMHs of single-spanners  (dashed red line). }
\centering
\begin{tabular}[t]{l|l}
\hline
parameter & value\\
\hline
p & 0.3457596\\
\hline
n & 8190\\
\hline
n\_success & 454\\
\hline
E(n\_success) & 462.6681\\
\hline
\end{tabular}
\end{table}

### 2021-03-02 to_doc.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+pandoc -o bbbq_article.docx bbbq_article.tex

### 2021-03-03 bbbq_article.tex
Plot SNP conservation in simpler bar plots, fix #186

@@
 
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
+  \caption{
+    Evolutionary conservation of SNPs in TMHs.
+  }
+  \label{fig:conservation}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
+  \caption{
+    Evolutionary conservation of SNPs in TMHs,
+    for either single- or multi-spanners
+  }
+  \label{fig:conservation_per_spanner}
+\end{figure}
+
 % \paragraph{Statistics}

### 2021-03-11 bbbq_article.tex
Plot number of SNPs per TMP, fix #143
Processed feedback Frans, thanks @fransbianchi. Fix #188

@@
 peptide fragments derived from transmembrane helices (TMHs) 
-are predicted to be presented more often than expected based on their abundance.
+are predicted to be presented more often than expected 
+based on their abundance in the proteome.
+We show that the over-presentation of TMH-derived peptides is likely general, 
+as it is predicted for diverse microbiota and viruses 
+and for both MHC-I and MHC-II
 However, the physiological reason of why membrane proteins might be 
@@
 compared to regions coding for extracellular and cytosolic protein regions. 
-Moreover, we show that the over-presentation of TMH-derived peptides is 
-likely general, as it is predicted for diverse microbial pathogens 
-and for both MHC-I and MHC-II.
 Thus, our findings suggest that T cells might respond 
 more to membrane proteins, because these are evolutionary more conserved.
-We speculate that TMHs might be less prone of the escape mutations 
+We speculate that TMHs therefor might be less prone to escape mutations 
 that enable pathogens to evade T cell responses.
@@
 
-Our immune system fights microbial pathogens, 
+Our immune system fights diseases and infections from pathogens, 
 such as fungi, bacteria or viruses. 
@@
 that develops specialized and more effective recognition of pathogens, 
-are T cells which recognize peptides derived from the pathogens 
-by use of Major Histocompatibility Complexes (MHC).
+are T cells which recognize peptides derived from 
+exogenous antigens presented on Major Histocompatibility Complexes (MHC).
 
@@
 For example, HLA-A and HLA-B have no overlap in which
-peptides they bind (\cite{lund2004definition}).
+peptides they bind (\cite{lund2004definition})
+\frans{Is dit zo?} 
+\richel{Volgens de (weggegevallen) referentie naar Lund et al., 2004, 'Definition of supertypes for {HLA} molecules using clustering of specificity matrices' wel}.
 The HLA region of humans is highly polymorphic, with hundreds 
@@
 the number of pathogenic peptides that can be detected is increased,
-as a wide variety of MHCs will be expressed,
+as a wide variety of MHCs: 
+6 MHC-I and 6-8 MHC-II molecules are expressed per individual
 each presenting their own subset of epitopes.
@@
 
-Much studies are aimed at determining which peptides are presented in MHC 
+Many studies are aimed at determining which peptides are presented in MHC 
 and will result in an immune response, 
@@
 
-This over-presentation is entirely explained by the fact 
+This over-presentation is attributed to the fact 
 that the peptide-binding groove of most MHC-I haplotypes 
@@
 Transmembrane helices (TMHs) are hydrophobic 
-as they need to span the hydrophobic lipid bilayer of cellular membranes, 
-and they consist of an alpha helix of usually 23 amino acids. 
+as they need to span the hydrophobic lipid bilayer of cellular membranes.
+They consist of an alpha helix on average 23 amino acids in length. 
 TMHs are common structures in the proteins of humans and pathogens. 
 Different TMH prediction tools estimate
-that 15-39% of the proteins in humans 
+that 15-39\% of the proteins in humans 
 contain at least one TMH (\cite{ahram2006estimation}).
@@
 stemming from soluble protein regions is unknown. 
-One possibility is that the presentation of 
+We hypothesized is that the presentation of 
 hydrophobic/TMH residues is evolutionary selected for, 
@@
 by the functional requirement to span a lipid bilayer. 
-Due to this, there are constraints in the types of amino acids 
-that can be present in the TMH (\cite{jones1994model}). 
+Due to these requirements, 
+the amino acids genuinely present in TMH 
+are limited to the ones with hydrophobic side chains (\cite{jones1994model}).
+\frans{Analyzing the occurrence of single nucleotide polymorphisms that were predicted to be inside ...}
+
 Therefore, the TMHs of pathogens 
@@
 \begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_snps_per_tmp.png}
+  \caption{
+    Histogram of the number of SNPs per trans-membrane protein.
+  }
+  \label{fig:n_tmhs_per_protein}
+\end{figure}
+
+\begin{figure}[!htbp]
   \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_tmhs_per_protein.png}

### 2021-03-11 bbbq_article.tex
Comments are visible in docx

@@
 peptide fragments derived from transmembrane helices (TMHs) 
-are predicted to be presented more often than expected 
+are predicted to be presented more often than expected,
 based on their abundance in the proteome.
@@
 
-% GvdB: Not really relevant. You don’t really need to explain the role of DCs. 
-% What is far more important is the concept of what an MHC is: 
-% a protein complex that contains a fold, called the peptide-binding groove, 
-% which presents peptides. Also immediately explain that MHC-I 
-% presents peptides of about 9-amino acids in length to cytolytic T cells, 
-% whereas MHC-II presents longer peptides to helper T cells.
-
 % \paragraph{Classification of HLA}
@@
 peptides they bind (\cite{lund2004definition})
-\frans{Is dit zo?} 
-\richel{Volgens de (weggegevallen) referentie naar Lund et al., 2004, 'Definition of supertypes for {HLA} molecules using clustering of specificity matrices' wel}.
+[FB: is that really so?]
+[RB: yes, according to the removed reference to Lund et al., 2004, 'Definition of supertypes for {HLA} molecules using clustering of specificity matrices'] 
 The HLA region of humans is highly polymorphic, with hundreds 
@@
 are limited to the ones with hydrophobic side chains (\cite{jones1994model}).
-\frans{Analyzing the occurrence of single nucleotide polymorphisms that were predicted to be inside ...}
+[FB: Analyzing the occurrence of single nucleotide polymorphisms 
+that were predicted to be inside ...]
 
@@
 This resulted in 1077 gene IDs
-\richel{
+[RJCB:
   This was December 2020.
@@
   has been obsoleted.
-}.
+].
 The next query was a call to the 'gene' database 
@@
 
-\richel{I suggest to move this section to the Supplementary Materials}
+[RJCB: I suggest to move this section to the Supplementary Materials]
 
@@
 
-\geert{A narrative was missing a bit. Why was this done?}
+[GVDB: A narrative was missing a bit. Why was this done?]
 To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
@@
 at least 100 epitopes were predicted to be derived from a TMH. 
-\geert{Also add a conclusion at the end of the paragraph.}
+[GVDB: Also add a conclusion at the end of the paragraph.]
 From these findings, we conclude that PureseqTM and TMHMM 
@@
 The chance to find, within TMHs, this amount or less SNPs 
-equals \richel{recalculate}
-% $1.6201 \cdot 10^{-9}$.
+equals [RJCB: recalculate]
+$1.6201 \cdot 10^{-9}$.
 
@@
 
-\richel{How many SNPs in TMHs?}
+[RJCB: How many SNPs in TMHs?]
 See figure \ref{fig:f_tmh_ncbi} to see which percentages
@@
 We decided to split this analysis based on the number of TMHs
-a protein has \richel{Why? Provide brief rationale again. Thus TMH-TMH interactions or something}. We hypothesize that single-spanners (i.e. proteins
+a protein has 
+[GVDB: Why? Provide brief rationale again. 
+Thus TMH-TMH interactions or something]. 
+We hypothesize that single-spanners (i.e. proteins
 with one TMH predicted) are less conserved, when compared to multi-spanners,
@@
 for example, \cite{eccleston2017host} for HIV) 
-\geert{Too vague. How?}.
+[GVDB: Too vague. How?].
 The bioinformatics pipeline developed in this
@@
 
-\geert{
+[GVDB: 
   Discussion section needs some work, 
@@
   The second part of the introduction would fit here well. I can also help with this.
-}
+]
 
@@
 software and confirm these have similar results.
-Two biological sources of noise \geert{What do you mean by noise?}
+Two biological sources of noise [GVDB: What do you mean by noise?]
 are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
@@
 
-\geert{This might be relevant, but you need to explain why it is relevant. Thus compare your findings with previous efforts to look at conservation of TMHs}
+[GVDB: 
+  This might be relevant, but you need to explain why it is relevant. 
+  Thus compare your findings with previous efforts 
+  to look at conservation of TMHs
+]
 % \paragraph{synonymous mutations}
@@
 for allowing us to read paywalled articles while working from home.
-\geert{Should we include Maxim? He did the alignment of Mycobacterium tuberculosis. Let’s also ask Frans.}
+[GVDB: 
+  Should we include Maxim? 
+  He did the alignment of Mycobacterium tuberculosis. 
+  Let’s also ask Frans.
+]
 

### 2021-03-11 bbbq_article.tex
Comments visible in doc

@@
 
-% GvdB:
-% Add one short paragraph, 
-% where you reiterate the most important findings:
-% In this study, we show that ...(and than basically 
-% the abstract in slightly different words and slightly 
-% more detail). End with a statement of why it is 
-% important (vaccin development based on TMHs?)...
+[GVDB:
+ Add one short paragraph, 
+ where you reiterate the most important findings:
+ In this study, we show that ...(and than basically 
+ the abstract in slightly different words and slightly 
+ more detail). End with a statement of why it is 
+ important (vaccine development based on TMHs?)...
+]
 
@@
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
-% GvdB: Not really relevant. As these are sequencing artefacts in the DNA, these should be randomly distributed and not affect the conclusions.
-% GvdB: In addition, you use the most frequent SNPs, which minimizes the chance of those sequencing artifacts.
-% It is estimated that approximately ten percent (ranging from 2 to 36 percent) 
-% are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}),
-% yet if this has been corrected for (if possible) in the last decade is unknown.
 
@@
 
-% RJCB: moved this from Introduction to Discussion.
-%
 % \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
-% 
-% It is important that MHC-I presents both the peptides of the 
-% (healthy) cell, as well as possible pathogen-derived peptides when the
-% cell is infected. As described above, MHC-I presents TMH-derived epitopes 
-% from the human host more often. It is unknown if MHC-I has the same
-% dedication to present epitopes that stem from TMHs derived from
-% proteins produced by pathogen, either would the pathogen
-% be a virus or 
-% a bacterium.
-% 
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+
+It is important that MHC-I presents both the peptides of the 
+(healthy) cell, as well as possible pathogen-derived peptides when the
+cell is infected. As described above, MHC-I presents TMH-derived epitopes 
+from the human host more often. It is unknown if MHC-I has the same
+dedication to present epitopes that stem from TMHs derived from
+proteins produced by pathogen, either would the pathogen
+be a virus or 
+a bacterium.
+
 % \paragraph{MHC-II is expected to present TMHs}
-% 
-% If presentation of TMHs on MHC-I would bring an evolutionary advantage 
-% in the recognition of pathogens by the immune system, 
-% it would follow that this is equally important for MHC-II, 
-% especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
-% response (\cite{novy2007cd4}). 
-% The mechanism to detect foreign TMHs by MHC-II would be unknown, 
-% similar to the discovery that MHC-I presents TMHs.
-% 
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+
+If presentation of TMHs on MHC-I would bring an evolutionary advantage 
+in the recognition of pathogens by the immune system, 
+it would follow that this is equally important for MHC-II, 
+especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
+response (\cite{novy2007cd4}). 
+The mechanism to detect foreign TMHs by MHC-II would be unknown, 
+similar to the discovery that MHC-I presents TMHs.
+
 % \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
-% 
-% It is unknown if MHC-II, like MHC-I, 
-% presents TMH-derived epitopes as often, either 
-% in humans,
-% bacteria or viruses.
-% 
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+
+It is unknown if MHC-II, like MHC-I, 
+presents TMH-derived epitopes as often, either 
+in humans,
+bacteria or viruses.
+
 % \paragraph{Selection undetectable in whole proteome}
-% 
-% In general, on would hope that evolutionary selection results in
-% an immune system that as most attentive for loci that are
-% essential for a virus, as these will be most conserved.
-% In SARS-CoV-2, for example, there is preliminary evidence that the strongest
-% selection pressure is upon residues that changes its 
-% virulence (\cite{velazquez2020positive}).
-% These loci, however, only account for a small part of a pathogen's proteome.
-% Additionally, these essential parts differ widely between pathogens.
-% Because of this scarcity and variance in targets, 
-% one can imagine that the human immune system 
-% is not tailored to detect these sites, 
-% as hinted by upon by the aforementioned influenza study.
-% 
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+
+In general, on would hope that evolutionary selection results in
+an immune system that as most attentive for loci that are
+essential for a virus, as these will be most conserved.
+In SARS-CoV-2, for example, there is preliminary evidence that the strongest
+selection pressure is upon residues that changes its 
+virulence (\cite{velazquez2020positive}).
+These loci, however, only account for a small part of a pathogen's proteome.
+Additionally, these essential parts differ widely between pathogens.
+Because of this scarcity and variance in targets, 
+one can imagine that the human immune system 
+is not tailored to detect these sites, 
+ as hinted by upon by the aforementioned influenza study.
+
 % \paragraph{Selection may be detectable in TMHs}
-% 
-% TMHs, on the other hand, also have their function constraints, 
-% yet can occur multiple time a pathogen's proteome.
-% One can safely assume a pathogen's proteome contains multiple TMHs.
-% Therefore, it may be beneficial for the host
-% if its immune system would be more attentive towards TMHs.
-% And maybe this has already happened: MHC-I already detects hydrophobic
-% peptides. This feature, however, may also be caused by selection
-% to detect hydrophobic regions in the soluble proteins of pathogens.
-% It is unknown, when focusing on TMHs only, if a signal of selection
-% can be detected.
-% 
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+
+TMHs, on the other hand, also have their function constraints, 
+yet can occur multiple time a pathogen's proteome.
+One can safely assume a pathogen's proteome contains multiple TMHs.
+Therefore, it may be beneficial for the host
+if its immune system would be more attentive towards TMHs.
+And maybe this has already happened: MHC-I already detects hydrophobic
+peptides. This feature, however, may also be caused by selection
+to detect hydrophobic regions in the soluble proteins of pathogens.
+It is unknown, when focusing on TMHs only, if a signal of selection
+can be detected.
+
 % \paragraph{Use of protein data in phylogenetics}
-% 
-% When using DNA sequences, one can use a skewed rate
-% in non- versus synonymous mutation
-% to detect the signature of selection (\cite{murrell2015gene}).
-% Using AA sequences, however, has its advantages,
-% as its is closer to the actual phenotype
-% selection acts upon: DNA may never be transcribed to RNA,
-% or its RNA may never be translated (\cite{diz2012proteomics}).
-% When using a proteome in phylogenetic research,
-% we know that the majority of proteins are selected to just
-% maintain their function most of the time, where
-% is the time spans there is selection, only a few AAs
-% can actually increase the 'fitness' of the
-% protein (\cite{anisimova2009investigating}).
-% There, when generalizing the dynamics of mostly purifying selection (to maintain
-% a protein's function) and a short duration of positive selection,
-% those genes that are selected cannot be detected (\cite{yang2000statistical}).
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+
+When using DNA sequences, one can use a skewed rate
+in non- versus synonymous mutation
+to detect the signature of selection (\cite{murrell2015gene}).
+Using AA sequences, however, has its advantages,
+as its is closer to the actual phenotype
+selection acts upon: DNA may never be transcribed to RNA,
+or its RNA may never be translated (\cite{diz2012proteomics}).
+When using a proteome in phylogenetic research,
+we know that the majority of proteins are selected to just
+maintain their function most of the time, where
+is the time spans there is selection, only a few AAs
+can actually increase the 'fitness' of the
+protein (\cite{anisimova2009investigating}).
+There, when generalizing the dynamics of mostly purifying selection (to maintain
+a protein's function) and a short duration of positive selection,
+those genes that are selected cannot be detected (\cite{yang2000statistical}).
+
+[RJCB: that was the last paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
 
@@
 
-% We concluded that the
-% epitopes that MHC-I presents are [as/not as] likely 
-% to be derived from TMH within either a human host and its bacterial pathogen.
-% Because a bacterium does not infect a cell, thus its peptides
-% will not be presented by MHC-I, this result is [unexpected/expected]
-% 
-% We aimed our evolutionary experiment at TMHs, because these can
-% be predicted well from a protein structure,
-% are common structures and are present in all pathogens. 
-% We could have done the same experiment on beta-turn,
-% as also these can be predicted well (\cite{petersen2010netturnp}),
-% are common structures and are present in all pathogens.
-% 
-% The human immune system and human pathogen are in an evolutionary
-% arms race: our immune systems is selected for the detection
-% of pathogens, whereas pathogens are selected to avoid detection.
-% From a pathogen's point of view, however, this struggle 
-% is of only minor importance:
-% in seasonal influenza, for example, the selection pressure
-% exerted by the immune system was only limited (\cite{han2019individual}).
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+We concluded that the
+epitopes that MHC-I presents are [as/not as] likely 
+to be derived from TMH within either a human host and its bacterial pathogen.
+Because a bacterium does not infect a cell, thus its peptides
+will not be presented by MHC-I, this result is [unexpected/expected]
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+We aimed our evolutionary experiment at TMHs, because these can
+be predicted well from a protein structure,
+are common structures and are present in all pathogens. 
+We could have done the same experiment on beta-turn,
+as also these can be predicted well (\cite{petersen2010netturnp}),
+are common structures and are present in all pathogens.
+
+[RJCB: this paragraph I ruthlessly moved
+from introduction to here, as was suggested by GVDB]
+The human immune system and human pathogen are in an evolutionary
+arms race: our immune systems is selected for the detection
+of pathogens, whereas pathogens are selected to avoid detection.
+From a pathogen's point of view, however, this struggle 
+is of only minor importance:
+in seasonal influenza, for example, the selection pressure
+exerted by the immune system was only limited (\cite{han2019individual}).
 

### 2021-03-11 ncbi_peregrine_results/ncbi_counts.tex
Add constants

\begin{table}

\caption{\label{tab:ncbi_counts}Amounts. map = MAP = membrane associated protein. tmp = TMP = transmembrane protein. in\_tmh = in transmembrane helix (TMH) of TMP. in\_sol = in soluble region of TMP. perc\_tmh = percentage transmembrane helix (TMH). }
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r}
\hline
what & raw & all\_proteins & map & tmp & in\_tmh & in\_sol\\
\hline
Number of variations & 60931 & 37831 & 16623 & 21208 & 3803 & 17405\\
\hline
Number of unique variations & 60544 & 37630 & 16606 & 21024 & 3789 & 17235\\
\hline
Number of unique SNPs & NA & 9621 & 4219 & 6026 & 1140 & 4936\\
\hline
Number of unique gene names & 953 & 911 & 457 & 605 & 325 & 590\\
\hline
Number of unique protein names & 5163 & 4780 & 2227 & 2553 & 1280 & 2467\\
\hline
Percentage TMH & NA & 10 & 0 & 19 & 26 & 18\\
\hline
\end{tabular}
\end{table}

### 2021-03-11 ncbi_peregrine_results/ncbi_counts.tex
Update constants

@@
 \caption{\label{tab:ncbi_counts}Amounts.
+raw = all variations, including DNA variations.
+all\_proteins = all proteins.
 map = MAP = membrane associated protein.
@@
 perc\_tmh = percentage transmembrane helix (TMH).
-}
+The number of unique SNPs and unique gene names does not add up for MAP and TMP, as one SNP may work on multiple isoforms, some of which can be MAP where others can be TMP.
+The number of unique SNPs and unique gene names does not add up for TMPs in TMH and TMP in soluble regions, as one SNP may work on multiple isoforms, some of which can be MAP where others can be TMP, and some SNPs fall in a TMH, where others are found in soluble regions}
 \centering

### 2021-03-18 bbbq_article.tex
Add constants
Move information from figures to table

@@
 
-[GVDB: A narrative was missing a bit. Why was this done?]
 To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
@@
 at least 100 epitopes were predicted to be derived from a TMH. 
-[GVDB: Also add a conclusion at the end of the paragraph.]
-From these findings, we conclude that PureseqTM and TMHMM 
-result in prediction of a similar number of TMHs 
-and that peptides derived from TMHs have been identified for both MHC-I and -II.
+From these findings, it is robustly predicted that
+epitopes derived from TMHs are presented by B cells [For Frans: please check
+that the article by Bergseng, 2015, 'Different binding motifs of the celiac 
+disease-associated HLA molecules DQ2. 5, DQ2. 2, and DQ7. 5 revealed by 
+relative quantitative proteomics of endogenous peptide repertoires'
+is appropriate. As far as I can see, it is, but I may overlook something. Richel].
 
@@
 would also be over-presented.
-Figure \ref{fig:1} shows the percentages of MHC-I epitopes predicted to be overlapping 
+Figure \ref{fig:1} shows the percentages of MHC-I epitopes 
+predicted to be overlapping 
 with TMHs for our human, viral and bacterial proteome.
@@
 Also for MHC-II, we observe that the extent the which TMH-derived
-epitopes are presented is similar between the haplotypes, regardless
-of the protein source.
+epitopes are presented is similar between the haplotypes, for
+a human, bacterial and viral proteome.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
+  \caption{
+    Evolutionary conservation of SNPs in TMHs.
+  }
+  \label{fig:conservation}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
+  \caption{
+    Evolutionary conservation of SNPs in TMHs,
+    for either single- or multi-spanners
+  }
+  \label{fig:conservation_per_spanner}
+\end{figure}
 
@@
 for TMHs and soluble protein regions within membrane proteins.
-We obtained 1,129 gene names associated with the phrase 'membrane protein',
+We obtained 1,129 [For Frans: I would enjoy to write 'approx 1k' here, would that
+be acceptable? I do refer to a table with the exact values
+later on anyways. Richel] gene names associated with the phrase 'membrane protein',
 which are genes for both membrane-associated proteins (MAPs, no TMH) and 
-transmembrane proteins (TMPs, at least 1 TMH).
-These genes are linked to 4,811 protein isoforms, 
-of which 2,568 are predicted to be TMPs (hence 
-2,243 proteins are predicted to be MAPs).
-We obtained 38,882 SNPs that resulted in an
-amino acids substitution, of which 21,738 were located 
+transmembrane proteins (TMPs, at least one TMH).
+These genes are linked to 4,780 protein isoforms, 
+of which 2,553 are predicted to be TMPs (hence 
+2,237 proteins are predicted to be MAPs).
+We obtained 37,630 SNPs that resulted in an
+amino acids substitution, of which 21,024 were located 
 in predicted TMPs.
+See table \ref{tab:ncbi_counts} for an overview of all amounts.
+
 Per protein, we calculated two percentages: 
@@
 overlap with the line of equality,
-this would be expected if SNPs occur just as likely within TMHs and soluble protein domains.
-Our analysis shows that SNPs occur more often in soluble domains
-than in TMHs.
-The effect size is that per 1 SNP found in soluble protein
-domains, one finds 0.925 SNPs in TMHs.
+this would be expected if SNPs occur just as likely within TMHs 
+and soluble protein domains.
+This analysis gives the first hint that SNPs 
+occur more often in soluble domains than in TMHs.
 
@@
 $E(n_{SNP\_TMH}) = n_{SNP} \cdot f_{TMH} \approx 4056$.
+
+The effect size is that, 
+when randomly selecting SNPs, 
+per 100 SNPs found in soluble protein domains, 
+one finds 91 SNPs in TMHs,
+as depicted in figure \ref{fig:conservation}.
 
@@
 both single- and multi-spanners are evolutionarily conserved.
+
+We conclude that,
+when randomly selecting SNPs, 
+for 100 SNPs found in soluble regions of multi-spanners,
+one finds 89.6 SNPs in TMHs, as shown by 
+figure \ref{fig:conservation_per_spanner}.
 
@@
 
+See table \ref{tab:ncbi_counts} for an overview of all amounts.
+
+% Label: tab:ncbi_counts
+\input{ncbi_peregrine_results/ncbi_counts.tex}
+
+Figure \ref{fig:snps_per_gene_name_ncbi} shows the distribution of the
+number of SNPs per gene name, at 2020-12-14.
+
 \begin{figure}[!htbp]
@@
 
+Figure \ref{fig:f_tmh_ncbi} shows the distribution 
+of the percentages of TMH of transmembrane proteins.
+
 \begin{figure}[!htbp]
@@
     Distribution of the percentages of TMH of transmembrane proteins.
+    Dashed vertical line: mean \% TMH in all proteins.
+    Dotted vertical line: mean \% TMH in TMPs.
   }
@@
     Histogram of the number of SNPs per trans-membrane protein.
+    Dashed vertical line: average number of SNPs per TMP
   }
@@
 We find no clear evidence of a bias.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
-  \caption{
-    Evolutionary conservation of SNPs in TMHs.
-  }
-  \label{fig:conservation}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
-  \caption{
-    Evolutionary conservation of SNPs in TMHs,
-    for either single- or multi-spanners
-  }
-  \label{fig:conservation_per_spanner}
-\end{figure}
 

### 2021-03-18 bbbq_article.tex
Removed statistics distributions

@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/ppoisbinom.png}
-  \caption{
-    Probabilities for all SNPs
-  }
-  \label{fig:ppoisbinom}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/ppoisbinom_single.png}
-  \caption{
-    Probabilities for SNPs in single-spanners
-  }
-  \label{fig:ppoisbinom_single}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/ppoisbinom_multi.png}
-  \caption{
-    Probabilities for SNPs in multi-spanners
-  }
-  \label{fig:ppoisbinom_multi}
-\end{figure}
-
 %

### 2021-03-18 bbbq_article.tex
All supmat figs have same style

@@
 
-See table \ref{tab:ncbi_counts} for an overview of all amounts.
-
-% Label: tab:ncbi_counts
-\input{ncbi_peregrine_results/ncbi_counts.tex}
+See tables \ref{tab:ncbi_counts_1} and tables \ref{tab:ncbi_counts_2}
+for an overview of all amounts.
+In table \ref{tab:ncbi_counts_1} one expects [RJCB: name these]
+numbers to add up. 
+The number of unique SNPs and unique gene names
+does not add up for MAP and TMP,
+as one SNP may work on multiple isoforms, some of which can be MAP
+where others can be TMP.
+The number of unique SNPs and unique gene names
+does not add up for TMPs in TMH and TMP in soluble regions
+as one SNP may work on multiple isoforms, some of which can be MAP
+where others can be TMP, and some SNPs fall in a TMH, where others
+are found in soluble regions
+In table \ref{tab:ncbi_counts_2} one expects [RJCB: name these]
+numbers to add up.
+
+% Label: tab:ncbi_counts_1
+\input{ncbi_peregrine_results/ncbi_counts_1.tex}
+
+% Label: tab:ncbi_counts_2
+\input{ncbi_peregrine_results/ncbi_counts_2.tex}
 

### 2021-03-18 ncbi_peregrine_results/ncbi_counts.tex
Move information from figures to table
All supmat figs have same style

--- 2021-03-11 ncbi_peregrine_results/ncbi_counts.tex
@@
 in\_sol = in soluble region of TMP.
-perc\_tmh = percentage transmembrane helix (TMH).
 }
 \centering
-\begin{tabular}[t]{l|r|r|r|r|r|r}
+\begin{tabular}[t]{l|r|r|r|r|r|r|r|r|r|r}
 \hline
-what & raw & all\_proteins & map & tmp & in\_tmh & in\_sol\\
+what & raw & all\_proteins & map & tmp & in\_tmh & in\_sol & single\_in\_tmh & single\_in\_sol & multi\_in\_tmh & multi\_in\_sol\\
 \hline
-Number of variations & 60931 & 37831 & 16623 & 21208 & 3803 & 17405\\
+Number of variations & 60931 & 37831 & 16623 & 21208 & 3803 & 17405 & 452 & 7734 & 3351 & 9671\\
 \hline
-Number of unique variations & 60544 & 37630 & 16606 & 21024 & 3789 & 17235\\
+Number of unique variations & 60544 & 37630 & 16606 & 21024 & 3789 & 17235 & 451 & 7733 & 3338 & 9502\\
 \hline
-Number of unique SNPs & NA & 9621 & 4219 & 6026 & 1140 & 4936\\
+Number of unique SNPs & NA & 9621 & 4219 & 6026 & 1140 & 4936 & 160 & 2393 & 994 & 2762\\
 \hline
-Number of unique gene names & 953 & 911 & 457 & 605 & 325 & 590\\
+Number of unique gene names & 953 & 911 & 457 & 605 & 325 & 590 & 96 & 282 & 243 & 344\\
 \hline
-Number of unique protein names & 5163 & 4780 & 2227 & 2553 & 1280 & 2467\\
+Number of unique protein names & 5163 & 4780 & 2227 & 2553 & 1280 & 2467 & 304 & 1032 & 976 & 1435\\
 \hline
-Percentage TMH & NA & 10 & 0 & 19 & 26 & 18\\
+Percentage TMH & NA & 10 & 0 & 19 & 26 & 18 & 11 & 5 & 35 & 26\\
 \hline

### 2021-03-18 ncbi_peregrine_results/ncbi_counts_1.tex
All supmat figs have same style

\begin{table}

\caption{\label{tab:ncbi_counts}Amounts. raw = all variations, including DNA variations. all\_proteins = all proteins. map = membrane associated protein. tmp = transmembrane protein. in\_tmh = in transmembrane helix of TMP. in\_sol = in soluble region of TMP. }
\centering
\begin{tabular}[t]{l|r|r|r|r|r|r}
\hline
what & raw & all\_proteins & map & tmp & in\_tmh & in\_sol\\
\hline
Number of variations & 60931 & 37831 & 16623 & 21208 & 3803 & 17405\\
\hline
Number of unique variations & 60544 & 37630 & 16606 & 21024 & 3789 & 17235\\
\hline
Number of unique SNPs & NA & 9621 & 4219 & 6026 & 1140 & 4936\\
\hline
Number of unique gene names & 953 & 911 & 457 & 605 & 325 & 590\\
\hline
Number of unique protein names & 5163 & 4780 & 2227 & 2553 & 1280 & 2467\\
\hline
Percentage TMH & NA & 10 & 0 & 19 & 26 & 18\\
\hline
\end{tabular}
\end{table}

### 2021-03-18 ncbi_peregrine_results/ncbi_counts_2.tex
All supmat figs have same style

\begin{table}

\caption{\label{tab:ncbi_counts}Amounts. single\_in\_tmh = in transmembrane helix of single-spanner. single\_in\_sol = in soluble region of single-spanner. multi\_in\_tmh = in transmembrane helix of multi-spanner. multi\_in\_sol = in soluble region of multi-spanner. }
\centering
\begin{tabular}[t]{l|r|r|r|r}
\hline
what & single\_in\_tmh & single\_in\_sol & multi\_in\_tmh & multi\_in\_sol\\
\hline
Number of variations & 452 & 7734 & 3351 & 9671\\
\hline
Number of unique variations & 451 & 7733 & 3338 & 9502\\
\hline
Number of unique SNPs & 160 & 2393 & 994 & 2762\\
\hline
Number of unique gene names & 96 & 282 & 243 & 344\\
\hline
Number of unique protein names & 304 & 1032 & 976 & 1435\\
\hline
Percentage TMH & 11 & 5 & 35 & 26\\
\hline
\end{tabular}
\end{table}

### 2021-03-18 bbbq_article.tex
Process feedback Geert

@@
 It is known that an individual's haplotype
-influences the dynamics of a pathogen (see,
-for example, \cite{eccleston2017host} for HIV) 
-[GVDB: Too vague. How?].
+influences the dynamics of a pathogen.
+For example, in the fast-mutating HIV virus, 
+HLA haplotypes that are likelier to induce a specific T-cell response
+for the Gag protein slow down the disease's 
+progression \cite{eccleston2017host}.
 The bioinformatics pipeline developed in this

### 2021-03-18 bbbq_article.tex
First decent conclusion section

@@
 at least in the human proteome.
+
+% \paragraph{Over-presentation in vivo in MHC-I is robust}
+
+We confirm that 1.4\% of epitopes that are naturally presented on MHC-I 
+can be mapped to stem from a TMH. Additionally, this 
+prediction is robust, as it holds for two TMH prediction tools, 
+as shown by table \ref{tab:elution}).
+
+% \paragraph{Over-presentation in vivo in MHC-II is robust}
+
+We predict that also MHC-II presents epitopes derived from TMHs
+in vivo, as we find that 3.9-4.0\% of sequenced epitopes
+can be uniquely traced onto a human reference 
+proteome (also shown in table \ref{tab:elution}).
+
+% \paragraph{Over-presentation in silico}
+
+Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
+are presented in vivo, we conclude that these are additionally
+presented more often than expected chance 
+alone (figures \ref{fig:1} and \ref{fig:2}).
+This in silico finding is a general pattern 
+that persists in both MHC-I and MHC-II, 
+for humans and two pathogens.
+
+% \paragraph{Evolutionary conservation}
+
+We conclude that mutations are more likely to occur
+in soluble regions of proteins, than in TMHs.
+When selecting SNPs at random, per 1000 SNPs found in a soluble domain, 
+one finds on average 918 SNPs in TMHs.
+This effect is present in TMPs with two or more TMHs, 
+but not in single-spanners.
 

### 2021-03-18 bbbq_article.tex
Moved section 'Interesting haplotype for SARS-CoV-2' to discussion
Discussion starts making sense

@@
 
-% \paragraph{Interesting haplotype for SARS-CoV-2}
-
-It is known that an individual's haplotype
-influences the dynamics of a pathogen.
-For example, in the fast-mutating HIV virus, 
-HLA haplotypes that are likelier to induce a specific T-cell response
-for the Gag protein slow down the disease's 
-progression \cite{eccleston2017host}.
-The bioinformatics pipeline developed in this
-study may be interesting to make in silico
-predictions of a SARS-CoV-2 infection.
-The most prominent example is the (MHC-I)
-haplotype HLA-A*24:02,
-which is predicted to present four times as much epitopes derived 
-from SARS-CoV-2 TMHs than expected by chance.
-This may help the immune system to detect a SARS-CoV-2 infection earlier,
-and change the infection's dynamics.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-% \paragraph{Does MHC-I present TMH-derived epitopes from pathogens as often?}
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
-
-It is important that MHC-I presents both the peptides of the 
-(healthy) cell, as well as possible pathogen-derived peptides when the
-cell is infected. As described above, MHC-I presents TMH-derived epitopes 
-from the human host more often. It is unknown if MHC-I has the same
-dedication to present epitopes that stem from TMHs derived from
-proteins produced by pathogen, either would the pathogen
-be a virus or 
-a bacterium.
+% \paragraph{General}
+
+Our understanding of the immune system and epitope prediction is 
+important to fight disease. By this paper, we suggest that epitopes
+derived from TMHs are needlessly overlooked and may inspire to investigate
+new potential targets for combating pathogens.
+
+\subsection{Elution studies}
+
+% \paragraph{TMH presentation in vivo}
+
+From elution studies and in silico predictions, 
+we suggest that TMH-derived epitopes are presented in vivo.
+The biochemical intracellular pathway to achieve this,
+however, is still unknown.
+
+% \paragraph{False positives}
+
+An additional unknown is, although our findings are robust,
+whether this is a false positive,
+due to additive methodological artifacts.
+An obvious artifact is that the elution studies used 
+will contain some unavoidable sequencing errors.
+Additionally, we assume that the epitopes naturally presented
+are derived from a cleanly translated proteome.
+Two known biological phenomena that violate this assumption
+are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
+and peptide fusion (\cite{delong2016pathogenic}),
+two processes that both result in peptide fragments
+that did not originate from genome directly.
+Also, the TMH prediction software is imperfect.
+This imperfection, however, appears to be of limited importance,
+as two different topology prediction software have similar results.
+Although those a couple of percent of epitopes that 
+can be mapped directly to the proteome,
+there is methodological bias caused by mass spectronomy,
+as hydrophobic fragments are detected less 
+by mass spectronomy. This means that the number of TMH-derived epitopes
+is underestimated.
+in vivo work in the lab will be needed to resolve this finding.
+
+\subsection{Presentation of epitopes derived from TMHs}
 
@@
 
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
-
-If presentation of TMHs on MHC-I would bring an evolutionary advantage 
-in the recognition of pathogens by the immune system, 
-it would follow that this is equally important for MHC-II, 
-especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
-response (\cite{novy2007cd4}). 
-The mechanism to detect foreign TMHs by MHC-II would be unknown, 
-similar to the discovery that MHC-I presents TMHs.
-
-% \paragraph{Does MHC-II present TMH-derived epitopes from pathogens as often?}
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
-
-It is unknown if MHC-II, like MHC-I, 
-presents TMH-derived epitopes as often, either 
-in humans,
-bacteria or viruses.
+This study predict that TMH-derived epitopes are over-presented
+in MHC-II as well. 
+The intracellular pathway of MHC-II differs from that of MHC-I,
+as it acquires epitopes from the exterior of the cell by phagocytosis.
+How TMHs derived from exterior TMPs are lysed and loaded onto MHC-II 
+is unknown. in vivo work is needed to elucidate this novel pathway.
+
+% \paragraph{Immunogenicity}
+
+In this study, we find that TMH-derived epitopes are over-presented.
+The presentation of an epitope, however, 
+is only a first step of an adaptive immune response.
+We assume that the presentation of epitopes is
+correlated with evoking a full immune response.
+TMH-derived epitopes, however, are more hydrophobic than the average
+epitope. If this evokes a stronger or weaker immune response is unknown,
+and in vivo work is needed to resolve this.
+
+\subsection{Evolutionary conservation of TMHs}
 
@@
 can be detected.
-
-% \paragraph{Use of protein data in phylogenetics}
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
-
-When using DNA sequences, one can use a skewed rate
-in non- versus synonymous mutation
-to detect the signature of selection (\cite{murrell2015gene}).
-Using AA sequences, however, has its advantages,
-as its is closer to the actual phenotype
-selection acts upon: DNA may never be transcribed to RNA,
-or its RNA may never be translated (\cite{diz2012proteomics}).
-When using a proteome in phylogenetic research,
-we know that the majority of proteins are selected to just
-maintain their function most of the time, where
-is the time spans there is selection, only a few AAs
-can actually increase the 'fitness' of the
-protein (\cite{anisimova2009investigating}).
-There, when generalizing the dynamics of mostly purifying selection (to maintain
-a protein's function) and a short duration of positive selection,
-those genes that are selected cannot be detected (\cite{yang2000statistical}).
-
-[RJCB: that was the last paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
-
-\subsection{Elution studies}
-
-% \paragraph{False positives}
-
-We assume that epitopes 
-derived from TMHs are actually presented by MHC-I and MHC-II.
-Our analysis (at subsection \ref{subsec:elution_studies}) 
-finds that 1.4\% (for MHC-I) and 4.0% (for MHC-II) 
-of the epitopes presented 
-overlap with a TMH.
-It is unsure if this a true finding or a false positive.
-One source of noise would stem from imperfect predictions
-by the topology prediction software used. 
-We have counteracted this by using two different topology prediction
-software and confirm these have similar results.
-Two biological sources of noise [GVDB: What do you mean by noise?]
-are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
-and peptide fusion \cite{delong2016pathogenic},
-two processes that both result in peptide fragments
-that did not originate from genome directly.
-Note however, that both elution studies have a
-methodological bias caused by mass spectronomy,
-as hydrophobic fragments are detected less 
-by mass spectronomy.
-
-\subsection{\% epitopes overlapping with TMHs}
 
@@
 
+% \paragraph{Interesting haplotype for SARS-CoV-2}
+
+It is known that an individual's haplotype
+influences the dynamics of a pathogen.
+For example, in the fast-mutating HIV virus, 
+HLA haplotypes that are likelier to induce a specific T-cell response
+for the Gag protein slow down the disease's 
+progression \cite{eccleston2017host}.
+The bioinformatics pipeline developed in this
+study may be interesting to make in silico
+predictions of a SARS-CoV-2 infection.
+The most prominent example is the (MHC-I)
+haplotype HLA-A*24:02,
+which is predicted to present four times as much epitopes derived 
+from SARS-CoV-2 TMHs than expected by chance.
+This may help the immune system to detect a SARS-CoV-2 infection earlier,
+and change the infection's dynamics.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-03-18 bbbq_article.tex
First step towards a proper discussion

@@
   What is the consequence of the findings? Why is it important. 
-
-  Now it is just a list of potential reasons why the results might be flawed. 
-  While this is in principle OK, 
-  it is also nitpicking as some things affect only a limited number of 
-  proteins (like selenoproteins) 
-  or you’d expect changes in the other direction (wrong prediction of TMHs for bacteria).
-
-  So offer interpretations. 
-  The second part of the introduction would fit here well. I can also help with this.
 ]
@@
 
+% \paragraph{Low number of SNPs}
+
+The NCBI dbSNP database contains millions of SNPs.
+As the NCBI API constrains its users to three calls per second
+to assure fair use, we had to limit the extent of our analysis.
+We assume we have taken a representative sample, but only
+an analysis using the full NCBI database can confirm this.
+
 % \paragraph{Selection undetectable in whole proteome}
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
 
@@
 is not tailored to detect these sites, 
- as hinted by upon by the aforementioned influenza study.
+as hinted by upon by a extensive study on influenza (\cite{han2019individual}),
+where it was found that the selection pressure
+exerted by the immune system was only limited.
 
 % \paragraph{Selection may be detectable in TMHs}
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
 
@@
 can be detected.
-
-\subsection{Evolutionary conservation}
-
-[GVDB: 
-  This might be relevant, but you need to explain why it is relevant. 
-  Thus compare your findings with previous efforts 
-  to look at conservation of TMHs
-]
-% \paragraph{synonymous mutations}
-%
-In our evolutionary experiment, 
-we removed variations that were synonymous mutations (i.e.
-resulted in the same amino acid, from a different genetic code) 
-from our analysis.
-There is evidence, however, that these synonymous mutations
-do have an effect and may even be evolutionary selected 
-for (\cite{hunt2009silent}).
-As the possible effect of synonymous mutations is ignored by our
-topology prediction software, we do so as well.
-
-% \paragraph{Low number of SNPs}
-
-The NCBI dbSNP database contains millions of SNPs.
-As the NCBI API constrains its users to three calls per second,
-we had to limited the extent of our analysis.
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
-We concluded that the
-epitopes that MHC-I presents are [as/not as] likely 
-to be derived from TMH within either a human host and its bacterial pathogen.
-Because a bacterium does not infect a cell, thus its peptides
-will not be presented by MHC-I, this result is [unexpected/expected]
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
+We speculate that the human immune system is more attentive 
+towards TMHs, as it is a general structure, that is evolutionarily conserved.
+
 We aimed our evolutionary experiment at TMHs, because these can
@@
 are common structures and are present in all pathogens.
-
-[RJCB: this paragraph I ruthlessly moved
-from introduction to here, as was suggested by GVDB]
-The human immune system and human pathogen are in an evolutionary
-arms race: our immune systems is selected for the detection
-of pathogens, whereas pathogens are selected to avoid detection.
-From a pathogen's point of view, however, this struggle 
-is of only minor importance:
-in seasonal influenza, for example, the selection pressure
-exerted by the immune system was only limited (\cite{han2019individual}).
-
-% \paragraph{Interesting haplotype for SARS-CoV-2}
-
-It is known that an individual's haplotype
-influences the dynamics of a pathogen.
-For example, in the fast-mutating HIV virus, 
-HLA haplotypes that are likelier to induce a specific T-cell response
-for the Gag protein slow down the disease's 
-progression \cite{eccleston2017host}.
-The bioinformatics pipeline developed in this
-study may be interesting to make in silico
-predictions of a SARS-CoV-2 infection.
-The most prominent example is the (MHC-I)
-haplotype HLA-A*24:02,
-which is predicted to present four times as much epitopes derived 
-from SARS-CoV-2 TMHs than expected by chance.
-This may help the immune system to detect a SARS-CoV-2 infection earlier,
-and change the infection's dynamics.
+We speculate that the human immune system may also be more attentive 
+towards beta turns for this reason.
+
+[GVDB: 
+  Also compare your findings with previous efforts 
+  to look at conservation of TMHs
+]
 
@@
 regions around these, hence we expect this not to affect our results.
+
+% \paragraph{synonymous mutations}
+%
+In our evolutionary experiment, 
+we removed variations that were synonymous mutations (i.e.
+resulted in the same amino acid, from a different genetic code) 
+from our analysis.
+There is evidence, however, that these synonymous mutations
+do have an effect and may even be evolutionary selected 
+for (\cite{hunt2009silent}).
+As the possible effect of synonymous mutations is ignored by our
+topology prediction software, we do so as well.
 

### 2021-03-22 bbbq_article.tex
No TiKZ, works on Overleaf
Updates from Overleaf
Re-read intro and fixed refs

@@
 
+\frans{Hier kun je je commentaar kwijt}
 Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
@@
 over-presented is unclear
-In this study, we show the TMHs are evolutionarily more conserved, 
+In this study, we show that TMHs are evolutionarily more conserved, 
 as relatively less single nucleotide polymorphisms (SNPs) 
@@
 [FB: is that really so?]
-[RB: yes, according to the removed reference to Lund et al., 2004, 'Definition of supertypes for {HLA} molecules using clustering of specificity matrices'] 
+[RB: yes, according to the reference that is absent in the docx version to Lund et al., 2004, 'Definition of supertypes for {HLA} molecules using clustering of specificity matrices'] 
 The HLA region of humans is highly polymorphic, with hundreds 
 to thousands of different alleles, 
-and each different HLA gene is called 
+and each different HLA allele is called 
 an MHC haplotype (\cite{marsh2010nomenclature}).
@@
 the number of pathogenic peptides that can be detected is increased,
-as a wide variety of MHCs: 
-6 MHC-I and 6-8 MHC-II molecules are expressed per individual
+as a wide variety of MHCs (6 MHC-I and 6-8 MHC-II molecules) 
+are expressed per individual
 each presenting their own subset of epitopes.
@@
 is relatively hydrophobic, 
-and therefore hydrophobic peptides have a higher affinity 
-than hydrophilic peptides.
+and therefore hydrophobic peptides have a higher affinity to bind
+than hydrophilic ones.
 
@@
 We hypothesized is that the presentation of 
-hydrophobic/TMH residues is evolutionary selected for, 
+TMH residues is evolutionary selected for, 
 because TMHs are less prone to undergo escape mutation. 
@@
 [FB: Analyzing the occurrence of single nucleotide polymorphisms 
-that were predicted to be inside ...]
-
+that were predicted to be inside ...][RJCB: Uh, you want to add another
+paragraph here?]
 Therefore, the TMHs of pathogens 
@@
 
-[GVDB:
- Add one short paragraph, 
- where you reiterate the most important findings:
- In this study, we show that ...(and than basically 
- the abstract in slightly different words and slightly 
- more detail). End with a statement of why it is 
- important (vaccine development based on TMHs?)...
-]
+From an elution study, we find that epitopes derived 
+from TMHs are presented on naturally occurring MHC-II,
+by finding the origin of these epitopes on a reference proteome,
+of which we predicted all proteins' topologies.
+From in silico predictions, we conclude that TMH-derived
+epitopes are presented more often than expected by chance,
+in a human, viral and bacterial proteome, for both MHC-I and II,
+as these epitopes have a higher affinity with the presenter molecule.
+From the generality of this pattern, we speculated that TMHs
+are likely to be evolutionarily conserved. From an analysis using human
+SNPs we found that these random point mutations are indeed less likely
+to occur within TMHs. These findings encourage us to investigate the
+overlooked importance of TMHs in the immune response and vaccine development.
 
@@
 % tab:elution
-\input{../bbbq_article_issue_157/results/results.tex}
+\input{bbbq_article_issue_157/results.tex}
 

### 2021-03-22 bbbq_article.tex
Describe presented TMH epitope counts table better

@@
 
-% Percentage of spots and spots that overlap with a TMH
-\input{bbbq_1_smart_results/table_f_tmh_2.latex}
-
 % \paragraph{IC50 prediction software}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Percentage of spots and spots that overlap with a TMH}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Table \ref{tab:f_tmh} shows an overview of the findings,
+where a target specifies the source of the proteome,
+where \verb;covid; denotes SARS-CoV-2 and \verb;covid; denotes
+Mycobacterium tuberculosis. \verb;mhc_class; denotes the MHC
+class, \verb;n_spots; the number of possible 9-mers (for MHC-I) 
+or 14-mers (for MHC-II) possible. \verb;n_spots_tmh; the
+number of epitopes that overlapped with a TMH that were binders. 
+\verb;f_tmh; the percentage of peptides that had at least 1 residue
+overlapping with a TMH.
+
+% Label: tab:f_tmh
+\input{bbbq_1_smart_results/table_f_tmh_2.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Minor methods}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+These are details that are removed from the 'Methods' section.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+These are details that are removed from the 'Discussion' section.
 

### 2021-03-22 bbbq_article.tex
Fix supmat
Remove duplicate label
Updates from Overleaf

@@
 The earlier study defines a peptide a binder (for a haplotype), 
-if \emp{within the peptide} in which it is found, 
+if \emph{within the peptide} in which it is found, 
 is within the peptides with the 2\% lowest IC50 values.
@@
 In this study, a peptide is defined a binder (for a haplotype), 
-if within a \emp{proteome} in which it is found, 
+if within a \emph{proteome} in which it is found, 
 is within the peptides with the 2\% lowest IC50 values,
 where our previous study used the lowest 2\% IC50 values
-within each \emp{protein}.
+within each \emph{protein}.
 Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
@@
 statistics for only single-spanners and multi-spanners respectively.
-These probabilities are also displayed visually in 
-figures \ref{fig:ppoisbinom} (all proteins), \ref{fig:ppoisbinom_single} (single-spanners)
-and \ref{fig:ppoisbinom_multi} (multi-spanners).
 

### 2021-03-22 bbbq_article.tex
Fix ref
Updates from Overleaf

@@
 
-\frans{Hier kun je je commentaar kwijt}
+\frans{Show off to Frans}
 Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
@@
 in predicted TMPs.
-See table \ref{tab:ncbi_counts} for an overview of all amounts.
+See tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
+for an overview of all amounts.
 
@@
   }
-  \label{fig:n_tmhs_per_protein}
+  \label{fig:fig_n_snps_per_tmp}
 \end{figure}
@@
   }
-  \label{fig:n_tmhs_per_protein}
+  \label{fig:fig_n_tmhs_per_protein}
 \end{figure}

### 2021-03-23 bbbq_article.tex
Made float smaller to remove warning
Feedback @fransbianchi

@@
 % \paragraph{Data sets for TMH epitopes}
-
-We used a human, viral and bacterial reference proteome to 
-determine the percentages of epitopes overlapping
-with TMHs.
-We used a human reference proteome
-with UniProt ID UP000005640\_9606, which
-is a representative set of the full proteome
-and used is our previous study (\cite{bianchi2017}).
-We used SARS-CoV-2 as our viral reference proteome
-with UniProt ID UP000464024 and
-Mycobacterium tuberculosis (MTb) 
-with UniProt ID UP000001584.
-We chose these pathogens based on their impact on the human population.
+A human, viral and bacterial reference proteome was used to determine the percentages of epitopes overlapping with TMHs.
+The human reference proteome has UniProt ID UP000005640_9606, which is a representative set of the full proteome and also used in our previous study (\cite{bianchi2017}).
+SARS-CoV-2 with reference proteome UniProt ID UP000464024 was used as a viral proteome and Mycobacterium tuberculosis (MTb) with UniProt ID UP000001584 was used .
+These pathogens were chosen based on their impact on the human population(references).
 
@@
 
-To determine if epitopes derived from TMHs are presented at all,
-we started from epitopes identified in elution studies
+To determine if epitopes derived from TMHs are presented we used epitopes identified in \frans{Welke studies is dit dan van SARS-COV2 en MTB ?}
 for MHC-I (\cite{schellens2015comprehensive}) 
-and MHC-II (\cite{bergseng2015different}).
-For each of these peptides, the human representative reference proteome
-was searched for its origin.
-We kept only the epitopes which were uniquely present
-in the proteome.
-We predicted the topology of the proteome
-and counted how often the epitopes overlapped
-with a predicted TMH.
-In this analysis, both TMHMM and 
-PureseqTM (see section 'Prediction software used') 
-were used to predict the topology.
+and MHC-II (\cite{bergseng2015different})peptide elution studies.
+For each of these peptides, the human representative reference proteome was searched for its origin\frans{wat bedoel je?}.
+We filtered for epitopes that were uniquely present in the proteome and correlated these for overlap with predicted transmembrane regions in the proteome.
+In this analysis, both TMHMM and PureseqTM (see section ’Prediction software used’) were used to predict the transmembrane topology. 
 The full analysis can be found
@@
 
-To determine the percentages of MHC-I and MHC-II epitopes overlapping
-with TMHs, we used mostly the same analysis as described in \cite{bianchi2017}.
+To determine the percentages of MHC-I and MHC-II epitopes overlapping with TMHs, we applied a similar analysis strategy as described in  \cite{bianchi2017}.
 To summarize: from a proteome all possible 9-mers are derived. For each
@@
 This study differs in some aspects from our previous work (\cite{bianchi2017}), 
-described below in more detail.
-The main differences are our definition of what a so-called binder is, 
-the inclusion of both MHC-I and MHC-II haplotypes, 
-the prediction software used, 
-and the significance level to determine if TMH-derived peptides will bind.
-These deviations are either a refinement of our previous approach or
-a pragmatic choice made due to the extension of the original experiment.
-Additionally, instead of only using a human proteome, this study
-also includes a viral and bacterial proteome.
-
 The definition of a binder differs from \cite{bianchi2017}:
-in our current study a peptide was called a binder if, within a certain haplotype, 
+in the current study a peptide is called a binder if, within a certain haplotype, 
 any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
@@
 The 13 MHC-I haplotypes used in this study are the same as 
-we used in our previous study (\cite{bianchi2017}).
+used in our previous study (\cite{bianchi2017}).
 The MHC-II haplotypes used additionally were selected 
@@
 used, as these are the most common MHC-II epitope size,
-as found in the elution study (\cite{bergseng2015different}).
+as found in the elution study (\cite{bergseng2015different}).\frans{what percentage is 9 mer or 14 mer?}
 
@@
 To detect the evolutionary conservation of TMHs, 
-we collected human mutations and 
-tallied their predicted location.
-To be more precise, we collected single nucleotide
+single nucleotide
 polymorphisms (SNPs) within the human population
-that resulted in the substitution of one amino acid.
-We then predicted that protein's topology and tallied if
-the SNP occurred within a TMH or not, 
-after which we used statistics to determine if there are
-more, less or an equal amount of SNPs in TMHs
+that resulted in the substitution of one amino acid were collected from ??? and tallied if
+the SNP occurred within a TMH or not. Using statistics is was determined if there are more, less or an equal amount of SNPs in TMHs
 as a measure of evolutionary conservation.
@@
 NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used,
-which are 'gene', to find the gene names of membrane proteins, 
+which are 'gene',\frans{snap ik niet which are gene?} to find the gene names of membrane proteins, 
 'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
@@
 
-To retrieve the data from these databases the
+\frans{ dit is relevant waarom: ? To retrieve the data from these databases the
 \verb;rentrez; R package (\cite{rentrez}) was used
@@
 We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
-bigger queries of multiple (and delayed) API calls.
+bigger queries of multiple (and delayed) API calls.}
 
@@
 resulting in 60,683 variations at the protein level.
-A variation needs not always be a SNP,
-as dbSNP also catalogs other DNA alterations, such as, among others, insertions,
-deletions and tandem repeats.
-We select only the variations that result in a SNP for
-a single amino acid substitution, resulting in 38,882 SNPs.
+Only variations that result in a SNP for
+a single amino acid substitution were analysed, resulting in 38,882 SNPs.
 
@@
 %
-We selected to analyze SNPs with the lowest SNP ID first.
-This means we have a bias for picking SNPs with
-an earlier discovery date.
-We expect this bias to be biologically irrelevant.
-To verify this expectation, we determined the
-relative position of each SNP in a protein,
-as shown in figure \ref{fig:snp_rel_pos}.
+SNPs were picked based on ID number, which is linked to their discovery date.
+To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
 
@@
 [RJCB: I suggest to move this section to the Supplementary Materials]
-
+\frans{lijkt mij een goed idee}
 \begin{table}[]
@@
 \begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
+  \includegraphics[height=0.8\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
   \caption{

### 2021-03-23 bbbq_article.tex
Processed feedback 'Measuring TMH epitopes', thanks @fransbianchi

@@
 
-% \paragraph{Data sets for TMH epitopes}
-A human, viral and bacterial reference proteome was used to determine the percentages of epitopes overlapping with TMHs.
-The human reference proteome has UniProt ID UP000005640_9606, which is a representative set of the full proteome and also used in our previous study (\cite{bianchi2017}).
-SARS-CoV-2 with reference proteome UniProt ID UP000464024 was used as a viral proteome and Mycobacterium tuberculosis (MTb) with UniProt ID UP000001584 was used .
-These pathogens were chosen based on their impact on the human population(references).
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-To determine if epitopes derived from TMHs are presented we used epitopes identified in \frans{Welke studies is dit dan van SARS-COV2 en MTB ?}
+To determine if epitopes derived from TMHs are presented 
+we used two elution studies, which both have sequenced
+epitopes found in naturally occurring human immune cells,
 for MHC-I (\cite{schellens2015comprehensive}) 
-and MHC-II (\cite{bergseng2015different})peptide elution studies.
-For each of these peptides, the human representative reference proteome was searched for its origin\frans{wat bedoel je?}.
-We filtered for epitopes that were uniquely present in the proteome and correlated these for overlap with predicted transmembrane regions in the proteome.
-In this analysis, both TMHMM and PureseqTM (see section ’Prediction software used’) were used to predict the transmembrane topology. 
+and MHC-II (\cite{bergseng2015different}).
+For each of the sequenced epitopes,
+we found its possible location(s) in 
+a human representative reference proteome,
+with UniProt ID UP000005640_9606.
+For the epitopes that were found exactly once in the proteome,
+we predicted the topology of the proteins in which the epitopes are found.
+We used both TMHMM and PureseqTM to do these predictions. 
+From this topology, we could determine if the epitope
+overlapped with a transmembrane regions.
+
 The full analysis can be found
@@
 
-To determine the percentages of MHC-I and MHC-II epitopes overlapping with TMHs, we applied a similar analysis strategy as described in  \cite{bianchi2017}.
-To summarize: from a proteome all possible 9-mers are derived. For each
-of these peptides, it was determined if it was part of a 
-TMH (at least overlapping with 1 residue), 
-and if the peptide bound to an MHC-I haplotype.
-For MHC-II, 14-mers were used, as these are the most frequently occurring
+To determine how frequently epitopes overlapping with TMHs are presented,
+we applied a similar analysis strategy as described in \cite{bianchi2017}.
+To investigate the generality of this, we applied this analysis
+for several haplotypes of both MHC-I and MHC-II, 
+as well as for a human, viral and bacterial proteome.
+To summarize, for each proteome, 
+all possible 9-mers (for MHC-I) or 14-mers (MHC-II) are derived. 
+For each of these peptides, we determined if it overlapped with a 
+TMH and if it was a binder for each haplotype.
+
+For MHC-I, 9-mers were used, as this is the length used in our earlier
+study (\cite{bianchi2017}). For MHC-II, 14-mers were used, 
+as these are the most frequently occurring
 epitope length (\cite{bergseng2015different}).
+We used a human (UniProt ID UP000005640_9606), 
+viral (SARS-CoV-2, UniProt ID UP000464024) 
+and bacterial (Mycobacterium tuberculosis (MTb), UniProt ID UP000001584) 
+reference proteome. The pathogens were chosen based 
+on their impact on the human 
+population \frans{Add references here}.
+We used TMHMM  \richel{ref} to predict the topology 
+of the proteins within the proteome.
+To predict the affinity of an epitope to a certain MHC molecule,
+we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
+and \verb;MHCnuggets; \richel{ref} for MHC-II.
+The 13 MHC-I haplotypes used in this study are the same as 
+used in our previous study (\cite{bianchi2017}).
+The MHC-II haplotypes used additionally were selected 
+to occur with a phenotypic frequency of at least 14\% in
+the human population (\cite{greenbaum2011functional}),
+resulting in 21 haplotypes.
 
@@
 that give rise to no or only very few MHC epitopes
-
-The 13 MHC-I haplotypes used in this study are the same as 
-used in our previous study (\cite{bianchi2017}).
-The MHC-II haplotypes used additionally were selected 
-to occur with a phenotypic frequency of at least 14\% in
-the human population (\cite{greenbaum2011functional}),
-resulting in 21 haplotypes.
-When using an MHC-II haplotype, instead of using 9-mers, 14-mers were
-used, as these are the most common MHC-II epitope size,
-as found in the elution study (\cite{bergseng2015different}).\frans{what percentage is 9 mer or 14 mer?}
 
@@
 where the change occurred.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Prediction software used}
-\label{subsec:prediction_software_used}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-[RJCB: I suggest to move this section to the Supplementary Materials]
-\frans{lijkt mij een goed idee}
-\begin{table}[]
-  \begin{tabular}{llll}
-    Goal & Tool & Reference \\ 
-    \hline
-    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
-    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
-    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
-    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
-    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
-    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
-    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
-    Combine all                       & bbbq                      & \cite{bbbq}
-  \end{tabular}
-  \caption{
-    Overview of all software used in this research.
-  }
-  \label{table:software_used}
-\end{table}
-
-For this research, the scientific literature was explored 
-to identify the most recent free and open source (FOSS) prediction software.
-This was done by searching for papers that (1) reference older
-prediction software, and (2) present a novel method to make predictions.
-As a starting point, a review paper was used.
-For all software needed, we found no studies that compared contemporary tools 
-in their prediction quality.
-
-% \paragraph{TMH prediction}
-
-There are multiple computational tools developed to predict which
-parts of a protein forms a TMH.
-In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
-of which TMHMM (\cite{krogh2001predicting}) turned out to be the best, 
-as is used in the previous study (\cite{bianchi2017}).
-However, TMHMM has a restrictive software license and is nearly two
-decades old.
-Therefore, PureseqTM (\cite{wang2019efficient}),
-was also used in this study, which has been more recently developed
-and has a free software license.
-
-% \paragraph{MHC-I epitope prediction}
-
-For MHC-I, there are multiple computational tools developed 
-to predict epitopes. 
-According to \cite{lundegaard2011prediction}, at that time,
-NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
-We used the same tool as used in our earlier study, \verb;epitope-prediction; (\cite{bianchi2017}),
-
-% \paragraph{MHC-II epitope prediction}
-
-Also for MHC-II, there are multiple computational tools developed 
-to predict epitopes,
-such as using a trained neural network (\cite{nielsen2003reliable})
-or a Gibbs sampling approach (\cite{nielsen2004improved}).
-According to \cite{lundegaard2011prediction}, in 2011,
-from a set of multiple tools, 
-NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
-gave rise to the most accurate predictions.
-The most recent and promising FOSS tool available now appears
-to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
-and MHC-II prediction, which we used for MHC-II predictions.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Prediction software used}
+\label{subsec:prediction_software_used}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+[RJCB: I suggest to move this section to the Supplementary Materials]
+\frans{lijkt mij een goed idee}
+\begin{table}[]
+  \begin{tabular}{llll}
+    Goal & Tool & Reference \\ 
+    \hline
+    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
+    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
+    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
+    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
+    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
+    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
+    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
+    Combine all                       & bbbq                      & \cite{bbbq}
+  \end{tabular}
+  \caption{
+    Overview of all software used in this research.
+  }
+  \label{table:software_used}
+\end{table}
+
+For this research, the scientific literature was explored 
+to identify the most recent free and open source (FOSS) prediction software.
+This was done by searching for papers that (1) reference older
+prediction software, and (2) present a novel method to make predictions.
+As a starting point, a review paper was used.
+For all software needed, we found no studies that compared contemporary tools 
+in their prediction quality.
+
+% \paragraph{TMH prediction}
+
+There are multiple computational tools developed to predict which
+parts of a protein forms a TMH.
+In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
+of which TMHMM (\cite{krogh2001predicting}) turned out to be the best, 
+as is used in the previous study (\cite{bianchi2017}).
+However, TMHMM has a restrictive software license and is nearly two
+decades old.
+Therefore, PureseqTM (\cite{wang2019efficient}),
+was also used in this study, which has been more recently developed
+and has a free software license.
+
+% \paragraph{MHC-I epitope prediction}
+
+For MHC-I, there are multiple computational tools developed 
+to predict epitopes. 
+According to \cite{lundegaard2011prediction}, at that time,
+NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
+We used the same tool as used in our earlier study, \verb;epitope-prediction; (\cite{bianchi2017}),
+
+% \paragraph{MHC-II epitope prediction}
+
+Also for MHC-II, there are multiple computational tools developed 
+to predict epitopes,
+such as using a trained neural network (\cite{nielsen2003reliable})
+or a Gibbs sampling approach (\cite{nielsen2004improved}).
+According to \cite{lundegaard2011prediction}, in 2011,
+from a set of multiple tools, 
+NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
+gave rise to the most accurate predictions.
+The most recent and promising FOSS tool available now appears
+to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
+and MHC-II prediction, which we used for MHC-II predictions.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Percentage of spots and spots that overlap with a TMH}

### 2021-03-23 bbbq_article.tex
Fix layout

@@
 a human representative reference proteome,
-with UniProt ID UP000005640_9606.
+with UniProt ID UP000005640\_9606.
 For the epitopes that were found exactly once in the proteome,
@@
 epitope length (\cite{bergseng2015different}).
-We used a human (UniProt ID UP000005640_9606), 
+We used a human (UniProt ID UP000005640\_9606), 
 viral (SARS-CoV-2, UniProt ID UP000464024) 
@@
 
-\frans{ dit is relevant waarom: ? To retrieve the data from these databases the
+\frans{ dit is relevant waarom: ?} To retrieve the data from these databases the
 \verb;rentrez; R package (\cite{rentrez}) was used
@@
 bigger queries of multiple (and delayed) API calls.}
+
+\richel{Processed all of Frans feedback here}
 

### 2021-03-23 bbbq_article.tex
Process feedback Frans. Thank @fransbianchi

@@
 on their impact on the human 
-population \frans{Add references here}.
+population \richel{ref}.
 We used TMHMM  \richel{ref} to predict the topology 
@@
 To detect the evolutionary conservation of TMHs,
-single nucleotide
-polymorphisms (SNPs) within the human population
-that resulted in the substitution of one amino acid were collected from ??? and tallied if
-the SNP occurred within a TMH or not. Using statistics is was determined if there are more, less or an equal amount of SNPs in TMHs
-as a measure of evolutionary conservation.
-This workflow is discussed in more detail below.
+we collected human single nucleotide polymorphisms (SNPs)
+that resulted in a single amino the substitution,
+and tallied if this substitution occurred within a TMH or not.
+From that tally, we determined how likely the number of SNPs
+found in TMHs is by chance.
 
@@
 As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used,
-which are 'gene',\frans{snap ik niet which are gene?} to find the gene names of membrane proteins, 
-'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
-and 'protein', to obtain the sequence of proteins that SNPs act upon.
-The 'dbSNP' contains 650 million 
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used.
+Pivotal was the 'dbSNP' (\cite{sherry2001dbsnp}) database,
+which contains 650 million 
 catalogued non-redundant humane variations (called RefSNPs,
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
-
-% \paragraph{Data retrieval}
-
-\frans{ dit is relevant waarom: ?} To retrieve the data from these databases the
-\verb;rentrez; R package (\cite{rentrez}) was used
-that calls the NCBI website's API. To provide for a 
-stable user experience for all users, 
-this API limits the user to 3 calls per second.
-Additionally, the API splits the result of a bigger
-query into multiple pages, each of which needs one API call.
-We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
-bigger queries of multiple (and delayed) API calls.}
-
-\richel{Processed all of Frans feedback here}
+Additionally, the aptly-named 'gene' (for gene names)
+and 'protein' (for proteins sequences) database were used.
 
@@
 Only variations that result in a SNP for
-a single amino acid substitution were analysed, resulting in 38,882 SNPs.
+a single amino acid substitution were analyzed, resulting in 38,882 SNPs.
 
@@
 SNPs were picked based on ID number, which is linked to their discovery date.
-To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
+To verify that ID number is unrelated to SNP positions the relative positions of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
 
@@
 
-[RJCB: I suggest to move this section to the Supplementary Materials]
-\frans{lijkt mij een goed idee}
 \begin{table}[]
@@
 and MHC-II prediction, which we used for MHC-II predictions.
+
+% \paragraph{NCBI data retrieval software}
+
+To retrieve the data from the NCBI databases the
+\verb;rentrez; R package (\cite{rentrez}) was used
+that calls the NCBI website's API. To provide for a 
+stable user experience for all users, 
+this API limits the user to 3 calls per second.
+Additionally, the API splits the result of a bigger
+query into multiple pages, each of which needs one API call.
+We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
+bigger queries of multiple (and delayed) API calls.}
 

### 2021-03-23 bbbq_article.tex
Process feedback Frans. Thank @fransbianchi
EOD

--- 2021-03-23 bbbq_article.tex
@@
 
-We aimed our evolutionary experiment at TMHs, because these can
-be predicted well from a protein structure,
-are common structures and are present in all pathogens. 
-We could have done the same experiment on beta-turn,
-as also these can be predicted well (\cite{petersen2010netturnp}),
-are common structures and are present in all pathogens.
-We speculate that the human immune system may also be more attentive 
-towards beta turns for this reason.
-
-[GVDB: 
-  Also compare your findings with previous efforts 
-  to look at conservation of TMHs
-]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Acknowledgments}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We thank the Center for Information Technology of the University 
+We aimed our evolutionary expy of the University 
 of Groningen for its support and for providing access to the Peregrine 
@@
 and MHC-II prediction, which we used for MHC-II predictions.
+
+% \paragraph{NCBI data retrieval software}
+
+To retrieve the data from the NCBI databases the
+\verb;rentrez; R package (\cite{rentrez}) was used
+that calls the NCBI website's API. To provide for a 
+stable user experience for all users, 
+this API limits the user to 3 calls per second.
+Additionally, the API splits the result of a bigger
+query into multiple pages, each of which needs one API call.
+We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
+bigger queries of multiple (and delayed) API calls.
 

### 2021-03-24 bbbq_article.tex
Feedback Frans, thanks @fransbianchi

--- 2021-03-23 bbbq_article.tex
@@
 using PureseqTM.
-From the topology and the known location of the SNP, 
-we score the location (i.e. TMH or soluble protein region) 
-where the change occurred.
+Using the protein topology SNPs were scored to be in or outside trans membrane domains.
 
@@
 [RJCB: I suggest to move this section to the Supplementary Materials]
+\frans{I agree move to supp}
 
@@
 as done in this study.
-'tmhmm' has been submitted to and is accepted by CRAN.
+'tmhmm' has been submitted to and is accepted by CRAN. \frans{wat is CRAN}
 
@@
 To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
-we reanalyzed published peptide elution studies.
+peptide elution studies were reanalyzed.
 Table \ref{tab:elution} shows the percentage of epitopes derived 
@@
 From these findings, it is robustly predicted that
-epitopes derived from TMHs are presented by B cells [For Frans: please check
-that the article by Bergseng, 2015, 'Different binding motifs of the celiac 
-disease-associated HLA molecules DQ2. 5, DQ2. 2, and DQ7. 5 revealed by 
-relative quantitative proteomics of endogenous peptide repertoires'
-is appropriate. As far as I can see, it is, but I may overlook something. Richel].
+epitopes derived from TMHs can and are presented on both MHC-I and MHC-II.
 
@@
 We recently showed that peptides derived from TMHs of human proteins 
-were predicted to be over-presented in MHC-I. 
+were predicted to be over-presented on HLA-A and HLA-B. 
 Here, we wondered if this finding would be general 
-and if peptides stemming from TMHs of microbial pathogens 
-would also be over-presented.
-Figure \ref{fig:1} shows the percentages of MHC-I epitopes 
-predicted to be overlapping 
-with TMHs for our human, viral and bacterial proteome.
+and if peptides stemming from predicted TMHs of microbial pathogens 
+would also be over-presented on HLA-A and HLA-B.We therefor analysed all superhaplotypes of the HLA-A and HLA-B class see Figure \ref{fig:1} and found that similar as for the human proteome the percentages of MHC-I epitopes predicted to be overlapping with TMHs for our human, viral and bacterial proteome was higher as expected by TMH abundance.
 See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
-for the exact TMH and epitope counts.
-We found that also for pathogen epitopes derived from TMHs are
-over-presented. 
-The extent of the over-presentation is similar between the haplotypes:
-if a haplotype if prone to present TMH-derived epitopes from
-human proteins, it is also prone to present such epitopes of
-bacterial and viral proteins.
+The extent of the over-presentation between the various different HLA supertypes is similar for the probed proteomes, which confirms that the hydrophopbicity of the MHC binding cleft is determinant.
 
@@
 
-Where TMH-derived epitopes are over-presented by MHC-I,
-we wondered if this finding would also generalize to MHC-II.
+We next wondered if this over-representation would also generalize to MHC-II. \frans{ you say MHC-II but this is very general I think you have to explain the choice for these haplotypes.}
+Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with ....
+variation in peptide binding repertoire, the MHC-II peptide binding groove is determined by both the alpha and beta chain.
+For HLA DQ both the alpha and beta chain is highly polymorphic  .....
+\frans{why is HLA-DP not predicted?}
+We took the alleles that represent percentage of the occurrence int he overall human population according to ... et al. 
 Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
@@
 We show that MHC-II has the same pattern of over-presentation
-as MHC-I, yet this over-presentation is more extreme.
-Also for MHC-II, we observe that the extent the which TMH-derived
+as MHC-I, yet this over-presentation is more pronounced. \frans{we should explain this does it like shown for mhc-I also linearly correlates with the hydrophopbicity of the MHC-II binding celft?}
+Also for MHC-II, we observe that the extent to which TMH-derived
 epitopes are presented is similar between the haplotypes, for
@@
 
-In the final set of experiments, 
-we addressed the question whether TMHs 
-would be more evolutionary conserved than soluble proteins regions 
-by comparing the occurrences of SNPs in the genome coding 
-for TMHs and soluble protein regions within membrane proteins.
-We obtained 1,129 [For Frans: I would enjoy to write 'approx 1k' here, would that
-be acceptable? I do refer to a table with the exact values
-later on anyways. Richel] gene names associated with the phrase 'membrane protein',
+In general the occurrence and expression levels of membrane proteins are low, therefor we addressed the question whether there is an evolutionary advantage for presenting TMHs.
+To this end we looked for conservation the conservation in TMH vs soluble regions by comparing the occurrences of SNPs in the genome coding for TMHs versus soluble protein regions within membrane proteins.
+We obtained >1000 gene names associated with the phrase 'membrane protein',
 which are genes for both membrane-associated proteins (MAPs, no TMH) and 
@@
 This analysis gives the first hint that SNPs 
-occur more often in soluble domains than in TMHs.
-
+occur more often in soluble domains than in TMHs.\frans{waar vind ik dan nu het verschil tussen met en zonder MAPs je zou hopen dat er geen verschil is toch omdat het aandeel TMH in de analyse met MAPS ook naar beneden gaat}
 To statistically test whether SNPs occur 
@@
 
-The actually number of SNPs found in TMHs
+The actual number of SNPs found in TMHs
 was $n_{SNP\_TMH} = 3,903$, which is lower than the expectation
@@
 for 100 SNPs found in soluble regions of multi-spanners,
-one finds 89.6 SNPs in TMHs, as shown by 
+one finds 89.6 SNPs in TMHs, as shown by \frans{it says 91,1 in fig 4}
 figure \ref{fig:conservation_per_spanner}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
+\section{Conclusion}\frans{why comes this section before discussion?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 are presented in vivo, we conclude that these are additionally
-presented more often than expected chance 
+presented more often than expected by chance 
 alone (figures \ref{fig:1} and \ref{fig:2}).
@@
 that persists in both MHC-I and MHC-II, 
-for humans and two pathogens.
+for humans, bacteria and viruses.
 
@@
 one finds on average 918 SNPs in TMHs.
-This effect is present in TMPs with two or more TMHs, 
-but not in single-spanners.
+This effect is more pronounced in multi spanners likely as these TMHs more often serve additional functions, besides spanning the lipid bilayer.
 
@@
 new potential targets for combating pathogens.
-
+\frans{ik zie hier verder geen argumentatie voor deze statement}
 \subsection{Elution studies}
@@
 From elution studies and in silico predictions, 
-we suggest that TMH-derived epitopes are presented in vivo.
-The biochemical intracellular pathway to achieve this,
-however, is still unknown.
+we suggest that TMH-derived epitopes are presented and cross-presented in vivo.
+How integral membrane proteins are presented and cross-presented on MHC complexes is however unclear.
+Most prominently because the fundamental principles of trans locating integral membrane proteins from there hydrophobic environment into the vacuole lumen and subsequent processing prior to loading on MHC complexes is unresolved.
 
@@
 This imperfection, however, appears to be of limited importance,
-as two different topology prediction software have similar results.
+as two different topology prediction software algorithms gave similar results.
 Although those a couple of percent of epitopes that 
@@
 % \paragraph{MHC-II is expected to present TMHs}
+\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevnet zijn de haplotypen die veel TMH presenteren?}
+
+The over representation of TMH on MHC-II haplotypes such as HLADQA!*0101/DQB*0501 is more pronounced as for MHC-I with a third of the predicted binders from the human and MtB proteome stem from TMH  human reference proteome and MtB reference proteome to be presented on MHC-II.
+\frans{ik zoek nog naar associaties}
 
@@
 is not tailored to detect these sites, 
-as hinted by upon by a extensive study on influenza (\cite{han2019individual}),
+as hinted upon by an extensive study on influenza (\cite{han2019individual}),
 where it was found that the selection pressure
-exerted by the immune system was only limited.
+exerted by the immune system was a non pro founding factor.
 
@@
 
-TMHs, on the other hand, also have their function constraints, 
-yet can occur multiple time a pathogen's proteome.
-One can safely assume a pathogen's proteome contains multiple TMHs.
-Therefore, it may be beneficial for the host
+TMHs, on the other hand, also have their functional constraints, 
+and make up ... of a human, bacterial or viral proteome.
+One can safely assume a proteome contains many TMHs,
+therefore it may be beneficial for the host
 if its immune system would be more attentive towards TMHs.
@@
 It is unknown, when focusing on TMHs only, if a signal of selection
-can be detected.
+can be detected.\frans{ik snap deze redenatie niet?}
 We speculate that the human immune system is more attentive 
@@
 be predicted well from a protein structure,
-are common structures and are present in all pathogens. 
-We could have done the same experiment on beta-turn,
+are common structures and are present in all pathogens. \frans{ dit doen we toch omdat we vinden dat er een preferred binding affinity is voor hydrophobe peptides?}
+We could have done the same experiment on beta-sheets,
 as also these can be predicted well (\cite{petersen2010netturnp}),
@@
 We speculate that the human immune system may also be more attentive 
-towards beta turns for this reason.
+towards beta sheets for this reason.
 
@@
 high performance computing cluster.
-Additionally, we would like to thank Sci-Hub (\cite{himmelstein2018sci})
-for allowing us to read paywalled articles while working from home.
-[GVDB: 
-  Should we include Maxim? 
-  He did the alignment of Mycobacterium tuberculosis. 
-  Let’s also ask Frans.
-]
 
@@
 We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
-bigger queries of multiple (and delayed) API calls.}
+bigger queries of multiple (and delayed) API calls.
 

### 2021-03-24 bbbq_article.tex

@@
 a human representative reference proteome,
-with UniProt ID UP000005640\_9606.
+with UniProt ID UP000005640_9606.
 For the epitopes that were found exactly once in the proteome,
@@
 epitope length (\cite{bergseng2015different}).
-We used a human (UniProt ID UP000005640\_9606), 
+We used a human (UniProt ID UP000005640_9606), 
 viral (SARS-CoV-2, UniProt ID UP000464024) 
@@
 on their impact on the human 
-population \richel{ref}.
+population \frans{Add references here}.
 We used TMHMM  \richel{ref} to predict the topology 
@@
 To detect the evolutionary conservation of TMHs, 
-we collected human single nucleotide polymorphisms (SNPs)
-that resulted in a single amino the substitution,
-and tallied if this substitution occurred within a TMH or not.
-From that tally, we determined how likely the number of SNPs
-found in TMHs is by chance.
+single nucleotide
+polymorphisms (SNPs) within the human population
+that resulted in the substitution of one amino acid were collected from ??? and tallied if
+the SNP occurred within a TMH or not. Using statistics is was determined if there are more, less or an equal amount of SNPs in TMHs
+as a measure of evolutionary conservation.
+This workflow is discussed in more detail below.
 
@@
 As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used.
-Pivotal was the 'dbSNP' (\cite{sherry2001dbsnp}) database,
-which contains 650 million 
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used,
+which are 'gene',\frans{snap ik niet which are gene?} to find the gene names of membrane proteins, 
+'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
+and 'protein', to obtain the sequence of proteins that SNPs act upon.
+The 'dbSNP' contains 650 million 
 catalogued non-redundant humane variations (called RefSNPs,
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
-Additionally, the aptly-named 'gene' (for gene names)
-and 'protein' (for proteins sequences) database were used.
+
+% \paragraph{Data retrieval}
+
+\frans{ dit is relevant waarom: ? To retrieve the data from these databases the
+\verb;rentrez; R package (\cite{rentrez}) was used
+that calls the NCBI website's API. To provide for a 
+stable user experience for all users, 
+this API limits the user to 3 calls per second.
+Additionally, the API splits the result of a bigger
+query into multiple pages, each of which needs one API call.
+We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
+bigger queries of multiple (and delayed) API calls.}
 
@@
 Only variations that result in a SNP for
-a single amino acid substitution were analyzed, resulting in 38,882 SNPs.
+a single amino acid substitution were analysed, resulting in 38,882 SNPs.
 
@@
 SNPs were picked based on ID number, which is linked to their discovery date.
-To verify that ID number is unrelated to SNP positions the relative positions of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
+To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
 

### 2021-03-24 bbbq_article.tex
Merge

--- 2021-03-24 bbbq_article.tex
@@
 From these findings, it is robustly predicted that
-epitopes derived from TMHs can and are presented on both MHC-I and MHC-II.
+epitopes derived from TMHs are presented by B cells [For Frans: please check
+that the article by Bergseng, 2015, 'Different binding motifs of the celiac 
+disease-associated HLA molecules DQ2. 5, DQ2. 2, and DQ7. 5 revealed by 
+relative quantitative proteomics of endogenous peptide repertoires'
+is appropriate. As far as I can see, it is, but I may overlook something. Richel].
 

### 2021-03-24 bbbq_article.tex
Moved 'Prediction software written' to SupMat
Moved Frans' text to Methods

@@
 resulting in 21 haplotypes.
+\richel{This is Frans' text: 
+Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with ....
+variation in peptide binding repertoire, the MHC-II peptide binding groove is determined by both the alpha and beta chain.
+For HLA DQ both the alpha and beta chain is highly polymorphic  .....
+\frans{why is HLA-DP not predicted?}
+We took the alleles that represent percentage of the occurrence int he overall human population according to ... et al. 
+}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Prediction software written}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-[RJCB: I suggest to move this section to the Supplementary Materials]
-\frans{I agree move to supp}
-
-The R programming language is used for the complete 
-experiment, including the analysis.
-The complete experiment is bundled in the 'bbbq' R package,
-which is dependent on 'tmhmm', 'pureseqtmr', 
-'epitope-prediction' and 'mhcnuggetsr'
-as described below.
-
-% \paragraph{tmhmm}
-
-The R package 'tmhmm' was developed to do the similar topology
-predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
-'TMHMM' has a restrictive software license (\cite{krogh2001predicting}) and allows a user
-to download a pre-compiled executable after confirmation he/she
-is in academia. The R package respects this restriction
-and allows the user to install and use TMHMM from within R,
-as done in this study.
-'tmhmm' has been submitted to and is accepted by CRAN. \frans{wat is CRAN}
-
-% \paragraph{pureseqtmr}
-
-To be able to call, from R, the TMH prediction 
-software 'PureseqTM' (\cite{wang2019efficient}),
-which is written in C, the package 'pureseqtmr' has been developed. 
-'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
-Excluded are the features that are used by the 'PureseqTM' 
-developers to verify the correctness of their work.
-'pureseqtmr' has been submitted to and is accepted by CRAN.
-
-% \paragraph{mhcnuggetsr}
-
-MHCnuggets is a free and open-source Python package to predict 
-epitope affinity for many MHC-I and MHC-II variants (\cite{shao2020high}).
-The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
-from within R.
-Also 'mhcnuggetsr' has been submitted to and is accepted by CRAN.
-
-% \paragraph{bbbq}
-
-To reproduce the full experiment presented in this paper,
-the functions needed are bundled in the 'bbbq' R package.
-This package is too specific to be submitted to CRAN.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
@@
 From these findings, it is robustly predicted that
-epitopes derived from TMHs are presented by B cells [For Frans: please check
-that the article by Bergseng, 2015, 'Different binding motifs of the celiac 
-disease-associated HLA molecules DQ2. 5, DQ2. 2, and DQ7. 5 revealed by 
-relative quantitative proteomics of endogenous peptide repertoires'
-is appropriate. As far as I can see, it is, but I may overlook something. Richel].
+epitopes derived from TMHs are presented by B cells.
 
@@
 
-We next wondered if this over-representation would also generalize to MHC-II. \frans{ you say MHC-II but this is very general I think you have to explain the choice for these haplotypes.}
-Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with ....
-variation in peptide binding repertoire, the MHC-II peptide binding groove is determined by both the alpha and beta chain.
-For HLA DQ both the alpha and beta chain is highly polymorphic  .....
-\frans{why is HLA-DP not predicted?}
-We took the alleles that represent percentage of the occurrence int he overall human population according to ... et al. 
+We next wondered if this over-representation would also generalize to MHC-II. 
 Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
@@
 We show that MHC-II has the same pattern of over-presentation
-as MHC-I, yet this over-presentation is more pronounced. \frans{we should explain this does it like shown for mhc-I also linearly correlates with the hydrophopbicity of the MHC-II binding celft?}
+as MHC-I, yet this over-presentation is more pronounced. 
+\frans{
+  we should explain this does it like shown for mhc-I also 
+  linearly correlates with the hydrophopbicity 
+  of the MHC-II binding cleft?
+}
 Also for MHC-II, we observe that the extent to which TMH-derived
@@
 figure \ref{fig:conservation_per_spanner}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}\frans{why comes this section before discussion?}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{General pattern in epitope presentation}
-
-From this study, two important conclusions can be drawn. 
-First, the MHC over-presentation of TMHs is likely a general feature 
-and predicted to occur for most haplotypes of both MHC-I and -II 
-and for humans as well as bacterial and viral pathogens. 
-Second, TMHs are more evolutionary conserved that soluble protein motifs, 
-at least in the human proteome.
-
-% \paragraph{Over-presentation in vivo in MHC-I is robust}
-
-We confirm that 1.4\% of epitopes that are naturally presented on MHC-I 
-can be mapped to stem from a TMH. Additionally, this 
-prediction is robust, as it holds for two TMH prediction tools, 
-as shown by table \ref{tab:elution}).
-
-% \paragraph{Over-presentation in vivo in MHC-II is robust}
-
-We predict that also MHC-II presents epitopes derived from TMHs
-in vivo, as we find that 3.9-4.0\% of sequenced epitopes
-can be uniquely traced onto a human reference 
-proteome (also shown in table \ref{tab:elution}).
-
-% \paragraph{Over-presentation in silico}
-
-Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
-are presented in vivo, we conclude that these are additionally
-presented more often than expected by chance 
-alone (figures \ref{fig:1} and \ref{fig:2}).
-This in silico finding is a general pattern 
-that persists in both MHC-I and MHC-II, 
-for humans, bacteria and viruses.
-
-% \paragraph{Evolutionary conservation}
-
-We conclude that mutations are more likely to occur
-in soluble regions of proteins, than in TMHs.
-When selecting SNPs at random, per 1000 SNPs found in a soluble domain, 
-one finds on average 918 SNPs in TMHs.
-This effect is more pronounced in multi spanners likely as these TMHs more often serve additional functions, besides spanning the lipid bilayer.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{General pattern in epitope presentation}
+
+From this study, two important conclusions can be drawn. 
+First, the MHC over-presentation of TMHs is likely a general feature 
+and predicted to occur for most haplotypes of both MHC-I and -II 
+and for humans as well as bacterial and viral pathogens. 
+Second, TMHs are more evolutionary conserved that soluble protein motifs, 
+at least in the human proteome.
+
+% \paragraph{Over-presentation in vivo in MHC-I is robust}
+
+We confirm that 1.4\% of epitopes that are naturally presented on MHC-I 
+can be mapped to stem from a TMH. Additionally, this 
+prediction is robust, as it holds for two TMH prediction tools, 
+as shown by table \ref{tab:elution}).
+
+% \paragraph{Over-presentation in vivo in MHC-II is robust}
+
+We predict that also MHC-II presents epitopes derived from TMHs
+in vivo, as we find that 3.9-4.0\% of sequenced epitopes
+can be uniquely traced onto a human reference 
+proteome (also shown in table \ref{tab:elution}).
+
+% \paragraph{Over-presentation in silico}
+
+Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
+are presented in vivo, we conclude that these are additionally
+presented more often than expected by chance 
+alone (figures \ref{fig:1} and \ref{fig:2}).
+This in silico finding is a general pattern 
+that persists in both MHC-I and MHC-II, 
+for humans, bacteria and viruses.
+
+% \paragraph{Evolutionary conservation}
+
+We conclude that mutations are more likely to occur
+in soluble regions of proteins, than in TMHs.
+When selecting SNPs at random, per 1000 SNPs found in a soluble domain, 
+one finds on average 918 SNPs in TMHs.
+This effect is more pronounced in multi spanners likely as these TMHs more often serve additional functions, besides spanning the lipid bilayer.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Acknowledgments}
@@
 bigger queries of multiple (and delayed) API calls.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Prediction software written}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The R programming language is used for the complete 
+experiment, including the analysis.
+The complete experiment is bundled in the 'bbbq' R package,
+which is dependent on 'tmhmm', 'pureseqtmr', 
+'epitope-prediction' and 'mhcnuggetsr'
+as described below.
+
+% \paragraph{tmhmm}
+
+The R package 'tmhmm' was developed to do the similar topology
+predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
+'TMHMM' has a restrictive software license (\cite{krogh2001predicting}) and allows a user
+to download a pre-compiled executable after confirmation he/she
+is in academia. The R package respects this restriction
+and allows the user to install and use TMHMM from within R,
+as done in this study.
+'tmhmm' has been submitted to and is accepted by CRAN. \frans{wat is CRAN}
+
+% \paragraph{pureseqtmr}
+
+To be able to call, from R, the TMH prediction 
+software 'PureseqTM' (\cite{wang2019efficient}),
+which is written in C, the package 'pureseqtmr' has been developed. 
+'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
+Excluded are the features that are used by the 'PureseqTM' 
+developers to verify the correctness of their work.
+'pureseqtmr' has been submitted to and is accepted by CRAN.
+
+% \paragraph{mhcnuggetsr}
+
+MHCnuggets is a free and open-source Python package to predict 
+epitope affinity for many MHC-I and MHC-II variants (\cite{shao2020high}).
+The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
+from within R.
+Also 'mhcnuggetsr' has been submitted to and is accepted by CRAN.
+
+% \paragraph{bbbq}
+
+To reproduce the full experiment presented in this paper,
+the functions needed are bundled in the 'bbbq' R package.
+This package is too specific to be submitted to CRAN.
 

### 2021-03-24 bbbq_article.tex
Process feedback Frans

@@
 
-\frans{Show off to Frans}
 Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
@@
 Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
-We performed a linear regression analyses on the data twice,
-one with and one without MAPs,
-and added a 95\% confidence interval for both of these regressions.
+We performed a linear regression analysis on the TMPs
+and added a 95\% confidence interval on this regression.
 Would the confidence interval of the regression line 
@@
 This analysis gives the first hint that SNPs 
-occur more often in soluble domains than in TMHs.\frans{waar vind ik dan nu het verschil tussen met en zonder MAPs je zou hopen dat er geen verschil is toch omdat het aandeel TMH in de analyse met MAPS ook naar beneden gaat}
+occur more often in soluble domains than in TMHs.
 To statistically test whether SNPs occur 
@@
 when randomly selecting SNPs,
+within TMPs that are 50\% TMH (or correcting for this percentage),
 per 100 SNPs found in soluble protein domains, 
@@
 when randomly selecting SNPs, 
+within TMPs that are 50\% TMH (or correcting for this percentage),
 for 100 SNPs found in soluble regions of multi-spanners,
-one finds 89.6 SNPs in TMHs, as shown by \frans{it says 91,1 in fig 4}
+one finds 91.1 SNPs in TMHs, as shown by
 figure \ref{fig:conservation_per_spanner}.
@@
 
-[GVDB: 
-  Discussion section needs some work, 
-  as here interpretation of the results needs to be given. 
-  What is the consequence of the findings? Why is it important. 
-]
-
 % \paragraph{General}
@@
 Our understanding of the immune system and epitope prediction is 
-important to fight disease. By this paper, we suggest that epitopes
+important to fight diseases. By this paper, we suggest that epitopes
 derived from TMHs are needlessly overlooked and may inspire to investigate
-new potential targets for combating pathogens.
-\frans{ik zie hier verder geen argumentatie voor deze statement}
+new potential targets for combating pathogens. We say 'needlessly',
+as TMH-derived epitopes are predicted to be 
+presented more often than expected by chance, 
+combined with are finding that TMHs are evolutionary conserved.
+
 \subsection{Elution studies}
@@
 
-The over representation of TMH on MHC-II haplotypes such as HLADQA!*0101/DQB*0501 is more pronounced as for MHC-I with a third of the predicted binders from the human and MtB proteome stem from TMH  human reference proteome and MtB reference proteome to be presented on MHC-II.
-\frans{ik zoek nog naar associaties}
+The over-presentation of TMHs for MHC-II haplotypes 
+is more pronounced than for MHC-I,
+as with a third of the predicted binders 
+HIERO
+from the human and MtB proteome stem from TMH human reference proteome and MtB reference proteome to be presented on MHC-II. \frans{ik zoek nog naar associaties}
+
+such as \verb;HLA-DQA1*0101/DQB1*0501;
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IC50s per haplotype}
+\subsection{IC50 values of binders per haplotype}
 \label{subsec:ic50s_per_haplotype}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MHC-I}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+show the exact number of binders, binders that overlap with TMHs
+and the percentage of binders that overlap with TMHs, as
+visualized by figures \ref{fig:1} and \ref{fig:2} respectively.
+
+% Label: tab:tmh_binders_mhc1
+\input{bbbq_1_smart_results/table_tmh_binders_mhc1_2.latex}
+
+% Label: tab:tmh_binders_mhc2
+\input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Relative presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To compare the over-presentation of TMH-derived epitopes between the
+different proteomes, we normalized this percentages in such a
+way that 1.0 is the percentage of TMH-derived epitopes that would 
+be expected by chance. 
+Figure \ref{fig:f_tmh_mhc1_normalized} and \ref{fig:f_tmh_mhc2_normalized}
+show these normalized values for the MHC-I and MHC-II haplotypes respectively.
 
@@
 
-% Label: tab:tmh_binders_mhc1
-\input{bbbq_1_smart_results/table_tmh_binders_mhc1_2.latex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MHC-II}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure}[!htbp]
@@
 
-% Label: tab:tmh_binders_mhc2
-\input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}
+To determine the additional over-presentation of TMH-derived epitopes 
+in MHC-II (as compared to MHC-I), we used the same normalized percentages
+for the human proteome only and compared the averages between MHC-I and MHC-II.
+See figure \ref{fig:fig_rel_presentation}.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation.png}
+  \caption{
+    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs
+    for human proteomes.
+  }
+  \label{fig:fig_rel_presentation}
+\end{figure}
 

### 2021-03-25 bbbq_article.tex
Process feedback
Macrh 25th

@@
 This resulted in 1077 gene IDs
-[RJCB:
-  This was December 2020.
-  At 2021-03-01, one will get 1130 gene IDs.
-  Also, one of the gene IDs that was valid back then,
-  has been obsoleted.
-].
+% 1077 gene IDs is correct for December 2020.
+% At 2021-03-01, one will get 1130 gene IDs.
+% Also, one of the gene IDs that was valid back then,
+% has been obsoleted.
 The next query was a call to the 'gene' database 
@@
 as MHC-I, yet this over-presentation is more pronounced. 
-\frans{
-  we should explain this does it like shown for mhc-I also 
-  linearly correlates with the hydrophopbicity 
-  of the MHC-II binding cleft?
+\richel{
+  Showing how general our finding is (as it is explained by hydrophobicity)
+  is important. See \url{https://github.com/richelbilderbeek/bbbq_article/issues/192}
 }
@@
 as with a third of the predicted binders 
-HIERO
+\richel{Hier ben ik nu met verwerken...}
 from the human and MtB proteome stem from TMH human reference proteome and MtB reference proteome to be presented on MHC-II. \frans{ik zoek nog naar associaties} 
-
 such as \verb;HLA-DQA1*0101/DQB1*0501;

### 2021-03-25 bbbq_article.tex

@@
 resulting in 21 haplotypes.
-\richel{This is Frans' text: 
+\richel{This is Frans' text, merge it with the current text: 
 Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with ....
@@
 \end{figure}
+
+HIERO
+Additionally, we investigated if MHC-I and MHC-II are equally prone
+to present TMH-derived epitopes. 
+See \ref{fig:rel_presentation}
+\ref{fig:rel_presentation_per_haplotype.png}
+
+fig_rel_presentation_per_haplotype.png
 
@@
 in TMHs as expected by chance.
-
-[RJCB: How many SNPs in TMHs?]
 See figure \ref{fig:f_tmh_ncbi} to see which percentages
@@
 % \paragraph{MHC-II is expected to present TMHs}
-\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevnet zijn de haplotypen die veel TMH presenteren?}
-
+\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevant zijn de haplotypen die veel TMH presenteren?}
 The over-presentation of TMHs for MHC-II haplotypes 
-is more pronounced than for MHC-I,
+is slight more pronounced than for MHC-I,
 as with a third of the predicted binders 
@@
 for the human proteome only and compared the averages between MHC-I and MHC-II.
+Figure \ref{fig:fig_rel_presentation_per_haplotype} shows the normalized
+TMH-derived epitope presentation per haplotype.
+
 See figure \ref{fig:fig_rel_presentation}.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation_per_haplotype.png}
+  \caption{
+    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs
+    for human proteomes.
+  }
+  \label{fig:fig_rel_presentation_per_haplotype}
+\end{figure}
 

### 2021-03-25 bbbq_article.tex
Progress https://github.com/richelbilderbeek/bbbq_article/issues/193

@@
 
-HIERO
 Additionally, we investigated if MHC-I and MHC-II are equally prone
 to present TMH-derived epitopes. 
-See \ref{fig:rel_presentation}
-\ref{fig:rel_presentation_per_haplotype.png}
-
-fig_rel_presentation_per_haplotype.png
+\richel{Add exact method}
+See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.
 
@@
 in MHC-II (as compared to MHC-I), we used the same normalized percentages
-for the human proteome only and compared the averages between MHC-I and MHC-II.
-Figure \ref{fig:fig_rel_presentation_per_haplotype} shows the normalized
+and compared the averages between MHC-I and MHC-II.
+Figure \ref{fig:rel_presentation_per_haplotype} shows the normalized
 TMH-derived epitope presentation per haplotype.
-
-See figure \ref{fig:fig_rel_presentation}.
+To compare the TMH-derived over-presentation per MHC class, 
+we plot the mean and standard error of the normalized values,
+grouped per haplotype in figure \ref{fig:rel_presentation}.
 
@@
   \caption{
-    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs
-    for human proteomes.
-  }
-  \label{fig:fig_rel_presentation_per_haplotype}
+    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs,
+    for the different haplotypes and proteomes
+  }
+  \label{fig:rel_presentation_per_haplotype}
 \end{figure}
@@
   \caption{
-    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs
-    for human proteomes.
-  }
-  \label{fig:fig_rel_presentation}
+    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs,
+    for the different MHC classes and proteomes. Error bars denote the
+    standard error.
+  }
+  \label{fig:rel_presentation}
 \end{figure}

### 2021-03-26 bbbq_article.tex
Added feedback Geert, thanks @bogeert

@@
 of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
-peptide fragments derived from transmembrane helices (TMHs) 
-are predicted to be presented more often than expected,
+peptide fragments derived from human transmembrane helices (TMHs) 
+are predicted to be presented more often than expected
 based on their abundance in the proteome.
-We show that the over-presentation of TMH-derived peptides is likely general, 
-as it is predicted for diverse microbiota and viruses 
-and for both MHC-I and MHC-II
-However, the physiological reason of why membrane proteins might be 
+We show that this over-presentation of TMH-derived peptides is  general, 
+as it is predicted for diverse bacteria and viruses 
+and for both MHC-I and MHC-II.  However, the physiological reason of why membrane proteins might be 
 over-presented is unclear
 In this study, we show that TMHs are evolutionarily more conserved, 
-as relatively less single nucleotide polymorphisms (SNPs) 
-are present in TMH-coding chromosomal regions 
+as single nucleotide polymorphisms (SNPs) 
+are present relatively less frequent in TMH-coding chromosomal regions 
 compared to regions coding for extracellular and cytosolic protein regions. 
@@
 An important part of the acquired immune response, 
-that develops specialized and more effective recognition of pathogens, 
+that develops specialized and more effective recognition of pathogens than the innate immune response, 
 are T cells which recognize peptides derived from 
-exogenous antigens presented on Major Histocompatibility Complexes (MHC).
+exogenous antigens presented on Major Histocompatibility Complexes (MHC) class I and II.
 
@@
 
-Any individual's immune system detects only a fraction of all possible
-peptide fragments.
-For humans, the MHC proteins are encoded by the
+The MHC proteins are heterodimeric proteins encoded by the
 HLA (Human Leukocyte Antigens) genes.
-There are three genes encoding for MHC-I, which are HLA-A, HLA-B and HLA-C,
-for MHC-II there are three major genes, which are HLA-DR, HLA-DQ, HLA-DP.
+In humans, the peptide-binding groove of MHC-I is made by the alpha chain, and there are three forms of alpha chains called HLA-A, HLA-B and HLA-C. 
+For MHC-II, both the alpha and the beta chains contribute to the binding groove, and there are three major forms as well, called HLA-DR, HLA-DQ and HLA-DP.
 Each MHC complex can only bind a subset of all possible peptides.
@@
 peptides they bind (\cite{lund2004definition})
-[FB: is that really so?]
-[RB: yes, according to the reference that is absent in the docx version to Lund et al., 2004, 'Definition of supertypes for {HLA} molecules using clustering of specificity matrices'] 
-The HLA region of humans is highly polymorphic, with hundreds 
+The HLA genes of humans are highly polymorphic, with hundreds 
 to thousands of different alleles, 
@@
 
-Because of these multiple and highly polymorphic MHC genes,
-the number of pathogenic peptides that can be detected is increased,
-as a wide variety of MHCs (6 MHC-I and 6-8 MHC-II molecules) 
-are expressed per individual
-each presenting their own subset of epitopes.
+While humans only express a limited number of MHC haplotypes, and therefore any individual's immune system detects only a fraction of all possible
+peptide fragments, the number of pathogenic peptides that can be detected is increased at the population level, because of the multiple and highly polymorphic MHC genes.
 It is believed that this improves immunity at the population level, 
@@
 will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
+[GEERT: I removed the nr of MHC molecules at is is more complex, as the alpha chains coded by the maternal chromosome of HLA-DQ,DR,DP can heterodimeriza with the beta chains coded by the paternal chromosome and vice versa]
 
@@
 These studies have led to the development 
-of very reliable prediction algorithms 
-that allows for in silico predictions 
-of the binding affinities of peptides (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
+of  prediction algorithms 
+that allow for very reliable in silico predictions 
+of the binding affinities of peptides (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).[GEERT: can you put 'very reliable' in more quantitative terms (e.g.
+with overall more than 90\% accuracy)]
 
@@
 Using these prediction algorithms, 
-we recently showed that peptides derived 
+we recently predicted that peptides derived 
 from transmembrane helices (TMHs) 
-are predicted to be more frequently presented by MHC-I 
+will be more frequently presented by MHC-I 
 than expected based on their abundance (\cite{bianchi2017}). 
-Moreover, we showed that some well-known immunodominant peptides stem from TMHs.
-
-This over-presentation is attributed to the fact 
+Moreover, we showed that some well-known immunodominant peptides stem from TMHs. This over-presentation is attributed to the fact 
 that the peptide-binding groove of most MHC-I haplotypes 
@@
 as they need to span the hydrophobic lipid bilayer of cellular membranes.
-They consist of an alpha helix on average 23 amino acids in length. 
-TMHs are common structures in the proteins of humans and pathogens. 
+They consist of an alpha helix of on average 23 amino acids in length. TMHs can also be predicted from a protein sequence 
+with high accuracy [GEERT: more quantitative?] by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
+TMHs are common structures in the proteins of humans and microbes. 
 Different TMH prediction tools estimate
-that 15-39\% of the proteins in humans 
+that 15-39\% of the human proteome 
 contain at least one TMH (\cite{ahram2006estimation}).
-TMHs can also be predicted from a protein sequence 
-with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
 
@@
 by the functional requirement to span a lipid bilayer. 
-Due to these requirements, 
-the amino acids genuinely present in TMH 
+Due to this requirement, 
+many of the amino acids genuinely present in TMH 
 are limited to the ones with hydrophobic side chains (\cite{jones1994model}).
@@
 that were predicted to be inside ...][RJCB: Uh, you want to add another
-paragraph here?]
-Therefore, the TMHs of pathogens 
-might have a lower chance to develop an escape mutations, 
+paragraph here?][GEERT: I think it is OK like Richel wrote]
+Therefore, we speculated that the TMHs of pathogens 
+might have a lower chance to develop escape mutations, 
 as many mutations will result in a dysfunctional TMH 
@@
 
-From an elution study, we find that epitopes derived 
-from TMHs are presented on naturally occurring MHC-II,
-by finding the origin of these epitopes on a reference proteome,
-of which we predicted all proteins' topologies.
-From in silico predictions, we conclude that TMH-derived
+This study has two objectives.
+First, we aimed to generalize our findings by predicting the presentation of peptides from different kingdoms of life and for both MHC-I and -II.
+From these in silico predictions, we conclude that TMH-derived
 epitopes are presented more often than expected by chance,
-in a human, viral and bacterial proteome, for both MHC-I and II,
-as these epitopes have a higher affinity with the presenter molecule.
-From the generality of this pattern, we speculated that TMHs
-are likely to be evolutionarily conserved. From an analysis using human
-SNPs we found that these random point mutations are indeed less likely
-to occur within TMHs. These findings encourage us to investigate the
-overlooked importance of TMHs in the immune response and vaccine development.
+in a human, viral and bacterial proteome, and for most haplotypes of both MHC-I and II.
+We confirmed this over-presentation of TMH-derived peptides by re-analyzis of a peptide elution study.
+Second, we tested our hypothesis that TMHs
+are more evolutionarily conserved than  solvent-exposed protein regions. Our analysis of human
+single nucleotide polymorphisms (SNPs) showed that random point mutations are indeed less likely
+to occur within TMHs. These findings strengthen the emerging notion that TMHs are important for the T cell branch of the adaptive immune system, and hence are of  
+overlooked importance in vaccine development.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Elution studies}\label{subsec:elution_studies}
+\subsection{Peptide elution studies}\label{subsec:elution_studies}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 with UniProt ID UP000005640\_9606.
-For the epitopes that were found exactly once in the proteome,
-we predicted the topology of the proteins in which the epitopes are found.
+For the epitopes that were present only a single time in the proteome,
+we predicted the topology of the proteins in which these epitopes are found.
 We used both TMHMM and PureseqTM to do these predictions. 
 From this topology, we could determine if the epitope
-overlapped with a transmembrane regions.
+overlapped with a transmembrane region.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Measuring TMH epitopes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-To determine how frequently epitopes overlapping with TMHs are presented,
-we applied a similar analysis strategy as described in \cite{bianchi2017}.
-To investigate the generality of this, we applied this analysis
-for several haplotypes of both MHC-I and MHC-II, 
+\subsection{Predicting TMH epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To predict how frequently epitopes overlapping with TMHs are presented,
+we applied a similar analysis strategy as described in \cite{bianchi2017} for several haplotypes of both MHC-I and MHC-II, 
 as well as for a human, viral and bacterial proteome.
 To summarize, for each proteome, 
-all possible 9-mers (for MHC-I) or 14-mers (MHC-II) are derived. 
-For each of these peptides, we determined if it overlapped with a 
-TMH and if it was a binder for each haplotype.
-
-For MHC-I, 9-mers were used, as this is the length used in our earlier
-study (\cite{bianchi2017}). For MHC-II, 14-mers were used, 
-as these are the most frequently occurring
+all possible 9-mers (for MHC-I) or 14-mers (MHC-II) were derived. 
+For each of these peptides, we determined if it overlapped with a predicted 
+TMH and if it was predicted to bind to each haplotype.
+
+For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I and was used in our earlier
+study (\cite{bianchi2017}). For MHC-II, 14-mers were used,  as these are the most frequently occurring
 epitope length (\cite{bergseng2015different}).
@@
 viral (SARS-CoV-2, UniProt ID UP000464024) 
-and bacterial (Mycobacterium tuberculosis (MTb), UniProt ID UP000001584) 
-reference proteome. The pathogens were chosen based 
-on their impact on the human 
-population \richel{ref}.
-We used TMHMM  \richel{ref} to predict the topology 
-of the proteins within the proteome.
+and bacterial (\emph{Mycobacterium tuberculosis} (MTb), UniProt ID UP000001584) 
+reference proteome. We used TMHMM  \richel{ref} to predict the topology 
+of the proteins within these proteomes.
 To predict the affinity of an epitope to a certain MHC molecule,
@@
 used in our previous study (\cite{bianchi2017}).
-The MHC-II haplotypes used additionally were selected 
-to occur with a phenotypic frequency of at least 14\% in
+The most frequent MHC-II haplotypes used  were selected, with a phenotypic frequency of at least 14\% in
 the human population (\cite{greenbaum2011functional}),
-resulting in 21 haplotypes.
+resulting in 21 MHC-II haplotypes.
 \richel{This is Frans' text, merge it with the current text: 
@@
 
-This study differs in some aspects from our previous work (\cite{bianchi2017}), 
+This study differs in one important aspect from our previous work (\cite{bianchi2017}), 
 The definition of a binder differs from \cite{bianchi2017}:
@@
 % See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
-We believe our revised definition precludes bias of proteins 
+We believe that our revised definition precludes bias of proteins 
 that give rise to no or only very few MHC epitopes
@@
 we collected human single nucleotide polymorphisms (SNPs)
-that resulted in a single amino the substitution,
-and tallied if this substitution occurred within a TMH or not.
-From that tally, we determined how likely the number of SNPs
-found in TMHs is by chance.
+that resulted in a single amino acid substitution,
+and determined if this substitution occurred within a predicted TMH or not.
 
@@
 As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used.
-Pivotal was the 'dbSNP' (\cite{sherry2001dbsnp}) database,
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: the 'dbSNP' (\cite{sherry2001dbsnp}) database,
 which contains 650 million 
 catalogued non-redundant humane variations (called RefSNPs,
-\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}).
-Additionally, the aptly-named 'gene' (for gene names)
-and 'protein' (for proteins sequences) database were used.
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases 'gene' (for gene names)
+and 'protein' (for proteins sequences). [GEERT: any references for databases gene and protein?]
 
@@
 The first query was a call to the 'gene' database for the 
-term 'membrane protein' (in all fields) for the organism \textbf{Homo sapiens}.
-This resulted in 1077 gene IDs
+term 'membrane protein' (in all fields) for the organism \emph{Homo sapiens}.
+This resulted in 1,077 gene IDs.
 % 1077 gene IDs is correct for December 2020.
@@
 to obtain the gene names from the gene IDs.
-Per gene name, the 'dbSNP' NCBI database is queried for 
+Per gene name, the 'dbSNP' NCBI database was queried for 
 variations associated with the gene name.
 
-The number of variations
+The number of SNPs
 was limited to the first 250 variations per gene (see figure
@@
 SNPs were picked based on ID number, which is linked to their discovery date.
-To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
+To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.[GEERT: rood practice is to present the data in a chronological fashion: thus move this to figure 1 (or put in supplement)]
 
@@
 protein sequence.
-Of each protein sequence, the protein topology was determined 
+For each protein sequence, the protein topology was determined 
 using PureseqTM.
-Using the protein topology SNPs were scored to be in or outside trans membrane domains.
+Using these predicted protein topologies, the SNPs were scored to be located within or outside trans membrane domains.
+[GEERT: for Table 1: Tables don't have a legend, but instead have a title and footnotes below the table.]
 
@@
 
-To confirm that peptides stemming from TMHs were presented in MHC-I and MHC-II,
+[GEERT: I would move this subsection to below: first explain the bioinformatics predictions, and then show this as an experimental confirmation]
+
+To obtain experimental confirmation that peptides stemming from TMHs are presented in MHC-I and MHC-II,
 peptide elution studies were reanalyzed.
-Table \ref{tab:elution} shows the percentage of epitopes derived 
-from a TMH
-found in MHC-I and MHC-II elution 
-studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} respectively),
-for the two topology prediction tools TMHMM and 
-PureseqTM (\cite{krogh2001predicting} and \cite{wang2019efficient} respectively). 
-Regardless of the prediction, 
-at least 100 epitopes were predicted to be derived from a TMH. 
+Table \ref{tab:elution} shows the percentages of epitopes derived 
+from TMHs
+found in the MHC-I (\cite{schellens2015comprehensive}) and MHC-II elution 
+studies  (\cite{bergseng2015different}),
+for the two topology prediction tools TMHMM (\cite{krogh2001predicting}) and PureseqTM (\cite{wang2019efficient}). 
+Regardless of the prediction tool, 
+at least 100 epitopes were predicted to be derived from a TMH for each condition. 
 From these findings, it is robustly predicted that
-epitopes derived from TMHs are presented by B cells.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MHC-I}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We recently showed that peptides derived from TMHs of human proteins 
-were predicted to be over-presented on HLA-A and HLA-B. 
-Here, we wondered if this finding would be general 
+epitopes derived from TMHs are presented in both MHC-I and MHC-II.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Prediction of TMH-derived peptide presentation in MHC-I}
+[GEERT: mooier is je subheadings samen te vatten met een conclusie. Dus: TMH-derived peptides are predicted to be over-presented in MHC-I (en dan ook voor de andere subheadings)]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We recently predicted that peptides derived from TMHs of human proteins 
+are over-presented on most haplotypes of HLA-A and HLA-B. Here, we wondered if this over-presentation is general 
 and if peptides stemming from predicted TMHs of microbial pathogens 
-would also be over-presented on HLA-A and HLA-B.We therefor analysed all superhaplotypes of the HLA-A and HLA-B class see Figure \ref{fig:1} and found that similar as for the human proteome the percentages of MHC-I epitopes predicted to be overlapping with TMHs for our human, viral and bacterial proteome was higher as expected by TMH abundance.
-See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) 
-The extent of the over-presentation between the various different HLA supertypes is similar for the probed proteomes, which confirms that the hydrophopbicity of the MHC binding cleft is determinant.
+would also be over-presented on HLA-A and HLA-B.
+We therefor analysed all superhaplotypes[GEERT: not explained, for instance in the Methods] of the HLA-A and HLA-B class (Figure \ref{fig:1}).
+Similar as for the human proteome, the percentages of MHC-I epitopes predicted to be overlapping with TMHs  was higher as expected by TMH abundance for the proteomes of SARS-CoV-2 and \emph{M.
+tuberculosis}.
+See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) [GEERT what is in supplementary table 6? The numbers?]
+The extent of the over-presentation between the various different HLA supertypes is similar for the probed proteomes, which strenghtens our previous conclusion [ref] that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
 
@@
   \caption{
-    Percentage of MHC-I epitopes overlapping with TMHs
-    for a human, viral and bacterial proteome.
-    The horizontal lines indicate the percentage as expected by chance.
-    See table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
+    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
+    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
+    The dashed horizontal lines indicate the percentage as expected by chance.
+    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{MHC-II}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We next wondered if this over-representation would also generalize to MHC-II. 
-Figure \ref{fig:2} shows the percentages of MHC-II epitopes overlapping 
-with TMHs for our human, viral and bacterial proteome.
-See the supplementary materials (table \ref{tab:tmh_binders_mhc2}) 
+\subsection{Prediction of TMH-derived peptide presentation in MHC-II}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We next wondered if the over-representation of TMH-derived peptided would also be present for MHC-II. 
+Figure \ref{fig:2} shows the percentages of MHC-II epitopes predicted to be overlapping 
+with TMHs for our human, viral and bacterial proteomes.
+See supplementary table \ref{tab:tmh_binders_mhc2} 
 for the TMH and epitope counts.
-We show that MHC-II has the same pattern of over-presentation
-as MHC-I, yet this over-presentation is more pronounced. 
+We predict that TMH-derived peptides are also over-presented in most MHC-II haplotypes, and this over-presentation seems even more pronounced than for MHC-I. 
 \richel{
@@
 }
-Also for MHC-II, we observe that the extent to which TMH-derived
-epitopes are presented is similar between the haplotypes, for
-a human, bacterial and viral proteome.
+Similar to our previous results with MHC-I, we predicted that the extent to which TMH-derived
+epitopes are presented by each MHC-II haplotype is similar for the human, bacterial and viral proteome.
 
@@
   \caption{
-    Percentage of MHC-II epitopes overlapping with TMHs
+    Percentage of MHC-II epitopes predicted to be overlapping with TMHs
     for a human, viral and bacterial proteome.
-    The horizontal lines indicate the percentage as expected by chance.
-    See table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
-    Note that for smaller proteomes a percentage of zero is likelier.
+    The dashed horizontal lines indicate the percentage as expected by chance.
+    See supplementary table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
   }
@@
 \richel{Add exact method}
-See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.
+See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}. [GEERT: Wat is er in 9 en 10? Wat is de conclusie? Dit moet nog wat verder uitgewerkt worden]
 
@@
 \subsection{Evolutionary conservation}
+[GEERT: zie mijn punt hierbove. Een korte conclusie als titel is veel sterker. Dus iets als: TMHs are more evolutionary conserved.]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \caption{
-    Evolutionary conservation of SNPs in TMHs.
+    Evolutionary conservation of SNPs in TMHs. [GEERT: legenad uitwerken. Wat is precies laten zien?]
   }
@@
     Evolutionary conservation of SNPs in TMHs,
-    for either single- or multi-spanners
+    for either single- or multi-spanners/ [GEERT: ook uitwerken]
   }
@@
     Percentage of SNPs found in TMHs.
-    Each point resembles one protein, with a predicted percentage of
-    TMH (x-axis) and an observed occurrence of SNPs being located
-    within a TMH (y-axis).
-    The dashed diagonal line shows the expected linear trend line
-    if TMHs and soluble protein regions are equally conserved.
-    The red line is a linear trend line that includes the membrane-associated
-    proteins at the origin. 
+    Each point shows the predicted percentage of
+    TMH (x-axis) and the observed occurrence of SNPs being located
+    within a TMH (y-axis) for a protein.
+    The dashed diagonal line shows the line of equality (i.e.
+    equal conservation of TMHs and soluble protein regions).
+    The red line [GEERT: where is the red line? Behind the blue line? That makes this useless and can be removed.]is a linear trend line that originates from the origin. 
     The blue line is a linear trend line that includes only the
@@
 
-In general the occurrence and expression levels of membrane proteins are low, therefor we addressed the question whether there is an evolutionary advantage for presenting TMHs.
-To this end we looked for conservation the conservation in TMH vs soluble regions by comparing the occurrences of SNPs in the genome coding for TMHs versus soluble protein regions within membrane proteins.
-We obtained >1000 gene names associated with the phrase 'membrane protein',
+We addressed the question whether there is an evolutionary advantage in presenting TMHs.
+We determined the conservation of TMHs by comparing the occurrences of SNPs located in TMHs or solvent-exposed protein regions for the genes coding for membrane proteins.
+We obtained $>$1,000 gene names associated with the phrase 'membrane protein',
 which are genes for both membrane-associated proteins (MAPs, no TMH) and 
@@
 These genes are linked to 4,780 protein isoforms, 
-of which 2,553 are predicted to be TMPs (hence 
-2,237 proteins are predicted to be MAPs).
+of which 2,553 are predicted to be TMPs and 
+2,237 proteins are predicted to be MAPs.
 We obtained 37,630 SNPs that resulted in an
@@
 See tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
-for an overview of all amounts.
+for the detailed numbers and distributions of SNPs.
 
 Per protein, we calculated two percentages: 
-(1) the percentage of the protein coding for TMHs, 
-and (2) the percentage of SNPs occurring in TMHs.
+(1) the percentage of the total protein predicted to be  TMHs, 
+and (2) the percentage of SNPs located within these predicted TMHs.
 Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
@@
 overlap with the line of equality,
-this would be expected if SNPs occur just as likely within TMHs 
-and soluble protein domains.
-This analysis gives the first hint that SNPs 
+this would indicate that SNPs occur just as likely within TMHs 
+and solvent-exposed protein domains (i.e., no higher conservation of TMHs). However, we observed that ...[GEERT give brief explanation in words of what is seen in the graph]
+This finding suggests that SNPs 
 occur more often in soluble domains than in TMHs.
+
 To statistically test whether SNPs occur 
@@
 in TMHs, we first calculated the number of SNPs that are expected to be
-found in TMHs of TMPs by chance, $E(n_{SNP\_TMH})$,
-equals the number of SNPs in TMHs, $n_{SNP} = 21,738$, 
-multiplied by the fraction of the TMPs that are TMH, $f_{TMH} = 18.66\%$,
-$E(n_{SNP\_TMH}) = n_{SNP} \cdot f_{TMH} \approx 4056$.
-
-The effect size is that, 
+found in TMHs of TMPs by chance ($E(n_{SNP\_TMH})= n_{SNP} \cdot f_{TMH} \approx 4,056$), by multiplying the number of SNPs predicted to be in TMHs ($n_{SNP} = 21,738$ ) with the fraction of the TMPs that are predicted TMH.
+The actual number of SNPs predicted to be located within TMHs
+is $n_{SNP\_TMH} = 3,903$, which is lower than the expectation
+of $E(n_{SNP\_TMH}) \approx 4,056$.
+
+To test whether this difference is significant,
+we used a binomial Poisson distribution, where the number of trails ($n$) equals the number of SNPs, which is 21,738. The probability of success
+for the $i$th TMP ($p\_i$), is the percentage of each TMP that is TMH.
+We calculated the probability of having 3,903 or less successes.
+We used the type I error cut-off value of $\alpha = 2.5\%$.
+The chance to find, within TMHs, this amount or less SNPs 
+equals [RJCB: recalculate]
+$1.6201 \cdot 10^{-9}$.
+
+Using a alpha value of $0.05$, as is common practice for an uninformed
+expectation, we conclude that there are significantly less SNPs
+in TMHs as expected by chance.
+See figure \ref{fig:f_tmh_ncbi} for the percentages of TMHs in the transmembrane proteins.[GEERT: staat deze figuur hier wel goed? Niet naar boven?] The effect size is that, 
 when randomly selecting SNPs,
@@
 as depicted in figure \ref{fig:conservation}.
-
-The actual number of SNPs found in TMHs
-was $n_{SNP\_TMH} = 3,903$, which is lower than the expectation
-of $E(n_{SNP\_TMH}) \approx 4056$.
-To test whether this difference is significant,
-we used a binomial Poisson distribution, where the number of trails,
-$n$ equals the number of SNPs, which is 21,738. The probability of success
-for TMP $i$, $p\_i$, is the percentage of each TMP that is TMH.
-We calculated the probability of having 3,903 or less successes.
-We used the type I error cut-off value of $\alpha = 2.5\%$.
-The chance to find, within TMHs, this amount or less SNPs 
-equals [RJCB: recalculate]
-$1.6201 \cdot 10^{-9}$.
-
-Using a alpha value of $0.05$, as is common practice for an uninformed
-expectation, we conclude that there are significantly less SNPs
-in TMHs as expected by chance.
-See figure \ref{fig:f_tmh_ncbi} to see which percentages
-are TMH of the transmembrane proteins.
+[GEERT: zie mijn eerder opmerking over panels.
+Ik zou met panel A, B, C, etc werken.] [GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}.
+Ik denk dat je deze hieronder kan samenvoegen]
 
@@
   \caption{
-    Percentage of SNPs found in TMHs per spanner.
-    The dashed diagonal lines show the expected linear trend line
-    if TMHs and soluble protein regions are equally conserved.
+    Percentage of SNPs found in TMPs predicted to have only a single (left) or multiple (right) TMHs.
+    The dashed diagonal lines show the expected line of equality (i.e., equal conservation of TMHs and solvent-exposed protein regions).
   }
@@
 
-We decided to split this analysis based on the number of TMHs
-a protein has 
-[GVDB: Why? Provide brief rationale again. 
-Thus TMH-TMH interactions or something]. 
-We hypothesize that single-spanners (i.e. proteins
-with one TMH predicted) are less conserved, when compared to multi-spanners,
-as single spanners just need to span a membrane, while multi-spanners
-might have interactions between their TMHs, 
+We splitted this analysis based on the number of TMHs
+of each TMP. 
+We hypothesized that single-membrane spanners (i.e., proteins
+with a single predicted TMH) are less conserved than multi-membrane spanners (with $>$1 TMH),
+because multi-membrane spanners
+might have protein-protein interactions between their TMHs, 
 for example to accommodate active sites, and 
@@
 As can be seen in figure \ref{fig:f_snps_found_and_expected_per_spanner}, 
-both single- and multi-spanners are evolutionarily conserved.
-
-We conclude that,
-when randomly selecting SNPs, 
-within TMPs that are 50\% TMH (or correcting for this percentage),
-for 100 SNPs found in soluble regions of multi-spanners,
-one finds 91.1 SNPs in TMHs, as shown by
-figure \ref{fig:conservation_per_spanner}.
+both single- and multi-spanners are evolutionarily conserved.[GEERT: maar een sterker dan de andere? Kun je die conclusie niet trekken?]
 
@@
 
-Our understanding of the immune system and epitope prediction is 
-important to fight diseases. By this paper, we suggest that epitopes
-derived from TMHs are needlessly overlooked and may inspire to investigate
-new potential targets for combating pathogens. We say 'needlessly',
-as TMH-derived epitopes are predicted to be 
-presented more often than expected by chance, 
-combined with are finding that TMHs are evolutionary conserved.
+Epitope prediction is 
+important to understand the immune system and for the design of vaccines. In this study, we provide evidence that epitopes
+derived from TMHs are a major but overlooked source of MHC epitopes. Both our bioinformatics predictions and reanalysis of peptide elution studies show that TMH-derived epitopes are 
+presented more often than expected by chance, regardless of the organism. Moreover, our SNP analysis shows  
+that TMHs are evolutionary more conserved than solvent-exposed protein regions.
 
 \subsection{Elution studies}
+[GEERT: ook hier: veel sterker is een korte conclusie, of boodschap]
 
@@
 
-From elution studies and in silico predictions, 
-we suggest that TMH-derived epitopes are presented and cross-presented in vivo.
-How integral membrane proteins are presented and cross-presented on MHC complexes is however unclear.
-Most prominently because the fundamental principles of trans locating integral membrane proteins from there hydrophobic environment into the vacuole lumen and subsequent processing prior to loading on MHC complexes is unresolved.
+Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown [GEERT: cite Bianchi].
+Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved.
+[GEERT: beetje kort.
+Suggesties?
+Experimental approaches?
+Hoeft niet super origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen.
+Frans kan vast helpen met mooie zin.]
 
@@
 
-An additional unknown is, although our findings are robust,
-whether this is a false positive,
-due to additive methodological artifacts.
-An obvious artifact is that the elution studies used 
-will contain some unavoidable sequencing errors.
-Additionally, we assume that the epitopes naturally presented
-are derived from a cleanly translated proteome.
-Two known biological phenomena that violate this assumption
+An important caveat of the elution studies is that we assume that the found epitopes are derived from their completely translated predecessor proteins.
+However, two known biological phenomena that violate this assumption
 are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
@@
 two processes that both result in peptide fragments
-that did not originate from genome directly.
-Also, the TMH prediction software is imperfect.
+that did not originate from the mature proteins.
+[GEERT: mooi argument, maar moet wel kort uitgewerkt worden.
+Uitleggen wat DRiPs zijn en wat peptide fusion is, en hoe het kan leiden tot false positives]
+
+Another caveat is that the TMH prediction tools are imperfect.
 This imperfection, however, appears to be of limited importance,
-as two different topology prediction software algorithms gave similar results.
-Although those a couple of percent of epitopes that 
-can be mapped directly to the proteome,
-there is methodological bias caused by mass spectronomy,
+as our data show that two different topology prediction software algorithms give similar results.
+
+A final caveat in the elution studies is that there is methodological bias caused by mass spectronomy,
 as hydrophobic fragments are detected less 
-by mass spectronomy. This means that the number of TMH-derived epitopes
-is underestimated.
-in vivo work in the lab will be needed to resolve this finding.
+by mass spectronomy [GEERT: referentie, zie paper Bianchi]. This means that the number of TMH-derived epitopes
+might be underestimated.
 
@@
 as with a third of the predicted binders 
-\richel{Hier ben ik nu met verwerken...}
+\richel{Hier ben ik nu met verwerken...} [GEERT: dan stop ik hier eventjes? Of de rest ook weer keer lezen? Laat maar weten]
 from the human and MtB proteome stem from TMH human reference proteome and MtB reference proteome to be presented on MHC-II. \frans{ik zoek nog naar associaties} 

### 2021-03-26 bbbq_article.tex
Process part of Geert's feedback

@@
 will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
-[GEERT: I removed the nr of MHC molecules at is is more complex, as the alpha chains coded by the maternal chromosome of HLA-DQ,DR,DP can heterodimeriza with the beta chains coded by the paternal chromosome and vice versa]
+% Geert: I removed the number of MHC molecules as is is more complex, 
+% as the alpha chains coded by the maternal chromosome 
+% of HLA-DQ, DR, DP can heterodimerize with the beta chains coded 
+% by the paternal chromosome and vice versa
 
@@
 that allow for very reliable in silico predictions 
-of the binding affinities of peptides (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).[GEERT: can you put 'very reliable' in more quantitative terms (e.g.
-with overall more than 90\% accuracy)]
+of the binding affinities of peptides (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
+For example, \cite{tang2011genome} found that, 
+of the 432 peptides that were predicted to bind to MHC,
+86\% were confirmed to do so in vivo.
 
@@
 They consist of an alpha helix of on average 23 amino acids in length. TMHs can also be predicted from a protein sequence 
-with high accuracy [GEERT: more quantitative?] by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}).
+with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}),
+for example, \cite{jones2007improving} found that,
+from 184 TMPs with known topology, 
+80\% of the TMH predictions replicated this finding.
+
 TMHs are common structures in the proteins of humans and microbes. 
@@
 contain at least one TMH (\cite{ahram2006estimation}).
-
 However, the physiological reason why peptides derived from TMHs 
@@
 are limited to the ones with hydrophobic side chains (\cite{jones1994model}).
-[FB: Analyzing the occurrence of single nucleotide polymorphisms 
-that were predicted to be inside ...][RJCB: Uh, you want to add another
-paragraph here?][GEERT: I think it is OK like Richel wrote]
 Therefore, we speculated that the TMHs of pathogens 
@@
 catalogued non-redundant humane variations (called RefSNPs,
-\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases 'gene' (for gene names)
-and 'protein' (for proteins sequences). [GEERT: any references for databases gene and protein?]
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases 'gene' (for gene names, \cite{brown2015gene})
+and 'protein' (for proteins sequences, \cite{sayers2010database}).
 
@@
 SNPs were picked based on ID number, which is linked to their discovery date.
-To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.[GEERT: rood practice is to present the data in a chronological fashion: thus move this to figure 1 (or put in supplement)]
+To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.[GEERT: good practice is to present the data in a chronological fashion: thus move this to figure 1 (or put in supplement)]
 

### 2021-03-26 bbbq_article.tex
Processed feedback Geert

@@
 \begin{abstract}
+
+\richel{TODO: Tables don't have a legend, but instead have a title and footnotes below the table. Modify the tables to make it so in the end}
+\richel{TODO: Condense figures into bigger ones with panels, following 1 message per figure. Order these figures chronologically}
 
@@
 will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
-% Geert: I removed the number of MHC molecules as is is more complex, 
+% Do noty give the number of MHC molecules, as is is more complex, 
 % as the alpha chains coded by the maternal chromosome 
@@
 SNPs were picked based on ID number, which is linked to their discovery date.
-To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.[GEERT: good practice is to present the data in a chronological fashion: thus move this to figure 1 (or put in supplement)]
+To verify that ID number is unrelated to SNP positions the relative positons of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
 
@@
 Using these predicted protein topologies, the SNPs were scored to be located within or outside trans membrane domains.
-[GEERT: for Table 1: Tables don't have a legend, but instead have a title and footnotes below the table.]
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Elution studies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH-derived peptides are predicted to be over-presented in MHC-I}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  First do the bioinformatics predictions,
+  then the experimental confirmation 
+}
+
+We recently predicted that peptides derived from TMHs of human proteins 
+are over-presented on most haplotypes of HLA-A and HLA-B. Here, we wondered if this over-presentation is general 
+and if peptides stemming from predicted TMHs of microbial pathogens 
+would also be over-presented on HLA-A and HLA-B.
+We therefor analysed all superhaplotypes[GEERT: not explained, for instance in the Methods] of the HLA-A and HLA-B class (Figure \ref{fig:1}).
+Similar as for the human proteome, the percentages of MHC-I epitopes predicted to be overlapping with TMHs  was higher as expected by TMH abundance for the proteomes of SARS-CoV-2 and \emph{M.
+tuberculosis}.
+See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) [GEERT what is in supplementary table 6? The numbers?]
+The extent of the over-presentation between the various different HLA supertypes is similar for the probed proteomes, which strenghtens our previous conclusion [ref] that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
+
+\begin{figure}[!htbp]
+  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
+  \caption{
+    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
+    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
+    The dashed horizontal lines indicate the percentage as expected by chance.
+    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
+  }
+  \label{fig:1}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH-derived peptides are predicted to be over-presented in MHC-II}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We next wondered if the over-representation of TMH-derived peptided would also be present for MHC-II. 
+Figure \ref{fig:2} shows the percentages of MHC-II epitopes predicted to be overlapping 
+with TMHs for our human, viral and bacterial proteomes.
+See supplementary table \ref{tab:tmh_binders_mhc2} 
+for the TMH and epitope counts.
+We predict that TMH-derived peptides are also over-presented in most MHC-II haplotypes, and this over-presentation seems even more pronounced than for MHC-I. 
+\richel{
+  Showing how general our finding is (as it is explained by hydrophobicity)
+  is important. See \url{https://github.com/richelbilderbeek/bbbq_article/issues/192}
+}
+Similar to our previous results with MHC-I, we predicted that the extent to which TMH-derived
+epitopes are presented by each MHC-II haplotype is similar for the human, bacterial and viral proteome.
+
+\begin{figure}[!htbp]
+  \includegraphics[height=0.8\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
+  \caption{
+    Percentage of MHC-II epitopes predicted to be overlapping with TMHs
+    for a human, viral and bacterial proteome.
+    The dashed horizontal lines indicate the percentage as expected by chance.
+    See supplementary table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
+  }
+  \label{fig:2}
+\end{figure}
+
+Additionally, we investigated if MHC-I and MHC-II are equally prone
+to present TMH-derived epitopes. 
+\richel{Add exact method}
+See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}. [GEERT: Wat is er in 9 en 10? Wat is de conclusie? Dit moet nog wat verder uitgewerkt worden]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH-derived peptides are predicted to be presented in vivo}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\richel{
+  Use this section as an experimental confirmation 
+  of the bioinformatics predictions
+}
 
@@
 \input{bbbq_article_issue_157/results.tex}
-
-[GEERT: I would move this subsection to below: first explain the bioinformatics predictions, and then show this as an experimental confirmation]
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Prediction of TMH-derived peptide presentation in MHC-I}
-[GEERT: mooier is je subheadings samen te vatten met een conclusie. Dus: TMH-derived peptides are predicted to be over-presented in MHC-I (en dan ook voor de andere subheadings)]
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We recently predicted that peptides derived from TMHs of human proteins 
-are over-presented on most haplotypes of HLA-A and HLA-B. Here, we wondered if this over-presentation is general 
-and if peptides stemming from predicted TMHs of microbial pathogens 
-would also be over-presented on HLA-A and HLA-B.
-We therefor analysed all superhaplotypes[GEERT: not explained, for instance in the Methods] of the HLA-A and HLA-B class (Figure \ref{fig:1}).
-Similar as for the human proteome, the percentages of MHC-I epitopes predicted to be overlapping with TMHs  was higher as expected by TMH abundance for the proteomes of SARS-CoV-2 and \emph{M.
-tuberculosis}.
-See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) [GEERT what is in supplementary table 6? The numbers?]
-The extent of the over-presentation between the various different HLA supertypes is similar for the probed proteomes, which strenghtens our previous conclusion [ref] that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
-
-\begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
-  \caption{
-    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
-    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
-    The dashed horizontal lines indicate the percentage as expected by chance.
-    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
-  }
-  \label{fig:1}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Prediction of TMH-derived peptide presentation in MHC-II}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We next wondered if the over-representation of TMH-derived peptided would also be present for MHC-II. 
-Figure \ref{fig:2} shows the percentages of MHC-II epitopes predicted to be overlapping 
-with TMHs for our human, viral and bacterial proteomes.
-See supplementary table \ref{tab:tmh_binders_mhc2} 
-for the TMH and epitope counts.
-We predict that TMH-derived peptides are also over-presented in most MHC-II haplotypes, and this over-presentation seems even more pronounced than for MHC-I. 
-\richel{
-  Showing how general our finding is (as it is explained by hydrophobicity)
-  is important. See \url{https://github.com/richelbilderbeek/bbbq_article/issues/192}
-}
-Similar to our previous results with MHC-I, we predicted that the extent to which TMH-derived
-epitopes are presented by each MHC-II haplotype is similar for the human, bacterial and viral proteome.
-
-\begin{figure}[!htbp]
-  \includegraphics[height=0.8\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
-  \caption{
-    Percentage of MHC-II epitopes predicted to be overlapping with TMHs
-    for a human, viral and bacterial proteome.
-    The dashed horizontal lines indicate the percentage as expected by chance.
-    See supplementary table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
-  }
-  \label{fig:2}
-\end{figure}
-
-Additionally, we investigated if MHC-I and MHC-II are equally prone
-to present TMH-derived epitopes. 
-\richel{Add exact method}
-See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}. [GEERT: Wat is er in 9 en 10? Wat is de conclusie? Dit moet nog wat verder uitgewerkt worden]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Evolutionary conservation}
-[GEERT: zie mijn punt hierbove. Een korte conclusie als titel is veel sterker. Dus iets als: TMHs are more evolutionary conserved.]
+\subsection{Human TMHs are evolutionarily conserved}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 as depicted in figure \ref{fig:conservation}.
-[GEERT: zie mijn eerder opmerking over panels.
-Ik zou met panel A, B, C, etc werken.] [GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}.
+[GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}.
 Ik denk dat je deze hieronder kan samenvoegen]
@@
 
-Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown [GEERT: cite Bianchi].
+Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown (\cite{bianchi2017}).
 Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved.
@@
 as hydrophobic fragments are detected less 
-by mass spectronomy [GEERT: referentie, zie paper Bianchi]. This means that the number of TMH-derived epitopes
+by mass spectronomy (\cite{bianchi2017}). This means that the number of TMH-derived epitopes
 might be underestimated.

### 2021-03-29 bbbq_article.tex
Process feedback Geert, thanks @bogeert, progress https://github.com/richelbilderbeek/bbbq_article/issues/196

@@
 
-\richel{TODO: Tables don't have a legend, but instead have a title and footnotes below the table. Modify the tables to make it so in the end}
-\richel{TODO: Condense figures into bigger ones with panels, following 1 message per figure. Order these figures chronologically}
+\richel{TODO: Tables don't have a legend, but instead have a title and footnotes below the table.
+Modify the tables to make it so in the end, \url{https://github.com/richelbilderbeek/bbbq_article/issues/181}}
+\richel{TODO: Condense figures into bigger ones with panels, following 1 message per figure. Order these figures chronologically,
+\url{https://github.com/richelbilderbeek/bbbq_article/issues/195}}
 
@@
 
+\richel{I wonder if the order of the Methods, should match the improved
+order of the Results as well, so first the MHC-I and MHC-II presentation,
+followed by the experimental confirmation}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 resulting in 21 MHC-II haplotypes.
-\richel{This is Frans' text, merge it with the current text: 
-Unlike MHC-I where only one variable chain determines the landscape of the epitope binding pocket allowing for the classification of 5 HLA-A and 8 HLA-B superhaplotypes with ....
-variation in peptide binding repertoire, the MHC-II peptide binding groove is determined by both the alpha and beta chain.
-For HLA DQ both the alpha and beta chain is highly polymorphic  .....
-\frans{why is HLA-DP not predicted?}
-We took the alleles that represent percentage of the occurrence int he overall human population according to ... et al. 
-}
+
+The MHC-I molecule presents epitopes of shorter length than MHC-II.
+We expect this to be of no influence in the presentation level
+of TMH-derived epitopes, for each of the three proteomes.
+We verified this, by normalizing the percentage of TMH-derived epitopes found
+to the percentage of TMH-derived epitopes expected, 
+where a value of $1.0$ denotes that the percentage of TMH-derived epitopes found
+matches the percentage as expected by chance.
+
+Additionally, we investigated if MHC-I and MHC-II are equally prone
+to present TMH-derived epitopes. 
+\richel{Add exact method}
+See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.
+
+In our previous work, we found that the observed pattern can be
+explained from the hydrophobicity of the MHC-I binding 
+cleft (\cite{bianchi2017}), we validated.
+Here we do the same analysis for MHC-II \richel{Explain the analysis here}.
 
@@
 a binder as having an IC50 in the lowest 2\% 
-of the peptides within a \emph{protein}.
+of the peptides within a \emph{protein}, 
+which put less strain on the bioinformatics pipeline.
 % See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
+\richel{Show side-by-side-comparison in supplementary materials}
 We believe that our revised definition precludes bias of proteins 
@@
 variations associated with the gene name. 
+As the NCBI API constrains its users to three calls per second
+to assure fair use, we had to limit the extent of our analysis.
 
@@
 
-We recently predicted that peptides derived from TMHs of human proteins 
-are over-presented on most haplotypes of HLA-A and HLA-B. Here, we wondered if this over-presentation is general 
-and if peptides stemming from predicted TMHs of microbial pathogens 
-would also be over-presented on HLA-A and HLA-B.
-We therefor analysed all superhaplotypes[GEERT: not explained, for instance in the Methods] of the HLA-A and HLA-B class (Figure \ref{fig:1}).
-Similar as for the human proteome, the percentages of MHC-I epitopes predicted to be overlapping with TMHs  was higher as expected by TMH abundance for the proteomes of SARS-CoV-2 and \emph{M.
-tuberculosis}.
-See the supplementary materials (table \ref{tab:tmh_binders_mhc1}) [GEERT what is in supplementary table 6? The numbers?]
-The extent of the over-presentation between the various different HLA supertypes is similar for the probed proteomes, which strenghtens our previous conclusion [ref] that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
+Figure \ref{fig:1} shows the predicted presentation of TMH-derived peptides in MHC-I,
+for a human, viral and bacterial proteome.
+Per MHC-I haplotype, it shows the percentage of binders that overlap with a TMH 
+for at least one residue.
+A dashed line indicates the expected percentage of TMH-derived epitopes 
+that would be presented, if TMH-derived epitopes would be presented just as 
+likely as soluble regions.
+In 12 out of 13 MHC-I haplotypes, TMH-derived epitopes are presented more often 
+then the null expectation, for a human proteome. 
+For the viral and bacterial proteome, 11 out of 13 haplotypes present
+TMH-derived epitopes more often than expected by chance.
+The extent of the over-presentation between the various different haplotypes 
+is similar for the probed proteomes, which strenghtens our previous conclusion (\cite{bianchi2017}) that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
 
@@
     See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
+    \richel{Add panel with hydrophobicity}
   }
@@
 
+Plotting the normalized values of presented TMH-derived epitopes (where
+$1.0$ denotes that the percentage of presented TMH-derived epitopes
+matches the values as expected by chance) in \ref{fig:rel_presentation}
+shows that MHC-II has a higher relative over-presentation of TMH-derived
+epitopes, when compared to MHC-I. \richel{HIERO}
+The MHC-I molecule presents epitopes of shorter length than MHC-II.
+We expect this to be of no influence in the presentation level
+of TMH-derived epitopes, for each of the three proteomes.
+We verified this, by normalizing the percentage of TMH-derived epitopes found
+to the percentage of TMH-derived epitopes expected, 
+where a value of $1.0$ denotes that the percentage of TMH-derived epitopes found
+matches the percentage as expected by chance.
+\richel{TODO: explain only the results found. Elaborate. Draw a conclusion}
 Additionally, we investigated if MHC-I and MHC-II are equally prone
@@
 \richel{Add exact method}
-See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}. [GEERT: Wat is er in 9 en 10? Wat is de conclusie? Dit moet nog wat verder uitgewerkt worden]
+See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.
 
@@
   \caption{
-    Evolutionary conservation of SNPs in TMHs. [GEERT: legenad uitwerken. Wat is precies laten zien?]
+    The number of SNPs in TMHs as expected by chance (left bar) 
+    and found in the dbSNP database (right bar).
+    Percentages are added to show the relative conservation
+    of SNPs in TMHs found.
   }
@@
   \caption{
-    Evolutionary conservation of SNPs in TMHs,
-    for either single- or multi-spanners/ [GEERT: ook uitwerken]
+    The number of SNPs in TMHs as expected by chance (red bars) 
+    and found in the dbSNP database (right bars), 
+    after splitting the data into single- and multi-spanners.
+    Percentages are added to show the relative conservation
+    of SNPs in TMHs found.
   }
@@
     equal conservation of TMHs and soluble protein regions).
-    The red line [GEERT: where is the red line? Behind the blue line? That makes this useless and can be removed.]is a linear trend line that originates from the origin. 
     The blue line is a linear trend line that includes only the
@@
 
-\subsection{Elution studies}
-[GEERT: ook hier: veel sterker is een korte conclusie, of boodschap]
+\subsection{TMH-derived epitopes are predicted to be presentated in vivo}
+
+% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
+We show that 1.4\% of epitopes that are naturally presented on MHC-I 
+can be mapped to stem from a TMH. We also find that MHC-II presents epitopes derived from TMHs in vivo, as we find that 3.9-4.0\% of sequenced epitopes
+can be uniquely traced onto a human reference 
+proteome (also shown in table \ref{tab:elution}).
 
@@
 An important caveat of the elution studies is that we assume that the found epitopes are derived from their completely translated predecessor proteins.
-However, two known biological phenomena that violate this assumption
-are defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
-and peptide fusion (\cite{delong2016pathogenic}),
+However,defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
+and peptide fusion (\cite{delong2016pathogenic}) are
 two processes that both result in peptide fragments
@@
 as our data show that two different topology prediction software algorithms give similar results.
+Additionally, this 
+prediction is robust, as it holds for two TMH prediction tools, 
+as shown by table \ref{tab:elution}).
 
@@
 
-\subsection{Presentation of epitopes derived from TMHs}
-
-% \paragraph{MHC-II is expected to present TMHs}
-\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevant zijn de haplotypen die veel TMH presenteren?}
-The over-presentation of TMHs for MHC-II haplotypes 
-is slight more pronounced than for MHC-I,
-as with a third of the predicted binders 
-\richel{Hier ben ik nu met verwerken...} [GEERT: dan stop ik hier eventjes? Of de rest ook weer keer lezen? Laat maar weten]
-from the human and MtB proteome stem from TMH human reference proteome and MtB reference proteome to be presented on MHC-II. \frans{ik zoek nog naar associaties} 
-such as \verb;HLA-DQA1*0101/DQB1*0501;
-
-This study predict that TMH-derived epitopes are over-presented
-in MHC-II as well. 
-The intracellular pathway of MHC-II differs from that of MHC-I,
-as it acquires epitopes from the exterior of the cell by phagocytosis.
-How TMHs derived from exterior TMPs are lysed and loaded onto MHC-II 
-is unknown. in vivo work is needed to elucidate this novel pathway.
-
-% \paragraph{Immunogenicity}
-
-In this study, we find that TMH-derived epitopes are over-presented.
-The presentation of an epitope, however, 
-is only a first step of an adaptive immune response.
-We assume that the presentation of epitopes is
-correlated with evoking a full immune response.
-TMH-derived epitopes, however, are more hydrophobic than the average
-epitope. If this evokes a stronger or weaker immune response is unknown,
-and in vivo work is needed to resolve this.
-
-\subsection{Evolutionary conservation of TMHs}
-
-% \paragraph{Low number of SNPs}
-
-The NCBI dbSNP database contains millions of SNPs.
-As the NCBI API constrains its users to three calls per second
-to assure fair use, we had to limit the extent of our analysis.
-We assume we have taken a representative sample, but only
-an analysis using the full NCBI database can confirm this.
-
-% \paragraph{Selection undetectable in whole proteome}
-
-In general, on would hope that evolutionary selection results in
-an immune system that as most attentive for loci that are
-essential for a virus, as these will be most conserved.
-In SARS-CoV-2, for example, there is preliminary evidence that the strongest
-selection pressure is upon residues that changes its 
-virulence (\cite{velazquez2020positive}).
-These loci, however, only account for a small part of a pathogen's proteome.
-Additionally, these essential parts differ widely between pathogens.
-Because of this scarcity and variance in targets, 
-one can imagine that the human immune system 
-is not tailored to detect these sites, 
-as hinted upon by an extensive study on influenza (\cite{han2019individual}),
-where it was found that the selection pressure
-exerted by the immune system was a non pro founding factor.
-
-% \paragraph{Selection may be detectable in TMHs}
-
-TMHs, on the other hand, also have their functional constraints, 
-and make up ... of a human, bacterial or viral proteome.
-One can safely assume a proteome contains many TMHs,
-therefore it may be beneficial for the host
-if its immune system would be more attentive towards TMHs.
-And maybe this has already happened: MHC-I already detects hydrophobic
-peptides. This feature, however, may also be caused by selection
-to detect hydrophobic regions in the soluble proteins of pathogens.
-It is unknown, when focusing on TMHs only, if a signal of selection
-can be detected.\frans{ik snap deze redenatie niet?}
-We speculate that the human immune system is more attentive 
-towards TMHs, as it is a general structure, that is evolutionarily conserved.
-
-We aimed our evolutionary experiment at TMHs, because these can
-be predicted well from a protein structure,
-are common structures and are present in all pathogens. \frans{ dit doen we toch omdat we vinden dat er een preferred binding affinity is voor hydrophobe peptides?}
-We could have done the same experiment on beta-sheets,
-as also these can be predicted well (\cite{petersen2010netturnp}),
-are common structures and are present in all pathogens.
-We speculate that the human immune system may also be more attentive 
-towards beta sheets for this reason.
-
-[GVDB: 
-  Also compare your findings with previous efforts 
-  to look at conservation of TMHs
-]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{General pattern in epitope presentation}
-
-From this study, two important conclusions can be drawn. 
-First, the MHC over-presentation of TMHs is likely a general feature 
-and predicted to occur for most haplotypes of both MHC-I and -II 
-and for humans as well as bacterial and viral pathogens. 
-Second, TMHs are more evolutionary conserved that soluble protein motifs, 
-at least in the human proteome.
-
-% \paragraph{Over-presentation in vivo in MHC-I is robust}
-
-We confirm that 1.4\% of epitopes that are naturally presented on MHC-I 
-can be mapped to stem from a TMH. Additionally, this 
-prediction is robust, as it holds for two TMH prediction tools, 
-as shown by table \ref{tab:elution}).
-
-% \paragraph{Over-presentation in vivo in MHC-II is robust}
-
-We predict that also MHC-II presents epitopes derived from TMHs
-in vivo, as we find that 3.9-4.0\% of sequenced epitopes
-can be uniquely traced onto a human reference 
-proteome (also shown in table \ref{tab:elution}).
+\subsection{TMH-derived epitopes are presented more often than expected by chance}
 
@@
 Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
-are presented in vivo, we conclude that these are additionally
+are presented , we conclude that these are additionally
 presented more often than expected by chance 
@@
 
-% \paragraph{Evolutionary conservation}
-
-We conclude that mutations are more likely to occur
+% \paragraph{MHC-II is expected to present TMHs}
+\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevant zijn de haplotypen die veel TMH presenteren?}
+The over-presentation of TMHs for most MHC-II haplotypes 
+is more pronounced than for MHC-I haplotypes,
+as up to a third of the predicted MHC-II binders 
+\richel{Hier ben ik nu met verwerken...} 
+from the human and MtB proteome stem from TMHs. \frans{ik zoek nog naar associaties} 
+such as \verb;HLA-DQA1*0101/DQB1*0501;
+
+This results from this study support our hypothesis that TMH-derived epitopes are over-presented
+in MHC-II, similarly to we previously showed for MHC-I [CITE BIANCHI].[GEERT: deel gewist, want was letterlijke herhaling van hierboven.
+Beter is zeggen dat (1) MHC-II presentatie belangrijk is voor B cell activatie.
+(2) B cellen antilichamen produceren, en dan antilichamen solvent-exposed protein regions herkennen.
+(3) Dat onze analyse suggereert HMC-II presentatie voornamelijk andere (TMH of buried) regio's betreft.
+(4) Dat T helper cellen dus andere delen van antigene eiwitten herkennen (MHC-II gepresenteerd door de B cellen) dan de BCR bindt.]
+
+\subsection{Less SNPs in TMHs found, hinting at evolutionary conservation}
+
+We find that mutations are more likely to occur
 in soluble regions of proteins, than in TMHs.
@@
 
+% \paragraph{Selection undetectable in whole proteome}
+
+In general, one would hope that evolutionary selection results in
+an immune system that as most attentive for protein regions that are
+essential for the survival, proliferation and/or virulence or pathogenic microbes, as these will be most conserved.
+In SARS-CoV-2, for example, there is preliminary evidence that the strongest
+selection pressure is upon residues that change its 
+virulence (\cite{velazquez2020positive}).
+These regions, however, only account for a small part of a pathogen's proteome.
+Additionally, these essential parts might differ widely between different pathogenic proteins.
+Because of this scarcity and variance in targets, 
+one can imagine that it will be mostly unfeasible to provide innate immune responses against such essential protein regions, 
+as hinted upon by a study on influenza (\cite{han2019individual}),
+where it was found that the selection pressure
+exerted by the immune system was a non profounding factor.
+
+% \paragraph{Selection may be detectable in TMHs}
+
+TMHs, on the other hand, also have strict functional constraints as they need to span through hydrophobic lipid bilayers.
+Moreover, TMHs are abundant and make up ...
+of a human, bacterial or viral proteome.
+Proteomes thus contain many TMHs,
+and it might therefore be beneficial for host organisms
+to tailor their immune systems towards immune responses against foreign TMHs.
+This feature, however, may also be caused by selection
+to detect hydrophobic regions in the soluble proteins of pathogens.
+It is unknown, when focusing on TMHs only, if a signal of selection
+can be detected.\frans{ik snap deze redenatie niet?}[GEERT: ik ook niet, vind beetje onnodig.
+Volstaat te zeggen dat TMHs niet de enige hydrofobe regio's zijn, en ook belang voor intra-protein en protein-protein interacties en catalytische sites.
+Daarnaast in deze paragraaf iets zeggen over single vs multi-TMH eiwitten]
+We speculate that the human immune system is more attentive 
+towards TMHs, as it is a general structure, that is evolutionarily conserved.
+
+[GVDB: 
+  Also compare your findings with previous efforts 
+  to look at conservation of TMHs
+]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{General pattern in epitope presentation}
+
+From this study, two important conclusions can be drawn. 
+First, the MHC over-presentation of TMHs is likely a general feature 
+and predicted to occur for most haplotypes of both MHC-I and -II 
+and for humans as well as bacterial and viral pathogens. 
+Second, TMHs are more evolutionary conserved than solvent-exposed protein motifs, 
+at least in the human proteome.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 high performance computing cluster.
+GvdB is funded by a Young Investigator Grant from the Human Frontier Science Program (HFSP; RGY0080/2018), and a Vidi grant from the Netherlands Organization for Scientific Research (NWO-ALW VIDI 864.14.001).
+GvdB has received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement No.
+862137.
 
@@
 
-RJCB and FB conceived the idea for this research. 
+RJCB and FB conceived the idea for this research. MVB helped with the proteome analysis of \emph{M. tuberculosis}.
 RJCB wrote the code.
-RJCB and FB wrote the article.
+RJCB, FB and GvdB wrote the article.
 
@@
 
-A part of this study does the same analysis as Bianchi et al., 2017.
-Here we describe the deviations, which are about the use of different
-software and the use of a different definition of what a binder is.
+A part of this study does the same analysis as Bianchi et al., 2017.mainly concern the use of different
+software and a different definition of what an MHC binder is.
 
@@
 
-The earlier study uses \verb;epitope-prediction; a hand-crafted method, 
+The earlier study used \verb;epitope-prediction; a hand-crafted method [GEERT: what is hand-crafted? They didn't use a computer? Why relevant?], 
 that has been trained on MHC-I haplotypes only,
 which is used here again. For MHC-II IC50 predictions, the
-'mhcnuggetsr' R package is used.
+'mhcnuggetsr' R package was used in this study.
 
@@
 
-The earlier study defines a peptide a binder (for a haplotype), 
-if \emph{within the peptide} in which it is found, 
-is within the peptides with the 2\% lowest IC50 values.
+The earlier study defined a peptide an MHC binder 
+if \emph{within the protein} in which it was found, 
+is was among the peptides with the 2\% lowest IC50 values.
 This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
@@
 
-In this study, a peptide is defined a binder (for a haplotype), 
-if within a \emph{proteome} in which it is found, 
-is within the peptides with the 2\% lowest IC50 values,
-where our previous study used the lowest 2\% IC50 values
-within each \emph{protein}.
+However, in this study, an MHC binder is defined as a peptide within a \emph{proteome} in which it is found, that is among the peptides with the 2\% lowest IC50 values.
 Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
-for a binder per haplotype.
+for a binder per haplotype. We believe that our revised definition is more correct, as it overcomes bias from proteins with very low numbers of peptides and/or MHC-predicted binders.
 
@@
 reference proteome.
-For the sake of reproducible research, we use the desktop version (as
+For the sake of reproducible research, we used the desktop version (as
 we can call it from scripts) and, due to this, we removed the
@@
 
-For this research, the scientific literature was explored 
-to identify the most recent free and open source (FOSS) prediction software.
-This was done by searching for papers that (1) reference older
+For this research, we needed software to predict protein
+topology, as well as the MHC-I and MHC-II binding affinities
+of epitopes. We selected our software, by
+searching the scientific literature 
+to identify the most recent free and open source (FOSS) 
+prediction software.
+This was done by searching for papers that (1) cite older
 prediction software, and (2) present a novel method to make predictions.
-As a starting point, a review paper was used.
-For all software needed, we found no studies that compared contemporary tools 
-in their prediction quality.
+As a starting point, per type of prediction software,
+a review paper was used (\cite{moller2001evaluation} for protein
+topology, \cite{lundegaard2011prediction} for MHC-I
+binding affinities and \cite{nielsen2003reliable} for MHC-II binding
+affinities).
 
@@
 In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
-of which TMHMM (\cite{krogh2001predicting}) turned out to be the best, 
+of which TMHMM (\cite{krogh2001predicting}) turned out to be the most accurate, 
 as is used in the previous study (\cite{bianchi2017}).
@@
 NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
-gave rise to the most accurate predictions.
-The most recent and promising FOSS tool available now appears
+made the most accurate predictions.
+The most recent FOSS tool available now appears
 to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
-and MHC-II prediction, which we used for MHC-II predictions.
+and MHC-II predictions. 
+As we already use \verb;epitope-prediction; (\cite{bianchi2017}) 
+for MHC-I predictions, we use MHCnuggets only for MHC-II predictions.
 
@@
 'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
-Excluded are the features that are used by the 'PureseqTM' 
-developers to verify the correctness of their work.
 'pureseqtmr' has been submitted to and is accepted by CRAN.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Percentage of spots and spots that overlap with a TMH}
+\subsection{Prediction of percentage of epitopes overlapping with a TMH}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 where a target specifies the source of the proteome,
-where \verb;covid; denotes SARS-CoV-2 and \verb;covid; denotes
+where \verb;covid; denotes SARS-CoV-2 and \verb;myco; denotes
 Mycobacterium tuberculosis. \verb;mhc_class; denotes the MHC
@@
 The TRDD1 ('T cell receptor delta diversity 1') protein,
-however, is two peptides long. 
+however, is two amino acids long. 
 The R package \verb;pureseqtmr;, however, 

### 2021-03-29 bbbq_article.tex
Process feedback Geert, thanks @bogeert, progress https://github.com/richelbilderbeek/bbbq_article/issues/196

@@
 when randomly selecting SNPs,
-within TMPs that are 50\% TMH (or correcting for this percentage),
 per 100 SNPs found in soluble protein domains, 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Discussion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{TMH-derived epitopes are predicted to be presentated in vivo}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved.
-[GEERT: beetje kort.
-Suggesties?
-Experimental approaches?
-Hoeft niet super origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen.
-Frans kan vast helpen met mooie zin.]
+\geert{
+  beetje kort. Suggesties? Experimental approaches? Hoeft niet super
+  origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. 
+  Frans kan vast helpen met mooie zin
+}
 
@@
 
-An important caveat of the elution studies is that we assume that the found epitopes are derived from their completely translated predecessor proteins.
-However,defective ribosomal products (DRiPs) (\cite{yewdell1996defective}) 
-and peptide fusion (\cite{delong2016pathogenic}) are
-two processes that both result in peptide fragments
-that did not originate from the mature proteins.
-[GEERT: mooi argument, maar moet wel kort uitgewerkt worden.
-Uitleggen wat DRiPs zijn en wat peptide fusion is, en hoe het kan leiden tot false positives]
+An important caveat of the elution studies is that we assume that 
+the found epitopes are derived from their completely 
+translated predecessor proteins.
+However, there are multiple cellular processes 
+that may result in peptide fragments that did not originate 
+from the mature proteins.
+One such process is the production of defective ribosomal 
+products (DRiPs) (\cite{yewdell1996defective}),
+in which the ribosome terminates its translation prematurely,
+intrudes a point mutation, 
+or cause a frame shift, at low frequencies (\cite{yewdell2011drips}).
+Another process that results in peptide fragments absent in the
+proteome is peptide fusion, 
+in which peptides form covalent bonds with other peptides,
+as is for example found in mice (\cite{delong2016pathogenic}).
+Due to these processes, we may incorrectly attribute an
+epitope to a specific proteome locations, 
+where in reality, it was created de novo.
+This may lead to false positives, as we incorrectly deduce
+epitopes to be derived from a TMH, instead of from one of
+these processes. Both processes, however, are thought to
+occur at low frequencies (\cite{yewdell2011drips,delong2016pathogenic}).
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{TMH-derived epitopes are presented more often than expected by chance}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Less SNPs in TMHs found, hinting at evolutionary conservation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 where it was found that the selection pressure
-exerted by the immune system was a non profounding factor.
+exerted by the immune system was either weak or absent.
 
@@
 
-TMHs, on the other hand, also have strict functional constraints as they need to span through hydrophobic lipid bilayers.
-Moreover, TMHs are abundant and make up ...
-of a human, bacterial or viral proteome.
-Proteomes thus contain many TMHs,
-and it might therefore be beneficial for host organisms
-to tailor their immune systems towards immune responses against foreign TMHs.
-This feature, however, may also be caused by selection
-to detect hydrophobic regions in the soluble proteins of pathogens.
-It is unknown, when focusing on TMHs only, if a signal of selection
-can be detected.\frans{ik snap deze redenatie niet?}[GEERT: ik ook niet, vind beetje onnodig.
-Volstaat te zeggen dat TMHs niet de enige hydrofobe regio's zijn, en ook belang voor intra-protein en protein-protein interacties en catalytische sites.
-Daarnaast in deze paragraaf iets zeggen over single vs multi-TMH eiwitten]
+Evolutionary selection of pathogens by a host's immune system,
+however, is likelier to occur for proteomic patterns that are general,
+over patterns that are rare.
+Where essential catalytic sites in a pathogenic proteome
+are rare, TMHs are common and thus a more feasible 
+target for evolution to respond to.
+We have found the signature of evolution when both factors,
+that is, TMHs and catalytic sites are likely to co-occur,
+which is in proteins that span the membrane at least twice.
+This may be explained additionally that multi-spoanners
+are not only likelier to have a catalytically active site,
+also there are interactions between the TMHs.
+For single-spanners, we found no evolutionary conservation.
 We speculate that the human immune system is more attentive 
-towards TMHs, as it is a general structure, that is evolutionarily conserved.
-
-[GVDB: 
+towards TMHs in multi-spanners, as these are evolutionarily conserved.
+
+\geert{
   Also compare your findings with previous efforts 
   to look at conservation of TMHs
-]
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Acknowledgments}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix
 \section{Supplementary materials}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Differences with Bianchi et al., 2017}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A part of this study does the same analysis as Bianchi et al., 2017.mainly concern the use of different
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A part of this study does the same analysis as Bianchi et al., 2017.
+mainly concern the use of different
 software and a different definition of what an MHC binder is.
-
-% \paragraph{IC50 prediction software}
-
-The earlier study used \verb;epitope-prediction; a hand-crafted method [GEERT: what is hand-crafted? They didn't use a computer? Why relevant?], 
-that has been trained on MHC-I haplotypes only,
-which is used here again. For MHC-II IC50 predictions, the
-'mhcnuggetsr' R package was used in this study.
 

### 2021-03-29 bbbq_article.tex
Combine figures into a 1-message-per-figure multi-panel figure. Fix https://github.com/richelbilderbeek/bbbq_article/issues/195

@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
-  \caption{
-    The number of SNPs in TMHs as expected by chance (left bar) 
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
+    \label{fig:conservation}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
+    \label{fig:f_snps_found_and_expected}
+  \end{subfigure}
+  \caption{
+      A. The number of SNPs in TMHs as expected by chance (left bar) 
       and found in the dbSNP database (right bar).
@@
       of SNPs in TMHs found.
-  }
-  \label{fig:conservation}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
-  \caption{
-    The number of SNPs in TMHs as expected by chance (red bars) 
-    and found in the dbSNP database (right bars), 
-    after splitting the data into single- and multi-spanners.
-    Percentages are added to show the relative conservation
-    of SNPs in TMHs found.
-  }
-  \label{fig:conservation_per_spanner}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
-  \caption{
+      B. 
       Percentage of SNPs found in TMHs.
@@
   }
-  \label{fig:f_snps_found_and_expected}
-\end{figure}
+\end{figure}
+
+%\begin{figure}[!htbp]
+%  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
+%  \caption{
+%    The number of SNPs in TMHs as expected by chance (left bar) 
+%    and found in the dbSNP database (right bar).
+%    Percentages are added to show the relative conservation
+%    of SNPs in TMHs found.
+%  }
+%  \label{fig:conservation}
+%\end{figure}
+
+%\begin{figure}[!htbp]
+%  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
+%  \caption{
+%    Percentage of SNPs found in TMHs.
+%    Each point shows the predicted percentage of
+%    TMH (x-axis) and the observed occurrence of SNPs being located
+%    within a TMH (y-axis) for a protein.
+%    The dashed diagonal line shows the line of equality (i.e.
+%    equal conservation of TMHs and soluble protein regions).
+%    The blue line is a linear trend line that includes only the
+%    transmembrane proteins.
+%  }
+%  \label{fig:f_snps_found_and_expected}
+%\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.3\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
+    \label{fig:conservation_per_spanner}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.55\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
+    \label{fig:f_snps_found_and_expected_per_spanner}
+  \end{subfigure}
+  \caption{
+      A. 
+      The number of SNPs in TMHs as expected by chance (red bars) 
+      and found in the dbSNP database (right bars), 
+      after splitting the data into single- and multi-spanners.
+      Percentages are added to show the relative conservation
+      of SNPs in TMHs found.
+      B. 
+      Percentage of SNPs found in TMPs predicted to have only a single
+      (left) or multiple (right) TMHs.
+      The dashed diagonal lines show the expected line of equality 
+      (i.e., equal conservation of TMHs and solvent-exposed 
+      protein regions).
+  }
+\end{figure}
+
+%\begin{figure}[!htbp]
+%  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
+%  \caption{
+%    The number of SNPs in TMHs as expected by chance (red bars) 
+%    and found in the dbSNP database (right bars), 
+%    after splitting the data into single- and multi-spanners.
+%    Percentages are added to show the relative conservation
+%    of SNPs in TMHs found.
+%  }
+%  \label{fig:conservation_per_spanner}
+%\end{figure}
+
+%\begin{figure}[!htbp]
+%  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
+%  \caption{
+%    Percentage of SNPs found in TMPs predicted to have only a single (left) or multiple (right) TMHs.
+%    The dashed diagonal lines show the expected line of equality (i.e., equal conservation of TMHs and solvent-exposed protein regions).
+%  }
+%  \label{fig:f_snps_found_and_expected_per_spanner}
+%\end{figure}
+)
 
@@
 Ik denk dat je deze hieronder kan samenvoegen]
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
-  \caption{
-    Percentage of SNPs found in TMPs predicted to have only a single (left) or multiple (right) TMHs.
-    The dashed diagonal lines show the expected line of equality (i.e., equal conservation of TMHs and solvent-exposed protein regions).
-  }
-  \label{fig:f_snps_found_and_expected_per_spanner}
-\end{figure}
-)
 

### 2021-03-29 bbbq_article.tex
Clean up single-figure-per-panel LaTex, add boldface to paneled figures

--- 2021-03-29 bbbq_article.tex
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation.png}
-  \caption{
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
+    \label{fig:conservation}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
+    \label{fig:f_snps_found_and_expected}
+  \end{subfigure}
+  \caption{
+      \textbf{(A)} 
       The number of SNPs in TMHs as expected by chance (left bar) 
@@
       of SNPs in TMHs found.
-  }
-  \label{fig:conservation}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
-  \caption{
-    The number of SNPs in TMHs as expected by chance (red bars) 
-    and found in the dbSNP database (right bars), 
-    after splitting the data into single- and multi-spanners.
-    Percentages are added to show the relative conservation
-    of SNPs in TMHs found.
-  }
-  \label{fig:conservation_per_spanner}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
-  \caption{
+      \textbf{(B)}
       Percentage of SNPs found in TMHs.
@@
   }
-  \label{fig:f_snps_found_and_expected}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.3\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
+    \label{fig:conservation_per_spanner}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.55\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
+    \label{fig:f_snps_found_and_expected_per_spanner}
+  \end{subfigure}
+  \caption{
+      \textbf{(A)} 
+      The number of SNPs in TMHs as expected by chance (red bars) 
+      and found in the dbSNP database (right bars), 
+      after splitting the data into single- and multi-spanners.
+      Percentages are added to show the relative conservation
+      of SNPs in TMHs found.
+      \textbf{(B)} 
+      Percentage of SNPs found in TMPs predicted to have only a single
+      (left) or multiple (right) TMHs.
+      The dashed diagonal lines show the expected line of equality 
+      (i.e., equal conservation of TMHs and solvent-exposed 
+      protein regions).
+  }
 \end{figure}

### 2021-03-29 bbbq_article.tex
Put TMH-derived epitopes figures into one with two panels.

@@
 
-\begin{figure}[!htbp]
-  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
-  \caption{
-    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
+    \label{fig:1}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
+    \label{fig:2}
+  \end{subfigure}
+  \caption{
+    TMH-derived peptides are predicted to be over-presented in MHC-I (left
+    panel) and MHC-II (right panel).
+    Both panels show the percentage of epitopes predicted to be 
+    overlapping with TMHs
     for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
     The dashed horizontal lines indicate the percentage as expected by chance.
-    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
+    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+    for the exact TMH and  epitope counts.
     \richel{Add panel with hydrophobicity}
   }
-  \label{fig:1}
-\end{figure}
+\end{figure}
+
+%\begin{figure}[!htbp]
+%  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
+%  \caption{
+%    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
+%    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
+%    The dashed horizontal lines indicate the percentage as expected by chance.
+%    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
+%    \richel{Add panel with hydrophobicity}
+%  }
+%  \label{fig:1}
+%\end{figure}
+%
+%\begin{figure}[!htbp]
+%  \includegraphics[height=0.8\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
+%  \caption{
+%    Percentage of MHC-II epitopes predicted to be overlapping with TMHs
+%    for a human, viral and bacterial proteome.
+%    The dashed horizontal lines indicate the percentage as expected by chance.
+%    See supplementary table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
+%  }
+%  \label{fig:2}
+%\end{figure}
 
@@
 epitopes are presented by each MHC-II haplotype is similar for the human, bacterial and viral proteome.
-
-\begin{figure}[!htbp]
-  \includegraphics[height=0.8\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
-  \caption{
-    Percentage of MHC-II epitopes predicted to be overlapping with TMHs
-    for a human, viral and bacterial proteome.
-    The dashed horizontal lines indicate the percentage as expected by chance.
-    See supplementary table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
-  }
-  \label{fig:2}
-\end{figure}
 

### 2021-03-29 bbbq_article.tex
Convert table to figure, fix #198

@@
 Modify the tables to make it so in the end, \url{https://github.com/richelbilderbeek/bbbq_article/issues/181}}
-\richel{TODO: Condense figures into bigger ones with panels, following 1 message per figure. Order these figures chronologically,
-\url{https://github.com/richelbilderbeek/bbbq_article/issues/195}}
 
@@
 will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
-% Do noty give the number of MHC molecules, as is is more complex, 
+% Do not give the number of MHC molecules, as is is more complex, 
 % as the alpha chains coded by the maternal chromosome 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methods}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Peptide elution studies}\label{subsec:elution_studies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Predicting TMH epitopes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection{Evolutionary conservation of TMHs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{TMH-derived peptides are predicted to be over-presented in MHC-I}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%\begin{figure}[!htbp]
-%  \includegraphics[height=0.9\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
-%  \caption{
-%    Percentage of MHC-I epitopes predicted to be overlapping with TMHs
-%    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
-%    The dashed horizontal lines indicate the percentage as expected by chance.
-%    See supplementary table \ref{tab:tmh_binders_mhc1} for the exact TMH and epitope counts.
-%    \richel{Add panel with hydrophobicity}
-%  }
-%  \label{fig:1}
-%\end{figure}
-%
-%\begin{figure}[!htbp]
-%  \includegraphics[height=0.8\textheight]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
-%  \caption{
-%    Percentage of MHC-II epitopes predicted to be overlapping with TMHs
-%    for a human, viral and bacterial proteome.
-%    The dashed horizontal lines indicate the percentage as expected by chance.
-%    See supplementary table \ref{tab:tmh_binders_mhc2} for the exact TMH and epitope counts.
-%  }
-%  \label{fig:2}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{TMH-derived peptides are predicted to be over-presented in MHC-II}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{TMH-derived peptides are predicted to be presented in vivo}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\richel{
-  Use this section as an experimental confirmation 
-  of the bioinformatics predictions
-}
-
-% tab:elution
-\input{bbbq_article_issue_157/results.tex}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
+  \caption{
+    TMH-derived epitopes are presented in vivo,
+    for both MHC-I and MHC-II, for two topology prediction
+    programs.
+    See table \ref{tab:elution} for the exact values.
+  }
+  \label{fig:elution}
+\end{figure}
 
@@
 peptide elution studies were reanalyzed.
-Table \ref{tab:elution} shows the percentages of epitopes derived 
+Figure \ref{fig:elution} shows the percentages of epitopes derived 
 from TMHs
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Elution studies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% tab:elution
+\input{bbbq_article_issue_157/results.tex}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-03-29 bbbq_article.tex
Fix spelling errors

@@
 in a human, viral and bacterial proteome, and for most haplotypes of both MHC-I and II.
-We confirmed this over-presentation of TMH-derived peptides by re-analyzis of a peptide elution study.
+We confirmed this over-presentation of TMH-derived peptides by re-analysis of a peptide elution study.
 Second, we tested our hypothesis that TMHs
@@
 The extent of the over-presentation between the various different haplotypes 
-is similar for the probed proteomes, which strenghtens our previous conclusion (\cite{bianchi2017}) that the hydrophopbicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
+is similar for the probed proteomes, which strengthens our previous conclusion (\cite{bianchi2017}) that the hydrophobicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
 
@@
 
-We next wondered if the over-representation of TMH-derived peptided would also be present for MHC-II. 
+We next wondered if the over-representation of TMH-derived peptides would also be present for MHC-II. 
 Figure \ref{fig:2} shows the percentages of MHC-II epitopes predicted to be overlapping 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TMH-derived epitopes are predicted to be presentated in vivo}
+\subsection{TMH-derived epitopes are predicted to be presented in vivo}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \richel{Hier ben ik nu met verwerken...} 
-from the human and MtB proteome stem from TMHs. \frans{ik zoek nog naar associaties} 
+from the human and MTb proteome stem from TMHs. \frans{ik zoek nog naar associaties} 
 such as \verb;HLA-DQA1*0101/DQB1*0501;
@@
 which is in proteins that span the membrane at least twice.
-This may be explained additionally that multi-spoanners
-are not only likelier to have a catalytically active site,
+This may be explained additionally that multi-spanners
+are not only likelier to have a catalytic active site,
 also there are interactions between the TMHs.

### 2021-03-29 bbbq_article.tex

@@
 
-\begin{figure}
+\begin{figure}[!htbp]
   \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
-    \label{fig:1}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
-    \label{fig:2}
-  \end{subfigure}
-  \caption{
-    TMH-derived peptides are predicted to be over-presented in MHC-I (left
-    panel) and MHC-II (right panel).
-    Both panels show the percentage of epitopes predicted to be 
-    overlapping with TMHs
-    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
-    The dashed horizontal lines indicate the percentage as expected by chance.
+  \includegraphics[width=0.95\textwidth]{bbbq_1_smart_results/fig_f_tmh_2_panel}
+  \caption{
+    TMH-derived peptides are predicted to be over-presented in general,
+    which is for MHC-I (left column) and MHC-II (right column),
+    as well as for the proteomes of SARS-CoV-2 (top row), human (middle 
+    row) and \emph{M. tuberculosis} (bottom row).
+    The bars show the percentage of epitopes predicted to be 
+    overlapping with TMHs.
+    The red lines indicate the percentage as expected by chance.
     See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
@@
   }
-\end{figure}
+  \label{fig:bbbq_1_smart_results}
+\end{figure}
+
+%\begin{figure}
+%  \centering
+%  \begin{subfigure}[t]{0.45\textwidth}
+%    \centering
+%    \caption{}
+%    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
+%    \label{fig:1}
+%  \end{subfigure}
+%  \hfill
+%  \begin{subfigure}[t]{0.45\textwidth}
+%    \centering
+%    \caption{}
+%    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
+%    \label{fig:2}
+%  \end{subfigure}
+%  \caption{
+%    TMH-derived peptides are predicted to be over-presented in MHC-I (left
+%    panel) and MHC-II (right panel).
+%    Both panels show the percentage of epitopes predicted to be 
+%    overlapping with TMHs
+%    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
+%    The dashed horizontal lines indicate the percentage as expected by chance.
+%    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+%    for the exact TMH and  epitope counts.
+%    \richel{Add panel with hydrophobicity}
+%  }
+%\end{figure}
 

### 2021-03-29 bbbq_article.tex
Make graphs more concise, fix #197

@@
 \begin{abstract}
-
-\richel{TODO: Tables don't have a legend, but instead have a title and footnotes below the table.
-Modify the tables to make it so in the end, \url{https://github.com/richelbilderbeek/bbbq_article/issues/181}}
 
@@
 
-\richel{
-  First do the bioinformatics predictions,
-  then the experimental confirmation 
-}
-
-Figure \ref{fig:1} shows the predicted presentation of TMH-derived peptides in MHC-I,
+Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
 for a human, viral and bacterial proteome.
@@
 
-%\begin{figure}
-%  \centering
-%  \begin{subfigure}[t]{0.45\textwidth}
-%    \centering
-%    \caption{}
-%    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_grid.png}
-%    \label{fig:1}
-%  \end{subfigure}
-%  \hfill
-%  \begin{subfigure}[t]{0.45\textwidth}
-%    \centering
-%    \caption{}
-%    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_grid.png}
-%    \label{fig:2}
-%  \end{subfigure}
-%  \caption{
-%    TMH-derived peptides are predicted to be over-presented in MHC-I (left
-%    panel) and MHC-II (right panel).
-%    Both panels show the percentage of epitopes predicted to be 
-%    overlapping with TMHs
-%    for the proteomes of human, SARS-CoV-2 and \emph{M. tuberculosis}.
-%    The dashed horizontal lines indicate the percentage as expected by chance.
-%    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
-%    for the exact TMH and  epitope counts.
-%    \richel{Add panel with hydrophobicity}
-%  }
-%\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We next wondered if the over-representation of TMH-derived peptides would also be present for MHC-II. 
-Figure \ref{fig:2} shows the percentages of MHC-II epitopes predicted to be overlapping 
+Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes predicted to be overlapping 
 with TMHs for our human, viral and bacterial proteomes.
@@
 presented more often than expected by chance 
-alone (figures \ref{fig:1} and \ref{fig:2}).
+alone (figure \ref{fig:bbbq_1_smart_results}).
 This in silico finding is a general pattern 
@@
 and the percentage of binders that overlap with TMHs, as
-visualized by figures \ref{fig:1} and \ref{fig:2} respectively.
+visualized by figure \ref{fig:bbbq_1_smart_results}.
 

### 2021-03-30 bbbq_article.tex
Processed feedback Geert 2021-03-24, fix #196

@@
 
-\richel{I wonder if the order of the Methods, should match the improved
-order of the Results as well, so first the MHC-I and MHC-II presentation,
-followed by the experimental confirmation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Peptide elution studies}\label{subsec:elution_studies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-To determine if epitopes derived from TMHs are presented 
-we used two elution studies, which both have sequenced
-epitopes found in naturally occurring human immune cells,
-for MHC-I (\cite{schellens2015comprehensive}) 
-and MHC-II (\cite{bergseng2015different}).
-For each of the sequenced epitopes,
-we found its possible location(s) in 
-a human representative reference proteome,
-with UniProt ID UP000005640\_9606.
-For the epitopes that were present only a single time in the proteome,
-we predicted the topology of the proteins in which these epitopes are found.
-We used both TMHMM and PureseqTM to do these predictions. 
-From this topology, we could determine if the epitope
-overlapped with a transmembrane region.
-
-The full analysis can be found
-at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 that give rise to no or only very few MHC epitopes
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Peptide elution studies}\label{subsec:elution_studies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To determine if epitopes derived from TMHs are presented 
+we used two elution studies, which both have sequenced
+epitopes found in naturally occurring human immune cells,
+for MHC-I (\cite{schellens2015comprehensive}) 
+and MHC-II (\cite{bergseng2015different}).
+For each of the sequenced epitopes,
+we found its possible location(s) in 
+a human representative reference proteome,
+with UniProt ID UP000005640\_9606.
+For the epitopes that were present only a single time in the proteome,
+we predicted the topology of the proteins in which these epitopes are found.
+We used both TMHMM and PureseqTM to do these predictions. 
+From this topology, we could determine if the epitope
+overlapped with a transmembrane region.
+
+The full analysis can be found
+at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.
 
@@
   \end{subfigure}
+
+   \vfill
+
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
+    \label{fig:f_tmh_ncbi}
+  \end{subfigure}
+
   \caption{
@@
     transmembrane proteins.
+    \textbf{(C)}
+    Distribution of the percentages of TMH of TMPs.
   }
@@
   \centering
-  \begin{subfigure}[t]{0.3\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
     \centering
@@
   \hfill
-  \begin{subfigure}[t]{0.55\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
+    \label{fig:f_tmh_ncbi_per_spanner}
+  \end{subfigure}
+
+  \vfill
+
+  \begin{subfigure}[t]{\textwidth}
     \centering
@@
       \textbf{(B)} 
+      \richel{THIS IS A PLACEHOLDER}
+      Distribution of the proportion of amino acids residing
+      in the plasma membrane.
+      \textbf{(C)} 
       Percentage of SNPs found in TMPs predicted to have only a single
@@
 Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
-We performed a linear regression analysis on the TMPs
+Most our TMPs had less than half their residues located in the 
+plasma membrane.
+The proportion of SNPs found in TMHs varied from 
+none (i.e. all SNPs were in
+soluble regions) to all.
+To determine if it is likelier to find SNPs within either TMHs or
+soluble-facing domains, we performed a linear regression analysis,
 and added a 95\% confidence interval on this regression.
-Would the confidence interval of the regression line 
-overlap with the line of equality,
-this would indicate that SNPs occur just as likely within TMHs 
-and solvent-exposed protein domains (i.e., no higher conservation of TMHs). However, we observed that ...[GEERT give brief explanation in words of what is seen in the graph]
-This finding suggests that SNPs 
-occur more often in soluble domains than in TMHs.
-
-To statistically test whether SNPs occur 
-just as likely in soluble domains as
-in TMHs, we first calculated the number of SNPs that are expected to be
-found in TMHs of TMPs by chance ($E(n_{SNP\_TMH})= n_{SNP} \cdot f_{TMH} \approx 4,056$), by multiplying the number of SNPs predicted to be in TMHs ($n_{SNP} = 21,738$ ) with the fraction of the TMPs that are predicted TMH.
-The actual number of SNPs predicted to be located within TMHs
-is $n_{SNP\_TMH} = 3,903$, which is lower than the expectation
-of $E(n_{SNP\_TMH}) \approx 4,056$.
-
-To test whether this difference is significant,
-we used a binomial Poisson distribution, where the number of trails ($n$) equals the number of SNPs, which is 21,738. The probability of success
-for the $i$th TMP ($p\_i$), is the percentage of each TMP that is TMH.
-We calculated the probability of having 3,903 or less successes.
+This linear fit nearly goes through the origin and has a slope
+below the line of equality,
+which shows that less SNPs are found in TMHs than expected by chance.
+
+We determined the probability to find the observed amount
+of SNPs in TMHs by chance, i.e. when assuming SNPs occur 
+just as likely in soluble domains as in TMHs.
+We used a binomial Poisson distribution, 
+where the number of trails ($n$) equals the number of SNPs, 
+which is 21,208. 
+The probability of success for the $i$th TMP ($p\_i$), 
+is the percentage of residues within a TMH per TMP. 
+These probabilities are shown as a histogram 
+in figure \ref{fig:f_tmh_ncbi}). 
+The expected number of SNPs expected to be found in 
+TMHs by chance equals $\sum{p} \approx 4,141$.
+As we observed 3,903 SNPs in TMHs, 
+we calculated the probability of having that amount or less successes.
 We used the type I error cut-off value of $\alpha = 2.5\%$.
 The chance to find, within TMHs, this amount or less SNPs 
-equals [RJCB: recalculate]
-$1.6201 \cdot 10^{-9}$.
-
-Using a alpha value of $0.05$, as is common practice for an uninformed
-expectation, we conclude that there are significantly less SNPs
-in TMHs as expected by chance.
-See figure \ref{fig:f_tmh_ncbi} for the percentages of TMHs in the transmembrane proteins.[GEERT: staat deze figuur hier wel goed? Niet naar boven?] The effect size is that, 
-when randomly selecting SNPs,
-per 100 SNPs found in soluble protein domains, 
-one finds 91 SNPs in TMHs,
+equals $6.8208 \cdot 10^{-11}$.
+We determined the relevance of this finding, by
+calculating how much less SNPs are found in TMHs,
+when compared to soluble regions, which is the
+ratio between the number of SNPs found in TMHs
+versus the number of SNPs as expected by chance.
+In effect, per 1000 SNPs found in soluble protein domains, 
+one finds 918 SNPs in TMHs,
 as depicted in figure \ref{fig:conservation}.
-[GEERT: hieronder had je indentieke tekst, maar verwees naar andere figuur \ref{fig:conservation_per_spanner}.
-Ik denk dat je deze hieronder kan samenvoegen]
-
-We splitted this analysis based on the number of TMHs
+
+We split this analysis based on the number of TMHs
 of each TMP. 
@@
 thus might have additional structural constraints.
-As can be seen in figure \ref{fig:f_snps_found_and_expected_per_spanner}, 
-both single- and multi-spanners are evolutionarily conserved.[GEERT: maar een sterker dan de andere? Kun je die conclusie niet trekken?]
+From the split data, we did the same analysis:
+figure \ref{fig:f_snps_found_and_expected_per_spanner} 
+shows, the percentage a protein resides in the membrane versus the
+percentage of SNPs found in those TMH regions, per spanner.
+In both cases, a linear regression shows that less
+SNPs are found in TMHs, than expected by chance.
+
+Also per spanner, we determined the probability to find the 
+observed amount of SNPs in TMHs by chance.
+For single-spanners, we found 452 SNPs in TMH, where
+$\approx462$ were expected by chance. 
+The chance to observe this or a lower number by chance is 
+$0.319$. As this chance was higher than our $\alpha = 0.025$
+and we considered this no significant effect.
+For the single-spanners, we found 3,351 SNPs in TMH, where
+$\approx3678$ were expected by chance. 
+The chance to observe this or a lower number by chance is 
+$8.315841 \cdot 10^{-12}$, 
+which means we attribute the lower number of SNPs found
+to an effect, instead of luck.
+
+Also, per spanner type, 
+we determined the relevance of this finding, by
+calculating how much less SNPs are found in TMHs,
+when compared to soluble regions,
+as depicted in figure \ref{fig:conservation_per_spanner}.
+In effect, per 1000 SNPs found in soluble protein domains, 
+one finds 978 SNPs in TMHs 
+in single spanners (although this is not significant)
+and 911 SNPs in TMHs in multi-spanners.
 
@@
   beetje kort. Suggesties? Experimental approaches? Hoeft niet super
-  origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. 
-  Frans kan vast helpen met mooie zin
+  origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. Frans kan vast helpen met mooie zin
 }
@@
 Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
-are presented , we conclude that these are additionally
+are presented, we conclude that these are additionally
 presented more often than expected by chance 
@@
 % \paragraph{MHC-II is expected to present TMHs}
-\frans{we moeten iets schrijven over de verschillende haplotypen wat is the occurence van de haplotypen met andere worden hoe relevant zijn de haplotypen die veel TMH presenteren?}
+The results from this study support our hypothesis 
+that TMH-derived epitopes are over-presented
+in MHC-II, similarly to we previously showed for MHC-I (\cite{bianchi2017}). 
 The over-presentation of TMHs for most MHC-II haplotypes 
-is more pronounced than for MHC-I haplotypes,
-as up to a third of the predicted MHC-II binders 
-\richel{Hier ben ik nu met verwerken...} 
-from the human and MTb proteome stem from TMHs. \frans{ik zoek nog naar associaties} 
-such as \verb;HLA-DQA1*0101/DQB1*0501;
-
-This results from this study support our hypothesis that TMH-derived epitopes are over-presented
-in MHC-II, similarly to we previously showed for MHC-I [CITE BIANCHI].[GEERT: deel gewist, want was letterlijke herhaling van hierboven.
+appears to follow a similar pattern as MHC-I haplotypes,
+although MHC-II presents slightly more TMH-derived
+epitopes (see Supplementary Materials, figure \ref{fig:rel_presentation}
+for a direct comparison). 
+Unlike MHC-I, the human proteome over-presents TMH-derived
+epitopes for all MHC-II haplotypes that we have investigated.
+Possible there are MHC-II haplotypes that present less TMH-derived
+epitopes than expected by chance.
+The haplotype \verb;HLA-DQA1*0101/DQB1*0501; had the highest
+over-presentation of TMH-derived epitopes,
+which we suggest is the haplotype of choice 
+to investigate the intracellular mechanism of TMH-derived
+antigen presentation.
+
+\richel{Geert, ik kan je hulp goed gebruiken bij dit stukje}
+Presentation of antigens by MHC-II is important for the activation of
+B cells \richel{je bedoelt denk ik T cellen hier?}. 
+B cell produce antibodies that recognize only solvent-exposed protein regions,
+thus the B cell receptor is inherently unable to present hydrophobic regions.
+Our analysis suggests, however, that MHC-II
+over-presents TMH regions. A helper T cell thus needs an 
+\geert{
   Beter is zeggen dat (1) MHC-II presentatie belangrijk is voor B cell activatie.
@@
 (3) Dat onze analyse suggereert HMC-II presentatie voornamelijk andere (TMH of buried) regio's betreft.
-(4) Dat T helper cellen dus andere delen van antigene eiwitten herkennen (MHC-II gepresenteerd door de B cellen) dan de BCR bindt.]
+(4) Dat T helper cellen dus andere delen van antigene eiwitten herkennen (MHC-II gepresenteerd door de B cellen) dan de BCR bindt.
+}
 
@@
 
-\geert{
-  Also compare your findings with previous efforts 
-  to look at conservation of TMHs
-}
+There have been more efforts to look at the conservation of TMHs,
+using different methodologies.
+One such example is \ref{stevens2001substitution}, 
+in which aligned protein sequence data was used.
+Also this study found that TMHs are evolutionarily more conserved,
+as the mean amino acid substitution rate in TMHs is about ten
+percent lower,
+which is the same value as this study found.
+Another example is a study that estimates the conservation
+scores for TMH and soluble regions based on 
+evolutionary related protein alignments,
+and found that TMHs are more conserved, 
+with a conservation score that was 17\% higher in 
+TMHs (\ref{oberai2009structural}).
+Note that the last study also used human SNPs
+and found that mutations in TMH are likelier to cause
+a disease, again underlining the importance of TMHs.
 
@@
 
-Figure \ref{fig:f_tmh_ncbi} shows the distribution 
-of the percentages of TMH of transmembrane proteins.
-
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
-  \caption{
-    Distribution of the percentages of TMH of transmembrane proteins.
-    Dashed vertical line: mean \% TMH in all proteins.
-    Dotted vertical line: mean \% TMH in TMPs.
-  }
-  \label{fig:f_tmh_ncbi}
-\end{figure}
+%Figure \ref{fig:f_tmh_ncbi} shows the distribution 
+%of the percentages of TMH of transmembrane proteins.
+%
+%\begin{figure}[!htbp]
+%  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
+%  \caption{
+%    Distribution of the percentages of TMH of transmembrane proteins.
+%    Dashed vertical line: mean \% TMH in all proteins.
+%    Dotted vertical line: mean \% TMH in TMPs.
+%  }
+%  \label{fig:f_tmh_ncbi}
+%\end{figure}
 

### 2021-03-30 bbbq_article.tex
Fix refs

@@
 shows that MHC-II has a higher relative over-presentation of TMH-derived
-epitopes, when compared to MHC-I. \richel{HIERO}
+epitopes, when compared to MHC-I.
 The MHC-I molecule presents epitopes of shorter length than MHC-II.
@@
 using different methodologies.
-One such example is \ref{stevens2001substitution}, 
+One such example is \cite{stevens2001substitution}, 
 in which aligned protein sequence data was used.
@@
 with a conservation score that was 17\% higher in 
-TMHs (\ref{oberai2009structural}).
+TMHs (\cite{oberai2009structural}).
 Note that the last study also used human SNPs

### 2021-03-30 bbbq_article.tex

@@
 and bacterial (\emph{Mycobacterium tuberculosis} (MTb), UniProt ID UP000001584) 
-reference proteome. We used TMHMM  \richel{ref} to predict the topology 
+reference proteome. We used TMHMM (\cite{krogh2001predicting}) to predict the topology 
 of the proteins within these proteomes.
@@
 we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
-and \verb;MHCnuggets; \richel{ref} for MHC-II.
+and \verb;MHCnuggets; (\cite{shao2020high}) for MHC-II.
 The 13 MHC-I haplotypes used in this study are the same as 
@@
 matches the percentage as expected by chance.
-\richel{TODO: explain only the results found. Elaborate. Draw a conclusion}
+\richel{explain only the results found. Elaborate. Draw a conclusion}
 Additionally, we investigated if MHC-I and MHC-II are equally prone

### 2021-03-30 bbbq_article.tex
Updates from Overleaf

@@
 
-\richel{Geert, ik kan je hulp goed gebruiken bij dit stukje}
 Presentation of antigens by MHC-II is important for the activation of
-B cells \richel{je bedoelt denk ik T cellen hier?}. 
-B cell produce antibodies that recognize only solvent-exposed protein regions,
+B cells, because B cells present peptides in MHC-II to helper T cells.  
+Interestingly, the CB cell produce antibodies that recognize only solvent-exposed protein regions,
 thus the B cell receptor is inherently unable to present hydrophobic regions.

### 2021-03-30 bbbq_article.tex

@@
 
-Presentation of antigens by MHC-II is important for the activation of
-B cells, because B cells present peptides in MHC-II to helper T cells.  
-Interestingly, the CB cell produce antibodies that recognize only solvent-exposed protein regions,
-thus the B cell receptor is inherently unable to present hydrophobic regions.
-Our analysis suggests, however, that MHC-II
-over-presents TMH regions. A helper T cell thus needs an 
-\geert{
-  Beter is zeggen dat (1) MHC-II presentatie belangrijk is voor B cell activatie.
-(2) B cellen antilichamen produceren, en dan antilichamen solvent-exposed protein regions herkennen.
-(3) Dat onze analyse suggereert HMC-II presentatie voornamelijk andere (TMH of buried) regio's betreft.
-(4) Dat T helper cellen dus andere delen van antigene eiwitten herkennen (MHC-II gepresenteerd door de B cellen) dan de BCR bindt.
-}
+Our study also suggests that T cells will largely recognize different protein regions than B cells. Presentation of antigens by MHC-II is important for the activation of
+naive B cells by helper T cells.
+For this activation, B cells first ingest antigen that is bound to their B cell receptor, and subsequently present peptides derived from this antigen in MHC-II to helper T cells.
+Following their activation by the T cells, B cells mature into plasma cells and release antibodies which recognize the same part of the antigen as the original B cell receptor.
+B cell receptors and antibodies will thus recognize solvent-exposed regions of antigens that are accessible for binding to the B cell receptor.
+However, the results from our study predict that most MHC-II haplotypes present relatively hydrophobic peptides, which are less likely to be solvent exposed.
+It is unknown why B and T cells seem to predominantly recognize different protein regions, but one possibility might be that this lowers the chance of B cell mediated autoimmune diseases, because auto-reactive B and T cells recognizing different part of the same antigen would need to be present for breakage of B cell tolerance.
 

### 2021-03-31 bbbq_article.tex
Geert wrote MHC-II to B cell to T cell part, fix #202
Process feedback

@@
 
-For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I and was used in our earlier
-study (\cite{bianchi2017}). For MHC-II, 14-mers were used,  as these are the most frequently occurring
-epitope length (\cite{bergseng2015different}).
+For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I 
+and was used in our earlier study (\cite{bianchi2017}). 
+For MHC-II, 14-mers were used, 
+as these are the most frequently occurring epitope length (\cite{bergseng2015different}).
 We used a human (UniProt ID UP000005640\_9606), 
@@
 used in our previous study (\cite{bianchi2017}).
-The most frequent MHC-II haplotypes used  were selected, with a phenotypic frequency of at least 14\% in
-the human population (\cite{greenbaum2011functional}),
+For MHC-II, we selected haplotypes with a phenotypic frequency 
+of at least 14\% in the human population (\cite{greenbaum2011functional}),
 resulting in 21 MHC-II haplotypes.
@@
 
-Additionally, we investigated if MHC-I and MHC-II are equally prone
-to present TMH-derived epitopes. 
-\richel{Add exact method}
-See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.
-
 In our previous work, we found that the observed pattern can be
@@
 cleft (\cite{bianchi2017}), we validated.
-Here we do the same analysis for MHC-II \richel{Explain the analysis here}.
+Here we do the same analysis for MHC-II 
+\richel{Really need to do? If yes, explain the analysis here}.
 
@@
     for the exact TMH and  epitope counts.
-    \richel{Add panel with hydrophobicity}
+    \richel{Decide in meeting: add panel with hydrophobicity}
   }
@@
 \richel{
-  Showing how general our finding is (as it is explained by hydrophobicity)
+  Decide in meeting: 
+  showing how general our finding is (as it is explained by hydrophobicity)
   is important. See \url{https://github.com/richelbilderbeek/bbbq_article/issues/192}
@@
 epitopes are presented by each MHC-II haplotype is similar for the human, bacterial and viral proteome.
-
-Plotting the normalized values of presented TMH-derived epitopes (where
-$1.0$ denotes that the percentage of presented TMH-derived epitopes
-matches the values as expected by chance) in \ref{fig:rel_presentation}
-shows that MHC-II has a higher relative over-presentation of TMH-derived
-epitopes, when compared to MHC-I.
-The MHC-I molecule presents epitopes of shorter length than MHC-II.
-We expect this to be of no influence in the presentation level
-of TMH-derived epitopes, for each of the three proteomes.
-We verified this, by normalizing the percentage of TMH-derived epitopes found
-to the percentage of TMH-derived epitopes expected, 
-where a value of $1.0$ denotes that the percentage of TMH-derived epitopes found
-matches the percentage as expected by chance.
-\richel{explain only the results found. Elaborate. Draw a conclusion}
-Additionally, we investigated if MHC-I and MHC-II are equally prone
-to present TMH-derived epitopes. 
-\richel{Add exact method}
-See \ref{fig:rel_presentation} and \ref{fig:rel_presentation_per_haplotype}.
 
@@
 
-Our study also suggests that T cells will largely recognize different protein regions than B cells. Presentation of antigens by MHC-II is important for the activation of
-naive B cells by helper T cells.
-For this activation, B cells first ingest antigen that is bound to their B cell receptor, and subsequently present peptides derived from this antigen in MHC-II to helper T cells.
-Following their activation by the T cells, B cells mature into plasma cells and release antibodies which recognize the same part of the antigen as the original B cell receptor.
-B cell receptors and antibodies will thus recognize solvent-exposed regions of antigens that are accessible for binding to the B cell receptor.
-However, the results from our study predict that most MHC-II haplotypes present relatively hydrophobic peptides, which are less likely to be solvent exposed.
-It is unknown why B and T cells seem to predominantly recognize different protein regions, but one possibility might be that this lowers the chance of B cell mediated autoimmune diseases, because auto-reactive B and T cells recognizing different part of the same antigen would need to be present for breakage of B cell tolerance.
+Our study also suggests that T cells will largely recognize 
+different protein regions than B cells. 
+Presentation of antigens by MHC-II is important 
+for the activation of naive B cells by helper T cells. 
+For this activation, 
+B cells first ingest antigen that is bound to their B cell receptor, 
+and subsequently present peptides derived from this antigen in MHC-II 
+to helper T cells. 
+Following their activation by the T cells, 
+B cells mature into plasma cells 
+and release antibodies which recognize the same part of the antigen 
+as the original B cell receptor. 
+B cell receptors and antibodies will thus recognize solvent-exposed regions 
+of antigens that are accessible for binding to the B cell receptor. 
+However, the results from our study predict that most MHC-II haplotypes 
+present relatively hydrophobic peptides, 
+which are less likely to be solvent exposed. 
+It is unknown why B and T cells seem to predominantly recognize 
+different protein regions, but one possibility might be that 
+this lowers the chance of B cell mediated autoimmune diseases, 
+because auto-reactive B and T cells recognizing different part 
+of the same antigen would need to be present for breakage of B cell tolerance.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{General pattern in epitope presentation}
+% \paragraph{Conclusion}
 
@@
 as done in this study.
-'tmhmm' has been submitted to and is accepted by CRAN. \frans{wat is CRAN}
+'tmhmm' has been submitted to and is accepted 
+by the Comprehensive R Archive Network (CRAN).
 
@@
 To determine the additional over-presentation of TMH-derived epitopes 
-in MHC-II (as compared to MHC-I), we used the same normalized percentages
-and compared the averages between MHC-I and MHC-II.
-Figure \ref{fig:rel_presentation_per_haplotype} shows the normalized
-TMH-derived epitope presentation per haplotype.
+in MHC-II (as compared to MHC-I), we normalized the data to enable
+a side-by-side comparison. 
+The percentage of TMH-derived epitopes presented was normalized
+to the expected percentage of TMH-derived epitopes,
+where $1.0$ denotes that the percentage of presented TMH-derived epitopes
+matches the values as expected by chance.
+The normalized values per haplotype are shown 
+in figure \ref{fig:rel_presentation_per_haplotype}.
 To compare the TMH-derived over-presentation per MHC class,
-we plot the mean and standard error of the normalized values,
-grouped per haplotype in figure \ref{fig:rel_presentation}.
+we grouped the normalized values per haplotype, 
+and plot the mean and standard error, as shown in figure \ref{fig:rel_presentation}.
 

### 2021-03-31 bbbq_article.tex
Add plot of old study for comparison, progress #203
Add plot of old study for comparison, fix #203

@@
 % See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
-\richel{Show side-by-side-comparison in supplementary materials}
 We believe that our revised definition precludes bias of proteins 
-that give rise to no or only very few MHC epitopes
+that give rise to no or only very few MHC epitopes.
+To verify our results are similar,
+we compare the previous and the
+current approach side by side in the supplementary materials.
 
@@
 
+% \paragraph{Compare results in figure}
+
+To verify if the previous and the current method give rise to
+notable difference, we show a side-by-side comparison
+in figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
+The figures that haplotypes that over-present or under-present TMH-derived epitopes,
+do so in both studies. The extent to which TMH-derived eptitopes are
+presented, however, is more extreme in our current setup.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/fig_1a.png}
+    \label{fig:bianch_et_al_2017_1a}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1.png}
+    \label{fig:bilderbeek_et_al_2021_1a}
+  \end{subfigure}
+
+  \caption{
+    \textbf{(A)} 
+    Results for \cite{bianchi2017}. 
+    Red lines denotes the coincidence interval.
+    \textbf{(B)}
+    Results for this study.
+    Red line denotes the percentage as expected by chance.
+  }
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-03-31 bbbq_article.tex
Improve figures, progress #199
Figures have readable texts, fix #199

@@
     for the exact TMH and  epitope counts.
-    \richel{Decide in meeting: add panel with hydrophobicity}
+    \richel{Add panel with hydrophobicity?}
   }
@@
     \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi_per_spanner.png}
     \label{fig:f_tmh_ncbi_per_spanner}
@@
       \textbf{(B)} 
-      \richel{THIS IS A PLACEHOLDER}
       Distribution of the proportion of amino acids residing

### 2021-04-02 bbbq_article.tex
Spellcheck
Updates from Overleaf

@@
 In our previous work, we found that the observed pattern can be
-explained from the hydrophobicity of the MHC-I binding 
-cleft (\cite{bianchi2017}), we validated.
-Here we do the same analysis for MHC-II 
-\richel{Really need to do? If yes, explain the analysis here}.
+explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
+Here we do a similar analysis, by correlating the percentage of TMH-derived
+epitopes presented versus the mean hydrophobicity of all presented
+peptides. From these points, we derive a linear trendline and measure
+the $R^2$, where $R^2$ is the proportion of the correlation explained by
+the assumed linear relation.
 
@@
 for the TMH and epitope counts.
-We predict that TMH-derived peptides are also over-presented in most MHC-II haplotypes, and this over-presentation seems even more pronounced than for MHC-I. 
+We predict that TMH-derived peptides are also over-presented in most MHC-II haplotypes, 
+and this over-presentation seems even more pronounced than for MHC-I.
+
+This over-presentation can be explained by the hydrophobicity of the MHC-I binding cleft.
 \richel{
-  Decide in meeting: 
-  showing how general our finding is (as it is explained by hydrophobicity)
-  is important. See \url{https://github.com/richelbilderbeek/bbbq_article/issues/192}
+  Elaborate.
+  See \url{https://github.com/richelbilderbeek/bbbq_article/issues/192}
 }
@@
 The figures that haplotypes that over-present or under-present TMH-derived epitopes,
-do so in both studies. The extent to which TMH-derived eptitopes are
+do so in both studies. The extent to which TMH-derived epitopes are
 presented, however, is more extreme in our current setup.

### 2021-04-02 bbbq_article.tex
Redo figure 1D for our study, progress #192

@@
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=0.95\textwidth]{bbbq_1_smart_results/fig_f_tmh_2_panel}
-  \caption{
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.9\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_panel.png}
+    \label{fig:bbbq_1_smart_results}
+  \end{subfigure}
+
+  \vfill
+
+  \begin{subfigure}[t]{0.4\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
+    \label{fig:hydrophobicity_2}
+  \end{subfigure}  
+  \hfill
+  \begin{subfigure}[t]{0.4\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
+    \label{fig:hydrophobicity_2}
+  \end{subfigure}  
+  \caption{
+    \textbf{(A)} 
     TMH-derived peptides are predicted to be over-presented in general,
@@
     for the exact TMH and  epitope counts.
-    \richel{Add panel with hydrophobicity?}
-  }
-  \label{fig:bbbq_1_smart_results}
+    \textbf{(B)} and \textbf{(C)}:
+    The relation between the percentage of TMH-derived epitopes presented,
+    and the hydrophobicity score of binders for MHC-I haplotypes (panel B)
+    and MHC-II haplotypes (panel C). Trendline assumes a linear relationship.
+  }
 \end{figure}

### 2021-04-05 bbbq_article.tex
Review paper on 2021-04-05, progress #212

@@
 peptide fragments derived from human transmembrane helices (TMHs) 
-are predicted to be presented more often than expected
+are predicted to be presented more often than expected,
 based on their abundance in the proteome.
@@
 as it is predicted for diverse bacteria and viruses 
-and for both MHC-I and MHC-II.  However, the physiological reason of why membrane proteins might be 
+and for both MHC-I and MHC-II.  
+However, the physiological reason of why membrane proteins might be 
 over-presented is unclear
@@
 An important part of the acquired immune response, 
-that develops specialized and more effective recognition of pathogens than the innate immune response, 
+that develops specialized and more effective recognition of pathogens 
+than the innate immune response, 
 are T cells which recognize peptides derived from 
@@
 HLA (Human Leukocyte Antigens) genes.
-In humans, the peptide-binding groove of MHC-I is made by the alpha chain, and there are three forms of alpha chains called HLA-A, HLA-B and HLA-C. 
-For MHC-II, both the alpha and the beta chains contribute to the binding groove, and there are three major forms as well, called HLA-DR, HLA-DQ and HLA-DP.
+In humans, the peptide-binding groove of MHC-I is made by the alpha chain, 
+and there are three forms of alpha chains called HLA-A, HLA-B and HLA-C. 
+For MHC-II, both the alpha and the beta chains contribute to the binding groove, 
+and there are three major forms as well, called HLA-DR, HLA-DQ and HLA-DP.
 Each MHC complex can only bind a subset of all possible peptides.
 For example, HLA-A and HLA-B have no overlap in which
-peptides they bind (\cite{lund2004definition})
+peptides they bind (\cite{lund2004definition}).
 The HLA genes of humans are highly polymorphic, with hundreds 
@@
 
-While humans only express a limited number of MHC haplotypes, and therefore any individual's immune system detects only a fraction of all possible
-peptide fragments, the number of pathogenic peptides that can be detected is increased at the population level, because of the multiple and highly polymorphic MHC genes.
+While humans only express a limited number of MHC haplotypes, 
+and therefore any individual's immune system detects 
+only a fraction of all possible peptide fragments, 
+the number of pathogenic peptides that can be detected 
+is increased at the population level, 
+because of the multiple and highly polymorphic MHC genes.
 It is believed that this improves immunity at the population level, 
@@
 will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
-% Do not give the number of MHC molecules, as is is more complex, 
-% as the alpha chains coded by the maternal chromosome 
-% of HLA-DQ, DR, DP can heterodimerize with the beta chains coded 
-% by the paternal chromosome and vice versa
 
@@
 as this will for instance aid the design of vaccines. 
-These studies have led to the development 
-of  prediction algorithms 
-that allow for very reliable in silico predictions 
-of the binding affinities of peptides (\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
+These studies have led to the development of prediction algorithms 
+that allow for very reliable \emph{in silico} predictions 
+of the binding affinities of peptides
+(\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
 For example, \cite{tang2011genome} found that, 
@@
 than expected based on their abundance (\cite{bianchi2017}). 
-Moreover, we showed that some well-known immunodominant peptides stem from TMHs. This over-presentation is attributed to the fact 
+Moreover, we showed that some well-known immunodominant peptides stem from TMHs. 
+This over-presentation is attributed to the fact 
 that the peptide-binding groove of most MHC-I haplotypes 
 is relatively hydrophobic, 
-and therefore hydrophobic peptides have a higher affinity to bind
-than hydrophilic ones.
-
-Transmembrane helices (TMHs) are hydrophobic 
+and therefore hydrophobic peptides have a higher affinity 
+to bind than hydrophilic ones.
+
+TMHs are hydrophobic 
 as they need to span the hydrophobic lipid bilayer of cellular membranes.
-They consist of an alpha helix of on average 23 amino acids in length. TMHs can also be predicted from a protein sequence 
+They consist of an alpha helix of, on average, 23 amino acids in length. 
+TMHs can also be predicted from a protein sequence 
 with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}),
@@
 TMHs are common structures in the proteins of humans and microbes. 
-Different TMH prediction tools estimate
-that 15-39\% of the human proteome 
-contain at least one TMH (\cite{ahram2006estimation}).
+Different TMH prediction tools estimate that 15-39\% of all proteins 
+in the human proteome contain at least one TMH (\cite{ahram2006estimation}).
 However, the physiological reason why peptides derived from TMHs 
@@
 stemming from soluble protein regions is unknown. 
-We hypothesized is that the presentation of 
+We hypothesized that the presentation of 
 TMH residues is evolutionary selected for, 
@@
 Due to this requirement, 
-many of the amino acids genuinely present in TMH 
+many of the amino acids genuinely present in TMHs 
 are limited to the ones with hydrophobic side chains (\cite{jones1994model}).
@@
 This study has two objectives. 
-First, we aimed to generalize our findings by predicting the presentation of peptides from different kingdoms of life and for both MHC-I and -II.
-From these in silico predictions, we conclude that TMH-derived
+First, we aimed to generalize our findings by predicting the presentation of peptides 
+from different kingdoms of life and for both MHC-I and -II. 
+From these \emph{in silico} predictions, we conclude that TMH-derived
 epitopes are presented more often than expected by chance,
-in a human, viral and bacterial proteome, and for most haplotypes of both MHC-I and II.
-We confirmed this over-presentation of TMH-derived peptides by re-analysis of a peptide elution study.
+in a human, viral and bacterial proteome, 
+and for most haplotypes of both MHC-I and II. 
+We confirmed this over-presentation of TMH-derived peptides 
+by re-analysis of a peptide elution study. 
 Second, we tested our hypothesis that TMHs 
-are more evolutionarily conserved than  solvent-exposed protein regions. Our analysis of human
-single nucleotide polymorphisms (SNPs) showed that random point mutations are indeed less likely
-to occur within TMHs. These findings strengthen the emerging notion that TMHs are important for the T cell branch of the adaptive immune system, and hence are of  
+are more evolutionarily conserved than solvent-exposed protein regions. 
+Our analysis of human single nucleotide polymorphisms (SNPs) showed 
+that random point mutations are indeed less likely
+to occur within TMHs. 
+These findings strengthen the emerging notion 
+that TMHs are important for the T cell branch of the adaptive immune system, 
+and hence are of  
 overlooked importance in vaccine development.
@@
 To predict how frequently epitopes overlapping with TMHs are presented,
-we applied a similar analysis strategy as described in \cite{bianchi2017} for several haplotypes of both MHC-I and MHC-II, 
+we applied a similar analysis strategy as described in \cite{bianchi2017} 
+for several haplotypes of both MHC-I and MHC-II, 
 as well as for a human, viral and bacterial proteome.
@@
 and bacterial (\emph{Mycobacterium tuberculosis} (MTb), UniProt ID UP000001584) 
-reference proteome. We used TMHMM (\cite{krogh2001predicting}) to predict the topology 
+reference proteome. 
+We used TMHMM (\cite{krogh2001predicting}) to predict the topology 
 of the proteins within these proteomes.
@@
 of TMH-derived epitopes, for each of the three proteomes.
-We verified this, by normalizing the percentage of TMH-derived epitopes found
-to the percentage of TMH-derived epitopes expected, 
-where a value of $1.0$ denotes that the percentage of TMH-derived epitopes found
-matches the percentage as expected by chance.
+We verified that MHC-II is just as prone to over-present TMH-derived
+epitopes by comparing the relative over-presentation of these epitopes.
 
@@
 peptides. From these points, we derive a linear trendline and measure
-the $R^2$, where $R^2$ is the proportion of the correlation explained by
-the assumed linear relation.
+the proportion of the correlation explained by the assumed linear relation.
 
@@
 any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
-the epitopes within a 
-\emph{proteome} (see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2}
-for values), whereas the original study defined
+the epitopes within a \emph{proteome} 
+(see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
+whereas the original study defined
 a binder as having an IC50 in the lowest 2\% 
@@
 
-To determine if epitopes derived from TMHs are presented 
-we used two elution studies, which both have sequenced
-epitopes found in naturally occurring human immune cells,
+To determine if epitopes derived from TMHs are presented \emph{in vivo}
+we used two elution studies, both of which have sequenced 
+the epitopes found in naturally occurring human immune cells,
 for MHC-I (\cite{schellens2015comprehensive}) 
@@
 with UniProt ID UP000005640\_9606.
-For the epitopes that were present only a single time in the proteome,
+For the epitopes that were present in the proteome exectly once,
 we predicted the topology of the proteins in which these epitopes are found.
-We used both TMHMM and PureseqTM to do these predictions. 
+We used both TMHMM (\cite{krogh2001predicting}) and PureseqTM (\cite{wang2019efficient}) to do these predictions. 
 From this topology, we could determine if the epitope
@@
 The first query was a call to the 'gene' database for the 
-term 'membrane protein' (in all fields) for the organism \emph{Homo sapiens}.
-This resulted in 1,077 gene IDs.
+term 'membrane protein' (in all fields) 
+for the organism \emph{Homo sapiens}.
+This resulted in 1,077 gene IDs (on December 2020).
 % 1077 gene IDs is correct for December 2020.
@@
 As the NCBI API constrains its users to three calls per second
-to assure fair use, we had to limit the extent of our analysis.
-
-The number of SNPs
-was limited to the first 250 variations per gene (see figure
-\ref{fig:snps_per_gene_name_ncbi} for the number of SNPs
-within the dbSNP database for these gene names),
-resulting in 60,683 variations at the protein level.
+(to assure fair use), we had to limit the extent of our analysis.
+
+The number of SNPs was limited to the first 250 variations per gene,
+resulting in $\approx$61k variations.
 Only variations that result in a SNP for
-a single amino acid substitution were analyzed, resulting in 38,882 SNPs.
+a single amino acid substitution were analyzed, resulting in $\approx$38k SNPs.
+The exact amounts can be found in the supplementary materials,
+tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
 
@@
 using PureseqTM.
-Using these predicted protein topologies, the SNPs were scored to be located within or outside trans membrane domains.
+Using these predicted protein topologies, 
+the SNPs were scored to be located within or outside trans membrane domains.
 
@@
 for at least one residue.
-A dashed line indicates the expected percentage of TMH-derived epitopes 
+A horizontal line shows the expected percentage of TMH-derived epitopes 
 that would be presented, if TMH-derived epitopes would be presented just as 
 likely as soluble regions.
-In 12 out of 13 MHC-I haplotypes, TMH-derived epitopes are presented more often 
-then the null expectation, for a human proteome. 
-For the viral and bacterial proteome, 11 out of 13 haplotypes present
+In 11 out of 13 MHC-I haplotypes, TMH-derived epitopes are presented more often 
+then the null expectation, for a human and bacterial proteome.
+For the viral proteome, 12 out of 13 haplotypes present
 TMH-derived epitopes more often than expected by chance.
 The extent of the over-presentation between the various different haplotypes 
-is similar for the probed proteomes, which strengthens our previous conclusion (\cite{bianchi2017}) that the hydrophobicity of the MHC-binding groove is the main factor responsible for the predicted over-presentation of TMH-derived peptides.
+is similar for the probed proteomes, 
+which strengthens our previous conclusion (\cite{bianchi2017}) 
+that the hydrophobicity of the MHC-binding groove 
+is the main factor responsible for the predicted over-presentation 
+of TMH-derived peptides.
 
@@
     \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
-    \label{fig:hydrophobicity_2}
+    \label{fig:hydrophobicity_1}
   \end{subfigure}  
@@
     and the hydrophobicity score of binders for MHC-I haplotypes (panel B)
-    and MHC-II haplotypes (panel C). Trendline assumes a linear relationship.
+    and MHC-II haplotypes (panel C). The trendline assumes a linear relationship.
   }
@@
 
-We next wondered if the over-representation of TMH-derived peptides would also be present for MHC-II. 
-Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes predicted to be overlapping 
-with TMHs for our human, viral and bacterial proteomes.
+We next wondered if the over-representation of TMH-derived peptides 
+would also be present for MHC-II. 
+Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
+predicted to be overlapping with TMHs for our human, viral and bacterial proteomes.
+We found that TMH-derived peptides are over-presented in all
+of the 21 MHC-II haplotypes, 
+in a human, bacterial and viral proteome,
+except for \verb;HLA-DRB3*0101; in \emph{MTb}
 See supplementary table \ref{tab:tmh_binders_mhc2} 
-for the TMH and epitope counts.
-We predict that TMH-derived peptides are also over-presented in most MHC-II haplotypes, 
-and this over-presentation seems even more pronounced than for MHC-I.
-
-This over-presentation can be explained by the hydrophobicity of the MHC-I binding cleft.
-\richel{
-  Elaborate.
-  See \url{https://github.com/richelbilderbeek/bbbq_article/issues/192}
-}
-Similar to our previous results with MHC-I, we predicted that the extent to which TMH-derived
-epitopes are presented by each MHC-II haplotype is similar for the human, bacterial and viral proteome.
+for the exact TMH and epitope counts.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Hydrophocity explains the generality of this finding}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Hydrophobicity may be the common reason of the consistent over-presentation
+of  (hydrophobic) TMH-derived epitopes.
+Figures \label{fig:hydrophobicity_1} and \label{fig:hydrophobicity_2}
+show the hydrophobicity preference score for the extent over-presentation
+of TMH-derived epitopes for the different haplotypes in a human proteome.
+An assumed linear correlation explains 88\% of the variability in MHC-I.
+For MHC-II, 62\% of the variability is explained by hydrophobicity.
 
@@
   \caption{
-    TMH-derived epitopes are presented in vivo,
+    TMH-derived epitopes are presented \emph{in vivo},
     for both MHC-I and MHC-II, for two topology prediction
     programs.
-    See table \ref{tab:elution} for the exact values.
   }
@@
 
-To obtain experimental confirmation that peptides stemming from TMHs are presented in MHC-I and MHC-II,
-peptide elution studies were reanalyzed.
+To obtain experimental confirmation that peptides stemming from TMHs 
+are presented in MHC-I and MHC-II,
+two peptide elution studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} for MHC-I and MHC-II respectively) 
+were reanalyzed.
 Figure \ref{fig:elution} shows the percentages of epitopes derived 
 from TMHs
-found in the MHC-I (\cite{schellens2015comprehensive}) and MHC-II elution 
-studies  (\cite{bergseng2015different}),
+found in the MHC-I and MHC-II elution 
+studies,
 for the two topology prediction tools TMHMM (\cite{krogh2001predicting}) and PureseqTM (\cite{wang2019efficient}). 
@@
 epitopes derived from TMHs are presented in both MHC-I and MHC-II.
+See the supplementary materials (table \ref{tab:elution}) for the exact values.
 
@@
     equal conservation of TMHs and soluble protein regions).
-    The blue line is a linear trend line that includes only the
-    transmembrane proteins.
     \textbf{(C)}
-    Distribution of the percentages of TMH of TMPs.
+    Distribution of the percentages of TMH in the TMPs used in this study.
   }
@@
 We addressed the question whether there is an evolutionary advantage in presenting TMHs. 
-We determined the conservation of TMHs by comparing the occurrences of SNPs located in TMHs or solvent-exposed protein regions for the genes coding for membrane proteins.
-We obtained $>$1,000 gene names associated with the phrase 'membrane protein',
-which are genes for both membrane-associated proteins (MAPs, no TMH) and 
-transmembrane proteins (TMPs, at least one TMH).
+We determined the conservation of TMHs 
+by comparing the occurrences of SNPs located in TMHs or solvent-exposed protein regions 
+for the genes coding for membrane proteins.
+We obtained 911 unique gene names associated with the phrase 'membrane protein',
+which are genes for both membrane-associated proteins (MAPs, which have no TMH) and 
+transmembrane proteins (TMPs, which have at least one TMH).
 These genes are linked to 4,780 protein isoforms, 
@@
 2,237 proteins are predicted to be MAPs.
-We obtained 37,630 SNPs that resulted in an
-amino acids substitution, of which 21,024 were located 
-in predicted TMPs.
+We obtained 37,630 unique variations, 
+of which 9,621 are SNPs that resulted in a straightforward amino acids substitution, 
+of which 6,062 were located in predicted TMPs.
 See tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
@@
 TMHs by chance equals $\sum{p} \approx 4,141$.
-As we observed 3,903 SNPs in TMHs, 
+As we observed 3,803 SNPs in TMHs, 
 we calculated the probability of having that amount or less successes.
@@
 of each TMP. 
-We hypothesized that single-membrane spanners (i.e., proteins
-with a single predicted TMH) are less conserved than multi-membrane spanners (with $>$1 TMH),
+We hypothesized that single-membrane spanners 
+(i.e., proteins with a single predicted TMH) 
+are less conserved than multi-membrane spanners (with $>$1 TMH),
 because multi-membrane spanners
@@
 
-Epitope prediction is 
-important to understand the immune system and for the design of vaccines. In this study, we provide evidence that epitopes
-derived from TMHs are a major but overlooked source of MHC epitopes. Both our bioinformatics predictions and reanalysis of peptide elution studies show that TMH-derived epitopes are 
-presented more often than expected by chance, regardless of the organism. Moreover, our SNP analysis shows  
-that TMHs are evolutionary more conserved than solvent-exposed protein regions.
+Epitope prediction is important to understand the immune system 
+and for the design of vaccines. 
+In this study, we provide evidence that epitopes
+derived from TMHs are a major but overlooked source of MHC epitopes. 
+Both our bioinformatics predictions and 
+reanalysis of peptide elution studies show that TMH-derived epitopes 
+are presented more often than expected by chance, 
+regardless of the organism. 
+Moreover, our SNP analysis shows that TMHs are evolutionary more conserved 
+than solvent-exposed protein regions.
 
@@
 We show that 1.4\% of epitopes that are naturally presented on MHC-I 
-can be mapped to stem from a TMH. We also find that MHC-II presents epitopes derived from TMHs in vivo, as we find that 3.9-4.0\% of sequenced epitopes
-can be uniquely traced onto a human reference 
-proteome (also shown in table \ref{tab:elution}).
+can be mapped to stem from a TMH. 
+We also find that MHC-II presents epitopes derived from TMHs \emph{in vivo}, 
+as we find that 3.9-4.0\% of sequenced epitopes
+can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}).
 
@@
 
-Although elution studies and in silico predictions both suggest that TMH-derived epitopes are presented in MHC-I and MHC-II, the molecular mechanisms of how integral membrane proteins are processed for MHC presentation are largely unknown (\cite{bianchi2017}).
-Most prominently, the fundamental principles of how TMHs are extracted from the  hydrophobic lipid environment into the aqueous vacuolar lumen, and their prior or subsequent proteolytic processing are unresolved.
+Although elution studies and \emph{in silico} predictions both suggest that 
+TMH-derived epitopes are presented in MHC-I and MHC-II, 
+the molecular mechanisms of how integral membrane proteins are processed 
+for MHC presentation are largely unknown (\cite{bianchi2017}). 
+Most prominently, the fundamental principles of 
+how TMHs are extracted from the hydrophobic lipid environment 
+into the aqueous vacuolar lumen, 
+and their prior or subsequent proteolytic processing are unresolved.
 \geert{
@@
 in which the ribosome terminates its translation prematurely,
-intrudes a point mutation, 
-or cause a frame shift, at low frequencies (\cite{yewdell2011drips}).
+or cause a point mutation or frame shift (\cite{yewdell2011drips}).
 Another process that results in peptide fragments absent in the
@@
 epitope to a specific proteome locations, 
-where in reality, it was created de novo.
+where in reality, it was created \emph{de novo}.
 This may lead to false positives, as we incorrectly deduce
@@
 
-A final caveat in the elution studies is that there is methodological bias caused by mass spectronomy,
+A final caveat in the elution studies is that 
+there is methodological bias caused by mass spectronomy,
 as hydrophobic fragments are detected less 
-by mass spectronomy (\cite{bianchi2017}). This means that the number of TMH-derived epitopes
+by mass spectronomy (\cite{bianchi2017}). 
+This means that the number of TMH-derived epitopes
 might be underestimated.
@@
 alone (figure \ref{fig:bbbq_1_smart_results}).
-This in silico finding is a general pattern 
+This \emph{in silico} finding is a general pattern 
 that persists in both MHC-I and MHC-II, 
@@
 that TMH-derived epitopes are over-presented
-in MHC-II, similarly to we previously showed for MHC-I (\cite{bianchi2017}). 
+in MHC-II, similar to what we previously showed for MHC-I (\cite{bianchi2017}). 
 The over-presentation of TMHs for most MHC-II haplotypes 
@@
 present relatively hydrophobic peptides, 
-which are less likely to be solvent exposed. 
+which are less likely to be solvent-exposed. 
 It is unknown why B and T cells seem to predominantly recognize 
@@
 an immune system that as most attentive for protein regions that are
-essential for the survival, proliferation and/or virulence or pathogenic microbes, as these will be most conserved.
+essential for the survival, proliferation and/or virulence or pathogenic microbes, 
+as these will be most conserved.
 In SARS-CoV-2, for example, there is preliminary evidence that the strongest
@@
 Because of this scarcity and variance in targets, 
-one can imagine that it will be mostly unfeasible to provide innate immune responses against such essential protein regions, 
+one can imagine that it will be mostly unfeasible 
+to provide innate immune responses against such rare essential protein regions, 
 as hinted upon by a study on influenza (\cite{han2019individual}),
@@
 high performance computing cluster. 
-GvdB is funded by a Young Investigator Grant from the Human Frontier Science Program (HFSP; RGY0080/2018), and a Vidi grant from the Netherlands Organization for Scientific Research (NWO-ALW VIDI 864.14.001).
-GvdB has received funding from the European Research Council (ERC) under the European Union’s Horizon 2020 research and innovation programme (grant agreement No.
-862137.
+GvdB is funded by a Young Investigator Grant from 
+the Human Frontier Science Program (HFSP; RGY0080/2018), 
+and a Vidi grant from 
+the Netherlands Organization for Scientific Research (NWO-ALW VIDI 864.14.001). 
+GvdB has received funding from the European Research Council (ERC) 
+under the European Union’s Horizon 2020 research and 
+innovation programme (grant agreement No. 862137.
 
@@
 
-RJCB and FB conceived the idea for this research. MVB helped with the proteome analysis of \emph{M. tuberculosis}.
+RJCB and FB conceived the idea for this research. 
+MVB helped with the proteome analysis of \emph{M. tuberculosis}.
 RJCB wrote the code.

### 2021-04-05 bbbq_article.tex
Review paper on 2021-04-05, progress #212

@@
 for an overview of all amounts.
-In table \ref{tab:ncbi_counts_1} one expects [RJCB: name these]
-numbers to add up. 
-The number of unique SNPs and unique gene names
-does not add up for MAP and TMP,
-as one SNP may work on multiple isoforms, some of which can be MAP
-where others can be TMP.
-The number of unique SNPs and unique gene names
-does not add up for TMPs in TMH and TMP in soluble regions
-as one SNP may work on multiple isoforms, some of which can be MAP
-where others can be TMP, and some SNPs fall in a TMH, where others
-are found in soluble regions
-In table \ref{tab:ncbi_counts_2} one expects [RJCB: name these]
-numbers to add up.
+In table \ref{tab:ncbi_counts_1} there are multiple instances where
+the amounts are expected to add up, yet don't, as one SNP can work on
+multiple isoforms. For example, there are 9,621 unique SNPs 
+found in all proteins, of which 4,219 around found in MAPs 
+and 6,026 in TMPs. Apparently, 624 SNPs work on a set of isoforms that
+contains both MAPs and TMPs.
 

### 2021-04-05 bbbq_article.tex
Spellcheck, progress #212

@@
 epitopes presented versus the mean hydrophobicity of all presented
-peptides. From these points, we derive a linear trendline and measure
+peptides. From these points, we derive a linear trend line and measure
 the proportion of the correlation explained by the assumed linear relation.
@@
 with UniProt ID UP000005640\_9606.
-For the epitopes that were present in the proteome exectly once,
+For the epitopes that were present in the proteome exactly once,
 we predicted the topology of the proteins in which these epitopes are found.
@@
     and the hydrophobicity score of binders for MHC-I haplotypes (panel B)
-    and MHC-II haplotypes (panel C). The trendline assumes a linear relationship.
+    and MHC-II haplotypes (panel C). The trend line assumes a linear relationship.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydrophocity explains the generality of this finding}
+\subsection{Hydrophobicity explains the generality of this finding}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-04-12 bbbq_article.tex
Updates from Overleaf
Use shorthand names for haplotypes

--- 2021-04-05 bbbq_article.tex
@@
 
-Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
-Because the peptide-binding grooves of most haplotypes 
-of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
-peptide fragments derived from human transmembrane helices (TMHs) 
-are predicted to be presented more often than expected,
-based on their abundance in the proteome.
-We show that this over-presentation of TMH-derived peptides is general, 
-as it is predicted for diverse bacteria and viruses 
+Cytolytic T cell responses are predicted to be biased towards membrane proteins. The peptide-binding grooves of most haplotypes 
+of histocompatibility complex class I (MHC-I) are relatively hydrophobic,  therefor peptide fragments derived from human transmembrane helices (TMHs) are predicted to be presented more often as would be expected
+based on their abundance in the proteome. However, the physiological reason of why membrane proteins might be over-presented is unclear. 
+In this study, we show that the over-presentation of TMH-derived peptides is general, as it is predicted for bacteria and viruses 
 and for both MHC-I and MHC-II.
-However, the physiological reason of why membrane proteins might be 
-over-presented is unclear
-In this study, we show that TMHs are evolutionarily more conserved, 
-as single nucleotide polymorphisms (SNPs) 
-are present relatively less frequent in TMH-coding chromosomal regions 
-compared to regions coding for extracellular and cytosolic protein regions. 
-Thus, our findings suggest that T cells might respond 
-more to membrane proteins, because these are evolutionary more conserved.
-We speculate that TMHs therefor might be less prone to escape mutations 
-that enable pathogens to evade T cell responses.
+Moreover, we show that TMHs are evolutionarily more conserved, because single nucleotide polymorphisms (SNPs) are present relatively less frequently in TMH-coding chromosomal regions compared to regions coding for extracellular and cytoplasmic protein regions.
+Thus, our findings suggest that both cytolytic and helper T cells respond more to membrane proteins, because these are evolutionary more conserved.
+We speculate that TMHs therefor are less prone to escape mutations that enable pathogens to evade T cell responses.
 
@@
 An important part of the acquired immune response, 
-that develops specialized and more effective recognition of pathogens 
+that develops specialized and more specific recognition of pathogens 
 than the innate immune response, 
-are T cells which recognize peptides derived from 
-exogenous antigens presented on Major Histocompatibility Complexes (MHC) class I and II.
+are T cells which recognize peptides, called epitopes, derived from antigenic proteins presented on Major Histocompatibility Complexes (MHC) class I and II on the cell surface.
 
@@
 
-The MHC proteins are heterodimeric proteins encoded by the
+The MHC proteins are heterodimeric complexes encoded by the
 HLA (Human Leukocyte Antigens) genes.
-In humans, the peptide-binding groove of MHC-I is made by the alpha chain, 
-and there are three forms of alpha chains called HLA-A, HLA-B and HLA-C. 
-For MHC-II, both the alpha and the beta chains contribute to the binding groove, 
-and there are three major forms as well, called HLA-DR, HLA-DQ and HLA-DP.
-Each MHC complex can only bind a subset of all possible peptides.
+In humans, the peptide binding groove of MHC-I is made by only the alpha subunit.
+There are three classical forms of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells.
+For MHC-II, both the alpha and the beta chains contribute to the peptide binding groove.
+There are three classical forms of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
+Each MHC complex can present a subset of all possible peptides.
 For example, HLA-A and HLA-B have no overlap in which
-peptides they bind (\cite{lund2004definition}).
-The HLA genes of humans are highly polymorphic, with hundreds 
+epitopes they bind (\cite{lund2004definition}).
+Moreover, the HLA genes of humans are highly polymorphic, with hundreds 
 to thousands of different alleles, 
 and each different HLA allele is called 
-an MHC haplotype (\cite{marsh2010nomenclature}).
+an MHC haplotype and presents a different subset of peptides (\cite{marsh2010nomenclature}).
 
@@
 
-While humans only express a limited number of MHC haplotypes, 
-and therefore any individual's immune system detects 
-only a fraction of all possible peptide fragments, 
-the number of pathogenic peptides that can be detected 
-is increased at the population level, 
-because of the multiple and highly polymorphic MHC genes.
-It is believed that this improves immunity at the population level, 
-as mutations in a protein that disrupt a particular MHC presentation, 
+Humans mostly express two haplotypes per MHC form, one from the parental and one from the maternal chromosome, and therefore an individual's immune system detects 
+only a fraction of all possible peptide fragments. 
+However, at the population level, the coverage of pathogenic peptides that are detected 
+is very high, because of the highly polymorphic MHC genes.
+It is therefor believed that MHC polymorphism improves immunity at the population level, 
+as mutations in a protein that disrupt a particular MHC presentation at the individual level, 
 so-called escape mutations, 
-will not affect MHC presentation for all haplotypes (\cite{sommer2005importance}).
+will not affect MHC presentation for all haplotypes present in the population  (\cite{sommer2005importance}).
 
@@
 
-Many studies are aimed at determining which peptides are presented in MHC 
-and will result in an immune response, 
+Many studies are aimed at identifying the repertoire of epitopes that are presented in any MHC haplotype and determining which epitopes will result in an immune response, 
 as this will for instance aid the design of vaccines. 
@@
 of the 432 peptides that were predicted to bind to MHC,
-86\% were confirmed to do so in vivo.
+86\% were experimentally confirmed to do so.
 
@@
 and therefore hydrophobic peptides have a higher affinity 
-to bind than hydrophilic ones.
+to bind than more hydrophilic peptides.
 
@@
 They consist of an alpha helix of, on average, 23 amino acids in length. 
-TMHs can also be predicted from a protein sequence 
-with high accuracy by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}),
+TMHs can also be predicted with high accuracy from a protein sequence 
+by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}),
 for example, \cite{jones2007improving} found that,
-from 184 TMPs with known topology, 
+from 184 transmembrane proteins (TMPs) with known topology, 
 80\% of the TMH predictions replicated this finding.
@@
 would be presented more often than peptides 
-stemming from soluble protein regions is unknown. 
+stemming from soluble (i.e., extracellular or cytoplasmic) protein regions is unknown. 
 We hypothesized that the presentation of 
 TMH residues is evolutionary selected for, 
-because TMHs are less prone to undergo escape mutation. 
+because TMHs are less prone to undergo escape mutations. 
 One reason to expect such a reduced 
@@
 
-This study has two objectives. 
+This study had two objectives. 
 First, we aimed to generalize our findings by predicting the presentation of peptides 
@@
 and for most haplotypes of both MHC-I and II. 
-We confirmed this over-presentation of TMH-derived peptides 
-by re-analysis of a peptide elution study. 
+We confirmed the presentation of TMH-derived peptides 
+by re-analysis of peptide elution studies. 
 Second, we tested our hypothesis that TMHs 
-are more evolutionarily conserved than solvent-exposed protein regions. 
+are more evolutionary conserved than soluble protein regions. 
 Our analysis of human single nucleotide polymorphisms (SNPs) showed 
@@
 To predict how frequently epitopes overlapping with TMHs are presented,
-we applied a similar analysis strategy as described in \cite{bianchi2017} 
+a similar analysis strategy was applied as described in \cite{bianchi2017} 
 for several haplotypes of both MHC-I and MHC-II, 
-as well as for a human, viral and bacterial proteome.
+and for a human, viral and bacterial proteome.
 To summarize, for each proteome, 
@@
 
-For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I 
-and was used in our earlier study (\cite{bianchi2017}). 
+For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I and was used in our earlier study (\cite{bianchi2017}). 
 For MHC-II, 14-mers were used, 
 as these are the most frequently occurring epitope length (\cite{bergseng2015different}).
-We used a human (UniProt ID UP000005640\_9606), 
+A human (UniProt ID UP000005640\_9606), 
 viral (SARS-CoV-2, UniProt ID UP000464024) 
-and bacterial (\emph{Mycobacterium tuberculosis} (MTb), UniProt ID UP000001584) 
-reference proteome. 
-We used TMHMM (\cite{krogh2001predicting}) to predict the topology 
+and bacterial (\emph{Mycobacterium tuberculosis}, UniProt ID UP000001584) 
+reference proteome was used. 
+\verb;TMHMM; (\cite{krogh2001predicting}) was used to predict the topology 
 of the proteins within these proteomes.
-To predict the affinity of an epitope to a certain MHC molecule,
-we used \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
-and \verb;MHCnuggets; (\cite{shao2020high}) for MHC-II.
+To predict the affinity of an epitope to a certain MHC haplotype,
+ \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
+and \verb;MHCnuggets; (\cite{shao2020high}) for MHC-II was used.
 The 13 MHC-I haplotypes used in this study are the same as 
-used in our previous study (\cite{bianchi2017}).
-For MHC-II, we selected haplotypes with a phenotypic frequency 
+used in the previous study (\cite{bianchi2017}).
+For MHC-II, haplotypes were selected with a phenotypic frequency 
 of at least 14\% in the human population (\cite{greenbaum2011functional}),
@@
 
-The MHC-I molecule presents epitopes of shorter length than MHC-II.
-We expect this to be of no influence in the presentation level
-of TMH-derived epitopes, for each of the three proteomes.
-We verified that MHC-II is just as prone to over-present TMH-derived
-epitopes by comparing the relative over-presentation of these epitopes.
-
-In our previous work, we found that the observed pattern can be
+\geert{wat text gewist want dit is de methods sectie}
+In previous work, it was found that the over-presentation of TMH-derived peptides can be
 explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
-Here we do a similar analysis, by correlating the percentage of TMH-derived
-epitopes presented versus the mean hydrophobicity of all presented
-peptides. From these points, we derive a linear trendline and measure
-the proportion of the correlation explained by the assumed linear relation.
-
-This study differs in one important aspect from our previous work (\cite{bianchi2017}), 
+Here, a similar analysis was applied, by correlating the percentage of predicted TMH-derived epitopes versus the mean hydrophobicity of all 
+peptides.
+
+This study differs in one important aspect from our previous work (\cite{bianchi2017}). 
 The definition of a binder differs from \cite{bianchi2017}:
-in the current study a peptide is called a binder if, within a certain haplotype, 
-any of its 9-mer peptides have an IC50 value in the lowest 2\% of 
-the epitopes within a \emph{proteome} 
+in the current study, a peptide is called a binder if, for a certain haplotype, 
+any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
+all peptides within a \emph{proteome} 
 (see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
-whereas the original study defined
+whereas the previous study defined
 a binder as having an IC50 in the lowest 2\% 
-of the peptides within a \emph{protein}, 
-which put less strain on the bioinformatics pipeline.
+of the peptides within a \emph{protein}.
 % See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
-We believe that our revised definition precludes bias of proteins 
+This revised definition precludes bias of proteins 
 that give rise to no or only very few MHC epitopes.
-To verify our results are similar,
-we compare the previous and the
-current approach side by side in the supplementary materials.
+To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials. \geert{refereer hier naar supplementary figuur}
 
@@
 
-To determine if epitopes derived from TMHs are presented \emph{in vivo}
-we used two elution studies, both of which have sequenced 
-the epitopes found in naturally occurring human immune cells,
-for MHC-I (\cite{schellens2015comprehensive}) 
-and MHC-II (\cite{bergseng2015different}).
-For each of the sequenced epitopes,
-we found its possible location(s) in 
-a human representative reference proteome,
-with UniProt ID UP000005640\_9606.
-For the epitopes that were present in the proteome exectly once,
-we predicted the topology of the proteins in which these epitopes are found.
-We used both TMHMM (\cite{krogh2001predicting}) and PureseqTM (\cite{wang2019efficient}) to do these predictions. 
-From this topology, we could determine if the epitope
-overlapped with a transmembrane region.
+To obtain experimental evidence that epitopes derived from TMHs are presented in MHC, peptide elution studies for MHC-I (\cite{schellens2015comprehensive}) 
+and MHC-II (\cite{bergseng2015different}) were reanalyzed.
+For each of the detected epitopes,
+its possible location(s) in 
+a human reference proteome,
+with UniProt ID UP000005640\_9606, was mapped.
+For the epitopes that were present in the proteome exactly once,
+the topology of the proteins in which these epitopes were located was predicted using both \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
+From this topology, we determined if the epitope
+overlapped with a TMH.
 
@@
 
-To detect the evolutionary conservation of TMHs,
-we collected human single nucleotide polymorphisms (SNPs)
+To determine the evolutionary conservation of TMHs,
+human single nucleotide polymorphisms (SNPs) were first collected
 that resulted in a single amino acid substitution,
-and determined if this substitution occurred within a predicted TMH or not.
+and we then determined if this substitution occurred within a predicted TMH or not.
 
@@
 As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: the 'dbSNP' (\cite{sherry2001dbsnp}) database,
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: the \emph{dbSNP} (\cite{sherry2001dbsnp}) database,
 which contains 650 million 
 catalogued non-redundant humane variations (called RefSNPs,
-\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases 'gene' (for gene names, \cite{brown2015gene})
-and 'protein' (for proteins sequences, \cite{sayers2010database}).
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names, \cite{brown2015gene})
+and \emph{protein} (for proteins sequences, \cite{sayers2010database}).
 
@@
 
-The first query was a call to the 'gene' database for the 
+The first query was a call to the \emph{gene} database for the 
 term 'membrane protein' (in all fields) 
@@
 % has been obsoleted.
-The next query was a call to the 'gene' database 
+The next query was a call to the \emph{gene} database 
 to obtain the gene names from the gene IDs.
-Per gene name, the 'dbSNP' NCBI database was queried for 
+Per gene name, the \emph{dbSNP} NCBI database was queried for 
 variations associated with the gene name. 
@@
 SNPs were picked based on ID number, which is linked to their discovery date.
-To verify that ID number is unrelated to SNP positions the relative positions of the analyzed SNPS in a protein was determined and showed no positional bias, shown in figure \ref{fig:snp_rel_pos}.
-
-Per SNP, the 'protein' NCBI database was queried for the
+To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined.
+This analysis showed no positional bias of the SNPs, as shown in \geert{dit is supplement toch?} supplementary figure \ref{fig:snp_rel_pos}.
+
+Per SNP, the \emph{protein} NCBI database was queried for the
 protein sequence.
-For each protein sequence, the protein topology was determined 
-using PureseqTM.
-Using these predicted protein topologies, 
-the SNPs were scored to be located within or outside trans membrane domains.
+For each protein sequence, the protein topology was determined using \verb;PureseqTM;.
+Using these predicted protein topologies, the SNPs were scored to be located within or outside TMHs.
 
@@
 Per MHC-I haplotype, it shows the percentage of binders that overlap with a TMH 
-for at least one residue.
-A horizontal line shows the expected percentage of TMH-derived epitopes 
+with at least one residue.
+The horizontal line shows the expected percentage of TMH-derived epitopes 
 that would be presented, if TMH-derived epitopes would be presented just as 
-likely as soluble regions.
-In 11 out of 13 MHC-I haplotypes, TMH-derived epitopes are presented more often 
-then the null expectation, for a human and bacterial proteome.
+likely as epitopes derived from soluble regions.
+For 11 out of 13 MHC-I haplotypes, TMH-derived epitopes are predicted to be presented more often 
+than the null expectation, for a human and bacterial proteome.
 For the viral proteome, 12 out of 13 haplotypes present
 TMH-derived epitopes more often than expected by chance.
-The extent of the over-presentation between the various different haplotypes 
+The extent of the over-presentation between the different haplotypes 
 is similar for the probed proteomes, 
@@
 of TMH-derived peptides.
+\frans{zoals eerder HLA-C mist en ook HLA-DP voor de nu gekozen haplotypes mis ik de rationale waarom we deze hebben genomen?
+volgens mij had ik een artikel gestuurd met de meest voorkomende HLA-DP haplotypes.
+De rationale voor HLA-C is dat de expressie hier naar verhousding heel laag van is en dat dus een minder explecitie rol speelt in antigen presentatie} \geert{in methods section geeft Richel al rationale voor de keuze van de haplotypes, waarin hij schrijft dat dit gebasseerd is op onze eerdere studie (voor MHC-I) en de frequentie van de haplotypes (voor MHC-II).
+Dit lijkt mij voldoende.}
 
@@
   \end{subfigure}  
-  \caption{
+  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes} \geert{een figuur heeft altijd een titel met de conclusie.
+De legenda zelf is feitelijke opsomming en biedt geen (of weinig) interpretatie}
     \textbf{(A)} 
-    TMH-derived peptides are predicted to be over-presented in general,
-    which is for MHC-I (left column) and MHC-II (right column),
-    as well as for the proteomes of SARS-CoV-2 (top row), human (middle 
+    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
+    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
     row) and \emph{M. tuberculosis} (bottom row).
-    The bars show the percentage of epitopes predicted to be 
-    overlapping with TMHs.
-    The red lines indicate the percentage as expected by chance.
+    The red lines indicate the percentages as expected by chance.
     See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
     for the exact TMH and  epitope counts.
-    \textbf{(B)} and \textbf{(C)}:
-    The relation between the percentage of TMH-derived epitopes presented,
-    and the hydrophobicity score of binders for MHC-I haplotypes (panel B)
-    and MHC-II haplotypes (panel C). The trendline assumes a linear relationship.
-  }
+    \textbf{(B-C)}
+    Correlation between the percentages of predicted TMH-derived epitopes
+    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
+    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. Labels are shorthand for the HLA haplotypes,
+    see Table \ref{tab:haplotype_abbreviations} for the names. \frans{wat is de reden dat nr 15 en 8 er buiten vallen? de rest is vrijwel linear)
+  }}
 \end{figure}
@@
 of the 21 MHC-II haplotypes, 
-in a human, bacterial and viral proteome,
-except for \verb;HLA-DRB3*0101; in \emph{MTb}
+for a human, bacterial and viral proteome,
+except for \verb;HLA-DRB3*0101; in \emph{M. tuberculosis}.
 See supplementary table \ref{tab:tmh_binders_mhc2} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Hydrophocity explains the generality of this finding}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Hydrophobicity may be the common reason of the consistent over-presentation
-of  (hydrophobic) TMH-derived epitopes.
-Figures \label{fig:hydrophobicity_1} and \label{fig:hydrophobicity_2}
-show the hydrophobicity preference score for the extent over-presentation
-of TMH-derived epitopes for the different haplotypes in a human proteome.
+\subsection{The over-presentation of TMH-derived peptides is caused by the hydrophobicity of the MHC peptide binding groove}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}.
+Figures \label{fig:hydrophobicity_1} and \label{fig:hydrophobicity_2} \frans{de verwijzing werkt denk ik niet}
+show the extent of over-presentation
+of TMH-derived epitopes as a function of the hydrophobicity preference score for the different haplotypes.
 An assumed linear correlation explains 88\% of the variability in MHC-I.
 For MHC-II, 62\% of the variability is explained by hydrophobicity.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TMH-derived peptides are predicted to be presented in vivo}
+\frans{dit is dus omdat 2 haplotypes buiten de boot vallen} This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experimental validation of presentation of TMH-derived peptides}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
-  \caption{
+  \caption{ \textbf{MHC-I and -II presentation of TMH-derived peptides in peptide elution studies.}
     TMH-derived epitopes are presented \emph{in vivo},
     for both MHC-I and MHC-II, for two topology prediction
-    programs.
+    programs. \geert{legenda is veel te summier. Wat is precies geplot? ook studies van de data citeren.}
   }
@@
 found in the MHC-I and MHC-II elution 
-studies,
-for the two topology prediction tools TMHMM (\cite{krogh2001predicting}) and PureseqTM (\cite{wang2019efficient}). 
+studies, \geert{is dit van totale peptides, of van humane peptides?
+Uiteraard van humane peptides gebruiken, want de rest heb je niet gemapped.
+Dat moet dan wel hier gezegd worden.
+Daarnaast: zijn de MHC-I en -II haplotypes bekend in deze studies?}
+for the two topology prediction tools \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
 Regardless of the prediction tool, 
@@
 
-  \caption{
+  \caption{ \textbf{Evolutionary conservation of human TMHs.}
     \textbf{(A)} 
@@
     and found in the dbSNP database (right bar).
-    Percentages are added to show the relative conservation
+    Percentages show the relative conservation
     of SNPs in TMHs found.
@@
     Percentage of SNPs found in TMHs.
-    Each point shows the predicted percentage of
-    TMH (x-axis) and the observed occurrence of SNPs being located
-    within a TMH (y-axis) for a protein.
-    The dashed diagonal line shows the line of equality (i.e.
+    Each point shows for one protein the predicted percentage of
+    TMH ($x$-axis) and the observed occurrence of SNPs being located
+    within a TMH ($y$-axis).
+    The dashed diagonal line shows the line of equality (i.e.,
     equal conservation of TMHs and soluble protein regions).
+The red line shows the linear regression analysis and the grey shaded area the 95\% confidence interval.
+\geert{please correct last sentence}
     \textbf{(C)}
@@
   \end{subfigure}
-  \caption{
+  \caption{ \textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
       \textbf{(A)} 
-      The number of SNPs in TMHs as expected by chance (red bars) 
-      and found in the dbSNP database (right bars), 
-      after splitting the data into single- and multi-spanners.
-      Percentages are added to show the relative conservation
+      The number of SNPs in TMHs as expected by chance 
+      and observed in the dbSNP database, 
+      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
+      Percentages show the relative conservation
       of SNPs in TMHs found.
@@
       (left) or multiple (right) TMHs.
-      The dashed diagonal lines show the expected line of equality 
-      (i.e., equal conservation of TMHs and solvent-exposed 
-      protein regions).
+        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
+The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
+The red lines shows the linear regression analyses and the grey shaded areas the 95\% confidence intervals.
+\geert{please correct last sentence}
      }
@@
 We determined the conservation of TMHs 
-by comparing the occurrences of SNPs located in TMHs or solvent-exposed protein regions 
+by comparing the occurrences of SNPs located in TMHs or soluble protein regions 
 for the genes coding for membrane proteins.
 We obtained 911 unique gene names associated with the phrase 'membrane protein',
-which are genes for both membrane-associated proteins (MAPs, which have no TMH) and 
+which are genes coding for both membrane-associated proteins (MAPs, which have no TMH) and 
 transmembrane proteins (TMPs, which have at least one TMH).
@@
 Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
-Most our TMPs had less than half their residues located in the 
-plasma membrane.
 The proportion of SNPs found in TMHs varied from 
@@
 soluble regions) to all.
-To determine if it is likelier to find SNPs within either TMHs or
-soluble-facing domains, we performed a linear regression analysis,
+To determine if SNPs were randomly distributed over the protein, we performed a linear regression analysis,
 and added a 95\% confidence interval on this regression.
@@
 We determined the probability to find the observed amount
-of SNPs in TMHs by chance, i.e. when assuming SNPs occur 
+of SNPs in TMHs by chance, i.e., when assuming SNPs occur 
 just as likely in soluble domains as in TMHs.
@@
 is the percentage of residues within a TMH per TMP. 
-These probabilities are shown as a histogram 
-in figure \ref{fig:f_tmh_ncbi}). 
+These percentages are shown as a histogram 
+in figure \ref{fig:f_tmh_ncbi}. 
 The expected number of SNPs expected to be found in 
@@
 
-We split this analysis based on the number of TMHs
-of each TMP. 
-We hypothesized that single-membrane spanners 
-(i.e., proteins with a single predicted TMH) 
-are less conserved than multi-membrane spanners (with $>$1 TMH),
+We split this analysis for TMPs containing only a single TMH (so-called single-membrane spanners) and TMPs containing multiple TMHs (multi-membrane spanners). 
+We hypothesized that single-membrane spanners are less conserved than multi-membrane spanners,
 because multi-membrane spanners
@@
 thus might have additional structural constraints.
-From the split data, we did the same analysis:
-figure \ref{fig:f_snps_found_and_expected_per_spanner} 
-shows, the percentage a protein resides in the membrane versus the
-percentage of SNPs found in those TMH regions, per spanner.
-In both cases, a linear regression shows that less
+From the split data, we did the same analysis as for the total TMPs.
+Figure \ref{fig:f_snps_found_and_expected_per_spanner} 
+shows the percentages of TMHs for individual proteins as a function of the
+percentage of SNPs located in TMHs.
+For both single- and multi-spanners, a linear regression shows that less
 SNPs are found in TMHs, than expected by chance.
 
-Also per spanner, we determined the probability to find the 
-observed amount of SNPs in TMHs by chance.
+We also determined the probability to find the 
+observed amount of SNPs  by chance in single- and multi-spanners.
 For single-spanners, we found 452 SNPs in TMH, where
@@
 The chance to observe this or a lower number by chance is 
-$0.319$. As this chance was higher than our $\alpha = 0.025$
-and we considered this no significant effect.
-For the single-spanners, we found 3,351 SNPs in TMH, where
-$\approx3678$ were expected by chance. 
+$0.319$. As this chance was higher than our $\alpha = 0.025$,
+we consider this no significant effect.
+For the single-spanners, we found 3,351 SNPs in TMH, where \frans{er staat 2keer single spanners dit zijn. toch de multispanners?}
+$\approx3,678$ were expected by chance. 
 The chance to observe this or a lower number by chance is 
 $8.315841 \cdot 10^{-12}$, 
-which means we attribute the lower number of SNPs found
-to an effect, instead of luck.
-
-Also, per spanner type, 
-we determined the relevance of this finding, by
-calculating how much less SNPs are found in TMHs,
+which means this number is significantly less as explained by variation.
+
+Also, for single- and multi-spanners, 
+we determined the relevance of this finding by
+calculating how much less SNPs are found in TMHs
 when compared to soluble regions,
 as depicted in figure \ref{fig:conservation_per_spanner}.
-In effect, per 1000 SNPs found in soluble protein domains, 
+In effect, per 1,000 SNPs found in soluble protein domains, 
 one finds 978 SNPs in TMHs 
-in single spanners (although this is not significant)
-and 911 SNPs in TMHs in multi-spanners.
+of single-spanners (although this is not significant)
+and 911 SNPs in TMHs of multi-spanners.
 
@@
 derived from TMHs are a major but overlooked source of MHC epitopes. 
-Both our bioinformatics predictions and 
-reanalysis of peptide elution studies show that TMH-derived epitopes 
-are presented more often than expected by chance, 
+Our bioinformatics predictions indicate that TMH-derived epitopes 
+are presented to both cytolytic and helper T cells more often than expected by chance, 
 regardless of the organism. 
-Moreover, our SNP analysis shows that TMHs are evolutionary more conserved 
+Moreover, reanalysis of peptide elution studies confirmed the presentation of TMH-derived epitopes.
+Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
 than solvent-exposed protein regions.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TMH-derived epitopes are predicted to be presented in vivo}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
-We show that 1.4\% of epitopes that are naturally presented on MHC-I 
-can be mapped to stem from a TMH. 
-We also find that MHC-II presents epitopes derived from TMHs \emph{in vivo}, 
-as we find that 3.9-4.0\% of sequenced epitopes
-can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}).
-
+\subsection{Mechanism of MHC presentation of TMH-derived epitopes}
 % \paragraph{TMH presentation in vivo}
 
-Although elution studies and \emph{in silico} predictions both suggest that 
+Although our data show that 
 TMH-derived epitopes are presented in MHC-I and MHC-II, 
@@
 into the aqueous vacuolar lumen, 
-and their prior or subsequent proteolytic processing are unresolved.
+and their prior or subsequent proteolytic processing are unresolved, as we discussed previously for MHC-I presentation \cite{bianchi2017}. 
 \geert{
-  beetje kort. Suggesties? Experimental approaches? Hoeft niet super
-  origineel, bijvoorbeeld suggereren dat lipases een rol kunnen spelen. Frans kan vast helpen met mooie zin
-}
+  Hier toevoegen wat probleem is MHC-II. Maxim had goed idee voor intraluminal vesicles in MHC-II loading.
+}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experimental evidence for MHC presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\geert{Deze gehele subsection kan wat mij betreft wel weg (of naar de 'minor discussion' in de supplement).
+Het is deels een herhaling van resultaten (1e en 3e paragraaf).
+DRiPs in paragraaf 2 kan echt niet alles verklaren, want MHC-II is opname van heel antigen.
+In parafraag 4, zal de methodological bias zal alleen maar resulteren in het onderschatten van presentatie van TMHs, en is dus ook niet echt relevant hier.}
+% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
+From reanalysis of peptide elution studies, we show that 1.4\% of epitopes that are naturally presented on MHC-I 
+can be mapped to stem from a TMH. 
+We also find that MHC-II presents epitopes derived from TMHs, 
+as we find that 3.9-4.0\% of sequenced epitopes
+can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}). \geert{zie eerdere vraag. Zijn deze percentages van totale peptides, of van humane peptides?}
 
@@
 
-An important caveat of the elution studies is that we assume that 
-the found epitopes are derived from their completely 
+A caveat of the detected TMH peptides in elution studies is that we assume that 
+the found epitopes are all derived from their completely 
 translated predecessor proteins.
-However, there are multiple cellular processes 
-that may result in peptide fragments that did not originate 
-from the mature proteins.
-One such process is the production of defective ribosomal 
+However, defective ribosomal 
 products (DRiPs) (\cite{yewdell1996defective}),
-in which the ribosome terminates its translation prematurely,
-or cause a point mutation or frame shift (\cite{yewdell2011drips}).
-Another process that results in peptide fragments absent in the
-proteome is peptide fusion, 
-in which peptides form covalent bonds with other peptides,
-as is for example found in mice (\cite{delong2016pathogenic}).
-Due to these processes, we may incorrectly attribute an
-epitope to a specific proteome locations, 
-where in reality, it was created \emph{de novo}.
-This may lead to false positives, as we incorrectly deduce
-epitopes to be derived from a TMH, instead of from one of
-these processes. Both processes, however, are thought to
-occur at low frequencies (\cite{yewdell2011drips,delong2016pathogenic}).
+in which the ribosome terminates its translation prematurely  (\cite{yewdell2011drips}), may result in peptide fragments that did not originate from the mature TMPs.
+DRiPs may lead to false positives, as we incorrectly deduce
+epitopes to be derived from a TMH, instead of from incompletely formed proteins.
 
@@
 This imperfection, however, appears to be of limited importance,
-as our data show that two different topology prediction software algorithms give similar results.
-Additionally, this 
-prediction is robust, as it holds for two TMH prediction tools, 
+as our data show that two different topology prediction software algorithms give similar results, 
 as shown by table \ref{tab:elution}).
@@
 there is methodological bias caused by mass spectronomy,
-as hydrophobic fragments are detected less 
+as hydrophobic fragments are detected less efficiently 
 by mass spectronomy (\cite{bianchi2017}). 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TMH-derived epitopes are presented more often than expected by chance}
+\subsection{T cells recognize different protein regions than B cells}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Not only do we predict that MHC-I and MHC-II TMH-derived epitopes 
-are presented, we conclude that these are additionally
-presented more often than expected by chance 
-alone (figure \ref{fig:bbbq_1_smart_results}).
-This \emph{in silico} finding is a general pattern 
-that persists in both MHC-I and MHC-II, 
-for humans, bacteria and viruses.
-
-% \paragraph{MHC-II is expected to present TMHs}
-The results from this study support our hypothesis 
-that TMH-derived epitopes are over-presented
-in MHC-II, similar to what we previously showed for MHC-I (\cite{bianchi2017}). 
-The over-presentation of TMHs for most MHC-II haplotypes 
-appears to follow a similar pattern as MHC-I haplotypes,
-although MHC-II presents slightly more TMH-derived
-epitopes (see Supplementary Materials, figure \ref{fig:rel_presentation}
-for a direct comparison). 
-Unlike MHC-I, the human proteome over-presents TMH-derived
-epitopes for all MHC-II haplotypes that we have investigated.
-Possible there are MHC-II haplotypes that present less TMH-derived
-epitopes than expected by chance.
-The haplotype \verb;HLA-DQA1*0101/DQB1*0501; had the highest
-over-presentation of TMH-derived epitopes,
-which we suggest is the haplotype of choice 
-to investigate the intracellular mechanism of TMH-derived
-antigen presentation.
-
-Our study also suggests that T cells will largely recognize 
+An important implication from the over-presentation of TMH-derived epitopes is that T cells will largely recognize 
 different protein regions than B cells. 
@@
 However, the results from our study predict that most MHC-II haplotypes 
-present relatively hydrophobic peptides, 
-which are less likely to be solvent-exposed. 
+present relatively hydrophobic peptides,  which are less likely to be solvent-exposed. 
 It is unknown why B and T cells seem to predominantly recognize 
@@
 this lowers the chance of B cell mediated autoimmune diseases, 
-because auto-reactive B and T cells recognizing different part 
+because auto-reactive B and T cells recognizing different parts 
 of the same antigen would need to be present for breakage of B cell tolerance.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Less SNPs in TMHs found, hinting at evolutionary conservation}
+\subsection{Evolutionary conservation of TMHs}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-We find that mutations are more likely to occur
-in soluble regions of proteins, than in TMHs.
-When selecting SNPs at random, per 1000 SNPs found in a soluble domain, 
-one finds on average 918 SNPs in TMHs.
-This effect is more pronounced in multi spanners likely as these TMHs more often serve additional functions, besides spanning the lipid bilayer.
-
 % \paragraph{Selection undetectable in whole proteome}
 
-In general, one would hope that evolutionary selection results in
+In general, one might expect that evolutionary selection results in
 an immune system that as most attentive for protein regions that are
@@
 virulence (\cite{velazquez2020positive}).
-These regions, however, only account for a small part of a pathogen's proteome.
-Additionally, these essential parts might differ widely between different pathogenic proteins.
+These regions, however, may only account for a small part of a pathogen's proteome.
+Additionally, the structure and function of these essential regions might differ widely between different pathogenic proteins.
 Because of this scarcity and variance in targets, 
@@
 to provide innate immune responses against such rare essential protein regions, 
-as hinted upon by a study on influenza (\cite{han2019individual}),
+as suggested in a study on influenza (\cite{han2019individual}),
 where it was found that the selection pressure
@@
 over patterns that are rare.
-Where essential catalytic sites in a pathogenic proteome
-are rare, TMHs are common and thus a more feasible 
+While essential catalytic sites in a pathogenic proteome
+might be relatively rare, TMHs are common and thus might be a more feasible 
 target for evolution to respond to.
-We have found the signature of evolution when both factors,
+Indeed, we have found the signature of evolution when both factors,
 that is, TMHs and catalytic sites are likely to co-occur,
-which is in proteins that span the membrane at least twice.
-This may be explained additionally that multi-spanners
-are not only likelier to have a catalytic active site,
-also there are interactions between the TMHs.
-For single-spanners, we found no evolutionary conservation.
-We speculate that the human immune system is more attentive 
-towards TMHs in multi-spanners, as these are evolutionarily conserved.
-
-There have been more efforts to look at the conservation of TMHs,
+which is in TMPs that span the membrane at least twice.
+In contrast to single-spanners, where we found no significant evolutionary conservation, the TMHs of multi-spanners are more evolutionary conserved than soluble protein regions.
+Likely, the TMHs in many multi-spanners need to interact which each other for correct protein structure and function and they might hence be more structurally constrained compared to the TMHs of single-spanners.
+Thus, we speculate that the human immune system is more attentive 
+towards TMHs in multi-spanners, as these are evolutionarily more conserved.
+
+There have been more efforts to assess the conservation of TMHs,
 using different methodologies.
@@
 percent lower,
-which is the same value as this study found.
-Another example is a study that estimates the conservation
-scores for TMH and soluble regions based on 
-evolutionary related protein alignments,
-and found that TMHs are more conserved, 
+which is a similar value as we found.
+Another example is a study that estimated the conservation
+scores for TMHs and soluble regions based on 
+alignments of evolutionary related proteins,
+and also found that TMHs are more conserved, 
 with a conservation score that was 17\% higher in 
 TMHs (\cite{oberai2009structural}).
-Note that the last study also used human SNPs
-and found that mutations in TMH are likelier to cause
-a disease, again underlining the importance of TMHs.
+Note that the last study also found that mutations in human TMHs are likelier to cause
+a disease, in line with our conclusion that TMHs are more conserved.
 
@@
 
-From this study, two important conclusions can be drawn. 
+Together, from this study, two important conclusions can be drawn. 
 First, the MHC over-presentation of TMHs is likely a general feature 
@@
 and for humans as well as bacterial and viral pathogens. 
-Second, TMHs are more evolutionary conserved than solvent-exposed protein motifs, 
+Second, TMHs are genuinely more evolutionary conserved than soluble protein motifs, 
 at least in the human proteome.
@@
 high performance computing cluster. 
+FB is funded by a Veni grant from the Netherlands Organization for Scientific
+Research (016.Veni.192.026) and an Off-Road Grant from the Dutch Medical Science Foundation (ZonMW 04510011910005).
 GvdB is funded by a Young Investigator Grant from 
@@
 
+Table \ref{tab:haplotype_abbreviations} shows the
+shorthand notation for the HLA haplotypes.
+
+% Label: tab:haplotype_abbreviations
+\input{bbbq_1_smart_results/haplotype_abbreviations.tex}
+
 Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}

### 2021-04-22 bbbq_article.tex
Changes Geert and Maxim

@@
 Most prominently, the fundamental principles of 
-how TMHs are extracted from the hydrophobic lipid environment 
+how TMHs are extracted from their hydrophobic lipid environments 
 into the aqueous vacuolar lumen, 
-and their prior or subsequent proteolytic processing are unresolved, as we discussed previously for MHC-I presentation \cite{bianchi2017}. 
-\geert{
-  Hier toevoegen wat probleem is MHC-II. Maxim had goed idee voor intraluminal vesicles in MHC-II loading.
-}
+and their prior or subsequent proteolytic processing are unresolved.
+
+A first possibility is that the extraction of TMPs from the membrane is mediated by the the ER-associated degradation (ERAD) machinery.
+For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmic reticulum (ER).
+The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) (Rock et al., 2016) first assemble and stabilize the heavy and light chains of MHC-I.
+Later, this complex binds to the transporter associated with antigen processing (TAP) leading to the formation of the so-called peptide-loading complex (PLC).
+The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I (Blees et al., 2017).
+Membrane proteins first will have to be extracted from the membrane before they become amenable to this MHC-I loading by the PLC.
+In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation (Preston GM, 2017, Meusser B 2005).
+Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading.
+This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{Bougnères2009}, since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo (Fujimoto and Ohsaki, 2006).
+\geert{missen nog bijna alle citaties in dit stuk}
+
+A second possibility is that TMPs are proteolytically processed by intramembrane proteases that cleave TMHs while they are still membrane embedded.
+Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E (Oliveira, 2015).
+The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER (Oliveira, 2015).
+However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein.
+Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well (Lautscham, 2001), suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
+\geert{missen nog bijna alle citaties in dit stuk}
+
+A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) (Oliveira, 2015).
+TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes.
+Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles.
+This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles.
+Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded (Gruenberg,  2019).
+Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs (Kleijmeer 2001, Peters 1991, Zwart 2005).
+However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs (Zwart 2005), indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
+This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact.
+A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading (Zwart et al 2005).
+\geert{check of correct zo, en weer referenties missen}.
+
+Alternatively to the enzymatic degradation of lipids in MVBs by lipases (Sander 2016; Gilleron 2016), they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 (Dingjan 2016).
+This oxidation can result in a destabilization and disruption of membranes (Dingjan 2016) and might thereby lead to the extraction of TMPs.
+Due to the hydrophobic nature of TMHs, however, the extracted proteins will likely aggregate and it is unclear how these aggregates would be processed further for MHC loading.
+\geert{referenties missen}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-04-24 bbbq_article.tex
Improved by Frans, thanks @fransbianchi

--- 2021-04-12 bbbq_article.tex
@@
 Most prominently, the fundamental principles of 
-how TMHs are extracted from the hydrophobic lipid environment 
+how TMHs are extracted from their hydrophobic lipid environments 
 into the aqueous vacuolar lumen, 
-and their prior or subsequent proteolytic processing are unresolved, as we discussed previously for MHC-I presentation \cite{bianchi2017}. 
-\geert{
-  Hier toevoegen wat probleem is MHC-II. Maxim had goed idee voor intraluminal vesicles in MHC-II loading.
-}
+and their prior or subsequent proteolytic processing are unresolved.
+
+A first possibility is that the extraction of TMPs from the membrane is mediated by the ER-associated degradation (ERAD) machinery.
+For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmic reticulum (ER).
+The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) \cite{rock2016present} first assemble and stabilize the heavy and light chains of MHC-I.
+Later, this complex binds to the transporter associated with antigen processing (TAP) leading to the formation of the so-called peptide-loading complex (PLC).
+The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I \cite{blees2017structure}.
+Membrane proteins first will have to be extracted from the membrane before they become amenable to this MHC-I loading by the PLC.
+In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation \cite{preston2017evolving} \cite{meusser2005erad}.
+Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading.
+This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{Bougnères2009}, since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.
+
+A second possibility is that TMPs are proteolytically processed by intramembrane proteases that cleave TMHs while they are still membrane embedded.
+Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}.
+The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}.
+However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein.
+Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well \cite{lautscham2001processing}, suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
+
+A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \cite{oliveira2015alternative}.
+TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes.
+Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles.
+This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles.
+Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded \cite{gruenberg2020life}.
+Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs \cite{kleijmeer2001reorganization}\cite{peters1991segregation}\cite{zwart2005spatial}.
+However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs \cite{zwart2005spatial}, indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
+This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact.
+A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading \cite{zwart2005spatial}.
+
+Alternatively to the enzymatic degradation of lipids in MVBs by lipases\cite{sander2016lipase} \cite{gilleron2016lysosomal}, they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}.
+This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} and might thereby lead to the extraction of TMPs.
+Due to the hydrophobic nature of TMHs, however, the extracted proteins will likely aggregate and it is unclear how these aggregates would be processed further for MHC loading.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-04-24 bbbq_article.tex
Use 'ref' not 'label' to refer to a figure

@@
 For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}.
-Figures \label{fig:hydrophobicity_1} and \label{fig:hydrophobicity_2} \frans{de verwijzing werkt denk ik niet}
+Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
 show the extent of over-presentation
@@
 Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading.
-This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{Bougnères2009}, since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.
+This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{bougneres2009role}, since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.
 

### 2021-04-24 bbbq_article.tex
Processed feedback, thanks Frans and Geert, thanks @bogeert and @fransbianchi

@@
 
-\geert{wat text gewist want dit is de methods sectie}
 In previous work, it was found that the over-presentation of TMH-derived peptides can be
@@
 that give rise to no or only very few MHC epitopes.
-To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials. \geert{refereer hier naar supplementary figuur}
+To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials.
 
@@
 To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined.
-This analysis showed no positional bias of the SNPs, as shown in \geert{dit is supplement toch?} supplementary figure \ref{fig:snp_rel_pos}.
+This analysis showed no positional bias of the SNPs, as shown in supplementary figure \ref{fig:snp_rel_pos}.
 
@@
 of TMH-derived peptides.
-\frans{zoals eerder HLA-C mist en ook HLA-DP voor de nu gekozen haplotypes mis ik de rationale waarom we deze hebben genomen?
-volgens mij had ik een artikel gestuurd met de meest voorkomende HLA-DP haplotypes.
-De rationale voor HLA-C is dat de expressie hier naar verhousding heel laag van is en dat dus een minder explecitie rol speelt in antigen presentatie} \geert{in methods section geeft Richel al rationale voor de keuze van de haplotypes, waarin hij schrijft dat dit gebasseerd is op onze eerdere studie (voor MHC-I) en de frequentie van de haplotypes (voor MHC-II).
-Dit lijkt mij voldoende.}
 
@@
   \end{subfigure}  
-  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes} \geert{een figuur heeft altijd een titel met de conclusie.
-De legenda zelf is feitelijke opsomming en biedt geen (of weinig) interpretatie}
+  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes}
     \textbf{(A)} 
@@
     and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. Labels are shorthand for the HLA haplotypes,
-    see Table \ref{tab:haplotype_abbreviations} for the names. \frans{wat is de reden dat nr 15 en 8 er buiten vallen? de rest is vrijwel linear)
-  }}
+    see Table \ref{tab:haplotype_abbreviations} for the names.
+  }
 \end{figure}
@@
 For MHC-II, 62\% of the variability is explained by hydrophobicity.
-\frans{dit is dus omdat 2 haplotypes buiten de boot vallen} This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.
+This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.
 
@@
   \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
-  \caption{ \textbf{MHC-I and -II presentation of TMH-derived peptides in peptide elution studies.}
-    TMH-derived epitopes are presented \emph{in vivo},
-    for both MHC-I and MHC-II, for two topology prediction
-    programs. \geert{legenda is veel te summier. Wat is precies geplot? ook studies van de data citeren.}
+  \caption{
+    \textbf{
+      Robust prediction that TMH epitopes are presented \emph{in vivo}.
+    }
+    Bars show the percentage of peptides obtained from elution
+    studies that is derived from a TMH, for MHC-I and -II,
+    for two TMH prediction tools.
   }
@@
 are presented in MHC-I and MHC-II,
-two peptide elution studies (\cite{schellens2015comprehensive} and \cite{bergseng2015different} for MHC-I and MHC-II respectively) 
-were reanalyzed.
+two peptide elution studies were reanalyzed:
+For MHC-I, peptides presented \emph{in vivo} 
+by the (humane) haplotypes HLA-A and B were sequenced (\cite{schellens2015comprehensive}),
+for MHC-II these were haplotypes DQ2.5, DQ2.2, and 
+DQ7.5 (\cite{bergseng2015different}).
 Figure \ref{fig:elution} shows the percentages of epitopes derived 
@@
 found in the MHC-I and MHC-II elution 
-studies, \geert{is dit van totale peptides, of van humane peptides?
-Uiteraard van humane peptides gebruiken, want de rest heb je niet gemapped.
-Dat moet dan wel hier gezegd worden.
-Daarnaast: zijn de MHC-I en -II haplotypes bekend in deze studies?}
+studies,
 for the two topology prediction tools \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
@@
     The dashed diagonal line shows the line of equality (i.e.,
-    equal conservation of TMHs and soluble protein regions).
-The red line shows the linear regression analysis and the grey shaded area the 95\% confidence interval.
-\geert{please correct last sentence}
+    equal conservation of TMHs and soluble protein regions). The red line indicates a linear fit, the gray area its 95\% confidence interval.
     \textbf{(C)}
@@
 The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
-The red lines shows the linear regression analyses and the grey shaded areas the 95\% confidence intervals.
-\geert{please correct last sentence}
+The red line indicates a linear fit, the gray area its 95\% confidence interval.
      }
@@
 we consider this no significant effect.
-For the single-spanners, we found 3,351 SNPs in TMH, where \frans{er staat 2keer single spanners dit zijn. toch de multispanners?}
+For the multi-spanners, we found 3,351 SNPs in TMH, where 
 $\approx3,678$ were expected by chance. 
@@
 For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmic reticulum (ER).
-The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) \cite{rock2016present} first assemble and stabilize the heavy and light chains of MHC-I.
+The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) (\cite{rock2016present}) first assemble and stabilize the heavy and light chains of MHC-I.
 Later, this complex binds to the transporter associated with antigen processing (TAP) leading to the formation of the so-called peptide-loading complex (PLC).
-The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I \cite{blees2017structure}.
+The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I (\cite{blees2017structure}).
 Membrane proteins first will have to be extracted from the membrane before they become amenable to this MHC-I loading by the PLC.
-In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation \cite{preston2017evolving} \cite{meusser2005erad}.
+In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation (\cite{preston2017evolving,meusser2005erad}).
 Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading.
-This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{bougneres2009role}, since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.
+This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) (\cite{bougneres2009role}), since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo (\cite{fujimoto2006proteasomal}).
 
 A second possibility is that TMPs are proteolytically processed by intramembrane proteases that cleave TMHs while they are still membrane embedded.
-Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}.
-The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}.
+Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E (\cite{oliveira2015alternative}).
+The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER (\cite{oliveira2015alternative}).
 However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein.
-Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well \cite{lautscham2001processing}, suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
-
-A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \cite{oliveira2015alternative}.
+Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well (\cite{lautscham2001processing}), suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
+
+A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) (\cite{oliveira2015alternative}).
 TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes.
@@
 This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles.
-Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded \cite{gruenberg2020life}.
-Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs \cite{kleijmeer2001reorganization}\cite{peters1991segregation}\cite{zwart2005spatial}.
-However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs \cite{zwart2005spatial}, indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
+Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded (\cite{gruenberg2020life}).
+Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs (\cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}).
+However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs (\cite{zwart2005spatial}), indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
 This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact.
-A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading \cite{zwart2005spatial}.
-
-Alternatively to the enzymatic degradation of lipids in MVBs by lipases\cite{sander2016lipase} \cite{gilleron2016lysosomal}, they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}.
-This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} and might thereby lead to the extraction of TMPs.
+A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading (\cite{zwart2005spatial}).
+
+Alternatively to the enzymatic degradation of lipids in MVBs by lipases (\cite{sander2016lipase,gilleron2016lysosomal}), they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 (\cite{dingjan2016lipid}).
+This oxidation can result in a destabilization and disruption of membranes (\cite{dingjan2016lipid}) and might thereby lead to the extraction of TMPs.
 Due to the hydrophobic nature of TMHs, however, the extracted proteins will likely aggregate and it is unclear how these aggregates would be processed further for MHC loading.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experimental evidence for MHC presentation of TMH-derived epitopes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\geert{Deze gehele subsection kan wat mij betreft wel weg (of naar de 'minor discussion' in de supplement).
-Het is deels een herhaling van resultaten (1e en 3e paragraaf).
-DRiPs in paragraaf 2 kan echt niet alles verklaren, want MHC-II is opname van heel antigen.
-In parafraag 4, zal de methodological bias zal alleen maar resulteren in het onderschatten van presentatie van TMHs, en is dus ook niet echt relevant hier.}
-% \paragraph{Over-presentation in vivo in MHC-I and MHC-II is robust
-From reanalysis of peptide elution studies, we show that 1.4\% of epitopes that are naturally presented on MHC-I 
-can be mapped to stem from a TMH. 
-We also find that MHC-II presents epitopes derived from TMHs, 
-as we find that 3.9-4.0\% of sequenced epitopes
-can be uniquely traced onto a human reference proteome (also shown in table \ref{tab:elution}). \geert{zie eerdere vraag. Zijn deze percentages van totale peptides, of van humane peptides?}
-
-% \paragraph{False positives}
-
-A caveat of the detected TMH peptides in elution studies is that we assume that 
-the found epitopes are all derived from their completely 
-translated predecessor proteins.
-However, defective ribosomal 
-products (DRiPs) (\cite{yewdell1996defective}),
-in which the ribosome terminates its translation prematurely  (\cite{yewdell2011drips}), may result in peptide fragments that did not originate from the mature TMPs.
-DRiPs may lead to false positives, as we incorrectly deduce
-epitopes to be derived from a TMH, instead of from incompletely formed proteins.
-
-Another caveat is that the TMH prediction tools are imperfect.
-This imperfection, however, appears to be of limited importance,
-as our data show that two different topology prediction software algorithms give similar results, 
-as shown by table \ref{tab:elution}).
-
-A final caveat in the elution studies is that 
-there is methodological bias caused by mass spectronomy,
-as hydrophobic fragments are detected less efficiently 
-by mass spectronomy (\cite{bianchi2017}). 
-This means that the number of TMH-derived epitopes
-might be underestimated.
 

### 2021-04-24 bbbq_article.tex
Spellcheck
Figure 1's caption does not overlap with page number
Use correctly formatted species names, fix #206

@@
   \centering
-  \begin{subfigure}[t]{0.9\textwidth}
+  \begin{subfigure}[t]{0.8\textwidth}
     \centering
@@
 
-  \begin{subfigure}[t]{0.4\textwidth}
+  \begin{subfigure}[t]{0.35\textwidth}
     \centering
@@
   \hfill
-  \begin{subfigure}[t]{0.4\textwidth}
+  \begin{subfigure}[t]{0.35\textwidth}
     \centering
@@
 A first possibility is that the extraction of TMPs from the membrane is mediated by the ER-associated degradation (ERAD) machinery.
-For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmic reticulum (ER).
+For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmatic reticulum (ER).
 The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) (\cite{rock2016present}) first assemble and stabilize the heavy and light chains of MHC-I.
@@
 where \verb;covid; denotes SARS-CoV-2 and \verb;myco; denotes
-Mycobacterium tuberculosis. \verb;mhc_class; denotes the MHC
+\emph{Mycobacterium tuberculosis}. \verb;mhc_class; denotes the MHC
 class, \verb;n_spots; the number of possible 9-mers (for MHC-I) 
@@
 
-Per target proteome (i.e. human, SARS-CoV-2, Mycobacterium tuberculosis),
+Per target proteome (i.e. human, SARS-CoV-2, \emph{M tuberculosis}),
 we collected all 9-mers (for MHC-I) and 14-mers (for MHC-II),
@@
     Legend: covid = SARS-CoV-2,
-    human = homo sapiens, myco = Mycobacterium tuberculosis
+    human = \emph{Homo sapiens}, 
+    myco = \emph{Mycobacterium tuberculosis}
   }
@@
     Legend: covid = SARS-CoV-2,
-    human = homo sapiens, myco = Mycobacterium tuberculosis
+    human = \emph{Homo sapiens}, myco = \emph{Mycobacterium tuberculosis}
   }

### 2021-04-24 bbbq_article.tex
Use unsrtnat/Vancouver bibliography style, progress #205
Use Vancouver style

@@
 For example, HLA-A and HLA-B have no overlap in which
-epitopes they bind (\cite{lund2004definition}).
+epitopes they bind \cite{lund2004definition}.
 Moreover, the HLA genes of humans are highly polymorphic, with hundreds 
@@
 and each different HLA allele is called 
-an MHC haplotype and presents a different subset of peptides (\cite{marsh2010nomenclature}).
+an MHC haplotype and presents a different subset of peptides \cite{marsh2010nomenclature}.
 
@@
 so-called escape mutations, 
-will not affect MHC presentation for all haplotypes present in the population  (\cite{sommer2005importance}).
+will not affect MHC presentation for all haplotypes present in the population \cite{sommer2005importance}.
 
@@
 of the binding affinities of peptides
-(\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}).
+\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
 For example, \cite{tang2011genome} found that, 
@@
 will be more frequently presented by MHC-I 
-than expected based on their abundance (\cite{bianchi2017}). 
+than expected based on their abundance \cite{bianchi2017}. 
 Moreover, we showed that some well-known immunodominant peptides stem from TMHs. 
@@
 TMHs can also be predicted with high accuracy from a protein sequence 
-by bioinformatics approaches (\cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}),
+by bioinformatics approaches \cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient},
 for example, \cite{jones2007improving} found that,
@@
 Different TMH prediction tools estimate that 15-39\% of all proteins 
-in the human proteome contain at least one TMH (\cite{ahram2006estimation}).
+in the human proteome contain at least one TMH \cite{ahram2006estimation}.
 However, the physiological reason why peptides derived from TMHs 
@@
 many of the amino acids genuinely present in TMHs 
-are limited to the ones with hydrophobic side chains (\cite{jones1994model}).
+are limited to the ones with hydrophobic side chains \cite{jones1994model}.
 Therefore, we speculated that the TMHs of pathogens 
@@
 
-For MHC-I, 9-mers were used, as this is the length most frequently presented in MHC-I and was used in our earlier study (\cite{bianchi2017}). 
+For MHC-I, 9-mers were used, 
+as this is the length most frequently presented in MHC-I 
+and was used in our earlier study \cite{bianchi2017}. 
 For MHC-II, 14-mers were used, 
-as these are the most frequently occurring epitope length (\cite{bergseng2015different}).
+as these are the most frequently occurring epitope length \cite{bergseng2015different}.
 A human (UniProt ID UP000005640\_9606), 
@@
 reference proteome was used. 
-\verb;TMHMM; (\cite{krogh2001predicting}) was used to predict the topology 
+\verb;TMHMM; \cite{krogh2001predicting} was used to predict the topology 
 of the proteins within these proteomes.
 To predict the affinity of an epitope to a certain MHC haplotype,
- \verb;EpitopePrediction; (\cite{bianchi2017}) for MHC-I 
-and \verb;MHCnuggets; (\cite{shao2020high}) for MHC-II was used.
+ \verb;EpitopePrediction; \cite{bianchi2017} for MHC-I 
+and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
 The 13 MHC-I haplotypes used in this study are the same as 
-used in the previous study (\cite{bianchi2017}).
+used in the previous study \cite{bianchi2017}.
 For MHC-II, haplotypes were selected with a phenotypic frequency 
-of at least 14\% in the human population (\cite{greenbaum2011functional}),
+of at least 14\% in the human population \cite{greenbaum2011functional},
 resulting in 21 MHC-II haplotypes.
@@
 In previous work, it was found that the over-presentation of TMH-derived peptides can be
-explained from the hydrophobicity of the MHC-I binding cleft (\cite{bianchi2017}).
+explained from the hydrophobicity of the MHC-I binding cleft \cite{bianchi2017}.
 Here, a similar analysis was applied, by correlating the percentage of predicted TMH-derived epitopes versus the mean hydrophobicity of all 
@@
 
-This study differs in one important aspect from our previous work (\cite{bianchi2017}). 
+This study differs in one important aspect from our previous work \cite{bianchi2017}. 
 The definition of a binder differs from \cite{bianchi2017}:
@@
 
-To obtain experimental evidence that epitopes derived from TMHs are presented in MHC, peptide elution studies for MHC-I (\cite{schellens2015comprehensive}) 
-and MHC-II (\cite{bergseng2015different}) were reanalyzed.
+To obtain experimental evidence that epitopes derived from TMHs are presented in MHC, 
+peptide elution studies for MHC-I \cite{schellens2015comprehensive} 
+and MHC-II \cite{bergseng2015different} were reanalyzed.
 For each of the detected epitopes,
@@
 For the epitopes that were present in the proteome exactly once,
-the topology of the proteins in which these epitopes were located was predicted using both \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
+the topology of the proteins in which these epitopes were located 
+was predicted using both \verb;TMHMM; \cite{krogh2001predicting} 
+and \verb;PureseqTM; \cite{wang2019efficient}. 
 From this topology, we determined if the epitope
@@
 As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: the \emph{dbSNP} (\cite{sherry2001dbsnp}) database,
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: 
+the \emph{dbSNP} \cite{sherry2001dbsnp} database,
 which contains 650 million 
@@
 is similar for the probed proteomes, 
-which strengthens our previous conclusion (\cite{bianchi2017}) 
+which strengthens our previous conclusion \cite{bianchi2017} 
 that the hydrophobicity of the MHC-binding groove 
@@
 For MHC-I, peptides presented \emph{in vivo} 
-by the (humane) haplotypes HLA-A and B were sequenced (\cite{schellens2015comprehensive}),
+by the (humane) haplotypes HLA-A and B 
+were sequenced \cite{schellens2015comprehensive},
 for MHC-II these were haplotypes DQ2.5, DQ2.2, and 
-DQ7.5 (\cite{bergseng2015different}).
+DQ7.5 \cite{bergseng2015different}.
 Figure \ref{fig:elution} shows the percentages of epitopes derived 
@@
 studies,
-for the two topology prediction tools \verb;TMHMM; (\cite{krogh2001predicting}) and \verb;PureseqTM; (\cite{wang2019efficient}). 
+for the two topology prediction tools \verb;TMHMM; \cite{krogh2001predicting} 
+and \verb;PureseqTM; \cite{wang2019efficient}. 
 Regardless of the prediction tool, 
@@
 the molecular mechanisms of how integral membrane proteins are processed 
-for MHC presentation are largely unknown (\cite{bianchi2017}). 
+for MHC presentation are largely unknown \cite{bianchi2017}. 
 Most prominently, the fundamental principles of 
@@
 
-A first possibility is that the extraction of TMPs from the membrane is mediated by the ER-associated degradation (ERAD) machinery.
-For MHC class I (MHC-I) antigen presentation of soluble proteins, the loading of the epitope primarily occurs at the endoplasmatic reticulum (ER).
-The chaperones tapasin (TAPBP), ERp57 (PDIA3), and calreticulin (CALR) (\cite{rock2016present}) first assemble and stabilize the heavy and light chains of MHC-I.
-Later, this complex binds to the transporter associated with antigen processing (TAP) leading to the formation of the so-called peptide-loading complex (PLC).
-The PLC drives import of peptides into the ER and mediates their subsequent loading into the peptide-binding groove of MHC-I (\cite{blees2017structure}).
-Membrane proteins first will have to be extracted from the membrane before they become amenable to this MHC-I loading by the PLC.
-In the ER, this process can be orchestrated by the ERAD machinery, consisting of several chaperones that recognize TMPs, ubiquitinate them, and extract them from the ER membrane into the cytosol (retrotranslocation) for proteasomal degradation (\cite{preston2017evolving,meusser2005erad}).
-Similar to the peptides generated from soluble proteins, the TMP-derived peptides might then be re-imported by TAP into the ER for MHC-I loading.
-This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) (\cite{bougneres2009role}), since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo (\cite{fujimoto2006proteasomal}).
-
-A second possibility is that TMPs are proteolytically processed by intramembrane proteases that cleave TMHs while they are still membrane embedded.
-Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences, are presented on a specialized class of MHC-I called HLA-E (\cite{oliveira2015alternative}).
-The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, possibly because the peptides are directly released into the lumen of the ER (\cite{oliveira2015alternative}).
+A first possibility is that the extraction of TMPs from the membrane 
+is mediated by the ER-associated degradation (ERAD) machinery. 
+For MHC class I (MHC-I) antigen presentation of soluble proteins, 
+the loading of the epitope primarily occurs at the endoplasmatic reticulum (ER). 
+The chaperones tapasin (TAPBP), ERp57 (PDIA3), 
+and calreticulin (CALR) \cite{rock2016present} first assemble 
+and stabilize the heavy and light chains of MHC-I. 
+Later, this complex binds to the transporter 
+associated with antigen processing (TAP) 
+leading to the formation of the so-called peptide-loading complex (PLC). 
+The PLC drives import of peptides into the ER 
+and mediates their subsequent loading into the peptide-binding groove of MHC-I \cite{blees2017structure}. 
+Membrane proteins first will have to be extracted from the membrane 
+before they become amenable to this MHC-I loading by the PLC. 
+In the ER, this process can be orchestrated by the ERAD machinery, 
+consisting of several chaperones that recognize TMPs, 
+ubiquitinate them, and extract them from the ER membrane 
+into the cytosol (retrotranslocation) for proteasomal degradation \cite{preston2017evolving,meusser2005erad}. 
+Similar to the peptides generated from soluble proteins, 
+the TMP-derived peptides might then be re-imported by TAP 
+into the ER for MHC-I loading. 
+This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{bougneres2009role}, 
+since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.
+
+A second possibility is that TMPs are proteolytically processed 
+by intramembrane proteases that cleave TMHs while they are still membrane embedded.
+Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences,
+are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}. 
+The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, 
+possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}. 
 However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein.
-Nevertheless, the presentation of peptides with a high hydrophobicity index was shown to be independent on TAP as well (\cite{lautscham2001processing}), suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
-
-A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) (\cite{oliveira2015alternative}).
+Nevertheless, the presentation of peptides with a high hydrophobicity index 
+was shown to be independent on TAP as well \cite{lautscham2001processing}, 
+suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
+
+A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \cite{oliveira2015alternative}. 
 TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes.
@@
 This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles.
-Upon the fusion of MVBs with lysosomes, the entire intraluminal vesicles including the TMPs are degraded (\cite{gruenberg2020life}).
-Via this mechanism, TMPs might well be processed for antigen presentation, particularly since the loading of MHC class II molecules is well understood to occur in MVBs (\cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}).
-However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles, but not on the limiting membranes of MVBs (\cite{zwart2005spatial}), indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
+Upon the fusion of MVBs with lysosomes,
+the entire intraluminal vesicles including the TMPs are degraded \cite{gruenberg2020life}. 
+Via this mechanism, TMPs might well be processed for antigen presentation, 
+particularly since the loading of MHC class II molecules is well understood 
+to occur in MVBs \cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}. 
+However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles,
+but not on the limiting membranes of MVBs \cite{zwart2005spatial}, 
+indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
 This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact.
-A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles, and it is hence unclear how TMPs from these microbes would be routed to MVBs for MHC-II loading (\cite{zwart2005spatial}).
-
-Alternatively to the enzymatic degradation of lipids in MVBs by lipases (\cite{sander2016lipase,gilleron2016lysosomal}), they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 (\cite{dingjan2016lipid}).
-This oxidation can result in a destabilization and disruption of membranes (\cite{dingjan2016lipid}) and might thereby lead to the extraction of TMPs.
-Due to the hydrophobic nature of TMHs, however, the extracted proteins will likely aggregate and it is unclear how these aggregates would be processed further for MHC loading.
+A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles,
+and it is hence unclear how TMPs from these microbes 
+would be routed to MVBs for MHC-II loading \cite{zwart2005spatial}.
+
+Alternatively to the enzymatic degradation of lipids in MVBs by lipases \cite{sander2016lipase,gilleron2016lysosomal}, 
+they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}. 
+This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} 
+and might thereby lead to the extraction of TMPs. 
+Due to the hydrophobic nature of TMHs, 
+however, the extracted proteins will likely aggregate 
+and it is unclear how these aggregates would be processed further for MHC loading.
 
@@
 selection pressure is upon residues that change its 
-virulence (\cite{velazquez2020positive}).
+virulence \cite{velazquez2020positive}.
 These regions, however, may only account for a small part of a pathogen's proteome.
@@
 to provide innate immune responses against such rare essential protein regions, 
-as suggested in a study on influenza (\cite{han2019individual}),
+as suggested in a study on influenza \cite{han2019individual},
 where it was found that the selection pressure
@@
 with a conservation score that was 17\% higher in 
-TMHs (\cite{oberai2009structural}).
+TMHs \cite{oberai2009structural}.
 Note that the last study also found that mutations in human TMHs are likelier to cause
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Vancouver style
+\bibliographystyle{unsrtnat}
 % MEE style
-\bibliographystyle{mee}
+%\bibliographystyle{mee}
 \bibliography{bbbq_article}
@@
 parts of a protein forms a TMH.
-In 2001, multiple of such prediction tools have been compared (\cite{moller2001evaluation}),
-of which TMHMM (\cite{krogh2001predicting}) turned out to be the most accurate, 
-as is used in the previous study (\cite{bianchi2017}).
+In 2001, multiple of such prediction tools have been compared \cite{moller2001evaluation},
+of which TMHMM \cite{krogh2001predicting} turned out to be the most accurate, 
+as is used in the previous study \cite{bianchi2017}.
 However, TMHMM has a restrictive software license and is nearly two
 decades old.
-Therefore, PureseqTM (\cite{wang2019efficient}),
+Therefore, PureseqTM \cite{wang2019efficient},
 was also used in this study, which has been more recently developed
@@
 According to \cite{lundegaard2011prediction}, at that time,
-NetMHCcons (\cite{karosiene2012netmhccons}) gave the best predictions.
-We used the same tool as used in our earlier study, \verb;epitope-prediction; (\cite{bianchi2017}),
+NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
+We used the same tool as used in our earlier study, 
+\verb;epitope-prediction; \cite{bianchi2017},
 
@@
 to predict epitopes,
-such as using a trained neural network (\cite{nielsen2003reliable})
-or a Gibbs sampling approach (\cite{nielsen2004improved}).
+such as using a trained neural network \cite{nielsen2003reliable}
+or a Gibbs sampling approach \cite{nielsen2004improved}.
 According to \cite{lundegaard2011prediction}, in 2011,
 from a set of multiple tools, 
-NetMHCIIpan (\cite{nielsen2008quantitative,karosiene2013netmhciipan})
+NetMHCIIpan \cite{nielsen2008quantitative,karosiene2013netmhciipan}
 made the most accurate predictions.
 The most recent FOSS tool available now appears
-to be MHCnuggets (\cite{shao2020high}), which can do both MHC-I 
+to be MHCnuggets \cite{shao2020high}, which can do both MHC-I 
 and MHC-II predictions. 
-As we already use \verb;epitope-prediction; (\cite{bianchi2017}) 
+As we already use \verb;epitope-prediction; \cite{bianchi2017} 
 for MHC-I predictions, we use MHCnuggets only for MHC-II predictions.
@@
 To retrieve the data from the NCBI databases the
-\verb;rentrez; R package (\cite{rentrez}) was used
+\verb;rentrez; R package \cite{rentrez} was used
 that calls the NCBI website's API. To provide for a 
@@
 query into multiple pages, each of which needs one API call.
-We wrote the \verb;sprentrez; package (\cite{sprentrez}) to provide for 
+We wrote the \verb;sprentrez; package \cite{sprentrez} to provide for 
 bigger queries of multiple (and delayed) API calls.
@@
 predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
-'TMHMM' has a restrictive software license (\cite{krogh2001predicting}) and allows a user
+'TMHMM' has a restrictive software license \cite{krogh2001predicting} 
+and allows a user
 to download a pre-compiled executable after confirmation he/she
@@
 To be able to call, from R, the TMH prediction 
-software 'PureseqTM' (\cite{wang2019efficient}),
+software 'PureseqTM' \cite{wang2019efficient},
 which is written in C, the package 'pureseqtmr' has been developed. 
@@
 MHCnuggets is a free and open-source Python package to predict 
-epitope affinity for many MHC-I and MHC-II variants (\cite{shao2020high}).
+epitope affinity for many MHC-I and MHC-II variants \cite{shao2020high}.
 The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
@@
 do have an effect and may even be evolutionary selected 
-for (\cite{hunt2009silent}).
+for \cite{hunt2009silent}.
 As the possible effect of synonymous mutations is ignored by our

### 2021-04-24 bbbq_article.tex
Processed feed Maksim, progress #217

@@
 Cytolytic T cell responses are predicted to be biased towards membrane proteins. The peptide-binding grooves of most haplotypes 
-of histocompatibility complex class I (MHC-I) are relatively hydrophobic,  therefor peptide fragments derived from human transmembrane helices (TMHs) are predicted to be presented more often as would be expected
+of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
+therefore peptide fragments derived from human transmembrane helices (TMHs) are predicted to be presented more often as would be expected 
 based on their abundance in the proteome. However, the physiological reason of why membrane proteins might be over-presented is unclear. 
@@
 Thus, our findings suggest that both cytolytic and helper T cells respond more to membrane proteins, because these are evolutionary more conserved.
-We speculate that TMHs therefor are less prone to escape mutations that enable pathogens to evade T cell responses.
+We speculate that TMHs therefore are less prone to escape mutations that enable pathogens to evade T cell responses.
 
@@
 
-Humans mostly express two haplotypes per MHC form, one from the parental and one from the maternal chromosome, and therefore an individual's immune system detects 
+Humans mostly express two haplotypes per MHC form, one from the parental and one from the maternal chromosome, 
+and therefore an individual's immune system detects 
 only a fraction of all possible peptide fragments. 
@@
 is very high, because of the highly polymorphic MHC genes.
-It is therefor believed that MHC polymorphism improves immunity at the population level, 
+It is therefore believed that MHC polymorphism improves immunity at the population level, 
 as mutations in a protein that disrupt a particular MHC presentation at the individual level, 
@@
 is relatively hydrophobic, 
-and therefore hydrophobic peptides have a higher affinity 
-to bind than more hydrophilic peptides.
+and therefore hydrophobic TMH-derived peptides have a higher affinity 
+to bind than their soluble hydrophilic counterparts.
 
@@
 Nevertheless, the presentation of peptides with a high hydrophobicity index 
-was shown to be independent on TAP as well \cite{lautscham2001processing}, 
+was shown to be independent of TAP as well \cite{lautscham2001processing}, 
 suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
@@
 RJCB wrote the code.
-RJCB, FB and GvdB wrote the article.
+RJCB, MB, GvdB and FB wrote the article.
 
@@
 and allows a user
-to download a pre-compiled executable after confirmation he/she
+to download a pre-compiled executable after confirmation that he/she
 is in academia. The R package respects this restriction

### 2021-04-24 bbbq_article.tex
Add non-standard abbreviations list, fix #204
Abreviations is an unnumbered section
Follow guidelines regarding table, progress #210

@@
 
-Cytolytic T cell responses are predicted to be biased towards membrane proteins. The peptide-binding grooves of most haplotypes 
+Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
+The peptide-binding grooves of most haplotypes 
 of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
@@
 In this study, we show that the over-presentation of TMH-derived peptides is general, as it is predicted for bacteria and viruses 
-and for both MHC-I and MHC-II.
-Moreover, we show that TMHs are evolutionarily more conserved, because single nucleotide polymorphisms (SNPs) are present relatively less frequently in TMH-coding chromosomal regions compared to regions coding for extracellular and cytoplasmic protein regions.
-Thus, our findings suggest that both cytolytic and helper T cells respond more to membrane proteins, because these are evolutionary more conserved.
-We speculate that TMHs therefore are less prone to escape mutations that enable pathogens to evade T cell responses.
+and for both MHC-I and MHC-II. Moreover, we show that TMHs are evolutionarily more conserved, 
+because single nucleotide polymorphisms (SNPs) are present 
+relatively less frequently in TMH-coding chromosomal regions 
+compared to regions coding for extracellular and cytoplasmic protein regions. 
+Thus, our findings suggest that 
+both cytolytic and helper T cells respond more to membrane proteins, 
+because these are evolutionary more conserved. 
+We speculate that TMHs therefore are less prone to escape mutations 
+that enable pathogens to evade T cell responses.
 
@@
 epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Abbreviations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[h]
+  \begin{tabular}{ll}
+    Abbreviation & Full               \\ 
+    \hline
+    MAP & Membrane-associated protein \\
+    TMH & Transmembrane helix         \\
+    TMP & Transmembrane protein
+  \end{tabular}
+\end{table}
 
@@
 all peptides within a \emph{proteome} 
-(see tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
+(see supplementary Tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
 whereas the previous study defined
@@
 The exact amounts can be found in the supplementary materials,
-tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
+Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
 
@@
     The red lines indicate the percentages as expected by chance.
-    See supplementary tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+    See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
     for the exact TMH and  epitope counts.
@@
     and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
-    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. Labels are shorthand for the HLA haplotypes,
-    see Table \ref{tab:haplotype_abbreviations} for the names.
+    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. 
+    Labels are shorthand for the HLA haplotypes,
+    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.
   }
@@
 except for \verb;HLA-DRB3*0101; in \emph{M. tuberculosis}.
-See supplementary table \ref{tab:tmh_binders_mhc2} 
+See supplementary Table \ref{tab:tmh_binders_mhc2} 
 for the exact TMH and epitope counts.
@@
 epitopes derived from TMHs are presented in both MHC-I and MHC-II.
-See the supplementary materials (table \ref{tab:elution}) for the exact values.
+See the supplementary Table \ref{tab:elution} for the exact values.
 
@@
 of which 6,062 were located in predicted TMPs.
-See tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
+See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
 for the detailed numbers and distributions of SNPs.
@@
 
-Table \ref{tab:f_tmh} shows an overview of the findings,
+Supplementary Table \ref{tab:f_tmh} shows an overview of the findings,
 where a target specifies the source of the proteome,
@@
 is used to determine if a peptide binds to the haplotype's MHC
-for MHC-I (see table \ref{tab:ic50_binders_mhc1}) and 
-MHC-II (see table \ref{tab:ic50_binders_mhc2}).
+for MHC-I (see supplementary Table \ref{tab:ic50_binders_mhc1}) and 
+MHC-II (see supplementary Table \ref{tab:ic50_binders_mhc2}).
 
@@
 
-Table \ref{tab:haplotype_abbreviations} shows the
+Supplementary Table \ref{tab:haplotype_abbreviations} shows the
 shorthand notation for the HLA haplotypes.
@@
 
-Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+Supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
 show the exact number of binders, binders that overlap with TMHs
@@
 
-See tables \ref{tab:ncbi_counts_1} and tables \ref{tab:ncbi_counts_2}
+See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}
 for an overview of all amounts.
-In table \ref{tab:ncbi_counts_1} there are multiple instances where
+In supplementary Table \ref{tab:ncbi_counts_1} there are multiple instances where
 the amounts are expected to add up, yet don't, as one SNP can work on
@@
 
-Table \ref{tab:snp_stats} shows the statistics for all
-SNPs, where tables \ref{tab:snp_stats_per_spanner_single}
+Supplementary Table \ref{tab:snp_stats} shows the statistics for all
+SNPs, where supplementary Tables \ref{tab:snp_stats_per_spanner_single}
 and \ref{tab:snp_stats_per_spanner_multi} show the

### 2021-05-02 bbbq_article.tex
Processed feedback Frans. Thanks @fransbianchi. Fix #218

@@
 TMHs can also be predicted with high accuracy from a protein sequence 
-by bioinformatics approaches \cite{krogh2001predicting,bianchi2017,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient},
+by bioinformatics approaches \cite{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient},
 for example, \cite{jones2007improving} found that,
@@
 many of the amino acids genuinely present in TMHs 
-are limited to the ones with hydrophobic side chains \cite{jones1994model}.
+are limited to the ones with hydrophobic side chains 
+\cite{hessa2007molecular,jones1994model}.
 Therefore, we speculated that the TMHs of pathogens 
@@
 one finds 978 SNPs in TMHs 
-of single-spanners (although this is not significant)
+of single-spanners
 and 911 SNPs in TMHs of multi-spanners.
@@
 from a set of multiple tools, 
-NetMHCIIpan \cite{nielsen2008quantitative,karosiene2013netmhciipan}
+\mbox{NetMHCIIpan} \cite{nielsen2008quantitative,karosiene2013netmhciipan}
 made the most accurate predictions.
@@
 Figure \ref{fig:snps_per_gene_name_ncbi} shows the distribution of the
-number of SNPs per gene name, at 2020-12-14.
+number of SNPs per gene name, at the date we started the experiment,
+at December 14th 2020.
 

### 2021-05-06 split_pdf.sh
Can split PDF

@@ -1,0 +1,13 @@
+#!/bin/bash
+#
+# From the PDF 'bbbq_article.pdf', create two new files:
+#
+# (1) 'bbbq_article_main.pdf', the paper,  
+# (2) 'bbbq_article_supplementary_materials.pdf', the supplementary materials
+#
+# usage:
+#
+#   ./split_pdf.sh
+#
+pdftk bbbq_article.pdf cat 1-31 output bbbq_article_main.pdf
+pdftk bbbq_article.pdf cat 32-54 output bbbq_article_supplementary_materials.pdf

### 2021-05-08 old/covid_dataset_strength/covid_dataset_strength.Rmd
Move obsolete files to 'old' folder

--- 2020-07-16 covid_dataset_strength/covid_dataset_strength.Rmd
@@
   mutate(perc = n_distinct / n)
-perc_unique <- mean(t_unique$perc)
+perc_unique_protein <- mean(t_unique$perc)
 ```
 
-Of all proteins, on average, `r perc_unique` are unique.
+Of all proteins, on average, `r perc_unique_protein` are unique.
+We are looking, however, for unique proteomes. The percentage
+of unique proteomes is the product of
 
 ```{r}
-n_unique_sequences <- perc_unique * n_sequences
+percs_duplicate_protein <- 1.0 - t_unique$perc
+perc_duplicate_proteome <- prod(percs_duplicate_protein)
+perc_unique_proteome <- 1.0 - perc_duplicate_proteome
+```
+
+```{r}
+n_unique_sequences <- perc_unique_proteome * n_sequences
 ```

### 2021-05-11 split_pdf.sh
PDF with figures

@@ -1,13 +1,15 @@
 #!/bin/bash
 #
-# From the PDF 'bbbq_article.pdf', create two new files:
+# From the PDF 'bbbq_article.pdf', create these new files:
 #
 # (1) 'bbbq_article_main.pdf', the paper,  
-# (2) 'bbbq_article_supplementary_materials.pdf', the supplementary materials
+# (2) 'bbbq_article_figures.pdf', the figures of the main part,  
+# (3) 'bbbq_article_supplementary_materials.pdf', the supplementary materials
 #
 # usage:
 #
 #   ./split_pdf.sh
 #
-pdftk bbbq_article.pdf cat 1-31 output bbbq_article_main.pdf
-pdftk bbbq_article.pdf cat 32-54 output bbbq_article_supplementary_materials.pdf
+pdftk bbbq_article.pdf cat 1-27 output bbbq_article_main.pdf
+pdftk bbbq_article.pdf cat 29-32 output bbbq_article_figures.pdf
+pdftk bbbq_article.pdf cat 33-55 output bbbq_article_supplementary_materials.pdf

### 2021-06-13 bbbq_article.tex
Can split PDF
Can make
Figures are separated from text
Main, figures and supmat is nicely separated
Add black-and-white imageses, progress https://github.com/richelbilderbeek/bbbq_article/issues/221

@@
 
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{0.8\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_panel.png}
-    \label{fig:bbbq_1_smart_results}
-  \end{subfigure}
-
-  \vfill
-
-  \begin{subfigure}[t]{0.35\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
-    \label{fig:hydrophobicity_1}
-  \end{subfigure}  
-  \hfill
-  \begin{subfigure}[t]{0.35\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
-    \label{fig:hydrophobicity_2}
-  \end{subfigure}  
-  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes}
-    \textbf{(A)} 
-    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
-    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
-    row) and \emph{M. tuberculosis} (bottom row).
-    The red lines indicate the percentages as expected by chance.
-    See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
-    for the exact TMH and  epitope counts.
-    \textbf{(B-C)}
-    Correlation between the percentages of predicted TMH-derived epitopes
-    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
-    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. 
-    Labels are shorthand for the HLA haplotypes,
-    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.
-  }
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
-  \caption{
-    \textbf{
-      Robust prediction that TMH epitopes are presented \emph{in vivo}.
-    }
-    Bars show the percentage of peptides obtained from elution
-    studies that is derived from a TMH, for MHC-I and -II,
-    for two TMH prediction tools.
-  }
-  \label{fig:elution}
-\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
-    \label{fig:conservation}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
-    \label{fig:f_snps_found_and_expected}
-  \end{subfigure}
-
-   \vfill
-
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
-    \label{fig:f_tmh_ncbi}
-  \end{subfigure}
-
-  \caption{ \textbf{Evolutionary conservation of human TMHs.}
-    \textbf{(A)} 
-    The number of SNPs in TMHs as expected by chance (left bar) 
-    and found in the dbSNP database (right bar).
-    Percentages show the relative conservation
-    of SNPs in TMHs found.
-    \textbf{(B)}
-    Percentage of SNPs found in TMHs.
-    Each point shows for one protein the predicted percentage of
-    TMH ($x$-axis) and the observed occurrence of SNPs being located
-    within a TMH ($y$-axis).
-    The dashed diagonal line shows the line of equality (i.e.,
-    equal conservation of TMHs and soluble protein regions). The red line indicates a linear fit, the gray area its 95\% confidence interval.
-    \textbf{(C)}
-    Distribution of the percentages of TMH in the TMPs used in this study.
-  }
-\end{figure}
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
-    \label{fig:conservation_per_spanner}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi_per_spanner.png}
-    \label{fig:f_tmh_ncbi_per_spanner}
-  \end{subfigure}
-
-  \vfill
-
-  \begin{subfigure}[t]{\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
-    \label{fig:f_snps_found_and_expected_per_spanner}
-  \end{subfigure}
-  \caption{ \textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
-      \textbf{(A)} 
-      The number of SNPs in TMHs as expected by chance 
-      and observed in the dbSNP database, 
-      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
-      Percentages show the relative conservation
-      of SNPs in TMHs found.
-      \textbf{(B)} 
-      Distribution of the proportion of amino acids residing
-      in the plasma membrane. 
-      \textbf{(C)} 
-      Percentage of SNPs found in TMPs predicted to have only a single
-      (left) or multiple (right) TMHs.
-        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
-The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
-The red line indicates a linear fit, the gray area its 95\% confidence interval.
-     }
-\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\section{Figures}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+\begin{figure}[!htbp]
+  \centering
+  \begin{subfigure}[t]{0.8\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_panel_bw.png}
+    \label{fig:bbbq_1_smart_results}
+  \end{subfigure}
+
+  \vfill
+
+  \begin{subfigure}[t]{0.35\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1_bw.png}
+    \label{fig:hydrophobicity_1}
+  \end{subfigure}  
+  \hfill
+  \begin{subfigure}[t]{0.35\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2_bw.png}
+    \label{fig:hydrophobicity_2}
+  \end{subfigure}  
+  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes}
+    \textbf{(A)} 
+    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
+    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
+    row) and \emph{M. tuberculosis} (bottom row).
+    The red lines indicate the percentages as expected by chance.
+    See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+    for the exact TMH and  epitope counts.
+    \textbf{(B-C)}
+    Correlation between the percentages of predicted TMH-derived epitopes
+    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
+    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. 
+    Labels are shorthand for the HLA haplotypes,
+    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.
+  }
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
+  \caption{
+    \textbf{
+      Robust prediction that TMH epitopes are presented \emph{in vivo}.
+    }
+    Bars show the percentage of peptides obtained from elution
+    studies that is derived from a TMH, for MHC-I and -II,
+    for two TMH prediction tools.
+  }
+  \label{fig:elution}
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+\begin{figure}[!htbp]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
+    \label{fig:conservation}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_bw.png}
+    \label{fig:f_snps_found_and_expected}
+  \end{subfigure}
+
+   \vfill
+
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
+    \label{fig:f_tmh_ncbi}
+  \end{subfigure}
+
+  \caption{ \textbf{Evolutionary conservation of human TMHs.}
+    \textbf{(A)} 
+    The number of SNPs in TMHs as expected by chance (left bar) 
+    and found in the dbSNP database (right bar).
+    Percentages show the relative conservation
+    of SNPs in TMHs found.
+    \textbf{(B)}
+    Percentage of SNPs found in TMHs.
+    Each point shows for one protein the predicted percentage of
+    TMH ($x$-axis) and the observed occurrence of SNPs being located
+    within a TMH ($y$-axis).
+    The dashed diagonal line shows the line of equality (i.e.,
+    equal conservation of TMHs and soluble protein regions). The red line indicates a linear fit, the gray area its 95\% confidence interval.
+    \textbf{(C)}
+    Distribution of the percentages of TMH in the TMPs used in this study.
+  }
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+\begin{figure}[!htbp]
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation_per_spanner.png}
+    \label{fig:conservation_per_spanner}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi_per_spanner.png}
+    \label{fig:f_tmh_ncbi_per_spanner}
+  \end{subfigure}
+
+  \vfill
+
+  \begin{subfigure}[t]{\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner_bw.png}
+    \label{fig:f_snps_found_and_expected_per_spanner}
+  \end{subfigure}
+  \caption{ \textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
+      \textbf{(A)} 
+      The number of SNPs in TMHs as expected by chance 
+      and observed in the dbSNP database, 
+      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
+      Percentages show the relative conservation
+      of SNPs in TMHs found.
+      \textbf{(B)} 
+      Distribution of the proportion of amino acids residing
+      in the plasma membrane. 
+      \textbf{(C)} 
+      Percentage of SNPs found in TMPs predicted to have only a single
+      (left) or multiple (right) TMHs.
+        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
+The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
+The red line indicates a linear fit, the gray area its 95\% confidence interval.
+     }
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
 \appendix
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/fig_1a.png}
+    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/fig_1a_bw.png}
     \label{fig:bianch_et_al_2017_1a}
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1.png}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1_bw.png}
     \label{fig:bilderbeek_et_al_2021_1a}
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_normalized.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_normalized_bw.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_normalized.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_normalized_bw.png}
   \caption{

### 2021-06-13 bbbq_article.tex
Add black-and-white images, progress https://github.com/richelbilderbeek/bbbq_article/issues/221
No reference to the color red

@@
     row) and \emph{M. tuberculosis} (bottom row).
-    The red lines indicate the percentages as expected by chance.
+    The dashed lines indicate the percentages as expected by chance.
     See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
@@
     and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
-    and MHC-II haplotypes \textbf{(C)}. Red curve: linear regression analysis. 
+    and MHC-II haplotypes \textbf{(C)}. Diagonal line: linear regression analysis. 
     Labels are shorthand for the HLA haplotypes,
@@
     The dashed diagonal line shows the line of equality (i.e.,
-    equal conservation of TMHs and soluble protein regions). The red line indicates a linear fit, the gray area its 95\% confidence interval.
+    equal conservation of TMHs and soluble protein regions). 
+    The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
     \textbf{(C)}
@@
 The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
-The red line indicates a linear fit, the gray area its 95\% confidence interval.
+        The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
      }
@@
     Results for \cite{bianchi2017}. 
-    Red lines denotes the coincidence interval.
+    Dashed lines denotes the coincidence interval.
     \textbf{(B)}
     Results for this study.
-    Red line denotes the percentage as expected by chance.
+    Dashed line denotes the percentage as expected by chance.
   }
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation_per_haplotype.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation_per_haplotype_bw.png}
   \caption{

### 2021-07-24 bbbq_article.tex
Depend on beautier v2.6.2

@@
 will be more frequently presented by MHC-I 
-than expected based on their abundance \cite{bianchi2017}. 
+than expected based on their abundance \cite{bianchi2017},
+which is in line with a previous study 
+by Istrail et al \cite{istrail2004comparative},
+demonstrating that N-terminal signal sequences 
+are likely to be presented within major histocompatibility complexes, 
+due their hydrophobic nature. 
 Moreover, we showed that some well-known immunodominant peptides stem from TMHs. 

### 2021-07-27 create_numbered_figure_files.R
Progress with creating numbered figures

@@ -1,0 +1,11 @@
+#
+
+tex_lines <- readr::read_lines("bbbq_article.tex")
+includegraphics_lines <- stringr::str_subset(tex_lines, "^  .*includegraphics")
+png_filenames <- stringr::str_match(includegraphics_lines, "includegraphics.*\\{(.*)\\}")[, 2]
+testthat::expect_true(all(file.exists(png_filenames)))
+tiff_filenames <- stringr::str_replace(png_filenames, "png$", "tiff")
+
+tiff_filenames[!file.exists(tiff_filenames)]
+testthat::expect_true(all(file.exists(tiff_filenames)))
+

### 2021-08-06 bbbq_article.tex
Supmat has figures starting from S1
Progress, thanks @fransbianchi

@@
   \centering
+
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_bw.png}
+    \label{fig:f_snps_found_and_expected}
+  \end{subfigure}
+  \hfill
   \begin{subfigure}[t]{0.45\textwidth}
@@
   \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_bw.png}
-    \label{fig:f_snps_found_and_expected}
-  \end{subfigure}
 
@@
     \textbf{(A)} 
-    The number of SNPs in TMHs as expected by chance (left bar) 
-    and found in the dbSNP database (right bar).
-    Percentages show the relative conservation
-    of SNPs in TMHs found.
-    \textbf{(B)}
     Percentage of SNPs found in TMHs.
@@
     The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
+    \textbf{(B)}
+    The number of SNPs in TMHs as expected by chance (left bar) 
+    and found in the dbSNP database (right bar).
+    Percentages show the relative conservation
+    of SNPs in TMHs found.
     \textbf{(C)}
@@
 \section{Supplementary materials}
+
+% Figures start from one and are prepended with an S
+\renewcommand{\thefigure}{S\arabic{figure}}
+\setcounter{figure}{0}
+
+% Tables start from one and are prepended with an S
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{table}{0}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{IC50 values of binders per haplotype}
+\label{subsec:ic50s_per_haplotype}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Per target proteome (i.e. human, SARS-CoV-2, \emph{M tuberculosis}),
+we collected all 9-mers (for MHC-I) and 14-mers (for MHC-II),
+after removing the selenoproteins and proteins that are shorter
+than the epitope length.
+From these epitopes, per MHC haplotype,
+we predicted the IC50 (in nM) using \verb;epitope-prediction; (for MHC-I)
+and MHCnuggets (for MHC-II). 
+Here, we show the IC50 value per haplotype that
+is used to determine if a peptide binds to the haplotype's MHC
+for MHC-I (see supplementary Table \ref{tab:ic50_binders_mhc1}) and 
+MHC-II (see supplementary Table \ref{tab:ic50_binders_mhc2}).
+
+% tab:ic50_binders_mhc1
+\input{bbbq_1_smart_results/table_ic50_binders_mhc1_2.latex}
+
+% tab:ic50_binders_mhc2
+\input{bbbq_1_smart_results/table_ic50_binders_mhc2_2.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Counts}
+\label{subsec:counts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}
+for an overview of all amounts.
+In supplementary Table \ref{tab:ncbi_counts_1} there are multiple instances where
+the amounts are expected to add up, yet don't, as one SNP can work on
+multiple isoforms. For example, there are 9,621 unique SNPs 
+found in all proteins, of which 4,219 around found in MAPs 
+and 6,026 in TMPs. Apparently, 624 SNPs work on a set of isoforms that
+contains both MAPs and TMPs.
+
+% Label: tab:ncbi_counts_1
+\input{ncbi_peregrine_results/ncbi_counts_1.tex}
+
+% Label: tab:ncbi_counts_2
+\input{ncbi_peregrine_results/ncbi_counts_2.tex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Relative positions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snp_rel_pos.png}
+  \caption{
+    Distribution of the relative position of the SNPs used,
+    where a relative position of zero denotes the first amino
+    acid at the N-terminus, where a relative position of one
+    indicates the last residue at the C-terminus.
+  }
+  \label{fig:snp_rel_pos}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Label: tab:tmh_binders_mhc2
+\input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Elution studies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% tab:elution
+\input{bbbq_article_issue_157/results.tex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Label: tab:tmh_binders_mhc1
+\input{bbbq_1_smart_results/table_tmh_binders_mhc1_2.latex}
+
+Supplementary Table \ref{tab:haplotype_abbreviations} shows the
+shorthand notation for the HLA haplotypes.
+
+% Label: tab:haplotype_abbreviations
+\input{bbbq_1_smart_results/haplotype_abbreviations.tex}
+
+Supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+show the exact number of binders, binders that overlap with TMHs
+and the percentage of binders that overlap with TMHs, as
+visualized by figure \ref{fig:bbbq_1_smart_results}.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Elution studies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% tab:elution
-\input{bbbq_article_issue_157/results.tex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IC50 values of binders per haplotype}
-\label{subsec:ic50s_per_haplotype}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Per target proteome (i.e. human, SARS-CoV-2, \emph{M tuberculosis}),
-we collected all 9-mers (for MHC-I) and 14-mers (for MHC-II),
-after removing the selenoproteins and proteins that are shorter
-than the epitope length.
-From these epitopes, per MHC haplotype,
-we predicted the IC50 (in nM) using \verb;epitope-prediction; (for MHC-I)
-and MHCnuggets (for MHC-II). 
-Here, we show the IC50 value per haplotype that
-is used to determine if a peptide binds to the haplotype's MHC
-for MHC-I (see supplementary Table \ref{tab:ic50_binders_mhc1}) and 
-MHC-II (see supplementary Table \ref{tab:ic50_binders_mhc2}).
-
-% tab:ic50_binders_mhc1
-\input{bbbq_1_smart_results/table_ic50_binders_mhc1_2.latex}
-
-% tab:ic50_binders_mhc2
-\input{bbbq_1_smart_results/table_ic50_binders_mhc2_2.latex}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Presentation of TMH-derived epitopes}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Supplementary Table \ref{tab:haplotype_abbreviations} shows the
-shorthand notation for the HLA haplotypes.
-
-% Label: tab:haplotype_abbreviations
-\input{bbbq_1_smart_results/haplotype_abbreviations.tex}
-
-Supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
-show the exact number of binders, binders that overlap with TMHs
-and the percentage of binders that overlap with TMHs, as
-visualized by figure \ref{fig:bbbq_1_smart_results}.
-
-% Label: tab:tmh_binders_mhc1
-\input{bbbq_1_smart_results/table_tmh_binders_mhc1_2.latex}
-
-% Label: tab:tmh_binders_mhc2
-\input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}
-for an overview of all amounts.
-In supplementary Table \ref{tab:ncbi_counts_1} there are multiple instances where
-the amounts are expected to add up, yet don't, as one SNP can work on
-multiple isoforms. For example, there are 9,621 unique SNPs 
-found in all proteins, of which 4,219 around found in MAPs 
-and 6,026 in TMPs. Apparently, 624 SNPs work on a set of isoforms that
-contains both MAPs and TMPs.
-
-% Label: tab:ncbi_counts_1
-\input{ncbi_peregrine_results/ncbi_counts_1.tex}
-
-% Label: tab:ncbi_counts_2
-\input{ncbi_peregrine_results/ncbi_counts_2.tex}
-
 Figure \ref{fig:snps_per_gene_name_ncbi} shows the distribution of the
@@
 \end{figure}
-
-%Figure \ref{fig:f_tmh_ncbi} shows the distribution 
-%of the percentages of TMH of transmembrane proteins.
-%
-%\begin{figure}[!htbp]
-%  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
-%  \caption{
-%    Distribution of the percentages of TMH of transmembrane proteins.
-%    Dashed vertical line: mean \% TMH in all proteins.
-%    Dotted vertical line: mean \% TMH in TMPs.
-%  }
-%  \label{fig:f_tmh_ncbi}
-%\end{figure}
 
@@
 
-\begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snp_rel_pos.png}
-  \caption{
-    Distribution of the relative position of the SNPs used,
-    where a relative position of zero denotes the first amino
-    acid at the N-terminus, where a relative position of one
-    indicates the last residue at the C-terminus.
-  }
-  \label{fig:snp_rel_pos}
-\end{figure}
-
 % \paragraph{Relative position of SNPs}
@@
 \input{ncbi_peregrine_results/snp_stats_per_spanner_multi.tex}
-
-%
-% Notes to self
-%
-
-%\begin{sidewaystable}
-%  \centering
-%  ... centered table here
-%\end{sidewaystable}
-
-% \scalebox{0.7}{
-%  ... scaled table here
-% }
-
-%{\tiny
-%  ... something with a tiny font
-%}
-
-% Instead of non-TMH use 'soluble protein region'
-% or 'cytosolic or extracellular protein region'

### 2021-08-06 bbbq_article.tex
Follow the correct numbering, thanks @fransbianchi

@@
 
+%
+% Figure 1
+%
 \begin{figure}[!htbp]
@@
 
+%
+% Figure 2
+%
 \begin{figure}[!htbp]
@@
 
+%
+% Figure 3
+%
 \begin{figure}[!htbp]
@@
 
+%
+% Figure 4
+%
 \begin{figure}[!htbp]
   \centering
+
+  \begin{subfigure}[t]{\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner_bw.png}
+    \label{fig:f_snps_found_and_expected_per_spanner}
+  \end{subfigure}
+
+  \vfill
+
   \begin{subfigure}[t]{0.45\textwidth}
@@
 
-  \vfill
-
-  \begin{subfigure}[t]{\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner_bw.png}
-    \label{fig:f_snps_found_and_expected_per_spanner}
-  \end{subfigure}
   \caption{ \textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
       \textbf{(A)} 
+      Percentage of SNPs found in TMPs predicted to have only a single
+      (left) or multiple (right) TMHs.
+        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
+The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
+        The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
+      \textbf{(B)} 
       The number of SNPs in TMHs as expected by chance 
@@
       of SNPs in TMHs found.
-      \textbf{(B)} 
+      \textbf{(C)} 
       Distribution of the proportion of amino acids residing
       in the plasma membrane. 
-      \textbf{(C)} 
-      Percentage of SNPs found in TMPs predicted to have only a single
-      (left) or multiple (right) TMHs.
-        Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
-The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
-        The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
      }

### 2021-08-18 bbbq_article.tex
'parse_beast_log' will be deprecated. Use 'parse_beast_tracelog_file' instead

@@
 and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
-The 13 MHC-I haplotypes used in this study are the same as 
-used in the previous study \cite{bianchi2017}.
 For MHC-II, haplotypes were selected with a phenotypic frequency 
@@
 
-In previous work, it was found that the over-presentation of TMH-derived peptides can be
-explained from the hydrophobicity of the MHC-I binding cleft \cite{bianchi2017}.
-Here, a similar analysis was applied, by correlating the percentage of predicted TMH-derived epitopes versus the mean hydrophobicity of all 
-peptides.
-
-This study differs in one important aspect from our previous work \cite{bianchi2017}. 
-The definition of a binder differs from \cite{bianchi2017}:
-in the current study, a peptide is called a binder if, for a certain haplotype, 
+[SOMETHING HERE] A peptide is called a binder if, for a certain haplotype, 
 any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 

### 2021-08-18 bbbq_article.tex
Process feedback, thanks @fransbianchi

@@
 
-[SOMETHING HERE] A peptide is called a binder if, for a certain haplotype, 
+We define a protein to be a binder if, for a certain haplotype, 
 any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
@@
 (see supplementary Tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
-whereas the previous study defined
+this differs from our previous study where we defined
 a binder as having an IC50 in the lowest 2\% 
@@
 For MHC-II, 62\% of the variability is explained by hydrophobicity. 
-This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.
+This indicates that TMH-derived peptides are over-presented, 
+because the peptide binding grooves of most MHC-I and -II haplotypes 
+are relatively hydrophobic.
 
@@
     \caption{}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
+    \label{fig:f_tmh_ncbi}
+  \end{subfigure}
+
+   \vfill
+
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
     \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
     \label{fig:conservation}
-  \end{subfigure}
-
-   \vfill
-
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
-    \label{fig:f_tmh_ncbi}
   \end{subfigure}
@@
     \textbf{(B)}
+    Distribution of the percentages of TMH in the TMPs used in this study.
+    \textbf{(C)}
     The number of SNPs in TMHs as expected by chance (left bar) 
@@
     of SNPs in TMHs found.
-    \textbf{(C)}
-    Distribution of the percentages of TMH in the TMPs used in this study.
   }

### 2021-08-18 bbbq_article.tex
No empty headers in supmat, thanks @fransbianchi

@@
 
+% Process all floats before going to a next page
+\clearpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See Supplementary Figure \ref{fig:snp_rel_pos}
+for the distribution of the relative position of the SNPs.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See supplementary Table \ref{tab:tmh_binders_mhc2} 
+for the percentage of MHC-II 14-mers overlapping with TMH.
 
@@
 \input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}
+
+% Process all floats before going to a next page
+\clearpage
 
@@
 
+See supplementary Table \ref{tab:elution} 
+for the percentages of epitopes derived from a TMH 
+found in the two elution studies, 
+for the two different kind of topology prediction tools.
+
 % tab:elution
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See supplementary Table \ref{tab:tmh_binders_mhc1} 
+for the percentage of MHC-I 9-mers overlapping with TMH.
 
@@
 visualized by figure \ref{fig:bbbq_1_smart_results}.
+
+% Process all floats before going to a next page
+\clearpage
 
@@
 \end{figure}
+
+% Process all floats before going to a next page
+\clearpage
 
@@
 \input{bbbq_1_smart_results/table_f_tmh_2.latex}
+
+% Process all floats before going to a next page
+\clearpage
 
@@
 \input{ncbi_peregrine_results/snp_stats_per_spanner_multi.tex}
+
+% Process all floats before going to a next page
+\clearpage

### 2021-08-27 bbbq_article.tex
Processed submat feedback, thanks @fransbianchi
Link to Dryad

@@
 
-All code is archived at \url{http://github.com/richelbilderbeek/someplace},
-with DOI \url{https://doi.org/12.3456/zenodo.1234567}.
+All code, intermediate and final results are archived at 
+Dryad at \url{https://doi.org/10.5061/dryad.fn2z34tv1}.
 
@@
 Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
-for a binder per haplotype. We believe that our revised definition is more correct, as it overcomes bias from proteins with very low numbers of peptides and/or MHC-predicted binders.
+for a binder per haplotype.
 

### 2021-09-28 bbbq_article.tex
Store data on GitHub
Process reviewers' feedback

@@
 \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
-For example, \cite{tang2011genome} found that, 
+For example, S. Tang et al. \cite{tang2011genome} found that, 
 of the 432 peptides that were predicted to bind to MHC,
@@
 All code, intermediate and final results are archived at 
-Dryad at \url{https://doi.org/10.5061/dryad.fn2z34tv1}.
+\url{https://github.com/richelbilderbeek/bbbq_article}.
 

### 2021-10-06 bbbq_article.tex
Explain how we chose the MHC-I allotypes, fix #234

@@
 
-Many studies are aimed at identifying the repertoire of epitopes that are presented in any MHC haplotype and determining which epitopes will result in an immune response, 
+Many studies are aimed at identifying the repertoire of epitopes that are presented 
+in any MHC haplotype and determining which epitopes will result in an immune response, 
 as this will for instance aid the design of vaccines. 
@@
 and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
-For MHC-II, haplotypes were selected with a phenotypic frequency 
-of at least 14\% in the human population \cite{greenbaum2011functional},
-resulting in 21 MHC-II haplotypes.
+Both MCH-I and MHC-II haplotypes were selected to have a high occurrance,
+where the MHC-I haplotypes are the same 13 haplothose used in our earlier study \cite{bianchi2017},
+and the 21 MHC-II, haplotypes have a phenotypic frequency 
+of 14\% or more in the human population \cite{greenbaum2011functional}.
 

### 2021-10-11 bbbq_article.tex
Use 80 chars per line
Use color, progress #237

@@
 
-For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}.
+For MHC-I, we previously showed that the over-presentation of TMH-derived 
+peptides is caused by the hydrophobicity of the peptide binding 
+grooves \cite{bianchi2017}. 
 Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
 show the extent of over-presentation
-of TMH-derived epitopes as a function of the hydrophobicity preference score for the different haplotypes.
+of TMH-derived epitopes as a function of the hydrophobicity preference score 
+for the different haplotypes.
 An assumed linear correlation explains 88\% of the variability in MHC-I.
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_panel_bw.png}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_panel.png}
     \label{fig:bbbq_1_smart_results}
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1_bw.png}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc1.png}
     \label{fig:hydrophobicity_1}
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2_bw.png}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_hydrophobicity_mhc2.png}
     \label{fig:hydrophobicity_2}
@@
     and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
-    and MHC-II haplotypes \textbf{(C)}. Diagonal line: linear regression analysis. 
+    and MHC-II haplotypes \textbf{(C)}. 
+    Diagonal line: linear regression analysis. 
     Labels are shorthand for the HLA haplotypes,
@@
     \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_bw.png}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected.png}
     \label{fig:f_snps_found_and_expected}
@@
     equal conservation of TMHs and soluble protein regions). 
-    The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
+    The diagonal lines indicates a linear fit, 
+    the gray area its 95\% confidence interval.
     \textbf{(B)}
@@
     \caption{}
-    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner_bw.png}
+    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_per_spanner.png}
     \label{fig:f_snps_found_and_expected_per_spanner}
@@
 The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
-        The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
+      The diagonal lines indicate a linear fit, the gray areas their 95\% confidence intervals.
       \textbf{(B)} 
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/fig_1a_bw.png}
+    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/fig_1a.png}
     \label{fig:bianch_et_al_2017_1a}
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1_bw.png}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1.png}
     \label{fig:bilderbeek_et_al_2021_1a}
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_normalized_bw.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_normalized.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_normalized_bw.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_normalized.png}
   \caption{
@@
 \begin{figure}[!htbp]
-  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation_per_haplotype_bw.png}
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation_per_haplotype.png}
   \caption{

### 2021-10-11 bbbq_article.tex
Fixed legends (for color pictures), fix #232

@@
     and MHC-II haplotypes \textbf{(C)}. 
-    Diagonal line: linear regression analysis. 
+    Diagonal red line: linear regression analysis. 
     Labels are shorthand for the HLA haplotypes,
@@
     equal conservation of TMHs and soluble protein regions). 
-    The diagonal lines indicates a linear fit, 
+    The diagonal red line indicates a linear fit, 
     the gray area its 95\% confidence interval.
@@
 The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
-      The diagonal lines indicate a linear fit, the gray areas their 95\% confidence intervals.
+      The diagonal red lines indicate a linear fit, the gray areas their 95\% confidence intervals.
       \textbf{(B)} 

### 2021-10-18 bbbq_article.tex
Update Bianchi et al., 2017 results, by re-doing the analysis

@@
 and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
-Both MCH-I and MHC-II haplotypes were selected to have a high occurrance,
-where the MHC-I haplotypes are the same 13 haplothose used in our earlier study \cite{bianchi2017},
+Both MCH-I and MHC-II haplotypes were selected to have a high occurrence,
+where the MHC-I haplotypes are the same 13 haplotypes as 
+those used in our earlier study \cite{bianchi2017},
 and the 21 MHC-II, haplotypes have a phenotypic frequency 
@@
 which contains 650 million 
-catalogued non-redundant humane variations (called RefSNPs,
+cataloged non-redundant humane variations (called RefSNPs,
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names, \cite{brown2015gene})
@@
     \caption{}
-    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/fig_1a.png}
+    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/figure-1-a.png} % use same naming as them
     \label{fig:bianch_et_al_2017_1a}

### 2021-10-18 create_numbered_figure_files.R

@@ -13,9 +13,15 @@ tiff_filenames <- stringr::str_replace(png_filenames, "png$", "tiff")
 tiff_filenames[!file.exists(tiff_filenames)]
 testthat::expect_true(all(file.exists(tiff_filenames)))
 
+
 figures <- tibble::tibble(
-  number = seq(1, 4),
-  n_sub = c(3, 1, 3, 4)
+  number = c(
+    paste0(1, letters[1:3]),
+    paste0(2, letters[1]),
+    paste0(3, letters[1:3]),
+    paste0(4, letters[1:4])
+  ),
+  filename = NA
 )
 HIERO
 

### 2021-10-18 create_numbered_figure_files.R
Add 'fig_f_tmh_2_human_mhc1.tiff'

@@ -12,7 +12,8 @@ tiff_filenames <- stringr::str_replace(png_filenames, "png$", "tiff")
 
 tiff_filenames[!file.exists(tiff_filenames)]
 testthat::expect_true(all(file.exists(tiff_filenames)))
-
+# fig_f_tmh_2_human_mhc1.png is created
+# fig_f_tmh_2_human_mhc1.tiff is not
 
 figures <- tibble::tibble(
   number = c(

### 2021-10-29 bbbq_article.tex
Analyses of chance versus observed epitopes must include uncertainly estimates, fix #233
Improvements by @fransbianchi

@@
 derived from TMHs are a major but overlooked source of MHC epitopes. 
-Our bioinformatics predictions indicate that TMH-derived epitopes 
-are presented to both cytolytic and helper T cells more often than expected by chance, 
+Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance, 
 regardless of the organism.
+Furthermore we see despite the bias towards hydrophilic peptides due to their enhanced ionization in mass spec that in several peptide elution studies still a higher fraction of TMH derived epitopes is found than predicted by chance.
+Therefor it's tempting to assume that this would also result in enhanced T-cell responses, assuming that the conservation of TMH's across organisms does not effect the deletion of such T-cells by central tolerance mechanisms.
 Moreover, reanalysis of peptide elution studies confirmed the presentation of TMH-derived epitopes.
@@
 % \paragraph{Over-presentation in silico}
-
-An important implication from the over-presentation of TMH-derived epitopes is that T cells will largely recognize 
-different protein regions than B cells. 
-Presentation of antigens by MHC-II is important 
-for the activation of naive B cells by helper T cells. 
-For this activation, 
-B cells first ingest antigen that is bound to their B cell receptor, 
-and subsequently present peptides derived from this antigen in MHC-II 
-to helper T cells. 
-Following their activation by the T cells, 
-B cells mature into plasma cells 
-and release antibodies which recognize the same part of the antigen 
-as the original B cell receptor. 
-B cell receptors and antibodies will thus recognize solvent-exposed regions 
-of antigens that are accessible for binding to the B cell receptor. 
-However, the results from our study predict that most MHC-II haplotypes 
-present relatively hydrophobic peptides,  which are less likely to be solvent-exposed. 
-It is unknown why B and T cells seem to predominantly recognize 
-different protein regions, but one possibility might be that 
-this lowers the chance of B cell mediated autoimmune diseases, 
-because auto-reactive B and T cells recognizing different parts 
-of the same antigen would need to be present for breakage of B cell tolerance.
+Although it's well known that T cells and B Cells seem to predominantly recognize different protein regions,
+this is enhanced by the over-presentation of TMH-derived epitopes, as TMH's buried in the lipid bilayer are unlikely to be recognized by BCR's.
 
@@
     row) and \emph{M. tuberculosis} (bottom row).
-    The dashed lines indicate the percentages as expected by chance.
+    The pair of dashed lines in each plot indicate the lower and upper bound of the 99\% confidence interval.
     See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}

### 2021-10-29 bbbq_article_issue_157/results.tex
Add two immunopeptidomes studies, progress #229

% latex table generated in R 4.1.1 by xtable 1.8-4 package
% Fri Oct 29 12:40:25 2021
\begin{table}[ht]
\centering
\begin{tabular}{llll}
  \hline
MHC class & Tool & Dataset & n \\ 
  \hline
I & PureseqTM & schellens & 1.38\% (109/7897) \\ 
  I & PureseqTM & iedb & 6.81\% (43/631) \\ 
  I & TMHMM & schellens & 1.43\% (113/7897) \\ 
  I & TMHMM & iedb & 7.13\% (45/631) \\ 
  II & PureseqTM & bergseng & 3.92\% (498/12712) \\ 
  II & PureseqTM & iedb & 0.29\% (4/1364) \\ 
  II & TMHMM & bergseng & 3.96\% (504/12712) \\ 
  II & TMHMM & iedb & 1.39\% (19/1364) \\ 
   \hline
\end{tabular}
\caption{Percentage of epitopes derived from a TMH found in the two elution studies, for the two different kind of topology prediction tools. The values between braces show the the number of epitopes that were predicted to overlapping with a TMH per all epitopes that could be uniquely mapped to the representative human reference proteome.} 
\label{tab:elution}
\end{table}

### 2021-10-29 bbbq_article.tex
Add two immunopeptidomes studies, progress #229

@@
     }
-    Bars show the percentage of peptides obtained from elution
-    studies that is derived from a TMH, for MHC-I and -II,
-    for two TMH prediction tools.
+    Bars show the percentage of naturally ocurring peptides
+    that are derived from a TMH. These percentages are shown for
+    both MHC-I and -II (the columns), two sources of data (the
+    rows) and two TMH prediction tools (the two bars). 
+    IEDB = the Immune Epitope Database, non-IEDB = 
+    \cite{schellens2015comprehensive} for MHC-I and
+    \cite{bergseng2015different} for MHC-II.
   }

### 2021-10-29 bbbq_article.tex
Added IEDB to methods, progress #229

@@
 To obtain experimental evidence that epitopes derived from TMHs are presented in MHC,
+we used two sources of data of naturally occurring epitopes.
+The first type of data are two randomly picked 
 peptide elution studies for MHC-I \cite{schellens2015comprehensive} 
-and MHC-II \cite{bergseng2015different} were reanalyzed.
+and MHC-II \cite{bergseng2015different}.
+The second type of data is obtained from IEDB, 
+the Immune Epitope Databse \cite{vita2019immune}.
+All these sources contain sequences of naturally occurring
+epitopes. 
 For each of the detected epitopes,
@@
 overlapped with a TMH.
+For IEDB, we queried for epitope sequences presented by B and T cells
+in healthy humans, for the haplotypes used in this study.
 

### 2021-11-01 bbbq_article.tex
Improvements @fransbianchi

@@
 Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
-The peptide-binding grooves of most haplotypes 
+The peptide-binding grooves of the alleles of most haplotypes
 of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
@@
 In humans, the peptide binding groove of MHC-I is made by only the alpha subunit.
-There are three classical forms of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells.
+There are three classical haplotypes of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells.
 For MHC-II, both the alpha and the beta chains contribute to the peptide binding groove.
-There are three classical forms of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
+There are three classical haplotypes of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
 Each MHC complex can present a subset of all possible peptides.
@@
 to thousands of different alleles, 
-and each different HLA allele is called 
-an MHC haplotype and presents a different subset of peptides \cite{marsh2010nomenclature}.
+and each different allele presents a different subset of peptides \cite{marsh2010nomenclature}.
 
@@
 so-called escape mutations, 
-will not affect MHC presentation for all haplotypes present in the population \cite{sommer2005importance}.
+will not affect MHC presentation for all alleles present in the population \cite{sommer2005importance}.
 
@@
 Many studies are aimed at identifying the repertoire of epitopes that are presented 
-in any MHC haplotype and determining which epitopes will result in an immune response, 
+in any MHC haplotype allele and determining which epitopes will result in an immune response, 
 as this will for instance aid the design of vaccines. 
@@
 From these \emph{in silico} predictions, we conclude that TMH-derived
-epitopes are presented more often than expected by chance,
+epitopes are likely to be presented more often than expected by chance,
 in a human, viral and bacterial proteome, 
-and for most haplotypes of both MHC-I and II. 
+and for most alleles of both MHC-I and II. 
 We confirmed the presentation of TMH-derived peptides 
@@
 For each of these peptides, we determined if it overlapped with a predicted 
-TMH and if it was predicted to bind to each haplotype.
+TMH and if it was predicted to bind to the moest frequent alleles of each haplotype.
 
@@
 of the proteins within these proteomes.
-To predict the affinity of an epitope to a certain MHC haplotype,
+To predict the affinity of an epitope to a certain HLA allele,
  \verb;EpitopePrediction; \cite{bianchi2017} for MHC-I 
 and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
-Both MCH-I and MHC-II haplotypes were selected to have a high occurrence,
-where the MHC-I haplotypes are the same 13 haplotypes as 
-those used in our earlier study \cite{bianchi2017},
-and the 21 MHC-II, haplotypes have a phenotypic frequency 
+Both MCH-I and MHC-II alleles were selected to have a high occurrence,
+where the alleles of the MHC-I haplotypes are the alleles representing the 13 supertypes 
+with over 99.6\% coverage of the populations MHC-I repertoire as defined by \cite{lund2004definition} \cite{sette1999},
+and the 21 MHC-II haplotype alleles, have a phenotypic frequency 
 of 14\% or more in the human population \cite{greenbaum2011functional}.
@@
 overlapped with a TMH.
-For IEDB, we queried for epitope sequences presented by B and T cells
+For the second we queried IEDB for epitope sequences presented by B and T cells
 in healthy humans, for the haplotypes used in this study.
@@
 for a human, viral and bacterial proteome.
-Per MHC-I haplotype, it shows the percentage of binders that overlap with a TMH 
+Per MHC-I allele, it shows the percentage of binders that overlap with a TMH 
 with at least one residue.
@@
 likely as epitopes derived from soluble regions.
-For 11 out of 13 MHC-I haplotypes, TMH-derived epitopes are predicted to be presented more often 
+For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
 than the null expectation, for a human and bacterial proteome.
-For the viral proteome, 12 out of 13 haplotypes present
+For the viral proteome, 12 out of 13 alleles present
 TMH-derived epitopes more often than expected by chance.
@@
 We found that TMH-derived peptides are over-presented in all
-of the 21 MHC-II haplotypes, 
+of the 21 MHC-II alleles, 
 for a human, bacterial and viral proteome,
@@
 of TMH-derived epitopes as a function of the hydrophobicity preference score 
-for the different haplotypes.
+for the different alleles.
 An assumed linear correlation explains 88\% of the variability in MHC-I.
@@
 For MHC-I, peptides presented \emph{in vivo} 
-by the (humane) haplotypes HLA-A and B 
+by the (human) haplotypes HLA-A and B 
 were sequenced \cite{schellens2015comprehensive},
-for MHC-II these were haplotypes DQ2.5, DQ2.2, and 
+for MHC-II these were alleles DQ2.5, DQ2.2, and 
 DQ7.5 \cite{bergseng2015different}.

### 2021-11-03 bbbq_article.tex
Improvements by @fransbianchi

@@
 Furthermore we see despite the bias towards hydrophilic peptides due to their enhanced ionization in mass spec that in several peptide elution studies still a higher fraction of TMH derived epitopes is found than predicted by chance.
-Therefor it's tempting to assume that this would also result in enhanced T-cell responses, assuming that the conservation of TMH's across organisms does not effect the deletion of such T-cells by central tolerance mechanisms.
+Therefor it's tempting to assume that this would also result in enhanced T-cell responses, assuming that the conservation of TMH's across organisms does not significantly effect the deletion of such T-cells by central tolerance mechanisms.
 Moreover, reanalysis of peptide elution studies confirmed the presentation of TMH-derived epitopes.
@@
 Although our data show that 
-TMH-derived epitopes are presented in MHC-I and MHC-II, 
+TMH-derived epitopes are presented in all classical MHC-I and MHC-II haplotypes, 
 the molecular mechanisms of how integral membrane proteins are processed 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{T cells recognize different protein regions than B cells}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{Over-presentation in silico}
-Although it's well known that T cells and B Cells seem to predominantly recognize different protein regions,
-this is enhanced by the over-presentation of TMH-derived epitopes, as TMH's buried in the lipid bilayer are unlikely to be recognized by BCR's.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Evolutionary conservation of TMHs}

### 2021-11-03 bbbq_article.tex
Added stubs of new figures

@@
   \centering
-  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2a.png}
   \caption{
@@
     }
-    Bars show the percentage of naturally ocurring peptides
-    that are derived from a TMH. These percentages are shown for
-    both MHC-I and -II (the columns), two sources of data (the
-    rows) and two TMH prediction tools (the two bars). 
-    IEDB = the Immune Epitope Database, non-IEDB = 
-    \cite{schellens2015comprehensive} for MHC-I and
-    \cite{bergseng2015different} for MHC-II.
+    The percentage of epitopes for MHC-I and -II alleles 
+    that overlap with TMHs for the proteomes of human
+    that is predicted 
+    (outer bars, same as figure 1A for human only) 
+    and is observed to occur naturally (inner bars/lines)
+    (from elution studies)
+    per MHC allele
   }
@@
 
-See supplementary Table \ref{tab:elution} 
+See figure \ref{fig:elution_per_software} or
+supplementary Table \ref{tab:elution} 
 for the percentages of epitopes derived from a TMH 
@@
 \clearpage
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
+  \caption{
+    \textbf{
+      Robust prediction that TMH epitopes are presented \emph{in vivo}.
+    }
+    Bars show the percentage of naturally ocurring peptides
+    that are derived from a TMH. These percentages are shown for
+    both MHC-I and -II (the columns), two sources of data (the
+    rows) and two TMH prediction tools (the two bars). 
+    IEDB = the Immune Epitope Database, non-IEDB = 
+    \cite{schellens2015comprehensive} for MHC-I and
+    \cite{bergseng2015different} for MHC-II.
+  }
+  \label{fig:elution_per_software}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2b.png}
+  \caption{
+    \textbf{
+      Relationship between TMH epitopes that are predicted to
+      be presented versus that are presented \emph{in vivo}.
+    }
+    \richel{Unsure if there is enough data}
+  }
+  \label{fig:elution_versus}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2c.png}
+  \caption{
+    \textbf{
+      Relationship between TMH epitopes that are presented \emph{in vivo}.
+    }
+    \richel{TODO: MHC ligands only}
+  }
+  \label{fig:elution_mhc_ligands}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2d.png}
+  \caption{
+    \textbf{
+      Relationship between TMH epitopes that are presented \emph{in vivo}
+      in T cells.
+    }
+    \richel{Must mere T cell data to get more data}
+  }
+  \label{fig:elution_versus}
+\end{figure}
 

### 2021-11-04 bbbq_article.tex

@@
     The pair of dashed lines in each plot indicate the lower and upper bound of the 99\% confidence interval.
-    See supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+    See supplementary Tables \ref{tab:tmh_binders_mhc2} and \ref{tab:tmh_binders_mhc1}
     for the exact TMH and  epitope counts.
@@
 \clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Presentation of TMH-derived epitopes when two amino acids overlap}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In our experiment, we define a TMH-derived epitope as a peptide that overlaps
+with a TMH for at least one amino acid. One could argue that we should use
+a higher number of overlapping amino acids, so to make the epitopes more
+'transmembrane helix-ey'. We chose not too, for two reason: (1) epitopes that overlap with a TMH
+for 1 AA already, cannot be processed by the proteasome in a known and conventional
+way (2) whatever number of overlapping amino acids we use, we expect the pattern to be the same.
+However, using only 1 AA gives the most TMH-derived epitopes and hence the highest statistical
+power.
+
+To prove this point, we did exactly the same analysis 
+as shown in Figure \ref{fig:bbbq_1_smart_results},
+yet with defining a TMH-derived epitope as an epitope that overlaps with a TMH
+for at least 2 AAs, as shown in Figure \ref{fig:bbbq_1_smart_results_2aas}.
+As these two figures look identical \richel{TODO: they are indentical,
+need to redo the calculation then}, 
+we also added the counts as numbers,
+with Table \ref{tab:tmh_binders_mhc2_2aas} showing the same data
+as \ref{tab:tmh_binders_mhc2}, except the former uses 2 AAs overlap.
+Likewise, Table \ref{tab:tmh_binders_mhc1_2aas} showing the same data
+as \ref{tab:tmh_binders_mhc1}, except the former uses 2 AAs overlap.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_2aa_results/fig_f_tmh_2_panel.png}
+  \caption{
+    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
+    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
+    row) and \emph{M. tuberculosis} (bottom row).
+    The pair of dashed lines in each plot indicate the lower and upper bound of the 99\% confidence interval.
+    See supplementary Tables \ref{tab:tmh_binders_mhc2_2aas} and \ref{tab:tmh_binders_mhc1_2aas}
+    for the exact TMH and  epitope counts.
+  }
+  \label{fig:bbbq_1_smart_results_2aas}
+\end{figure}
+
+% Label: tab:tmh_binders_mhc2_2aas
+\input{bbbq_1_smart_2aa_results/table_tmh_binders_mhc2_2.latex}
+
+% Label: tab:tmh_binders_mhc1_2aas
+\input{bbbq_1_smart_2aa_results/table_tmh_binders_mhc1_2.latex}
+
+% Process all floats before going to a next page
+\clearpage

### 2021-11-04 bbbq_article.tex
Is it significant the overlapping of one residue for MHC binding? No. Fix https://github.com/richelbilderbeek/bbbq_article/issues/242

@@
 for at least 2 AAs, as shown in Figure \ref{fig:bbbq_1_smart_results_2aas}.
-As these two figures look identical \richel{TODO: they are indentical,
-need to redo the calculation then}, 
+As these two figures look identical, 
 we also added the counts as numbers,
@@
     The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
-    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
+    overlap with TMHs (for at least two amino acids) 
+    for the proteomes of SARS-CoV-2 (top row), human (middle 
     row) and \emph{M. tuberculosis} (bottom row).

### 2021-11-09 bbbq_article.tex
Describe IEDB and 'iedbr'

@@
 and MHC-II \cite{bergseng2015different}.
-The second type of data is obtained from IEDB, 
-the Immune Epitope Databse \cite{vita2019immune}.
+The second type of data is obtained from , 
+the Immune Epitope Database (IEDB) \cite{vita2019immune}.
 All these sources contain sequences of naturally occurring
@@
     Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
-    Call TMHMM from R                 & tmhmm                     & \cite{tmhmm} \\
-    Call PureseqTM from R             & pureseqtmr                & \cite{pureseqtmr} \\
-    Call NetMHCIIpan from R           & netmhc2pan                & \cite{netmhc2pan} \\
-    Combine all                       & bbbq                      & \cite{bbbq}
+    Call TMHMM from R                 & \verb;tmhmm;              & \cite{tmhmm} \\
+    Call PureseqTM from R             & \verb;pureseqtmr;         & \cite{pureseqtmr} \\
+    Call NetMHCIIpan from R           & \verb;netmhc2pan;         & \cite{netmhc2pan} \\
+    Work with IEDB                    & \verb;iedbr;              & \cite{iedbr} \\
+    Work with \verb;rentrez;          & \verb;sprentrez;          & \cite{sprentrez} \\
+    Combine all                       & \verb;bbbq;               & \cite{bbbq}
   \end{tabular}
@@
 \verb;rentrez; R package \cite{rentrez} was used
-that calls the NCBI website's API. To provide for a 
-stable user experience for all users, 
-this API limits the user to 3 calls per second.
+that calls the NCBI database's API. 
+The NCBI database provides a stable user experience for all users, 
+by limiting its API to 3 calls per second per user.
 Additionally, the API splits the result of a bigger
 query into multiple pages, each of which needs one API call.
-We wrote the \verb;sprentrez; package \cite{sprentrez} to provide for 
+The \verb;sprentrez; package \cite{sprentrez} provides for 
 bigger queries of multiple (and delayed) API calls.
+
+% \paragraph{IEDB software}
+
+To retrieve the data from the IEDB  databases \cite{vita2019immune}, 
+the \verb;iedbr; R package \cite{iedbr} was written,
+to calls the IEDB database's API. 
+Similar to the NCBI database,
+the IEDB has a limit to 1 call per second per user
+and allows a query results to return 10k results maximally.
+The \verb;iedbr; package \cite{iedbr} allows for bigger queries.
 

### 2021-11-09 bbbq_article.tex
Explain extensions to editor

@@
 the Immune Epitope Database (IEDB) \cite{vita2019immune}.
-All these sources contain sequences of naturally occurring
-epitopes. 
+All these sources contain sequences of naturally occurring epitopes. 
 For each of the detected epitopes,
@@
 overlapped with a TMH.
-For the second we queried IEDB for epitope sequences presented by B and T cells
-in healthy humans, for the haplotypes used in this study.
+As a second source of naturally occurring epitopes, 
+we used the IEDB \cite{vita2019immune} for epitope sequences presented in healthy humans
+by T cells, or as found in MHC ligand studies.
 

### 2021-11-10 bbbq_article.tex

@@
   \centering
+  \begin{subfigure}[t]{0.35\textwidth}
+    \centering
+    \caption{}
     \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2a.png}
+    \label{fig:2a}
+  \end{subfigure}  
+  \hfill
+  \begin{subfigure}[t]{0.35\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2b.png}
+    \label{fig:2b}
+  \end{subfigure}
+
+  \vfill
+
+  \begin{subfigure}[t]{0.35\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2c.png}
+    \label{fig:2c}
+  \end{subfigure}  
+  \hfill
+  \begin{subfigure}[t]{0.35\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2d.png}
+    \label{fig:2d}
+  \end{subfigure}  
   \caption{
@@
     The percentage of epitopes for MHC-I and -II alleles 
-    that overlap with TMHs for the proteomes of human
-    that is predicted 
-    (outer bars, same as figure 1A for human only) 
-    and is observed to occur naturally (inner bars/lines)
-    (from elution studies)
-    per MHC allele
+    that overlap with TMHs that are presented
+    (A) x
+    (B) x
+    (C) IEDB MHC ligands
+    (D) IEDB T cells
   }

### 2021-11-10 bbbq_article.tex
First sketch of figure 2

@@
   \centering
-  \begin{subfigure}[t]{0.35\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
     \centering
     \caption{}
-    \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2a.png}
+    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2a.png}
     \label{fig:2a}
@@
   \hfill
-  \begin{subfigure}[t]{0.35\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
     \centering
     \caption{}
-    \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2b.png}
+    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2b.png}
     \label{fig:2b}
@@
 
-  \begin{subfigure}[t]{0.35\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
     \centering
     \caption{}
-    \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2c.png}
+    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2c.png}
     \label{fig:2c}
@@
   \hfill
-  \begin{subfigure}[t]{0.35\textwidth}
+  \begin{subfigure}[t]{0.45\textwidth}
     \centering
     \caption{}
-    \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2d.png}
+    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2d.png}
     \label{fig:2d}
@@
     that overlap with TMHs that are presented
-    (A) x
-    (B) x
-    (C) IEDB MHC ligands
-    (D) IEDB T cells
+    \textbf{(A)} 
+    Observed and predicted percentage of TMH-derived epitopes
+    \textbf{(B)} 
+    Observed and predicted percentage of TMH-derived epitopes,
+    with a red linear regression line and 95\% confidence interval
+    \textbf{(C)} 
+    MHC ligands from the IEDB bind to TMH-derived epitopes.
+    The numbers above the bars denotes the number of epitopes
+    obtained using an MHC ligand assay
+    \textbf{(D)} 
+    T cells from the IEDB presnt TMH-derived epitopes.
+    The numbers above the bars denotes the number of epitopes
+    obtained using a T cell assay
   }

### 2021-11-13 bbbq_article.tex
Remove S2, S3, S4, S5

@@
 
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/results.png}
-  \caption{
-    \textbf{
-      Robust prediction that TMH epitopes are presented \emph{in vivo}.
-    }
-    Bars show the percentage of naturally ocurring peptides
-    that are derived from a TMH. These percentages are shown for
-    both MHC-I and -II (the columns), two sources of data (the
-    rows) and two TMH prediction tools (the two bars). 
-    IEDB = the Immune Epitope Database, non-IEDB = 
-    \cite{schellens2015comprehensive} for MHC-I and
-    \cite{bergseng2015different} for MHC-II.
-  }
-  \label{fig:elution_per_software}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2b.png}
-  \caption{
-    \textbf{
-      Relationship between TMH epitopes that are predicted to
-      be presented versus that are presented \emph{in vivo}.
-    }
-    \richel{Unsure if there is enough data}
-  }
-  \label{fig:elution_versus}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2c.png}
-  \caption{
-    \textbf{
-      Relationship between TMH epitopes that are presented \emph{in vivo}.
-    }
-    \richel{TODO: MHC ligands only}
-  }
-  \label{fig:elution_mhc_ligands}
-\end{figure}
-
-\begin{figure}[!htbp]
-  \centering
-  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2d.png}
-  \caption{
-    \textbf{
-      Relationship between TMH epitopes that are presented \emph{in vivo}
-      in T cells.
-    }
-    \richel{Must mere T cell data to get more data}
-  }
-  \label{fig:elution_versus}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-11-13 bbbq_article.tex
Move in vivo presentation of TMH-derived epitopes in T cells to SupMat and explain the rationale there

@@
   \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2d.png}
-    \label{fig:2d}
-  \end{subfigure}  
+
   \caption{
@@
     obtained using an MHC ligand assay
-    \textbf{(D)} 
-    T cells from the IEDB presnt TMH-derived epitopes.
-    The numbers above the bars denotes the number of epitopes
-    obtained using a T cell assay
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Presentation of TMH-derived epitopes \emph{in vivo}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Figure \ref{fig:elution_versus} shows the percentage of TMH-derived
+epitopes as presented \emph{in vivo} in T cells. 
+The data was obtained from the IEDB and includes only the MHC
+alleles as used in this study. As there are many (especially class II)
+MHC alleles, only a small percentage of the full IEDB data could be used.
+Although this figure demonstrates that TMH-derived epitopes are presented
+\emph{in vivo}, we excluded this figure from the main article, as these
+protocols select epitopes in a non-random fashion (unlike the
+MHC ligand assays, as presented in the main article).
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2d.png}
+  \caption{
+    \textbf{
+      T cells present TMH-derived epitopes \emph{in vivo}
+    }
+    The numbers above the bars denotes the number of epitopes
+    found in the IEDB for the MHC alleles used in this study
+  }
+  \label{fig:elution_versus}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Presentation of TMH-derived epitopes}

### 2021-11-13 bbbq_article.tex
Remove figure 2b, as figure 2a displays the same info, yet clearer

@@
     \caption{}
-    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2b.png}
-    \label{fig:2b}
-  \end{subfigure}
-
-  \vfill
-
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
     \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2c.png}
@@
     \textbf{(B)} 
-    Observed and predicted percentage of TMH-derived epitopes,
-    with a red linear regression line and 95\% confidence interval
-    \textbf{(C)} 
     MHC ligands from the IEDB bind to TMH-derived epitopes.

### 2021-11-13 bbbq_article.tex
Remove elution studies by Schellens and Bergseng

@@
 To obtain experimental evidence that epitopes derived from TMHs are presented in MHC,
-we used two sources of data of naturally occurring epitopes.  
-The first type of data are two randomly picked 
-peptide elution studies for MHC-I \cite{schellens2015comprehensive} 
-and MHC-II \cite{bergseng2015different}.
-The second type of data is obtained from , 
-the Immune Epitope Database (IEDB) \cite{vita2019immune}.
-All these sources contain sequences of naturally occurring epitopes. 
-For each of the detected epitopes,
+we used the Immune Epitope Database (IEDB) \cite{vita2019immune}.
+This database contains the sequences of thousands of experimentally
+established epitope sequences.
+For each of these epitopes,
 its possible location(s) in 
@@
 the topology of the proteins in which these epitopes were located 
-was predicted using both \verb;TMHMM; \cite{krogh2001predicting} 
-and \verb;PureseqTM; \cite{wang2019efficient}. 
+was predicted using \verb;TMHMM; \cite{krogh2001predicting} . 
 From this topology, we determined if the epitope
 overlapped with a TMH.
-As a second source of naturally occurring epitopes, 
-we used the IEDB \cite{vita2019immune} for epitope sequences presented in healthy humans
-by T cells, or as found in MHC ligand studies.
 
@@
 To obtain experimental confirmation that peptides stemming from TMHs 
-are presented in MHC-I and MHC-II,
-two peptide elution studies were reanalyzed:
-For MHC-I, peptides presented \emph{in vivo} 
-by the (human) haplotypes HLA-A and B 
-were sequenced \cite{schellens2015comprehensive},
-for MHC-II these were alleles DQ2.5, DQ2.2, and 
-DQ7.5 \cite{bergseng2015different}.
-Figure \ref{fig:elution} shows the percentages of epitopes derived 
-from TMHs
-found in the MHC-I and MHC-II elution 
-studies,
-for the two topology prediction tools \verb;TMHMM; \cite{krogh2001predicting} 
-and \verb;PureseqTM; \cite{wang2019efficient}. 
-Regardless of the prediction tool, 
-at least 100 epitopes were predicted to be derived from a TMH for each condition. 
+are presented \emph{in vivo} in MHC-I and MHC-II,
+we used epitope data from the IEDB.
+We queried the IEDB for all linear epitopes obtained
+from MHC ligand assays in healthy humans, 
+for the MHC alleles as used in this study.
+From these epitopes, we kept those that are found
+exactly once in the human reference proteome.
+We predicted the topology of the protein each epitope
+was found in, using \verb;TMHMM; \cite{krogh2001predicting},
+from which we conclude if the epitope is overlapping with a TMH
+for at least 1 amino acid.
+
 From these findings, it is robustly predicted that

### 2021-11-13 bbbq_article.tex
Explain IEDB Method and Results better

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Peptide elution studies}\label{subsec:elution_studies}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-To obtain experimental evidence that epitopes derived from TMHs are presented in MHC,
+\subsection{TMH epitopes obtained from experimental data}\label{subsec:elution_studies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To obtain experimental confirmation that peptides stemming from TMHs 
+are presented \emph{in vivo} in MHC-I and MHC-II,
 we used the Immune Epitope Database (IEDB) \cite{vita2019immune}.
-This database contains the sequences of thousands of experimentally
-established epitope sequences.
-For each of these epitopes,
-its possible location(s) in 
-a human reference proteome,
-with UniProt ID UP000005640\_9606, was mapped.
-For the epitopes that were present in the proteome exactly once,
-the topology of the proteins in which these epitopes were located 
-was predicted using \verb;TMHMM; \cite{krogh2001predicting} . 
-From this topology, we determined if the epitope
-overlapped with a TMH.
+We queried the IEDB for all linear epitopes obtained
+from MHC ligand assays in healthy humans, 
+for the MHC alleles as used in this study.
+From these epitopes, we kept those that are found
+exactly once in the human reference proteome
+with UniProt ID UP000005640\_9606.
+We predicted the topology of the protein each epitope
+was found in, using \verb;TMHMM; \cite{krogh2001predicting},
+from which we conclude if the epitope is overlapping with a TMH
+for at least 1 amino acid.
 
@@
 
-To obtain experimental confirmation that peptides stemming from TMHs 
-are presented \emph{in vivo} in MHC-I and MHC-II,
-we used epitope data from the IEDB.
-We queried the IEDB for all linear epitopes obtained
-from MHC ligand assays in healthy humans, 
-for the MHC alleles as used in this study.
-From these epitopes, we kept those that are found
-exactly once in the human reference proteome.
-We predicted the topology of the protein each epitope
-was found in, using \verb;TMHMM; \cite{krogh2001predicting},
-from which we conclude if the epitope is overlapping with a TMH
-for at least 1 amino acid.
+From the IEDB,
+we obtained \richel{53,041}
+and \richel{2,495} linear epitope sequences for our MHC-I 
+and MHC-II alleles respectively.
+There are relatively few epitopes for MHC-II, 
+as MHC-II has many more different alleles than MHC-I \richel{describe how an
+MHC-II allele is a combation of two alleles}, 
+where we select only the 21 MHC-II alleles used in this study.
+
+Figure \ref{fig:2a} shows that there is a correlation between the
+percentage of TMH-derived epitopes that is presented \emph{in vivo}
+versus the percentage of TMH-derived epitopes that is predicted to be presented
+in MHC-I alleles. For MHC-II alleles, there were too few epitopes per MHC allele
+to result in a useful plot.
+
+We grouped all epitopes presented by MHC-I and MHC-II alleles and show 
+the percentage of which these were TMH-derived in figure \ref{fig:2c}.
+\richel{value}\% and \richel{value}\% were
+predicted to be TMH-derived epitopes.
+\richel{Indeed, what is the null hypothesis? The original one is that
+zero percent of all TMH derived epitopes is presented, rendering statistics
+useless. However, what about testing if TMH-derived epitopes are presented
+just as often as TMHs are part of the proteom?}.
 

### 2021-11-13 bbbq_article.tex
Add suggested text bu @fransbianchi, progress #249

@@
 
+\frans{Start}
+
+3.4 Experimental validation of presentation and T-cell
+activation of TMH-derived peptides
+
+To obtain experimental confirmation t hat peptides stemming from TMHs are
+presented in MHC-I and MHC-II, we mined the IEDB database for confirmed human
+MHC-ligands from the same MHC-I and MHC-II alleles as in fig1. Figure 2A? shows the
+percentages of epitopes derived from TMHs found for each of the most common MHC-
+I alleles incomparison to the predictions, although ...... A single allele comparison for
+MHC-II is not yet feasible as IEDB has limited reports on human MHC-II ligands for the
+representative alleles.
+The overall percentage of TMH derived epitopes found for
+representative MHC-I and MHC-II alleles was ... and ... , respectively see Fig 2b. The
+percentage of TMH derived epitopes is lower than expected from both the predictions
+(Fig1) and statistical chance, however these results are affected by (I) the probed proteins/
+peptides (II) the bias in mass spectrometry towards hydrophillic peptides (refs) and (III)
+availablilty of data for the different alleles.
+In fact these findings robustly confirm that
+epitopes derived from human TMHs are presented in both MHC-I and MHC-II. See t he
+supplementary Table S? for the exact values. We next mined the IEDB database for in
+which T-cell responses were confirmed against specific human epitopes from the specified
+alleles, from the total reports ...\% and ...\% concerned TMH derived epitopes in MHC class I
+and II, respectively Fig S?. The data in Fig S? thus confirms that not only TMH derived
+epitopes are presented on MHC, but also result in Immune responses.
+\frans{End}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Presentation of TMH-derived epitopes \emph{in vivo}}
+\subsection{Presentation of TMH-derived epitopes \emph{in vivo} in T cells}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2021-11-13 bbbq_article.tex
Merged @fransbianchi's Methods part for IEDB into Methods, moved rest to IEDB Results section

@@
 are presented \emph{in vivo} in MHC-I and MHC-II,
-we used the Immune Epitope Database (IEDB) \cite{vita2019immune}.
+we mined the Immune Epitope Database (IEDB) \cite{vita2019immune}
+for confirmed human MHC-ligands.
 We queried the IEDB for all linear epitopes obtained
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Evolutionary conservation of TMHs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{Introduction}
+
+To determine the evolutionary conservation of TMHs,
+human single nucleotide polymorphisms (SNPs) were first collected
+that resulted in a single amino acid substitution,
+and we then determined if this substitution occurred within a predicted TMH or not.
+
+% \paragraph{Data}
+
+As a data source, multiple
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: 
+the \emph{dbSNP} \cite{sherry2001dbsnp} database,
+which contains 650 million 
+cataloged non-redundant humane variations (called RefSNPs,
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names, \cite{brown2015gene})
+and \emph{protein} (for proteins sequences, \cite{sayers2010database}).
+
+% \paragraph{Pipeline}
+
+The first query was a call to the \emph{gene} database for the 
+term 'membrane protein' (in all fields) 
+for the organism \emph{Homo sapiens}.
+This resulted in 1,077 gene IDs (on December 2020).
+% 1077 gene IDs is correct for December 2020.
+% At 2021-03-01, one will get 1130 gene IDs.
+% Also, one of the gene IDs that was valid back then,
+% has been obsoleted.
+The next query was a call to the \emph{gene} database 
+to obtain the gene names from the gene IDs.
+Per gene name, the \emph{dbSNP} NCBI database was queried for 
+variations associated with the gene name. 
+As the NCBI API constrains its users to three calls per second
+(to assure fair use), we had to limit the extent of our analysis.
+
+The number of SNPs was limited to the first 250 variations per gene,
+resulting in $\approx$61k variations.
+Only variations that result in a SNP for
+a single amino acid substitution were analyzed, resulting in $\approx$38k SNPs.
+The exact amounts can be found in the supplementary materials,
+Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
+
+% \paragraph{Selection of SNPs}
+%
+SNPs were picked based on ID number, which is linked to their discovery date.
+To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined.
+This analysis showed no positional bias of the SNPs, as shown in supplementary figure \ref{fig:snp_rel_pos}.
+
+Per SNP, the \emph{protein} NCBI database was queried for the
+protein sequence.
+For each protein sequence, the protein topology was determined using \verb;PureseqTM;.
+Using these predicted protein topologies, the SNPs were scored to be located within or outside TMHs.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH-derived peptides are predicted to be over-presented in MHC-I}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
+for a human, viral and bacterial proteome.
+Per MHC-I allele, it shows the percentage of binders that overlap with a TMH 
+with at least one residue.
+The horizontal line shows the expected percentage of TMH-derived epitopes 
+that would be presented, if TMH-derived epitopes would be presented just as 
+likely as epitopes derived from soluble regions.
+For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
+than the null expectation, for a human and bacterial proteome.
+For the viral proteome, 12 out of 13 alleles present
+TMH-derived epitopes more often than expected by chance.
+The extent of the over-presentation between the different haplotypes 
+is similar for the probed proteomes, 
+which strengthens our previous conclusion \cite{bianchi2017} 
+that the hydrophobicity of the MHC-binding groove 
+is the main factor responsible for the predicted over-presentation 
+of TMH-derived peptides.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH-derived peptides are predicted to be over-presented in MHC-II}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We next wondered if the over-representation of TMH-derived peptides 
+would also be present for MHC-II. 
+Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
+predicted to be overlapping with TMHs for our human, viral and bacterial proteomes.
+We found that TMH-derived peptides are over-presented in all
+of the 21 MHC-II alleles, 
+for a human, bacterial and viral proteome,
+except for \verb;HLA-DRB3*0101; in \emph{M. tuberculosis}.
+See supplementary Table \ref{tab:tmh_binders_mhc2} 
+for the exact TMH and epitope counts.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The over-presentation of TMH-derived peptides is caused by the hydrophobicity of the MHC peptide binding groove}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For MHC-I, we previously showed that the over-presentation of TMH-derived 
+peptides is caused by the hydrophobicity of the peptide binding 
+grooves \cite{bianchi2017}. 
+Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
+show the extent of over-presentation
+of TMH-derived epitopes as a function of the hydrophobicity preference score 
+for the different alleles.
+An assumed linear correlation explains 88\% of the variability in MHC-I.
+For MHC-II, 62\% of the variability is explained by hydrophobicity. 
+This indicates that TMH-derived peptides are over-presented, 
+because the peptide binding grooves of most MHC-I and -II haplotypes 
+are relatively hydrophobic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experimental validation of presentation of TMH-derived peptides}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+From the IEDB,
+we obtained \richel{53,041}
+and \richel{2,495} linear epitope sequences for our MHC-I 
+and MHC-II alleles respectively.
+There are relatively few epitopes for MHC-II, 
+as MHC-II has many more different alleles than MHC-I \richel{describe how an
+MHC-II allele is a combation of two alleles}, 
+where we select only the 21 MHC-II alleles used in this study.
+
+Figure \ref{fig:2a} shows that there is a correlation between the
+percentage of TMH-derived epitopes that is presented \emph{in vivo}
+versus the percentage of TMH-derived epitopes that is predicted to be presented
+in MHC-I alleles. For MHC-II alleles, there were too few epitopes per MHC allele
+to result in a useful plot.
+
+We grouped all epitopes presented by MHC-I and MHC-II alleles and show 
+the percentage of which these were TMH-derived in figure \ref{fig:2c}.
+\richel{value}\% and \richel{value}\% were
+predicted to be TMH-derived epitopes.
+\richel{Indeed, what is the null hypothesis? The original one is that
+zero percent of all TMH derived epitopes is presented, rendering statistics
+useless. However, what about testing if TMH-derived epitopes are presented
+just as often as TMHs are part of the proteom?}.
+
+From these findings, it is robustly predicted that
+epitopes derived from TMHs are presented in both MHC-I and MHC-II.
+See the supplementary Table \ref{tab:elution} for the exact values.
+
 \frans{Start}
-
-3.4 Experimental validation of presentation and T-cell
-activation of TMH-derived peptides
-
-To obtain experimental confirmation t hat peptides stemming from TMHs are
-presented in MHC-I and MHC-II, we mined the IEDB database for confirmed human
-MHC-ligands from the same MHC-I and MHC-II alleles as in fig1. Figure 2A? shows the
+Figure 2A? shows the
 percentages of epitopes derived from TMHs found for each of the most common MHC-
@@
 \frans{End}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsubsection{Evolutionary conservation of TMHs}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \paragraph{Introduction}
-
-To determine the evolutionary conservation of TMHs,
-human single nucleotide polymorphisms (SNPs) were first collected
-that resulted in a single amino acid substitution,
-and we then determined if this substitution occurred within a predicted TMH or not.
-
-% \paragraph{Data}
-
-As a data source, multiple
-NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: 
-the \emph{dbSNP} \cite{sherry2001dbsnp} database,
-which contains 650 million 
-cataloged non-redundant humane variations (called RefSNPs,
-\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names, \cite{brown2015gene})
-and \emph{protein} (for proteins sequences, \cite{sayers2010database}).
-
-% \paragraph{Pipeline}
-
-The first query was a call to the \emph{gene} database for the 
-term 'membrane protein' (in all fields) 
-for the organism \emph{Homo sapiens}.
-This resulted in 1,077 gene IDs (on December 2020).
-% 1077 gene IDs is correct for December 2020.
-% At 2021-03-01, one will get 1130 gene IDs.
-% Also, one of the gene IDs that was valid back then,
-% has been obsoleted.
-The next query was a call to the \emph{gene} database 
-to obtain the gene names from the gene IDs.
-Per gene name, the \emph{dbSNP} NCBI database was queried for 
-variations associated with the gene name. 
-As the NCBI API constrains its users to three calls per second
-(to assure fair use), we had to limit the extent of our analysis.
-
-The number of SNPs was limited to the first 250 variations per gene,
-resulting in $\approx$61k variations.
-Only variations that result in a SNP for
-a single amino acid substitution were analyzed, resulting in $\approx$38k SNPs.
-The exact amounts can be found in the supplementary materials,
-Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
-
-% \paragraph{Selection of SNPs}
-%
-SNPs were picked based on ID number, which is linked to their discovery date.
-To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined.
-This analysis showed no positional bias of the SNPs, as shown in supplementary figure \ref{fig:snp_rel_pos}.
-
-Per SNP, the \emph{protein} NCBI database was queried for the
-protein sequence.
-For each protein sequence, the protein topology was determined using \verb;PureseqTM;.
-Using these predicted protein topologies, the SNPs were scored to be located within or outside TMHs.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TMH-derived peptides are predicted to be over-presented in MHC-I}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
-for a human, viral and bacterial proteome.
-Per MHC-I allele, it shows the percentage of binders that overlap with a TMH 
-with at least one residue.
-The horizontal line shows the expected percentage of TMH-derived epitopes 
-that would be presented, if TMH-derived epitopes would be presented just as 
-likely as epitopes derived from soluble regions.
-For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
-than the null expectation, for a human and bacterial proteome.
-For the viral proteome, 12 out of 13 alleles present
-TMH-derived epitopes more often than expected by chance.
-The extent of the over-presentation between the different haplotypes 
-is similar for the probed proteomes, 
-which strengthens our previous conclusion \cite{bianchi2017} 
-that the hydrophobicity of the MHC-binding groove 
-is the main factor responsible for the predicted over-presentation 
-of TMH-derived peptides.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{TMH-derived peptides are predicted to be over-presented in MHC-II}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-We next wondered if the over-representation of TMH-derived peptides 
-would also be present for MHC-II. 
-Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
-predicted to be overlapping with TMHs for our human, viral and bacterial proteomes.
-We found that TMH-derived peptides are over-presented in all
-of the 21 MHC-II alleles, 
-for a human, bacterial and viral proteome,
-except for \verb;HLA-DRB3*0101; in \emph{M. tuberculosis}.
-See supplementary Table \ref{tab:tmh_binders_mhc2} 
-for the exact TMH and epitope counts.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{The over-presentation of TMH-derived peptides is caused by the hydrophobicity of the MHC peptide binding groove}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-For MHC-I, we previously showed that the over-presentation of TMH-derived 
-peptides is caused by the hydrophobicity of the peptide binding 
-grooves \cite{bianchi2017}. 
-Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
-show the extent of over-presentation
-of TMH-derived epitopes as a function of the hydrophobicity preference score 
-for the different alleles.
-An assumed linear correlation explains 88\% of the variability in MHC-I.
-For MHC-II, 62\% of the variability is explained by hydrophobicity. 
-This indicates that TMH-derived peptides are over-presented, 
-because the peptide binding grooves of most MHC-I and -II haplotypes 
-are relatively hydrophobic.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experimental validation of presentation of TMH-derived peptides}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-From the IEDB,
-we obtained \richel{53,041}
-and \richel{2,495} linear epitope sequences for our MHC-I 
-and MHC-II alleles respectively.
-There are relatively few epitopes for MHC-II, 
-as MHC-II has many more different alleles than MHC-I \richel{describe how an
-MHC-II allele is a combation of two alleles}, 
-where we select only the 21 MHC-II alleles used in this study.
-
-Figure \ref{fig:2a} shows that there is a correlation between the
-percentage of TMH-derived epitopes that is presented \emph{in vivo}
-versus the percentage of TMH-derived epitopes that is predicted to be presented
-in MHC-I alleles. For MHC-II alleles, there were too few epitopes per MHC allele
-to result in a useful plot.
-
-We grouped all epitopes presented by MHC-I and MHC-II alleles and show 
-the percentage of which these were TMH-derived in figure \ref{fig:2c}.
-\richel{value}\% and \richel{value}\% were
-predicted to be TMH-derived epitopes.
-\richel{Indeed, what is the null hypothesis? The original one is that
-zero percent of all TMH derived epitopes is presented, rendering statistics
-useless. However, what about testing if TMH-derived epitopes are presented
-just as often as TMHs are part of the proteom?}.
-
-From these findings, it is robustly predicted that
-epitopes derived from TMHs are presented in both MHC-I and MHC-II.
-See the supplementary Table \ref{tab:elution} for the exact values.
 

### 2021-11-13 bbbq_article.tex
Merged @fransbianchi's work into Results section of IEDB study

--- 2021-11-13 bbbq_article.tex
@@
 
+\frans{This section requires review}
+\richel{This section requires review indeed}
+
 From the IEDB,
@@
 in MHC-I alleles. For MHC-II alleles, there were too few epitopes per MHC allele
-to result in a useful plot.
+to result in an informative figure.
 
@@
 
-From these findings, it is robustly predicted that
-epitopes derived from TMHs are presented in both MHC-I and MHC-II.
+The percentage of TMH derived epitopes is lower than expected, 
+for both the predictions \richel{Figure 1} 
+and statistical chance \richel{indeed, need stats}, 
+however these results are affected by 
+(I) the probed proteins/peptides 
+(II) the bias in mass spectrometry towards hydrophillic peptides (refs) and (III) availablilty of data for the different alleles.
+
+In fact, these findings robustly confirm that
+epitopes derived from human TMHs are presented in both MHC-I and MHC-II.
 See the supplementary Table \ref{tab:elution} for the exact values.
+
+We also mined the IEDB database for in
+which T-cell responses were confirmed against specific human epitopes from the specified
+alleles, from the total reports \richel{value}\% and \richel{value}\% concerned TMH-derived epitopes in MHC class I
+and II, respectively 
+(see Figure \ref{fig:t_cells_present_tmh_derived_epitopes}). 
+This data confirms that not only TMH derived epitopes are presented on MHC, but also result in immune responses.
 
@@
   }
-  \label{fig:elution_versus}
+  \label{fig:t_cells_present_tmh_derived_epitopes}
 \end{figure}

### 2021-11-13 bbbq_article.tex
Fix ref
Found which numbers are correct

@@
     obtained using an MHC ligand assay
+    \richel{These numbers (and probably, due to that, the
+    percentages) are incorrect}
   }
@@
 
-Figure \ref{fig:elution_versus} shows the percentage of TMH-derived
+Figure \ref{fig:t_cells_present_tmh_derived_epitopes} 
+shows the percentage of TMH-derived
 epitopes as presented \emph{in vivo} in T cells. 
@@
     found in the IEDB for the MHC alleles used in this study
+    \richel{Numbers must be approx 259 and 651}
   }

### 2021-11-14 bbbq_article.tex
Found error in calculation

@@
 
-Humans mostly express two haplotypes per MHC form, one from the parental and one from the maternal chromosome, 
+Humans mostly express two alleles per MHC haplotype, one from the parental and one from the maternal chromosome, 
 and therefore an individual's immune system detects 
@@
 Many studies are aimed at identifying the repertoire of epitopes that are presented 
-in any MHC haplotype allele and determining which epitopes will result in an immune response, 
+in any of the different alleles to determine which epitopes will result in an immune response, 
 as this will for instance aid the design of vaccines. 
@@
 For example, S. Tang et al. \cite{tang2011genome} found that, 
-of the 432 peptides that were predicted to bind to MHC,
+of the 432 peptides that were predicted to bind to an MHC allele,
 86\% were experimentally confirmed to do so.
@@
 from transmembrane helices (TMHs) 
-will be more frequently presented by MHC-I 
+are likely to be more frequently presented by MHC-I 
 than expected based on their abundance \cite{bianchi2017},
@@
 This over-presentation is attributed to the fact 
-that the peptide-binding groove of most MHC-I haplotypes 
+that the peptide-binding groove of most MHC-I alleles
 is relatively hydrophobic, 
@@
 TMH-derived epitopes more often than expected by chance.
-The extent of the over-presentation between the different haplotypes 
+The extent of the over-presentation between the different alleles
 is similar for the probed proteomes, 
@@
 
-The percentage of TMH derived epitopes is lower than expected, 
+The percentage of TMH derived epitopes is relatively high
 for both the predictions \richel{Figure 1} 
 and statistical chance \richel{indeed, need stats}, 
-however these results are affected by 
-(I) the probed proteins/peptides 
+however these results are hard to compare with the predictions as they are biased by
+(I) the expression of the probed proteins and cleavage sites, 
 (II) the bias in mass spectrometry towards hydrophillic peptides (refs) and (III) availablilty of data for the different alleles.
@@
 Therefor it's tempting to assume that this would also result in enhanced T-cell responses, assuming that the conservation of TMH's across organisms does not significantly effect the deletion of such T-cells by central tolerance mechanisms.
-Moreover, reanalysis of peptide elution studies confirmed the presentation of TMH-derived epitopes.
+Moreover, reanalysis of MHC-Ligands from the IEDB database and T-cell studies confirmed the presentation and immunicity of TMH-derived epitopes.
 Finally, our SNP analysis shows that TMHs are evolutionary more conserved 
@@
     obtained using an MHC ligand assay
-    \richel{These numbers (and probably, due to that, the
-    percentages) are incorrect}
+    \richel{Numbers must be 8031 and 698, percentage will then match 2A}
   }

### 2021-11-14 bbbq_article_issue_157/results.tex
Update table

@@
 % latex table generated in R 4.1.1 by xtable 1.8-4 package
-% Fri Oct 29 12:40:25 2021
+% Sun Nov 14 19:33:23 2021
 \begin{table}[ht]
 \centering
-\begin{tabular}{llll}
+\begin{tabular}{lll}
   \hline
-MHC class & Tool & Dataset & n \\ 
+MHC class & Dataset & n \\ 
   \hline
-I & PureseqTM & schellens & 1.38\% (109/7897) \\ 
-  I & PureseqTM & iedb & 6.81\% (43/631) \\ 
-  I & TMHMM & schellens & 1.43\% (113/7897) \\ 
-  I & TMHMM & iedb & 7.13\% (45/631) \\ 
-  II & PureseqTM & bergseng & 3.92\% (498/12712) \\ 
-  II & PureseqTM & iedb & 0.29\% (4/1364) \\ 
-  II & TMHMM & bergseng & 3.96\% (504/12712) \\ 
-  II & TMHMM & iedb & 1.39\% (19/1364) \\ 
+I & iedb\_mhc\_ligand & 22.28\% (1789/8030) \\ 
+  I & iedb\_t\_cell & 35.91\% (93/259) \\ 
+  II & iedb\_mhc\_ligand & 10.46\% (73/698) \\ 
+  II & iedb\_t\_cell &  6.66\% (42/631) \\ 
    \hline
 \end{tabular}
-\caption{Percentage of epitopes derived from a TMH found in the two elution studies, for the two different kind of topology prediction tools.
+\caption{Percentage of epitopes derived from a TMH for epitopes taken from the IEDB, for two different types of assays:  an MHC ligand assay, as well as a T cell assay.
 The values between braces show the the number of epitopes that were predicted to overlapping with a TMH per all epitopes that could be uniquely mapped to the representative human reference proteome.}

### 2021-11-14 bbbq_article.tex
Updates from Overleaf

@@
 
-\frans{This section requires review}
-\richel{This section requires review indeed}
-
-From the IEDB,
-we obtained \richel{53,041}
-and \richel{2,495} linear epitope sequences for our MHC-I 
+The IEDB contains millions of linear epitope sequences obtained
+by an MHC ligand assay.
+For the MHC alleles used in this study, 
+we obtained 54,303 and 2,484 linear epitope sequences for the MHC-I
 and MHC-II alleles respectively.
 There are relatively few epitopes for MHC-II, 
-as MHC-II has many more different alleles than MHC-I \richel{describe how an
-MHC-II allele is a combation of two alleles}, 
+as MHC-II has many more different alleles than MHC-I, 
 where we select only the 21 MHC-II alleles used in this study.
 
-Figure \ref{fig:2a} shows that there is a correlation between the
+Figure \ref{fig:2a} shows that there are similar levels of
+over-presentation of TMH-derived epitopes between (1) the
 percentage of TMH-derived epitopes that is presented \emph{in vivo}
-versus the percentage of TMH-derived epitopes that is predicted to be presented
+versus (2) the percentage of TMH-derived epitopes that is predicted to be presented
 in MHC-I alleles. For MHC-II alleles, there were too few epitopes per MHC allele
@@
 We grouped all epitopes presented by MHC-I and MHC-II alleles and show 
-the percentage of which these were TMH-derived in figure \ref{fig:2c}.
+the percentage of which these were TMH-derived in figure \ref{fig:2b}.
 \richel{value}\% and \richel{value}\% were
@@
     \caption{}
-    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2c.png}
-    \label{fig:2c}
+    \includegraphics[width=1.0\textwidth]{bbbq_article_issue_157/figure_2b.png}
+    \label{fig:2b}
   \end{subfigure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Correlation of epitope presentation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In the main text of this research, we use two sources of epitopes
+to determine if TMH-derived epitopes are presented.
+The first source of epitopes are all the 9-mers (for MHC-I) (and 15-mers for MHC-II)
+derived from a human reference proteome, where this over-presentation
+is displayed in figure \ref{fig:bbbq_1_smart_results}.
+The second source of epitopes are those that are present in the IEDB
+that are obtained from MHC ligand assays,
+as displayed in figure \ref{fig:2a}.
+
+Here we correlate between the over-presentation of TMH-derived
+epitopes between these two sources of data.
+Figure \ref{fig:thm_presentation_correlation} shows per haplotype
+the percentage of TMH-derived epitopes, with a linear trendline.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2c.png}
+  \caption{
+    \textbf{
+      TMH-derived epitopes are over-presented when using
+      predicted as well as \emph{in vivo} data
+    }
+    For the MHC class I alleles, the over-presentation of
+    TMH-derived epitopes is correlated between 
+    IEDB MHC ligand epitopes (horizontal axis) and the 9-mers
+    derived from a human reference proteome (vertical axis).
+    The trendline shows the linear correlation between these
+    percentages, where the gray area is the 95\% confidence interval.
+  }
+  \label{fig:thm_presentation_correlation}
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Presentation of TMH-derived epitopes \emph{in vivo} in T cells}
@@
     The numbers above the bars denotes the number of epitopes
-    found in the IEDB for the MHC alleles used in this study
-    \richel{Numbers must be approx 259 and 651}
+    found in the IEDB for the MHC alleles used in this study.
   }

### 2021-11-14 bbbq_article.tex
Processed all Richel's notes-to-self

@@
 We grouped all epitopes presented by MHC-I and MHC-II alleles and show 
-the percentage of which these were TMH-derived in figure \ref{fig:2b}.
-\richel{value}\% and \richel{value}\% were
-predicted to be TMH-derived epitopes.
-\richel{Indeed, what is the null hypothesis? The original one is that
-zero percent of all TMH derived epitopes is presented, rendering statistics
-useless. However, what about testing if TMH-derived epitopes are presented
-just as often as TMHs are part of the proteom?}.
+the percentage of which these were TMH-derived in figure \ref{fig:2b},
+which are 22\% and 10\%.
 
 The percentage of TMH derived epitopes is relatively high
-for both the predictions \richel{Figure 1} 
-and statistical chance \richel{indeed, need stats}, 
-however these results are hard to compare with the predictions as they are biased by
+for both sources of epitopes, which are (1) the predicted 
+9-mers (for MHC-I) and 14-mers (for MHC-II) from
+a reference human proteome  (see Figure \ref{fig:bbbq_1_smart_results}),
+(2) the IEBD epitopes from MHC ligand assays (Figure \ref{fig:2a}).
+These results cannot be compared with an \emp{a priori}
+expactation, as the predicted percentages are biased by
 (I) the expression of the probed proteins and cleavage sites, 
-(II) the bias in mass spectrometry towards hydrophillic peptides (refs) and (III) availablilty of data for the different alleles.
-
-In fact, these findings robustly confirm that
+(II) the bias in mass spectrometry towards hydrophillic peptides 
+and (III) availablilty of data for the different alleles.
+
+Regardless, these findings robustly confirm that
 epitopes derived from human TMHs are presented in both MHC-I and MHC-II.
@@
 which T-cell responses were confirmed against specific human epitopes from the specified
-alleles, from the total reports \richel{value}\% and \richel{value}\% concerned TMH-derived epitopes in MHC class I
-and II, respectively 
+alleles, from the total reports 36\% and 7\% 
+concerned TMH-derived epitopes in MHC class I and II, respectively 
 (see Figure \ref{fig:t_cells_present_tmh_derived_epitopes}).
+
 This data confirms that not only TMH derived epitopes are presented on MHC, but also result in immune responses.
@@
     obtained using an MHC ligand assay
-    \richel{Numbers must be 8031 and 698, percentage will then match 2A}
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Elution studies}
+\subsection{The percentage of TMH-derived epitopes from IEDB epitopes}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-See figure \ref{fig:elution_per_software} or
-supplementary Table \ref{tab:elution} 
-for the percentages of epitopes derived from a TMH 
-found in the two elution studies, 
-for the two different kind of topology prediction tools.
+We display the over-presentation of epitopes taken from the
+IEDB database, for two assays: an MHC ligand assay (Figure \ref{fig:2a})
+and a T cell assay (see figure \ref{fig:t_cells_present_tmh_derived_epitopes}),
+as a bar plot.
+Supplementary Table \ref{tab:elution} below shows the
+exact numbers.
 
@@
 to determine if TMH-derived epitopes are presented.
-The first source of epitopes are all the 9-mers (for MHC-I) (and 15-mers for MHC-II)
+The first source of epitopes are all the 9-mers (for MHC-I) (and 14-mers for MHC-II)
 derived from a human reference proteome, where this over-presentation

### 2021-11-18 bbbq_article.tex
Feedback @bogeert

@@
 based on their abundance in the proteome. However, the physiological reason of why membrane proteins might be over-presented is unclear. 
-In this study, we show that the over-presentation of TMH-derived peptides is general, as it is predicted for bacteria and viruses 
-and for both MHC-I and MHC-II. Moreover, we show that TMHs are evolutionarily more conserved, 
+In this study, we show that the predicted over-presentation of TMH-derived peptides is general, as it is predicted for bacteria and viruses 
+and for both MHC-I and MHC-II, and confirmed from re-analysis of epitope databases. Moreover, we show that TMHs are evolutionarily more conserved, 
 because single nucleotide polymorphisms (SNPs) are present 
@@
     \hline
+ER & Endoplasmatic reticulum \\
+ERAD & ER-associated degradation \\
+HLA & Human leukocyte antigen \\
+IEDB & Immune Epitope Database  \\
+LB & lipid body \\
 MAP & Membrane-associated protein \\
+MHC & Major histocompatibility complex \\
+MVB & Multivesicular body \\
+PLC & Peptide-loading complex \\
+SNP & Single nucleotide polymorphism \\
 TMH & Transmembrane helix         \\
@@
 than the innate immune response, 
-are T cells which recognize peptides, called epitopes, derived from antigenic proteins presented on Major Histocompatibility Complexes (MHC) class I and II on the cell surface.
+is T cells which recognize peptides, called epitopes, derived from antigenic proteins presented on Major Histocompatibility Complexes (MHC) class I and II on the cell surface.
 
@@
 TMHs can also be predicted with high accuracy from a protein sequence 
-by bioinformatics approaches \cite{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient},
-for example, \cite{jones2007improving} found that,
+by bioinformatics approaches \cite{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}.
+For example, a study by Jones \cite{jones2007improving} found that,
 from 184 transmembrane proteins (TMPs) with known topology, 
@@
 stemming from soluble (i.e., extracellular or cytoplasmic) protein regions is unknown. 
-We hypothesized that the presentation of 
+In this study, we hypothesized that the presentation of 
 TMH residues is evolutionary selected for, 
@@
 where the alleles of the MHC-I haplotypes are the alleles representing the 13 supertypes 
-with over 99.6\% coverage of the populations MHC-I repertoire as defined by \cite{lund2004definition} \cite{sette1999},
+with over 99.6\% coverage of the population's MHC-I repertoire as defined by \cite{lund2004definition} \cite{sette1999},
 and the 21 MHC-II haplotype alleles, have a phenotypic frequency 
@@
 that give rise to no or only very few MHC epitopes.
-To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials.
+To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials. \textbf{GEERT: where? give reference to figure}
 
@@
 from MHC ligand assays in healthy humans, 
-for the MHC alleles as used in this study.
-From these epitopes, we kept those that are found
+carrying the MHC alleles as used in this study.
+From these epitopes, we kept those that were present
 exactly once in the human reference proteome
@@
 was found in, using \verb;TMHMM; \cite{krogh2001predicting},
-from which we conclude if the epitope is overlapping with a TMH
-for at least 1 amino acid.
+from which we concluded if the epitope is overlapping with a TMH
+with at least 1 amino acid.
 
@@
 cataloged non-redundant humane variations (called RefSNPs,
-\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names, \cite{brown2015gene})
-and \emph{protein} (for proteins sequences, \cite{sayers2010database}).
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \cite{brown2015gene})
+and \emph{protein} (for proteins sequences \cite{sayers2010database}).
 
@@
 than the null expectation, for a human and bacterial proteome.
-For the viral proteome, 12 out of 13 alleles present
+For the viral proteome, 12 out of 13 MHC-I alleles present
 TMH-derived epitopes more often than expected by chance.
@@
 of TMH-derived epitopes as a function of the hydrophobicity preference score 
-for the different alleles.
+for the different alleles. \textbf{GEERT: is this done for human, SARS-CoV-2, or MtB?}
 An assumed linear correlation explains 88\% of the variability in MHC-I.
@@
 
-The IEDB contains millions of linear epitope sequences obtained
+The Immune Epitope Database (IEDB) from the National Institutes of Health contains millions of linear epitope sequences obtained
 by an MHC ligand assay.
@@
 There are relatively few epitopes for MHC-II, 
-as MHC-II has many more different alleles than MHC-I, 
-where we select only the 21 MHC-II alleles used in this study.
+as MHC-II has many more different alleles than MHC-I and we selected only the epitopes found for the 21 MHC-II alleles used in this study.
 
@@
 over-presentation of TMH-derived epitopes between (1) the
-percentage of TMH-derived epitopes that is presented \emph{in vivo}
+percentage of TMH-derived epitopes that is reported in the IEDB database
 versus (2) the percentage of TMH-derived epitopes that is predicted to be presented
@@
 the percentage of which these were TMH-derived in figure \ref{fig:2b},
-which are 22\% and 10\%.
-
-The percentage of TMH derived epitopes is relatively high
+which are 22\% and 10\%. \textbf{GEERT: zou je ook geen horizontale rode streep kunnen trekken op de verwachtingswaardes, net al in Fig 1A? Dan zie je dat het hoger is}
+
+\textbf{GEERT: zie voorgaande punt aangaande verwachtingwaarde.
+De gevonden waardes zijn hoger, dus zoek je in volgende stuk tekst naar problemen die er niet zijn, en zou dit dus weg laten.
+Is onduidelijk en wollig stuk: The percentage of TMH derived epitopes is relatively high
 for both sources of epitopes, which are (1) the predicted 
@@
 (II) the bias in mass spectrometry towards hydrophillic peptides 
-and (III) availablilty of data for the different alleles.
-
-Regardless, these findings robustly confirm that
-epitopes derived from human TMHs are presented in both MHC-I and MHC-II.
+and (III) availablilty of data for the different alleles.}
+
+These findings robustly confirm that
+epitopes derived from human TMHs are presented in both MHC-I and MHC-II, and support that they are over-presented.
 See the supplementary Table \ref{tab:elution} for the exact values.
 
-We also mined the IEDB database for in
-which T-cell responses were confirmed against specific human epitopes from the specified
+We also mined the IEDB database for epitopes for
+which T-cell responses were confirmed from the specified
 alleles, from the total reports 36\% and 7\% 
@@
 The proportion of SNPs found in TMHs varied from 
-none (i.e. all SNPs were in
-soluble regions) to all.
+none (i.e., all SNPs were in
+soluble regions) to all (i.e., all SNPs were in
+TMHs).
 To determine if SNPs were randomly distributed over the protein, we performed a linear regression analysis,
@@
 which means this number is significantly less as explained by variation.
+The TMHs of multi-spanners are thus significantly more conserved than soluble protein regions, whereas this is not the case for single-spanners.
 
@@
 derived from TMHs are a major but overlooked source of MHC epitopes. 
-Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance, 
-regardless of the organism.
-Furthermore we see despite the bias towards hydrophilic peptides due to their enhanced ionization in mass spec that in several peptide elution studies still a higher fraction of TMH derived epitopes is found than predicted by chance.
-Therefor it's tempting to assume that this would also result in enhanced T-cell responses, assuming that the conservation of TMH's across organisms does not significantly effect the deletion of such T-cells by central tolerance mechanisms.
-Moreover, reanalysis of MHC-Ligands from the IEDB database and T-cell studies confirmed the presentation and immunicity of TMH-derived epitopes.
+Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance for both MHC-I and MHC-II, regardless of the organism.
+Moreover, reanalysis of MHC-ligands from the IEDB database confirmed the presentation of TMH-derived epitopes.
+Therefor, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms.
 Finally, our SNP analysis shows that TMHs are evolutionary more conserved
@@
 by intramembrane proteases that cleave TMHs while they are still membrane embedded.
-Supporting this hypothesis is the well established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences,
+Supporting this hypothesis is the well-established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences,
 are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}. 
@@
 possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}. 
-However, this mechanism would not explain how multispanner polytopic membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at the N-terminus of a protein.
+However, this mechanism cannot explain how most membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at their C-termini, and N-terminal domains will hence not be removed.
 Nevertheless, the presentation of peptides with a high hydrophobicity index 
 was shown to be independent of TAP as well \cite{lautscham2001processing}, 
-suggesting the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
+suggesting that the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
 
@@
 Via this mechanism, TMPs might well be processed for antigen presentation, 
-particularly since the loading of MHC class II molecules is well understood 
+particularly since the loading of MHC-II molecules is well understood 
 to occur in MVBs \cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}. 
@@
 Alternatively to the enzymatic degradation of lipids in MVBs by lipases \cite{sander2016lipase,gilleron2016lysosomal}, 
-they might be oxidatively degraded by reactions with radical oxygen species (ROS) produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}. 
+they might be oxidatively degraded by reactions with radical oxygen species produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}. 
 This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} 
@@
 In general, one might expect that evolutionary selection results in
-an immune system that as most attentive for protein regions that are
+an immune system that is most attentive for protein regions that are
 essential for the survival, proliferation and/or virulence or pathogenic microbes, 
@@
 Evolutionary selection of pathogens by a host's immune system,
-however, is likelier to occur for proteomic patterns that are general,
+however, is likelier to occur for protein patterns that are general,
 over patterns that are rare.
@@
 using different methodologies.
-One such example is \cite{stevens2001substitution}, 
+One such example is a study by Stevens and Arkin \cite{stevens2001substitution}, 
 in which aligned protein sequence data was used.
@@
 which is a similar value as we found.
-Another example is a study that estimated the conservation
+Another example is a study by Oberai, et al. \cite{oberai2009structural} that estimated the conservation
 scores for TMHs and soluble regions based on 
@@
 with a conservation score that was 17\% higher in 
-TMHs \cite{oberai2009structural}.
+TMHs.
 Note that the last study also found that mutations in human TMHs are likelier to cause
@@
     overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
-    row) and \emph{M. tuberculosis} (bottom row).
+    row) and \emph{M. tuberculosis} (MtB; bottom row).
     The pair of dashed lines in each plot indicate the lower and upper bound of the 99\% confidence interval.
@@
     and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
-    and MHC-II haplotypes \textbf{(C)}. 
+    and MHC-II haplotypes \textbf{(C)}. \textbf{GEERT: are panels B-C done with human, SARS-CoV-2 or MtB?}
     Diagonal red line: linear regression analysis. 
@@
     The percentage of epitopes for MHC-I and -II alleles 
-    that overlap with TMHs that are presented
+    that overlap with TMHs that are presented.
     \textbf{(A)} 
@@
     \textbf{(B)} 
-    MHC ligands from the IEDB bind to TMH-derived epitopes.
+    MHC ligands from IEDB binding to TMH-derived epitopes.
     The numbers above the bars denotes the number of epitopes
-    obtained using an MHC ligand assay
+    obtained using an MHC ligand assay. \textbf{GEER: zie comment in main text: wellicht rode horizontale streep bij verwachtingswaarde, net als in Fig 1A? Maakt manuscript veel duidelijker}
   }
@@
     Each point shows for one protein the predicted percentage of
-    TMH ($x$-axis) and the observed occurrence of SNPs being located
+    amino acids that are part of a TMH ($x$-axis) and the observed occurrence of SNPs being located
     within a TMH ($y$-axis).
@@
     Percentages show the relative conservation
-    of SNPs in TMHs found.
+    of SNPs in TMHs found relative to stochastic chance.
   }
@@
       (left) or multiple (right) TMHs.
-      Each point shows for one protein the predicted percentage of TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
+      Each point shows for one protein the predicted percentage of amino acids that are part of a TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
 The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
@@
       Percentages show the relative conservation
-      of SNPs in TMHs found.
+      of SNPs in TMHs found relative to the stochastic chances.
       \textbf{(C)} 
@@
 See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}
-for an overview of all amounts.
+for an overview of all amounts. \textbf{GEERT in Table S1 en S2 schrijven SARS-CoV-2 ipv covid (want covid is de ziekte, SARS-CoV-2 het virus)}
 In supplementary Table \ref{tab:ncbi_counts_1} there are multiple instances where
@@
 Supplementary Table \ref{tab:elution} below shows the
-exact numbers.
+exact numbers. \textbf{GEERT: spelfout in legenda: brackets ipv braces}
 
@@
       TMH-derived epitopes are over-presented when using
-      predicted as well as \emph{in vivo} data
+      predicted as well as experimental data
     }
@@
 shows the percentage of TMH-derived
-epitopes as presented \emph{in vivo} in T cells. 
+epitopes of the reported epitopes that for which T-cell responses were established. 
 The data was obtained from the IEDB and includes only the MHC
-alleles as used in this study. As there are many (especially class II)
+alleles used in this study. As there are many (especially class II)
 MHC alleles, only a small percentage of the full IEDB data could be used.
-Although this figure demonstrates that TMH-derived epitopes are presented
-\emph{in vivo}, we excluded this figure from the main article, as these
-protocols select epitopes in a non-random fashion (unlike the
-MHC ligand assays, as presented in the main article).
 
@@
     \textbf{
-      T cells present TMH-derived epitopes \emph{in vivo}
+      TMH-derived epitopes evoke T-cell responses 
       }

### 2021-11-18 bbbq_article.tex
Fix LaTeX

@@
 (2) the IEBD epitopes from MHC ligand assays (Figure \ref{fig:2a}).
-These results cannot be compared with an \emp{a priori}
+These results cannot be compared with an \emph{a priori}
 expactation, as the predicted percentages are biased by
@@
 (II) the bias in mass spectrometry towards hydrophillic peptides 
-and (III) availablilty of data for the different alleles.}
+and (III) availablilty of data for the different alleles.
+}
 

### 2021-11-18 bbbq_article.tex
Add red horizontal lines for confidence interval in figure 2a

@@
     row) and \emph{M. tuberculosis} (MtB; bottom row).
-    The pair of dashed lines in each plot indicate the lower and upper bound of the 99\% confidence interval.
+    The pair of horizontal red lines in each plot indicate the lower and upper bound 
+    of the 99\% confidence interval.
     See supplementary Tables \ref{tab:tmh_binders_mhc2} and \ref{tab:tmh_binders_mhc1}
@@
     that overlap with TMHs that are presented.
+    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
     \textbf{(A)} 
@@
     obtained using an MHC ligand assay. \textbf{GEER: zie comment in main text: wellicht rode horizontale streep bij verwachtingswaarde, net als in Fig 1A? Maakt manuscript veel duidelijker}
+    The pair of horizontal red lines in each plot indicate the lower and upper bound 
+    of the 99\% confidence interval.
   }
@@
     derived from a human reference proteome (vertical axis).
+    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
     The trendline shows the linear correlation between these

### 2021-11-18 bbbq_article.tex
Confidence interval for plot2a only

@@
     Observed and predicted percentage of TMH-derived epitopes
+    The pair of horizontal red lines in each plot indicate the lower and upper bound 
+    of the 99\% confidence interval.
     \textbf{(B)} 
@@
     The numbers above the bars denotes the number of epitopes
-    obtained using an MHC ligand assay. \textbf{GEER: zie comment in main text: wellicht rode horizontale streep bij verwachtingswaarde, net als in Fig 1A? Maakt manuscript veel duidelijker}
-    The pair of horizontal red lines in each plot indicate the lower and upper bound 
-    of the 99\% confidence interval.
+    obtained using an MHC ligand assay. 
   }

### 2021-11-18 bbbq_article.tex

@@
 that give rise to no or only very few MHC epitopes.
-To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials. \textbf{GEERT: where? give reference to figure}
+To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials, Figures \ref{fig:bianch_et_al_2017_1a}
+and \ref{fig:bilderbeek_et_al_2021_1a}.
 
@@
     The pair of horizontal red lines in each plot indicate the lower and upper bound 
-    of the 99\% confidence interval.
+    of the 99\% confidence interval. Note that only one line is visible as this
+    interval is relatively narrow.
     \textbf{(B)} 

### 2021-11-18 bbbq_article.tex
Figure 2b has coincidence intervals

@@
 of TMH-derived epitopes as a function of the hydrophobicity preference score 
-for the different alleles. \textbf{GEERT: is this done for human, SARS-CoV-2, or MtB?}
+for the different human MHC alleles.
 An assumed linear correlation explains 88\% of the variability in MHC-I.

### 2021-11-18 bbbq_article.tex
Remove wooly part, thanks Geert

@@
 the percentage of which these were TMH-derived in figure \ref{fig:2b},
-which are 22\% and 10\%. \textbf{GEERT: zou je ook geen horizontale rode streep kunnen trekken op de verwachtingswaardes, net al in Fig 1A? Dan zie je dat het hoger is}
-
-\textbf{GEERT: zie voorgaande punt aangaande verwachtingwaarde.
-De gevonden waardes zijn hoger, dus zoek je in volgende stuk tekst naar problemen die er niet zijn, en zou dit dus weg laten.
-Is onduidelijk en wollig stuk: The percentage of TMH derived epitopes is relatively high
-for both sources of epitopes, which are (1) the predicted 
-9-mers (for MHC-I) and 14-mers (for MHC-II) from
-a reference human proteome  (see Figure \ref{fig:bbbq_1_smart_results}),
-(2) the IEBD epitopes from MHC ligand assays (Figure \ref{fig:2a}).
-These results cannot be compared with an \emph{a priori}
-expactation, as the predicted percentages are biased by
-(I) the expression of the probed proteins and cleavage sites, 
-(II) the bias in mass spectrometry towards hydrophillic peptides 
-and (III) availablilty of data for the different alleles.
-}
+which are 22\% and 10\%.
 
@@
     that overlap with TMHs that are presented.
+    The pair of horizontal red lines in each plot indicate the lower and upper bound 
+    of the 99\% confidence interval. Note that only one line is visible as this
+    interval is relatively narrow.
     Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
@@
     Observed and predicted percentage of TMH-derived epitopes
-    The pair of horizontal red lines in each plot indicate the lower and upper bound 
-    of the 99\% confidence interval. Note that only one line is visible as this
-    interval is relatively narrow.
     \textbf{(B)} 

### 2021-11-18 bbbq_article_issue_157/results.tex
Indeed, '(' is a round bracket, not a round brace, thanks Geert

@@
 \caption{Percentage of epitopes derived from a TMH for epitopes taken from the IEDB, for two different types of assays:  an MHC ligand assay, as well as a T cell assay.
-The values between braces show the the number of epitopes that were predicted to overlapping with a TMH per all epitopes that could be uniquely mapped to the representative human reference proteome.}
+The values between brackets show the the number of epitopes that were predicted to overlapping with a TMH per all epitopes that could be uniquely mapped to the representative human reference proteome.}
 \label{tab:elution}

### 2021-11-18 bbbq_article.tex

@@
     Correlation between the percentages of predicted TMH-derived epitopes
-    and the hydrophobicity score of all predicted epitopes for MHC-I \textbf{(B)}
-    and MHC-II haplotypes \textbf{(C)}. \textbf{GEERT: are panels B-C done with human, SARS-CoV-2 or MtB?}
+    and the hydrophobicity score of all predicted epitopes for 
+    human MHC-I \textbf{(B)} and MHC-II alleles \textbf{(C)}.
     Diagonal red line: linear regression analysis. 
@@
 See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}
-for an overview of all amounts. \textbf{GEERT in Table S1 en S2 schrijven SARS-CoV-2 ipv covid (want covid is de ziekte, SARS-CoV-2 het virus)}
+for an overview of all amounts. 
+Note that, for the analyses using the SARS-CoV-2 virus proteome,
+we labeled this by its disease (\verb;covid;) to prevent typos.
 In supplementary Table \ref{tab:ncbi_counts_1} there are multiple instances where

### 2021-11-18 bbbq_article.tex
Correct name for Maksim
Process feedback Maksim, progress #253

@@
 Thus, our findings suggest that 
-both cytolytic and helper T cells respond more to membrane proteins, 
+both cytolytic and helper T cells are more tuned to respond to membrane proteins, 
 because these are evolutionary more conserved. 
-We speculate that TMHs therefore are less prone to escape mutations 
+We speculate that TMHs are less prone to mutations 
 that enable pathogens to evade T cell responses.
@@
 epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation
+
+% Process all floats before going to a next page
+\clearpage
 
@@
 
+\maksim{
+  I would change the whole sentence: 
+  T cells are an important part of the acquired immunity, they specialize in recognition of pathogen-derived peptides, called epitopes, presented on the cells surface of antigen presenting cells (APC) within the Major Histocompatibility Complexes (MHC) class I and II.
+}
 The MHC proteins are heterodimeric complexes encoded by the
@@
 These studies have led to the development of prediction algorithms 
-that allow for very reliable \emph{in silico} predictions 
-of the binding affinities of peptides
+that allow for very reliable \emph{in silico} 
+\maksim{predictions of the peptide binding affinities}
+predictions of the binding affinities of peptides
 \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
@@
 Using these prediction algorithms, 
-we recently predicted that peptides derived 
+we recently showed that peptides derived 
 from transmembrane helices (TMHs) 
@@
 and therefore hydrophobic TMH-derived peptides have a higher affinity 
-to bind than their soluble hydrophilic counterparts.
+to bind than their soluble hydrophilic 
+\maxim{shouldn't be hydrophobic? 
+  A hydrophobic peptide binds to its hydrophobic counterpart - hyprophobic grove of MHC-I}
+counterparts.
 
@@
 from 184 transmembrane proteins (TMPs) with known topology, 
-80\% of the TMH predictions replicated this finding.
+80\% of the TMH predictions 
+\maxim{this part is confusing: 
+  correct maybe:
+  for example, TMH prediction algorithms identified 80\% of TMH in 184 transmembrane proteins (TMPs) with known topology [12].}
+replicated this finding.
 
@@
 In this study, we hypothesized that the presentation of 
-TMH residues is evolutionary selected for, 
-because TMHs are less prone to undergo escape mutations. 
+TMH residues is evolutionarily preferred, 
+since TMHs are less prone to undergo escape mutations. 
 One reason to expect such a reduced 
 variability (and hence evolutionary conservation) in TMHs, 
-is that these are restricted in their evolution 
+is that these are restricted in their variability 
 by the functional requirement to span a lipid bilayer. 
-Due to this requirement, 
-many of the amino acids genuinely present in TMHs 
-are limited to the ones with hydrophobic side chains 
+This limits 
+many of the amino acids present in TMHs 
+to have hydrophobic side chains 
 \cite{hessa2007molecular,jones1994model}.
@@
 might have a lower chance to develop escape mutations, 
-as many mutations will result in a dysfunctional TMH 
+as that will result in a dysfunctional TMH 
 and render the protein inactive.
@@
 This study had two objectives. 
-First, we aimed to generalize our findings by predicting the presentation of peptides 
-from different kingdoms of life and for both MHC-I and -II. 
+First, we aimed to generalize our findings by predicting
+\maxim{the antigenic presentation}
+the presentation of peptides 
+from different kingdoms of life in both MHC-I and -II. 
 From these \emph{in silico} predictions, we conclude that TMH-derived
-epitopes are likely to be presented more often than expected by chance,
-in a human, viral and bacterial proteome, 
-and for most alleles of both MHC-I and II. 
+epitopes from a human, viral and bacterial proteome
+are likely to be presented more often than expected by chance 
+for most alleles of MHC-I and II. 
 We confirmed the presentation of TMH-derived peptides 
@@
 These findings strengthen the emerging notion 
-that TMHs are important for the T cell branch of the adaptive immune system, 
+that TMHs are important for the T cell-mediated adaptive immune system, 
 and hence are of  
@@
 Supplementary Table \ref{tab:elution} below shows the
-exact numbers. \textbf{GEERT: spelfout in legenda: brackets ipv braces}
+exact numbers.
 

### 2021-11-18 bbbq_article.tex
Process feedback Maksim until (and excluding) Discussion

@@
 For each of these peptides, we determined if it overlapped with a predicted 
-TMH and if it was predicted to bind to the moest frequent alleles of each haplotype.
+TMH and if it was predicted to bind to the most frequent alleles of each haplotype.
 
@@
 For MHC-II, 14-mers were used, 
-as these are the most frequently occurring epitope length \cite{bergseng2015different}.
+as this is the most frequently occurring epitope length \cite{bergseng2015different}.
 A human (UniProt ID UP000005640\_9606), 
@@
 and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
-Both MCH-I and MHC-II alleles were selected to have a high occurrence,
+Both MCH-I and MHC-II alleles were selected to have a high occurrence
+\maxim{was it meant to be: 
+were selected based on their occurrence scores? or based in their prevalence in the population?},
 where the alleles of the MHC-I haplotypes are the alleles representing the 13 supertypes 
@@
 that give rise to no or only very few MHC epitopes.
-To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials, Figures \ref{fig:bianch_et_al_2017_1a}
+To verify that the results are similar \maxim{does it mean: that the new thresholding does not obscure the previous results, a side...}, a side by side comparison was performed shown in the supplementary materials, Figures \ref{fig:bianch_et_al_2017_1a}
 and \ref{fig:bilderbeek_et_al_2021_1a}.
@@
 from which we concluded if the epitope is overlapping with a TMH
-with at least 1 amino acid.
+with at least 1 amino acid \maxim{1 amino acid overlap is already a good enough evidence to label something as TMH?}.
 
@@
 human single nucleotide polymorphisms (SNPs) were first collected
+\maxim{..TMHs, we first collected (SNPs) resulting in a single amino acid substitution to determine if this occurred within a predicted TMH or not.}
 that resulted in a single amino acid substitution,
@@
 The exact amounts can be found in the supplementary materials,
-Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
+Tables \maxim{earlier you used brackets to refer to these tables} \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
 
@@
 that would be presented, if TMH-derived epitopes would be presented just as 
-likely as epitopes derived from soluble regions.
+likely as epitopes derived from soluble regions \maxim{assuming equal incidence of soluble and TMH-derived epitope presentation.}.
 For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
@@
 We next wondered if the over-representation of TMH-derived peptides 
-would also be present for MHC-II. 
+would also be confirmed for MHC-II. 
 Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
@@
 There are relatively few epitopes for MHC-II, 
-as MHC-II has many more different alleles than MHC-I and we selected only the epitopes found for the 21 MHC-II alleles used in this study.
+as MHC-II has many more different alleles than MHC-I,
+whereas we selected only the epitopes found for the 21 MHC-II alleles used in this study.
 
@@
 
-We grouped all epitopes presented by MHC-I and MHC-II alleles and show 
-the percentage of which these were TMH-derived in figure \ref{fig:2b},
+In figure \ref{fig:2b} we grouped all the epitopes 
+presented by MHC-I and MHC-II alleles 
+by the percentage of TMH-derived epitopes,
 which are 22\% and 10\%.
@@
 We also mined the IEDB database for epitopes for
-which T-cell responses were confirmed from the specified
+any type of T-cell response from the specified
 alleles, from the total reports 36\% and 7\% 
@@
 
-This data confirms that not only TMH derived epitopes are presented on MHC, but also result in immune responses.
+This data confirms that not only TMH derived epitopes are presented on MHC, 
+but this also elicits T-cell mediated immune responses.
 
@@
 Per protein, we calculated two percentages: 
-(1) the percentage of the total protein predicted to be TMHs, 
+(1) the percentage of the total protein predicted to be TMHs \maxim{is it percentage of a protein sequence length bearing TMHs? 
+if it is true - it would be a better wording}, 
 and (2) the percentage of SNPs located within these predicted TMHs.
@@
 one finds 918 SNPs in TMHs,
+\maxim{your current figure does not describe it}
 as depicted in figure \ref{fig:conservation}.

### 2021-11-18 bbbq_article.tex
All Maksim's feedback is in the document, progress #253

@@
 
-Epitope prediction is important to understand the immune system 
+Epitope prediction is important to understand the immune system function
 and for the design of vaccines.
@@
 Moreover, reanalysis of MHC-ligands from the IEDB database confirmed the presentation of TMH-derived epitopes.
-Therefor, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms.
+Therefore, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms.
 Finally, our SNP analysis shows that TMHs are evolutionary more conserved
@@
 into the aqueous vacuolar lumen, 
-and their prior or subsequent proteolytic processing are unresolved.
+leading to subsequent proteolytic processing are unresolved.
 
@@
 
-In general, one might expect that evolutionary selection results in
-an immune system that is most attentive for protein regions that are
+In general, one might expect that evolutionary selection shapes an immune system 
+where surveillance is directed towards protein regions 
 essential for the survival, proliferation and/or virulence or pathogenic microbes, 
@@
 In SARS-CoV-2, for example, there is preliminary evidence that the strongest
-selection pressure is upon residues that change its 
+selection pressure is directed upon residues that change its 
 virulence \cite{velazquez2020positive}.
@@
 Evolutionary selection of pathogens by a host's immune system,
-however, is likelier to occur for protein patterns that are general,
+however, is more likely to occur for protein patterns that are general,
 over patterns that are rare.

### 2021-11-18 bbbq_article.tex
Processed feedback Maksim. Thanks Maksim!

@@
 
-\maksim{
-  I would change the whole sentence: 
-  T cells are an important part of the acquired immunity, they specialize in recognition of pathogen-derived peptides, called epitopes, presented on the cells surface of antigen presenting cells (APC) within the Major Histocompatibility Complexes (MHC) class I and II.
-}
 The MHC proteins are heterodimeric complexes encoded by the
@@
 that allow for very reliable \emph{in silico} 
-\maksim{predictions of the peptide binding affinities}
-predictions of the binding affinities of peptides
+predictions of the peptide binding affinities
 \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
@@
 and therefore hydrophobic TMH-derived peptides have a higher affinity 
-to bind than their soluble hydrophilic 
-\maxim{shouldn't be hydrophobic? 
-  A hydrophobic peptide binds to its hydrophobic counterpart - hyprophobic grove of MHC-I}
+to bind than their soluble hydrophobic 
 counterparts.
@@
 For example, a study by Jones \cite{jones2007improving} found that,
-from 184 transmembrane proteins (TMPs) with known topology, 
-80\% of the TMH predictions 
-\maxim{this part is confusing: 
-  correct maybe:
-  for example, TMH prediction algorithms identified 80\% of TMH in 184 transmembrane proteins (TMPs) with known topology [12].}
-replicated this finding.
-
+from 184 transmembrane proteins (TMPs) with known \emph{in vitro} topology, 
+80\% of the TMH predictions of these proteins
+matched the experimental findings.
 TMHs are common structures in the proteins of humans and microbes. 
@@
 First, we aimed to generalize our findings by predicting
-\maxim{the antigenic presentation}
-the presentation of peptides 
+the antigenic presentation
 from different kingdoms of life in both MHC-I and -II. 
@@
 and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
-Both MCH-I and MHC-II alleles were selected to have a high occurrence
-\maxim{was it meant to be: 
-were selected based on their occurrence scores? or based in their prevalence in the population?},
+Both MCH-I and MHC-II alleles were selected 
+to have a high prevalence in the population,
 where the alleles of the MHC-I haplotypes are the alleles representing the 13 supertypes 
@@
 that give rise to no or only very few MHC epitopes.
-To verify that the results are similar \maxim{does it mean: that the new thresholding does not obscure the previous results, a side...}, a side by side comparison was performed shown in the supplementary materials, Figures \ref{fig:bianch_et_al_2017_1a}
-and \ref{fig:bilderbeek_et_al_2021_1a}.
+To verify that the slight change in method yields similar results,
+a side by side comparison is shown in the supplementary materials, 
+Figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
 
@@
 from which we concluded if the epitope is overlapping with a TMH
-with at least 1 amino acid \maxim{1 amino acid overlap is already a good enough evidence to label something as TMH?}.
+with at least 1 amino acid.
 
@@
 To determine the evolutionary conservation of TMHs,
-human single nucleotide polymorphisms (SNPs) were first collected
-\maxim{..TMHs, we first collected (SNPs) resulting in a single amino acid substitution to determine if this occurred within a predicted TMH or not.}
-that resulted in a single amino acid substitution,
-and we then determined if this substitution occurred within a predicted TMH or not.
+we first collected human single nucleotide polymorphisms (SNPs) 
+resulting in a single amino acid substitution to determine if this occurred within a predicted TMH or not.
 
@@
 The exact amounts can be found in the supplementary materials,
-Tables \maxim{earlier you used brackets to refer to these tables} \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
+Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
 
@@
 that would be presented, if TMH-derived epitopes would be presented just as 
-likely as epitopes derived from soluble regions \maxim{assuming equal incidence of soluble and TMH-derived epitope presentation.}.
+likely as epitopes derived from soluble regions,
+when assuming equal incidence of soluble and TMH-derived epitope presentation.
 For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
@@
 Per protein, we calculated two percentages: 
-(1) the percentage of the total protein predicted to be TMHs \maxim{is it percentage of a protein sequence length bearing TMHs? 
-if it is true - it would be a better wording}, 
+(1) the percentage of a protein sequence length bearing TMHs, 
 and (2) the percentage of SNPs located within these predicted TMHs.
@@
 one finds 918 SNPs in TMHs,
-\maxim{your current figure does not describe it}
-as depicted in figure \ref{fig:conservation}.
+as depicted (as percentages) in figure \ref{fig:conservation}.
 

### 2021-11-18 bbbq_article.tex

@@
 we obtained 54,303 and 2,484 linear epitope sequences for the MHC-I
-and MHC-II alleles respectively.
+and MHC-II alleles from human origin respectively.
 There are relatively few epitopes for MHC-II, 
 as MHC-II has many more different alleles than MHC-I,
-whereas we selected only the epitopes found for the 21 MHC-II alleles used in this study.
+whereas we selected only the human epitopes found for the 21 MHC-II alleles used in this study.
 
@@
 to result in an informative figure. 
+We showed that there is a correlation between epitopes taken from
+predicted 9-mers (for MHC-I) and 14-mers (for MHC-II) from a human
+reference proteome and epitopes taken from the IEDB for MHC ligands
+essays (see in Supplementary figure \ref{fig:thm_presentation_correlation}).
 

### 2021-11-19 bbbq_article.tex
Changes @fransbianchi, thanks Frans

@@
 In this study, we show that the predicted over-presentation of TMH-derived peptides is general, as it is predicted for bacteria and viruses 
-and for both MHC-I and MHC-II, and confirmed from re-analysis of epitope databases. Moreover, we show that TMHs are evolutionarily more conserved, 
+and for both MHC-I and MHC-II, and confirmed by re-analysis of epitope databases. Moreover, we show that TMHs are evolutionarily more conserved, 
 because single nucleotide polymorphisms (SNPs) are present 
@@
 than the innate immune response, 
-is T cells which recognize peptides, called epitopes, derived from antigenic proteins presented on Major Histocompatibility Complexes (MHC) class I and II on the cell surface.
+are T cells which recognize peptides, called epitopes, derived from antigenic proteins presented on Major Histocompatibility Complexes (MHC) class I and II on the cell surface.
 
@@
 For example, a study by Jones \cite{jones2007improving} found that,
-from 184 transmembrane proteins (TMPs) with known \emph{in vitro} topology, 
+from 184 transmembrane proteins (TMPs) with known topology, 
 80\% of the TMH predictions of these proteins
@@
 We confirmed the presentation of TMH-derived peptides 
-by re-analysis of peptide elution studies. 
+by re-analysis of peptides from The Immune Epitope Database (IEDB) \cite{vita2019immune}.
 Second, we tested our hypothesis that TMHs 
@@
 To obtain experimental confirmation that peptides stemming from TMHs 
-are presented \emph{in vivo} in MHC-I and MHC-II,
-we mined the Immune Epitope Database (IEDB) \cite{vita2019immune}
+are presented by MHC-I and MHC-II,
+we mined the IEDB \cite{vita2019immune}
 for confirmed human MHC-ligands.
@@
 which contains 650 million 
-cataloged non-redundant humane variations (called RefSNPs,
+cataloged non-redundant human variations (called RefSNPs,
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \cite{brown2015gene})
@@
 The Immune Epitope Database (IEDB) from the National Institutes of Health contains millions of linear epitope sequences obtained
-by an MHC ligand assay.
+by MHC ligand assays.
 For the MHC alleles used in this study, 
@@
 
-Figure \ref{fig:2a} shows that there are similar levels of
+Figure (\ref{fig:2a} and \ref{fig:thm_presentation_correlation}) shows there are similar levels of
 over-presentation of TMH-derived epitopes between (1) the
@@
 versus (2) the percentage of TMH-derived epitopes that is predicted to be presented
-in MHC-I alleles. For MHC-II alleles, there were too few epitopes per MHC allele
+in MHC-I alleles.
+For MHC-II alleles, there were too few epitopes per MHC allele
 to result in an informative figure.
-We showed that there is a correlation between epitopes taken from
-predicted 9-mers (for MHC-I) and 14-mers (for MHC-II) from a human
-reference proteome and epitopes taken from the IEDB for MHC ligands
-essays (see in Supplementary figure \ref{fig:thm_presentation_correlation}).
 
@@
 by the percentage of TMH-derived epitopes,
-which are 22\% and 10\%.
+which are 22\% and 10\%, respectively.
 
@@
 \subsection{Mechanism of MHC presentation of TMH-derived epitopes}
-% \paragraph{TMH presentation in vivo}
+% \paragraph{TMH presentation}
 
@@
     \textbf{
-      Robust prediction that TMH epitopes are presented \emph{in vivo}.
-    }
-    The percentage of epitopes for MHC-I and -II alleles
-    that overlap with TMHs that are presented.
-    The pair of horizontal red lines in each plot indicate the lower and upper bound 
-    of the 99\% confidence interval. Note that only one line is visible as this
-    interval is relatively narrow.
-    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
-    \textbf{(A)} 
-    Observed and predicted percentage of TMH-derived epitopes
-    \textbf{(B)} 
-    MHC ligands from IEDB binding to TMH-derived epitopes.
-    The numbers above the bars denotes the number of epitopes
-    obtained using an MHC ligand assay. 
+      Robust prediction that TMH epitopes are presented.
+    }
+    The percentage of epitopes for MHC-I and -II alleles that overlap with TMHs
+that are presented. The pair of horizontal red lines in each plot indicate the
+lower and upper bound of the 99\% confidence interval. Note that only one line
+is visible as this interval is relatively narrow. A\ref{tab:haplotype_abbreviations}).
+ \textbf{(A)} Observed and predicted percentage of TMH-derived epitopes for MHC-I alleles. \textbf{(B)} MHC
+ligands from IEDB corresponding to TMH-derived epitopes. The numbers above the
+bars denotes the number of TMH derived epitopes obtained.
+
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Presentation of TMH-derived epitopes \emph{in vivo} in T cells}
+\subsection{Presentation of TMH-derived epitopes result in T cell responses}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 shows the percentage of TMH-derived
-epitopes of the reported epitopes that for which T-cell responses were established. 
+epitopes of the reported epitopes from human origin for which T-cell responses were established. 
 The data was obtained from the IEDB and includes only the MHC

### 2021-11-19 split_pdf.sh
Split PDF correctly

@@ -10,6 +10,6 @@
 #
 #   ./split_pdf.sh
 #
-pdftk bbbq_article.pdf cat 1-27 output bbbq_article_main.pdf
-pdftk bbbq_article.pdf cat 29-32 output bbbq_article_figures.pdf
-pdftk bbbq_article.pdf cat 33-55 output bbbq_article_supplementary_materials.pdf
+pdftk bbbq_article.pdf cat 1-28 output bbbq_article_main.pdf
+pdftk bbbq_article.pdf cat 30-33 output bbbq_article_figures.pdf
+pdftk bbbq_article.pdf cat 34-65 output bbbq_article_supplementary_materials.pdf

### 2021-11-19 bbbq_article.tex
Fix LaTeX doc

@@
 Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
-The peptide-binding grooves of the alleles of most haplotypes
+The peptide-binding grooves of most alleles
 of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
@@
 In humans, the peptide binding groove of MHC-I is made by only the alpha subunit.
-There are three classical haplotypes of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells.
+There are three classical alleles of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells.
 For MHC-II, both the alpha and the beta chains contribute to the peptide binding groove.
-There are three classical haplotypes of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
+There are three classical alleles of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
 Each MHC complex can present a subset of all possible peptides.
@@
 
-Humans mostly express two alleles per MHC haplotype, one from the parental and one from the maternal chromosome, 
+Humans express a limited set of MHC alleles
 and therefore an individual's immune system detects 
@@
 a similar analysis strategy was applied as described in \cite{bianchi2017} 
-for several haplotypes of both MHC-I and MHC-II, 
+for several alleles of both MHC-I and MHC-II, 
 and for a human, viral and bacterial proteome.
@@
 For each of these peptides, we determined if it overlapped with a predicted 
-TMH and if it was predicted to bind to the most frequent alleles of each haplotype.
+TMH and if it was predicted to bind to the most frequent alleles of each MHC allele.
 
@@
 to have a high prevalence in the population,
-where the alleles of the MHC-I haplotypes are the alleles representing the 13 supertypes 
+where the alleles of MHC-I are the alleles representing the 13 supertypes 
 with over 99.6\% coverage of the population's MHC-I repertoire as defined by \cite{lund2004definition} \cite{sette1999},
-and the 21 MHC-II haplotype alleles, have a phenotypic frequency 
+and the 21 MHC-II alleles, have a phenotypic frequency 
 of 14\% or more in the human population \cite{greenbaum2011functional}.
 
-We define a protein to be a binder if, for a certain haplotype, 
+We define a protein to be a binder if, for a certain MHC allele, 
 any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
@@
 This indicates that TMH-derived peptides are over-presented, 
-because the peptide binding grooves of most MHC-I and -II haplotypes 
+because the peptide binding grooves of most MHC-I and -II alleles 
 are relatively hydrophobic.
@@
 Although our data show that 
-TMH-derived epitopes are presented in all classical MHC-I and MHC-II haplotypes, 
+TMH-derived epitopes are presented in all classical MHC-I and MHC-II alleles, 
 the molecular mechanisms of how integral membrane proteins are processed 
@@
 which is in TMPs that span the membrane at least twice.
-In contrast to single-spanners, where we found no significant evolutionary conservation, the TMHs of multi-spanners are more evolutionary conserved than soluble protein regions.
-Likely, the TMHs in many multi-spanners need to interact which each other for correct protein structure and function and they might hence be more structurally constrained compared to the TMHs of single-spanners.
+In contrast to single-spanners, where we found no significant evolutionary conservation, 
+the TMHs of multi-spanners are more evolutionary conserved than soluble protein regions. 
+Likely, the TMHs in many multi-spanners need to interact which each other 
+for correct protein structure and function 
+and they might hence be more structurally constrained 
+compared to the TMHs of single-spanners.
 Thus, we speculate that the human immune system is more attentive 
@@
 First, the MHC over-presentation of TMHs is likely a general feature 
-and predicted to occur for most haplotypes of both MHC-I and -II 
+and predicted to occur for most alleles of both MHC-I and -II 
 and for humans as well as bacterial and viral pathogens. 
@@
   \end{subfigure}  
-  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II haplotypes}
+  \caption{ \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II alleles}
     \textbf{(A)} 
-    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
+    The percentage of epitopes for MHC-I and -II alleles that are predicted to 
     overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
@@
     Diagonal red line: linear regression analysis. 
-    Labels are shorthand for the HLA haplotypes,
+    Labels are shorthand for the HLA alleles,
     see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.
@@
     lower and upper bound of the 99\% confidence interval. Note that only one line
-is visible as this interval is relatively narrow. A\ref{tab:haplotype_abbreviations}).
- \textbf{(A)} Observed and predicted percentage of TMH-derived epitopes for MHC-I alleles. \textbf{(B)} MHC
-ligands from IEDB corresponding to TMH-derived epitopes. The numbers above the
-bars denotes the number of TMH derived epitopes obtained.
-
+    is visible as this interval is relatively narrow. 
+    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
+    \textbf{(A)} 
+    Observed and predicted percentage of TMH-derived epitopes for MHC-I alleles. 
+    \textbf{(B)} 
+    MHC ligands from IEDB corresponding to TMH-derived epitopes. 
+    The numbers above the bars denotes the number of TMH derived epitopes obtained.
   }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{IC50 values of binders per haplotype}
+\subsection{IC50 values of binders per MHC allele}
 \label{subsec:ic50s_per_haplotype}
@@
 than the epitope length.
-From these epitopes, per MHC haplotype,
+From these epitopes, per MHC allele,
 we predicted the IC50 (in nM) using \verb;epitope-prediction; (for MHC-I)
 and MHCnuggets (for MHC-II). 
-Here, we show the IC50 value per haplotype that
-is used to determine if a peptide binds to the haplotype's MHC
+Here, we show the IC50 value per MHC allele that
+is used to determine if a peptide binds to the allele's MHC
 for MHC-I (see supplementary Table \ref{tab:ic50_binders_mhc1}) and 
@@
 epitopes between these two sources of data.
-Figure \ref{fig:thm_presentation_correlation} shows per haplotype
+Figure \ref{fig:thm_presentation_correlation} shows per MHC allele
 the percentage of TMH-derived epitopes, with a linear trendline.
@@
 Supplementary Table \ref{tab:haplotype_abbreviations} shows the
-shorthand notation for the HLA haplotypes.
+shorthand notation for the HLA alleles.
 
@@
 Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
-for a binder per haplotype.
+for a binder per MHC allele.
 
@@
 in figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
-The figures that haplotypes that over-present or under-present TMH-derived epitopes,
+The figures that MHC molecules that over-present or under-present TMH-derived epitopes,
 do so in both studies. The extent to which TMH-derived epitopes are
@@
 Figure \ref{fig:f_tmh_mhc1_normalized} and \ref{fig:f_tmh_mhc2_normalized}
-show these normalized values for the MHC-I and MHC-II haplotypes respectively.
+show these normalized values for the MHC-I and MHC-II alleles respectively.
 
@@
 matches the values as expected by chance.
-The normalized values per haplotype are shown 
+The normalized values per MHC allele are shown 
 in figure \ref{fig:rel_presentation_per_haplotype}.
 To compare the TMH-derived over-presentation per MHC class,
-we grouped the normalized values per haplotype, 
+we grouped the normalized values per allele, 
 and plot the mean and standard error, as shown in figure \ref{fig:rel_presentation}.
@@
     Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs,
-    for the different haplotypes and proteomes
+    for the different MHC alleles and proteomes
   }
@@
   \caption{
-    The percentage of epitopes for MHC-I and -II haplotypes that are predicted to 
+    The percentage of epitopes for MHC-I and -II alleles that are predicted to 
     overlap with TMHs (for at least two amino acids) 

### 2021-11-19 zip_for_frontiers.sh
Zip on GHA

@@ -14,7 +14,7 @@ chmod +x split_pdf.sh
 # Creates the numbered TIFF files in 'numbered_figure_files'
 Rscript create_numbered_figure_files.R
 
-zip ~/frontiers.zip \
+zip frontiers.zip \
   bbbq_article.bib \
   bbbq_article.tex \
   bbbq_header.bbl \

### 2021-11-19 create_merged_tex.R
Merge TeX into one big file, progress #261

@@ -1,0 +1,25 @@
+# Create one merged .tex file from the multiple one
+#
+# Output:
+# * manuscript.tex
+# * manuscript.bib
+output_bib_filename <- "manuscript.bib"
+output_tex_filename <- "manuscript.tex"
+
+# Input
+# * bbbq_header.tex
+# * bbbq_article.tex
+# * bbbq_article.bib
+input_bib_filename <- "bbbq_article.bib"
+input_header_tex_filename <- "bbbq_header.tex"
+input_main_tex_filename <- "bbbq_article.tex"
+testthat::expect_true(file.exists(input_bib_filename))
+testthat::expect_true(file.exists(input_header_tex_filename))
+testthat::expect_true(file.exists(input_main_tex_filename))
+
+# Steps
+#
+# * 1. Copy .bib file
+# * 2. Merge .tex files
+
+file.copy(from = input_bib_filename, to = output_bib_filename, overwrite = TRUE)

### 2021-11-19 create_merged_tex.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+Rscript create_merged_tex.R

### 2021-11-19 zip_for_frontiers.sh
Frontiers wants manuscript.tex and manuscript.bib, not multiple files, progress #261

@@ -18,6 +18,9 @@ zip frontiers.zip \
   revision/README.md \
   revision/revision.docx \
   revision/revision.pdf \
+  manuscript.bib \
+  manuscript.tex \
+  bbbq_article.tex \
   bbbq_article.bib \
   bbbq_article.tex \
   bbbq_header.bbl \

### 2021-11-19 manuscript.tex
Hack together 'manuscript.*'

--- 2021-11-19 bbbq_article.tex
@@
+\documentclass{article}
+
+% Just to have them already
+% \begin{sloppypar}
+% \end{sloppypar}
+
+% Just try to parse, do not ask for input
+\nonstopmode
+
+% Bibliography
+% For Vancounter style:
+\usepackage[numbers]{natbib}
+% For regular style:
+% \usepackage{natbib}
+
+%\bibpunct{(}{)}{;}{a}{}{;}
+
+% Use 'It was found that A is B (Name 1234)' style
+%\setcitestyle{authoryear,open={},close={}}
+
+% Affiliations
+\usepackage{authblk}
+\title{
+  Transmembrane helices are 
+  an overlooked and evolutionarily conserved
+  source of major histocompatibility complex class I and II epitopes
+}
+\author[1]{Richèl J.C. Bilderbeek}
+\author[1]{Maksim V. Baranov}
+\author[1]{Geert van den Bogaart}
+\author[1]{Frans Bianchi}
+\affil[1]{GBB, University of Groningen, Groningen, The Netherlands}
+
+% Use double spacing
+\usepackage{setspace}
+\doublespacing
+
+\usepackage{listings}
+\usepackage{hyperref}
+\usepackage{todonotes}
+\usepackage{verbatim}
+\usepackage{pgf}
+\usepackage{bm}
+\usepackage{multirow}
+\usepackage{amsfonts}
+\usepackage{array}
+\usepackage{booktabs}
+\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
+\usepackage{longtable}
+
+% sidewaysfigure
+\usepackage{rotating}
+
+% Style of listings
+% From http://r.789695.n4.nabble.com/
+%   How-to-nicely-display-R-code-with-the-LaTeX-package-listings-tp4648110.html
+\usepackage{fancyvrb} 
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolor}{rgb}{0.95,0.95,0.92}
+\lstdefinestyle{mystyle}{
+  language=R,% set programming language
+  basicstyle=\ttfamily\small,% basic font style
+  commentstyle=\color{gray},% comment style
+  numberstyle=\scriptsize,% use small line numbers
+  numbersep=10pt,% space between line numbers and code
+  tabsize=2,% sizes of tabs
+  showstringspaces=false,
+  captionpos=b,% positioning of the caption below
+  breaklines=true,% automatic line breaking
+  escapeinside={(*}{*)},% escaping to LaTeX
+  fancyvrb=true,% verbatim code is typset by listings
+  extendedchars=false,% prohibit extended chars (chars of codes 128--255)
+  alsoletter={.<-},% becomes a letter
+  alsoother={$},% becomes other
+  otherkeywords={!=, ~, $, \&, \%/\%, \%*\%, \%\%, <-, <<-, /},
+  deletekeywords={c}% remove keywords 
+}
+\lstset{style=mystyle}
+
+% Adds numbered lines
+\usepackage{lineno}
+\linenumbers
+
+% Subfigures
+\usepackage{subcaption}
+
+% Subfigure numbers use uppercase, 
+% from https://tex.stackexchange.com/a/278140
+\renewcommand{\thesubfigure}{\Alph{subfigure}}
+
+%comments
+\newcommand{\frans}[1]{\textcolor{blue}{\textbf{[FB: #1]}}}
+\newcommand{\richel}[1]{\textcolor{orange}{\textbf{[RB: #1]}}}
+\newcommand{\geert}[1]{\textcolor{red}{\textbf{[GVDB: #1]}}}
+\newcommand{\maksim}[1]{\textcolor{blue}{\textbf{[MVB: #1]}}}
+\newcommand{\maxim}[1]{\textcolor{blue}{\textbf{[MVB: #1]}}}
+
+\begin{document}
+
+\maketitle
+
 \begin{abstract}
@@
     \textbf{
-      Robust prediction that TMH epitopes are presented.
+      Analysis of epitope database shows that TMH derived epitopes are over presented.
     }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Differences with Bianchi et al., 2017}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+A part of this study does the same analysis as Bianchi et al., 2017.
+mainly concern the use of different
+software and a different definition of what an MHC binder is.
+
+% \paragraph{Definition of what a binder is}
+
+The earlier study defined a peptide an MHC binder 
+if \emph{within the protein} in which it was found, 
+is was among the peptides with the 2\% lowest IC50 values.
+This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
+where the binders are written to file.
+
+However, in this study, an MHC binder is defined as a peptide within a \emph{proteome} in which it is found, that is among the peptides with the 2\% lowest IC50 values.
+Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
+for a binder per MHC allele.
+
+% \paragraph{Selenoproteins}
+
+Our previous study used the TMHMM web server
+to predict TMHs.
+The desktop version of TMHMM, however, gives an
+error message on the 25 selenoproteins found in the human
+reference proteome.
+For the sake of reproducible research, we used the desktop version (as
+we can call it from scripts) and, due to this, we removed the
+selenoproteins from this analysis.
+
+% \paragraph{Compare results in figure}
+
+To verify if the previous and the current method give rise to
+notable difference, we show a side-by-side comparison
+in figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
+The figures that MHC molecules that over-present or under-present TMH-derived epitopes,
+do so in both studies. The extent to which TMH-derived epitopes are
+presented, however, is more extreme in our current setup.
+
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/figure-1-a.png} % use same naming as them
+    \label{fig:bianch_et_al_2017_1a}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \caption{}
+    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1.png}
+    \label{fig:bilderbeek_et_al_2021_1a}
+  \end{subfigure}
+
+  \caption{
+    \textbf{(A)} 
+    Results for \cite{bianchi2017}. 
+    Dashed lines denotes the coincidence interval.
+    \textbf{(B)}
+    Results for this study.
+    Dashed line denotes the percentage as expected by chance.
+  }
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 visualized by figure \ref{fig:bbbq_1_smart_results}.
-
-% Process all floats before going to a next page
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Differences with Bianchi et al., 2017}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-A part of this study does the same analysis as Bianchi et al., 2017.
-mainly concern the use of different
-software and a different definition of what an MHC binder is.
-
-% \paragraph{Definition of what a binder is}
-
-The earlier study defined a peptide an MHC binder 
-if \emph{within the protein} in which it was found, 
-is was among the peptides with the 2\% lowest IC50 values.
-This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
-where the binders are written to file.
-
-However, in this study, an MHC binder is defined as a peptide within a \emph{proteome} in which it is found, that is among the peptides with the 2\% lowest IC50 values.
-Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
-for a binder per MHC allele.
-
-% \paragraph{Selenoproteins}
-
-Our previous study used the TMHMM web server
-to predict TMHs.
-The desktop version of TMHMM, however, gives an
-error message on the 25 selenoproteins found in the human
-reference proteome.
-For the sake of reproducible research, we used the desktop version (as
-we can call it from scripts) and, due to this, we removed the
-selenoproteins from this analysis.
-
-% \paragraph{Compare results in figure}
-
-To verify if the previous and the current method give rise to
-notable difference, we show a side-by-side comparison
-in figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
-The figures that MHC molecules that over-present or under-present TMH-derived epitopes,
-do so in both studies. The extent to which TMH-derived epitopes are
-presented, however, is more extreme in our current setup.
-
-\begin{figure}
-  \centering
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{bianchi_et_al_2017_results/figure-1-a.png} % use same naming as them
-    \label{fig:bianch_et_al_2017_1a}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \caption{}
-    \includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1.png}
-    \label{fig:bilderbeek_et_al_2021_1a}
-  \end{subfigure}
-
-  \caption{
-    \textbf{(A)} 
-    Results for \cite{bianchi2017}. 
-    Dashed lines denotes the coincidence interval.
-    \textbf{(B)}
-    Results for this study.
-    Dashed line denotes the percentage as expected by chance.
-  }
-\end{figure}
 
@@
 \clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\section{Figures}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+%
+% Figure 1
+%
+\textbf{Figure 1: Over-presentation of TMH-derived epitopes on most MHC-I and -II alleles}
+    \textbf{(A)} 
+    The percentage of epitopes for MHC-I and -II alleles that are predicted to 
+    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
+    row) and \emph{M. tuberculosis} (MtB; bottom row).
+    The pair of horizontal red lines in each plot indicate the lower and upper bound 
+    of the 99\% confidence interval.
+    See supplementary Tables \ref{tab:tmh_binders_mhc2} and \ref{tab:tmh_binders_mhc1}
+    for the exact TMH and  epitope counts.
+    \textbf{(B-C)}
+    Correlation between the percentages of predicted TMH-derived epitopes
+    and the hydrophobicity score of all predicted epitopes for 
+    human MHC-I \textbf{(B)} and MHC-II alleles \textbf{(C)}.
+    Diagonal red line: linear regression analysis. 
+    Labels are shorthand for the HLA alleles,
+    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.
+
+% Process all floats before going to a next page
+\clearpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+%
+% Figure 2
+%
+\textbf{Figure 2: Analysis of epitope database shows that TMH derived epitopes are over presented.}
+    The percentage of epitopes for MHC-I and -II alleles that overlap with TMHs
+    that are presented. The pair of horizontal red lines in each plot indicate the
+    lower and upper bound of the 99\% confidence interval. Note that only one line
+    is visible as this interval is relatively narrow. 
+    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
+    \textbf{(A)} 
+    Observed and predicted percentage of TMH-derived epitopes for MHC-I alleles. 
+    \textbf{(B)} 
+    MHC ligands from IEDB corresponding to TMH-derived epitopes. 
+    The numbers above the bars denotes the number of TMH derived epitopes obtained.
+
+% Process all floats before going to a next page
+\clearpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+%
+% Figure 3
+%
+\textbf{Figure 3: Evolutionary conservation of human TMHs.}
+    \textbf{(A)} 
+    Percentage of SNPs found in TMHs.
+    Each point shows for one protein the predicted percentage of
+    amino acids that are part of a TMH ($x$-axis) and the observed occurrence of SNPs being located
+    within a TMH ($y$-axis).
+    The dashed diagonal line shows the line of equality (i.e.,
+    equal conservation of TMHs and soluble protein regions). 
+    The diagonal red line indicates a linear fit, 
+    the gray area its 95\% confidence interval.
+    \textbf{(B)}
+    Distribution of the percentages of TMH in the TMPs used in this study.
+    \textbf{(C)}
+    The number of SNPs in TMHs as expected by chance (left bar) 
+    and found in the dbSNP database (right bar).
+    Percentages show the relative conservation
+    of SNPs in TMHs found relative to stochastic chance.
+
+% Process all floats before going to a next page
+\clearpage
+
+% No page number on this page
+\thispagestyle{empty}
+
+%
+% Figure 4
+%
+\textbf{Figure 4: Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
+      \textbf{(A)} 
+      Percentage of SNPs found in TMPs predicted to have only a single
+      (left) or multiple (right) TMHs.
+      Each point shows for one protein the predicted percentage of amino acids that are part of a TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
+The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
+      The diagonal red lines indicate a linear fit, the gray areas their 95\% confidence intervals.
+      \textbf{(B)} 
+      The number of SNPs in TMHs as expected by chance 
+      and observed in the dbSNP database, 
+      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
+      Percentages show the relative conservation
+      of SNPs in TMHs found relative to the stochastic chances.
+      \textbf{(C)} 
+      Distribution of the proportion of amino acids residing
+      in the plasma membrane. 
+     }
+
+% Process all floats before going to a next page
+\clearpage
+
+\end{document}

### 2021-11-19 create_manuscript.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+#
+# Create the PDF 'manuscript.pdf'
+#
+# It does so by converting the TeX to PDF, 
+# as well as stitching some other PDFs to it
+#
+# Usage:
+#
+#   ./create_manuscript.sh
+#
+pdflatex manuscript.tex
+bibtex manuscript >/dev/null
+pdflatex manuscript.tex >/dev/null
+pdflatex manuscript.tex
+

### 2021-11-19 create_merged_figures.R

@@ -1,0 +1,16 @@
+filenames <- list.files("numbered_figure_files", pattern = "tiff", full.names = TRUE)
+filenames
+testthat::expect_equal(length(filenames), 12)
+
+fig_1_filenames <- stringr::str_subset(filenames, pattern = "figure_1")
+fig_1_output_filename <- file.path(dirname(fig_1_filenames[1]), "figure_1.tiff")
+
+args <- c(
+  "montage",
+  "-density", "300x300", 
+  "-units", "PixelsPerInch",
+  "-size", "3072x2048",
+  fig_1_filenames, fig_1_output_filename
+)
+system2(command = args[1], args = args[-1])
+

### 2021-11-19 create_merged_tex.sh

@@ -1,2 +1,0 @@
-#!/bin/bash
-Rscript create_merged_tex.R

### 2021-11-19 create_numbered_figure_files.R

@@ -9,9 +9,14 @@ includegraphics_lines <- stringr::str_subset(tex_lines, "^  .*includegraphics")
 png_filenames <- stringr::str_match(includegraphics_lines, "includegraphics.*\\{(.*)\\}")[, 2]
 testthat::expect_true(all(file.exists(png_filenames)))
 tiff_filenames <- stringr::str_replace(png_filenames, "png$", "tiff")
+eps_filenames <- stringr::str_replace(png_filenames, "png$", "eps")
 
 tiff_filenames[!file.exists(tiff_filenames)]
 testthat::expect_true(all(file.exists(tiff_filenames)))
+
+message("Not all .eps files can be found")
+eps_filenames[!file.exists(eps_filenames)]
+# testthat::expect_true(all(file.exists(eps_filenames)))
 
 figure_numbers <- c(
   paste0(1, letters[1:3]),

### 2021-11-19 split_pdf_manuscript.sh
Submit 'manuscript_*.pdf'

@@ -1,0 +1,15 @@
+#!/bin/bash
+#
+# From the PDF 'manuscript.pdf', create these new files:
+#
+# (1) 'manuscript_main.pdf', the paper,  
+# (2) 'manuscript_figures.pdf', the figures of the main part,  
+# (3) 'manuscript_supplementary_materials.pdf', the supplementary materials
+#
+# usage:
+#
+#   ./split_pdf.sh
+#
+pdftk manuscript.pdf cat 1-29 68-71 output manuscript_main.pdf
+pdftk manuscript.pdf cat 31-34 output manuscript_figures.pdf
+pdftk manuscript.pdf cat 35-66 output manuscript_supplementary_materials.pdf

### 2021-11-30 bbbq_article.tex
Article refer to figures ascendingly
Process reviewer feedback, thanks @fransbianchi and @bogeert

@@
 that TMHs are important for the T cell-mediated adaptive immune system, 
-and hence are of  
-overlooked importance in vaccine development.
+and hence are of importance in vaccine development.
 
@@
 In this study, we provide evidence that epitopes
-derived from TMHs are a major but overlooked source of MHC epitopes. 
+derived from TMHs are a major source of MHC epitopes. 
 Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance for both MHC-I and MHC-II, regardless of the organism.

### 2021-12-12 frontiers.tex
Use 'create_frontiers_manuscript.sh' to create the Frontiers manuscript

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.   
%                                              %
%                                                                                                                                                         %
% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 3.4 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
%\documentclass[utf8]{frontiersHLTH} % for Health articles
%\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles

%\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

\linenumbers

% Leave a blank line between paragraphs instead of using \\

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Sample {et~al.}} %use et al only if is more than 1 author
\def\Authors{First Author\,$^{1,*}$, Co-Author\,$^{2}$ and Co-Author\,$^{1,2}$}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{$^{1}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X \\
$^{2}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X  }
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Corresponding Author}

\def\corrEmail{email@uni.edu}

\begin{document}
\onecolumn
\firstpage{1}

\title[Transmembrane helices are 
  an overlooked and evolutionarily conserved
  source of major histocompatibility complex class I and II epitopes
]{Transmembrane helices are 
  an overlooked and evolutionarily conserved
  source of major histocompatibility complex class I and II epitopes
}

\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
\address{} %This field will be automatically populated
\correspondance{} %This field will be automatically populated

\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}

\maketitle

\begin{abstract}

%%% Leave the Abstract empty if your article does not require one, please see the Summary Table for full details.
\section{}
For full guidelines regarding your manuscript please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines}{Author Guidelines}.

As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract. Leave the Abstract empty if your article does not require one, please see \href{http://www.frontiersin.org/about/AuthorGuidelines#SummaryTable}{Summary Table} for details according to article type.

\tiny
 \keyFont{ \section{Keywords:} keyword, keyword, keyword, keyword, keyword, keyword, keyword, keyword} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
\end{abstract}

\section{Introduction}

For Original Research Articles \citep{conference}, Clinical Trial Articles \citep{article}, and Technology Reports \citep{patent}, the introduction should be succinct, with no subheadings \citep{book}. For Case Reports the Introduction should include symptoms at presentation \citep{chapter}, physical exams and lab results \citep{dataset}.

\section{Article types}

For requirements for a specific article type please refer to the Article Types on any Frontiers journal page. Please also refer to  \href{http://home.frontiersin.org/about/author-guidelines#Sections}{Author Guidelines} for further information on how to organize your manuscript in the required sections or their equivalents for your field

% For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.

\section{Manuscript Formatting}

\subsection{Heading Levels}

%There are 5 heading levels

\subsection{Level 2}
\subsubsection{Level 3}
\paragraph{Level 4}
\subparagraph{Level 5}

\subsection{Equations}
Equations should be inserted in editable format from the equation editor.

\begin{equation}
\sum x+ y =Z\label{eq:01}
\end{equation}

\subsection{Figures}
Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript. Figures will then be automatically embedded at the bottom of the submitted manuscript. Kindly ensure that each table and figure is mentioned in the text and in numerical order. Figures must be of sufficient resolution for publication \href{http://home.frontiersin.org/about/author-guidelines#ResolutionRequirements}{see here for examples and minimum requirements}. Figures which are not according to the guidelines will cause substantial delay during the production process. Please see \href{http://home.frontiersin.org/about/author-guidelines#GeneralStyleGuidelinesforFigures}{here} for full figure guidelines. Cite figures with subfigures as figure \ref{fig:2}B.

\subsubsection{Permission to Reuse and Copyright}
Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet). It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
%%Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet). It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.

\subsection{Tables}
Tables should be inserted at the end of the manuscript. Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted. Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space. These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance. The author will be notified during the typesetting of the final article if this is the case.

\section{Nomenclature}

\subsection{Resource Identification Initiative}
To take part in the Resource Identification Initiative, please use the corresponding catalog number and RRID in your current manuscript. For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.

\subsection{Life Science Identifiers}
Life Science Identifiers (LSIDs) for ZOOBANK registered names or nomenclatural acts should be listed in the manuscript before the keywords. For more information on LSIDs please see \href{http://www.frontiersin.org/about/AuthorGuidelines#InclusionofZoologicalNomenclature}{Inclusion of Zoological Nomenclature} section of the guidelines.

\section{Additional Requirements}

For additional requirements for specific article types and further information please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines#AdditionalRequirements}{Author Guidelines}.

\section*{Conflict of Interest Statement}
%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed. If no such relationship exists, authors will be asked to confirm the following statement:

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

The Author Contributions section is mandatory for all articles, including articles by sole authors. If an appropriate statement is not provided on submission, a standard one will be inserted during the production process. The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work. Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.

\section*{Funding}
Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.

\section*{Acknowledgments}
This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.

\section*{Supplemental Data}
 \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure. LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.

\section*{Data Availability Statement}
The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData

\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
\bibliography{test}

%%% Make sure to upload the bib file along with the tex file and PDF
%%% Please see the test.bib file for some examples of references

\section*{Figure captions}

%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png
%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%

\begin{figure}[h!]
\begin{center}
\includegraphics[width=10cm]{logo1}% This is a *.eps file
\end{center}
\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
\end{figure}

\begin{figure}[h!]
\begin{center}
\includegraphics[width=15cm]{logos}
\end{center}
\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
\end{figure}

%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
%%% Frontiers will add the figures at the end of the provisional pdf automatically
%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

\end{document}

### 2021-12-12 create_frontiers_manuscript.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+#
+# Create the PDF 'frontiers.pdf'
+#
+# It does so by converting the TeX to PDF, 
+# as well as stitching some other PDFs to it
+#
+# Usage:
+#
+#   ./create_frontiers_manuscript.sh
+#
+pdflatex frontiers.tex
+bibtex frontiers >/dev/null
+pdflatex frontiers.tex >/dev/null
+pdflatex frontiers.tex
+

### 2021-12-12 frontiers.tex
Authors correct, progress #263

@@
 \def\firstAuthorLast{Sample {et~al.}} %use et al only if is more than 1 author
-\def\Authors{First Author\,$^{1,*}$, Co-Author\,$^{2}$ and Co-Author\,$^{1,2}$}
+\def\Authors{Richèl J.C. Bilderbeek\,$^{1}$, Maksim V. Baranov\,$^{1}$, Geert van den Bogaart\,$^{1}$ and Frans Bianchi\,$^{1,*}$}
+
 % Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
 % If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{$^{1}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X \\
-$^{2}$Laboratory X, Institute X, Department X, Organization X, City X , State XX (only USA, Canada and Australia), Country X  }
-% The Corresponding Author should be marked with an asterisk
+\def\Address{$^{1}$Molecular Immunology, GBB, University of Groningen, Groningen, The Netherlands}
 % Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
-\def\corrAuthor{Corresponding Author}
+\def\corrAuthor{Frans Bianchi, Nijenborgh 7, 9747 AG Groningen}
 
-\def\corrEmail{email@uni.edu}
+\def\corrEmail{f.bianchi@rug.nl}
 

### 2021-12-12 frontiers.tex
Add abstract, progress #263

@@
 \section{}
-For full guidelines regarding your manuscript please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines}{Author Guidelines}.
-
-As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership.
-References should not be cited in the abstract.
-Leave the Abstract empty if your article does not require one, please see \href{http://www.frontiersin.org/about/AuthorGuidelines#SummaryTable}{Summary Table} for details according to article type.
+Cytolytic T cell responses are predicted to be biased towards membrane proteins. 
+The peptide-binding grooves of most alleles
+of histocompatibility complex class I (MHC-I) are relatively hydrophobic, 
+therefore peptide fragments derived from human transmembrane helices (TMHs) are predicted to be presented more often as would be expected 
+based on their abundance in the proteome. However, the physiological reason of why membrane proteins might be over-presented is unclear. 
+In this study, we show that the predicted over-presentation of TMH-derived peptides is general, as it is predicted for bacteria and viruses 
+and for both MHC-I and MHC-II, and confirmed by re-analysis of epitope databases. Moreover, we show that TMHs are evolutionarily more conserved, 
+because single nucleotide polymorphisms (SNPs) are present 
+relatively less frequently in TMH-coding chromosomal regions 
+compared to regions coding for extracellular and cytoplasmic protein regions. 
+Thus, our findings suggest that 
+both cytolytic and helper T cells are more tuned to respond to membrane proteins, 
+because these are evolutionary more conserved. 
+We speculate that TMHs are less prone to mutations 
+that enable pathogens to evade T cell responses.
 

### 2021-12-12 frontiers.tex
Add keywords, progress #263

@@
 \tiny
- \keyFont{ \section{Keywords:} keyword, keyword, keyword, keyword, keyword, keyword, keyword, keyword} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
+ \keyFont{ \section{Keywords:} antigen presentation, membrane proteins, bioinformatics, adaptive immunity, transmembrane domain, transmembrane helix, epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation}
 \end{abstract}

### 2021-12-12 frontiers.tex
Add Introduction and bibliography, progress #263

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-For Original Research Articles \citep{conference}, Clinical Trial Articles \citep{article}, and Technology Reports \citep{patent}, the introduction should be succinct, with no subheadings \citep{book}.
+% For Original Research Articles \citep{conference}, Clinical Trial Articles \citep{article}, and Technology Reports \citep{patent}, the introduction should be succinct, with no subheadings \citep{book}.
 For Case Reports the Introduction should include symptoms at presentation \citep{chapter}, physical exams and lab results \citep{dataset}.
 
+% \paragraph{Immune response}
+
+Our immune system fights diseases and infections from pathogens, 
+such as fungi, bacteria or viruses. 
+An important part of the acquired immune response, 
+that develops specialized and more specific recognition of pathogens 
+than the innate immune response, 
+are T cells which recognize peptides, called epitopes, derived from antigenic proteins presented on Major Histocompatibility Complexes (MHC) class I and II on the cell surface.
+
+% \paragraph{Classification of HLA}
+
+The MHC proteins are heterodimeric complexes encoded by the
+HLA (Human Leukocyte Antigens) genes.
+In humans, the peptide binding groove of MHC-I is made by only the alpha subunit.
+There are three classical alleles of MHC-I, hallmarked by a highly polymorphic alpha chain called HLA-A, HLA-B and HLA-C, that all present epitopes to cytolytic T cells.
+For MHC-II, both the alpha and the beta chains contribute to the peptide binding groove.
+There are three classical alleles of MHC-II as well, called HLA-DR, HLA-DQ and HLA-DP, that all present epitopes to helper T cells.
+Each MHC complex can present a subset of all possible peptides.
+For example, HLA-A and HLA-B have no overlap in which
+epitopes they bind \cite{lund2004definition}.
+Moreover, the HLA genes of humans are highly polymorphic, with hundreds 
+to thousands of different alleles, 
+and each different allele presents a different subset of peptides \cite{marsh2010nomenclature}.
+
+% \paragraph{HLAs increase detection range}
+
+Humans express a limited set of MHC alleles
+and therefore an individual's immune system detects 
+only a fraction of all possible peptide fragments. 
+However, at the population level, the coverage of pathogenic peptides that are detected 
+is very high, because of the highly polymorphic MHC genes.
+It is therefore believed that MHC polymorphism improves immunity at the population level, 
+as mutations in a protein that disrupt a particular MHC presentation at the individual level, 
+so-called escape mutations, 
+will not affect MHC presentation for all alleles present in the population \cite{sommer2005importance}.
+
+% \paragraph{Epitope prediction}
+
+Many studies are aimed at identifying the repertoire of epitopes that are presented 
+in any of the different alleles to determine which epitopes will result in an immune response, 
+as this will for instance aid the design of vaccines. 
+These studies have led to the development of prediction algorithms 
+that allow for very reliable \emph{in silico} 
+predictions of the peptide binding affinities
+\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+For example, S. Tang et al. \cite{tang2011genome} found that, 
+of the 432 peptides that were predicted to bind to an MHC allele,
+86\% were experimentally confirmed to do so.
+
+% \paragraph{TMHs}
+
+Using these prediction algorithms, 
+we recently showed that peptides derived 
+from transmembrane helices (TMHs) 
+are likely to be more frequently presented by MHC-I 
+than expected based on their abundance \cite{bianchi2017},
+which is in line with a previous study 
+by Istrail et al \cite{istrail2004comparative},
+demonstrating that N-terminal signal sequences 
+are likely to be presented within major histocompatibility complexes, 
+due their hydrophobic nature. 
+Moreover, we showed that some well-known immunodominant peptides stem from TMHs. 
+This over-presentation is attributed to the fact 
+that the peptide-binding groove of most MHC-I alleles
+is relatively hydrophobic, 
+and therefore hydrophobic TMH-derived peptides have a higher affinity 
+to bind than their soluble hydrophobic 
+counterparts.
+
+TMHs are hydrophobic 
+as they need to span the hydrophobic lipid bilayer of cellular membranes.
+They consist of an alpha helix of, on average, 23 amino acids in length. 
+TMHs can also be predicted with high accuracy from a protein sequence 
+by bioinformatics approaches \cite{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}. 
+For example, a study by Jones \cite{jones2007improving} found that,
+from 184 transmembrane proteins (TMPs) with known topology, 
+80\% of the TMH predictions of these proteins
+matched the experimental findings.
+TMHs are common structures in the proteins of humans and microbes. 
+Different TMH prediction tools estimate that 15-39\% of all proteins 
+in the human proteome contain at least one TMH \cite{ahram2006estimation}.
+However, the physiological reason why peptides derived from TMHs 
+would be presented more often than peptides 
+stemming from soluble (i.e., extracellular or cytoplasmic) protein regions is unknown. 
+In this study, we hypothesized that the presentation of 
+TMH residues is evolutionarily preferred, 
+since TMHs are less prone to undergo escape mutations. 
+One reason to expect such a reduced 
+variability (and hence evolutionary conservation) in TMHs, 
+is that these are restricted in their variability 
+by the functional requirement to span a lipid bilayer. 
+This limits 
+many of the amino acids present in TMHs 
+to have hydrophobic side chains 
+\cite{hessa2007molecular,jones1994model}.
+Therefore, we speculated that the TMHs of pathogens 
+might have a lower chance to develop escape mutations, 
+as that will result in a dysfunctional TMH 
+and render the protein inactive.
+
+This study had two objectives. 
+First, we aimed to generalize our findings by predicting
+the antigenic presentation
+from different kingdoms of life in both MHC-I and -II. 
+From these \emph{in silico} predictions, we conclude that TMH-derived
+epitopes from a human, viral and bacterial proteome
+are likely to be presented more often than expected by chance 
+for most alleles of MHC-I and II. 
+We confirmed the presentation of TMH-derived peptides 
+by re-analysis of peptides from The Immune Epitope Database (IEDB) \cite{vita2019immune}.
+Second, we tested our hypothesis that TMHs 
+are more evolutionary conserved than soluble protein regions.
+Our analysis of human single nucleotide polymorphisms (SNPs) showed 
+that random point mutations are indeed less likely
+to occur within TMHs. 
+These findings strengthen the emerging notion 
+that TMHs are important for the T cell-mediated adaptive immune system, 
+and hence are of  
+overlooked importance in vaccine development.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Article types}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
-\bibliography{test}
+\bibliography{bbbq_article}
 

### 2021-12-12 frontiers.tex
Add main text, progress #263

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Article types}
+\section{Methods}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-For requirements for a specific article type please refer to the Article Types on any Frontiers journal page.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Predicting TMH epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To predict how frequently epitopes overlapping with TMHs are presented,
+a similar analysis strategy was applied as described in \cite{bianchi2017} 
+for several alleles of both MHC-I and MHC-II, 
+and for a human, viral and bacterial proteome.
+To summarize, for each proteome, 
+all possible 9-mers (for MHC-I) or 14-mers (MHC-II) were derived. 
+For each of these peptides, we determined if it overlapped with a predicted 
+TMH and if it was predicted to bind to the most frequent alleles of each MHC allele.
+
+For MHC-I, 9-mers were used, 
+as this is the length most frequently presented in MHC-I 
+and was used in our earlier study \cite{bianchi2017}. 
+For MHC-II, 14-mers were used, 
+as this is the most frequently occurring epitope length \cite{bergseng2015different}.
+A human (UniProt ID UP000005640\_9606), 
+viral (SARS-CoV-2, UniProt ID UP000464024) 
+and bacterial (\emph{Mycobacterium tuberculosis}, UniProt ID UP000001584) 
+reference proteome was used. 
+\verb;TMHMM; \cite{krogh2001predicting} was used to predict the topology 
+of the proteins within these proteomes.
+To predict the affinity of an epitope to a certain HLA allele,
+ \verb;EpitopePrediction; \cite{bianchi2017} for MHC-I 
+and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
+Both MCH-I and MHC-II alleles were selected 
+to have a high prevalence in the population,
+where the alleles of MHC-I are the alleles representing the 13 supertypes 
+with over 99.6\% coverage of the population's MHC-I repertoire as defined by \cite{lund2004definition} \cite{sette1999},
+and the 21 MHC-II alleles, have a phenotypic frequency 
+of 14\% or more in the human population \cite{greenbaum2011functional}.
+
+We define a protein to be a binder if, for a certain MHC allele, 
+any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
+all peptides within a \emph{proteome} 
+(see supplementary Tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
+this differs from our previous study where we defined
+a binder as having an IC50 in the lowest 2\% 
+of the peptides within a \emph{protein}.
+% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
+This revised definition precludes bias of proteins 
+that give rise to no or only very few MHC epitopes.
+To verify that the slight change in method yields similar results,
+a side by side comparison is shown in the supplementary materials, 
+Figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH epitopes obtained from experimental data}\label{subsec:elution_studies}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To obtain experimental confirmation that peptides stemming from TMHs 
+are presented by MHC-I and MHC-II,
+we mined the IEDB \cite{vita2019immune}
+for confirmed human MHC-ligands.
+We queried the IEDB for all linear epitopes obtained
+from MHC ligand assays in healthy humans, 
+carrying the MHC alleles as used in this study.
+From these epitopes, we kept those that were present
+exactly once in the human reference proteome
+with UniProt ID UP000005640\_9606.
+We predicted the topology of the protein each epitope
+was found in, using \verb;TMHMM; \cite{krogh2001predicting},
+from which we concluded if the epitope is overlapping with a TMH
+with at least 1 amino acid.
+
+The full analysis can be found
+at \url{https://github.com/richelbilderbeek/bbbq_article_issue_157}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Evolutionary conservation of TMHs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{Introduction}
+
+To determine the evolutionary conservation of TMHs,
+we first collected human single nucleotide polymorphisms (SNPs) 
+resulting in a single amino acid substitution to determine if this occurred within a predicted TMH or not.
+
+% \paragraph{Data}
+
+As a data source, multiple
+NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: 
+the \emph{dbSNP} \cite{sherry2001dbsnp} database,
+which contains 650 million 
+cataloged non-redundant human variations (called RefSNPs,
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \cite{brown2015gene})
+and \emph{protein} (for proteins sequences \cite{sayers2010database}).
+
+% \paragraph{Pipeline}
+
+The first query was a call to the \emph{gene} database for the 
+term 'membrane protein' (in all fields) 
+for the organism \emph{Homo sapiens}.
+This resulted in 1,077 gene IDs (on December 2020).
+% 1077 gene IDs is correct for December 2020.
+% At 2021-03-01, one will get 1130 gene IDs.
+% Also, one of the gene IDs that was valid back then,
+% has been obsoleted.
+The next query was a call to the \emph{gene} database 
+to obtain the gene names from the gene IDs.
+Per gene name, the \emph{dbSNP} NCBI database was queried for 
+variations associated with the gene name. 
+As the NCBI API constrains its users to three calls per second
+(to assure fair use), we had to limit the extent of our analysis.
+
+The number of SNPs was limited to the first 250 variations per gene,
+resulting in $\approx$61k variations.
+Only variations that result in a SNP for
+a single amino acid substitution were analyzed, resulting in $\approx$38k SNPs.
+The exact amounts can be found in the supplementary materials,
+Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
+
+% \paragraph{Selection of SNPs}
+%
+SNPs were picked based on ID number, which is linked to their discovery date.
+To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined.
+This analysis showed no positional bias of the SNPs, as shown in supplementary figure \ref{fig:snp_rel_pos}.
+
+Per SNP, the \emph{protein} NCBI database was queried for the
+protein sequence.
+For each protein sequence, the protein topology was determined using \verb;PureseqTM;.
+Using these predicted protein topologies, the SNPs were scored to be located within or outside TMHs.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH-derived peptides are predicted to be over-presented in MHC-I}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
+for a human, viral and bacterial proteome.
+Per MHC-I allele, it shows the percentage of binders that overlap with a TMH 
+with at least one residue.
+The horizontal line shows the expected percentage of TMH-derived epitopes 
+that would be presented, if TMH-derived epitopes would be presented just as 
+likely as epitopes derived from soluble regions,
+when assuming equal incidence of soluble and TMH-derived epitope presentation.
+For 11 out of 13 MHC-I alleles, TMH-derived epitopes are predicted to be presented more often 
+than the null expectation, for a human and bacterial proteome.
+For the viral proteome, 12 out of 13 MHC-I alleles present
+TMH-derived epitopes more often than expected by chance.
+The extent of the over-presentation between the different alleles
+is similar for the probed proteomes, 
+which strengthens our previous conclusion \cite{bianchi2017} 
+that the hydrophobicity of the MHC-binding groove 
+is the main factor responsible for the predicted over-presentation 
+of TMH-derived peptides.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{TMH-derived peptides are predicted to be over-presented in MHC-II}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We next wondered if the over-representation of TMH-derived peptides 
+would also be confirmed for MHC-II. 
+Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
+predicted to be overlapping with TMHs for our human, viral and bacterial proteomes.
+We found that TMH-derived peptides are over-presented in all
+of the 21 MHC-II alleles, 
+for a human, bacterial and viral proteome,
+except for \verb;HLA-DRB3*0101; in \emph{M. tuberculosis}.
+See supplementary Table \ref{tab:tmh_binders_mhc2} 
+for the exact TMH and epitope counts.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The over-presentation of TMH-derived peptides is caused by the hydrophobicity of the MHC peptide binding groove}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+For MHC-I, we previously showed that the over-presentation of TMH-derived 
+peptides is caused by the hydrophobicity of the peptide binding 
+grooves \cite{bianchi2017}. 
+Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
+show the extent of over-presentation
+of TMH-derived epitopes as a function of the hydrophobicity preference score 
+for the different human MHC alleles.
+An assumed linear correlation explains 88\% of the variability in MHC-I.
+For MHC-II, 62\% of the variability is explained by hydrophobicity. 
+This indicates that TMH-derived peptides are over-presented, 
+because the peptide binding grooves of most MHC-I and -II alleles 
+are relatively hydrophobic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experimental validation of presentation of TMH-derived peptides}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The Immune Epitope Database (IEDB) from the National Institutes of Health contains millions of linear epitope sequences obtained
+by MHC ligand assays.
+For the MHC alleles used in this study, 
+we obtained 54,303 and 2,484 linear epitope sequences for the MHC-I
+and MHC-II alleles from human origin respectively.
+There are relatively few epitopes for MHC-II, 
+as MHC-II has many more different alleles than MHC-I,
+whereas we selected only the human epitopes found for the 21 MHC-II alleles used in this study.
+
+Figure (\ref{fig:2a} and \ref{fig:thm_presentation_correlation}) shows there are similar levels of
+over-presentation of TMH-derived epitopes between (1) the
+percentage of TMH-derived epitopes that is reported in the IEDB database
+versus (2) the percentage of TMH-derived epitopes that is predicted to be presented
+in MHC-I alleles.
+For MHC-II alleles, there were too few epitopes per MHC allele
+to result in an informative figure.
+
+In figure \ref{fig:2b} we grouped all the epitopes 
+presented by MHC-I and MHC-II alleles 
+by the percentage of TMH-derived epitopes,
+which are 22\% and 10\%, respectively.
+
+These findings robustly confirm that
+epitopes derived from human TMHs are presented in both MHC-I and MHC-II, and support that they are over-presented.
+See the supplementary Table \ref{tab:elution} for the exact values.
+
+We also mined the IEDB database for epitopes for
+any type of T-cell response from the specified
+alleles, from the total reports 36\% and 7\% 
+concerned TMH-derived epitopes in MHC class I and II, respectively 
+(see Figure \ref{fig:t_cells_present_tmh_derived_epitopes}).
+
+This data confirms that not only TMH derived epitopes are presented on MHC, 
+but this also elicits T-cell mediated immune responses.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Human TMHs are evolutionarily conserved}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We addressed the question whether there is an evolutionary advantage in presenting TMHs.
+We determined the conservation of TMHs 
+by comparing the occurrences of SNPs located in TMHs or soluble protein regions 
+for the genes coding for membrane proteins.
+We obtained 911 unique gene names associated with the phrase 'membrane protein',
+which are genes coding for both membrane-associated proteins (MAPs, which have no TMH) and 
+transmembrane proteins (TMPs, which have at least one TMH).
+These genes are linked to 4,780 protein isoforms, 
+of which 2,553 are predicted to be TMPs and 
+2,237 proteins are predicted to be MAPs.
+We obtained 37,630 unique variations, 
+of which 9,621 are SNPs that resulted in a straightforward amino acids substitution, 
+of which 6,062 were located in predicted TMPs.
+See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
+for the detailed numbers and distributions of SNPs.
+
+Per protein, we calculated two percentages: 
+(1) the percentage of a protein sequence length bearing TMHs, 
+and (2) the percentage of SNPs located within these predicted TMHs.
+Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
+The proportion of SNPs found in TMHs varied from 
+none (i.e., all SNPs were in
+soluble regions) to all (i.e., all SNPs were in
+TMHs).
+To determine if SNPs were randomly distributed over the protein, we performed a linear regression analysis,
+and added a 95\% confidence interval on this regression.
+This linear fit nearly goes through the origin and has a slope
+below the line of equality,
+which shows that less SNPs are found in TMHs than expected by chance.
+
+We determined the probability to find the observed amount
+of SNPs in TMHs by chance, i.e., when assuming SNPs occur 
+just as likely in soluble domains as in TMHs.
+We used a binomial Poisson distribution, 
+where the number of trails ($n$) equals the number of SNPs, 
+which is 21,208. 
+The probability of success for the $i$th TMP ($p\_i$), 
+is the percentage of residues within a TMH per TMP. 
+These percentages are shown as a histogram 
+in figure \ref{fig:f_tmh_ncbi}. 
+The expected number of SNPs expected to be found in 
+TMHs by chance equals $\sum{p} \approx 4,141$.
+As we observed 3,803 SNPs in TMHs, 
+we calculated the probability of having that amount or less successes.
+We used the type I error cut-off value of $\alpha = 2.5\%$.
+The chance to find, within TMHs, this amount or less SNPs 
+equals $6.8208 \cdot 10^{-11}$.
+We determined the relevance of this finding, by
+calculating how much less SNPs are found in TMHs,
+when compared to soluble regions, which is the
+ratio between the number of SNPs found in TMHs
+versus the number of SNPs as expected by chance.
+In effect, per 1000 SNPs found in soluble protein domains, 
+one finds 918 SNPs in TMHs,
+as depicted (as percentages) in figure \ref{fig:conservation}.
+
+We split this analysis for TMPs containing only a single TMH (so-called single-membrane spanners) and TMPs containing multiple TMHs (multi-membrane spanners). 
+We hypothesized that single-membrane spanners are less conserved than multi-membrane spanners,
+because multi-membrane spanners
+might have protein-protein interactions between their TMHs, 
+for example to accommodate active sites, and 
+thus might have additional structural constraints.
+From the split data, we did the same analysis as for the total TMPs.
+Figure \ref{fig:f_snps_found_and_expected_per_spanner} 
+shows the percentages of TMHs for individual proteins as a function of the
+percentage of SNPs located in TMHs.
+For both single- and multi-spanners, a linear regression shows that less
+SNPs are found in TMHs, than expected by chance.
+
+We also determined the probability to find the 
+observed amount of SNPs  by chance in single- and multi-spanners.
+For single-spanners, we found 452 SNPs in TMH, where
+$\approx462$ were expected by chance. 
+The chance to observe this or a lower number by chance is 
+$0.319$. As this chance was higher than our $\alpha = 0.025$,
+we consider this no significant effect.
+For the multi-spanners, we found 3,351 SNPs in TMH, where 
+$\approx3,678$ were expected by chance. 
+The chance to observe this or a lower number by chance is 
+$8.315841 \cdot 10^{-12}$, 
+which means this number is significantly less as explained by variation.
+The TMHs of multi-spanners are thus significantly more conserved than soluble protein regions, whereas this is not the case for single-spanners.
+
+Also, for single- and multi-spanners, 
+we determined the relevance of this finding by
+calculating how much less SNPs are found in TMHs
+when compared to soluble regions,
+as depicted in figure \ref{fig:conservation_per_spanner}.
+In effect, per 1,000 SNPs found in soluble protein domains, 
+one finds 978 SNPs in TMHs 
+of single-spanners
+and 911 SNPs in TMHs of multi-spanners.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{General}
+
+Epitope prediction is important to understand the immune system function
+and for the design of vaccines.
+In this study, we provide evidence that epitopes
+derived from TMHs are a major but overlooked source of MHC epitopes. 
+Our bioinformatics predictions indicate that the TMH-derived epitope repertoire is larger than expected by chance for both MHC-I and MHC-II, regardless of the organism.
+Moreover, reanalysis of MHC-ligands from the IEDB database confirmed the presentation of TMH-derived epitopes.
+Therefore, it seems likely that TMH-derived epitopes would also result in enhanced T cell responses, although the conservation of TMHs might promote the deletion of T cells responsive to TMH-derived epitopes by central tolerance mechanisms.
+Finally, our SNP analysis shows that TMHs are evolutionary more conserved
+than solvent-exposed protein regions.
+
+\subsection{Mechanism of MHC presentation of TMH-derived epitopes}
+% \paragraph{TMH presentation}
+
+Although our data show that 
+TMH-derived epitopes are presented in all classical MHC-I and MHC-II alleles, 
+the molecular mechanisms of how integral membrane proteins are processed 
+for MHC presentation are largely unknown \cite{bianchi2017}. 
+Most prominently, the fundamental principles of 
+how TMHs are extracted from their hydrophobic lipid environments 
+into the aqueous vacuolar lumen, 
+leading to subsequent proteolytic processing are unresolved.
+
+A first possibility is that the extraction of TMPs from the membrane 
+is mediated by the ER-associated degradation (ERAD) machinery. 
+For MHC class I (MHC-I) antigen presentation of soluble proteins, 
+the loading of the epitope primarily occurs at the endoplasmatic reticulum (ER). 
+The chaperones tapasin (TAPBP), ERp57 (PDIA3), 
+and calreticulin (CALR) \cite{rock2016present} first assemble 
+and stabilize the heavy and light chains of MHC-I. 
+Later, this complex binds to the transporter 
+associated with antigen processing (TAP) 
+leading to the formation of the so-called peptide-loading complex (PLC). 
+The PLC drives import of peptides into the ER 
+and mediates their subsequent loading into the peptide-binding groove of MHC-I \cite{blees2017structure}. 
+Membrane proteins first will have to be extracted from the membrane 
+before they become amenable to this MHC-I loading by the PLC. 
+In the ER, this process can be orchestrated by the ERAD machinery, 
+consisting of several chaperones that recognize TMPs, 
+ubiquitinate them, and extract them from the ER membrane 
+into the cytosol (retrotranslocation) for proteasomal degradation \cite{preston2017evolving,meusser2005erad}. 
+Similar to the peptides generated from soluble proteins, 
+the TMP-derived peptides might then be re-imported by TAP 
+into the ER for MHC-I loading. 
+This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{bougneres2009role}, 
+since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.
+
+A second possibility is that TMPs are proteolytically processed 
+by intramembrane proteases that cleave TMHs while they are still membrane embedded.
+Supporting this hypothesis is the well-established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences,
+are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}. 
+The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, 
+possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}. 
+However, this mechanism cannot explain how most membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at their C-termini, and N-terminal domains will hence not be removed.
+Nevertheless, the presentation of peptides with a high hydrophobicity index 
+was shown to be independent of TAP as well \cite{lautscham2001processing}, 
+suggesting that the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
+
+A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \cite{oliveira2015alternative}. 
+TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes.
+Eventually, these TMPs can be sorted by the endosomal sorting complexes required for transport (ESCRT) pathway into luminal invaginations that pinch off from the limiting membrane and form intraluminal vesicles.
+This thus results in MVBs where the membrane proteins destined for degradation are located in intraluminal vesicles.
+Upon the fusion of MVBs with lysosomes,
+the entire intraluminal vesicles including the TMPs are degraded \cite{gruenberg2020life}. 
+Via this mechanism, TMPs might well be processed for antigen presentation, 
+particularly since the loading of MHC-II molecules is well understood 
+to occur in MVBs \cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}. 
+However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles,
+but not on the limiting membranes of MVBs \cite{zwart2005spatial}, 
+indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
+This observation hence raises the question how the intraluminal vesicles carrying the TMPs destined for antigen presentation can be selectively degraded, while the intraluminal vesicles carrying the MHC-II remain intact.
+A second problem is that phagosomes carrying internalized microbes lack intraluminal vesicles,
+and it is hence unclear how TMPs from these microbes 
+would be routed to MVBs for MHC-II loading \cite{zwart2005spatial}.
+
+Alternatively to the enzymatic degradation of lipids in MVBs by lipases \cite{sander2016lipase,gilleron2016lysosomal}, 
+they might be oxidatively degraded by reactions with radical oxygen species produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}. 
+This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} 
+and might thereby lead to the extraction of TMPs. 
+Due to the hydrophobic nature of TMHs, 
+however, the extracted proteins will likely aggregate 
+and it is unclear how these aggregates would be processed further for MHC loading.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Evolutionary conservation of TMHs}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% \paragraph{Selection undetectable in whole proteome}
+
+In general, one might expect that evolutionary selection shapes an immune system 
+where surveillance is directed towards protein regions 
+essential for the survival, proliferation and/or virulence or pathogenic microbes, 
+as these will be most conserved.
+In SARS-CoV-2, for example, there is preliminary evidence that the strongest
+selection pressure is directed upon residues that change its 
+virulence \cite{velazquez2020positive}.
+These regions, however, may only account for a small part of a pathogen's proteome.
+Additionally, the structure and function of these essential regions might differ widely between different pathogenic proteins.
+Because of this scarcity and variance in targets, 
+one can imagine that it will be mostly unfeasible 
+to provide innate immune responses against such rare essential protein regions, 
+as suggested in a study on influenza \cite{han2019individual},
+where it was found that the selection pressure
+exerted by the immune system was either weak or absent.
+
+% \paragraph{Selection may be detectable in TMHs}
+
+Evolutionary selection of pathogens by a host's immune system,
+however, is more likely to occur for protein patterns that are general,
+over patterns that are rare.
+While essential catalytic sites in a pathogenic proteome
+might be relatively rare, TMHs are common and thus might be a more feasible 
+target for evolution to respond to.
+Indeed, we have found the signature of evolution when both factors,
+that is, TMHs and catalytic sites are likely to co-occur,
+which is in TMPs that span the membrane at least twice.
+In contrast to single-spanners, where we found no significant evolutionary conservation, 
+the TMHs of multi-spanners are more evolutionary conserved than soluble protein regions. 
+Likely, the TMHs in many multi-spanners need to interact which each other 
+for correct protein structure and function 
+and they might hence be more structurally constrained 
+compared to the TMHs of single-spanners.
+Thus, we speculate that the human immune system is more attentive 
+towards TMHs in multi-spanners, as these are evolutionarily more conserved.
+
+There have been more efforts to assess the conservation of TMHs,
+using different methodologies.
+One such example is a study by Stevens and Arkin \cite{stevens2001substitution}, 
+in which aligned protein sequence data was used.
+Also this study found that TMHs are evolutionarily more conserved,
+as the mean amino acid substitution rate in TMHs is about ten
+percent lower,
+which is a similar value as we found.
+Another example is a study by Oberai, et al. \cite{oberai2009structural} that estimated the conservation
+scores for TMHs and soluble regions based on 
+alignments of evolutionary related proteins,
+and also found that TMHs are more conserved, 
+with a conservation score that was 17\% higher in 
+TMHs.
+Note that the last study also found that mutations in human TMHs are likelier to cause
+a disease, in line with our conclusion that TMHs are more conserved.
+
+% \paragraph{Conclusion}
+
+Together, from this study, two important conclusions can be drawn. 
+First, the MHC over-presentation of TMHs is likely a general feature 
+and predicted to occur for most alleles of both MHC-I and -II 
+and for humans as well as bacterial and viral pathogens. 
+Second, TMHs are genuinely more evolutionary conserved than soluble protein motifs, 
+at least in the human proteome.
+
+%
+% Frontiers text, just keep it in for now to be sure.
+%
+%\section{Article types}
+
+% For requirements for a specific article type please refer to the Article Types on any Frontiers journal page.
 Please also refer to  \href{http://home.frontiersin.org/about/author-guidelines#Sections}{Author Guidelines} for further information on how to organize your manuscript in the required sections or their equivalents for your field

### 2021-12-12 frontiers.tex
Add sections below main text, progress #263

@@
 %\section{Article types}
-
+%
 % For requirements for a specific article type please refer to the Article Types on any Frontiers journal page.
 Please also refer to  \href{http://home.frontiersin.org/about/author-guidelines#Sections}{Author Guidelines} for further information on how to organize your manuscript in the required sections or their equivalents for your field
-
+%
 % For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
-
-\section{Manuscript Formatting}
-
-\subsection{Heading Levels}
-
+%
+%\section{Manuscript Formatting}
+%
+%\subsection{Heading Levels}
+%
 %There are 5 heading levels
-
-\subsection{Level 2}
-\subsubsection{Level 3}
-\paragraph{Level 4}
-\subparagraph{Level 5}
-
-\subsection{Equations}
-Equations should be inserted in editable format from the equation editor.
-
-\begin{equation}
-\sum x+ y =Z\label{eq:01}
-\end{equation}
-
-\subsection{Figures}
-Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript.
+%
+%\subsection{Level 2}
+%\subsubsection{Level 3}
+%\paragraph{Level 4}
+%\subparagraph{Level 5}
+%
+%\subsection{Equations}
+%Equations should be inserted in editable format from the equation editor.
+%
+%\begin{equation}
+%\sum x+ y =Z\label{eq:01}
+%\end{equation}
+%
+%\subsection{Figures}
+%Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript.
 Figures will then be automatically embedded at the bottom of the submitted manuscript.
@@
 Cite figures with subfigures as figure \ref{fig:2}B.
-
-\subsubsection{Permission to Reuse and Copyright}
-Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
+%
+%
+%\subsubsection{Permission to Reuse and Copyright}
+%Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
 It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
@@
 It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
-
-\subsection{Tables}
-Tables should be inserted at the end of the manuscript.
+%
+%\subsection{Tables}
+%Tables should be inserted at the end of the manuscript.
 Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
@@
 The author will be notified during the typesetting of the final article if this is the case.
-
-\section{Nomenclature}
-
-\subsection{Resource Identification Initiative}
-To take part in the Resource Identification Initiative, please use the corresponding catalog number and RRID in your current manuscript.
+%
+%\section{Nomenclature}
+%
+%\subsection{Resource Identification Initiative}
+%To take part in the Resource Identification Initiative, please use the corresponding catalog number and RRID in your current manuscript.
 For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.
-
-\subsection{Life Science Identifiers}
-Life Science Identifiers (LSIDs) for ZOOBANK registered names or nomenclatural acts should be listed in the manuscript before the keywords.
+%
+%\subsection{Life Science Identifiers}
+%Life Science Identifiers (LSIDs) for ZOOBANK registered names or nomenclatural acts should be listed in the manuscript before the keywords.
 For more information on LSIDs please see \href{http://www.frontiersin.org/about/AuthorGuidelines#InclusionofZoologicalNomenclature}{Inclusion of Zoological Nomenclature} section of the guidelines.
-
-\section{Additional Requirements}
-
-For additional requirements for specific article types and further information please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines#AdditionalRequirements}{Author Guidelines}.
+%
+%
+%\section{Additional Requirements}
+%
+%For additional requirements for specific article types and further information please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines#AdditionalRequirements}{Author Guidelines}.
+%
 
 \section*{Conflict of Interest Statement}
+
 %All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
@@
 
-The Author Contributions section is mandatory for all articles, including articles by sole authors.
+% The Author Contributions section is mandatory for all articles, including articles by sole authors.
 If an appropriate statement is not provided on submission, a standard one will be inserted during the production process.
@@
 
+RJCB and FB conceived the idea for this research. 
+MVB helped with the proteome analysis of \emph{M. tuberculosis}.
+RJCB wrote the code.
+RJCB, MB, GvdB and FB wrote the article.
+
 \section*{Funding}
-Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
+
+% Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
+
+FB is funded by a Veni grant from the Netherlands Organization for Scientific
+Research (016.Veni.192.026) and an Off-Road Grant from the Dutch Medical Science Foundation (ZonMW 04510011910005).
+GvdB is funded by a Young Investigator Grant from 
+the Human Frontier Science Program (HFSP; RGY0080/2018), 
+and a Vidi grant from 
+the Netherlands Organization for Scientific Research (NWO-ALW VIDI 864.14.001). 
+GvdB has received funding from the European Research Council (ERC) 
+under the European Union’s Horizon 2020 research and 
+innovation programme (grant agreement No. 862137.
 
 \section*{Acknowledgments}
-This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.
+
+% This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.
+
+We thank the Center for Information Technology of the University 
+of Groningen for its support and for providing access to the Peregrine 
+high performance computing cluster.
 
@@
 \section*{Data Availability Statement}
-The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
+%The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
 % Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData
+All code, intermediate and final results are archived at 
+\url{https://github.com/richelbilderbeek/bbbq_article}.
 

### 2021-12-12 frontiers.tex
Have figure 1 and 2

@@
 \linenumbers
+
+% Added by RJCB
+% Subfigures
+%\usepackage{subcaption} % Already added
+
+% Added by RJCB
+% Subfigure numbers use uppercase, 
+% from https://tex.stackexchange.com/a/278140
+\renewcommand{\thesubfigure}{\Alph{subfigure}}
+
+% Added by RJCB
+\usepackage{listings}
+\usepackage{hyperref}
+\usepackage{todonotes}
+\usepackage{verbatim}
+\usepackage{pgf}
+\usepackage{bm}
+\usepackage{multirow}
+\usepackage{amsfonts}
+\usepackage{array}
+\usepackage{booktabs}
+\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
+\usepackage{longtable}
 
@@
 
+%
+% Figure 1
+%
 \begin{figure}[h!]
 \begin{center}
-\includegraphics[width=10cm]{logo1}% This is a *.eps file
+\includegraphics[width=0.8\textwidth]{bbbq_article_figures_Page_1}% This is a *.eps file
 \end{center}
-\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
+\caption{
+    \textbf{Over-presentation of TMH-derived epitopes on most MHC-I and -II alleles}
+    \textbf{(A)} 
+    The percentage of epitopes for MHC-I and -II alleles that are predicted to 
+    overlap with TMHs for the proteomes of SARS-CoV-2 (top row), human (middle 
+    row) and \emph{M. tuberculosis} (MtB; bottom row).
+    The pair of horizontal red lines in each plot indicate the lower and upper bound 
+    of the 99\% confidence interval.
+    See supplementary Tables \ref{tab:tmh_binders_mhc2} and \ref{tab:tmh_binders_mhc1}
+    for the exact TMH and  epitope counts.
+    \textbf{(B-C)}
+    Correlation between the percentages of predicted TMH-derived epitopes
+    and the hydrophobicity score of all predicted epitopes for 
+    human MHC-I \textbf{(B)} and MHC-II alleles \textbf{(C)}.
+    Diagonal red line: linear regression analysis. 
+    Labels are shorthand for the HLA alleles,
+    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.}
+  \label{fig:1}
 \end{figure}
+% 1A: \label{fig:bbbq_1_smart_results}
+% 1B: \label{fig:hydrophobicity_1}
+% 1C: \label{fig:hydrophobicity_2}
 
+%
+% Figure 2
+%
 \begin{figure}[h!]
 \begin{center}
-\includegraphics[width=15cm]{logos}
+\includegraphics[width=\textwidth]{bbbq_article_figures_Page_2}% This is a *.eps file
 \end{center}
-\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
+\caption{
+    \textbf{
+      Analysis of epitope database shows that TMH derived epitopes are over presented.
+    }
+    The percentage of epitopes for MHC-I and -II alleles that overlap with TMHs
+    that are presented. The pair of horizontal red lines in each plot indicate the
+    lower and upper bound of the 99\% confidence interval. Note that only one line
+    is visible as this interval is relatively narrow. 
+    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
+    \textbf{(A)} 
+    Observed and predicted percentage of TMH-derived epitopes for MHC-I alleles. 
+    \textbf{(B)} 
+    MHC ligands from IEDB corresponding to TMH-derived epitopes. 
+    The numbers above the bars denotes the number of TMH derived epitopes obtained.
+}
+\label{fig:elution}
 \end{figure}
+% Figure 2a: \label{fig:2a}
+% Figure 2b: \label{fig:2b}
 

### 2021-12-12 frontiers.tex
Add figure 3

@@
 
+%
+% Figure 3
+%
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=\textwidth]{bbbq_article_figures_Page_3}% This is a *.eps file
+\end{center}
+\caption{
+    \textbf{Evolutionary conservation of human TMHs.}
+    \textbf{(A)} 
+    Percentage of SNPs found in TMHs.
+    Each point shows for one protein the predicted percentage of
+    amino acids that are part of a TMH ($x$-axis) and the observed occurrence of SNPs being located
+    within a TMH ($y$-axis).
+    The dashed diagonal line shows the line of equality (i.e.,
+    equal conservation of TMHs and soluble protein regions). 
+    The diagonal red line indicates a linear fit, 
+    the gray area its 95\% confidence interval.
+    \textbf{(B)}
+    Distribution of the percentages of TMH in the TMPs used in this study.
+    \textbf{(C)}
+    The number of SNPs in TMHs as expected by chance (left bar) 
+    and found in the dbSNP database (right bar).
+    Percentages show the relative conservation
+    of SNPs in TMHs found relative to stochastic chance.
+}
+\label{fig:3}
+\end{figure}
+% Figure 3A: \label{fig:f_snps_found_and_expected}
+% Figure 3B: \label{fig:f_tmh_ncbi}
+% Figure 3C: \label{fig:conservation}
+
 %%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.

### 2021-12-12 frontiers.tex
Add figure 4

@@
 
+%
+% Figure 4
+%
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=\textwidth]{bbbq_article_figures_Page_4}% This is a *.eps file
+\end{center}
+\caption{\textbf{Membrane proteins with multiple TMHs are evolutionary more conserved than proteins with only a single TMH.}
+      \textbf{(A)} 
+      Percentage of SNPs found in TMPs predicted to have only a single
+      (left) or multiple (right) TMHs.
+      Each point shows for one protein the predicted percentage of amino acids that are part of a TMH ($x$-axis) and the observed occurrence of SNPs being located within a TMH ($y$-axis).
+The dashed diagonal lines show the line of equality (i.e., equal conservation of TMHs and soluble protein regions).
+      The diagonal red lines indicate a linear fit, the gray areas their 95\% confidence intervals.
+      \textbf{(B)} 
+      The number of SNPs in TMHs as expected by chance 
+      and observed in the dbSNP database, 
+      for TMPs with one TMH (single-spanners) and multiple TMHs (multi-spanners).
+      Percentages show the relative conservation
+      of SNPs in TMHs found relative to the stochastic chances.
+      \textbf{(C)} 
+      Distribution of the proportion of amino acids residing
+      in the plasma membrane. 
+}
+\label{fig:4}
+\end{figure}
+% Figure 4A: \label{fig:f_snps_found_and_expected_per_spanner}
+% Figure 4B: \label{fig:conservation_per_spanner}
+% Figure 4C: \label{fig:f_tmh_ncbi_per_spanner}
+
 %%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.

### 2021-12-12 frontiers_SupplementaryMaterial.tex
First step in supmat

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when producing your supplementary material for your Frontiers article.                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 2.5 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

\documentclass[utf8]{frontiers_suppmat} % for all articles
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}

% Leave a blank line between paragraphs instead of using \\

\begin{document}
\onecolumn
\firstpage{1}

\title[Supplementary Material]{{\helveticaitalic{Supplementary Material}}}

\maketitle

\section{Supplementary Data}

Supplementary Material should be uploaded separately on submission. Please include any supplementary data, figures and/or tables. All supplementary files are deposited to FigShare for permanent storage and receive a DOI.

Supplementary material is not typeset so please ensure that all information is clearly presented, the appropriate caption is included in the file and not in the manuscript, and that the style conforms to the rest of the article. To avoid discrepancies between the published article and the supplementary material, please do not add the title, author list, affiliations or correspondence in the supplementary files.

\section{Supplementary Tables and Figures}

For more information on Supplementary Material and for details on the different file types accepted, please see \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{the Supplementary Material section} of the Author Guidelines.

Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet). It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.

%% Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet). It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.

\subsection{Figures}

%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
%%% If using *.tif files convert them to .jpg or .png
%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=9cm]{logo1}% This is a *.eps file
\end{center}
\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=10cm]{logos}
\end{center}
\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
\end{figure}

%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
%%% Frontiers will add the figures at the end of the provisional pdf automatically
%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.

%\bibliographystyle{frontiersinSCNS_ENG_HUMS} %  for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
%\bibliography{bbbq_article}

\end{document}

### 2021-12-12 create_frontiers_supmat.sh

@@ -1,0 +1,17 @@
+#!/bin/bash
+#
+# Create the PDF 'frontiers_supmat.pdf'
+#
+# It does so by converting the TeX to PDF, 
+# as well as stitching some other PDFs to it
+#
+# Usage:
+#
+#   ./create_frontiers_supmat.sh
+#
+pdflatex frontiers_SupplementaryMaterial.tex
+bibtex frontiers_SupplementaryMaterial.tex >/dev/null
+pdflatex frontiers_SupplementaryMaterial.tex >/dev/null
+pdflatex frontiers_SupplementaryMaterial.tex
+mv frontiers_SupplementaryMaterial.pdf frontiers_supmat.pdf
+

### 2021-12-12 frontiers.tex
Have a reasonable supmat

@@
 \usepackage{longtable}
+
+% Added by RJCB
+%\usepackage{xr}
+%\externaldocument{frontiers_SupplementaryMaterial.tex}
 
@@
 
-Figure \ref{fig:bbbq_1_smart_results} shows the predicted presentation of TMH-derived peptides in MHC-I,
+Figure \ref{fig:1}A shows the predicted presentation of TMH-derived peptides in MHC-I,
 for a human, viral and bacterial proteome.
@@
 would also be confirmed for MHC-II. 
-Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
+Figure \ref{fig:1}A shows the percentages of MHC-II epitopes 
 predicted to be overlapping with TMHs for our human, viral and bacterial proteomes.
@@
 grooves \cite{bianchi2017}. 
-Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
+Figures \ref{fig:1}B and \ref{fig:1}C
 show the extent of over-presentation
@@
 and (2) the percentage of SNPs located within these predicted TMHs.
-Each percentage pair was plotted in figure \ref{fig:f_snps_found_and_expected}.
+Each percentage pair was plotted in figure \ref{fig:3}A.
 The proportion of SNPs found in TMHs varied from 
@@
 These percentages are shown as a histogram 
-in figure \ref{fig:f_tmh_ncbi}. 
+in figure \ref{fig:3}B. 
 The expected number of SNPs expected to be found in 
@@
 one finds 918 SNPs in TMHs,
-as depicted (as percentages) in figure \ref{fig:conservation}.
+as depicted (as percentages) in figure \ref{fig:3}C.
 
@@
 From the split data, we did the same analysis as for the total TMPs.
-Figure \ref{fig:f_snps_found_and_expected_per_spanner} 
+Figure \ref{fig:4}A 
 shows the percentages of TMHs for individual proteins as a function of the
@@
 when compared to soluble regions,
-as depicted in figure \ref{fig:conservation_per_spanner}.
+as depicted in figure \ref{fig:4}B.
 In effect, per 1,000 SNPs found in soluble protein domains, 

### 2021-12-12 frontiers_SupplementaryMaterial.tex
Have a reasonable supmat

@@
 
-\section{Supplementary Data}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Differences with Bianchi et al., 2017}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-Supplementary Material should be uploaded separately on submission.
-Please include any supplementary data, figures and/or tables.
-All supplementary files are deposited to FigShare for permanent storage and receive a DOI.
+A part of this study does the same analysis as Bianchi et al., 2017.
+mainly concern the use of different
+software and a different definition of what an MHC binder is.
 
-Supplementary material is not typeset so please ensure that all information is clearly presented, the appropriate caption is included in the file and not in the manuscript, and that the style conforms to the rest of the article.
-To avoid discrepancies between the published article and the supplementary material, please do not add the title, author list, affiliations or correspondence in the supplementary files.
+% \paragraph{Definition of what a binder is}
 
-\section{Supplementary Tables and Figures}
+The earlier study defined a peptide an MHC binder 
+if \emph{within the protein} in which it was found, 
+is was among the peptides with the 2\% lowest IC50 values.
+This can be seen at \url{https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/master/predict-binders.R},
+where the binders are written to file.
 
-For more information on Supplementary Material and for details on the different file types accepted, please see \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{the Supplementary Material section} of the Author Guidelines.
+However, in this study, an MHC binder is defined as a peptide within a \emph{proteome} in which it is found, that is among the peptides with the 2\% lowest IC50 values.
+Subsection \ref{subsec:ic50s_per_haplotype} shows the IC50 values
+for a binder per MHC allele.
 
-Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
-It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
+% \paragraph{Selenoproteins}
 
-%% Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
-It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
+Our previous study used the TMHMM web server
+to predict TMHs.
+The desktop version of TMHMM, however, gives an
+error message on the 25 selenoproteins found in the human
+reference proteome.
+For the sake of reproducible research, we used the desktop version (as
+we can call it from scripts) and, due to this, we removed the
+selenoproteins from this analysis.
 
-\subsection{Figures}
+% \paragraph{Compare results in figure}
 
-%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
-%%% If using *.tif files convert them to .jpg or .png
-%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%
+To verify if the previous and the current method give rise to
+notable difference, we show a side-by-side comparison
+in figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
+The figures that MHC molecules that over-present or under-present TMH-derived epitopes,
+do so in both studies. The extent to which TMH-derived epitopes are
+presented, however, is more extreme in our current setup.
 
-\begin{figure}[htbp]
+\begin{figure}[h!]
 \begin{center}
-\includegraphics[width=9cm]{logo1}% This is a *.eps file
+\includegraphics[width=\linewidth]{bianchi_et_al_2017_results/figure-1-a.png} % use same naming as them
 \end{center}
-\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
+\caption{
+    Results for \cite{bianchi2017}. 
+    Dashed lines denotes the coincidence interval.
+}
+\label{fig:bianch_et_al_2017_1a}
 \end{figure}
 
-\begin{figure}[htbp]
+\begin{figure}[h!]
 \begin{center}
-\includegraphics[width=10cm]{logos}
+\includegraphics[width=\linewidth]{bbbq_1_smart_results/fig_f_tmh_2_human_mhc1.png}
 \end{center}
-\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
+\caption{
+    Results for this study.
+    Dashed line denotes the percentage as expected by chance.
+}
+\label{fig:bilderbeek_et_al_2021_1a}
 \end{figure}
 
-%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
-%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
-%%% Frontiers will add the figures at the end of the provisional pdf automatically
-%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.
+% Process all floats before going to a next page
+\clearpage
 
-%\bibliographystyle{frontiersinSCNS_ENG_HUMS} %  for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
-%\bibliographystyle{frontiersinHLTH&FPHY} % for Health and Physics articles
-%\bibliography{bbbq_article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{IC50 values of binders per MHC allele}
+\label{subsec:ic50s_per_haplotype}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Per target proteome (i.e. human, SARS-CoV-2, \emph{M tuberculosis}),
+we collected all 9-mers (for MHC-I) and 14-mers (for MHC-II),
+after removing the selenoproteins and proteins that are shorter
+than the epitope length.
+From these epitopes, per MHC allele,
+we predicted the IC50 (in nM) using \verb;epitope-prediction; (for MHC-I)
+and MHCnuggets (for MHC-II). 
+Here, we show the IC50 value per MHC allele that
+is used to determine if a peptide binds to the allele's MHC
+for MHC-I (see supplementary Table \ref{tab:ic50_binders_mhc1}) and 
+MHC-II (see supplementary Table \ref{tab:ic50_binders_mhc2}).
+
+% tab:ic50_binders_mhc1
+\input{bbbq_1_smart_results/table_ic50_binders_mhc1_2.latex}
+
+% tab:ic50_binders_mhc2
+\input{bbbq_1_smart_results/table_ic50_binders_mhc2_2.latex}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Counts}
+\label{subsec:counts}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}
+for an overview of all amounts. 
+Note that, for the analyses using the SARS-CoV-2 virus proteome,
+we labeled this by its disease (\verb;covid;) to prevent typos.
+In supplementary Table \ref{tab:ncbi_counts_1} there are multiple instances where
+the amounts are expected to add up, yet don't, as one SNP can work on
+multiple isoforms. For example, there are 9,621 unique SNPs 
+found in all proteins, of which 4,219 around found in MAPs 
+and 6,026 in TMPs. Apparently, 624 SNPs work on a set of isoforms that
+contains both MAPs and TMPs.
+
+% Label: tab:ncbi_counts_1
+\input{ncbi_peregrine_results/ncbi_counts_1.tex}
+
+% Label: tab:ncbi_counts_2
+\input{ncbi_peregrine_results/ncbi_counts_2.tex}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Relative positions}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See Supplementary Figure \ref{fig:snp_rel_pos}
+for the distribution of the relative position of the SNPs.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snp_rel_pos.png}
+  \caption{
+    Distribution of the relative position of the SNPs used,
+    where a relative position of zero denotes the first amino
+    acid at the N-terminus, where a relative position of one
+    indicates the last residue at the C-terminus.
+  }
+  \label{fig:snp_rel_pos}
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See supplementary Table \ref{tab:tmh_binders_mhc2} 
+for the percentage of MHC-II 14-mers overlapping with TMH.
+
+% Label: tab:tmh_binders_mhc2
+\input{bbbq_1_smart_results/table_tmh_binders_mhc2_2.latex}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The percentage of TMH-derived epitopes from IEDB epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+We display the over-presentation of epitopes taken from the
+IEDB database, for two assays: an MHC ligand assay (Figure \ref{fig:2a})
+and a T cell assay (see figure \ref{fig:t_cells_present_tmh_derived_epitopes}),
+as a bar plot.
+Supplementary Table \ref{tab:elution} below shows the
+exact numbers.
+
+% tab:elution
+\input{bbbq_article_issue_157/results.tex}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Correlation of epitope presentation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In the main text of this research, we use two sources of epitopes
+to determine if TMH-derived epitopes are presented.
+The first source of epitopes are all the 9-mers (for MHC-I) (and 14-mers for MHC-II)
+derived from a human reference proteome, where this over-presentation
+is displayed in figure \ref{fig:bbbq_1_smart_results}.
+The second source of epitopes are those that are present in the IEDB
+that are obtained from MHC ligand assays,
+as displayed in figure \ref{fig:2a}.
+
+Here we correlate between the over-presentation of TMH-derived
+epitopes between these two sources of data.
+Figure \ref{fig:thm_presentation_correlation} shows per MHC allele
+the percentage of TMH-derived epitopes, with a linear trendline.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2c.png}
+  \caption{
+    \textbf{
+      TMH-derived epitopes are over-presented when using
+      predicted as well as experimental data
+    }
+    For the MHC class I alleles, the over-presentation of
+    TMH-derived epitopes is correlated between 
+    IEDB MHC ligand epitopes (horizontal axis) and the 9-mers
+    derived from a human reference proteome (vertical axis).
+    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
+    The trendline shows the linear correlation between these
+    percentages, where the gray area is the 95\% confidence interval.
+  }
+  \label{fig:thm_presentation_correlation}
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Presentation of TMH-derived epitopes result in T cell responses}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Figure \ref{fig:t_cells_present_tmh_derived_epitopes} 
+shows the percentage of TMH-derived
+epitopes of the reported epitopes from human origin for which T-cell responses were established. 
+The data was obtained from the IEDB and includes only the MHC
+alleles used in this study. As there are many (especially class II)
+MHC alleles, only a small percentage of the full IEDB data could be used.
+
+\begin{figure}[!htbp]
+  \centering
+  \includegraphics[width=0.5\textwidth]{bbbq_article_issue_157/figure_2d.png}
+  \caption{
+    \textbf{
+      TMH-derived epitopes evoke T-cell responses 
+      }
+    The numbers above the bars denotes the number of epitopes
+    found in the IEDB for the MHC alleles used in this study.
+  }
+  \label{fig:t_cells_present_tmh_derived_epitopes}
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+See supplementary Table \ref{tab:tmh_binders_mhc1} 
+for the percentage of MHC-I 9-mers overlapping with TMH.
+
+% Label: tab:tmh_binders_mhc1
+\input{bbbq_1_smart_results/table_tmh_binders_mhc1_2.latex}
+
+Supplementary Table \ref{tab:haplotype_abbreviations} shows the
+shorthand notation for the HLA alleles.
+
+% Label: tab:haplotype_abbreviations
+\input{bbbq_1_smart_results/haplotype_abbreviations.tex}
+
+Supplementary Tables \ref{tab:tmh_binders_mhc1} and \ref{tab:tmh_binders_mhc2}
+show the exact number of binders, binders that overlap with TMHs
+and the percentage of binders that overlap with TMHs, as
+visualized by figure \ref{fig:bbbq_1_smart_results}.
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Prediction software used}
+\label{subsec:prediction_software_used}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[]
+  \begin{tabular}{llll}
+    Goal & Tool & Reference \\ 
+    \hline
+    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
+    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
+    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
+    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
+    Call TMHMM from R                 & \verb;tmhmm;              & \cite{tmhmm} \\
+    Call PureseqTM from R             & \verb;pureseqtmr;         & \cite{pureseqtmr} \\
+    Call NetMHCIIpan from R           & \verb;netmhc2pan;         & \cite{netmhc2pan} \\
+    Work with IEDB                    & \verb;iedbr;              & \cite{iedbr} \\
+    Work with \verb;rentrez;          & \verb;sprentrez;          & \cite{sprentrez} \\
+    Combine all                       & \verb;bbbq;               & \cite{bbbq}
+  \end{tabular}
+  \caption{
+    Overview of all software used in this research.
+  }
+  \label{table:software_used}
+\end{table}
+
+For this research, we needed software to predict protein
+topology, as well as the MHC-I and MHC-II binding affinities
+of epitopes. We selected our software, by
+searching the scientific literature 
+to identify the most recent free and open source (FOSS) 
+prediction software.
+This was done by searching for papers that (1) cite older
+prediction software, and (2) present a novel method to make predictions.
+As a starting point, per type of prediction software,
+a review paper was used (\cite{moller2001evaluation} for protein
+topology, \cite{lundegaard2011prediction} for MHC-I
+binding affinities and \cite{nielsen2003reliable} for MHC-II binding
+affinities).
+
+% \paragraph{TMH prediction}
+
+There are multiple computational tools developed to predict which
+parts of a protein forms a TMH.
+In 2001, multiple of such prediction tools have been compared \cite{moller2001evaluation},
+of which TMHMM \cite{krogh2001predicting} turned out to be the most accurate, 
+as is used in the previous study \cite{bianchi2017}.
+However, TMHMM has a restrictive software license and is nearly two
+decades old.
+Therefore, PureseqTM \cite{wang2019efficient},
+was also used in this study, which has been more recently developed
+and has a free software license.
+
+% \paragraph{MHC-I epitope prediction}
+
+For MHC-I, there are multiple computational tools developed 
+to predict epitopes. 
+According to \cite{lundegaard2011prediction}, at that time,
+NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
+We used the same tool as used in our earlier study, 
+\verb;epitope-prediction; \cite{bianchi2017},
+
+% \paragraph{MHC-II epitope prediction}
+
+Also for MHC-II, there are multiple computational tools developed 
+to predict epitopes,
+such as using a trained neural network \cite{nielsen2003reliable}
+or a Gibbs sampling approach \cite{nielsen2004improved}.
+According to \cite{lundegaard2011prediction}, in 2011,
+from a set of multiple tools, 
+\mbox{NetMHCIIpan} \cite{nielsen2008quantitative,karosiene2013netmhciipan}
+made the most accurate predictions.
+The most recent FOSS tool available now appears
+to be MHCnuggets \cite{shao2020high}, which can do both MHC-I 
+and MHC-II predictions. 
+As we already use \verb;epitope-prediction; \cite{bianchi2017} 
+for MHC-I predictions, we use MHCnuggets only for MHC-II predictions.
+
+% \paragraph{NCBI data retrieval software}
+
+To retrieve the data from the NCBI databases the
+\verb;rentrez; R package \cite{rentrez} was used
+that calls the NCBI database's API. 
+The NCBI database provides a stable user experience for all users, 
+by limiting its API to 3 calls per second per user.
+Additionally, the API splits the result of a bigger
+query into multiple pages, each of which needs one API call.
+The \verb;sprentrez; package \cite{sprentrez} provides for 
+bigger queries of multiple (and delayed) API calls.
+
+% \paragraph{IEDB software}
+
+To retrieve the data from the IEDB  databases \cite{vita2019immune}, 
+the \verb;iedbr; R package \cite{iedbr} was written,
+to calls the IEDB database's API. 
+Similar to the NCBI database,
+the IEDB has a limit to 1 call per second per user
+and allows a query results to return 10k results maximally.
+The \verb;iedbr; package \cite{iedbr} allows for bigger queries.
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Prediction software written}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+The R programming language is used for the complete 
+experiment, including the analysis.
+The complete experiment is bundled in the 'bbbq' R package,
+which is dependent on 'tmhmm', 'pureseqtmr', 
+'epitope-prediction' and 'mhcnuggetsr'
+as described below.
+
+% \paragraph{tmhmm}
+
+The R package 'tmhmm' was developed to do the similar topology
+predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
+'TMHMM' has a restrictive software license \cite{krogh2001predicting} 
+and allows a user
+to download a pre-compiled executable after confirmation that he/she
+is in academia. The R package respects this restriction
+and allows the user to install and use TMHMM from within R,
+as done in this study.
+'tmhmm' has been submitted to and is accepted 
+by the Comprehensive R Archive Network (CRAN).
+
+% \paragraph{pureseqtmr}
+
+To be able to call, from R, the TMH prediction 
+software 'PureseqTM' \cite{wang2019efficient},
+which is written in C, the package 'pureseqtmr' has been developed. 
+'pureseqtmr' allows to install 'PureseqTM' and use most of its features.
+'pureseqtmr' has been submitted to and is accepted by CRAN.
+
+% \paragraph{mhcnuggetsr}
+
+MHCnuggets is a free and open-source Python package to predict 
+epitope affinity for many MHC-I and MHC-II variants \cite{shao2020high}.
+The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
+from within R.
+Also 'mhcnuggetsr' has been submitted to and is accepted by CRAN.
+
+% \paragraph{bbbq}
+
+To reproduce the full experiment presented in this paper,
+the functions needed are bundled in the 'bbbq' R package.
+This package is too specific to be submitted to CRAN.
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Prediction of percentage of epitopes overlapping with a TMH}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Supplementary Table \ref{tab:f_tmh} shows an overview of the findings,
+where a target specifies the source of the proteome,
+where \verb;covid; denotes SARS-CoV-2 and \verb;myco; denotes
+\emph{Mycobacterium tuberculosis}. \verb;mhc_class; denotes the MHC
+class, \verb;n_spots; the number of possible 9-mers (for MHC-I) 
+or 14-mers (for MHC-II) possible. \verb;n_spots_tmh; the
+number of epitopes that overlapped with a TMH that were binders. 
+\verb;f_tmh; the percentage of peptides that had at least 1 residue
+overlapping with a TMH.
+
+% Label: tab:f_tmh
+\input{bbbq_1_smart_results/table_f_tmh_2.latex}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Minor methods}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+These are details that are removed from the 'Methods' section.
+
+PureseqTM does not predict the topology
+of proteins that have less than three amino acids. 
+The TRDD1 ('T cell receptor delta diversity 1') protein,
+however, is two amino acids long. 
+The R package \verb;pureseqtmr;, however, 
+predicts that mono- and di-peptides are cytosolic.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Minor discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+These are details that are removed from the 'Discussion' section.
+
+% \paragraph{Bacteria have a different cell membrane}
+
+In this experiment we predicted epitopes that overlap with 
+TMHs from a human, bacterial and viral proteome,
+would these proteins be expressed in a human host.
+Bacteria, however have different cell membranes and cell walls, 
+hence different structural requirements for a TMH.
+Both topology prediction tools were trained to recognize
+human TMHs, thus we cannot be sure that
+the transmembrane regions predicted in bacterial proteins
+are actually part of a TMH.
+For the purpose of this study, we assume the 
+error in topology predictions to be unbiased way towards topology.
+In other words: that a bacterial TMH is incorrectly
+predicted to be absent just as often as it is incorrectly
+predicted to be present elsewhere.
+
+% \paragraph{False positives in SNPs}
+
+Regarding the evolutionary conservation of TMHs using SNPs,
+again, it is estimated that approximately ten percent
+of SNPs is a false positive that result from the methods to determine
+a SNP. One example is that sequence variations are incorrectly
+detected due to highly similar duplicated sequences \cite{musumeci2010single}.
+We assume that these duplications occur as often in TMHs as in
+regions around these, hence we expect this not to affect our results.
+
+% \paragraph{synonymous mutations}
+%
+In our evolutionary experiment, 
+we removed variations that were synonymous mutations (i.e.
+resulted in the same amino acid, from a different genetic code) 
+from our analysis.
+There is evidence, however, that these synonymous mutations
+do have an effect and may even be evolutionary selected 
+for \cite{hunt2009silent}.
+As the possible effect of synonymous mutations is ignored by our
+topology prediction software, we do so as well.
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Relative presentation of TMH-derived epitopes}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+To compare the over-presentation of TMH-derived epitopes between the
+different proteomes, we normalized this percentages in such a
+way that 1.0 is the percentage of TMH-derived epitopes that would 
+be expected by chance. 
+Figure \ref{fig:f_tmh_mhc1_normalized} and \ref{fig:f_tmh_mhc2_normalized}
+show these normalized values for the MHC-I and MHC-II alleles respectively.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc1_2_normalized.png}
+  \caption{
+    Normalized proportion of MHC-I epitopes overlapping with TMHs
+    for human, viral and bacterial proteomes.
+    Legend: covid = SARS-CoV-2,
+    human = \emph{Homo sapiens}, 
+    myco = \emph{Mycobacterium tuberculosis}
+  }
+  \label{fig:f_tmh_mhc1_normalized}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_f_tmh_mhc2_2_normalized.png}
+  \caption{
+    Normalized proportion of MHC-II epitopes overlapping with TMHs
+    for human, viral and bacterial proteomes.
+    Legend: covid = SARS-CoV-2,
+    human = \emph{Homo sapiens}, myco = \emph{Mycobacterium tuberculosis}
+  }
+  \label{fig:f_tmh_mhc2_normalized}
+\end{figure}
+
+To determine the additional over-presentation of TMH-derived epitopes 
+in MHC-II (as compared to MHC-I), we normalized the data to enable
+a side-by-side comparison. 
+The percentage of TMH-derived epitopes presented was normalized
+to the expected percentage of TMH-derived epitopes,
+where $1.0$ denotes that the percentage of presented TMH-derived epitopes
+matches the values as expected by chance.
+The normalized values per MHC allele are shown 
+in figure \ref{fig:rel_presentation_per_haplotype}.
+To compare the TMH-derived over-presentation per MHC class,
+we grouped the normalized values per allele, 
+and plot the mean and standard error, as shown in figure \ref{fig:rel_presentation}.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation_per_haplotype.png}
+  \caption{
+    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs,
+    for the different MHC alleles and proteomes
+  }
+  \label{fig:rel_presentation_per_haplotype}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_results/fig_rel_presentation.png}
+  \caption{
+    Normalized proportion of MHC-I and MHC-II epitopes overlapping with TMHs,
+    for the different MHC classes and proteomes. Error bars denote the
+    standard error.
+  }
+  \label{fig:rel_presentation}
+\end{figure}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Evolutionary conservation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Figure \ref{fig:snps_per_gene_name_ncbi} shows the distribution of the
+number of SNPs per gene name, at the date we started the experiment,
+at December 14th 2020.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_ncbi.png}
+  \caption{
+    Distribution of the number of SNPs per gene name in the NCBI database.
+  }
+  \label{fig:snps_per_gene_name_ncbi}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_snps_per_gene_name_processed.png}
+  \caption{
+    Distribution of the number of protein variations and SNPs per gene name processed.
+  }
+  \label{fig:snps_per_gene_name_processed}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_proteins_per_gene_name.png}
+  \caption{
+    Histogram of the number of proteins found per gene name.
+    Most often, a gene name is associated with one proteins. 
+  }
+  \label{fig:n_proteins_per_gene_name}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_snps_per_tmp.png}
+  \caption{
+    Histogram of the number of SNPs per trans-membrane protein.
+    Dashed vertical line: average number of SNPs per TMP
+  }
+  \label{fig:fig_n_snps_per_tmp}
+\end{figure}
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{ncbi_peregrine_results/fig_n_tmhs_per_protein.png}
+  \caption{
+    Histogram of the number of TMHs predicted per protein,
+    for the trans-membrane proteins used.
+  }
+  \label{fig:fig_n_tmhs_per_protein}
+\end{figure}
+
+% \paragraph{Relative position of SNPs}
+
+To verify if SNPs were sampled uniformly
+over proteins, we show the distribution 
+of the relative position in figure \ref{fig:snp_rel_pos}.
+We find no clear evidence of a bias.
+
+% \paragraph{Statistics}
+
+Supplementary Table \ref{tab:snp_stats} shows the statistics for all
+SNPs, where supplementary Tables \ref{tab:snp_stats_per_spanner_single}
+and \ref{tab:snp_stats_per_spanner_multi} show the
+statistics for only single-spanners and multi-spanners respectively.
+
+% Label: tab:snp_stats
+\input{ncbi_peregrine_results/snp_stats.tex}
+
+% Label: tab:snp_stats_per_spanner_single
+\input{ncbi_peregrine_results/snp_stats_per_spanner_single.tex}
+
+% Label: tab:snp_stats_per_spanner_multi
+\input{ncbi_peregrine_results/snp_stats_per_spanner_multi.tex}
+
+% Process all floats before going to a next page
+\clearpage
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Presentation of TMH-derived epitopes when two amino acids overlap}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In our experiment, we define a TMH-derived epitope as a peptide that overlaps
+with a TMH for at least one amino acid. One could argue that we should use
+a higher number of overlapping amino acids, so to make the epitopes more
+'transmembrane helix-ey'. We chose not too, for two reason: (1) epitopes that overlap with a TMH
+for 1 AA already, cannot be processed by the proteasome in a known and conventional
+way (2) whatever number of overlapping amino acids we use, we expect the pattern to be the same.
+However, using only 1 AA gives the most TMH-derived epitopes and hence the highest statistical
+power.
+
+To prove this point, we did exactly the same analysis 
+as shown in Figure \ref{fig:bbbq_1_smart_results},
+yet with defining a TMH-derived epitope as an epitope that overlaps with a TMH
+for at least 2 AAs, as shown in Figure \ref{fig:bbbq_1_smart_results_2aas}.
+As these two figures look identical, 
+we also added the counts as numbers,
+with Table \ref{tab:tmh_binders_mhc2_2aas} showing the same data
+as \ref{tab:tmh_binders_mhc2}, except the former uses 2 AAs overlap.
+Likewise, Table \ref{tab:tmh_binders_mhc1_2aas} showing the same data
+as \ref{tab:tmh_binders_mhc1}, except the former uses 2 AAs overlap.
+
+\begin{figure}[!htbp]
+  \includegraphics[width=\textwidth]{bbbq_1_smart_2aa_results/fig_f_tmh_2_panel.png}
+  \caption{
+    The percentage of epitopes for MHC-I and -II alleles that are predicted to 
+    overlap with TMHs (for at least two amino acids) 
+    for the proteomes of SARS-CoV-2 (top row), human (middle 
+    row) and \emph{M. tuberculosis} (bottom row).
+    The pair of dashed lines in each plot indicate the lower and upper bound of the 99\% confidence interval.
+    See supplementary Tables \ref{tab:tmh_binders_mhc2_2aas} and \ref{tab:tmh_binders_mhc1_2aas}
+    for the exact TMH and  epitope counts.
+  }
+  \label{fig:bbbq_1_smart_results_2aas}
+\end{figure}
+
+% Label: tab:tmh_binders_mhc2_2aas
+\input{bbbq_1_smart_2aa_results/table_tmh_binders_mhc2_2.latex}
+
+% Label: tab:tmh_binders_mhc1_2aas
+\input{bbbq_1_smart_2aa_results/table_tmh_binders_mhc1_2.latex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Abbreviations}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{table}[h]
+  \begin{tabular}{ll}
+    Abbreviation & Full               \\ 
+    \hline
+ER & Endoplasmatic reticulum \\
+ERAD & ER-associated degradation \\
+HLA & Human leukocyte antigen \\
+IEDB & Immune Epitope Database  \\
+LB & lipid body \\
+MAP & Membrane-associated protein \\
+MHC & Major histocompatibility complex \\
+MVB & Multivesicular body \\
+PLC & Peptide-loading complex \\
+SNP & Single nucleotide polymorphism \\
+TMH & Transmembrane helix         \\
+    TMP & Transmembrane protein
+  \end{tabular}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
+%\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
+\bibliography{bbbq_article}
 

### 2021-12-12 create_frontiers_supmat.sh

@@ -10,8 +10,7 @@
 #   ./create_frontiers_supmat.sh
 #
 pdflatex frontiers_SupplementaryMaterial.tex
-bibtex frontiers_SupplementaryMaterial.tex >/dev/null
+bibtex frontiers_SupplementaryMaterial >/dev/null
 pdflatex frontiers_SupplementaryMaterial.tex >/dev/null
 pdflatex frontiers_SupplementaryMaterial.tex
-mv frontiers_SupplementaryMaterial.pdf frontiers_supmat.pdf
 

### 2021-12-12 frontiers.tex
'frontiers.pdf' is done

@@
 all peptides within a \emph{proteome} 
-(see supplementary Tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
+(see supplementary Tables ST1 and ST2 for values), 
 this differs from our previous study where we defined
@@
 a side by side comparison is shown in the supplementary materials, 
-Figures \ref{fig:bianch_et_al_2017_1a} and \ref{fig:bilderbeek_et_al_2021_1a}.
+Figures S1 and S2.
 
@@
 The exact amounts can be found in the supplementary materials,
-Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2}.
+Tables ST3 and ST4.
 
@@
 To verify that these ID numbers are unrelated to SNP positions, the relative positions of all analyzed SNPs in a protein were determined.
-This analysis showed no positional bias of the SNPs, as shown in supplementary figure \ref{fig:snp_rel_pos}.
+This analysis showed no positional bias of the SNPs, as shown in supplementary figure S3.
 
@@
 except for \verb;HLA-DRB3*0101; in \emph{M. tuberculosis}.
-See supplementary Table \ref{tab:tmh_binders_mhc2} 
+See supplementary Table ST5 
 for the exact TMH and epitope counts.
@@
 
-Figure (\ref{fig:2a} and \ref{fig:thm_presentation_correlation}) shows there are similar levels of
+Figure (\ref{fig:2}A and S4) shows there are similar levels of
 over-presentation of TMH-derived epitopes between (1) the
@@
 
-In figure \ref{fig:2b} we grouped all the epitopes 
+In figure \ref{fig:2}B we grouped all the epitopes 
 presented by MHC-I and MHC-II alleles 
@@
 epitopes derived from human TMHs are presented in both MHC-I and MHC-II, and support that they are over-presented.
-See the supplementary Table \ref{tab:elution} for the exact values.
+See the supplementary Table ST6 for the exact values.
 
@@
 concerned TMH-derived epitopes in MHC class I and II, respectively 
-(see Figure \ref{fig:t_cells_present_tmh_derived_epitopes}).
+(see Figure S5).
 
@@
 of which 6,062 were located in predicted TMPs.
-See supplementary Tables \ref{tab:ncbi_counts_1} and \ref{tab:ncbi_counts_2} 
+See supplementary Tables ST3 and ST4 
 for the detailed numbers and distributions of SNPs.
@@
     of the 99\% confidence interval.
-    See supplementary Tables \ref{tab:tmh_binders_mhc2} and \ref{tab:tmh_binders_mhc1}
+    See supplementary Tables ST5 and ST7
     for the exact TMH and  epitope counts.
@@
     Labels are shorthand for the HLA alleles,
-    see the supplementary Table \ref{tab:haplotype_abbreviations} for the names.}
+    see the supplementary Table ST8 for the names.}
   \label{fig:1}
@@
     is visible as this interval is relatively narrow. 
-    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
+    Alleles are listed in Table ST8).
     \textbf{(A)} 
@@
 }
-\label{fig:elution}
+\label{fig:2}
 \end{figure}

### 2021-12-12 frontiers_SupplementaryMaterial.tex
Supmat done

@@
 We display the over-presentation of epitopes taken from the
-IEDB database, for two assays: an MHC ligand assay (Figure \ref{fig:2a})
+IEDB database, for two assays: an MHC ligand assay (Figure 2A)
 and a T cell assay (see figure \ref{fig:t_cells_present_tmh_derived_epitopes}),
@@
 derived from a human reference proteome, where this over-presentation
-is displayed in figure \ref{fig:bbbq_1_smart_results}.
+is displayed in figure 1A.
 The second source of epitopes are those that are present in the IEDB
 that are obtained from MHC ligand assays,
-as displayed in figure \ref{fig:2a}.
+as displayed in figure 2A.
 
@@
 and the percentage of binders that overlap with TMHs, as
-visualized by figure \ref{fig:bbbq_1_smart_results}.
+visualized by figure 1A.
 
@@
 To prove this point, we did exactly the same analysis 
-as shown in Figure \ref{fig:bbbq_1_smart_results},
+as shown in Figure 1A,
 yet with defining a TMH-derived epitope as an epitope that overlaps with a TMH

### 2021-12-12 frontiers.tex
Use 'citep' instead of 'cite' in LaTeX

@@
 For example, HLA-A and HLA-B have no overlap in which
-epitopes they bind \cite{lund2004definition}.
+epitopes they bind \citep{lund2004definition}.
 Moreover, the HLA genes of humans are highly polymorphic, with hundreds 
 to thousands of different alleles, 
-and each different allele presents a different subset of peptides \cite{marsh2010nomenclature}.
+and each different allele presents a different subset of peptides \citep{marsh2010nomenclature}.
 
@@
 so-called escape mutations, 
-will not affect MHC presentation for all alleles present in the population \cite{sommer2005importance}.
+will not affect MHC presentation for all alleles present in the population \citep{sommer2005importance}.
 
@@
 predictions of the peptide binding affinities
-\cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.
-For example, S. Tang et al. \cite{tang2011genome} found that, 
+\citep{larsen2010identification,schellens2008unanticipated,tang2011genome}.
+For example, S. Tang et al. \citep{tang2011genome} found that, 
 of the 432 peptides that were predicted to bind to an MHC allele,
@@
 are likely to be more frequently presented by MHC-I 
-than expected based on their abundance \cite{bianchi2017},
+than expected based on their abundance \citep{bianchi2017},
 which is in line with a previous study 
-by Istrail et al \cite{istrail2004comparative},
+by Istrail et al \citep{istrail2004comparative},
 demonstrating that N-terminal signal sequences 
@@
 TMHs can also be predicted with high accuracy from a protein sequence 
-by bioinformatics approaches \cite{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}. 
-For example, a study by Jones \cite{jones2007improving} found that,
+by bioinformatics approaches \citep{krogh2001predicting,kall2004combined,arai2004conpred,jones2007improving,klammer2009metatm,wang2019efficient}. 
+For example, a study by Jones \citep{jones2007improving} found that,
 from 184 transmembrane proteins (TMPs) with known topology, 
@@
 Different TMH prediction tools estimate that 15-39\% of all proteins 
-in the human proteome contain at least one TMH \cite{ahram2006estimation}.
+in the human proteome contain at least one TMH \citep{ahram2006estimation}.
 However, the physiological reason why peptides derived from TMHs 
@@
 to have hydrophobic side chains 
-\cite{hessa2007molecular,jones1994model}.
+\citep{hessa2007molecular,jones1994model}.
 Therefore, we speculated that the TMHs of pathogens 
@@
 We confirmed the presentation of TMH-derived peptides 
-by re-analysis of peptides from The Immune Epitope Database (IEDB) \cite{vita2019immune}.
+by re-analysis of peptides from The Immune Epitope Database (IEDB) \citep{vita2019immune}.
 Second, we tested our hypothesis that TMHs 
@@
 To predict how frequently epitopes overlapping with TMHs are presented,
-a similar analysis strategy was applied as described in \cite{bianchi2017} 
+a similar analysis strategy was applied as described in \citep{bianchi2017} 
 for several alleles of both MHC-I and MHC-II, 
@@
 as this is the length most frequently presented in MHC-I 
-and was used in our earlier study \cite{bianchi2017}. 
+and was used in our earlier study \citep{bianchi2017}. 
 For MHC-II, 14-mers were used, 
-as this is the most frequently occurring epitope length \cite{bergseng2015different}.
+as this is the most frequently occurring epitope length \citep{bergseng2015different}.
 A human (UniProt ID UP000005640\_9606), 
@@
 reference proteome was used. 
-\verb;TMHMM; \cite{krogh2001predicting} was used to predict the topology 
+\verb;TMHMM; \citep{krogh2001predicting} was used to predict the topology 
 of the proteins within these proteomes.
 To predict the affinity of an epitope to a certain HLA allele,
- \verb;EpitopePrediction; \cite{bianchi2017} for MHC-I 
-and \verb;MHCnuggets; \cite{shao2020high} for MHC-II was used.
+ \verb;EpitopePrediction; \citep{bianchi2017} for MHC-I 
+and \verb;MHCnuggets; \citep{shao2020high} for MHC-II was used.
 Both MCH-I and MHC-II alleles were selected 
@@
 where the alleles of MHC-I are the alleles representing the 13 supertypes 
-with over 99.6\% coverage of the population's MHC-I repertoire as defined by \cite{lund2004definition} \cite{sette1999},
+with over 99.6\% coverage of the population's MHC-I repertoire as defined by \citep{lund2004definition} \citep{sette1999},
 and the 21 MHC-II alleles, have a phenotypic frequency 
-of 14\% or more in the human population \cite{greenbaum2011functional}.
+of 14\% or more in the human population \citep{greenbaum2011functional}.
 
@@
 are presented by MHC-I and MHC-II,
-we mined the IEDB \cite{vita2019immune}
+we mined the IEDB \citep{vita2019immune}
 for confirmed human MHC-ligands.
@@
 We predicted the topology of the protein each epitope
-was found in, using \verb;TMHMM; \cite{krogh2001predicting},
+was found in, using \verb;TMHMM; \citep{krogh2001predicting},
 from which we concluded if the epitope is overlapping with a TMH
@@
 NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases were used: 
-the \emph{dbSNP} \cite{sherry2001dbsnp} database,
+the \emph{dbSNP} \citep{sherry2001dbsnp} database,
 which contains 650 million 
 cataloged non-redundant human variations (called RefSNPs,
-\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \cite{brown2015gene})
-and \emph{protein} (for proteins sequences \cite{sayers2010database}).
+\url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \citep{brown2015gene})
+and \emph{protein} (for proteins sequences \citep{sayers2010database}).
 
@@
 is similar for the probed proteomes, 
-which strengthens our previous conclusion \cite{bianchi2017} 
+which strengthens our previous conclusion \citep{bianchi2017} 
 that the hydrophobicity of the MHC-binding groove 
@@
 peptides is caused by the hydrophobicity of the peptide binding 
-grooves \cite{bianchi2017}. 
+grooves \citep{bianchi2017}. 
 Figures \ref{fig:1}B and \ref{fig:1}C
@@
 the molecular mechanisms of how integral membrane proteins are processed 
-for MHC presentation are largely unknown \cite{bianchi2017}. 
+for MHC presentation are largely unknown \citep{bianchi2017}. 
 Most prominently, the fundamental principles of 
@@
 The chaperones tapasin (TAPBP), ERp57 (PDIA3), 
-and calreticulin (CALR) \cite{rock2016present} first assemble 
+and calreticulin (CALR) \citep{rock2016present} first assemble 
 and stabilize the heavy and light chains of MHC-I. 
@@
 The PLC drives import of peptides into the ER 
-and mediates their subsequent loading into the peptide-binding groove of MHC-I \cite{blees2017structure}. 
+and mediates their subsequent loading into the peptide-binding groove of MHC-I \citep{blees2017structure}. 
 Membrane proteins first will have to be extracted from the membrane 
@@
 ubiquitinate them, and extract them from the ER membrane 
-into the cytosol (retrotranslocation) for proteasomal degradation \cite{preston2017evolving,meusser2005erad}. 
+into the cytosol (retrotranslocation) for proteasomal degradation \citep{preston2017evolving,meusser2005erad}. 
 Similar to the peptides generated from soluble proteins, 
@@
 into the ER for MHC-I loading. 
-This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \cite{bougneres2009role}, 
-since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \cite{fujimoto2006proteasomal}.
+This ERAD-driven antigen retrotranslocation might be facilitated by lipid bodies (LBs) \citep{bougneres2009role}, 
+since LBs can serve as cytosolic sites for ubiquitination of ER-derived cargo \citep{fujimoto2006proteasomal}.
 
@@
 Supporting this hypothesis is the well-established notion that peptides generated by signal peptide peptidases (SPPs), an important class of intramembrane proteases that cleave TMH-like signal sequences,
-are presented on a specialized class of MHC-I called HLA-E \cite{oliveira2015alternative}. 
+are presented on a specialized class of MHC-I called HLA-E \citep{oliveira2015alternative}. 
 The loading of peptides generated by SPP onto MHC-I does not depend on the proteosome and TAP, 
-possibly because the peptides are directly released into the lumen of the ER \cite{oliveira2015alternative}. 
+possibly because the peptides are directly released into the lumen of the ER \citep{oliveira2015alternative}. 
 However, this mechanism cannot explain how most membrane proteins can be processed for antigen presentation, because SPPs only cleave TMH-like signal sequences at their C-termini, and N-terminal domains will hence not be removed.
 Nevertheless, the presentation of peptides with a high hydrophobicity index 
-was shown to be independent of TAP as well \cite{lautscham2001processing}, 
+was shown to be independent of TAP as well \citep{lautscham2001processing}, 
 suggesting that the TMH peptides might perhaps be released directly in the ER lumen by other intramembrane proteases.
 
-A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \cite{oliveira2015alternative}. 
+A third possibility is that peptide processing and MHC-loading occur in multivesicular bodies (MVBs) \citep{oliveira2015alternative}. 
 TMPs can be routed from the plasma membrane and other organelles by vesicular trafficking to endosomes.
@@
 Upon the fusion of MVBs with lysosomes,
-the entire intraluminal vesicles including the TMPs are degraded \cite{gruenberg2020life}. 
+the entire intraluminal vesicles including the TMPs are degraded \citep{gruenberg2020life}. 
 Via this mechanism, TMPs might well be processed for antigen presentation, 
 particularly since the loading of MHC-II molecules is well understood 
-to occur in MVBs \cite{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}. 
+to occur in MVBs \citep{kleijmeer2001reorganization,peters1991segregation,zwart2005spatial}. 
 However, such processing of membrane proteins in MVBs for antigen presentation poses a problem, because complexes of HLA-DR with its antigen-loading chaperon HLA-DM were only observed on intraluminal vesicles,
-but not on the limiting membranes of MVBs \cite{zwart2005spatial}, 
+but not on the limiting membranes of MVBs \citep{zwart2005spatial}, 
 indicating that epitope loading of MHC-II also occurs at intraluminal vesicles.
@@
 and it is hence unclear how TMPs from these microbes 
-would be routed to MVBs for MHC-II loading \cite{zwart2005spatial}.
+would be routed to MVBs for MHC-II loading \citep{zwart2005spatial}.
 
-Alternatively to the enzymatic degradation of lipids in MVBs by lipases \cite{sander2016lipase,gilleron2016lysosomal}, 
-they might be oxidatively degraded by reactions with radical oxygen species produced by the NADPH oxidase NOX2 \cite{dingjan2016lipid}. 
-This oxidation can result in a destabilization and disruption of membranes \cite{dingjan2016lipid} 
+Alternatively to the enzymatic degradation of lipids in MVBs by lipases \citep{sander2016lipase,gilleron2016lysosomal}, 
+they might be oxidatively degraded by reactions with radical oxygen species produced by the NADPH oxidase NOX2 \citep{dingjan2016lipid}. 
+This oxidation can result in a destabilization and disruption of membranes \citep{dingjan2016lipid} 
 and might thereby lead to the extraction of TMPs. 
@@
 selection pressure is directed upon residues that change its 
-virulence \cite{velazquez2020positive}.
+virulence \citep{velazquez2020positive}.
 These regions, however, may only account for a small part of a pathogen's proteome.
@@
 to provide innate immune responses against such rare essential protein regions, 
-as suggested in a study on influenza \cite{han2019individual},
+as suggested in a study on influenza \citep{han2019individual},
 where it was found that the selection pressure
@@
 using different methodologies.
-One such example is a study by Stevens and Arkin \cite{stevens2001substitution}, 
+One such example is a study by Stevens and Arkin \citep{stevens2001substitution}, 
 in which aligned protein sequence data was used.
@@
 which is a similar value as we found.
-Another example is a study by Oberai, et al. \cite{oberai2009structural} that estimated the conservation
+Another example is a study by Oberai, et al. \citep{oberai2009structural} that estimated the conservation
 scores for TMHs and soluble regions based on 

### 2021-12-12 frontiers_SupplementaryMaterial.tex
Use 'citep' instead of 'cite' in LaTeX

@@
 \caption{
-    Results for \cite{bianchi2017}. 
+    Results for \citep{bianchi2017}. 
     Dashed lines denotes the coincidence interval.
@@
     \hline
-    Predict topology                  & TMHMM                     & \cite{krogh2001predicting} \\
-    Predict topology                  & PureseqTM                 & \cite{wang2019efficient} \\
-    Predict epitopes MHC-I            & \verb;epitope-prediction; & \cite{bianchi2017} \\
-    Predict epitopes MHC-II           & NetMHCIIpan               & \cite{nielsen2008quantitative,karosiene2013netmhciipan} \\
-    Call TMHMM from R                 & \verb;tmhmm;              & \cite{tmhmm} \\
-    Call PureseqTM from R             & \verb;pureseqtmr;         & \cite{pureseqtmr} \\
-    Call NetMHCIIpan from R           & \verb;netmhc2pan;         & \cite{netmhc2pan} \\
-    Work with IEDB                    & \verb;iedbr;              & \cite{iedbr} \\
-    Work with \verb;rentrez;          & \verb;sprentrez;          & \cite{sprentrez} \\
-    Combine all                       & \verb;bbbq;               & \cite{bbbq}
+    Predict topology                  & TMHMM                     & \citep{krogh2001predicting} \\
+    Predict topology                  & PureseqTM                 & \citep{wang2019efficient} \\
+    Predict epitopes MHC-I            & \verb;epitope-prediction; & \citep{bianchi2017} \\
+    Predict epitopes MHC-II           & NetMHCIIpan               & \citep{nielsen2008quantitative,karosiene2013netmhciipan} \\
+    Call TMHMM from R                 & \verb;tmhmm;              & \citep{tmhmm} \\
+    Call PureseqTM from R             & \verb;pureseqtmr;         & \citep{pureseqtmr} \\
+    Call NetMHCIIpan from R           & \verb;netmhc2pan;         & \citep{netmhc2pan} \\
+    Work with IEDB                    & \verb;iedbr;              & \citep{iedbr} \\
+    Work with \verb;rentrez;          & \verb;sprentrez;          & \citep{sprentrez} \\
+    Combine all                       & \verb;bbbq;               & \citep{bbbq}
   \end{tabular}
@@
 As a starting point, per type of prediction software,
-a review paper was used (\cite{moller2001evaluation} for protein
-topology, \cite{lundegaard2011prediction} for MHC-I
-binding affinities and \cite{nielsen2003reliable} for MHC-II binding
+a review paper was used (\citep{moller2001evaluation} for protein
+topology, \citep{lundegaard2011prediction} for MHC-I
+binding affinities and \citep{nielsen2003reliable} for MHC-II binding
 affinities).
@@
 parts of a protein forms a TMH.
-In 2001, multiple of such prediction tools have been compared \cite{moller2001evaluation},
-of which TMHMM \cite{krogh2001predicting} turned out to be the most accurate, 
-as is used in the previous study \cite{bianchi2017}.
+In 2001, multiple of such prediction tools have been compared \citep{moller2001evaluation},
+of which TMHMM \citep{krogh2001predicting} turned out to be the most accurate, 
+as is used in the previous study \citep{bianchi2017}.
 However, TMHMM has a restrictive software license and is nearly two
 decades old.
-Therefore, PureseqTM \cite{wang2019efficient},
+Therefore, PureseqTM \citep{wang2019efficient},
 was also used in this study, which has been more recently developed
@@
 to predict epitopes. 
-According to \cite{lundegaard2011prediction}, at that time,
-NetMHCcons \cite{karosiene2012netmhccons} gave the best predictions.
+According to \citep{lundegaard2011prediction}, at that time,
+NetMHCcons \citep{karosiene2012netmhccons} gave the best predictions.
 We used the same tool as used in our earlier study, 
-\verb;epitope-prediction; \cite{bianchi2017},
+\verb;epitope-prediction; \citep{bianchi2017},
 
@@
 to predict epitopes,
-such as using a trained neural network \cite{nielsen2003reliable}
-or a Gibbs sampling approach \cite{nielsen2004improved}.
-According to \cite{lundegaard2011prediction}, in 2011,
+such as using a trained neural network \citep{nielsen2003reliable}
+or a Gibbs sampling approach \citep{nielsen2004improved}.
+According to \citep{lundegaard2011prediction}, in 2011,
 from a set of multiple tools, 
-\mbox{NetMHCIIpan} \cite{nielsen2008quantitative,karosiene2013netmhciipan}
+\mbox{NetMHCIIpan} \citep{nielsen2008quantitative,karosiene2013netmhciipan}
 made the most accurate predictions.
 The most recent FOSS tool available now appears
-to be MHCnuggets \cite{shao2020high}, which can do both MHC-I 
+to be MHCnuggets \citep{shao2020high}, which can do both MHC-I 
 and MHC-II predictions. 
-As we already use \verb;epitope-prediction; \cite{bianchi2017} 
+As we already use \verb;epitope-prediction; \citep{bianchi2017} 
 for MHC-I predictions, we use MHCnuggets only for MHC-II predictions.
@@
 To retrieve the data from the NCBI databases the
-\verb;rentrez; R package \cite{rentrez} was used
+\verb;rentrez; R package \citep{rentrez} was used
 that calls the NCBI database's API. 
@@
 query into multiple pages, each of which needs one API call.
-The \verb;sprentrez; package \cite{sprentrez} provides for 
+The \verb;sprentrez; package \citep{sprentrez} provides for 
 bigger queries of multiple (and delayed) API calls.
@@
 
-To retrieve the data from the IEDB  databases \cite{vita2019immune}, 
-the \verb;iedbr; R package \cite{iedbr} was written,
+To retrieve the data from the IEDB  databases \citep{vita2019immune}, 
+the \verb;iedbr; R package \citep{iedbr} was written,
 to calls the IEDB database's API. 
@@
 and allows a query results to return 10k results maximally.
-The \verb;iedbr; package \cite{iedbr} allows for bigger queries.
+The \verb;iedbr; package \citep{iedbr} allows for bigger queries.
 
@@
 predictions as our earlier study (that used 'TMHMM'), yet in an automated way.
-'TMHMM' has a restrictive software license \cite{krogh2001predicting} 
+'TMHMM' has a restrictive software license \citep{krogh2001predicting} 
 and allows a user
@@
 To be able to call, from R, the TMH prediction 
-software 'PureseqTM' \cite{wang2019efficient},
+software 'PureseqTM' \citep{wang2019efficient},
 which is written in C, the package 'pureseqtmr' has been developed. 
@@
 MHCnuggets is a free and open-source Python package to predict 
-epitope affinity for many MHC-I and MHC-II variants \cite{shao2020high}.
+epitope affinity for many MHC-I and MHC-II variants \citep{shao2020high}.
 The R package 'mhcnuggetsr' allows one to install and use MHCnuggets
@@
 a SNP. One example is that sequence variations are incorrectly
-detected due to highly similar duplicated sequences \cite{musumeci2010single}.
+detected due to highly similar duplicated sequences \citep{musumeci2010single}.
 We assume that these duplications occur as often in TMHs as in
@@
 do have an effect and may even be evolutionary selected 
-for \cite{hunt2009silent}.
+for \citep{hunt2009silent}.
 As the possible effect of synonymous mutations is ignored by our

### 2021-12-12 zip_frontiers_template.sh
Zip Frontiers template

@@ -1,0 +1,10 @@
+#!/bin/bash
+#
+# Create a zip files with exactly those files that Frontiers needs
+#
+# Usage:
+#
+#   ./zip_frontiers_template.sh
+#
+
+zip frontiers_template.zip bbbq_article.bib frontiers_SupplementaryMaterial.tex frontiers.tex *.bst *.cls *.sty

### 2021-12-29 bbbq_article.tex
Fix typo, thanks Frans
Reword sentence, thanks Frans

@@
     Histogram of the number of proteins found per gene name.
-    Most often, a gene name is associated with one proteins. 
+    Most often, a gene name is associated with one protein.
   }
@@
 In our experiment, we define a TMH-derived epitope as a peptide that overlaps
-with a TMH for at least one amino acid. One could argue that we should use
-a higher number of overlapping amino acids, so to make the epitopes more
-'transmembrane helix-ey'. We chose not too, for two reason: (1) epitopes that overlap with a TMH
+with a TMH for at least one amino acid. 
+One could argue that we should use a higher number of overlapping amino acids, so that the epitope has a higher TMH coverage.
+We chose not too, for two reason: (1) epitopes that overlap with a TMH
 for 1 AA already, cannot be processed by the proteasome in a known and conventional

### 2021-12-29 frontiers_SupplementaryMaterial.tex
Fix typo, thanks Frans
Reword sentence, thanks Frans

@@
     Histogram of the number of proteins found per gene name.
-    Most often, a gene name is associated with one proteins. 
+    Most often, a gene name is associated with one protein. 
   }
@@
     Histogram of the number of SNPs per trans-membrane protein.
-    Dashed vertical line: average number of SNPs per TMP
   }

### 2021-12-29 bbbq_article.tex
Reword sentence, thanks Frans

@@
 One could argue that we should use a higher number of overlapping amino acids, so that the epitope has a higher TMH coverage.
-We chose not too, for two reason: (1) epitopes that overlap with a TMH
-for 1 AA already, cannot be processed by the proteasome in a known and conventional
-way (2) whatever number of overlapping amino acids we use, we expect the pattern to be the same.
+We chose not to, for two reason: (1) epitopes that overlap
+with a TMH for 1 AA already, cannot be processed by the proteasome in a known and conventional way as it still requires extraction from the membrane
+(2) whatever number of overlapping amino acids we use, we expect the pattern to be the same as the chance that an epitope stems from a TMH is equally reduced.
 However, using only 1 AA gives the most TMH-derived epitopes and hence the highest statistical

### 2021-12-29 frontiers.tex
Update keywords, fix Q9

@@
 \tiny
- \keyFont{ \section{Keywords:} antigen presentation, membrane proteins, bioinformatics, adaptive immunity, transmembrane domain, transmembrane helix, epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation}
+ \keyFont{ \section{Keywords:} antigen presentation, membrane proteins,adaptive immunity, transmembrane domain, epitopes, MHC-I, MHC-II, evolutionary conservation}
 \end{abstract}

### 2021-12-29 manuscript.tex
Fix typo, thanks Frans
Update keywords, fix Q9

@@
 
-{\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
-adaptive immunity, transmembrane domain, transmembrane helix, 
-epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation
+{\bf Keywords:} antigen presentation, membrane proteins,adaptive immunity, transmembrane domain, epitopes, MHC-I, MHC-II, evolutionary conservation
 

### 2021-12-29 frontiers.tex
Fix title

@@
 
-\title[Transmembrane helices are 
-  an overlooked and evolutionarily conserved
-  source of major histocompatibility complex class I and II epitopes
-]{Transmembrane helices are 
-  an overlooked and evolutionarily conserved
-  source of major histocompatibility complex class I and II epitopes
+\title[Transmembrane helices are an over-presented and evolutionarily conserved source of major histocompatibility complex class I and II epitopes
+]{Transmembrane helices are an over-presented and evolutionarily conserved source of major histocompatibility complex class I and II epitopes
 }

### 2021-12-29 frontiers.tex
Refer to fig 4C, fix Q15

@@
 
-Also, for single- and multi-spanners, 
-we determined the relevance of this finding by
-calculating how much less SNPs are found in TMHs
-when compared to soluble regions,
-as depicted in figure \ref{fig:4}B.
+Also, for single- and multi-spanners, we determined the
+relevance of this finding by calculating where and how much less SNPs
+are found in TMHs when compared to soluble regions, as
+depicted in Figure \ref{fig:4}B 
+and \ref{fig:4}C.
 In effect, per 1,000 SNPs found in soluble protein domains, 

### 2021-12-29 manuscript.tex
Fix title
Refer to fig 4C, fix Q15

@@
 \title{
-  Transmembrane helices are 
-  an overlooked and evolutionarily conserved
-  source of major histocompatibility complex class I and II epitopes
+  Transmembrane helices are an over-presented and evolutionarily conserved source of major histocompatibility complex class I and II epitopes
 }
@@
 
-Also, for single- and multi-spanners, 
-we determined the relevance of this finding by
-calculating how much less SNPs are found in TMHs
-when compared to soluble regions,
-as depicted in figure \ref{fig:conservation_per_spanner}.
+Also, for single- and multi-spanners, we determined the
+relevance of this finding by calculating where and how much less SNPs
+are found in TMHs when compared to soluble regions, as
+depicted in Figure \ref{fig:conservation_per_spanner} 
+and \ref{fig:f_tmh_ncbi_per_spanner}
 In effect, per 1,000 SNPs found in soluble protein domains, 

### 2021-12-29 bbbq_article.tex
Fix Q16
Fix typo, thanks Geert

@@
 and therefore hydrophobic TMH-derived peptides have a higher affinity 
-to bind than their soluble hydrophobic 
+to bind than their soluble hydrophilic 
 counterparts.
@@
 
-All code, intermediate and final results are archived at 
+The data presented in the study are deposited in the Zenodo repository, accession number 10.5281/zenodo.5809139.
+Additionally, all code, intermediate and final results are archived at 
 \url{https://github.com/richelbilderbeek/bbbq_article}.

### 2021-12-29 frontiers.tex
Fix Q16
Fix typo, thanks Geert

@@
 \section*{Data Availability Statement}
-%The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK].
-% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData
-All code, intermediate and final results are archived at 
+The data presented in the study are deposited in the Zenodo repository, accession number 10.5281/zenodo.5809139.
+Additionally, all code, intermediate and final results are archived at 
 \url{https://github.com/richelbilderbeek/bbbq_article}.

### 2021-12-29 bbbq_article.tex
Fix typo, thanks Geert

@@
 with UniProt ID UP000005640\_9606.
-We predicted the topology of the protein each epitope
-was found in, using \verb;TMHMM; \cite{krogh2001predicting},
-from which we concluded if the epitope is overlapping with a TMH
-with at least 1 amino acid.
-
+We concluded that the epitope overlapped with a TMH if at least 1 amino acid was overlapping with a TMH, as predicted with \verb;TMHMM; \cite{krogh2001predicting}.
 The full analysis can be found

### 2021-12-29 frontiers.tex
Fix typo, thanks Geert

@@
 with UniProt ID UP000005640\_9606.
-We predicted the topology of the protein each epitope
-was found in, using \verb;TMHMM; \citep{krogh2001predicting},
-from which we concluded if the epitope is overlapping with a TMH
-with at least 1 amino acid.
+We concluded that the epitope overlapped with a TMH if at least 1 amino acid was overlapping with a TMH, as predicted with \verb;TMHMM; \cite{krogh2001predicting}.
 

### 2021-12-29 bbbq_article.tex
Fix typo, thanks Geert

@@
 which contains 650 million 
-cataloged non-redundant human variations (called RefSNPs,
+cataloged non-redundant human variations (called \textit{RefSNPs},
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \cite{brown2015gene})

### 2021-12-29 frontiers.tex
Fix typo, thanks Geert

@@
 which contains 650 million 
-cataloged non-redundant human variations (called RefSNPs,
+cataloged non-redundant human variations (called \textit{RefSNPs},
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \citep{brown2015gene})

### 2021-12-29 bbbq_article.tex
Fix wording, thanks Geert

@@
 
-To determine the evolutionary conservation of TMHs,
-we first collected human single nucleotide polymorphisms (SNPs) 
-resulting in a single amino acid substitution to determine if this occurred within a predicted TMH or not.
+To determine the evolutionary conservation of TMHs, we first
+collected human single nucleotide polymorphisms (SNPs)
+resulting in a single amino acid substitution and determined if
+this occurred within a predicted TMH or not.
 

### 2021-12-29 bbbq_article.tex
Database names in italics, thanks Geert

@@
 which contains 650 million 
-cataloged non-redundant human variations (called \textit{RefSNPs},
+cataloged non-redundant human variations (called \emph{RefSNPs},
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \cite{brown2015gene})

### 2021-12-29 frontiers.tex
Database names in italics, thanks Geert

@@
 which contains 650 million 
-cataloged non-redundant human variations (called \textit{RefSNPs},
+cataloged non-redundant human variations (called \emph{RefSNPs},
 \url{https://www.ncbi.nlm.nih.gov/snp/docs/RefSNP_about/}), and the databases \emph{gene} (for gene names \citep{brown2015gene})

### 2021-12-29 bbbq_article.tex
Improve wording, thanks Geert
Fix typ, thanks Geert

@@
 
-We next wondered if the over-representation of TMH-derived peptides 
-would also be confirmed for MHC-II. 
+We next wondered if the over-representation of TMH-derived peptides would also be present for MHC-II.
 Figure \ref{fig:bbbq_1_smart_results} shows the percentages of MHC-II epitopes 
@@
 
-Figure (\ref{fig:2a} and \ref{fig:thm_presentation_correlation}) shows there are similar levels of
+Figure \ref{fig:2a} and \ref{fig:thm_presentation_correlation} show there are similar levels of
 over-presentation of TMH-derived epitopes between (1) the

### 2021-12-29 frontiers.tex
Improve wording, thanks Geert
Fix typ, thanks Geert

@@
 
-We next wondered if the over-representation of TMH-derived peptides 
-would also be confirmed for MHC-II. 
+We next wondered if the over-representation of TMH-derived peptides would also be present for MHC-II.
 Figure \ref{fig:1}A shows the percentages of MHC-II epitopes 
@@
 
-Figure (\ref{fig:2}A and S4) shows there are similar levels of
+Figure \ref{fig:2}A and S4 show there are similar levels of
 over-presentation of TMH-derived epitopes between (1) the

### 2021-12-29 bbbq_article.tex
Fix typo, thanks Geert
Split up sentence in two, thanks Geert

@@
 
-We also mined the IEDB database for epitopes for
-any type of T-cell response from the specified
-alleles, from the total reports 36\% and 7\% 
-concerned TMH-derived epitopes in MHC class I and II, respectively 
-(see Figure \ref{fig:t_cells_present_tmh_derived_epitopes}).
+We also mined the IEDB database for epitopes for any type of
+T cell response from the specified alleles.
+From the total reports, 36\% and 7\% concerned TMH-derived epitopes in MHC class I and II, respectively (see Figure \ref{fig:t_cells_present_tmh_derived_epitopes}).
 
@@
     \textbf{
-      Analysis of epitope database shows that TMH derived epitopes are over presented.
+      Analysis of epitope database shows that TMH derived epitopes are over-presented.
     }
@@
     is visible as this interval is relatively narrow. 
-    Alleles are listed in Table \ref{tab:haplotype_abbreviations}).
+    Alleles are listed in Table \ref{tab:haplotype_abbreviations}.
     \textbf{(A)} 

### 2021-12-29 frontiers.tex
Fix typo, thanks Geert
Split up sentence in two, thanks Geert

@@
 
-We also mined the IEDB database for epitopes for
-any type of T-cell response from the specified
-alleles, from the total reports 36\% and 7\% 
-concerned TMH-derived epitopes in MHC class I and II, respectively 
-(see Figure S5).
+We also mined the IEDB database for epitopes for any type of
+T cell response from the specified alleles. From the total reports, 36\% and 7\% concerned TMH-derived epitopes in MHC class I and II, respectively (see Figure S5).
 
@@
     is visible as this interval is relatively narrow. 
-    Alleles are listed in Table ST8).
+    Alleles are listed in Table ST8.
     \textbf{(A)} 

### 2021-12-29 bbbq_article.tex
Add bracket, thanks Geert
Use Meld to be (more) sure versions match

@@
 
-Also, for single- and multi-spanners, 
-we determined the relevance of this finding by
-calculating how much less SNPs are found in TMHs
-when compared to soluble regions,
-as depicted in figure \ref{fig:conservation_per_spanner}.
+Also, for single- and multi-spanners, we determined the
+relevance of this finding by calculating where and how much less SNPs
+are found in TMHs when compared to soluble regions, as
+depicted in Figure \ref{fig:conservation_per_spanner} 
+and \ref{fig:conservation_per_spanner}.
 In effect, per 1,000 SNPs found in soluble protein domains, 
@@
 under the European Union’s Horizon 2020 research and 
-innovation programme (grant agreement No. 862137.
+innovation programme (grant agreement No. 862137).
 

### 2021-12-29 frontiers.tex
Add bracket, thanks Geert
Use Meld to be (more) sure versions match

@@
 We concluded that the epitope overlapped with a TMH if at least 1 amino acid was overlapping with a TMH, as predicted with \verb;TMHMM; \cite{krogh2001predicting}.
-
 The full analysis can be found
@@
 
-%
-% Frontiers text, just keep it in for now to be sure.
-%
-%\section{Article types}
-%
-% For requirements for a specific article type please refer to the Article Types on any Frontiers journal page.
-Please also refer to  \href{http://home.frontiersin.org/about/author-guidelines#Sections}{Author Guidelines} for further information on how to organize your manuscript in the required sections or their equivalents for your field
-%
-% For Original Research articles, please note that the Material and Methods section can be placed in any of the following ways: before Results, before Discussion or after Discussion.
-%
-%\section{Manuscript Formatting}
-%
-%\subsection{Heading Levels}
-%
-%There are 5 heading levels
-%
-%\subsection{Level 2}
-%\subsubsection{Level 3}
-%\paragraph{Level 4}
-%\subparagraph{Level 5}
-%
-%\subsection{Equations}
-%Equations should be inserted in editable format from the equation editor.
-%
-%\begin{equation}
-%\sum x+ y =Z\label{eq:01}
-%\end{equation}
-%
-%\subsection{Figures}
-%Frontiers requires figures to be submitted individually, in the same order as they are referred to in the manuscript.
-Figures will then be automatically embedded at the bottom of the submitted manuscript.
-Kindly ensure that each table and figure is mentioned in the text and in numerical order.
-Figures must be of sufficient resolution for publication \href{http://home.frontiersin.org/about/author-guidelines#ResolutionRequirements}{see here for examples and minimum requirements}.
-Figures which are not according to the guidelines will cause substantial delay during the production process.
-Please see \href{http://home.frontiersin.org/about/author-guidelines#GeneralStyleGuidelinesforFigures}{here} for full figure guidelines.
-Cite figures with subfigures as figure \ref{fig:2}B.
-%
-%
-%\subsubsection{Permission to Reuse and Copyright}
-%Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
-It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
-%%Figures, tables, and images will be published under a Creative Commons CC-BY licence and permission must be obtained for use of copyrighted material from other sources (including re-published/adapted/modified/partial figures and images from the internet).
-It is the responsibility of the authors to acquire the licenses, to follow any citation instructions requested by third-party rights holders, and cover any supplementary charges.
-%
-%\subsection{Tables}
-%Tables should be inserted at the end of the manuscript.
-Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
-Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
-These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance.
-The author will be notified during the typesetting of the final article if this is the case.
-%
-%\section{Nomenclature}
-%
-%\subsection{Resource Identification Initiative}
-%To take part in the Resource Identification Initiative, please use the corresponding catalog number and RRID in your current manuscript.
-For more information about the project and for steps on how to search for an RRID, please click \href{http://www.frontiersin.org/files/pdf/letter_to_author.pdf}{here}.
-%
-%\subsection{Life Science Identifiers}
-%Life Science Identifiers (LSIDs) for ZOOBANK registered names or nomenclatural acts should be listed in the manuscript before the keywords.
-For more information on LSIDs please see \href{http://www.frontiersin.org/about/AuthorGuidelines#InclusionofZoologicalNomenclature}{Inclusion of Zoological Nomenclature} section of the guidelines.
-%
-%
-%\section{Additional Requirements}
-%
-%For additional requirements for specific article types and further information please refer to \href{http://www.frontiersin.org/about/AuthorGuidelines#AdditionalRequirements}{Author Guidelines}.
-%
-
 \section*{Conflict of Interest Statement}
-
-%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
-If no such relationship exists, authors will be asked to confirm the following statement:
 
@@
 \section*{Author Contributions}
-
-% The Author Contributions section is mandatory for all articles, including articles by sole authors.
-If an appropriate statement is not provided on submission, a standard one will be inserted during the production process.
-The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work.
-Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.
 
@@
 \section*{Funding}
-
-% Details of all funding sources should be provided, including grant numbers if applicable. Please ensure to add all necessary funding information, as after publication this is no longer possible.
 
@@
 under the European Union’s Horizon 2020 research and 
-innovation programme (grant agreement No. 862137.
+innovation programme (grant agreement No. 862137).
 
 \section*{Acknowledgments}
-
-% This is a short text to acknowledge the contributions of specific colleagues, institutions, or agencies that aided the efforts of the authors.
 
@@
 \section*{Supplemental Data}
- \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
+
+%\href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
 LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.

### 2021-12-29 bbbq_article.tex
Compare versions

@@
 
-{\bf Keywords:} antigen presentation, membrane proteins, bioinformatics, 
-adaptive immunity, transmembrane domain, transmembrane helix, 
-epitopes, T lymphocyte, MHC-I, MHC-II, evolutionary conservation
+{\bf Keywords:} antigen presentation, membrane proteins, adaptive immunity, transmembrane domain, epitopes, MHC-I, MHC-II, evolutionary conservation
 
